### 2013-01-22 doc/numerical-instability-rda.tex
Initial commit.

\documentclass[11pt]{article}
\usepackage{graphicx, amsmath, amssymb, bm, url, mathtools, natbib, amsthm, setspace}

\pagestyle{plain}
%----------------Page dimensions ----------------
\oddsidemargin 0.0in
\evensidemargin 0.0in
\topmargin -0.75in
\leftmargin 0in
\headheight 0.0in
\headsep 0.5in
%\footheight 0.0in
\footskip 0.5in
\footnotesep 0.0in
\textwidth 6.7in
\textheight 9.5in
%-----------Define Pictures---------------------
\def\picture #1 by #2 (#3){
 \vbox to #2{
   \hrule width #1 height 0pt depth 0pt
   \vfill
   \special{picture #3} % this is the low-level interface
   }
 }
\def\scaledpicture #1 by #2 (#3 scaled #4){{
 \dimen0=#1 \dimen1=#2
 \divide\dimen0 by 1000 \multiply\dimen0 by #4
 \divide\dimen1 by 1000 \multiply\dimen1 by #4
 \picture \dimen0 by \dimen1 (#3 scaled #4)}
 }

\newcommand{\xbar}{\bar{\bm x}}
\newcommand{\tr}{\text{tr}}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}

\title{Numerical Instability of the RDA Classifier}

\author{John A. Ramey and Dean M. Young}

\begin{document}

\newtheorem*{thm}{Theorem}
\newtheorem*{cor}{Corollary}

\bibpunct{(}{)}{;}{a}{,}{,}

\doublespacing

\maketitle

\textbf{The derivation below was initially in my paper entitled ``Another Look at Regularized Discriminant Analysis.'' The reasoning below is flawed though. In my proof, I assume implicitly that the eigenvalues of $\bm A$ are constant as $p$ is incremented. This is not true though: As $p$ increases, the trace of $\bm A$ must increase, so the sum of the eigenvalues of $\bm A$ must also increase. Because the first $q$ eigenvalues of $\bm A$ are positive, while the remainder of them are 0, it must be that the eigenvalues change as $p$ increases. It may be that the reasoning below is solid in many cases, or perhaps in all cases if the constant $c$ is selected correctly in the proof of the theorem below. No matter, in its current form, the proof is flawed.}

\textbf{Holy Grail Question: Assuming $\bm \Gamma$ is a diagonal matrix as defined below, what are the necessary and sufficient conditions for $|\bm A + \bm \Gamma| \rightarrow 0$ so that $\bm A + \bm \Gamma$ is nearly singular as the number of variables included in the model is increased for a fixed rank (sample size)?}

The usual focus of regularization methods employed to stabilize the inverse covariance matrix is often effective in producing well-performing classifiers when $p >> N$. As we have discussed, the shrinkage factor employed in \eqref{eq:sig-rda} involves the average of the eigenvalues of \eqref{eq:sig-lambda}. In this section, we demonstrate that the \emph{RDA} classifier becomes numerically incalculable as $p$ increases for a fixed sample size. First, we present a general result  that if a positive semidefinite matrix $\bm A$ is shrunken towards a diagonal matrix $\bm \Gamma$ consisting of positive diagonal elements less than 1, the shrunken matrix $\bm A + \bm \Gamma$ is nearly positive semidefinite because its determinant goes to 0 as $p \rightarrow \infty$. Next, we demonstrate briefly that the average of the eigenvalues goes to 0 as $p \rightarrow \infty$ for a fixed sample size. The result is that the covariance-matrix shrinkage employed in the \emph{RDA} classifier is ineffective at producing a positive definite covariance matrix estimator, resulting in a numerically incalculable classifier.

\begin{thm}
Let $\bm \Gamma = \text{diag}(\gamma_1, \gamma_2, \ldots, \gamma_p)$ such that $0 < \gamma_j < 1$ for $j = 1, 2, \ldots, p$, and let $\bm A \in \mathbb{R}_{p \times p}^{\ge}$ be symmetric such that $\text{rank}(\bm A) = q < p$ and with eigenvalues $e_1 \ge e_2 \ge \ldots \ge e_p \ge 0$. Then, with $q$ fixed,
\begin{align*}
	\lim_{p \rightarrow \infty} |\bm A + \bm \Gamma| = 0.
\end{align*}
\end{thm}
\begin{proof}
Let $\gamma = \max_j \gamma_j$, and let $c \in \mathbb{R}$ such that $e_1 \le (c - 1) \gamma$. Then, we have that 
	\begin{align*}
		|\bm A + \bm \Gamma| &= \prod_{j = 1}^p (e_j + \gamma_j)\\
%		&\le \prod_{j = 1}^p (e_j + \gamma)\\
		&\le \gamma^{p-q} \prod_{j=1}^q (e_j + \gamma)\\
		&\le c^q\gamma^p.
	\end{align*}
Therefore, as $p \rightarrow \infty$, $|\bm A + \bm \Gamma| \rightarrow 0$ because $c^q\gamma^p \rightarrow 0$.

\end{proof}

\begin{cor}
	 Let $\bm A$ be defined as in the above theorem, and let $\gamma \in (0,1)$. Then, with $q$ fixed,
\begin{align}
	\lim_{p \rightarrow \infty} |\bm A + \gamma \bm I_p| = 0.\label{eq:corollary}
\end{align}
\end{cor}
\begin{proof}
The proof follows directly from the above theorem.
\end{proof}

From \eqref{eq:corollary}, we see that shrinking a large singular covariance matrix $\bm A$ can have no regularization effect on $|\bm A|$.  Supervised classifiers that are functions of the determinant of the covariance matrix, $\widehat{\bm \Sigma}$, can yield invalid results when $p >> N$ because $\text{rank}(\widehat{\bm \Sigma}) \le N - 1$. Now, consider again the RDA classifier in \eqref{eq:rda}. We can see that $D(\bm x) \rightarrow \infty$ as $p \rightarrow \infty$ because $\widehat{\bm\Sigma}_k \rightarrow 0$, $k = 1, \ldots, K$. The natural logarithm is a nuisance function with this classifier.

As we have shown above, $\lim_{p \rightarrow \infty} \gamma = 0$. Hence, assuming the rank of $\widehat{\bm\Sigma}_k = q_k$, the last $p-q_k$ eigenvalues are 0. This implies that the RDA classifier is not applicable for high-dimensional data.

\section{Example}

To demonstrate the corollary, we first construct a $\bm A \in \mathbb{R}_{p \times p}^{\ge}$, having its first $q = 25$ eigenvalues equal to 1 and the remaining $p - q$ eigenvalues equal to 0. In Figure \ref{fig:log-det}, we see that $\ln |\bm\Sigma + \gamma \bm I_p|$ approaches $-\infty$ quickly for small values of $\gamma$ but is still finite for larger values of $\gamma$. In Figure 2, we see further supportive evidence of our main result for larger values of $\gamma$. For example, $\ln |\bm\Sigma + \gamma \bm I_p|$ is numerically $-\infty$ for $p \ge 400$ with $\gamma = 0.1$.

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale = 0.45]{log-determinant}
\caption{For various values of $\gamma$, we plot $\ln |\bm\Sigma + \gamma \bm I_p|$ as a function of $p$ with rank$(\bm\Sigma)$ fixed at 25. The points on the lower edge of the graph are numerically $-\infty$.}
\label{fig:log-det}
\end{center}
\end{figure}

% The following graphics are produced with the R code:
%library('plyr')
%library('ggplot2')
%
%eps <- c(0.1, 0.3, 0.5, 0.7, 0.9)
%p <- seq.int(100, 2000, by = 100)
%grid <- expand.grid(eps = eps, p = p)
%
%r <- 25
%results_25 <- ddply(grid, .(eps, p), function(g) {
%  log(det(diag(g$eps + c(rep(1, r), rep(0, g$p - r)))))
%})
%names(results_25) <- c("eps", "p", "det")
%results_25$eps <- factor(results_25$eps)
%gp <- ggplot(results_25, aes(x = p, y = det, group = eps, color = eps, linetype = eps))
%gp <- gp + geom_line(size = 2) + geom_point(aes(shape = eps), size = 2)
%gp <- gp + ylab("Log-Determinant")
%gp + theme_set(theme_bw())
%

\end{document}

### 2013-01-22 reports/paper-etrda.tex
Initial commit.

\documentclass[11pt]{article}
\usepackage{graphicx, amsmath, amssymb, bm, url, mathtools, natbib, amsthm, setspace}

\pagestyle{plain}
%----------------Page dimensions ----------------
\oddsidemargin 0.0in
\evensidemargin 0.0in
\topmargin -0.75in
\leftmargin 0in
\headheight 0.0in
\headsep 0.5in
%\footheight 0.0in
\footskip 0.5in
\footnotesep 0.0in
\textwidth 6.7in
\textheight 9.5in
%-----------Define Pictures---------------------
\def\picture #1 by #2 (#3){
 \vbox to #2{
   \hrule width #1 height 0pt depth 0pt
   \vfill
   \special{picture #3} % this is the low-level interface
   }
 }
\def\scaledpicture #1 by #2 (#3 scaled #4){{
 \dimen0=#1 \dimen1=#2
 \divide\dimen0 by 1000 \multiply\dimen0 by #4
 \divide\dimen1 by 1000 \multiply\dimen1 by #4
 \picture \dimen0 by \dimen1 (#3 scaled #4)}
 }

\newcommand{\xbar}{\bar{\bm x}}
\newcommand{\tr}{\text{tr}}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}

\title{A Generalization of Regularized Discriminant Analysis with Applications to High-dimensional Microarray Data}

\author{John A. Ramey and Dean M. Young}

\begin{document}

\newtheorem*{thm}{Theorem}
\newtheorem*{cor}{Corollary}

\bibpunct{(}{)}{;}{a}{,}{,}

\doublespacing

\maketitle

\begin{abstract}
\cite{Friedman:1989tm} proposed the popular regularized discriminant analysis (\emph{RDA}) classifier for the small-sample, high-dimensional setting by employing a regularized covariance matrix estimator, which is shrunken towards a scaled identity matrix. We demonstrate empirically, however, that the RDA classifier is often numerically incalculable or unstable for small-sample, high-dimensional data because the shrinkage factor is a function of the average of the eigenvalues, which tends to zero in this setting. Alternatively, \cite{Srivastava:2007ww} propose an empirical-Bayesian estimator that employs a similar shrinkage factor using the average of the nonzero eigenvalues.  Because the eigenvalues are often highly skewed, this average is often excessively large, so that the near-zero eigenvalues are substantially overcompensated. A comparable eigenvalue adjustment method from \cite{Thomaz:2006ef} often yields a shrinkage factor that is approximately equal to the average of the eigenvalues. By writing each shrinkage method as a matrix function, we generalize the shrinkage method of \cite{Thomaz:2006ef} to allow a powerful estimation technique that does not over inflate the near-zero eigenvalues. By incorporating our proposed covariance-matrix regularization with the RDA classifier, we overcome the numerical incalculability issues of the \emph{RDA} classifier. Moreover, we show that our proposed classifier is often superior to the regularized classifiers mentioned above in terms of classification accuracy using three simulated and five small-sample, high-dimensional microarray data sets. Additionally, we provide implementations of the competing classifiers in the {\tt regdiscrim} R package, available on CRAN.
\end{abstract}

\section{Introduction}

Microarrays have enabled researchers to measure gene-expression levels of thousands of genes in a single experiment to identify a subset of genes and their expression patterns that either characterize a particular cell state or facilitate an accurate prediction of future cell state.  We focus on the classification of tissues and other biological samples into one of $K$ disjoint, predefined cancer types, where our goal is to determine accurately the class membership of a new tissue sample based on gene-expression microarray data. However, microarray data typically consist of a large number of gene expressions $p$ relative to the sample size $n$. Moreover, standard benchmark microarray data sets consist of between $p = 10,000$ and $55,000$ probes and fewer than $n = 100$ observations, in which case standard supervised-learning methods exhibit poor classification performance, due to the \emph{curse of dimensionality} \citep{Bellman:1961tn}. Furthermore, well-known classifiers, including linear discriminant analysis \emph{(LDA)} and quadratic discriminant analysis \emph{(QDA)}, are incalculable when $p > n$ because the sample covariance matrices employed in both classifiers are singular. As a result, many researchers, including \cite*{Merchante:2012vk}, \cite{Witten:2011kc}, \cite*{Pang:2009ik}, \cite*{Huang:2010ju}, and \cite*{Clemmensen:2011kr}, have restricted the \emph{LDA} and \emph{QDA} models to reduce number of covariance parameters that must be estimated to ensure that the resulting classifiers are calculable.  For instance, \cite{Dudoit:2002ev} have proposed a modified \emph{LDA} classifier where the class covariance matrices are assumed to be diagonal.

Alternatively, several researchers, including \cite*{Zhang:2010va}, \cite*{Guo:2007te}, \cite{Srivastava:2007ud}, and \cite{Mkhadri:1995jp}, have proposed regularization methods to improve the estimation of the population covariance matrices before applying the \emph{LDA} and \emph{QDA} classifiers. Covariance-matrix regularization methods often are based on the well-known ridge-regression approach from \cite{Hoerl:1970cd} and adjust the eigenvalues of the sample covariance matrices to ensure that the estimator is positive definite. This shrinkage stabilizes eigenvalues that are near zero to ensure that the inverse of the the sample covariance matrix is well-posed. For an excellent overview of the regularization problem in discriminant analysis, see \cite{Murphy:2012uq} and \cite{Mkhadri:1997gy}.

Here, we consider specifically the  popular regularized discriminant analysis (\emph{RDA}) classifier from \cite{Friedman:1989tm} that incorporates a convex combination of the sample covariance matrices from the \emph{LDA} and \emph{QDA} classifiers. Because the resulting biased covariance-matrix estimator often remains singular in the small-sample, high-dimensional setting, \cite{Friedman:1989tm} applies covariance-matrix shrinkage, where the resulting covariance-matrix estimator is shrunken towards a scaled identity matrix. \cite{Friedman:1989tm} have shown that the \emph{RDA} classifier can attain improved classification accuracy, compared to the \emph{LDA} and \emph{QDA} classifiers when $n > p$.

(\textbf{Remove this paragraph?})
Despite its popularity, we demonstrate empirically that the covariance-matrix shrinkage employed in the \emph{RDA} classifier is often ineffective in adjusting the eigenvalues of the \emph{RDA} covariance-matrix estimator when $p \gg n$. The \emph{RDA} classifier is numerically unstable as the number of features incorporated into the \emph{RDA} classifier is increased because the covariance-matrix estimator is nearly singular. Specifically, the shrinkage applied in the \emph{RDA} classifier is a function of the average eigenvalues of the convex combination of the class sample covariance matrix and the sample covariance matrix pooled across all classes. Because the average of the eigenvalues is often nearly zero when $p \gg n$, the resulting covariance-matrix estimator has two issues. First, the determinant of the covariance-matrix estimator approaches zero, suggesting that the \emph{RDA} covariance-matrix estimator is nearly singular in the small-sample, high-dimensional setting. Second, the smallest eigenvalues are near zero, which causes the \emph{RDA} classifier to have an incalculable or numerically unstable discriminant function.

In this paper we present a generalization of the \emph{RDA} classifier in terms of matrix functions. We show that several common covariance-matrix regularization methods can be specified in terms of a matrix function including the \emph{RDA} classifier as well as methods from \cite{Srivastava:2007ww} and \cite*{Thomaz:2006ef} and several sparse diagonal classifiers, including the diagonal \emph{LDA} (\emph{DLDA}) and \emph{QDA} (\emph{DQDA}) classifiers from \cite{Dudoit:2002ev}, and some of its variants including those proposed by \cite{Pang:2009ik}.  Our generalization elucidates some properties of these covariance-regularization methods and the relationships among them. Moreover, our generalization allows us to view the covariance-matrix regularization methods in terms of eigenvalue thresholds, which further allows us to identify situations where each method is ineffective and ineffective. From a perspective of eigenvalue thresholds, we propose two new classifiers that are often superior to the \emph{RDA} classifier in terms of classification performance because both methods employ a more robust eigenvalue thresholding to regularize the convex combination of covariance matrix estimators employed in the \emph{RDA} classifier. Our two proposed methods also have an advantage over the \emph{RDA} classifier because both methods employ dimension reduction by reducing the number of basis functions included in the \emph{RDA} classifier, thereby reducing the variance of the resulting classifiers.

In our first classifier we generalize the eigenvalue replacement employed in the \emph{NLDA} classifier and apply a soft threshold to the eigenvalues of the \emph{RDA} covariance-matrix estimator. We refer to our first classifier as the \emph{RDA} with soft eigenvalue-thresholding (\emph{Soft-ETRDA}) classifier.  As we show, the eigenvalue replacement employed in the \emph{NLDA} classifier is often approximately equal to the shrinkage applied in the \emph{RDA} classifier for small-sample, high-dimensional data. Furthermore, a special case of the \emph{Soft-ETRDA} classifier is the \emph{NLDA} classifier as well as the \emph{RDA} classifier.  We remark that our soft eigenvalue threshold is similar to and partially motivated by the shrunken centroids classifier from \cite*{Tibshirani:2002ht}. 	In our second proposed classifier, we apply a hard threshold to the eigenvalues of the \emph{RDA} covariance-matrix estimator, so that the number of covariance parameters incorporated into the \emph{RDA} classifier is reduced substantially. We refer to this classifier as the \emph{RDA} with hard eigenvalue-thresholding (\emph{Hard-ETRDA}) classifier. We remark that the inverse of \emph{Hard-ETRDA} covariance-matrix estimator includes as a special case, the Moore-Penrose pseudo inverse of the \emph{RDA} covariance-matrix estimator.

Using \textbf{five} small-sample, high-dimensional microarray data sets, we demonstrate that our two proposed classifiers often yield superior classification accuracy to the \emph{RDA}, \emph{MDEB}, and \emph{NLDA} classifiers. We also include in our comparison the alternatively-defined \emph{RDA} classifier from \cite*{Hastie:2008dt}, the regularized \emph{LDA} classifier from \cite{Guo:2007te}, a recently proposed variant of diagonal discriminant analysis from \cite*{Tong:2012hw}, and the penalized \emph{LDA} classifier from \cite{Witten:2011kc}. Additionally, we examine the classification accuracy of the classifiers considered using \textbf{three} simulation configurations. Furthermore, we show that our proposed eigenvalue-threshold methods are more numerically stable than the other considered regularization methods using the simulated configurations.

We have organized the remainder of the paper as follows. In Section 2 we review the \emph{RDA}, the \emph{MDEB}, and \emph{NLDA} classifiers and define each in terms of matrix functions. In Section 3 we generalize the eigenvalue adjustment employed in the \emph{NLDA} classifier and define our proposed \emph{Soft-ETRDA} and \emph{Hard-ETRDA} classifiers. In Section 4 we describe the \textbf{three} simulation configurations and \textbf{five} small-sample, high-dimensional microarray data sets and then demonstrate that our two proposed classifiers are superior to the competing classifiers in terms of classification accuracy. We then conclude with a brief discussion in Section 6.

\section{Covariance-Matrix Regularization in Discriminant Analysis}

In discriminant analysis we wish to correctly assign an unlabeled $p$-dimensional observation vector $\bm x$ to one of $K$ unique, known classes (or populations) by constructing a classifier from $n$ training observations that can accurately predict the class membership of $\bm x$. Let $\bm x_i = (x_{i1}, x_{i2}, \ldots, x_{ip}) \in \mathbb{R}_{p \times 1}$ be the $i$th observation $(i = 1, \ldots, n)$ with true, unique membership $y_i \in \mathcal{K} = \{\omega_1, \ldots, \omega_K\}$, where $\mathbb{R}_{m \times n}$ denotes the matrix space of all $m \times n$ matrices over the real field $\mathbb{R}$. We assume that $(\bm x_i, y_i)$ is a realization from a mixture distribution $p(\bm x) = \sum_{k=1}^K p(\bm x | \omega_k) p(\omega_k)$, where $p(\bm x | \omega_k)$ is the probability density function (PDF) of the $k$th class, $p(\omega_k)$ is prior probability of class membership of the $k$th class.

The \emph{QDA} classifier is the optimal Bayesian decision rule with respect to a $0-1$ loss function when $p(\bm x | \omega_k)$ is the PDF of the multivariate normal distribution with known mean vector  $\bm\mu_k \in \mathbb{R}_{p \times 1}$ and known covariance matrix $\bm\Sigma_k \in \mathbb{R}_{p \times p}^{>}$, $k = 1, 2, \ldots, K$, where $\mathbb{R}_{p \times p}^{>}$ denotes the cone of real $p \times p$ positive-definite matrices. Because the parameters are typically unknown, we estimate the unknown parameters $\bm \mu_k$ and $\bm\Sigma_k$ with their maximum likelihood estimators (MLEs) and substitute the MLEs into the \emph{QDA} classifier. Assuming, for simplicity, that the prior probabilities of class membership $p(\omega_k)$ are equal for $k = 1, \ldots, K$, we assign an unlabeled observation $\bm x$ to class $\omega_k$ using the sample \emph{QDA} classifier:
\begin{align}
	D_{QDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\widehat{\bm\Sigma}_k^{-1}(\bm x - \xbar_k)  + \ln |\widehat{\bm\Sigma}_k|, \label{eq:qda}
\end{align}
where $\xbar_k$ is the MLE for $\bm \mu_k$ and  $\widehat{\bm\Sigma}_k$ is the MLE for $\bm \Sigma_k$. Notice that if we assume further that the covariance matrix parameters are equal for each class (i.e., $\bm\Sigma_k = \bm\Sigma$ for all $k$), then \eqref{eq:qda} reduces to Fisher's linear discriminant analysis (LDA) classifier,
\begin{align}
	D_{LDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\widehat{\bm\Sigma}^{-1}(\bm x - \xbar_k), \label{eq:lda}
\end{align}
where $\widehat{\bm\Sigma}$ is the pooled sample covariance matrix
\begin{align}
	\widehat{\bm\Sigma} = \frac{1}{n} \sum_{k=1}^K n_k \widehat{\bm\Sigma}_k. \label{eq:pooled-cov}
\end{align}

In the small-sample, high-dimensional setting, if $p > n_k$ for any $k$, \eqref{eq:qda} is incalculable because the sample covariance matrix $\widehat{\bm\Sigma}_k$ for class $\omega_k$ is singular, and, thus, its inverse $\widehat{\bm\Sigma}_k^{-1}$ does not exist.

Although \eqref{eq:qda} is in a standard form, we prefer to express the \emph{QDA} classifier explicitly in terms of the eigenvalues and eigenvectors of $\widehat{\bm \Sigma}_k$. Let $e_{jk}$ denote the $j$th largest eigenvalue of $\widehat{\bm\Sigma}_k$ and $\bm v_{jk}$ be the corresponding eigenvector such that $e_{1k} \ge e_{2k} \ge \ldots \ge e_{pk} \ge 0$. Equivalently, we write \eqref{eq:qda} as
\begin{align}
  D(\bm x) = \argmin_{k} \sum_{j = 1}^p \frac{[\bm v_{jk}' (\bm x - \bm \xbar_k)]^2}{e_{jk}} + \sum_{j=1}^p \ln e_{jk}. \label{eq:qda-eigen}
\end{align}
The smallest eigenvalues and the directions associated with their eigenvectors highly influence \eqref{eq:qda-eigen} . In fact, the eigenvalues of $\widehat{\bm \Sigma}_k$ are well-known to be biased if $p \ge n_k$ such that the smallest eigenvalues are underestimated \citep{Seber:2004uh}. This bias increases as $p$ increases relative to $n_k$ \citep{TODONeedCitation}. Moreover, if $p > n_k$, then rank$(\widehat{\bm \Sigma}_k) \le n_k$, which implies that the last $p - n_k$ eigenvalues of $\widehat{\bm \Sigma}_k$ are 0, which implies that \eqref{eq:qda-eigen} is incalculable. Thus, although more feature information is available to discriminate among the $K$ classes as $p$ increases, classification accuracy decreases unless one obtains enough training-sample observations to reliably estimate the increased number of parameters.

Several regularization methods, such as the methods considered by \citet*{Guo:2007te}, \cite{Mkhadri:1995jp}, and \citet*{Xu:2009fl}, have been proposed in the literature to stabilize the eigenvalues of $\widehat{\bm \Sigma}_k$ used in \eqref{eq:qda}. These methods often can be written concisely as a matrix function, which we define here using the notation of \cite{Izenman:2008gm}. Let $\bm A \in \mathbb{R}_{p \times p}$ be symmetric, and let $\phi:\mathbb{R} \rightarrow \mathbb{R}$. Then
\begin{align}
	\phi(\bm A) = \sum_{j = 1}^p \phi(e_j) \bm v_j \bm v_j'\label{eq:spectral-decomp}
\end{align}
is a matrix function defined explicitly in terms of the eigenvalues of $\bm A$. For this reason, $\phi$ is often called a spectral transform. Special case of \eqref{eq:spectral-decomp} include the spectral decomposition of $\bm A$ when $\phi(e_j) = e_j$, the matrix square-root of $\bm A$ when $\phi(e_j) = e_j^{1/2}$, and, if it exists, the inverse of $\bm A$ when $\phi(e_j) = e_j^{-1}$ \citep{Harville:2008wja}.

Viewing regularized covariance-matrix estimators as matrix functions, we can write a more general form of the \emph{QDA} classifier as
\begin{align}
  D(\bm x) = \argmin_{k} \sum_{j = 1}^p \phi(e_{jk})[\bm v_{jk}' (\bm x - \bm \mu_k)]^2 + \sum_{j=1}^p \ln \{ \phi(e_{jk}) \}. \label{eq:generalized-qda}
\end{align}
The generalized \emph{QDA} classifier in \eqref{eq:generalized-qda} includes several common regularized classifiers. Notice that if  $\phi(e_{jk}) = e_{jk}$, $k = 1, \ldots, K$, \eqref{eq:generalized-qda} reduces to the \emph{QDA} classifier in \eqref{eq:qda-eigen}. Furthermore, if $\phi(e_{jk}) = e_{j}$, $k = 1, \ldots, K$, \eqref{eq:generalized-qda} reduces to the \emph{LDA} classifier in \eqref{eq:lda}, and the second summand in \eqref{eq:generalized-qda} is constant across all classes and can be omitted. Additionally, one common approach is to ignore the zero eigenvalues when $p > n$ by employing the Moore-Penrose pseudoinverse $\widehat{\bm \Sigma}_k^{+}$, which employs the matrix function $\phi(e_j) = e_j^{+}$, where $z^{+} = 1/z$ if $z > 0$ and $0$ otherwise. As noted by \cite{Hoyle:2011vt}, \cite{Raudys:1998dd}, and \cite{Ramey:2011ji}, the pseudoinverse is often inadvisable because the resulting classifier often yields a sharp decrease in classification accuracy if $p \approx n$.

Another common form of the covariance-matrix regularization applies a shrinkage factor $\gamma_k > 0$, so that
\begin{align}
	\phi(e_{jk}) = (e_{jk} + \gamma_k)^{-1}, \label{eq:ridge-estimator}
\end{align}
is a ridge-like estimator, similar to a method employed in ridge regression. The estimator in \eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance matrix $\widehat{\bm\Sigma}_k$ towards the $p$-dimensional identity matrix $\bm I_p$, thereby increasing the eigenvalues of $\widehat{\bm\Sigma}_k$ by $\gamma_k$ so that they are positive, resulting in $\widehat{\bm \Sigma}_k(\gamma_k)$ being positive definite.

We emphasize the regularized estimator in \eqref{eq:ridge-estimator} because the \emph{RDA} and \emph{MDEB} classifiers can be written in this form as we show below. Also, we present the \emph{NLDA} and \emph{ETRDA} classifiers in terms of matrix functions using \eqref{eq:generalized-qda}. For additional covariance-matrix regularization methods, see \cite{Ye:2009gd} and \cite{Ramey:2011ji}.

\subsection{Regularized Discriminant Analysis}

\cite{Friedman:1989tm} has extended the shrinkage technique by first computing a
weighted average of the sample covariance matrix $\widehat{\bm \Sigma}_k$ for class
$\omega_k$ and the pooled sample covariance matrix $\widehat{\bm\Sigma}$ to
estimate the covariance matrix for class $\omega_k$ with $\widehat{\bm\Sigma}_k(\lambda)$,
where $\lambda \in [0, 1]$, $n_k(\lambda) = (1 - \lambda) n_k + \lambda n$, $\bm S_k = n_k \widehat{\bm\Sigma}_k$, $\bm S = \sum_{k=1}^K \bm S_k$, $\bm S_k(\lambda) = (1 - \lambda) \bm S_k + \lambda \bm S$, and
\begin{align}
  \widehat{\bm\Sigma}_k(\lambda) = n_k^{-1}(\lambda) \bm S_k(\lambda).\label{eq:sig-lambda}
\end{align}
We can interpret \eqref{eq:sig-lambda} as a
covariance matrix estimator for class $\omega_k$ that borrows from $\widehat{\bm\Sigma}$ in \eqref{eq:pooled-cov} to better estimate $\bm \Sigma_k$. Notice that for $\lambda = 0$,
\eqref{eq:sig-lambda} is equal to $\widehat{\bm\Sigma}_k$ used in the \emph{QDA} classifier in \eqref{eq:qda}. Similarly, notice that for $\lambda = 1$, \eqref{eq:sig-lambda} is equal to $\widehat{\bm\Sigma}$ used in the \emph{LDA} classifier in \eqref{eq:lda}, in which case we assume implicitly that $\bm\Sigma_k = \bm \Sigma$ for $k = 1, \ldots, K$.

\cite{Friedman:1989tm} also uses a regularization parameter $\gamma$ to shrink the
eigenvalues of \eqref{eq:sig-lambda} towards the $p$-dimensional identity matrix $\bm I_p$
in order to stabilize the inverse of \eqref{eq:sig-lambda}, resulting in a biased
covariance-matrix estimator with matrix function
\begin{align}
	\phi(e_{jk}) = (\{1 - \gamma\} e_{jk} + \gamma \bar{e}_k)^{-1},\label{eq:rda-eigenvalue-function}
\end{align}
for class $\omega_k$, where $\gamma \in [0, 1]$, $e_{jk}$ be the $j$th largest eigenvalue of $\widehat{\bm\Sigma}_k(\lambda)$ in \eqref{eq:sig-lambda}, $\bm v_{jk}$ be the corresponding eigenvector, and $\bar{e}_k = p^{-1} \sum_{j=1}^p e_{jk}$ is the average of the eigenvalues of \eqref{eq:sig-lambda}. In summary, the \emph{pooling} parameter $\lambda$ controls the amount that we borrow from $\widehat{\bm\Sigma}$ to estimate $\bm \Sigma_k$, and the \emph{shrinkage} parameter $\gamma$ determines the amount of applied shrinkage.

\subsection{Minimum Distance Empirical Bayes}

As an alternative shrinkage approach, we first consider the minimum distance empirical Bayes (\emph{MDEB}) classifier from \cite{Srivastava:2007ww}, who incorporate a similar regularized covariance-matrix estimator that shrinks a covariance-matrix estimator towards a scaled identity, where the shrinkage factor is approximately equal to the average of the nonzero eigenvalues of the original estimator. Furthermore, \cite{Srivastava:2007ww} claim that the \emph{MDEB} classifier is best regularization method they have encountered in the literature. When $p \le n$, the shrinkage factors of the \emph{RDA} and \emph{MDEB} classifiers are equal, but for $p > n$ this shrinkage can differ substantially. In fact, \cite{Ramey:2011ji} show empirically that the \emph{MDEB} classifier can yield superior classification accuracy to the \emph{RDA} classifier. Moreover, as the number of features incorporated in the model is increased, the \emph{MDEB} covariance-matrix estimator remains positive definite and is not nearly singular like the \emph{RDA} classifier. However, for $p \gg n$, the eigenvalue adjustment employed in the \emph{MDEB} covariance-matrix estimator can be too large, thereby dominating the distance of an unlabeled observation to the sample mean.

\cite{Srivastava:2007ww} derive an empirical-Bayes covariance-matrix estimator under the assumption that the data are independent and identically distributed multivariate normal observations. \cite{Srivastava:2007ww} then substitute the estimator into the \eqref{eq:lda} to attain the \emph{MDEB} classifier. To derive the empirical-Bayes covariance-matrix estimator, the authors first assume that $\bm \Sigma_k^{-1}$ follows a Wishart distribution \emph{a priori} with mean $\lambda^{-1} \bm I_p$, $\lambda > 0$, and degrees of freedom, $l \ge p$. The resulting posterior estimator for $\bm \Sigma_k^{-1}$ is $\frac{n_k + l}{n_k}(n_k^{-1}\lambda \bm I_p + \widehat{\bm \Sigma}_k)^{-1}$. Using an empirical Bayes argument to estimate $\lambda$ for $p > n_k$, \cite{Srivastava:2007ww} propose an empirical-Bayes covariance-matrix estimator with matrix function
\begin{align}
	\phi(e_{jk}) = \left(e_{jk} + n_k^{-1} \sum_{j = 1}^p e_{jk}\right)^{-1},\label{eq:cov-mdeb}
\end{align}
where the shrinkage factor in \eqref{eq:cov-mdeb} is approximately equal to the average of the nonzero eigenvalues of $\widehat{\bm \Sigma}_k$. Substituting \eqref{eq:cov-mdeb} into \eqref{eq:generalized-qda}, we have the \emph{MDEB} classifier. We note that \cite{Srivastava:2007ww} set assumed that $\bm \Sigma_k = \bm \Sigma$ so that $\phi(e_{jk}) = \phi(e_k)$, $k = 1, \ldots, K$. We will return to this assumption below.

\subsection{New LDA}

We also consider the covariance-matrix shrinkage method from \cite{Thomaz:2006ef}, who propose the New \emph{LDA} (\emph{NLDA}) classifier. \cite*{Xu:2009fl} and \cite{Ramey:2011ji} demonstrate that the \emph{NLDA} classifier can attain classification accuracies comparable to the \emph{RDA} and \emph{MDEB} classifiers for small-sample, high-dimensional microarray data sets. These results are surprising because the \emph{NLDA} covariance matrix estimator employs a simple eigenvalue-replacement method, where any eigenvalue smaller than the average of the eigenvalues is replaced with the average of the eigenvalues.

Whereas shrinkage methods often have the form given in \eqref{eq:ridge-estimator}, the covariance matrix estimator employed in the \emph{NLDA} classifier stabilizes the inverse by replacing the smallest eigenvalues of \eqref{eq:pooled-cov} with the average of the eigenvalues of \eqref{eq:pooled-cov}. Let $e_{jk}$ denote the $j$th largest eigenvalue of $\widehat{\bm \Sigma}_k$, and let $\bar{e}_k$ be the average of the eigenvalues of \eqref{eq:pooled-cov}. Then, the inverse of the \emph{NLDA} covariance-matrix estimator has the matrix function
\begin{align}
	\phi(e_{jk}) = (\bar{e}_k + \{e_j - \bar{e}_k\}_+)^{-1},\label{eq:cov-nlda}
\end{align}
with $z_+ = z$ if $z > 0$ and 0 otherwise. Substituting \eqref{eq:cov-nlda} into \eqref{eq:generalized-qda}, we have the \emph{NLDA} classifier. Similar to the \emph{MDEB} classifier, \cite{Thomaz:2006ef} assume that $\bm \Sigma_k = \bm \Sigma$ so that $\phi(e_{jk}) = \phi(e_k)$, $k = 1, \ldots, K$.

\section{Regularized Discriminant Analysis with Eigenvalue Thresholding}

\subsection{The Hard-ETRDA Classifier}

Although the shrinkage applied in \eqref{eq:cov-nlda} is ineffective, the form in \eqref{eq:cov-nlda} allows us to consider a more general eigenvalue thresholding approach to resolve the numerical issues of the \emph{RDA} classifier discussed below. We are interested in a general shrinkage value $\Delta$ instead of the average of the eigenvalues $\bar{e}_k$. We propose to shrink the pooled covariance matrix estimator in \eqref{eq:sig-lambda} with a more general eigenvalue threshold, motivated by the \emph{NLDA} classifier. That is, for $\Delta_0, \Delta > 0$, we define the matrix function
\begin{align}
	\phi(e_{jk}) = (\Delta_0 + e_{jk} \cdot I\{e_{jk} > \Delta\} )^{+},\label{eq:cov-hard-etrda}
\end{align}
where $e_{jk}$ are the eigenvalues of $\widehat{\bm \Sigma}_k(\lambda)$ in \eqref{eq:sig-lambda} and $\bm v_{jk}$ are the corresponding eigenvectors. Substituting \eqref{eq:cov-soft-etrda} into \eqref{eq:generalized-qda}, we arrive at the \emph{ETRDA} classifier.

For $\Delta_0 = \Delta = \gamma$, the shrinkage employed in \eqref{eq:cov-hard-etrda} is equal to \eqref{eq:ridge-estimator}. Hence, \eqref{eq:cov-hard-etrda} includes the regularization methods of \cite{Friedman:1989tm} and \cite{Srivastava:2007ww} as special cases. To see this, notice that if $\gamma$ is the average of the eigenvalues, \eqref{eq:cov-hard-etrda} is equal to \eqref{eq:rda-eigenvalue-function}. Furthermore, if $\gamma$ is the average of the nonzero eigenvalues \eqref{eq:cov-hard-etrda} is equal to \eqref{eq:cov-mdeb}. Notice also that if $\Delta_0 = \Delta = 0$, then \eqref{eq:cov-hard-etrda} reduces to the Moore-Penrose pseudoinverse \citep{Harville:2008wja}.

\subsection{The Soft-ETRDA Classifier}

As an alternative eigenvalue threshold, we consider a soft threshold. Although the shrinkage applied in \eqref{eq:cov-nlda} is ineffective, the form in \eqref{eq:cov-nlda} allows us to consider a more general eigenvalue thresholding approach to resolve the numerical issues of the \emph{RDA} classifier discussed below. We are interested in a general shrinkage value $\Delta$ instead of the average of the eigenvalues $\bar{e}_k$. We propose to shrink the pooled covariance matrix estimator in \eqref{eq:sig-lambda} with a more general eigenvalue threshold, motivated by the \emph{NLDA} classifier. That is, for $\Delta > 0$, we define the matrix function
\begin{align}
	\phi(e_{jk}) = (\Delta_0 + \{e_{jk} - \Delta\}_+)^{+},\label{eq:cov-soft-etrda}
\end{align}
where $e_{jk}$ are the eigenvalues of $\widehat{\bm \Sigma}_k(\lambda)$ in \eqref{eq:sig-lambda} and $\bm v_{jk}$ are the corresponding eigenvectors. Substituting \eqref{eq:cov-soft-etrda} into \eqref{eq:generalized-qda}, we arrive at the \emph{ETRDA} classifier.

For small values of $\gamma$ such that $\Delta_0 = \Delta = \gamma$, the shrinkage employed in \eqref{eq:cov-soft-etrda} is approximately equal to \eqref{eq:ridge-estimator} because the near-zero are adjusted by $\gamma$, while the larger eigenvalues are hardly adjusted by $\gamma$. Because of this approximation, the regularization methods of \cite{Friedman:1989tm} and \cite{Srivastava:2007ww} are approximately special cases of the \eqref{eq:cov-soft-etrda}. To see this, notice that if $\gamma$ is the average of the eigenvalues, \eqref{eq:cov-soft-etrda} is equal to \eqref{eq:cov-nlda}, which is approximately equal to \eqref{eq:rda-eigenvalue-function}. Furthermore, if $\gamma$ is the average of the nonzero eigenvalues \eqref{eq:cov-soft-etrda} is approximately equal to \eqref{eq:cov-mdeb}.

Notice that \eqref{eq:cov-soft-etrda} has a form similar to the soft thresholding employed in the nearest shrunken centroids classifier from \cite{Tibshirani:2002ht} in the case that the matrix of eigenvectors of $\widehat{\bm \Sigma}_k$ is the identity matrix and $\Delta_0 = 0$. Our choice of $\Delta_0 = 0$ is also advantageous because it can reduce the number of dimensions included in the \emph{Soft-ETRDA} classifier, thereby decreasing the classification reducing. Noting the similarity of \eqref{eq:cov-soft-etrda} to the soft threshold of \cite{Tibshirani:2002ht}, we follow \cite{Tibshirani:2002ht} and select $\Delta$ by cross-validation, resulting in a superior classifier to the \emph{RDA} classifier proposed by \cite{Friedman:1989tm}.

For small values of $\gamma$, the eigenvalue thresholding employed in the \emph{Soft-ETRDA} and \emph{Hard-ETRDA} classifiers are approximately equal. Both methods are effective regularization methods in stabilizing the smallest eigenvalues. The soft eigenvalue threshold employed in the \emph{Soft-ETRDA} classifier has the distinct advantage that the largest eigenvalues are reduced, which smooths the distribution of sample eigenvalues, yielding an effective classifier that is often superior to the proposed sparse and regularized classifiers considered.

(\textbf{TODO: Emphasize that for small $\Delta$, Soft and Hard are approximately equal. This provides a connection between NLDA and RDA.})

\section{Monte Carlo Simulations}

TODO

\section{Discussion}

It is interesting to note that our approach is similar to an aside made by Friedman (1989) in the case that the RDA covariance matrix estimator is singular. Friedman (1989) recommended that the zero eigenvalues be replaced with a small value such as 0.001. However, in Theorem 1, we saw that even small adjustments can be inappropriate because they result in ineffective shrinkage.

\section{Next Steps}

\subsection{Simulations}

\begin{itemize}
	\item  Design an experiment similar to Friedman's initial simulations, where the the majority of the variances are nearly 0. Fix the sample size. Consider the error rate as a function of $p$. This should cause the RDA regularization to be ineffective because the average eigenvalue is near 0.

	\item  Design an experiment where the average eigenvalue goes to 0, but the average nonzero eigenvalue blows up for a fixed sample size. Consider the error rate as a function of $p$. This should cause the RDA and MDEB regularizations to be ineffective. However, I am banking on the ETRDA methods doing well.

	\item In our simulations, we need to report the estimated values of $c$, the average of the eigenvalues, and the average of the nonzero eigenvalues.

\end{itemize}

\subsection{Notation}

\begin{itemize}
	\item Define the majority of notation at the beginning of Section 2, including $z^+$, $z_+$, and $I[z]$.
	\item Define the generalization of RDA in terms of the eigenvectors and eigenvalues of the RDA estimator using $\lambda$ as an index. This should remove the need to redefine the eigendecomposition each time we change regularization methods. A possible formulation is $\bm v_{jk}^{(\lambda)}$ and $e_{jk}^{(\lambda)}$.
	\item In \eqref{eq:generalized-qda} we have a slight issue in the formulation of the log-determinant in the second sum. The eigenvalue function $\phi$ in the first sum should be the inverse of the function in the second sum. Furthermore, we need to define the log-determinant so that the last $p - q$ terms are ignored when we reduce the dimension to $q$.
\end{itemize}

\subsection{Integrate Section 4 Better}

We should create another figure that shows the shrinkage factor as a function of $q$ for each of the shrinkage factors using one or two data microarray data sets with the Dudoit variable selection method. Show that as more features are incorporated into the model, the competing shrinkage methods fail.

Moreover, show the determinants or log-determinants of each covariance matrix. Then discuss the near-equivalence of the shrinkage in both NLDA and RDA.

Perhaps, we should not consider the NLDA classifier at all. Alternatively, we recall simply the regularization method of the authors and use that as motivation for the Soft-ETRDA classifier.

\subsection{ETRDA}

\begin{itemize}
	\item Look at the shrinkage factors as a function of $q$. We may find that the optimal value for $c$ is approximately equal to the average of the nonzero eigenvalues. If this is the case, we could note in the discussion that to improve the speed, simply use the average of the nonzero eigenvalues.
	\item Investigate a quick way to calculate the ETRDA classifier for each value of $c$, similar to RDA. For instance, use the Fast SVD with Hard-ETRDA and write enough squiggles to explain this thoroughly.
\end{itemize}

\subsection{Discuss Replacing RDA Shrinkage with MDEB Shrinkage}

The next step could be to discuss replacing the RDA shrinkage with the MDEB shrinkage, which uses essentially the average nonzero eigenvalues rather than the average eigenvalues. However, as DMY and I have discovered, the average of the nonzero eigenvalues often can be too large, resulting in extremely inflated denominators in \eqref{eq:generalized-qda}. This has the negative effect of decreasing the scaled differences in the means. Show a histogram of the eigenvalues for a data set, such as the Alon data set. Draw vertical lines showing the average of the 1) eigenvalues and 2) nonzero eigenvalues.

Discuss that the MDEB approach to regularizing the RDA estimator is inappropriate. While it works for the linear case, it does not for the quadratic case because of the determinant. Because the average of the nonzero eigenvalues is often very large, the last $p - q$ terms in the sum in \eqref{eq:generalized-qda} are approximately 0 (perhaps, not quite 0). In the linear case, the second sum is ignored, so that the MDEB estimator results in a classifier very similar to the one where the M-P pseudo inverse is used. However, the rank does not need to be estimated using the MDEB estimator, and the terms in the first sum with eigenvalues near 0 are no longer stable.

Previously, I investigated applying MDEB to the QDA classifier as well but was discouraged because the classification accuracy was quite poor. It seems that the second sum in \eqref{eq:generalized-qda} is likely to be blamed. Letting rank$(\widehat{\bm\Sigma}_k) = q_k$ and $\gamma_k$ be the average of the nonzero eigenvalues of $\widehat{\bm \Sigma}_k$, the second sum in \eqref{eq:generalized-qda} can be partitioned as
\begin{align*}
	\sum_{j=1}^p \ln (e_{jk} + \gamma_k) &= \sum_{j=1}^{q_k} \ln (e_{jk} + \gamma) + \sum_{j=q_k + 1}^p \ln (\gamma_k)\\
	&= \sum_{j=1}^{q_k} \ln (e_{jk} + \gamma_k) + (p - q_k - 1) \ln (\gamma_k).
\end{align*}
My guess is that $(p - q_k - 1) \ln (\gamma_k)$ tends to dominate for the class with either the largest or the smallest sample size. Thus, the shrinkage has too much influence on the classification. With this in mind, our proposed selection of $c$ may yield a significant improvement to the QDA classifier compared with MDEB applied to the QDA classifier.

While the \emph{RDA} classifier suffers from numerical incalculability as the data dimensionality $p$ increases, the shrinkage factor of the \emph{MDEB} classifier tends to dominate the estimator. Because the trace of the covariance-matrix estimator is equal to the sum of its eigenvalues, the average of the nonzero eigenvalues can become very large. For each feature added, the variance of that feature must be positive, so it adds to the trace.  Even if the average of the nonzero eigenvalues yields a convergent series, the shrinkage factor is substantially large.

\bibliographystyle{plainnat}
\bibliography{rda}

\end{document}

### 2013-01-22 lib/helpers.R

@@ -1,0 +1,103 @@
+#' Generates the Cartesian product of all simulation experiment factors.
+#' Then, replicates the Cartesian product a specified number of times.
+#' This function is especially useful for simulations run with parallel
+#' processing, where we need to run the same set of simulation configurations
+#' multiples times.
+#'
+#' If num_reps = 1, the output is nearly identical to that of the 'expand.grid' function.
+#' In fact, much of the documentation for this function has been copied
+#' directly from the 'expand.grid' function.
+#'
+#' @param ... vectors, factors or a list containing these.
+#' @param num_reps the number of times to replicate each simulation configuration
+#' @param stringsAsFactors logical specifying if character vectors are converted
+#'   to factors.
+#' @return A data frame containing D rows for each combination of the supplied factors.
+#'  The first factors vary fastest. The columns are labelled by the factors if these
+#'  are supplied as named arguments or named components of a list. The row names are
+#'  'automatic.'
+sim_factors <- function(..., num_reps = 1, stringsAsFactors = FALSE) {
+  grid <- expand.grid(..., stringsAsFactors = stringsAsFactors)
+  do.call(rbind, replicate(num_reps, grid, simplify = FALSE))
+}
+
+#' Helper function to return predictions from the list element 'class'
+#'
+#' A common convention for classifiers in R is to return a list from the
+#' 'predict' with the classifications (class labels) of each test observation
+#' along with a variety of information about each specific test
+#' classification, such as discriminant scores and posterior probabilities.
+#' For our 'classify' function, we need only the class labels. This helper
+#' function is a wrapper function that returns only the class labels when
+#' the aforementioned convention is followed.
+predict_helper <- function(obj, newdata) {
+  predict(obj, newdata)$class
+}
+
+# This function rounds to the nearest digits and does not truncated leading
+# zeroes, unlike the 'round' function.
+round_digits <- function(x, num_digits = 3) {
+  sprintf(paste("%.", num_digits, "f", sep = ""), round(x, num_digits))
+}
+
+
+# Classifier from Witten and Tibshirani (2011) - JRSS B
+Witten_Tibshirani <- function(train_x, train_y, test_x) {
+  # Uses cross-validation to select the best lambda and K
+  cv_out <- PenalizedLDA.cv(x = train_x, y = as.integer(train_y))
+
[…]

### 2013-01-22 shrinkage.r

@@ -1,0 +1,33 @@
+library('ProjectTemplate')
+load.project()
+
+data(alon)
+q_vals <- seq.int(50, 2000, by = 50)
+
+log_det <- lapply(q_vals, function(q) {
+  x <- with(alon, dudoit(x, y, q = q))$train_x
+  Sigma <- cov_pool(x = x, y = alon$y)
+  Sigma_eigenvalues <- eigen(Sigma, symmetric = TRUE, only.values = TRUE)$values
+
+  shrinkage_rda <- mean(Sigma_eigenvalues)
+  shrinkage_mdeb <- sum(Sigma_eigenvalues) / length(alon$y)
+  
+  eigen_rda <- Sigma_eigenvalues + shrinkage_rda
+  eigen_mdeb <- Sigma_eigenvalues + shrinkage_mdeb
+  eigen_nlda <- replace(Sigma_eigenvalues, Sigma_eigenvalues < shrinkage_rda, shrinkage_rda)
+  
+  log_determinants <- list(MDEB = sum(log(eigen_mdeb)),
+                           NLDA = sum(log(eigen_nlda)),
+                           RDA = sum(log(eigen_rda)),
+                           q = q)
+  log_determinants
+})
+log_det <- do.call(rbind.data.frame, log_det)
+m_log_det <- melt(log_det, id.vars = 'q', variable.name = 'Classifier',
+                  value.name = 'Log_Determinant')
+
+# Log-determinants of the covariance matrices for each of the three shrinkage methods
+# We used the pooled sample covariance matrix
+# In the figure caption, say that this corresponds to $\lambda = 1$
+p <- ggplot(m_log_det, aes(x = q, y = Log_Determinant, group = Classifier, color = Classifier))
+p + geom_line(aes(linetype = Classifier))

### 2013-01-22 sim-microarray.r

@@ -1,0 +1,67 @@
+library('ProjectTemplate')
+load.project()
+
+set.seed(42)
+num_cores <- 1
+
+data_sets <- c("burczynski", "christensen", "khan", "nakayama", "sorlie", "su", "sun", "yeoh")
+
+
+results <- mclapply(data_sets, function(data_set) {
+  data <- load_microarray(data_set)
+  data$x <- as.matrix(data$x)
+
+  cv_folds <- cv_partition(y = data$y, num_folds = 10)
+
+  cv_results <- lapply(cv_folds, function(cv_fold) {
+    train_x <- as.matrix(data$x[cv_fold$training, ])
+    train_y <- data$y[cv_fold$training]
+    test_x <- as.matrix(data$x[cv_fold$test, ])
+    test_y <- data$y[cv_fold$test]
+
+    # Sets the prior probabilities equal
+    num_classes <- nlevels(train_y)
+    prior_probs <- rep(1, num_classes) / num_classes
+
+    # Witten and Tibshirani (2011) - JRSS B
+    Witten_out <- Witten_Tibshirani(train_x, train_y, test_x)
+    Witten_errors <- sum(Witten_out$predictions != test_y)
+
+    # Applies the variable selection from Witten and Tibshirani
+    train_x <- train_x[, Witten_out$variables]
+    test_x <- test_x[, Witten_out$variables]
+
+    # Hard-ETRDA
+    cv_out <- etrda_cv(x = train_x, y = train_y, threshold = "hard", num_delta = 10, prior = prior_probs)
+    etrda_out <- with(cv_out, etrda(x = train_x, y = train_y, threshold = "hard", lambda = lambda, delta = delta, prior = prior_probs))
+    etrda_hard_errors <- sum(predict(etrda_out, test_x)$class != test_y)
+
+    # RDA from Friedman (1989)
+    cv_out <- rda_cv(x = train_x, y = train_y, prior = prior_probs)
+    rda_out <- with(cv_out, rda(x = train_x, y = train_y, lambda = lambda, gamma = gamma, prior = prior_probs))
+    rda_errors <- sum(predict(rda_out, test_x)$class != test_y)
+    
+    # DLDA and DQDA
+    dlda_errors <- sum(predict(dlda(x = train_x, y = train_y, prior = prior_probs), test_x)$class != test_y)
+    dqda_errors <- sum(predict(dqda(x = train_x, y = train_y, prior = prior_probs), test_x)$class != test_y)
+
+    list(
[…]

### 2013-01-25 sim-microarray.r
Removed RDA classifier from simulation.

@@ -2,9 +2,9 @@ library('ProjectTemplate')
 load.project()
 
 set.seed(42)
-num_cores <- 1
+num_cores <- 7
 
-data_sets <- c("burczynski", "christensen", "khan", "nakayama", "sorlie", "su", "sun", "yeoh")
+data_sets <- c("burczynski", "christensen", "khan", "nakayama", "sorlie", "su", "yeoh")
 
 
 results <- mclapply(data_sets, function(data_set) {
@@ -27,20 +27,15 @@ results <- mclapply(data_sets, function(data_set) {
     Witten_out <- Witten_Tibshirani(train_x, train_y, test_x)
     Witten_errors <- sum(Witten_out$predictions != test_y)
 
-    # Applies the variable selection from Witten and Tibshirani
+    # Applies the variable selection from Witten and Tibshirani (2011)
     train_x <- train_x[, Witten_out$variables]
     test_x <- test_x[, Witten_out$variables]
 
-    # Hard-ETRDA
-    cv_out <- etrda_cv(x = train_x, y = train_y, threshold = "hard", num_delta = 10, prior = prior_probs)
-    etrda_out <- with(cv_out, etrda(x = train_x, y = train_y, threshold = "hard", lambda = lambda, delta = delta, prior = prior_probs))
-    etrda_hard_errors <- sum(predict(etrda_out, test_x)$class != test_y)
+    # ETRDA
+    cv_out <- etrda_cv(x = train_x, y = train_y, prior = prior_probs)
+    etrda_out <- with(cv_out, etrda(x = train_x, y = train_y, lambda = lambda, delta = delta, prior = prior_probs))
+    etrda_errors <- sum(predict(etrda_out, test_x)$class != test_y)
 
-    # RDA from Friedman (1989)
-    cv_out <- rda_cv(x = train_x, y = train_y, prior = prior_probs)
-    rda_out <- with(cv_out, rda(x = train_x, y = train_y, lambda = lambda, gamma = gamma, prior = prior_probs))
-    rda_errors <- sum(predict(rda_out, test_x)$class != test_y)
-    
     # DLDA and DQDA
     dlda_errors <- sum(predict(dlda(x = train_x, y = train_y, prior = prior_probs), test_x)$class != test_y)
     dqda_errors <- sum(predict(dqda(x = train_x, y = train_y, prior = prior_probs), test_x)$class != test_y)
[…]

### 2013-02-05 reports/paper-etrda.tex
Added Caleb to authors list. Added data set descriptions. Also, shortened the description of the other classifiers.

@@
 
-\author{John A. Ramey and Dean M. Young}
+\author{John A. Ramey, Caleb K. Stein, and Dean M. Young}
 
@@
 
-\subsection{Minimum Distance Empirical Bayes}
-
-As an alternative shrinkage approach, we first consider the minimum distance empirical Bayes (\emph{MDEB}) classifier from \cite{Srivastava:2007ww}, who incorporate a similar regularized covariance-matrix estimator that shrinks a covariance-matrix estimator towards a scaled identity, where the shrinkage factor is approximately equal to the average of the nonzero eigenvalues of the original estimator.
-Furthermore, \cite{Srivastava:2007ww} claim that the \emph{MDEB} classifier is best regularization method they have encountered in the literature.
-When $p \le n$, the shrinkage factors of the \emph{RDA} and \emph{MDEB} classifiers are equal, but for $p > n$ this shrinkage can differ substantially.
-In fact, \cite{Ramey:2011ji} show empirically that the \emph{MDEB} classifier can yield superior classification accuracy to the \emph{RDA} classifier.
-Moreover, as the number of features incorporated in the model is increased, the \emph{MDEB} covariance-matrix estimator remains positive definite and is not nearly singular like the \emph{RDA} classifier.
-However, for $p \gg n$, the eigenvalue adjustment employed in the \emph{MDEB} covariance-matrix estimator can be too large, thereby dominating the distance of an unlabeled observation to the sample mean.
-
-\cite{Srivastava:2007ww} derive an empirical-Bayes covariance-matrix estimator under the assumption that the data are independent and identically distributed multivariate normal observations.
-\cite{Srivastava:2007ww} then substitute the estimator into the \eqref{eq:lda} to attain the \emph{MDEB} classifier.
-To derive the empirical-Bayes covariance-matrix estimator, the authors first assume that $\bm \Sigma_k^{-1}$ follows a Wishart distribution \emph{a priori} with mean $\lambda^{-1} \bm I_p$, $\lambda > 0$, and degrees of freedom, $l \ge p$.
-The resulting posterior estimator for $\bm \Sigma_k^{-1}$ is $\frac{n_k + l}{n_k}(n_k^{-1}\lambda \bm I_p + \widehat{\bm \Sigma}_k)^{-1}$.
-Using an empirical Bayes argument to estimate $\lambda$ for $p > n_k$, \cite{Srivastava:2007ww} propose an empirical-Bayes covariance-matrix estimator with matrix function
-\begin{align}
-	\phi(e_{jk}) = \left(e_{jk} + n_k^{-1} \sum_{j = 1}^p e_{jk}\right)^{-1},\label{eq:cov-mdeb}
-\end{align}
-where the shrinkage factor in \eqref{eq:cov-mdeb} is approximately equal to the average of the nonzero eigenvalues of $\widehat{\bm \Sigma}_k$.
-Substituting \eqref{eq:cov-mdeb} into \eqref{eq:generalized-qda}, we have the \emph{MDEB} classifier.
-We note that \cite{Srivastava:2007ww} set assumed that $\bm \Sigma_k = \bm \Sigma$ so that $\phi(e_{jk}) = \phi(e_k)$, $k = 1, \ldots, K$.
-We will return to this assumption below.
-
-\subsection{New LDA}
-
-We also consider the covariance-matrix shrinkage method from \cite{Thomaz:2006ef}, who propose the New \emph{LDA} (\emph{NLDA}) classifier.
-\cite*{Xu:2009fl} and \cite{Ramey:2011ji} demonstrate that the \emph{NLDA} classifier can attain classification accuracies comparable to the \emph{RDA} and \emph{MDEB} classifiers for small-sample, high-dimensional microarray data sets.
-These results are surprising because the \emph{NLDA} covariance matrix estimator employs a simple eigenvalue-replacement method, where any eigenvalue smaller than the average of the eigenvalues is replaced with the average of the eigenvalues.
-
-Whereas shrinkage methods often have the form given in \eqref{eq:ridge-estimator}, the covariance matrix estimator employed in the \emph{NLDA} classifier stabilizes the inverse by replacing the smallest eigenvalues of \eqref{eq:pooled-cov} with the average of the eigenvalues of \eqref{eq:pooled-cov}.
-Let $e_{jk}$ denote the $j$th largest eigenvalue of $\widehat{\bm \Sigma}_k$, and let $\bar{e}_k$ be the average of the eigenvalues of \eqref{eq:pooled-cov}.
-Then, the inverse of the \emph{NLDA} covariance-matrix estimator has the matrix function
-\begin{align}
-	\phi(e_{jk}) = (\bar{e}_k + \{e_j - \bar{e}_k\}_+)^{-1},\label{eq:cov-nlda}
-\end{align}
-with $z_+ = z$ if $z > 0$ and 0 otherwise.
-Substituting \eqref{eq:cov-nlda} into \eqref{eq:generalized-qda}, we have the \emph{NLDA} classifier.
-Similar to the \emph{MDEB} classifier, \cite{Thomaz:2006ef} assume that $\bm \Sigma_k = \bm \Sigma$ so that $\phi(e_{jk}) = \phi(e_k)$, $k = 1, \ldots, K$.
-
 \section{Regularized Discriminant Analysis with Eigenvalue Thresholding}
@@
 
+\section{Special Cases}
+
+\subsection{Minimum Distance Empirical Bayes}
+
+\cite{Srivastava:2007ww} derive an empirical-Bayes covariance-matrix estimator under the assumption that the data are independent and identically distributed multivariate normal observations.
+\cite{Srivastava:2007ww} then substitute the estimator into the \eqref{eq:lda} to attain the minimum distance empirical Bayes (\emph{MDEB}) classifier.
+To derive the empirical-Bayes covariance-matrix estimator, \cite{Srivastava:2007ww} first assume that $\bm \Sigma_k^{-1}$ follows a Wishart distribution \emph{a priori} with mean $\lambda^{-1} \bm I_p$, $\lambda > 0$, and degrees of freedom, $l \ge p$.
+The resulting posterior estimator for $\bm \Sigma_k^{-1}$ is $\frac{n_k + l}{n_k}(n_k^{-1}\lambda \bm I_p + \widehat{\bm \Sigma}_k)^{-1}$.
+Using an empirical Bayes argument to estimate $\lambda$ for $p > n_k$, \cite{Srivastava:2007ww} propose an empirical-Bayes covariance-matrix estimator with matrix function
+\begin{align}
+	\phi(e_{jk}) = \left(e_{jk} + n_k^{-1} \sum_{j = 1}^p e_{jk}\right)^{-1},\label{eq:cov-mdeb}
+\end{align}
+where the shrinkage factor in \eqref{eq:cov-mdeb} is approximately equal to the average of the nonzero eigenvalues of $\widehat{\bm \Sigma}_k$.
+Substituting \eqref{eq:cov-mdeb} into \eqref{eq:generalized-qda}, we have the \emph{MDEB} classifier.
+We note that \cite{Srivastava:2007ww} set assumed that $\bm \Sigma_k = \bm \Sigma$ so that $\phi(e_{jk}) = \phi(e_k)$, $k = 1, \ldots, K$.
+
+\cite{Srivastava:2007ww} claim that the \emph{MDEB} classifier is best regularization method they have encountered in the literature.
+When $p \le n$, the shrinkage factors of the \emph{RDA} and \emph{MDEB} classifiers are equal, but for $p > n$ this shrinkage can differ substantially.
+In fact, \cite{Ramey:2011ji} show empirically that the \emph{MDEB} classifier can yield superior classification accuracy to the \emph{RDA} classifier.
+
+\subsection{New LDA}
+Whereas shrinkage methods often have the form given in \eqref{eq:ridge-estimator}, the covariance matrix estimator from \cite{Thomaz:2006ef} stabilizes the inverse by replacing the smallest eigenvalues of \eqref{eq:pooled-cov} with the average of the eigenvalues of \eqref{eq:pooled-cov}.
+Let $e_{jk}$ denote the $j$th largest eigenvalue of $\widehat{\bm \Sigma}_k$, and let $\bar{e}_k$ be the average of the eigenvalues of \eqref{eq:pooled-cov}.
+Then, the inverse of the covariance-matrix estimator has the matrix function
+\begin{align}
+	\phi(e_{jk}) = (\bar{e}_k + \{e_j - \bar{e}_k\}_+)^{-1},\label{eq:cov-nlda}
+\end{align}
+with $z_+ = z$ if $z > 0$ and 0 otherwise.
+\cite*{Xu:2009fl} and \cite{Ramey:2011ji} demonstrate that the substituting \eqref{eq:cov-nlda} into \eqref{eq:generalized-qda} yields classification accuracies comparable to the \emph{RDA} and \emph{MDEB} classifiers for small-sample, high-dimensional microarray data sets.
+Similar to the \emph{MDEB} classifier, \cite{Thomaz:2006ef} assume that $\bm \Sigma_k = \bm \Sigma$ so that $\phi(e_{jk}) = \phi(e_k)$, $k = 1, \ldots, K$.
+
 \section{Monte Carlo Simulations}
@@
 TODO
+
+\section{Data Sets}
+
+\subsection{\cite{Burczynski:2006ik} Data Set}
+
+\cite{Burczynski:2006ik} acquired the peripheral blood mononuclear cells (PBMC) through hybridization to microarrays from 127 individuals resulting in 22,283 sequences.
+Of the 127 individuals, 42 were healthy, 59 had Crohn's disease (CD), and 26 had ulcerative colitis (UC).
+The goal of \cite{Burczynski:2006ik} was to improve accuracy in the discrimination of Inflammatory Bowel Disease (IBD) using the PBMC-based gene expression signature of a patient.
+
+\subsection{\cite{Nakayama:2007fl} Data Set}
+
+\cite{Nakayama:2007fl} acquired the gene expression through an oligonucleotide microarray from 105 samples of 10 types of soft tissue tumors.
+This included 16 samples of synovial sarcoma (SS), 19 samples of myxoid/round cell liposarcoma (MLS), 3 samples of lipoma, 3 samples of well-differentiated liposarcoma (WDLS), 15 samples of dedifferentiated liposarcoma (DDLS), 15 samples of myxofibrosarcoma (MFS), 6 samples of leiomyosarcoma (LMS), 3 samples of malignant nerve sheathe tumor (MPNST), 4 samples of fibrosarcoma (FS), and 21 samples of malignant fibrous histiocytoma (MFH).
+\cite{Nakayama:2007fl} determined from their data that these 10 types fell into 4 broader groups: (1) SS; (2) MLS; (3) Lipoma, WDLS, and part of DDLS; (4) Spindle cell and pleomorophic sarcomas including DDLS, MFS, LMS, MPNST, FS, and MFH.
+Following \cite{Witten:2011kc}, we restrict our analysis to the five tumor types having at least 15 samples observed.
+
+\subsection{\cite{Singh:2002fh} Data Set}
+
+\cite{Singh:2002fh} have examined 235 radical prostatectomy specimens from surgery patients between 1995 and 1997.
+The authors used oligonucleotide microarrays containing probes for approximately 12,600 genes and expressed sequence tags.
+They have reported that 102 of the radical prostatectomy specimens are of high quality: 52 prostate tumour samples and 50 non-tumour prostate samples.
 

### 2013-02-05 doc/alternative-RDA.tex
Added document containing alternative definition of RDA. Integrated the document into the latest draft of the paper.

\documentclass[11pt]{article}
\usepackage{graphicx, amsmath, amssymb, bm, url, mathtools, natbib, amsthm, setspace}

\pagestyle{plain}
%----------------Page dimensions ----------------
\oddsidemargin 0.0in
\evensidemargin 0.0in
\topmargin -0.75in
\leftmargin 0in
\headheight 0.0in
\headsep 0.5in
%\footheight 0.0in
\footskip 0.5in
\footnotesep 0.0in
\textwidth 6.7in
\textheight 9.5in
%-----------Define Pictures---------------------
\def\picture #1 by #2 (#3){
 \vbox to #2{
   \hrule width #1 height 0pt depth 0pt
   \vfill
   \special{picture #3} % this is the low-level interface
   }
 }
\def\scaledpicture #1 by #2 (#3 scaled #4){{
 \dimen0=#1 \dimen1=#2
 \divide\dimen0 by 1000 \multiply\dimen0 by #4
 \divide\dimen1 by 1000 \multiply\dimen1 by #4
 \picture \dimen0 by \dimen1 (#3 scaled #4)}
 }

\newcommand{\xbar}{\bar{\bm x}}
\newcommand{\tr}{\text{tr}}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}

\title{An Alternative Parameterization of Regularized Discriminant Analysis}

\author{John A. Ramey and Dean M. Young}

\begin{document}

\newtheorem*{thm}{Theorem}
\newtheorem*{cor}{Corollary}

\bibpunct{(}{)}{;}{a}{,}{,}

\doublespacing

\maketitle

\begin{abstract}
We introduce an alternative parameterization of regularized discriminant analysis (\emph{RDA}) from \cite{Friedman:1989tm} that is simpler and has a more intuitive interpretation. Additionally, by exploiting properties of the singular value decomposition (SVD) as well as the so-called Fast SVD algorithm, we propose a classifier that is faster to compute for small-sample, high-dimensional data sets.
\end{abstract}

\section{Introduction}

We wish to correctly assign an unlabeled $p$-dimensional observation vector $\bm x$ to one of $K$ unique, known classes (or populations) by constructing a classifier from $n$ training observations that can accurately predict the class membership of $\bm x$. Let $\bm x_i = (x_{i1}, x_{i2}, \ldots, x_{ip}) \in \mathbb{R}_{p \times 1}$ be the $i$th observation $(i = 1, \ldots, n)$ with true, unique membership $y_i \in \mathcal{K} = \{\omega_1, \ldots, \omega_K\}$, where $\mathbb{R}_{m \times n}$ denotes the matrix space of all $m \times n$ matrices over the real field $\mathbb{R}$. We assume that $(\bm x_i, y_i)$ is a realization from a mixture distribution $p(\bm x) = \sum_{k=1}^K p(\bm x | \omega_k) p(\omega_k)$, where $p(\bm x | \omega_k)$ is the probability density function (PDF) of the $k$th class, $p(\omega_k)$ is prior probability of class membership of the $k$th class.

Quadratic discriminant analysis (\emph{QDA}) is the optimal Bayesian decision rule with respect to a $0-1$ loss function when $p(\bm x | \omega_k)$ is the PDF of the multivariate normal distribution with known mean vector  $\bm\mu_k \in \mathbb{R}_{p \times 1}$ and known covariance matrix $\bm\Sigma_k \in \mathbb{R}_{p \times p}^{>}$, $k = 1, 2, \ldots, K$, where $\mathbb{R}_{p \times p}^{>}$ denotes the cone of real $p \times p$ positive-definite matrices. Because the parameters are typically unknown, we estimate the unknown parameters $\bm \mu_k$ and $\bm\Sigma_k$ with their maximum likelihood estimators (MLEs) and substitute the MLEs into the \emph{QDA} classifier. Assuming, for simplicity, that the prior probabilities of class membership $p(\omega_k)$ are equal for $k = 1, \ldots, K$, we assign an unlabeled observation $\bm x$ to class $\omega_k$ using the sample \emph{QDA} classifier:
\begin{align}
	D_{QDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\widehat{\bm\Sigma}_k^{-1}(\bm x - \xbar_k)  + \ln |\widehat{\bm\Sigma}_k|, \label{eq:qda}
\end{align}
where $\xbar_k$ and $\widehat{\bm\Sigma}_k$ are the MLEs for $\bm \mu_k$ and $\bm \Sigma_k$, respectively. If we assume further that the covariance matrix parameters are equal for each class (i.e., $\bm\Sigma_k = \bm\Sigma$ for all $k$), then \eqref{eq:qda} reduces to Fisher's linear discriminant analysis (LDA) classifier,
\begin{align}
	D_{LDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\widehat{\bm\Sigma}^{-1}(\bm x - \xbar_k), \label{eq:lda}
\end{align}
where $\widehat{\bm\Sigma}$ is the pooled sample covariance matrix
\begin{align}
	\widehat{\bm\Sigma} = \frac{1}{n} \sum_{k=1}^K n_k \widehat{\bm\Sigma}_k. \label{eq:pooled-cov}
\end{align}

\section{Regularized Discriminant Analysis}

\cite{Friedman:1989tm} has proposed the \emph{RDA} classifier by incorporating a weighted average of the sample covariance matrix $\widehat{\bm \Sigma}_k$ for class $\omega_k$ and the pooled sample covariance matrix $\widehat{\bm\Sigma}$ to estimate the covariance matrix for class $\omega_k$ with
\begin{align}
  \widehat{\bm\Sigma}_k(\lambda) = n_k^{-1}(\lambda) \bm S_k(\lambda),\label{eq:sig-lambda}
\end{align}
where $\lambda \in [0, 1]$, $n_k(\lambda) = (1 - \lambda) n_k + \lambda n$, $\bm S_k = n_k \widehat{\bm\Sigma}_k$, $\bm S = \sum_{k=1}^K \bm S_k$, and $\bm S_k(\lambda) = (1 - \lambda) \bm S_k + \lambda \bm S$. We can interpret \eqref{eq:sig-lambda} as a covariance matrix estimator for class $\omega_k$ that borrows from $\widehat{\bm\Sigma}$ in \eqref{eq:pooled-cov} to better estimate $\bm \Sigma_k$. If $\lambda = 0$, \eqref{eq:sig-lambda} is equal to $\widehat{\bm\Sigma}_k$ used in the \emph{QDA} classifier in \eqref{eq:qda}. Alternatively, if $\lambda = 1$, \eqref{eq:sig-lambda} is equal to $\widehat{\bm\Sigma}$ used in the \emph{LDA} classifier in \eqref{eq:lda}, in which case we assume implicitly that $\bm\Sigma_k = \bm \Sigma$ for $k = 1, \ldots, K$.

To further improve the estimation of $\bm \Sigma_k$ and to stabilize the inverse of \eqref{eq:sig-lambda}, \cite{Friedman:1989tm} has proposed the biased \emph{RDA} covariance-matrix estimator
\begin{align}
	\widehat{\bm\Sigma}_k(\lambda, \gamma) = (1 - \gamma) \widehat{\bm\Sigma}_k(\lambda) + \gamma \frac{\tr\{\widehat{\bm\Sigma}_k(\lambda)\}}{p} \bm I_p,\label{eq:sig-rda}
\end{align}
for class $\omega_k$, where $\gamma \in [0, 1]$ is a regularization parameter that controls the shrinkage of \eqref{eq:sig-rda}  towards the $p$-dimensional identity matrix $\bm I_p$, which is weighted by the the average of the eigenvalues of \eqref{eq:sig-lambda}. Thus, the \emph{pooling} parameter $\lambda$ controls the amount that we borrow from $\widehat{\bm\Sigma}$ to estimate $\bm \Sigma_k$, and the \emph{shrinkage} parameter $\gamma$ determines the amount of shrinkage applied.

Substituting \eqref{eq:sig-rda} into \eqref{eq:qda}, we have the \emph{RDA} classifier, where we assign an unlabeled observation $\bm x$ to class $\omega_k$:
\begin{align}
	D_{RDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\widehat{\bm\Sigma}_k(\lambda, \gamma)^{-1}(\bm x - \xbar_k)  + \ln |\widehat{\bm\Sigma}_k(\lambda, \gamma)|. \label{eq:rda}
\end{align}

\section{An Alternative Approach to Regularized Discriminant Analysis}

Here, we present a more straightforward, alternative parameterization of the \emph{RDA} classifier. We first define explicitly
\begin{align}
  \widehat{\bm\Sigma}_k(\lambda) = (1 - \lambda) \widehat{\bm\Sigma}_k + \lambda \widehat{\bm\Sigma}.\label{eq:sig-lambda-alternative}
\end{align}
as a convex combination of the MLEs of $\bm \Sigma_k$ and $\bm \Sigma$, where $\lambda \in [0, 1]$ is the \emph{pooling} parameter. Thus, we no longer have need for the convex combination of $n_k$ and $N$ as defined by \cite{Friedman:1989tm}. Before examining the shrinkage applied to \eqref{eq:sig-lambda-alternative}, we examine some of its properties and write \eqref{eq:sig-lambda-alternative} as
\begin{align}
	\widehat{\bm\Sigma}_k(\lambda) &= \widehat{\bm\Sigma}_k + \lambda (\widehat{\bm\Sigma} - \widehat{\bm\Sigma}_k) \nonumber\\
	&= \widehat{\bm\Sigma}_k + \lambda \left(\sum_{k' = 1}^K \frac{n_k}{N} \widehat{\bm\Sigma}_{k'}  - \widehat{\bm\Sigma}_k \right) \nonumber\\
	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right) \widehat{\bm\Sigma}_k +  \frac{\lambda}{N} \sum_{\substack{k' = 1\\k' \ne k}} n_{k'} \widehat{\bm\Sigma}_{k'}.\label{eq:sig-lambda-alternative2}
\end{align}
From \eqref{eq:sig-lambda-alternative2}, we see the weight $\lambda$ places on the covariance matrices from the remaining $K - 1$ classes. The interesting part here is that $\lambda$ controls the contribution of each of the $N$ observations to estimating $\Sigma_k$. Effectively, this states that for $0 < \lambda \le 1$, we use $N$ observations to estimate $\bm \Sigma_k$ rather than $n_k$ as done in the \emph{QDA} classifier. To see this further, suppose that we have centered each observation $\bm x_i$, $i = 1, \ldots, N$, by its class sample mean. Then, we write \eqref{eq:sig-lambda-alternative2} as
\begin{align}
	\widehat{\bm\Sigma}_k(\lambda) &= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N I(y_i = k) \bm x_i \bm x_i' +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k) \bm x_i \bm x_i' \nonumber \\
	&= \sum_{i=1}^N c_{ik}(\lambda) \bm x_i \bm x_i',\label{eq:sig-lambda-alternative3}
\intertext{where}
	c_{ik}(\lambda) &= \begin{cases}
		(1 - \lambda)n_k^{-1} + \lambda N^{-1}, & y_i = k,\\
		\lambda N^{-1}, & y_i \ne k.\\
	\end{cases}\nonumber
\end{align}
Therefore, after centering each observation by its corresponding class sample mean, we have that $\widehat{\bm\Sigma}_k(\lambda)$ in \eqref{eq:sig-lambda-alternative3} incorporates each centered observation into the estimation of $\bm \Sigma_k$. Notice that if $\lambda  = 0$, then the observations from the remaining $K - 1$ classes do not contribute to the estimation of $\bm \Sigma_k$, corresponding to $\widehat{\bm \Sigma}_k$. Furthermore, if $\lambda = 1$, the weights in \eqref{eq:sig-lambda-alternative3} reduce to $1/N$, corresponding to $\widehat{\bm\Sigma}$.

\subsection{Computational Advantages}

Thus far, our alternative parameterization simplifies the definition of the \emph{RDA} classifier and has an improved interpretation in terms of the contribution of each observation. Here, we demonstrate that our parameterization yields significant computational advantages over the originally proposed \emph{RDA} classifier from \cite{Friedman:1989tm} by writing \eqref{eq:sig-lambda-alternative3} as
\begin{align*}
	\widehat{\bm\Sigma}_k(\lambda) &= \bm X_k(\lambda)' \bm X_k(\lambda),
\end{align*}
where $\bm X_k(\lambda) = [\sqrt{c_{1k}(\lambda)} \bm x_1', \ldots,  \sqrt{c_{Nk}(\lambda)} \bm x_N']'$. Now, we can apply the Fast SVD to compute the eigenvalue decomposition of $\widehat{\bm\Sigma}_k(\lambda)$ efficiently. By writing the SVD of $\bm X_k(\lambda) = \bm U_k \bm D_k \bm V_k'$, we have the spectral decomposition $\widehat{\bm\Sigma}_k(\lambda) = \bm X_k(\lambda)' \bm X_k(\lambda) = \bm V_k \bm D_k^2 \bm V_k'$, where $\bm U_k \in \mathbb{R}_{N \times N}$ is orthogonal, $\bm D_k \in \mathbb{R}_{N \times N}$ is a diagonal matrix consisting of the singular values of $\bm X_k(\lambda)$, and $\bm V_k \in \mathbb{R}_{p \times N}$ is orthogonal. Notice that $\bm X_k(\lambda)\bm X_k(\lambda)' = \bm U_k \bm D_k^2 \bm U_k'$, so that the eigenvalue decomposition of $\bm X_k(\lambda)\bm X_k(\lambda)'$ yields $\bm U$ and $\bm D$. This is much faster when $p \gg N$ because $\bm X_k(\lambda)\bm X_k(\lambda)' \in \mathbb{R}_{N \times N}$. Also, notice that the singular values of $\bm D$ are nonzero, so that $\bm D^{-1}$ exists. After obtaining $\bm U_k$ and $\bm D_k$, we compute the matrix consisting of the first $N$ eigenvectors of $\widehat{\bm\Sigma}_k(\lambda)$ by simply computing\footnote{TODO: Update this paragraph to theorem/proof.}
\begin{align*}
	\bm V_k = \bm X_k(\lambda)' \bm U_k \bm D_k^{-1}.
\end{align*}

\subsection{Covariance-Matrix Regularization}

We are now ready to discuss the shrinkage employed in the \emph{RDA} covariance-matrix estimator. As we are interested primarily in the case that $p \gg N$, rank$\{\widehat{\bm\Sigma}_k(\lambda)\} = q_k < p$, which implies that $p - q_k$ of the eigenvalues of $\widehat{\bm\Sigma}_k(\lambda)$ are equal to zero. Thus, for $\gamma > 0$, shrinking $\widehat{\bm\Sigma}_k(\lambda)$ towards $\bm I_p$ as in \eqref{eq:sig-rda} increases the rank of $\widehat{\bm\Sigma}_k(\lambda)$ from $q_k$ to $p$. However, \cite{Ye:2009gd} and \cite{YeOtherPaperCitedInTheFirst} have shown that shrinking the zero eigenvalues does not improve the classification performance of the \emph{RDA} classifier. Consequently, we instead shrink only the nonzero eigenvalues of $\widehat{\bm\Sigma}_k(\lambda)$, which has two additional benefits. First, by considering only the nonzero eigenvalues of $\widehat{\bm\Sigma}_k(\lambda)$, we effectively ignore the eigenvectors corresponding to the zero eigenvalues, which reduces significantly the number of parameters that we incorporate into our \emph{RDA} classifier because $p \gg q_k$ if $p \gg N$. Furthermore, the size of the matrix $\bm V_k$ is $p \times N$ as opposed to a $p \times p$ matrix in the original definition of the \emph{RDA} classifier. For significantly large $p$, the reduced size of $\bm V_k$ expedites calculation of the \emph{RDA} classifier and reduces its memory usage.\footnote{TODO: Provide an example for a realistic $p$ and $N$.}

Specifically, we shrink only the nonzero eigenvalues of $\widehat{\bm \Sigma}_k(\lambda)$ by utilizing the estimator

\begin{align}
	\widehat{\bm\Sigma}_k(\lambda, \gamma) = (1 - \gamma) \widehat{\bm\Sigma}_k(\lambda) + \gamma \frac{\tr\{\widehat{\bm\Sigma}_k(\lambda)\}}{p} (\bm I_{q_k} \oplus \bm 0_{p - q_k}),\label{eq:sig-rda-alternative}
\end{align}
where $\bm A \oplus \bm B$ denote the direct sum of $\bm A \in \mathbb{R}_{r \times r}$ and $\bm B \in \mathbb{R}_{s \times s}$ \citep[Chapter 1]{Lutkepohl:1996uz}. Therefore, substituting \eqref{eq:sig-rda-alternative} into \eqref{eq:rda}, we have our alternative parameterization of the \emph{RDA} classifier.

\bibliographystyle{plainnat}
\bibliography{../reports/rda}

\end{document}

### 2013-02-05 reports/paper-etrda.tex
Added document containing alternative definition of RDA. Integrated the document into the latest draft of the paper.

@@
 
-\title{A Generalization of Regularized Discriminant Analysis with Applications to High-dimensional Microarray Data}
+\title{Another Look at Regularized Discriminant Analysis for High-dimensional Classification}
 
@@
 \begin{abstract}
-\cite{Friedman:1989tm} proposed the popular regularized discriminant analysis (\emph{RDA}) classifier for the small-sample, high-dimensional setting by employing a regularized covariance matrix estimator, which is shrunken towards a scaled identity matrix.
-We demonstrate empirically, however, that the RDA classifier is often numerically incalculable or unstable for small-sample, high-dimensional data because the shrinkage factor is a function of the average of the eigenvalues, which tends to zero in this setting.
+\cite{Friedman:1989tm} proposed the popular regularized discriminant analysis (\emph{RDA}) classifier for the small-sample, high-dimensional setting by employing a regularized covariance matrix estimator, which employs a ridge-like shrinkage factor to ensure positive definiteness.
+We demonstrate that the \emph{RDA} classifier, in its original form, is inadequate for the small-sample, high-dimensional setting.
+We redefine the \emph{RDA} classifier to increase its computational runtime and its interpretation.
+We show that this new definition includes several well-known regularized classifiers as special cases.
+We also consider several properties of the newly defined \emph{RDA} classifier.
 Alternatively, \cite{Srivastava:2007ww} propose an empirical-Bayesian estimator that employs a similar shrinkage factor using the average of the nonzero eigenvalues.
@@
 By incorporating our proposed covariance-matrix regularization with the RDA classifier, we overcome the numerical incalculability issues of the \emph{RDA} classifier.
-Moreover, we show that our proposed classifier is often superior to the regularized classifiers mentioned above in terms of classification accuracy using three simulated and five small-sample, high-dimensional microarray data sets.
+Moreover, we show that our proposed classifier is often superior to the regularized classifiers mentioned above in terms of classification accuracy using three simulated and three small-sample, high-dimensional microarray data sets.
 Additionally, we provide implementations of the competing classifiers in the {\tt regdiscrim} R package, available on CRAN.
@@
 
-(\textbf{Remove this paragraph?})
-Despite its popularity, we demonstrate empirically that the covariance-matrix shrinkage employed in the \emph{RDA} classifier is often ineffective in adjusting the eigenvalues of the \emph{RDA} covariance-matrix estimator when $p \gg n$.
-The \emph{RDA} classifier is numerically unstable as the number of features incorporated into the \emph{RDA} classifier is increased because the covariance-matrix estimator is nearly singular.
-Specifically, the shrinkage applied in the \emph{RDA} classifier is a function of the average eigenvalues of the convex combination of the class sample covariance matrix and the sample covariance matrix pooled across all classes.
-Because the average of the eigenvalues is often nearly zero when $p \gg n$, the resulting covariance-matrix estimator has two issues.
-First, the determinant of the covariance-matrix estimator approaches zero, suggesting that the \emph{RDA} covariance-matrix estimator is nearly singular in the small-sample, high-dimensional setting.
-Second, the smallest eigenvalues are near zero, which causes the \emph{RDA} classifier to have an incalculable or numerically unstable discriminant function.
-
 In this paper we present a generalization of the \emph{RDA} classifier in terms of matrix functions.
@@
 
-Using \textbf{five} small-sample, high-dimensional microarray data sets, we demonstrate that our two proposed classifiers often yield superior classification accuracy to the \emph{RDA}, \emph{MDEB}, and \emph{NLDA} classifiers.
+Using three small-sample, high-dimensional microarray data sets, we demonstrate that our two proposed classifiers often yield superior classification accuracy to the \emph{RDA}, \emph{MDEB}, and \emph{NLDA} classifiers.
 We also include in our comparison the alternatively-defined \emph{RDA} classifier from \cite*{Hastie:2008dt}, the regularized \emph{LDA} classifier from \cite{Guo:2007te}, a recently proposed variant of diagonal discriminant analysis from \cite*{Tong:2012hw}, and the penalized \emph{LDA} classifier from \cite{Witten:2011kc}.
@@
 In Section 3 we generalize the eigenvalue adjustment employed in the \emph{NLDA} classifier and define our proposed \emph{Soft-ETRDA} and \emph{Hard-ETRDA} classifiers.
-In Section 4 we describe the \textbf{three} simulation configurations and \textbf{five} small-sample, high-dimensional microarray data sets and then demonstrate that our two proposed classifiers are superior to the competing classifiers in terms of classification accuracy.
+In Section 4 we describe the \textbf{three} simulation configurations and three small-sample, high-dimensional microarray data sets and then demonstrate that our two proposed classifiers are superior to the competing classifiers in terms of classification accuracy.
 We then conclude with a brief discussion in Section 6.
 
-\section{Covariance-Matrix Regularization in Discriminant Analysis}
+\section{Regularized Discriminant Analysis}
 
-In discriminant analysis we wish to correctly assign an unlabeled $p$-dimensional observation vector $\bm x$ to one of $K$ unique, known classes (or populations) by constructing a classifier from $n$ training observations that can accurately predict the class membership of $\bm x$.
+We\footnote{Define the majority of notation here, including $z^+$, $z_+$, and $I[z]$.} wish to assign correctly an unlabeled $p$-dimensional observation vector $\bm x$ to one of $K$ unique, known classes (or populations) by constructing a classifier from $n$ training observations that can accurately predict the class membership of $\bm x$.
 Let $\bm x_i = (x_{i1}, x_{i2}, \ldots, x_{ip}) \in \mathbb{R}_{p \times 1}$ be the $i$th observation $(i = 1, \ldots, n)$ with true, unique membership $y_i \in \mathcal{K} = \{\omega_1, \ldots, \omega_K\}$, where $\mathbb{R}_{m \times n}$ denotes the matrix space of all $m \times n$ matrices over the real field $\mathbb{R}$.
@@
 \end{align}
-where $\xbar_k$ is the MLE for $\bm \mu_k$ and  $\widehat{\bm\Sigma}_k$ is the MLE for $\bm \Sigma_k$.
-Notice that if we assume further that the covariance matrix parameters are equal for each class (i.e., $\bm\Sigma_k = \bm\Sigma$ for all $k$), then \eqref{eq:qda} reduces to Fisher's linear discriminant analysis (LDA) classifier,
+where $\xbar_k$ and $\widehat{\bm\Sigma}_k$ are the MLEs for $\bm \mu_k$ and $\bm \Sigma_k$, respectively.
+If we assume further that the covariance matrix parameters are equal for each class (i.e., $\bm\Sigma_k = \bm\Sigma$ for all $k$), then \eqref{eq:qda} reduces to Fisher's linear discriminant analysis (LDA) classifier,
 \begin{align}
@@
 
-In the small-sample, high-dimensional setting, if $p > n_k$ for any $k$, \eqref{eq:qda} is incalculable because the sample covariance matrix $\widehat{\bm\Sigma}_k$ for class $\omega_k$ is singular, and, thus, its inverse $\widehat{\bm\Sigma}_k^{-1}$ does not exist.
+\subsection{Original Definition of Regularized Discriminant Analysis}
+
+\cite{Friedman:1989tm} has proposed the \emph{RDA} classifier by incorporating a weighted average of the sample covariance matrix $\widehat{\bm \Sigma}_k$ for class $\omega_k$ and the pooled sample covariance matrix $\widehat{\bm\Sigma}$ to estimate the covariance matrix for class $\omega_k$ with
+\begin{align}
+  \widehat{\bm\Sigma}_k(\lambda) = n_k^{-1}(\lambda) \bm S_k(\lambda),\label{eq:sig-lambda}
+\end{align}
+where $\lambda \in [0, 1]$, $n_k(\lambda) = (1 - \lambda) n_k + \lambda n$, $\bm S_k = n_k \widehat{\bm\Sigma}_k$, $\bm S = \sum_{k=1}^K \bm S_k$, and $\bm S_k(\lambda) = (1 - \lambda) \bm S_k + \lambda \bm S$.
+We can interpret \eqref{eq:sig-lambda} as a covariance matrix estimator for class $\omega_k$ that borrows from $\widehat{\bm\Sigma}$ in \eqref{eq:pooled-cov} to better estimate $\bm \Sigma_k$.
+If $\lambda = 0$, \eqref{eq:sig-lambda} is equal to $\widehat{\bm\Sigma}_k$ used in the \emph{QDA} classifier in \eqref{eq:qda}.
+Alternatively, if $\lambda = 1$, \eqref{eq:sig-lambda} is equal to $\widehat{\bm\Sigma}$ used in the \emph{LDA} classifier in \eqref{eq:lda}, in which case we assume implicitly that $\bm\Sigma_k = \bm \Sigma$ for $k = 1, \ldots, K$.
+
+To further improve the estimation of $\bm \Sigma_k$ and to stabilize the inverse of \eqref{eq:sig-lambda}, \cite{Friedman:1989tm} has proposed the biased \emph{RDA} covariance-matrix estimator
+\begin{align}
+	\widehat{\bm\Sigma}_k(\lambda, \gamma) = (1 - \gamma) \widehat{\bm\Sigma}_k(\lambda) + \gamma \frac{\tr\{\widehat{\bm\Sigma}_k(\lambda)\}}{p} \bm I_p,\label{eq:sig-rda}
+\end{align}
+for class $\omega_k$, where $\gamma \in [0, 1]$ is a regularization parameter that controls the shrinkage of \eqref{eq:sig-rda}  towards the $p$-dimensional identity matrix $\bm I_p$, which is weighted by the the average of the eigenvalues of \eqref{eq:sig-lambda}.
+Thus, the \emph{pooling} parameter $\lambda$ controls the amount that we borrow from $\widehat{\bm\Sigma}$ to estimate $\bm \Sigma_k$, and the \emph{shrinkage} parameter $\gamma$ determines the amount of shrinkage applied.
+
+Substituting \eqref{eq:sig-rda} into \eqref{eq:qda}, we have the \emph{RDA} classifier, where we assign an unlabeled observation $\bm x$ to class $\omega_k$:
+\begin{align}
+	D_{RDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\widehat{\bm\Sigma}_k(\lambda, \gamma)^{-1}(\bm x - \xbar_k)  + \ln |\widehat{\bm\Sigma}_k(\lambda, \gamma)|. \label{eq:rda}
+\end{align}
+
+\subsection{An Alternative Approach to Regularized Discriminant Analysis}
+
+Here, we present a more straightforward, alternative parameterization of the \emph{RDA} classifier. We first define explicitly
+\begin{align}
+  \widehat{\bm\Sigma}_k(\lambda) = (1 - \lambda) \widehat{\bm\Sigma}_k + \lambda \widehat{\bm\Sigma}.\label{eq:sig-lambda-alternative}
+\end{align}
+as a convex combination of the MLEs of $\bm \Sigma_k$ and $\bm \Sigma$, where $\lambda \in [0, 1]$ is the \emph{pooling} parameter.
+Thus, we no longer have need for the convex combination of $n_k$ and $N$ as defined by \cite{Friedman:1989tm}.
+Before examining the shrinkage applied to \eqref{eq:sig-lambda-alternative}, we examine some of its properties and write \eqref{eq:sig-lambda-alternative} as
+\begin{align}
+	\widehat{\bm\Sigma}_k(\lambda) &= \widehat{\bm\Sigma}_k + \lambda (\widehat{\bm\Sigma} - \widehat{\bm\Sigma}_k) \nonumber\\
+	&= \widehat{\bm\Sigma}_k + \lambda \left(\sum_{k' = 1}^K \frac{n_k}{N} \widehat{\bm\Sigma}_{k'}  - \widehat{\bm\Sigma}_k \right) \nonumber\\
+	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right) \widehat{\bm\Sigma}_k +  \frac{\lambda}{N} \sum_{\substack{k' = 1\\k' \ne k}} n_{k'} \widehat{\bm\Sigma}_{k'}.\label{eq:sig-lambda-alternative2}
+\end{align}
+From \eqref{eq:sig-lambda-alternative2}, we see the weight $\lambda$ places on the covariance matrices from the remaining $K - 1$ classes.
+The interesting part here is that $\lambda$ controls the contribution of each of the $N$ observations to estimating $\Sigma_k$.
+Effectively, this states that for $0 < \lambda \le 1$, we use $N$ observations to estimate $\bm \Sigma_k$ rather than $n_k$ as done in the \emph{QDA} classifier.
+As we will discuss in-depth later, the pooling operation is advantageous in increasing the rank of each $\widehat{\bm\Sigma}_k(\lambda)$ from $n_k - 1$ to $N - 1$ for $0 < \lambda \le 1$.
+Moreover, suppose that we have centered each observation $\bm x_i$, $i = 1, \ldots, N$, by its class sample mean.
+Then, we write \eqref{eq:sig-lambda-alternative2} as
+\begin{align}
+	\widehat{\bm\Sigma}_k(\lambda) &= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N I(y_i = k) \bm x_i \bm x_i' +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k) \bm x_i \bm x_i' \nonumber \\
+	&= \sum_{i=1}^N c_{ik}(\lambda) \bm x_i \bm x_i',\label{eq:sig-lambda-alternative3}
+\intertext{where}
+	c_{ik}(\lambda) &= \begin{cases}
+		(1 - \lambda)n_k^{-1} + \lambda N^{-1}, & y_i = k,\\
+		\lambda N^{-1}, & y_i \ne k.\\
+	\end{cases}\nonumber
+\end{align}
+Therefore, after centering each observation by its corresponding class sample mean, we have that $\widehat{\bm\Sigma}_k(\lambda)$ in \eqref{eq:sig-lambda-alternative3} incorporates each centered observation into the estimation of $\bm \Sigma_k$.
+Notice that if $\lambda  = 0$, then the observations from the remaining $K - 1$ classes do not contribute to the estimation of $\bm \Sigma_k$, corresponding to $\widehat{\bm \Sigma}_k$.
+Furthermore, if $\lambda = 1$, the weights in \eqref{eq:sig-lambda-alternative3} reduce to $1/N$, corresponding to $\widehat{\bm\Sigma}$.
+
+\subsection{Covariance Matrix Regularization}
 
 Although \eqref{eq:qda} is in a standard form, we prefer to express the \emph{QDA} classifier explicitly in terms of the eigenvalues and eigenvectors of $\widehat{\bm \Sigma}_k$.
-Let $e_{jk}$ denote the $j$th largest eigenvalue of $\widehat{\bm\Sigma}_k$ and $\bm v_{jk}$ be the corresponding eigenvector such that $e_{1k} \ge e_{2k} \ge \ldots \ge e_{pk} \ge 0$.
+Let $e_{jk}(\lambda)$ denote the $j$th largest eigenvalue of $\widehat{\bm\Sigma}_k$ and $\bm v_{jk}(\lambda)$ be the corresponding eigenvector such that $e_{1k}(\lambda) \ge e_{2k}(\lambda) \ge \ldots \ge e_{pk}(\lambda) \ge 0$.
 Equivalently, we write \eqref{eq:qda} as
 \begin{align}
-  D(\bm x) = \argmin_{k} \sum_{j = 1}^p \frac{[\bm v_{jk}' (\bm x - \bm \xbar_k)]^2}{e_{jk}} + \sum_{j=1}^p \ln e_{jk}. \label{eq:qda-eigen}
+  D(\bm x) = \argmin_{k} \sum_{j = 1}^p \frac{[\bm v_{jk}(\lambda)' (\bm x - \bm \xbar_k)]^2}{e_{jk}(\lambda)} + \sum_{j=1}^p \ln \{e_{jk}(\lambda)\}. \label{eq:rda-eigen}
 \end{align}
-The smallest eigenvalues and the directions associated with their eigenvectors highly influence \eqref{eq:qda-eigen} .
-In fact, the eigenvalues of $\widehat{\bm \Sigma}_k$ are well-known to be biased if $p \ge n_k$ such that the smallest eigenvalues are underestimated \citep{Seber:2004uh}.
+The smallest eigenvalues and the directions associated with their eigenvectors highly influence \eqref{eq:rda-eigen}.
+In fact, the eigenvalues of $\widehat{\bm \Sigma}_k(\lambda)$ are well-known to be biased if $p \ge n_k$ such that the smallest eigenvalues are underestimated \citep{Seber:2004uh}.
 This bias increases as $p$ increases relative to $n_k$ \citep{TODONeedCitation}.
-Moreover, if $p > n_k$, then rank$(\widehat{\bm \Sigma}_k) \le n_k$, which implies that the last $p - n_k$ eigenvalues of $\widehat{\bm \Sigma}_k$ are 0, which implies that \eqref{eq:qda-eigen} is incalculable.
+Moreover, if $p > n_k$, then rank$(\widehat{\bm \Sigma}_k(\lambda)) \le n_k$, which implies that the last $p - n_k$ eigenvalues of $\widehat{\bm \Sigma}_k(\lambda)$ are 0, which implies that \eqref{eq:rda-eigen} is incalculable.
 Thus, although more feature information is available to discriminate among the $K$ classes as $p$ increases, classification accuracy decreases unless one obtains enough training-sample observations to reliably estimate the increased number of parameters.
@@
 is a matrix function defined explicitly in terms of the eigenvalues of $\bm A$.
-For this reason, $\phi$ is often called a spectral transform.
+For this reason, $\phi$ is often called a spectral transform \citep{Ch15SSLBook} or a transfer function \citep{Ye:2009gd, Chapelle2002SSLChapter15Reference} .
 Special case of \eqref{eq:spectral-decomp} include the spectral decomposition of $\bm A$ when $\phi(e_j) = e_j$, the matrix square-root of $\bm A$ when $\phi(e_j) = e_j^{1/2}$, and, if it exists, the inverse of $\bm A$ when $\phi(e_j) = e_j^{-1}$ \citep{Harville:2008wja}.
@@
 \begin{align}
-  D(\bm x) = \argmin_{k} \sum_{j = 1}^p \phi(e_{jk})[\bm v_{jk}' (\bm x - \bm \mu_k)]^2 + \sum_{j=1}^p \ln \{ \phi(e_{jk}) \}. \label{eq:generalized-qda}
+  D(\bm x) = \argmin_{k} \sum_{j = 1}^p \phi(e_{jk}(\lambda))[\bm v_{jk}(\lambda)' (\bm x - \bm \mu_k)]^2 + \sum_{j=1}^p \ln \{ \phi(e_{jk}(\lambda)) \}. \label{eq:generalized-qda}
 \end{align}
-The generalized \emph{QDA} classifier in \eqref{eq:generalized-qda} includes several common regularized classifiers.
-Notice that if  $\phi(e_{jk}) = e_{jk}$, $k = 1, \ldots, K$, \eqref{eq:generalized-qda} reduces to the \emph{QDA} classifier in \eqref{eq:qda-eigen}.
-Furthermore, if $\phi(e_{jk}) = e_{j}$, $k = 1, \ldots, K$, \eqref{eq:generalized-qda} reduces to the \emph{LDA} classifier in \eqref{eq:lda}, and the second summand in \eqref{eq:generalized-qda} is constant across all classes and can be omitted.
+The\footnote{In \eqref{eq:generalized-qda} we have a slight issue in the formulation of the log-determinant in the second sum.
+The eigenvalue function $\phi$ in the first sum should be the inverse of the function in the second sum.
+Furthermore, we need to define the log-determinant so that the last $p - q$ terms are ignored when we reduce the dimension to $q$.} generalized \emph{QDA} classifier in \eqref{eq:generalized-qda} includes several common regularized classifiers.
+Notice that if  $\phi(e_{jk}(\lambda)) = e_{jk}(\lambda)$, $k = 1, \ldots, K$, \eqref{eq:generalized-qda} reduces to the \emph{QDA} classifier in \eqref{eq:rda-eigen}.
+Furthermore, if $\phi(e_{jk}(\lambda)) = e_{j}$, $k = 1, \ldots, K$, \eqref{eq:generalized-qda} reduces to the \emph{LDA} classifier in \eqref{eq:lda}, and the second summand in \eqref{eq:generalized-qda} is constant across all classes and can be omitted.
 Additionally, one common approach is to ignore the zero eigenvalues when $p > n$ by employing the Moore-Penrose pseudoinverse $\widehat{\bm \Sigma}_k^{+}$, which employs the matrix function $\phi(e_j) = e_j^{+}$, where $z^{+} = 1/z$ if $z > 0$ and $0$ otherwise.
@@
 \begin{align}
-	\phi(e_{jk}) = (e_{jk} + \gamma_k)^{-1}, \label{eq:ridge-estimator}
+	\phi(e_{jk}(\lambda)) = (e_{jk}(\lambda) + \gamma_k)^{-1}, \label{eq:ridge-estimator}
 \end{align}
 is a ridge-like estimator, similar to a method employed in ridge regression.
-The estimator in \eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance matrix $\widehat{\bm\Sigma}_k$ towards the $p$-dimensional identity matrix $\bm I_p$, thereby increasing the eigenvalues of $\widehat{\bm\Sigma}_k$ by $\gamma_k$ so that they are positive, resulting in $\widehat{\bm \Sigma}_k(\gamma_k)$ being positive definite.
+The estimator in \eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance matrix $\widehat{\bm\Sigma}_k(\lambda)$ towards the $p$-dimensional identity matrix $\bm I_p$, thereby increasing the eigenvalues of $\widehat{\bm\Sigma}_k(\lambda)$ by $\gamma_k$ so that they are positive, resulting in $\widehat{\bm \Sigma}_k(\lambda) + \gamma_k \bm I_p$ being positive definite.
 
@@
 
-\subsection{Regularized Discriminant Analysis}
+\section{Properties of the Redefined RDA Classifier}
 
-\cite{Friedman:1989tm} has extended the shrinkage technique by first computing a
-weighted average of the sample covariance matrix $\widehat{\bm \Sigma}_k$ for class
-$\omega_k$ and the pooled sample covariance matrix $\widehat{\bm\Sigma}$ to
-estimate the covariance matrix for class $\omega_k$ with $\widehat{\bm\Sigma}_k(\lambda)$,
-where $\lambda \in [0, 1]$, $n_k(\lambda) = (1 - \lambda) n_k + \lambda n$, $\bm S_k = n_k \widehat{\bm\Sigma}_k$, $\bm S = \sum_{k=1}^K \bm S_k$, $\bm S_k(\lambda) = (1 - \lambda) \bm S_k + \lambda \bm S$, and
+\subsection{Computational Advantages}
+
+Thus far, our alternative parameterization simplifies the definition of the \emph{RDA} classifier and has an improved interpretation in terms of the contribution of each observation.
+Here, we demonstrate that our parameterization yields significant computational advantages over the originally proposed \emph{RDA} classifier from \cite{Friedman:1989tm} by writing \eqref{eq:sig-lambda-alternative3} as
+\begin{align*}
+	\widehat{\bm\Sigma}_k(\lambda) &= \bm X_k(\lambda)' \bm X_k(\lambda),
+\end{align*}
+where $\bm X_k(\lambda) = [\sqrt{c_{1k}(\lambda)} \bm x_1', \ldots,  \sqrt{c_{Nk}(\lambda)} \bm x_N']'$.
+Now, we can apply the Fast SVD to compute the eigenvalue decomposition of $\widehat{\bm\Sigma}_k(\lambda)$ efficiently.
+By writing the SVD of $\bm X_k(\lambda) = \bm U_k \bm D_k \bm V_k'$, we have the spectral decomposition $\widehat{\bm\Sigma}_k(\lambda) = \bm X_k(\lambda)' \bm X_k(\lambda) = \bm V_k \bm D_k^2 \bm V_k'$, where $\bm U_k \in \mathbb{R}_{N \times N}$ is orthogonal, $\bm D_k \in \mathbb{R}_{N \times N}$ is a diagonal matrix consisting of the singular values of $\bm X_k(\lambda)$, and $\bm V_k \in \mathbb{R}_{p \times N}$ is orthogonal.
+Notice that $\bm X_k(\lambda)\bm X_k(\lambda)' = \bm U_k \bm D_k^2 \bm U_k'$, so that the eigenvalue decomposition of $\bm X_k(\lambda)\bm X_k(\lambda)'$ yields $\bm U_k$ and $\bm D_k$.
+This is much faster when $p \gg N$ because $\bm X_k(\lambda)\bm X_k(\lambda)' \in \mathbb{R}_{N \times N}$.
+Also, notice that the singular values of $\bm D_k$ are nonzero, so that $\bm D_k^{-1}$ exists.
+After obtaining $\bm U_k$ and $\bm D_k$, we compute the matrix consisting of the first $N$ eigenvectors of $\widehat{\bm\Sigma}_k(\lambda)$ by simply computing\footnote{TODO: Update this paragraph to theorem/proof.}
+\begin{align*}
+	\bm V_k = \bm X_k(\lambda)' \bm U_k \bm D_k^{-1}.
+\end{align*}
+
+\subsection{Covariance-Matrix Regularization}
+
+We are now ready to discuss the shrinkage employed in the \emph{RDA} covariance-matrix estimator.
+As we are interested primarily in the case that $p \gg N$, rank$\{\widehat{\bm\Sigma}_k(\lambda)\} = q_k < p$, which implies that $p - q_k$ of the eigenvalues of $\widehat{\bm\Sigma}_k(\lambda)$ are equal to zero.
+Thus, for $\gamma > 0$, shrinking $\widehat{\bm\Sigma}_k(\lambda)$ towards $\bm I_p$ as in \eqref{eq:sig-rda} increases the rank of $\widehat{\bm\Sigma}_k(\lambda)$ from $q_k$ to $p$.
+However, \cite{Ye:2009gd} and \cite{YeOtherPaperCitedInTheFirst} have shown that shrinking the zero eigenvalues does not improve the classification performance of the \emph{RDA} classifier.
+Consequently, we instead shrink only the nonzero eigenvalues of $\widehat{\bm\Sigma}_k(\lambda)$, which has two additional benefits.
+First, by considering only the nonzero eigenvalues of $\widehat{\bm\Sigma}_k(\lambda)$, we effectively ignore the eigenvectors corresponding to the zero eigenvalues, which reduces significantly the number of parameters that we incorporate into our \emph{RDA} classifier because $p \gg q_k$ if $p \gg N$.
+Furthermore, the size of the matrix $\bm V_k$ is $p \times N$ as opposed to a $p \times p$ matrix in the original definition of the \emph{RDA} classifier.
+For significantly large $p$, the reduced size of $\bm V_k$ expedites calculation of the \emph{RDA} classifier and reduces its memory usage.\footnote{TODO: Provide an example for a realistic $p$ and $N$.}
+
+Specifically, we shrink only the nonzero eigenvalues of $\widehat{\bm \Sigma}_k(\lambda)$ by utilizing the estimator
+
 \begin{align}
-  \widehat{\bm\Sigma}_k(\lambda) = n_k^{-1}(\lambda) \bm S_k(\lambda).\label{eq:sig-lambda}
+	\widehat{\bm\Sigma}_k(\lambda, \gamma) = (1 - \gamma) \widehat{\bm\Sigma}_k(\lambda) + \gamma \frac{\tr\{\widehat{\bm\Sigma}_k(\lambda)\}}{p} (\bm I_{q_k} \oplus \bm 0_{p - q_k}),\label{eq:sig-rda-alternative}
 \end{align}
-We can interpret \eqref{eq:sig-lambda} as a
-covariance matrix estimator for class $\omega_k$ that borrows from $\widehat{\bm\Sigma}$ in \eqref{eq:pooled-cov} to better estimate $\bm \Sigma_k$. Notice that for $\lambda = 0$,
-\eqref{eq:sig-lambda} is equal to $\widehat{\bm\Sigma}_k$ used in the \emph{QDA} classifier in \eqref{eq:qda}.
-Similarly, notice that for $\lambda = 1$, \eqref{eq:sig-lambda} is equal to $\widehat{\bm\Sigma}$ used in the \emph{LDA} classifier in \eqref{eq:lda}, in which case we assume implicitly that $\bm\Sigma_k = \bm \Sigma$ for $k = 1, \ldots, K$.
-
-\cite{Friedman:1989tm} also uses a regularization parameter $\gamma$ to shrink the
-eigenvalues of \eqref{eq:sig-lambda} towards the $p$-dimensional identity matrix $\bm I_p$
-in order to stabilize the inverse of \eqref{eq:sig-lambda}, resulting in a biased
-covariance-matrix estimator with matrix function
-\begin{align}
-	\phi(e_{jk}) = (\{1 - \gamma\} e_{jk} + \gamma \bar{e}_k)^{-1},\label{eq:rda-eigenvalue-function}
-\end{align}
-for class $\omega_k$, where $\gamma \in [0, 1]$, $e_{jk}$ be the $j$th largest eigenvalue of $\widehat{\bm\Sigma}_k(\lambda)$ in \eqref{eq:sig-lambda}, $\bm v_{jk}$ be the corresponding eigenvector, and $\bar{e}_k = p^{-1} \sum_{j=1}^p e_{jk}$ is the average of the eigenvalues of \eqref{eq:sig-lambda}.
-In summary, the \emph{pooling} parameter $\lambda$ controls the amount that we borrow from $\widehat{\bm\Sigma}$ to estimate $\bm \Sigma_k$, and the \emph{shrinkage} parameter $\gamma$ determines the amount of applied shrinkage.
+where $\bm A \oplus \bm B$ denote the direct sum of $\bm A \in \mathbb{R}_{r \times r}$ and $\bm B \in \mathbb{R}_{s \times s}$ \citep[Chapter 1]{Lutkepohl:1996uz}.
+Therefore, substituting \eqref{eq:sig-rda-alternative} into \eqref{eq:rda}, we have our alternative parameterization of the \emph{RDA} classifier.
 
@@
 
-(\textbf{TODO: Emphasize that for small $\Delta$, Soft and Hard are approximately equal. This provides a connection between NLDA and RDA.})
+\section{Special Cases}
 
-\section{Special Cases}
+\subsection{RDA}
+
+\cite{Friedman:1989tm} also uses a regularization parameter $\gamma$ to shrink the
+eigenvalues of \eqref{eq:sig-lambda} towards the $p$-dimensional identity matrix $\bm I_p$
+in order to stabilize the inverse of \eqref{eq:sig-lambda}, resulting in a biased
+covariance-matrix estimator with matrix function
+\begin{align}
+	\phi(e_{jk}(\lambda)) = (\{1 - \gamma\} e_{jk}(\lambda) + \gamma \bar{e}_k)^{-1},\label{eq:rda-eigenvalue-function}
+\end{align}
+for class $\omega_k$, where $\gamma \in [0, 1]$, $e_{jk}(\lambda)$ be the $j$th largest eigenvalue of $\widehat{\bm\Sigma}_k(\lambda)$ in \eqref{eq:sig-lambda}, $\bm v_{jk}(\lambda)$ be the corresponding eigenvector, and $\bar{e}_k = p^{-1} \sum_{j=1}^p e_{jk}(\lambda)$ is the average of the eigenvalues of \eqref{eq:sig-lambda}.
 
@@
 
-TODO
+\begin{itemize}
+	\item TODO: Design an experiment similar to Friedman's initial simulations, where the the majority of the variances are nearly 0.
+Fix the sample size.
+Consider the error rate as a function of $p$.
+This should cause the RDA regularization to be ineffective because the average eigenvalue is near 0.
 
-\section{Data Sets}
+	\item TODO: Design an experiment where the average eigenvalue goes to 0, but the average nonzero eigenvalue blows up for a fixed sample size.
+Consider the error rate as a function of $p$.
+This should cause the RDA and MDEB regularizations to be ineffective.
+However, I am banking on the ETRDA methods doing well.
+
+\end{itemize}
+
+\section{High-Dimensional Microarray Data Sets}
+
+In this section, we describe three high-dimensional microarray data sets and compare our proposed \emph{ETRDA} classifier with two recently proposed classifiers for small-sample, high-dimensional data: the penalized \emph{LDA} (\emph{PLDA}) classifier from \cite{Witten:2011kc}  and the shrinkage-mean-based DLDA (\emph{SmDLDA}) classifier from \cite{Tong:2012hw}.
+We calculated 10-fold cross-validation error rates \citep{Hastie:2008dt} to evaluate the performance of each classifier considered.
+Because the \emph{PLDA} classifier has a built-in variable selection method, for fair comparison we compared each of the classifiers using the variables selected from each cross-validation fold.
 
@@
 
+\subsection{Classification Results}
+
+TODO\footnote{TODO: Add results discussion and table here.}
+
 \section{Discussion}
@@
 
-\section{Next Steps}
-
-\subsection{Simulations}
-
-\begin{itemize}
-	\item  Design an experiment similar to Friedman's initial simulations, where the the majority of the variances are nearly 0.
-Fix the sample size.
-Consider the error rate as a function of $p$.
-This should cause the RDA regularization to be ineffective because the average eigenvalue is near 0.
-
-	\item  Design an experiment where the average eigenvalue goes to 0, but the average nonzero eigenvalue blows up for a fixed sample size.
-Consider the error rate as a function of $p$.
-This should cause the RDA and MDEB regularizations to be ineffective.
-However, I am banking on the ETRDA methods doing well.
-
-	\item In our simulations, we need to report the estimated values of $c$, the average of the eigenvalues, and the average of the nonzero eigenvalues.
-
-\end{itemize}
-
-\subsection{Notation}
-
-\begin{itemize}
-	\item Define the majority of notation at the beginning of Section 2, including $z^+$, $z_+$, and $I[z]$.
-	\item Define the generalization of RDA in terms of the eigenvectors and eigenvalues of the RDA estimator using $\lambda$ as an index.
-This should remove the need to redefine the eigendecomposition each time we change regularization methods.
-A possible formulation is $\bm v_{jk}^{(\lambda)}$ and $e_{jk}^{(\lambda)}$.
-	\item In \eqref{eq:generalized-qda} we have a slight issue in the formulation of the log-determinant in the second sum.
-The eigenvalue function $\phi$ in the first sum should be the inverse of the function in the second sum.
-Furthermore, we need to define the log-determinant so that the last $p - q$ terms are ignored when we reduce the dimension to $q$.
-\end{itemize}
-
-\subsection{Integrate Section 4 Better}
-
-We should create another figure that shows the shrinkage factor as a function of $q$ for each of the shrinkage factors using one or two data microarray data sets with the Dudoit variable selection method.
-Show that as more features are incorporated into the model, the competing shrinkage methods fail.
-
-Moreover, show the determinants or log-determinants of each covariance matrix. Then discuss the near-equivalence of the shrinkage in both NLDA and RDA.
-
-Perhaps, we should not consider the NLDA classifier at all. Alternatively, we recall simply the regularization method of the authors and use that as motivation for the Soft-ETRDA classifier.
-
-\subsection{ETRDA}
-
-\begin{itemize}
-	\item Look at the shrinkage factors as a function of $q$.
-We may find that the optimal value for $c$ is approximately equal to the average of the nonzero eigenvalues.
-If this is the case, we could note in the discussion that to improve the speed, simply use the average of the nonzero eigenvalues.
-	\item Investigate a quick way to calculate the ETRDA classifier for each value of $c$, similar to RDA.
-For instance, use the Fast SVD with Hard-ETRDA and write enough squiggles to explain this thoroughly.
-\end{itemize}
-
-\subsection{Discuss Replacing RDA Shrinkage with MDEB Shrinkage}
-
-The next step could be to discuss replacing the RDA shrinkage with the MDEB shrinkage, which uses essentially the average nonzero eigenvalues rather than the average eigenvalues.
-However, as DMY and I have discovered, the average of the nonzero eigenvalues often can be too large, resulting in extremely inflated denominators in \eqref{eq:generalized-qda}.
-This has the negative effect of decreasing the scaled differences in the means.
-Show a histogram of the eigenvalues for a data set, such as the Alon data set.
-Draw vertical lines showing the average of the 1) eigenvalues and 2) nonzero eigenvalues.
-
-Discuss that the MDEB approach to regularizing the RDA estimator is inappropriate.
-While it works for the linear case, it does not for the quadratic case because of the determinant.
-Because the average of the nonzero eigenvalues is often very large, the last $p - q$ terms in the sum in \eqref{eq:generalized-qda} are approximately 0 (perhaps, not quite 0).
-In the linear case, the second sum is ignored, so that the MDEB estimator results in a classifier very similar to the one where the M-P pseudo inverse is used.
-However, the rank does not need to be estimated using the MDEB estimator, and the terms in the first sum with eigenvalues near 0 are no longer stable.
-
-Previously, I investigated applying MDEB to the QDA classifier as well but was discouraged because the classification accuracy was quite poor.
-It seems that the second sum in \eqref{eq:generalized-qda} is likely to be blamed.
-Letting rank$(\widehat{\bm\Sigma}_k) = q_k$ and $\gamma_k$ be the average of the nonzero eigenvalues of $\widehat{\bm \Sigma}_k$, the second sum in \eqref{eq:generalized-qda} can be partitioned as
-\begin{align*}
-	\sum_{j=1}^p \ln (e_{jk} + \gamma_k) &= \sum_{j=1}^{q_k} \ln (e_{jk} + \gamma) + \sum_{j=q_k + 1}^p \ln (\gamma_k)\\
-	&= \sum_{j=1}^{q_k} \ln (e_{jk} + \gamma_k) + (p - q_k - 1) \ln (\gamma_k).
-\end{align*}
-My guess is that $(p - q_k - 1) \ln (\gamma_k)$ tends to dominate for the class with either the largest or the smallest sample size.
-Thus, the shrinkage has too much influence on the classification.
-With this in mind, our proposed selection of $c$ may yield a significant improvement to the QDA classifier compared with MDEB applied to the QDA classifier.
-
-While the \emph{RDA} classifier suffers from numerical incalculability as the data dimensionality $p$ increases, the shrinkage factor of the \emph{MDEB} classifier tends to dominate the estimator.
-Because the trace of the covariance-matrix estimator is equal to the sum of its eigenvalues, the average of the nonzero eigenvalues can become very large.
-For each feature added, the variance of that feature must be positive, so it adds to the trace.
-Even if the average of the nonzero eigenvalues yields a convergent series, the shrinkage factor is substantially large.
-
 \bibliographystyle{plainnat}

### 2013-02-11 reports/paper-rda.tex
Updated the paper to include a lot of the theory based on Ye and Wang (2006). Also, reorganized a lot of the paper to be more coherent.

--- 2013-02-05 reports/paper-etrda.tex
@@
 
-\title{Another Look at Regularized Discriminant Analysis for High-dimensional Classification}
+\title{Another Look at Regularized Discriminant Analysis for High-Dimensional Classification}
 
@@
 
-\newtheorem*{thm}{Theorem}
-\newtheorem*{cor}{Corollary}
+\newtheorem{cor}{Corollary}
+\newtheorem{lemma}{Lemma}
+\newtheorem{thm}{Theorem}
 
@@
 \begin{abstract}
-\cite{Friedman:1989tm} proposed the popular regularized discriminant analysis (\emph{RDA}) classifier for the small-sample, high-dimensional setting by employing a regularized covariance matrix estimator, which employs a ridge-like shrinkage factor to ensure positive definiteness.
-We demonstrate that the \emph{RDA} classifier, in its original form, is inadequate for the small-sample, high-dimensional setting.
-We redefine the \emph{RDA} classifier to increase its computational runtime and its interpretation.
-We show that this new definition includes several well-known regularized classifiers as special cases.
-We also consider several properties of the newly defined \emph{RDA} classifier.
-Alternatively, \cite{Srivastava:2007ww} propose an empirical-Bayesian estimator that employs a similar shrinkage factor using the average of the nonzero eigenvalues.
-Because the eigenvalues are often highly skewed, this average is often excessively large, so that the near-zero eigenvalues are substantially overcompensated.
-A comparable eigenvalue adjustment method from \cite{Thomaz:2006ef} often yields a shrinkage factor that is approximately equal to the average of the eigenvalues.
-By writing each shrinkage method as a matrix function, we generalize the shrinkage method of \cite{Thomaz:2006ef} to allow a powerful estimation technique that does not over inflate the near-zero eigenvalues.
-By incorporating our proposed covariance-matrix regularization with the RDA classifier, we overcome the numerical incalculability issues of the \emph{RDA} classifier.
-Moreover, we show that our proposed classifier is often superior to the regularized classifiers mentioned above in terms of classification accuracy using three simulated and three small-sample, high-dimensional microarray data sets.
-Additionally, we provide implementations of the competing classifiers in the {\tt regdiscrim} R package, available on CRAN.
+\cite{Friedman:1989tm} proposed the popular regularized discriminant analysis (\emph{RDA}) classifier for the small-sample, high-dimensional setting by employing a regularized covariance-matrix estimator, which incorporates a ridge-like shrinkage factor by adding a positive constant to the zero-eigenvalue estimates to ensure positive definiteness.
+We show that this shrinkage is unnecessary because shrinking the zero eigenvalues corresponding to the null space of the pooled sample covariance matrix has no effect on the \emph{RDA} classifier.
+Accordingly, we propose a novel generalization of the \emph{RDA} classifier that is more computationally efficient than its original counterpart because the inverses and determinants computed in the decision rule can be reduced from the original feature space to a linear subspace with dimensions corresponding to the individual class sample sizes.
+Additionally, our formulation of the \emph{RDA} classifier allows us to estimate efficiently optimal tuning parameters from a large set of candidate values.
+Furthermore, our generalization simplifies the construction of the covariance-matrix estimator to improve its interpretation and allows for a more flexible classification model by viewing shrinkage applied to the nonzero eigenvalue estimates in terms of matrix functions.
+Fortuitously, this matrix-function formulation includes as special cases the original \emph{RDA} classifier and several other well-known regularized classifiers.
+Moreover, we demonstrate with several high-dimensional microarray data sets that our proposed classifier is competitive and often superior to several recently proposed sparse and regularized classifiers.
+Finally, we provide implementations of the competing classifiers in the {\tt regdiscrim} R package, available on CRAN.
 \end{abstract}
@@
 
-In our first classifier we generalize the eigenvalue replacement employed in the \emph{NLDA} classifier and apply a soft threshold to the eigenvalues of the \emph{RDA} covariance-matrix estimator.
-We refer to our first classifier as the \emph{RDA} with soft eigenvalue-thresholding (\emph{Soft-ETRDA}) classifier.
-As we show, the eigenvalue replacement employed in the \emph{NLDA} classifier is often approximately equal to the shrinkage applied in the \emph{RDA} classifier for small-sample, high-dimensional data.
-Furthermore, a special case of the \emph{Soft-ETRDA} classifier is the \emph{NLDA} classifier as well as the \emph{RDA} classifier.
-We remark that our soft eigenvalue threshold is similar to and partially motivated by the shrunken centroids classifier from \cite*{Tibshirani:2002ht}.
-	In our second proposed classifier, we apply a hard threshold to the eigenvalues of the \emph{RDA} covariance-matrix estimator, so that the number of covariance parameters incorporated into the \emph{RDA} classifier is reduced substantially.
-We refer to this classifier as the \emph{RDA} with hard eigenvalue-thresholding (\emph{Hard-ETRDA}) classifier.
-We remark that the inverse of \emph{Hard-ETRDA} covariance-matrix estimator includes as a special case, the Moore-Penrose pseudo inverse of the \emph{RDA} covariance-matrix estimator.
-
 Using three small-sample, high-dimensional microarray data sets, we demonstrate that our two proposed classifiers often yield superior classification accuracy to the \emph{RDA}, \emph{MDEB}, and \emph{NLDA} classifiers.
@@
 
+Quote from \cite{Ye:2009gd}: ``There is a lack of a systematic study to explore the commonalities and differences of these algorithms, as well as their intrinsic relation- ship.
+This has been a challenging task, since different algorithms apply completely different schemes when dealing with the singularity problem.'' Cite this as an issue with LDA and that regularization has been less explored with QDA.
+Our formulation provides the necessary tools to study this is in detail.
+
+\cite{Ye:2006jm} have shown that shrinking the zero eigenvalues does not improve the classification performance of the \emph{RDA} classifier when $\lambda = 1$.
+Consequently, we instead shrink only the nonzero eigenvalues of $\widehat{\bm\Sigma}_k(\lambda)$, which has two additional benefits.
+First, by considering only the nonzero eigenvalues of $\widehat{\bm\Sigma}_k(\lambda)$, we effectively ignore the eigenvectors corresponding to the zero eigenvalues, which reduces significantly the number of parameters that we incorporate into our \emph{RDA} classifier because $p \gg q_k$ if $p \gg N$.
+Furthermore, the size of the matrix $\bm V_k$ is $p \times N$ as opposed to a $p \times p$ matrix in the original definition of the \emph{RDA} classifier.
+For significantly large $p$, the reduced size of $\bm V_k$ expedites calculation of the \emph{RDA} classifier and reduces its memory usage.\footnote{TODO: Provide an example for a realistic $p$ and $N$.}
+
+As noted by \cite{Hoyle:2011vt}, \cite{Raudys:1998dd}, and \cite{Ramey:2011ji}, the pseudoinverse is often inadvisable because the resulting classifier often yields a sharp decrease in classification accuracy if $p \approx n$.\footnote{TODO: Contrast the computational efficiency we gain with the need to stabilize the smallest eigenvalues (those near $q$).}
+
+We prove several results using arguments similar to \cite{Ye:2006tq}.
+
 We have organized the remainder of the paper as follows.
-In Section 2 we review the \emph{RDA}, the \emph{MDEB}, and \emph{NLDA} classifiers and define each in terms of matrix functions.
-In Section 3 we generalize the eigenvalue adjustment employed in the \emph{NLDA} classifier and define our proposed \emph{Soft-ETRDA} and \emph{Hard-ETRDA} classifiers.
+In Section 2 we present our generalization of the \emph{RDA} classifier after reviewing its original counterpart from \cite{Friedman:1989tm}.
+In Section 3 we show that our proposed classifier includes several well-known regularized classifiers as special cases and briefly discuss some commonalities among them.
 In Section 4 we describe the \textbf{three} simulation configurations and three small-sample, high-dimensional microarray data sets and then demonstrate that our two proposed classifiers are superior to the competing classifiers in terms of classification accuracy.
-We then conclude with a brief discussion in Section 6.
+We then conclude with a brief discussion in Section 5.
 
@@
 
-We\footnote{Define the majority of notation here, including $z^+$, $z_+$, and $I[z]$.} wish to assign correctly an unlabeled $p$-dimensional observation vector $\bm x$ to one of $K$ unique, known classes (or populations) by constructing a classifier from $n$ training observations that can accurately predict the class membership of $\bm x$.
-Let $\bm x_i = (x_{i1}, x_{i2}, \ldots, x_{ip}) \in \mathbb{R}_{p \times 1}$ be the $i$th observation $(i = 1, \ldots, n)$ with true, unique membership $y_i \in \mathcal{K} = \{\omega_1, \ldots, \omega_K\}$, where $\mathbb{R}_{m \times n}$ denotes the matrix space of all $m \times n$ matrices over the real field $\mathbb{R}$.
+In discriminant analysis we wish to assign correctly an unlabeled $p$-dimensional observation vector $\bm x$ to one of $K$ unique, known classes (or populations) by constructing a classifier from $n$ training observations that can accurately predict the class membership of $\bm x$.
+Let $\bm x_i = (x_{i1}, x_{i2}, \ldots, x_{ip}) \in \mathbb{R}_{p \times 1}$ be the $i$th observation $(i = 1, \ldots, n)$ with true, unique membership $y_i \in \{\omega_1, \ldots, \omega_K\}$, where $\mathbb{R}_{a \times b}$ denotes the matrix space of all $a \times b$ matrices over the real field $\mathbb{R}$.
 We assume that $(\bm x_i, y_i)$ is a realization from a mixture distribution $p(\bm x) = \sum_{k=1}^K p(\bm x | \omega_k) p(\omega_k)$, where $p(\bm x | \omega_k)$ is the probability density function (PDF) of the $k$th class, $p(\omega_k)$ is prior probability of class membership of the $k$th class.
@@
 where $\xbar_k$ and $\widehat{\bm\Sigma}_k$ are the MLEs for $\bm \mu_k$ and $\bm \Sigma_k$, respectively.
-If we assume further that the covariance matrix parameters are equal for each class (i.e., $\bm\Sigma_k = \bm\Sigma$ for all $k$), then \eqref{eq:qda} reduces to Fisher's linear discriminant analysis (LDA) classifier,
-\begin{align}
-	D_{LDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\widehat{\bm\Sigma}^{-1}(\bm x - \xbar_k), \label{eq:lda}
-\end{align}
-where $\widehat{\bm\Sigma}$ is the pooled sample covariance matrix
+If we assume further that the covariance matrix parameters are equal for each class (i.e., $\bm\Sigma_k = \bm\Sigma$ for all $k$), then the pooled sample covariance matrix $\widehat{\bm\Sigma}$ is applied instead for each class, where
 \begin{align}
@@
 \end{align}
+In this case \eqref{eq:qda} reduces to Fisher's linear discriminant analysis (LDA) classifier, and the log-determinant is omitted because it is constant across all classes.
+
+\subsection{Covariance Matrix Regularization}
+
+The smallest eigenvalues and the directions associated with their eigenvectors highly influence \eqref{eq:qda}.
+In fact, the eigenvalues of $\widehat{\bm \Sigma}_k$ are well-known to be biased if $p \ge n_k$ such that the smallest eigenvalues are underestimated \citep{Seber:2004uh}, and this bias increases as $p$ increases relative to $n_k$ \citep{TODONeedCitation}.
+Moreover, if $p > n_k$, then rank$(\widehat{\bm \Sigma}_k) \le n_k$, which implies that at least $p - n_k$ eigenvalues of $\widehat{\bm \Sigma}_k$ are 0.
+Therefore, although more feature information is available to discriminate among the $K$ classes, if $p > n_k$, \eqref{eq:qda} is incalculable because $\widehat{\bm \Sigma}_k^{-1}$ does not exist.
+
+Several regularization methods, such as the methods considered by \citet*{Guo:2007te}, \cite{Mkhadri:1995jp}, and \citet*{Xu:2009fl}, have been proposed in the literature to adjust the eigenvalues of $\widehat{\bm \Sigma}_k$ so that \eqref{eq:qda} is calculable and has reduced variability in the standard bias-variance tradeoff.
+These methods often can be written concisely as a matrix function, which we define here using the notation of \cite{Izenman:2008gm}.
+Let $\bm A \in \mathbb{R}_{p \times p}$ be symmetric, and let $\phi:\mathbb{R} \rightarrow \mathbb{R}$ be monotonic.
+Then
+\begin{align}
+	\phi(\bm A) = \sum_{j = 1}^p \phi(e_j) \bm v_j \bm v_j'\label{eq:spectral-decomp}
+\end{align}
+is a matrix function defined explicitly in terms of the eigenvalues of $\bm A$.
+For this reason, $\phi$ is often called a spectral transform \citep{Zhu:2010vu} or a transfer function \citep{Ye:2009gd}.
+Special cases of \eqref{eq:spectral-decomp} include the spectral decomposition of $\bm A$ when $\phi(e_j) = e_j$, the matrix square-root of $\bm A$ when $\phi(e_j) = e_j^{1/2}$, and, if it exists, the inverse of $\bm A$ when $\phi(e_j) = e_j^{-1}$ \citep{Harville:2008wja}.
+
+Notice that if  $\phi(e_{jk}) = e_{jk}$, $k = 1, \ldots, K$, no eigenvalue adjustment is performed, and \eqref{eq:qda} remains incalculable.
+Similarly, if $\phi(e_{jk}) = e_{j}$, $k = 1, \ldots, K$, the \emph{LDA} classifier is employed without eigenvalue shrinkage.
+Additionally, setting $\phi(e_j) = e_j^{+}$ with $z^{+} = 1/z$ if $z > 0$ and $0$ otherwise, we ignore the zero eigenvalues, so that $\phi(\widehat{\bm \Sigma}_k) = \widehat{\bm \Sigma}_k^{+}$, the Moore-Penrose pseudoinverse of $\widehat{\bm \Sigma}_k$.
+
+Another common form of the covariance-matrix regularization applies a shrinkage factor $\gamma_k > 0$, so that
+\begin{align}
+	\phi(e_{jk}) = e_{jk} + \gamma_k, \label{eq:ridge-estimator}
+\end{align}
+similar to a method employed in ridge regression.
+Equation \eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance matrix $\widehat{\bm\Sigma}_k$ towards the $p$-dimensional identity matrix $\bm I_p$, thereby increasing the eigenvalues of $\widehat{\bm\Sigma}_k$ by $\gamma_k$.
+Specifically, the zero eigenvalues are replaced with $\gamma_k$, resulting in $\widehat{\bm \Sigma}_k + \gamma_k \bm I_p$ being positive definite.
+For additional covariance-matrix regularization methods, see \cite{Ye:2009gd} and \cite{Ramey:2011ji}.
 
@@
 If $\lambda = 0$, \eqref{eq:sig-lambda} is equal to $\widehat{\bm\Sigma}_k$ used in the \emph{QDA} classifier in \eqref{eq:qda}.
-Alternatively, if $\lambda = 1$, \eqref{eq:sig-lambda} is equal to $\widehat{\bm\Sigma}$ used in the \emph{LDA} classifier in \eqref{eq:lda}, in which case we assume implicitly that $\bm\Sigma_k = \bm \Sigma$ for $k = 1, \ldots, K$.
+Alternatively, if $\lambda = 1$, \eqref{eq:sig-lambda} reduces to \eqref{eq:pooled-cov} used in the \emph{LDA} classifier.
 
@@
 
-Substituting \eqref{eq:sig-rda} into \eqref{eq:qda}, we have the \emph{RDA} classifier, where we assign an unlabeled observation $\bm x$ to class $\omega_k$:
+\cite{Friedman:1989tm} substituted \eqref{eq:sig-rda} into \eqref{eq:qda} to obtain the \emph{RDA} classifier
 \begin{align}
@@
 \end{align}
+Notice that we can write the shrinkage term in \eqref{eq:sig-rda} as the matrix function
+\begin{align*}
+	\phi(e_{jk}(\lambda)) = (1 - \gamma) e_{jk}(\lambda) + \gamma \bar{e}_k,
+\end{align*}
+for class $\omega_k$, where $e_{jk}(\lambda)$ is the $j$th largest eigenvalue of $\widehat{\bm\Sigma}_k(\lambda)$ and $\bar{e}_k = p^{-1} \sum_{j=1}^p e_{jk}(\lambda)$ is the average of the eigenvalues of \eqref{eq:sig-lambda}.
 
-\subsection{An Alternative Approach to Regularized Discriminant Analysis}
+\subsection{A Generalization of Regularized Discriminant Analysis}
 
-Here, we present a more straightforward, alternative parameterization of the \emph{RDA} classifier. We first define explicitly
+Here, we define our generalization of the \emph{RDA} classifier by first simplifying the formulation of $\widehat{\bm \Sigma}_k(\lambda)$ in \eqref{eq:sig-lambda} and demonstrating its clear interpretation as a linear combination of the crossproducts of the training observations centered by their respective class sample means.
+Then, we generalize the adjustment of the eigenvalues of $\widehat{\bm \Sigma}_k(\lambda)$ by introducing a more general matrix-function formulation.
+This formulation allows us to adjust the eigenvalues of $\widehat{\bm \Sigma}_k(\lambda)$ in a more general manner so that we are not limited to ridge-like estimators.
 \begin{align}
@@
 Thus, we no longer have need for the convex combination of $n_k$ and $N$ as defined by \cite{Friedman:1989tm}.
-Before examining the shrinkage applied to \eqref{eq:sig-lambda-alternative}, we examine some of its properties and write \eqref{eq:sig-lambda-alternative} as
+By rewriting \eqref{eq:sig-lambda-alternative} in terms of the observations $\bm x_i$, $i = 1, \ldots, N$, each centered by its class sample mean, we attain a clear interpretation of $\widehat{\bm\Sigma}_k(\lambda)$.
+We have that
 \begin{align}
@@
 	&= \widehat{\bm\Sigma}_k + \lambda \left(\sum_{k' = 1}^K \frac{n_k}{N} \widehat{\bm\Sigma}_{k'}  - \widehat{\bm\Sigma}_k \right) \nonumber\\
-	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right) \widehat{\bm\Sigma}_k +  \frac{\lambda}{N} \sum_{\substack{k' = 1\\k' \ne k}} n_{k'} \widehat{\bm\Sigma}_{k'}.\label{eq:sig-lambda-alternative2}
+	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right) \widehat{\bm\Sigma}_k +  \frac{\lambda}{N} \sum_{\substack{k' = 1\\k' \ne k}} n_{k'} \widehat{\bm\Sigma}_{k'} \nonumber \\
+	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N I(y_i = k) \bm x_i \bm x_i' +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k) \bm x_i \bm x_i' \nonumber \\
+	&= \sum_{i=1}^N c_{ik}(\lambda) \bm x_i \bm x_i',\label{eq:sig-lambda-alternative2}
 \end{align}
-From \eqref{eq:sig-lambda-alternative2}, we see the weight $\lambda$ places on the covariance matrices from the remaining $K - 1$ classes.
-The interesting part here is that $\lambda$ controls the contribution of each of the $N$ observations to estimating $\Sigma_k$.
-Effectively, this states that for $0 < \lambda \le 1$, we use $N$ observations to estimate $\bm \Sigma_k$ rather than $n_k$ as done in the \emph{QDA} classifier.
-As we will discuss in-depth later, the pooling operation is advantageous in increasing the rank of each $\widehat{\bm\Sigma}_k(\lambda)$ from $n_k - 1$ to $N - 1$ for $0 < \lambda \le 1$.
-Moreover, suppose that we have centered each observation $\bm x_i$, $i = 1, \ldots, N$, by its class sample mean.
-Then, we write \eqref{eq:sig-lambda-alternative2} as
+where $c_{ik}(\lambda) = \lambda N^{-1} + (1 - \lambda)n_k^{-1}I(y_i = k)$.
+From \eqref{eq:sig-lambda-alternative2}, $\lambda$ weights the contribution of each of the $N$ observations to estimating $\bm \Sigma_k$.
+That is, for $0 < \lambda \le 1$, we estimate $\bm \Sigma_k$ with $N$ rather than $n_k$ observations.
+Below, we will show that the pooling operation is advantageous in increasing the rank of each $\widehat{\bm\Sigma}_k(\lambda)$ from rank$(\widehat{\bm\Sigma}_k)$ to rank$(\widehat{\bm\Sigma}_k)$ for $0 < \lambda \le 1$.
+Notice that if $\lambda  = 0$, then the observations from the remaining $K - 1$ classes do not contribute to the estimation of $\bm \Sigma_k$, corresponding to $\widehat{\bm \Sigma}_k$.
+Furthermore, if $\lambda = 1$, the weights in \eqref{eq:sig-lambda-alternative2} reduce to $1/N$, corresponding to $\widehat{\bm\Sigma}$.
+
+As we have discussed, several eigenvalue adjustment methods have been proposed to increase eigenvalues (approximately) equal to 0.
+Here, we define a more general eigenvalue adjustment of \eqref{eq:sig-lambda-alternative} by applying the matrix function $\phi_k(\widehat{\bm \Sigma}_k(\lambda))$.
+To simplify the notation, we write $\tilde{\bm \Sigma}_k = \phi_k(\widehat{\bm \Sigma}_k(\lambda))$.
+Thus, we have the generalized \emph{RDA} (\emph{GRDA}) classifier
 \begin{align}
-	\widehat{\bm\Sigma}_k(\lambda) &= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N I(y_i = k) \bm x_i \bm x_i' +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k) \bm x_i \bm x_i' \nonumber \\
-	&= \sum_{i=1}^N c_{ik}(\lambda) \bm x_i \bm x_i',\label{eq:sig-lambda-alternative3}
+	D_{GRDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\tilde{\bm\Sigma}_k^{-1}(\bm x - \xbar_k)  + \ln |\tilde{\bm\Sigma}_k|, \label{eq:grda}
+\end{align}
+
+\section{Squiggles}
+We require the following relationship about the null spaces of $\widehat{\bm \Sigma}_k(\lambda)$, $\widehat{\bm \Sigma}$, and $\widehat{\bm \Sigma}_k$ in order to attain improved computational efficiency of our generalized \emph{RDA} classifier.
+
+\begin{lemma}\label{lemma:null-spaces}
+Let $\widehat{\bm \Sigma}_k(\lambda)$, $\widehat{\bm \Sigma}$, and $\widehat{\bm \Sigma}_k$ be defined as above.
+Then, $\mathcal{N}(\widehat{\bm \Sigma}_k(\lambda)) \subset \mathcal{N}(\widehat{\bm \Sigma}) \subset \mathcal{N}(\widehat{\bm \Sigma}_k)$, $k = 1, \ldots, K$.
+\end{lemma}
+\begin{proof}
+Let $\bm z \in \mathcal{N}(\widehat{\bm \Sigma}_k(\lambda))$ for some $k = 1, \ldots, K$.
+Hence, $0 = \bm z' \widehat{\bm \Sigma}_k(\lambda) \bm z = (1 - \lambda) \bm z' \widehat{\bm \Sigma}_k \bm z + \lambda \bm z' \widehat{\bm \Sigma} \bm z$.
+Because $\widehat{\bm \Sigma}_k$ and $\widehat{\bm \Sigma}$ are each positive semidefinite, we have $\bm z \in \mathcal{N}(\widehat{\bm \Sigma})$ and $\bm z \in \mathcal{N}(\widehat{\bm \Sigma}_k)$.
+Now, suppose $\bm z \in \mathcal{N}(\widehat{\bm \Sigma})$.
+We have that $0 = \bm z' \widehat{\bm \Sigma} \bm z = N^{-1} \sum_{k = 1}^K n_k \bm z' \widehat{\bm \Sigma}_k \bm z$, which implies that $z \in \mathcal{N}(\widehat{\bm \Sigma}_k)$ once again because $\widehat{\bm \Sigma}_k \in \mathbb{R}_{p \times p}^{\ge}$ for $k = 1, \ldots, K$.
+\end{proof}
+
+Here, we derive an alternative expression for $\tilde{\bm \Sigma}_k$ to decompose the decision rule of our proposed classifier.
+In particular, the alternative expression is terms of the first $q$ regularized eigenvalues of $\tilde{\bm \Sigma}_k$ and the adjustments applied to the $p - q$ zero eigenvalues.
+
+\begin{lemma}\label{lemma:rda-tilde-Sigma_k}
+Suppose that rank$(\widehat{\bm \Sigma}) = q \le p$.
+Let $\widehat{\bm \Sigma} = \bm U \bm D \bm U'$ be the eigendecomposition of $\widehat{\bm \Sigma}$ such that $\bm D \in \mathbb{R}_{p \times p}$ is the diagonal matrix of eigenvalues of $\widehat{\bm \Sigma}$ with
+\begin{align*}
+	\bm D = \begin{bmatrix}
+		\bm D_q & 0\\
+		\bm 0 & \bm 0
+	\end{bmatrix}
+\end{align*}
+and $\bm U \in \mathbb{R}_{p \times p}$ is the matrix whose columns are the corresponding eigenvectors of $\widehat{\bm \Sigma}$.
+Then, partitioning $\bm U = (\bm U_1, \bm U_2)$ so that $\bm U_1 \in \mathbb{R}_{p \times q}$ and $\bm U_2 \in \mathbb{R}_{p \times (p - q)}$, we have the equivalent expression\footnote{To generalize the shrinkage, we define the diagonal matrix of adjusted eigenvalues as $\bm \Phi$ and partition it into two parts: $\bm \Phi_1$ corresponding to the first $q$ eigenvalues of $\widehat{\bm \Sigma}$, and $\bm \Phi_2$ corresponding to the remaining $p - q$ eigenvalues of $\widehat{\bm \Sigma}$.}
+\begin{align}
+	\tilde{\bm \Sigma}_k &= \bm U \begin{bmatrix}
+		\bm W_k & \bm 0\\
+		\bm 0 & \gamma \bm I_{p - q}
+	\end{bmatrix}\bm U',\label{eq:rda-matrix}
 \intertext{where}
-	c_{ik}(\lambda) &= \begin{cases}
-		(1 - \lambda)n_k^{-1} + \lambda N^{-1}, & y_i = k,\\
-		\lambda N^{-1}, & y_i \ne k.\\
-	\end{cases}\nonumber
+\bm W_k &= (1-\gamma) \{(1 - \lambda) \bm U_1' \widehat{\bm \Sigma}_k \bm U_1 + \lambda \bm D_q\} + \gamma \bm I_q.
 \end{align}
-Therefore, after centering each observation by its corresponding class sample mean, we have that $\widehat{\bm\Sigma}_k(\lambda)$ in \eqref{eq:sig-lambda-alternative3} incorporates each centered observation into the estimation of $\bm \Sigma_k$.
-Notice that if $\lambda  = 0$, then the observations from the remaining $K - 1$ classes do not contribute to the estimation of $\bm \Sigma_k$, corresponding to $\widehat{\bm \Sigma}_k$.
-Furthermore, if $\lambda = 1$, the weights in \eqref{eq:sig-lambda-alternative3} reduce to $1/N$, corresponding to $\widehat{\bm\Sigma}$.
+\end{lemma}
+\begin{proof}
+First, notice that
+\begin{align*}
+	\bm U' \tilde{\bm \Sigma}_k \bm U = (1-\gamma) \{(1 - \lambda) \bm U_1' \widehat{\bm \Sigma}_k \bm U_1 + \lambda \bm D_q\} + \gamma \bm U' \bm U.
+\end{align*}
+From Lemma \ref{lemma:null-spaces}, we have that $\widehat{\bm \Sigma}_k \bm U_2 = \bm 0$, which implies that 
+\begin{align*}
+	\bm U' \widehat{\bm \Sigma}_k \bm U = \begin{bmatrix}
+		\bm U_1' \widehat{\bm \Sigma}_k \bm U_1 & \bm 0\\
+		\bm 0 & \bm 0
+	\end{bmatrix},
+\end{align*}
+for $k = 1, \ldots, K$. Thus, recalling that $\bm U$ is orthogonal, we have that 
+\begin{align*}
+	\tilde{\bm \Sigma}_k = \bm U \begin{bmatrix}
+		\bm W_k & \bm 0\\
+		\bm 0 & \gamma \bm I_{p - q}
+	\end{bmatrix}\bm U'.
+\end{align*}
+\end{proof}
 
-\subsection{Covariance Matrix Regularization}
+An immediate consequence of Lemma \ref{lemma:rda-tilde-Sigma_k} is that ranks of the \emph{RDA} covariance matrices for all classes are each equal to the rank of the pooled sample covariance matrix if the pooling parameter $\lambda$ is positive.
+Thus, by incorporating each $\bm x_i$ into the estimation of $\bm \Sigma_k$, we increase the rank.
+We state this formally here.
 
-Although \eqref{eq:qda} is in a standard form, we prefer to express the \emph{QDA} classifier explicitly in terms of the eigenvalues and eigenvectors of $\widehat{\bm \Sigma}_k$.
-Let $e_{jk}(\lambda)$ denote the $j$th largest eigenvalue of $\widehat{\bm\Sigma}_k$ and $\bm v_{jk}(\lambda)$ be the corresponding eigenvector such that $e_{1k}(\lambda) \ge e_{2k}(\lambda) \ge \ldots \ge e_{pk}(\lambda) \ge 0$.
-Equivalently, we write \eqref{eq:qda} as
+\begin{cor}
+Let $\widehat{\bm \Sigma}_k(\lambda)$ be defined as in \eqref{eq:sig-lambda-alternative}. Then, for $\lambda \in (0, 1]$, rank$(\widehat{\bm \Sigma}_k(\lambda)) = q$ for all $k = 1, \ldots, K$.
+\end{cor}
+\begin{proof}
+The proof follows by setting $\gamma = 0$ in Lemma \ref{lemma:rda-tilde-Sigma_k}.
+\end{proof}
+
+In our following result we show that the inverse and determinants of $\bm W_k$ can be written so that they are performed on $n_k \times n_k$ matrices rather than $q \times q$ matrices.
+Given that $n_k$ is often quite small, say, between 15 and 50, this can reduce the amount of computation substantially.
+We apply the well-known Sherman-Woodbury-Morrison formula and matrix determinant lemma.\footnote{TODO: This is the term given on Wikipedia, but I have not seen this phrase elsewhere.
+However, the Wikipedia page references the Harville text.}
+
+\begin{lemma}\label{lemma:rda-W_k}
+Let $\bm W_k$ be defined as above. Then, $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$ and
 	\begin{align}
-  D(\bm x) = \argmin_{k} \sum_{j = 1}^p \frac{[\bm v_{jk}(\lambda)' (\bm x - \bm \xbar_k)]^2}{e_{jk}(\lambda)} + \sum_{j=1}^p \ln \{e_{jk}(\lambda)\}. \label{eq:rda-eigen}
+		\bm W_k^{-1} &= \bm \Gamma_k^{-1} - (1 - \gamma)(1 - \lambda) \bm \Gamma_k^{-1} \bm U_1' \bm X_k' \bm Q_k^{-1} \bm X_k \bm U_1 \bm \Gamma_k^{-1},
 	\end{align}
-The smallest eigenvalues and the directions associated with their eigenvectors highly influence \eqref{eq:rda-eigen}.
-In fact, the eigenvalues of $\widehat{\bm \Sigma}_k(\lambda)$ are well-known to be biased if $p \ge n_k$ such that the smallest eigenvalues are underestimated \citep{Seber:2004uh}.
-This bias increases as $p$ increases relative to $n_k$ \citep{TODONeedCitation}.
-Moreover, if $p > n_k$, then rank$(\widehat{\bm \Sigma}_k(\lambda)) \le n_k$, which implies that the last $p - n_k$ eigenvalues of $\widehat{\bm \Sigma}_k(\lambda)$ are 0, which implies that \eqref{eq:rda-eigen} is incalculable.
-Thus, although more feature information is available to discriminate among the $K$ classes as $p$ increases, classification accuracy decreases unless one obtains enough training-sample observations to reliably estimate the increased number of parameters.
+where $\bm Q_k = \bm I_{n_k} + (1 - \gamma)(1- \lambda) \bm X_k \bm U_1 \bm \Gamma_k^{-1} \bm U_1' \bm X_k'$ and $\bm \Gamma_k = (1 - \gamma) \lambda \bm D_q + \gamma \bm I_q$.
+\end{lemma}
+\begin{proof}
+	Set $\bm A = \bm \Gamma_k$, $\bm B = (1 - \gamma)(1 - \lambda) \bm U_1' \bm X_k'$, $\bm T = \bm I_{n_k}$, and $\bm C = \bm X_k \bm U_1$.
+To calculate $|\bm W_k|$, from \cite{Harville:2008wja} we apply Theorem 18.1.1, which states that $|\bm A + \bm B \bm T \bm C| = |\bm A| |\bm T| |\bm T^{-1} + \bm C \bm A^{-1} \bm B|$.
+Thus, we have $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$.
+Also from \cite{Harville:2008wja}\footnote{TODO: Cite the appropriate theorem from Harville's text.}, we apply the Sherman-Woodbury-Morrison formula $(\bm A + \bm B \bm C)^{-1} = \bm A^{-1} - \bm A^{-1} \bm B(\bm I + \bm C \bm A^{-1} \bm B)^{-1} \bm C \bm A^{-1}$, which implies that $\bm W_k^{-1} = \bm \Gamma_k^{-1} - (1 - \gamma)(1 - \lambda) \bm \Gamma_k^{-1} \bm U_1' \bm X_k' \bm Q_k^{-1} \bm X_k \bm U_1 \bm \Gamma_k^{-1}$.
+\end{proof}
 
-Several regularization methods, such as the methods considered by \citet*{Guo:2007te}, \cite{Mkhadri:1995jp}, and \citet*{Xu:2009fl}, have been proposed in the literature to stabilize the eigenvalues of $\widehat{\bm \Sigma}_k$ used in \eqref{eq:qda}.
-These methods often can be written concisely as a matrix function, which we define here using the notation of \cite{Izenman:2008gm}.
-Let $\bm A \in \mathbb{R}_{p \times p}$ be symmetric, and let $\phi:\mathbb{R} \rightarrow \mathbb{R}$.
-Then
+The following result is essential to our main result.
+Here, we show that the difference of a vector $\bm x \in \mathbb{R}_{p \times 1}$ to each class sample mean is constant across all classes by showing that the difference is in the null space of $\bm U_2'$.
+
+\begin{lemma}\label{lemma:RDA-constant-term}
+Let $\bm U_2$ be defined as above. Then, for all $\bm x \in \mathbb{R}_{p \times 1}$, $\bm U_2' (\bm x - \xbar_k) = \bm U_2' (\bm x - \xbar_{k'})$, $1 \le k, k' \le K$.
+\end{lemma}
+\begin{proof}
+	First, we have that $\xbar_k - \xbar_{k'} = \sum_{i=1}^N \alpha_i \bm x_i$, where $\alpha_i = (n_k n_{k'})^{-1} \{ I(y_i = k) n_{k'} - I(y_i = k') n_k \}$.
+Let $\bm x \in \mathbb{R}_{p \times 1}$.
+Then, $\bm U_2' (\bm x - \xbar_k) = \bm U_2' (\bm x - \xbar_{k'})$ is equivalent to $\bm U_2'(\xbar_k - \xbar_{k'}) = \sum_{i=1}^N \alpha_i \bm U_2' \bm x_i$.
+Recall that $\bm U_2 \in \mathcal{N}(\widehat{\bm \Sigma})$, which implies that $\bm U_2' \in \mathcal{C}(\widehat{\bm \Sigma})^{\perp}$ \citep[Lemma 1.2.5]{Kollo:2005vp}.
+Now, because $\bm x_i \in \mathcal{C}(\widehat{\bm \Sigma})$, for $i = 1, \ldots, N$, we have that $\bm U_2' \bm x_i = \bm 0$, which implies that $\bm U_2'(\xbar_k - \xbar_{k'}) = \bm 0$.
+\end{proof}
+
+Here, we present our main result. We partition the \emph{RDA} decision rule into a sum.
+
+\begin{thm}
+	The \emph{RDA} decision rule in \eqref{eq:rda} is equivalent to
 	\begin{align}
-	\phi(\bm A) = \sum_{j = 1}^p \phi(e_j) \bm v_j \bm v_j'\label{eq:spectral-decomp}
+		D_{RDA}(\bm x) &= \argmin_k (\bm x - \xbar_k)' \bm U_1 \bm W_k^{-1} \bm U_1' (\bm x - \xbar_k)' + \ln |\bm Q_k| + (p - q) \ln \gamma + q \ln \{ (1 - \gamma) \lambda \} \label{eq:rda-decomposed}
 	\end{align}
-is a matrix function defined explicitly in terms of the eigenvalues of $\bm A$.
-For this reason, $\phi$ is often called a spectral transform \citep{Ch15SSLBook} or a transfer function \citep{Ye:2009gd, Chapelle2002SSLChapter15Reference} .
-Special case of \eqref{eq:spectral-decomp} include the spectral decomposition of $\bm A$ when $\phi(e_j) = e_j$, the matrix square-root of $\bm A$ when $\phi(e_j) = e_j^{1/2}$, and, if it exists, the inverse of $\bm A$ when $\phi(e_j) = e_j^{-1}$ \citep{Harville:2008wja}.
 
-Viewing regularized covariance-matrix estimators as matrix functions, we can write a more general form of the \emph{QDA} classifier as
-\begin{align}
-  D(\bm x) = \argmin_{k} \sum_{j = 1}^p \phi(e_{jk}(\lambda))[\bm v_{jk}(\lambda)' (\bm x - \bm \mu_k)]^2 + \sum_{j=1}^p \ln \{ \phi(e_{jk}(\lambda)) \}. \label{eq:generalized-qda}
-\end{align}
-The\footnote{In \eqref{eq:generalized-qda} we have a slight issue in the formulation of the log-determinant in the second sum.
-The eigenvalue function $\phi$ in the first sum should be the inverse of the function in the second sum.
-Furthermore, we need to define the log-determinant so that the last $p - q$ terms are ignored when we reduce the dimension to $q$.} generalized \emph{QDA} classifier in \eqref{eq:generalized-qda} includes several common regularized classifiers.
-Notice that if  $\phi(e_{jk}(\lambda)) = e_{jk}(\lambda)$, $k = 1, \ldots, K$, \eqref{eq:generalized-qda} reduces to the \emph{QDA} classifier in \eqref{eq:rda-eigen}.
-Furthermore, if $\phi(e_{jk}(\lambda)) = e_{j}$, $k = 1, \ldots, K$, \eqref{eq:generalized-qda} reduces to the \emph{LDA} classifier in \eqref{eq:lda}, and the second summand in \eqref{eq:generalized-qda} is constant across all classes and can be omitted.
-Additionally, one common approach is to ignore the zero eigenvalues when $p > n$ by employing the Moore-Penrose pseudoinverse $\widehat{\bm \Sigma}_k^{+}$, which employs the matrix function $\phi(e_j) = e_j^{+}$, where $z^{+} = 1/z$ if $z > 0$ and $0$ otherwise.
-As noted by \cite{Hoyle:2011vt}, \cite{Raudys:1998dd}, and \cite{Ramey:2011ji}, the pseudoinverse is often inadvisable because the resulting classifier often yields a sharp decrease in classification accuracy if $p \approx n$.
+\end{thm}
+\begin{proof}
+	From \eqref{eq:rda-matrix}, we have that
+	\begin{align*}
+		\tilde{\bm \Sigma}_k^{-1} = \bm U \begin{bmatrix}
+			\bm W_k^{-1} & \bm 0\\
+			\bm 0 & \gamma^{-1} \bm I_{p - q}
+		\end{bmatrix} \bm U'
+	\end{align*}
+	and $|\tilde{\bm \Sigma}_k| = \gamma^{p-q} | \bm W_k |$. Therefore, for all $\bm x \in \mathbb{R}_{p \times 1}$, we have that
+	\begin{align*}
+	(\bm x - \xbar_k)' \tilde{\bm\Sigma}_k^{-1}(\bm x - \xbar_k)  + \ln |\tilde{\bm\Sigma}_k| &= (\bm x - \xbar_k)' \bm U_1 \bm W_k^{-1} \bm U_1' (\bm x - \xbar_k)'\\
+	&+ \gamma^{-1} (\bm x - \xbar_k)' \bm U_2 \bm U_2' (\bm x - \xbar_k) + \ln | \bm W_k | + (p - q) \ln \gamma.
+	\end{align*}
+Applying Lemma \ref{lemma:RDA-constant-term}, $\bm U_2' (\bm x - \xbar_k)$ is constant for all $k = 1, \ldots, K$.
+Also, from Lemma \ref{lemma:rda-W_k}, we have that $\ln |\bm W_k| = \ln |\bm \Gamma_k| + \ln |\bm Q_k|$, and because \eqref{eq:rda-decomposed} is independent of $\ln |\bm D_q|$, the proof follows.
+\end{proof}
 
-Another common form of the covariance-matrix regularization applies a shrinkage factor $\gamma_k > 0$, so that
-\begin{align}
-	\phi(e_{jk}(\lambda)) = (e_{jk}(\lambda) + \gamma_k)^{-1}, \label{eq:ridge-estimator}
-\end{align}
-is a ridge-like estimator, similar to a method employed in ridge regression.
-The estimator in \eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance matrix $\widehat{\bm\Sigma}_k(\lambda)$ towards the $p$-dimensional identity matrix $\bm I_p$, thereby increasing the eigenvalues of $\widehat{\bm\Sigma}_k(\lambda)$ by $\gamma_k$ so that they are positive, resulting in $\widehat{\bm \Sigma}_k(\lambda) + \gamma_k \bm I_p$ being positive definite.
-
-We emphasize the regularized estimator in \eqref{eq:ridge-estimator} because the \emph{RDA} and \emph{MDEB} classifiers can be written in this form as we show below.
-Also, we present the \emph{NLDA} and \emph{ETRDA} classifiers in terms of matrix functions using \eqref{eq:generalized-qda}.
-For additional covariance-matrix regularization methods, see \cite{Ye:2009gd} and \cite{Ramey:2011ji}.
-
-\section{Properties of the Redefined RDA Classifier}
+\section{Properties of the Generalized RDA Classifier}
 
@@
 
-\subsection{Covariance-Matrix Regularization}
-
-We are now ready to discuss the shrinkage employed in the \emph{RDA} covariance-matrix estimator.
-As we are interested primarily in the case that $p \gg N$, rank$\{\widehat{\bm\Sigma}_k(\lambda)\} = q_k < p$, which implies that $p - q_k$ of the eigenvalues of $\widehat{\bm\Sigma}_k(\lambda)$ are equal to zero.
-Thus, for $\gamma > 0$, shrinking $\widehat{\bm\Sigma}_k(\lambda)$ towards $\bm I_p$ as in \eqref{eq:sig-rda} increases the rank of $\widehat{\bm\Sigma}_k(\lambda)$ from $q_k$ to $p$.
-However, \cite{Ye:2009gd} and \cite{YeOtherPaperCitedInTheFirst} have shown that shrinking the zero eigenvalues does not improve the classification performance of the \emph{RDA} classifier.
-Consequently, we instead shrink only the nonzero eigenvalues of $\widehat{\bm\Sigma}_k(\lambda)$, which has two additional benefits.
-First, by considering only the nonzero eigenvalues of $\widehat{\bm\Sigma}_k(\lambda)$, we effectively ignore the eigenvectors corresponding to the zero eigenvalues, which reduces significantly the number of parameters that we incorporate into our \emph{RDA} classifier because $p \gg q_k$ if $p \gg N$.
-Furthermore, the size of the matrix $\bm V_k$ is $p \times N$ as opposed to a $p \times p$ matrix in the original definition of the \emph{RDA} classifier.
-For significantly large $p$, the reduced size of $\bm V_k$ expedites calculation of the \emph{RDA} classifier and reduces its memory usage.\footnote{TODO: Provide an example for a realistic $p$ and $N$.}
-
-Specifically, we shrink only the nonzero eigenvalues of $\widehat{\bm \Sigma}_k(\lambda)$ by utilizing the estimator
-
-\begin{align}
-	\widehat{\bm\Sigma}_k(\lambda, \gamma) = (1 - \gamma) \widehat{\bm\Sigma}_k(\lambda) + \gamma \frac{\tr\{\widehat{\bm\Sigma}_k(\lambda)\}}{p} (\bm I_{q_k} \oplus \bm 0_{p - q_k}),\label{eq:sig-rda-alternative}
-\end{align}
-where $\bm A \oplus \bm B$ denote the direct sum of $\bm A \in \mathbb{R}_{r \times r}$ and $\bm B \in \mathbb{R}_{s \times s}$ \citep[Chapter 1]{Lutkepohl:1996uz}.
-Therefore, substituting \eqref{eq:sig-rda-alternative} into \eqref{eq:rda}, we have our alternative parameterization of the \emph{RDA} classifier.
-
-\section{Regularized Discriminant Analysis with Eigenvalue Thresholding}
-
-\subsection{The Hard-ETRDA Classifier}
-
-Although the shrinkage applied in \eqref{eq:cov-nlda} is ineffective, the form in \eqref{eq:cov-nlda} allows us to consider a more general eigenvalue thresholding approach to resolve the numerical issues of the \emph{RDA} classifier discussed below.
-We are interested in a general shrinkage value $\Delta$ instead of the average of the eigenvalues $\bar{e}_k$.
-We propose to shrink the pooled covariance matrix estimator in \eqref{eq:sig-lambda} with a more general eigenvalue threshold, motivated by the \emph{NLDA} classifier.
-That is, for $\Delta_0, \Delta > 0$, we define the matrix function
-\begin{align}
-	\phi(e_{jk}) = (\Delta_0 + e_{jk} \cdot I\{e_{jk} > \Delta\} )^{+},\label{eq:cov-hard-etrda}
-\end{align}
-where $e_{jk}$ are the eigenvalues of $\widehat{\bm \Sigma}_k(\lambda)$ in \eqref{eq:sig-lambda} and $\bm v_{jk}$ are the corresponding eigenvectors.
-Substituting \eqref{eq:cov-soft-etrda} into \eqref{eq:generalized-qda}, we arrive at the \emph{ETRDA} classifier.
-
-For $\Delta_0 = \Delta = \gamma$, the shrinkage employed in \eqref{eq:cov-hard-etrda} is equal to \eqref{eq:ridge-estimator}.
-Hence, \eqref{eq:cov-hard-etrda} includes the regularization methods of \cite{Friedman:1989tm} and \cite{Srivastava:2007ww} as special cases.
-To see this, notice that if $\gamma$ is the average of the eigenvalues, \eqref{eq:cov-hard-etrda} is equal to \eqref{eq:rda-eigenvalue-function}.
-Furthermore, if $\gamma$ is the average of the nonzero eigenvalues \eqref{eq:cov-hard-etrda} is equal to \eqref{eq:cov-mdeb}.
-Notice also that if $\Delta_0 = \Delta = 0$, then \eqref{eq:cov-hard-etrda} reduces to the Moore-Penrose pseudoinverse \citep{Harville:2008wja}.
-
-\subsection{The Soft-ETRDA Classifier}
-
-As an alternative eigenvalue threshold, we consider a soft threshold.
-Although the shrinkage applied in \eqref{eq:cov-nlda} is ineffective, the form in \eqref{eq:cov-nlda} allows us to consider a more general eigenvalue thresholding approach to resolve the numerical issues of the \emph{RDA} classifier discussed below.
-We are interested in a general shrinkage value $\Delta$ instead of the average of the eigenvalues $\bar{e}_k$.
-We propose to shrink the pooled covariance matrix estimator in \eqref{eq:sig-lambda} with a more general eigenvalue threshold, motivated by the \emph{NLDA} classifier.
-That is, for $\Delta > 0$, we define the matrix function
-\begin{align}
-	\phi(e_{jk}) = (\Delta_0 + \{e_{jk} - \Delta\}_+)^{+},\label{eq:cov-soft-etrda}
-\end{align}
-where $e_{jk}$ are the eigenvalues of $\widehat{\bm \Sigma}_k(\lambda)$ in \eqref{eq:sig-lambda} and $\bm v_{jk}$ are the corresponding eigenvectors.
-Substituting \eqref{eq:cov-soft-etrda} into \eqref{eq:generalized-qda}, we arrive at the \emph{ETRDA} classifier.
-
-For small values of $\gamma$ such that $\Delta_0 = \Delta = \gamma$, the shrinkage employed in \eqref{eq:cov-soft-etrda} is approximately equal to \eqref{eq:ridge-estimator} because the near-zero are adjusted by $\gamma$, while the larger eigenvalues are hardly adjusted by $\gamma$.
-Because of this approximation, the regularization methods of \cite{Friedman:1989tm} and \cite{Srivastava:2007ww} are approximately special cases of the \eqref{eq:cov-soft-etrda}.
-To see this, notice that if $\gamma$ is the average of the eigenvalues, \eqref{eq:cov-soft-etrda} is equal to \eqref{eq:cov-nlda}, which is approximately equal to \eqref{eq:rda-eigenvalue-function}.
-Furthermore, if $\gamma$ is the average of the nonzero eigenvalues \eqref{eq:cov-soft-etrda} is approximately equal to \eqref{eq:cov-mdeb}.
-
-Notice that \eqref{eq:cov-soft-etrda} has a form similar to the soft thresholding employed in the nearest shrunken centroids classifier from \cite{Tibshirani:2002ht} in the case that the matrix of eigenvectors of $\widehat{\bm \Sigma}_k$ is the identity matrix and $\Delta_0 = 0$.
-Our choice of $\Delta_0 = 0$ is also advantageous because it can reduce the number of dimensions included in the \emph{Soft-ETRDA} classifier, thereby decreasing the classification reducing.
-Noting the similarity of \eqref{eq:cov-soft-etrda} to the soft threshold of \cite{Tibshirani:2002ht}, we follow \cite{Tibshirani:2002ht} and select $\Delta$ by cross-validation, resulting in a superior classifier to the \emph{RDA} classifier proposed by \cite{Friedman:1989tm}.
-
-For small values of $\gamma$, the eigenvalue thresholding employed in the \emph{Soft-ETRDA} and \emph{Hard-ETRDA} classifiers are approximately equal.
-Both methods are effective regularization methods in stabilizing the smallest eigenvalues.
-The soft eigenvalue threshold employed in the \emph{Soft-ETRDA} classifier has the distinct advantage that the largest eigenvalues are reduced, which smooths the distribution of sample eigenvalues, yielding an effective classifier that is often superior to the proposed sparse and regularized classifiers considered.
-
 \section{Special Cases}
-
-\subsection{RDA}
-
-\cite{Friedman:1989tm} also uses a regularization parameter $\gamma$ to shrink the
-eigenvalues of \eqref{eq:sig-lambda} towards the $p$-dimensional identity matrix $\bm I_p$
-in order to stabilize the inverse of \eqref{eq:sig-lambda}, resulting in a biased
-covariance-matrix estimator with matrix function
-\begin{align}
-	\phi(e_{jk}(\lambda)) = (\{1 - \gamma\} e_{jk}(\lambda) + \gamma \bar{e}_k)^{-1},\label{eq:rda-eigenvalue-function}
-\end{align}
-for class $\omega_k$, where $\gamma \in [0, 1]$, $e_{jk}(\lambda)$ be the $j$th largest eigenvalue of $\widehat{\bm\Sigma}_k(\lambda)$ in \eqref{eq:sig-lambda}, $\bm v_{jk}(\lambda)$ be the corresponding eigenvector, and $\bar{e}_k = p^{-1} \sum_{j=1}^p e_{jk}(\lambda)$ is the average of the eigenvalues of \eqref{eq:sig-lambda}.
 
@@
 \begin{align}
-	\phi(e_{jk}) = \left(e_{jk} + n_k^{-1} \sum_{j = 1}^p e_{jk}\right)^{-1},\label{eq:cov-mdeb}
+	\phi(e_{jk}) = e_{jk} + n_k^{-1} \sum_{j = 1}^p e_{jk},\label{eq:cov-mdeb}
 \end{align}
@@
 \begin{align}
-	\phi(e_{jk}) = (\bar{e}_k + \{e_j - \bar{e}_k\}_+)^{-1},\label{eq:cov-nlda}
+	\phi(e_{jk}) = \bar{e}_k + (e_j - \bar{e}_k)_+,\label{eq:cov-nlda}
 \end{align}
@@
 
+\section{Notation}
+
+$\bm A \oplus \bm B$ denote the direct sum of $\bm A \in \mathbb{R}_{r \times r}$ and $\bm B \in \mathbb{R}_{s \times s}$ \citep[Chapter 1]{Lutkepohl:1996uz}.
+
 \section{Monte Carlo Simulations}
-
-\begin{itemize}
-	\item TODO: Design an experiment similar to Friedman's initial simulations, where the the majority of the variances are nearly 0.
-Fix the sample size.
-Consider the error rate as a function of $p$.
-This should cause the RDA regularization to be ineffective because the average eigenvalue is near 0.
-
-	\item TODO: Design an experiment where the average eigenvalue goes to 0, but the average nonzero eigenvalue blows up for a fixed sample size.
-Consider the error rate as a function of $p$.
-This should cause the RDA and MDEB regularizations to be ineffective.
-However, I am banking on the ETRDA methods doing well.
-
-\end{itemize}
 
@@
 
-In this section, we describe three high-dimensional microarray data sets and compare our proposed \emph{ETRDA} classifier with two recently proposed classifiers for small-sample, high-dimensional data: the penalized \emph{LDA} (\emph{PLDA}) classifier from \cite{Witten:2011kc}  and the shrinkage-mean-based DLDA (\emph{SmDLDA}) classifier from \cite{Tong:2012hw}.
+In this section, we describe three high-dimensional microarray data sets and compare our proposed \emph{GRDA} classifier with two recently proposed classifiers for small-sample, high-dimensional data: the penalized \emph{LDA} (\emph{PLDA}) classifier from \cite{Witten:2011kc}, the shrinkage-mean-based DLDA (\emph{SmDLDA}) classifier from \cite{Tong:2012hw}, the \emph{RDA} classifier as defined in \cite{Hastie:2008dt}, and the sparse discriminant analysis (\emph{SDA}) classifier from \cite{Clemmensen:2011kr}.
 We calculated 10-fold cross-validation error rates \citep{Hastie:2008dt} to evaluate the performance of each classifier considered.

### 2013-02-14 reports/paper-rda.tex
Rewrote introduction. Significant changes to the discussion on RDA and our proposed generalization.

@@
 \newtheorem{lemma}{Lemma}
+\newtheorem{proposition}{Proposition}
 \newtheorem{thm}{Theorem}
@@
 Microarrays have enabled researchers to measure gene-expression levels of thousands of genes in a single experiment to identify a subset of genes and their expression patterns that either characterize a particular cell state or facilitate an accurate prediction of future cell state.
-We focus on the classification of tissues and other biological samples into one of $K$ disjoint, predefined cancer types, where our goal is to determine accurately the class membership of a new tissue sample based on gene-expression microarray data.
-However, microarray data typically consist of a large number of gene expressions $p$ relative to the sample size $n$.
-Moreover, standard benchmark microarray data sets consist of between $p = 10,000$ and $55,000$ probes and fewer than $n = 100$ observations, in which case standard supervised-learning methods exhibit poor classification performance, due to the \emph{curse of dimensionality} \citep{Bellman:1961tn}.
-Furthermore, well-known classifiers, including linear discriminant analysis \emph{(LDA)} and quadratic discriminant analysis \emph{(QDA)}, are incalculable when $p > n$ because the sample covariance matrices employed in both classifiers are singular.
-As a result, many researchers, including \cite*{Merchante:2012vk}, \cite{Witten:2011kc}, \cite*{Pang:2009ik}, \cite*{Huang:2010ju}, and \cite*{Clemmensen:2011kr}, have restricted the \emph{LDA} and \emph{QDA} models to reduce number of covariance parameters that must be estimated to ensure that the resulting classifiers are calculable.
-For instance, \cite{Dudoit:2002ev} have proposed a modified \emph{LDA} classifier where the class covariance matrices are assumed to be diagonal.
+We focus on the classification of tissues and other biological samples into one of $K$ disjoint, predefined cancer types, where our goal is to ascertain the class membership of a new tissue sample based on gene-expression microarray data.
+However, microarray data typically consist of a large number of gene expressions $p$ relative to the sample size $N$.
+Moreover, standard benchmark microarray data sets usually consist of between $p = 10,000$ and $55,000$ probes and fewer than $N = 100$ observations, in which case standard supervised-learning methods exhibit poor classification performance, due to the \emph{curse of dimensionality} \citep{Bellman:1961tn}.
+Furthermore, well-known classifiers, including linear discriminant analysis \emph{(LDA)} and quadratic discriminant analysis \emph{(QDA)}, are incalculable when $p > N$ because the sample covariance matrices employed in both classifiers are singular.
+As a result, many researchers, including \cite*{Merchante:2012vk}, \cite{Witten:2011kc}, \cite*{Pang:2009ik}, \cite*{Huang:2010ju}, and \cite*{Clemmensen:2011kr}, have introduced restrictions on the population covariance matrices in the \emph{LDA} and \emph{QDA} models to reduce number of covariance parameters that must be estimated to ensure that the resulting classifiers are calculable.
+For instance, \cite{Dudoit:2002ev} have proposed the diagonal \emph{LDA} (\emph{DLDA}) classifier, a modification of the \emph{LDA} classifier where the class covariance matrices are assumed to be diagonal.
 
-Alternatively, several researchers, including \cite*{Zhang:2010va}, \cite*{Guo:2007te}, \cite{Srivastava:2007ud}, and \cite{Mkhadri:1995jp}, have proposed regularization methods to improve the estimation of the population covariance matrices before applying the \emph{LDA} and \emph{QDA} classifiers.
+Alternatively, several researchers, including \cite*{Zhang:2010va}, \cite{Ye:2009gd}, \cite{Ji:2008wp}, \cite*{Guo:2007te}, \cite{Srivastava:2007ud}, \cite{Ye:2006vx}, \cite*{Thomaz:2006ef}, \cite{Ye:2005uu}, \cite{Mkhadri:1995jp}, and \cite{Peck:1982tt}, have proposed regularization methods to improve the estimation of the population covariance matrices before applying the \emph{LDA} and \emph{QDA} classifiers.
 Covariance-matrix regularization methods often are based on the well-known ridge-regression approach from \cite{Hoerl:1970cd} and adjust the eigenvalues of the sample covariance matrices to ensure that the estimator is positive definite.
-This shrinkage stabilizes eigenvalues that are near zero to ensure that the inverse of the the sample covariance matrix is well-posed.
+Typically, a positive shrinkage value is added to each eigenvalue estimate, replacing the zero eigenvalues with the shrinkage constant and stabilizing the eigenvalues near zero to ensure that the inverse of the the sample covariance matrix is well-posed.
+This common form of covariance-matrix regularization is equivalent to summing the original covariance-matrix estimator with the identity matrix scaled by the shrinkage constant.
 For an excellent overview of the regularization problem in discriminant analysis, see \cite{Murphy:2012uq} and \cite{Mkhadri:1997gy}.
 
-Here, we consider specifically the  popular regularized discriminant analysis (\emph{RDA}) classifier from \cite{Friedman:1989tm} that incorporates a convex combination of the sample covariance matrices from the \emph{LDA} and \emph{QDA} classifiers.
-Because the resulting biased covariance-matrix estimator often remains singular in the small-sample, high-dimensional setting, \cite{Friedman:1989tm} applies covariance-matrix shrinkage, where the resulting covariance-matrix estimator is shrunken towards a scaled identity matrix.
-\cite{Friedman:1989tm} have shown that the \emph{RDA} classifier can attain improved classification accuracy, compared to the \emph{LDA} and \emph{QDA} classifiers when $n > p$.
+Here, we consider specifically the popular regularized discriminant analysis (\emph{RDA}) classifier from \cite{Friedman:1989tm}, who has proposed a weighted average of the covariance-matrix estimators from the \emph{LDA} and \emph{QDA} classifiers.
+Because the resulting biased covariance-matrix estimator can remain singular or nearly singular, \cite{Friedman:1989tm} has additionally incorporated applied covariance-matrix shrinkage, where the shrinkage constant is a function of the average eigenvalue of the original estimator.
+\cite{Friedman:1989tm} has shown that the \emph{RDA} classifier can attain improved classification accuracy, compared to the \emph{LDA} and \emph{QDA} classifiers when $p > N$.
+However, in its original form, the \emph{RDA} classifier is impractical when $p \gg N$ because the \emph{RDA} decision rule requires the calculation of the inverses and determinants of the $p \times p$ covariance-matrix estimators for each class.
+Furthermore, the \emph{RDA} classifier has two tuning parameters that are selected via cross-validation, which adds a substantial computational burden to the model-selection process.
+Despite the introduction of a computational shortcut by \cite{Friedman:1989tm} for model selection to alleviate some of the computation, the \emph{RDA} classifier is ill-advised on modern data sets because the model selection requires on the order of $Kp^3$ calculations for each tuning-parameter pair for each fold in the cross-validation scheme.
 
-In this paper we present a generalization of the \emph{RDA} classifier in terms of matrix functions.
-We show that several common covariance-matrix regularization methods can be specified in terms of a matrix function including the \emph{RDA} classifier as well as methods from \cite{Srivastava:2007ww} and \cite*{Thomaz:2006ef} and several sparse diagonal classifiers, including the diagonal \emph{LDA} (\emph{DLDA}) and \emph{QDA} (\emph{DQDA}) classifiers from \cite{Dudoit:2002ev}, and some of its variants including those proposed by \cite{Pang:2009ik}.
-Our generalization elucidates some properties of these covariance-regularization methods and the relationships among them.
-Moreover, our generalization allows us to view the covariance-matrix regularization methods in terms of eigenvalue thresholds, which further allows us to identify situations where each method is ineffective and ineffective.
-From a perspective of eigenvalue thresholds, we propose two new classifiers that are often superior to the \emph{RDA} classifier in terms of classification performance because both methods employ a more robust eigenvalue thresholding to regularize the convex combination of covariance matrix estimators employed in the \emph{RDA} classifier.
-Our two proposed methods also have an advantage over the \emph{RDA} classifier because both methods employ dimension reduction by reducing the number of basis functions included in the \emph{RDA} classifier, thereby reducing the variance of the resulting classifiers.
+In this paper we address the computational shortcomings of the \emph{RDA} classifier.
+Applying a technique similar to \cite{Ye:2006tq}, we decompose the \emph{RDA} classifier and show that the sum related to the null space of the pooled sample covariance matrix can be discarded.
+In turn, this omission provides a substantial increase in the computational efficiency of the \emph{RDA} classifier and its model-selection.
+In fact, we show rigorously that the inverse and determinants can be reduced to calculations of $n_k \times n_k$ matrices, where $n_k$ is the sample size of the $k$th class, $k = 1, \ldots, K$.
+This reduction in calculation is clearly advantageous because $n_k$ is often quite small, say, 15--20, in microarray data sets.
+Thus, we demonstrate that the calculation of the \emph{RDA} decision rule involving inverses and determinants of extremely large covariance matrices is equivalent a decision rule involving inverses and determinants on much smaller matrices.
 
-Using three small-sample, high-dimensional microarray data sets, we demonstrate that our two proposed classifiers often yield superior classification accuracy to the \emph{RDA}, \emph{MDEB}, and \emph{NLDA} classifiers.
-We also include in our comparison the alternatively-defined \emph{RDA} classifier from \cite*{Hastie:2008dt}, the regularized \emph{LDA} classifier from \cite{Guo:2007te}, a recently proposed variant of diagonal discriminant analysis from \cite*{Tong:2012hw}, and the penalized \emph{LDA} classifier from \cite{Witten:2011kc}.
-Additionally, we examine the classification accuracy of the classifiers considered using \textbf{three} simulation configurations.
-Furthermore, we show that our proposed eigenvalue-threshold methods are more numerically stable than the other considered regularization methods using the simulated configurations.
+Our improvements to the calculation and model selection of the \emph{RDA} classifier yields a natural generalization of the \emph{RDA} classifier.
+In our proposed generalization we first simplify the pooling matrix in the \emph{RDA} classifier by constructing instead a convex combination of the covariance-matrix estimators employed in the \emph{LDA} and \emph{QDA} classifiers.
+This simplification is advantageous in that it improves the interpretation of the pooling operation of the \emph{RDA} classifier.
+Next, we consider covariance-matrix regularization in terms of matrix functions, which includes a large family of shrinkage methods, such as the methods from \cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and a wide variety of other shrinkage methods studied by \cite{Ramey:2011ji} and \cite*{Xu:2009fl}.
+While a large number of these methods can be viewed as variants of ridge regression, we consider the more general case, where a real monotonic function is applied to the eigenvalue estimates.
+We show that adjusting the zero eigenvalues with a general matrix function has no effect on the \emph{RDA} classifier.
+That is, we show that the \emph{RDA} classifier is invariant to any adjustments of the eigenvalues corresponding to the null space of the pooled sample covariance matrix.
+Moreover, there is no classificatory information available in the null space of the pooled sample covariance matrix.
+Hence, no gain in classification performance can be obtained by adjusting these zero eigenvalues.
+Although \cite{Ji:2008wp} and \cite{Ye:2006jm} have provided a similar invariance property for the \emph{LDA} decision rule, the \emph{LDA} classifier involves the assumption that the population covariance matrices are equal, which is often too stringent \citep{Clemmensen:2011kr}.
+The \emph{RDA} classifier allows us to relax this assumption by including unrestricted covariance matrices in the model.
+The \emph{RDA} classifier is then a compromise between the two approaches, favoring the stronger linear assumption if a larger value of the pooling parameter is selected as optimal.
+Additionally, in the linear case  \cite{Ye:2009gd} have remarked that  a lack of a systematic study has been conduced to explore the relationships among various shrinkage methods.
+Our formulation provides a general framework from which we can elucidate properties of covariance-regularization methods and the relationships among them in the linear case, the quadratic case, and the cases in between.
 
-Quote from \cite{Ye:2009gd}: ``There is a lack of a systematic study to explore the commonalities and differences of these algorithms, as well as their intrinsic relation- ship.
-This has been a challenging task, since different algorithms apply completely different schemes when dealing with the singularity problem.'' Cite this as an issue with LDA and that regularization has been less explored with QDA.
-Our formulation provides the necessary tools to study this is in detail.
-
-\cite{Ye:2006jm} have shown that shrinking the zero eigenvalues does not improve the classification performance of the \emph{RDA} classifier when $\lambda = 1$.
-Consequently, we instead shrink only the nonzero eigenvalues of $\widehat{\bm\Sigma}_k(\lambda)$, which has two additional benefits.
-First, by considering only the nonzero eigenvalues of $\widehat{\bm\Sigma}_k(\lambda)$, we effectively ignore the eigenvectors corresponding to the zero eigenvalues, which reduces significantly the number of parameters that we incorporate into our \emph{RDA} classifier because $p \gg q_k$ if $p \gg N$.
-Furthermore, the size of the matrix $\bm V_k$ is $p \times N$ as opposed to a $p \times p$ matrix in the original definition of the \emph{RDA} classifier.
-For significantly large $p$, the reduced size of $\bm V_k$ expedites calculation of the \emph{RDA} classifier and reduces its memory usage.\footnote{TODO: Provide an example for a realistic $p$ and $N$.}
-
-As noted by \cite{Hoyle:2011vt}, \cite{Raudys:1998dd}, and \cite{Ramey:2011ji}, the pseudoinverse is often inadvisable because the resulting classifier often yields a sharp decrease in classification accuracy if $p \approx n$.\footnote{TODO: Contrast the computational efficiency we gain with the need to stabilize the smallest eigenvalues (those near $q$).}
-
-We prove several results using arguments similar to \cite{Ye:2006tq}.
+In addition to the improvement in computational performance of classifiers with regularized covariance matrices, we demonstrate that our generalized \emph{RDA} classifier is competitive and often superior to several recent classifiers designed for small-sample, high-dimensional data.
+In particular, we compare our generalized RDA classifier with an alternatively-defined \emph{RDA} classifier from \cite*{Hastie:2008dt}, the regularized \emph{LDA} classifier from \cite{Guo:2007te}, a recently proposed variant of the \emph{DLDA} classifier from \cite*{Tong:2012hw}, and the penalized \emph{LDA} classifier from \cite{Witten:2011kc}.
+We conduct our comparison of the classification accuracy of the competing classifiers using three small-sample, high-dimensional microarray data sets and examine the classification accuracy of the competing classifiers using \textbf{three} simulation configurations.
 
@@
 \section{Regularized Discriminant Analysis}
+\label{sec:rda}
 
-In discriminant analysis we wish to assign correctly an unlabeled $p$-dimensional observation vector $\bm x$ to one of $K$ unique, known classes (or populations) by constructing a classifier from $n$ training observations that can accurately predict the class membership of $\bm x$.
+In\footnote{TODO: Add paragraph of other notation here, including null space, orthogonal complement, and direct sum.
+Example: $\bm A \oplus \bm B$ denote the direct sum of $\bm A \in \mathbb{R}_{r \times r}$ and $\bm B \in \mathbb{R}_{s \times s}$ \citep[Chapter 1]{Lutkepohl:1996uz}.} discriminant analysis we wish to assign correctly an unlabeled $p$-dimensional observation vector $\bm x$ to one of $K$ unique, known classes (or populations) by constructing a classifier from $n$ training observations that can accurately predict the class membership of $\bm x$.
 Let $\bm x_i = (x_{i1}, x_{i2}, \ldots, x_{ip}) \in \mathbb{R}_{p \times 1}$ be the $i$th observation $(i = 1, \ldots, n)$ with true, unique membership $y_i \in \{\omega_1, \ldots, \omega_K\}$, where $\mathbb{R}_{a \times b}$ denotes the matrix space of all $a \times b$ matrices over the real field $\mathbb{R}$.
@@
 Because the parameters are typically unknown, we estimate the unknown parameters $\bm \mu_k$ and $\bm\Sigma_k$ with their maximum likelihood estimators (MLEs) and substitute the MLEs into the \emph{QDA} classifier.
-Assuming, for simplicity, that the prior probabilities of class membership $p(\omega_k)$ are equal for $k = 1, \ldots, K$, we assign an unlabeled observation $\bm x$ to class $\omega_k$ using the sample \emph{QDA} classifier:
+Assuming that the prior probabilities of class membership $p(\omega_k)$ are equal for $k = 1, \ldots, K$, we assign an unlabeled observation $\bm x$ to class $\omega_k$ with the sample \emph{QDA} classifier
 \begin{align}
@@
 where $\xbar_k$ and $\widehat{\bm\Sigma}_k$ are the MLEs for $\bm \mu_k$ and $\bm \Sigma_k$, respectively.
-If we assume further that the covariance matrix parameters are equal for each class (i.e., $\bm\Sigma_k = \bm\Sigma$ for all $k$), then the pooled sample covariance matrix $\widehat{\bm\Sigma}$ is applied instead for each class, where
+If we assume further that the population covariance matrices are equal for each class (i.e., $\bm\Sigma_k = \bm\Sigma$ for all $k$), then the pooled sample covariance matrix $\widehat{\bm\Sigma}$ is substituted for $\widehat{\bm \Sigma}_k$ in \eqref{eq:qda}, where
 \begin{align}
@@
 \end{align}
-In this case \eqref{eq:qda} reduces to Fisher's linear discriminant analysis (LDA) classifier, and the log-determinant is omitted because it is constant across all classes.
+is the MLE for $\bm \Sigma$. In this case \eqref{eq:qda} reduces to Fisher's \emph{LDA} classifier, and the log-determinant is omitted because it is constant across all classes.
 
@@
 The smallest eigenvalues and the directions associated with their eigenvectors highly influence \eqref{eq:qda}.
-In fact, the eigenvalues of $\widehat{\bm \Sigma}_k$ are well-known to be biased if $p \ge n_k$ such that the smallest eigenvalues are underestimated \citep{Seber:2004uh}, and this bias increases as $p$ increases relative to $n_k$ \citep{TODONeedCitation}.
-Moreover, if $p > n_k$, then rank$(\widehat{\bm \Sigma}_k) \le n_k$, which implies that at least $p - n_k$ eigenvalues of $\widehat{\bm \Sigma}_k$ are 0.
-Therefore, although more feature information is available to discriminate among the $K$ classes, if $p > n_k$, \eqref{eq:qda} is incalculable because $\widehat{\bm \Sigma}_k^{-1}$ does not exist.
+In fact, the eigenvalues of $\widehat{\bm \Sigma}_k$ are well-known to be biased if $p \ge n_k$ such that the smallest eigenvalues are underestimated \citep{Seber:2004uh}, Moreover, if $p > n_k$, then rank$(\widehat{\bm \Sigma}_k) \le n_k$, which implies that at least $p - n_k$ eigenvalues of $\widehat{\bm \Sigma}_k$ are 0.
+Hence, the bias of the eigenvalue estimators increases as $p$ increases relative to $n_k$.
+Furthermore, although more feature information is available to discriminate among the $K$ classes, if $p > n_k$, \eqref{eq:qda} is incalculable because $\widehat{\bm \Sigma}_k^{-1}$ does not exist.
 
@@
 is a matrix function defined explicitly in terms of the eigenvalues of $\bm A$.
-For this reason, $\phi$ is often called a spectral transform \citep{Zhu:2010vu} or a transfer function \citep{Ye:2009gd}.
+For this reason, $\phi$ is often called a spectral transform \citep*{Zhu:2010vu} or a transfer function \citep{Ye:2009gd}.
 Special cases of \eqref{eq:spectral-decomp} include the spectral decomposition of $\bm A$ when $\phi(e_j) = e_j$, the matrix square-root of $\bm A$ when $\phi(e_j) = e_j^{1/2}$, and, if it exists, the inverse of $\bm A$ when $\phi(e_j) = e_j^{-1}$ \citep{Harville:2008wja}.
@@
 Specifically, the zero eigenvalues are replaced with $\gamma_k$, resulting in $\widehat{\bm \Sigma}_k + \gamma_k \bm I_p$ being positive definite.
-For additional covariance-matrix regularization methods, see \cite{Ye:2009gd} and \cite{Ramey:2011ji}.
+For additional covariance-matrix regularization methods, see \cite{Ramey:2011ji}, \cite{Xu:2009fl}, and \cite{Ye:2009gd}.
 
@@
 Then, we generalize the adjustment of the eigenvalues of $\widehat{\bm \Sigma}_k(\lambda)$ by introducing a more general matrix-function formulation.
-This formulation allows us to adjust the eigenvalues of $\widehat{\bm \Sigma}_k(\lambda)$ in a more general manner so that we are not limited to ridge-like estimators.
+We define
 \begin{align}
@@
 	\widehat{\bm\Sigma}_k(\lambda) &= \widehat{\bm\Sigma}_k + \lambda (\widehat{\bm\Sigma} - \widehat{\bm\Sigma}_k) \nonumber\\
-	&= \widehat{\bm\Sigma}_k + \lambda \left(\sum_{k' = 1}^K \frac{n_k}{N} \widehat{\bm\Sigma}_{k'}  - \widehat{\bm\Sigma}_k \right) \nonumber\\
+%	&= \widehat{\bm\Sigma}_k + \lambda \left(\sum_{k' = 1}^K \frac{n_k}{N} \widehat{\bm\Sigma}_{k'}  - \widehat{\bm\Sigma}_k \right) \nonumber\\
 	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right) \widehat{\bm\Sigma}_k +  \frac{\lambda}{N} \sum_{\substack{k' = 1\\k' \ne k}} n_{k'} \widehat{\bm\Sigma}_{k'} \nonumber \\
-	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N I(y_i = k) \bm x_i \bm x_i' +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k) \bm x_i \bm x_i' \nonumber \\
+%	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N I(y_i = k) \bm x_i \bm x_i' +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k) \bm x_i \bm x_i' \nonumber \\
 	&= \sum_{i=1}^N c_{ik}(\lambda) \bm x_i \bm x_i',\label{eq:sig-lambda-alternative2}
@@
 As we have discussed, several eigenvalue adjustment methods have been proposed to increase eigenvalues (approximately) equal to 0.
-Here, we define a more general eigenvalue adjustment of \eqref{eq:sig-lambda-alternative} by applying the matrix function $\phi_k(\widehat{\bm \Sigma}_k(\lambda))$.
+In this section we define a more general eigenvalue adjustment of \eqref{eq:sig-lambda-alternative} by applying the matrix function $\phi_k(\widehat{\bm \Sigma}_k(\lambda))$.
 To simplify the notation, we write $\tilde{\bm \Sigma}_k = \phi_k(\widehat{\bm \Sigma}_k(\lambda))$.
@@
 \end{align}
+The specification of $\phi_k$ in \eqref{eq:grda} allows for a large family of eigenvalue shrinkage methods to be examined and for their relationships to be examined in depth.
 
-\section{Squiggles}
-We require the following relationship about the null spaces of $\widehat{\bm \Sigma}_k(\lambda)$, $\widehat{\bm \Sigma}$, and $\widehat{\bm \Sigma}_k$ in order to attain improved computational efficiency of our generalized \emph{RDA} classifier.
+\section{The Generalized RDA Classifier}
+In this section, we demonstrate that our proposed generalization of the \emph{RDA} classifier facilitates an efficient calculation of \eqref{eq:grda} as well as a rapid model selection of the pooling parameter $\lambda$.
+Our approach is to decompose \eqref{eq:grda} into a sum of two summands, where the first summand consists of matrix operations applied to low-dimensional matrices.
+The second summand corresponds to the null space of $\widehat{\bm \Sigma}$ in \eqref{eq:pooled-cov}.
+For $p \gg N$, the operation applied to this null space involve determinants and inverses of high-dimensional matrices.
+We show that these operations are constant across all classes and can be omitted.
+A byproduct of this omission is that adjustments to the associated eigenvalues have no effect on \eqref{eq:grda}.
+That is, \eqref{eq:grda} is invariant to adjustments of these eigenvalues.
+We then proceed by showing that matrix operations applied to the low-dimensional matrices in the first summand can be further improved to yield significantly lower computational costs than the \emph{RDA} classifier originally proposed by \cite{Friedman:1989tm}.
+Lastly, we exploit the so-called Fast singular value decomposition (SVD) to construct the eigenvalue decomposition of $\widehat{\bm \Sigma}$.
+
+First, we require the following relationship regarding the null spaces of $\widehat{\bm \Sigma}_k(\lambda)$, $\widehat{\bm \Sigma}$, and $\widehat{\bm \Sigma}_k$.
 
 \begin{lemma}\label{lemma:null-spaces}
-Let $\widehat{\bm \Sigma}_k(\lambda)$, $\widehat{\bm \Sigma}$, and $\widehat{\bm \Sigma}_k$ be defined as above.
-Then, $\mathcal{N}(\widehat{\bm \Sigma}_k(\lambda)) \subset \mathcal{N}(\widehat{\bm \Sigma}) \subset \mathcal{N}(\widehat{\bm \Sigma}_k)$, $k = 1, \ldots, K$.
+Let $\widehat{\bm \Sigma}_k$ and $\widehat{\bm \Sigma}$ be the MLEs of $\bm \Sigma_k$ and $\bm \Sigma$ as discussed in Section \ref{sec:rda}, and let $\widehat{\bm \Sigma}_k(\lambda)$ be defined as in \eqref{eq:sig-lambda-alternative}.
+Then, for $k = 1, \ldots, K$, $\mathcal{N}(\widehat{\bm \Sigma}_k(\lambda)) \subset \mathcal{N}(\widehat{\bm \Sigma}) \subset \mathcal{N}(\widehat{\bm \Sigma}_k)$.
 \end{lemma}
@@
 Hence, $0 = \bm z' \widehat{\bm \Sigma}_k(\lambda) \bm z = (1 - \lambda) \bm z' \widehat{\bm \Sigma}_k \bm z + \lambda \bm z' \widehat{\bm \Sigma} \bm z$.
-Because $\widehat{\bm \Sigma}_k$ and $\widehat{\bm \Sigma}$ are each positive semidefinite, we have $\bm z \in \mathcal{N}(\widehat{\bm \Sigma})$ and $\bm z \in \mathcal{N}(\widehat{\bm \Sigma}_k)$.
+Because $\widehat{\bm \Sigma}_k, \widehat{\bm \Sigma}\in \mathbb{R}_{p \times p}^{\ge}$, we have $\bm z \in \mathcal{N}(\widehat{\bm \Sigma})$ and $\bm z \in \mathcal{N}(\widehat{\bm \Sigma}_k)$.
+In particular, we have that $\mathcal{N}(\widehat{\bm \Sigma}_k(\lambda)) \subset \mathcal{N}(\widehat{\bm \Sigma})$.
 Now, suppose $\bm z \in \mathcal{N}(\widehat{\bm \Sigma})$.
-We have that $0 = \bm z' \widehat{\bm \Sigma} \bm z = N^{-1} \sum_{k = 1}^K n_k \bm z' \widehat{\bm \Sigma}_k \bm z$, which implies that $z \in \mathcal{N}(\widehat{\bm \Sigma}_k)$ once again because $\widehat{\bm \Sigma}_k \in \mathbb{R}_{p \times p}^{\ge}$ for $k = 1, \ldots, K$.
+Similarly, we have that $0 = \bm z' \widehat{\bm \Sigma} \bm z = N^{-1} \sum_{k = 1}^K n_k \bm z' \widehat{\bm \Sigma}_k \bm z$, which implies that $z \in \mathcal{N}(\widehat{\bm \Sigma}_k)$ because $\widehat{\bm \Sigma}_k \in \mathbb{R}_{p \times p}^{\ge}$ for $k = 1, \ldots, K$.
+Therefore, $\mathcal{N}(\widehat{\bm \Sigma}) \subset \mathcal{N}(\widehat{\bm \Sigma}_k)$.
 \end{proof}
@@
 
-In our following result we show that the inverse and determinants of $\bm W_k$ can be written so that they are performed on $n_k \times n_k$ matrices rather than $q \times q$ matrices.
-Given that $n_k$ is often quite small, say, between 15 and 50, this can reduce the amount of computation substantially.
-We apply the well-known Sherman-Woodbury-Morrison formula and matrix determinant lemma.\footnote{TODO: This is the term given on Wikipedia, but I have not seen this phrase elsewhere.
-However, the Wikipedia page references the Harville text.}
-
-\begin{lemma}\label{lemma:rda-W_k}
-Let $\bm W_k$ be defined as above. Then, $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$ and
-	\begin{align}
-		\bm W_k^{-1} &= \bm \Gamma_k^{-1} - (1 - \gamma)(1 - \lambda) \bm \Gamma_k^{-1} \bm U_1' \bm X_k' \bm Q_k^{-1} \bm X_k \bm U_1 \bm \Gamma_k^{-1},
-	\end{align}
-where $\bm Q_k = \bm I_{n_k} + (1 - \gamma)(1- \lambda) \bm X_k \bm U_1 \bm \Gamma_k^{-1} \bm U_1' \bm X_k'$ and $\bm \Gamma_k = (1 - \gamma) \lambda \bm D_q + \gamma \bm I_q$.
-\end{lemma}
-\begin{proof}
-	Set $\bm A = \bm \Gamma_k$, $\bm B = (1 - \gamma)(1 - \lambda) \bm U_1' \bm X_k'$, $\bm T = \bm I_{n_k}$, and $\bm C = \bm X_k \bm U_1$.
-To calculate $|\bm W_k|$, from \cite{Harville:2008wja} we apply Theorem 18.1.1, which states that $|\bm A + \bm B \bm T \bm C| = |\bm A| |\bm T| |\bm T^{-1} + \bm C \bm A^{-1} \bm B|$.
-Thus, we have $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$.
-Also from \cite{Harville:2008wja}\footnote{TODO: Cite the appropriate theorem from Harville's text.}, we apply the Sherman-Woodbury-Morrison formula $(\bm A + \bm B \bm C)^{-1} = \bm A^{-1} - \bm A^{-1} \bm B(\bm I + \bm C \bm A^{-1} \bm B)^{-1} \bm C \bm A^{-1}$, which implies that $\bm W_k^{-1} = \bm \Gamma_k^{-1} - (1 - \gamma)(1 - \lambda) \bm \Gamma_k^{-1} \bm U_1' \bm X_k' \bm Q_k^{-1} \bm X_k \bm U_1 \bm \Gamma_k^{-1}$.
-\end{proof}
-
 The following result is essential to our main result.
@@
 
-\section{Properties of the Generalized RDA Classifier}
+\subsection{Model Selection}
 
-\subsection{Computational Advantages}
+In our following result we show that the inverse and determinants of $\bm W_k$ can be written so that they are performed on $n_k \times n_k$ matrices rather than $q \times q$ matrices.
+Given that $n_k$ is often quite small, say, between 15 and 50, this can reduce the amount of computation substantially.
+We apply the well-known Woodbury formula and a related determinant lemma.
 
-Thus far, our alternative parameterization simplifies the definition of the \emph{RDA} classifier and has an improved interpretation in terms of the contribution of each observation.
-Here, we demonstrate that our parameterization yields significant computational advantages over the originally proposed \emph{RDA} classifier from \cite{Friedman:1989tm} by writing \eqref{eq:sig-lambda-alternative3} as
+\begin{proposition}\label{proposition:rda-W_k}
+Let $\bm W_k$ be defined as above. Then, $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$ and
+	\begin{align}
+		\bm W_k^{-1} &= \bm \Gamma_k^{-1} - (1 - \gamma)(1 - \lambda) \bm \Gamma_k^{-1} \bm U_1' \bm X_k' \bm Q_k^{-1} \bm X_k \bm U_1 \bm \Gamma_k^{-1},
+	\end{align}
+where $\bm Q_k = \bm I_{n_k} + (1 - \gamma)(1- \lambda) \bm X_k \bm U_1 \bm \Gamma_k^{-1} \bm U_1' \bm X_k'$ and $\bm \Gamma_k = (1 - \gamma) \lambda \bm D_q + \gamma \bm I_q$.
+\end{proposition}
+\begin{proof}
+	Set $\bm A = \bm \Gamma_k$, $\bm B = (1 - \gamma)(1 - \lambda) \bm U_1' \bm X_k'$, $\bm T = \bm I_{n_k}$, and $\bm C = \bm X_k \bm U_1$.
+To calculate $|\bm W_k|$, from \cite{Harville:2008wja} we apply Theorem 18.1.1, which states that $|\bm A + \bm B \bm T \bm C| = |\bm A| |\bm T| |\bm T^{-1} + \bm C \bm A^{-1} \bm B|$.
+Thus, we have $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$.
+Also from Theorem 18.2.8 in \cite{Harville:2008wja}, we apply the Woodbury formula $(\bm A + \bm B \bm C)^{-1} = \bm A^{-1} - \bm A^{-1} \bm B(\bm I + \bm C \bm A^{-1} \bm B)^{-1} \bm C \bm A^{-1}$, which implies that $\bm W_k^{-1} = \bm \Gamma_k^{-1} - (1 - \gamma)(1 - \lambda) \bm \Gamma_k^{-1} \bm U_1' \bm X_k' \bm Q_k^{-1} \bm X_k \bm U_1 \bm \Gamma_k^{-1}$.
+\end{proof}
+
+Note that we can write the pooled sample covariance matrix in \eqref{eq:pooled-cov} as
 \begin{align*}
-	\widehat{\bm\Sigma}_k(\lambda) &= \bm X_k(\lambda)' \bm X_k(\lambda),
+	\widehat{\bm\Sigma} = \bm X' \bm X,
 \end{align*}
-where $\bm X_k(\lambda) = [\sqrt{c_{1k}(\lambda)} \bm x_1', \ldots,  \sqrt{c_{Nk}(\lambda)} \bm x_N']'$.
-Now, we can apply the Fast SVD to compute the eigenvalue decomposition of $\widehat{\bm\Sigma}_k(\lambda)$ efficiently.
-By writing the SVD of $\bm X_k(\lambda) = \bm U_k \bm D_k \bm V_k'$, we have the spectral decomposition $\widehat{\bm\Sigma}_k(\lambda) = \bm X_k(\lambda)' \bm X_k(\lambda) = \bm V_k \bm D_k^2 \bm V_k'$, where $\bm U_k \in \mathbb{R}_{N \times N}$ is orthogonal, $\bm D_k \in \mathbb{R}_{N \times N}$ is a diagonal matrix consisting of the singular values of $\bm X_k(\lambda)$, and $\bm V_k \in \mathbb{R}_{p \times N}$ is orthogonal.
-Notice that $\bm X_k(\lambda)\bm X_k(\lambda)' = \bm U_k \bm D_k^2 \bm U_k'$, so that the eigenvalue decomposition of $\bm X_k(\lambda)\bm X_k(\lambda)'$ yields $\bm U_k$ and $\bm D_k$.
-This is much faster when $p \gg N$ because $\bm X_k(\lambda)\bm X_k(\lambda)' \in \mathbb{R}_{N \times N}$.
-Also, notice that the singular values of $\bm D_k$ are nonzero, so that $\bm D_k^{-1}$ exists.
-After obtaining $\bm U_k$ and $\bm D_k$, we compute the matrix consisting of the first $N$ eigenvectors of $\widehat{\bm\Sigma}_k(\lambda)$ by simply computing\footnote{TODO: Update this paragraph to theorem/proof.}
+where we define $\bm X = [\sqrt{c_{1k}(1)} \bm x_1', \ldots,  \sqrt{c_{Nk}(1)} \bm x_N']'$.
+Now, we can apply the Fast SVD to compute the eigenvalue decomposition of $\widehat{\bm\Sigma}$ efficiently.
+By writing the SVD of $\bm X = \bm U \bm D \bm V'$, we have the spectral decomposition $\widehat{\bm\Sigma} = \bm X' \bm X = \bm V \bm D^2 \bm V'$, where $\bm U \in \mathbb{R}_{N \times N}$ is orthogonal, $\bm D \in \mathbb{R}_{N \times N}$ is a diagonal matrix consisting of the singular values of $\bm X$, and $\bm V \in \mathbb{R}_{p \times N}$ is orthogonal.
+Notice that $\bm X \bm X' = \bm U \bm D^2 \bm U'$, so that the eigenvalue decomposition of $\bm X \bm X'$ yields $\bm U$ and $\bm D$.
+This is much faster when $p \gg N$ because $\bm X \bm X' \in \mathbb{R}_{N \times N}$.
+Also, notice that rank$(\widehat{\bm \Sigma}) = q < N$ so that $\bm D^{-1}$ does not exist.
+After obtaining $\bm U$ and $\bm D$, we compute the matrix consisting of the first $N$ eigenvectors of $\widehat{\bm\Sigma}$ by simply computing\footnote{TODO: Update this paragraph to a proposition along with proof.}
 \begin{align*}
-	\bm V_k = \bm X_k(\lambda)' \bm U_k \bm D_k^{-1}.
+	\bm V = \bm X' \bm U \bm D^{-1}.
 \end{align*}
 
-\section{Special Cases}
+We can determine $q$ efficiently by computing the number of nonzero eigenvalues.
+However, in our experience, we must take care because the $N-q$ smallest eigenvalues are numerically nonzero, albeit they are near zero.
+Thus, we select $q$ to be the number of eigenvalues that exceeds some pre specified tolerance value, say, $10^{-6}$.
 
-\subsection{Minimum Distance Empirical Bayes}
+As noted by \cite{Hoyle:2011vt}, \cite{Raudys:1998dd}, and \cite{Ramey:2011ji}, the pseudoinverse is often inadvisable because the resulting classifier often yields a sharp decrease in classification accuracy if $p \approx N$.\footnote{TODO: Mention that the care recommended above is related to the pseudoinverse and the issue with the smallest eigenvalues.}
 
-\cite{Srivastava:2007ww} derive an empirical-Bayes covariance-matrix estimator under the assumption that the data are independent and identically distributed multivariate normal observations.
+Cross-Validation\footnote{TODO: Use the Sherman-Morrison formula for cross-validation.}
+
+\section{Monte Carlo Simulations}
+
+Our\footnote{TODO: Look at Ye's (2005) JMLR paper on his choice of the $\bm M$ matrix.
+He has a nice way of saying that he considers special cases.} generalization of the \emph{RDA} classifier facilitates a large number of potential models based on the specification of the matrix functions to shrink the eigenvalues.
+In our simulations below, we consider two such eigenvalue value functions.
+First, ...
+Second, we consider the matrix function from \cite{Srivastava:2007ww}, who derive an empirical-Bayes covariance-matrix estimator under the assumption that the data are independent and identically distributed multivariate normal observations.
 \cite{Srivastava:2007ww} then substitute the estimator into the \eqref{eq:lda} to attain the minimum distance empirical Bayes (\emph{MDEB}) classifier.
@@
 where the shrinkage factor in \eqref{eq:cov-mdeb} is approximately equal to the average of the nonzero eigenvalues of $\widehat{\bm \Sigma}_k$.
-Substituting \eqref{eq:cov-mdeb} into \eqref{eq:generalized-qda}, we have the \emph{MDEB} classifier.
 We note that \cite{Srivastava:2007ww} set assumed that $\bm \Sigma_k = \bm \Sigma$ so that $\phi(e_{jk}) = \phi(e_k)$, $k = 1, \ldots, K$.
 
-\cite{Srivastava:2007ww} claim that the \emph{MDEB} classifier is best regularization method they have encountered in the literature.
+\cite{Srivastava:2007ww} claim that the \emph{MDEB} classifier is the best regularization method they have encountered in the literature.
 When $p \le n$, the shrinkage factors of the \emph{RDA} and \emph{MDEB} classifiers are equal, but for $p > n$ this shrinkage can differ substantially.
 In fact, \cite{Ramey:2011ji} show empirically that the \emph{MDEB} classifier can yield superior classification accuracy to the \emph{RDA} classifier.
-
-\subsection{New LDA}
-Whereas shrinkage methods often have the form given in \eqref{eq:ridge-estimator}, the covariance matrix estimator from \cite{Thomaz:2006ef} stabilizes the inverse by replacing the smallest eigenvalues of \eqref{eq:pooled-cov} with the average of the eigenvalues of \eqref{eq:pooled-cov}.
-Let $e_{jk}$ denote the $j$th largest eigenvalue of $\widehat{\bm \Sigma}_k$, and let $\bar{e}_k$ be the average of the eigenvalues of \eqref{eq:pooled-cov}.
-Then, the inverse of the covariance-matrix estimator has the matrix function
-\begin{align}
-	\phi(e_{jk}) = \bar{e}_k + (e_j - \bar{e}_k)_+,\label{eq:cov-nlda}
-\end{align}
-with $z_+ = z$ if $z > 0$ and 0 otherwise.
-\cite*{Xu:2009fl} and \cite{Ramey:2011ji} demonstrate that the substituting \eqref{eq:cov-nlda} into \eqref{eq:generalized-qda} yields classification accuracies comparable to the \emph{RDA} and \emph{MDEB} classifiers for small-sample, high-dimensional microarray data sets.
-Similar to the \emph{MDEB} classifier, \cite{Thomaz:2006ef} assume that $\bm \Sigma_k = \bm \Sigma$ so that $\phi(e_{jk}) = \phi(e_k)$, $k = 1, \ldots, K$.
-
-\section{Notation}
-
-$\bm A \oplus \bm B$ denote the direct sum of $\bm A \in \mathbb{R}_{r \times r}$ and $\bm B \in \mathbb{R}_{s \times s}$ \citep[Chapter 1]{Lutkepohl:1996uz}.
-
-\section{Monte Carlo Simulations}
 
@@
 
+Question to answer: What do we do when $\lambda = 0$, corresponding to \emph{QDA}?
+
 It is interesting to note that our approach is similar to an aside made by Friedman (1989) in the case that the RDA covariance matrix estimator is singular.

### 2013-02-17 reports/paper-rda.tex
Cleaned up much of the discussion of the rigor in the paper.

@@
 
-In\footnote{TODO: Add paragraph of other notation here, including null space, orthogonal complement, and direct sum.
-Example: $\bm A \oplus \bm B$ denote the direct sum of $\bm A \in \mathbb{R}_{r \times r}$ and $\bm B \in \mathbb{R}_{s \times s}$ \citep[Chapter 1]{Lutkepohl:1996uz}.} discriminant analysis we wish to assign correctly an unlabeled $p$-dimensional observation vector $\bm x$ to one of $K$ unique, known classes (or populations) by constructing a classifier from $n$ training observations that can accurately predict the class membership of $\bm x$.
+In discriminant analysis we wish to assign correctly an unlabeled $p$-dimensional observation vector $\bm x$ to one of $K$ unique, known classes (or populations) by constructing a classifier from $n$ training observations that can accurately predict the class membership of $\bm x$.
 Let $\bm x_i = (x_{i1}, x_{i2}, \ldots, x_{ip}) \in \mathbb{R}_{p \times 1}$ be the $i$th observation $(i = 1, \ldots, n)$ with true, unique membership $y_i \in \{\omega_1, \ldots, \omega_K\}$, where $\mathbb{R}_{a \times b}$ denotes the matrix space of all $a \times b$ matrices over the real field $\mathbb{R}$.
@@
 \section{The Generalized RDA Classifier}
-In this section, we demonstrate that our proposed generalization of the \emph{RDA} classifier facilitates an efficient calculation of \eqref{eq:grda} as well as a rapid model selection of the pooling parameter $\lambda$.
-Our approach is to decompose \eqref{eq:grda} into a sum of two summands, where the first summand consists of matrix operations applied to low-dimensional matrices.
-The second summand corresponds to the null space of $\widehat{\bm \Sigma}$ in \eqref{eq:pooled-cov}.
-For $p \gg N$, the operation applied to this null space involve determinants and inverses of high-dimensional matrices.
-We show that these operations are constant across all classes and can be omitted.
-A byproduct of this omission is that adjustments to the associated eigenvalues have no effect on \eqref{eq:grda}.
-That is, \eqref{eq:grda} is invariant to adjustments of these eigenvalues.
-We then proceed by showing that matrix operations applied to the low-dimensional matrices in the first summand can be further improved to yield significantly lower computational costs than the \emph{RDA} classifier originally proposed by \cite{Friedman:1989tm}.
-Lastly, we exploit the so-called Fast singular value decomposition (SVD) to construct the eigenvalue decomposition of $\widehat{\bm \Sigma}$.
+In this section, we demonstrate that our proposed generalization of the \emph{RDA} classifier facilitates an efficient calculation of \eqref{eq:grda} and a rapid model selection of the pooling parameter $\lambda$.
+Our approach is to decompose \eqref{eq:grda} into a sum of two components, where the first summand consists of matrix operations applied to low-dimensional matrices and the second summand corresponds to the null space of $\widehat{\bm \Sigma}$ in \eqref{eq:pooled-cov}.
+We show that the matrix operations performed on the null space of $\widehat{\bm \Sigma}$ yield constant quadratic forms across all classes and can be omitted.
+For $p \gg N$, the constant component involves determinants and inverses of high-dimensional matrices, and by ignoring these calculations, a substantial reduction in computational costs is achieved relative to the \emph{RDA} classifier originally proposed by \cite{Friedman:1989tm}.
+Furthermore, a byproduct of this omission is that adjustments to the associated eigenvalues have no effect on \eqref{eq:grda}.
+Lastly, we exploit the so-called Fast singular value decomposition (SVD) to construct the eigenvalue decomposition of $\widehat{\bm \Sigma}$, further reducing the computational costs of our proposed \emph{RDA} classifier.
 
@@
 
-Here, we derive an alternative expression for $\tilde{\bm \Sigma}_k$ to decompose the decision rule of our proposed classifier.
-In particular, the alternative expression is terms of the first $q$ regularized eigenvalues of $\tilde{\bm \Sigma}_k$ and the adjustments applied to the $p - q$ zero eigenvalues.
+In Lemma \ref{lemma:rda-tilde-Sigma_k} below, we derive an alternative expression for $\tilde{\bm \Sigma}_k$ in terms of the eigenvectors of $\widehat{\bm \Sigma}$.
+Let $\widehat{\bm \Sigma} = \bm U \bm D \bm U'$ be the eigendecomposition of $\widehat{\bm \Sigma}$ such that $\bm D \in \mathbb{R}_{p \times p}$ is the diagonal matrix of eigenvalues of $\widehat{\bm \Sigma}$ with $\bm D = \bm D_q \oplus \bm 0_{p-q}$, the columns of $\bm U \in \mathbb{R}_{p \times p}$ are the corresponding eigenvectors of $\widehat{\bm \Sigma}$, and rank$(\widehat{\bm \Sigma}) = q$, where $\bm A \oplus \bm B$ denotes the direct sum of $\bm A \in \mathbb{R}_{r \times r}$ and $\bm B \in \mathbb{R}_{s \times s}$ \citep[Chapter 1]{Lutkepohl:1996uz}.
+Then, partitioning $\bm U = (\bm U_1, \bm U_2)$ such that $\bm U_1 \in \mathbb{R}_{p \times q}$ and $\bm U_2 \in \mathbb{R}_{p \times (p - q)}$, we see that the component of $\tilde{\bm \Sigma}_k$ corresponding to the null space of $\widehat{\bm \Sigma}$ is in terms of the matrix function $\phi_k$ alone.
 
 \begin{lemma}\label{lemma:rda-tilde-Sigma_k}
-Suppose that rank$(\widehat{\bm \Sigma}) = q \le p$.
-Let $\widehat{\bm \Sigma} = \bm U \bm D \bm U'$ be the eigendecomposition of $\widehat{\bm \Sigma}$ such that $\bm D \in \mathbb{R}_{p \times p}$ is the diagonal matrix of eigenvalues of $\widehat{\bm \Sigma}$ with
-\begin{align*}
-	\bm D = \begin{bmatrix}
-		\bm D_q & 0\\
-		\bm 0 & \bm 0
-	\end{bmatrix}
-\end{align*}
-and $\bm U \in \mathbb{R}_{p \times p}$ is the matrix whose columns are the corresponding eigenvectors of $\widehat{\bm \Sigma}$.
-Then, partitioning $\bm U = (\bm U_1, \bm U_2)$ so that $\bm U_1 \in \mathbb{R}_{p \times q}$ and $\bm U_2 \in \mathbb{R}_{p \times (p - q)}$, we have the equivalent expression\footnote{To generalize the shrinkage, we define the diagonal matrix of adjusted eigenvalues as $\bm \Phi$ and partition it into two parts: $\bm \Phi_1$ corresponding to the first $q$ eigenvalues of $\widehat{\bm \Sigma}$, and $\bm \Phi_2$ corresponding to the remaining $p - q$ eigenvalues of $\widehat{\bm \Sigma}$.}
+Let $\widehat{\bm \Sigma} = \bm U \bm D \bm U'$ be the eigendecomposition of $\widehat{\bm \Sigma}$ as above, and suppose that rank$(\widehat{\bm \Sigma}) = q \le p$. Then, we have
 \begin{align}
-	\tilde{\bm \Sigma}_k &= \bm U \begin{bmatrix}
-		\bm W_k & \bm 0\\
-		\bm 0 & \gamma \bm I_{p - q}
-	\end{bmatrix}\bm U',\label{eq:rda-matrix}
+	\tilde{\bm \Sigma}_k &= \bm U(\bm W_k \oplus \gamma \bm I_{p - q})\bm U',\label{eq:rda-matrix}
 \intertext{where}
-\bm W_k &= (1-\gamma) \{(1 - \lambda) \bm U_1' \widehat{\bm \Sigma}_k \bm U_1 + \lambda \bm D_q\} + \gamma \bm I_q.
+\bm W_k &= (1-\gamma) \{(1 - \lambda) \bm U_1' \widehat{\bm \Sigma}_k \bm U_1 + \lambda \bm D_q\} + \gamma \bm I_q.\label{eq:Wk}
 \end{align}
@@
 \end{align*}
-From Lemma \ref{lemma:null-spaces}, we have that $\widehat{\bm \Sigma}_k \bm U_2 = \bm 0$, which implies that 
-\begin{align*}
-	\bm U' \widehat{\bm \Sigma}_k \bm U = \begin{bmatrix}
-		\bm U_1' \widehat{\bm \Sigma}_k \bm U_1 & \bm 0\\
-		\bm 0 & \bm 0
-	\end{bmatrix},
-\end{align*}
-for $k = 1, \ldots, K$. Thus, recalling that $\bm U$ is orthogonal, we have that 
-\begin{align*}
-	\tilde{\bm \Sigma}_k = \bm U \begin{bmatrix}
-		\bm W_k & \bm 0\\
-		\bm 0 & \gamma \bm I_{p - q}
-	\end{bmatrix}\bm U'.
-\end{align*}
+From Lemma \ref{lemma:null-spaces}, we have that the columns of $\bm U_2$ span the null space of $\widehat{\bm \Sigma}$, which implies that $\widehat{\bm \Sigma}_k \bm U_2 = \bm 0$.
+Hence, for $k = 1, \ldots, K$, $\bm U' \widehat{\bm \Sigma}_k \bm U = \bm U_1' \widehat{\bm \Sigma}_k \bm U_1 \oplus \bm 0_{p-q}$.
+Thus, \eqref{eq:rda-matrix} follows because $\bm U$ is orthogonal.
 \end{proof}
 
-An immediate consequence of Lemma \ref{lemma:rda-tilde-Sigma_k} is that ranks of the \emph{RDA} covariance matrices for all classes are each equal to the rank of the pooled sample covariance matrix if the pooling parameter $\lambda$ is positive.
-Thus, by incorporating each $\bm x_i$ into the estimation of $\bm \Sigma_k$, we increase the rank.
+An immediate consequence of Lemma \ref{lemma:rda-tilde-Sigma_k} is rank$(\widehat{\bm \Sigma}_k(\lambda)) = q$, $k = 1, \ldots, K$ if the pooling parameter $\lambda \ne 0$.
+Thus, by incorporating each $\bm x_i$ into the estimation of $\bm \Sigma_k$, we increase the rank of $\widehat{\bm \Sigma}_k(\lambda)$ and, therefore, $\tilde{\bm \Sigma}_k$.
 We state this formally here.
@@
 
-The following result is essential to our main result.
-Here, we show that the difference of a vector $\bm x \in \mathbb{R}_{p \times 1}$ to each class sample mean is constant across all classes by showing that the difference is in the null space of $\bm U_2'$.
+Also from Lemma 2, we see that $\tilde{\bm \Sigma}_k$ has a similar form, even if the matrix function $\phi_k$ is unique.
+Next, we provide an essential result that will allow us to prove that for any unlabeled observation $\bm x \in \mathbb{R}_{p \times 1}$ that \eqref{eq:grda} is invariant to adjustments to the eigenvalues of $\tilde{\bm \Sigma}_k$ corresponding to the null space of $\widehat{\bm \Sigma}$.
 
@@
 \begin{proof}
-	First, we have that $\xbar_k - \xbar_{k'} = \sum_{i=1}^N \alpha_i \bm x_i$, where $\alpha_i = (n_k n_{k'})^{-1} \{ I(y_i = k) n_{k'} - I(y_i = k') n_k \}$.
-Let $\bm x \in \mathbb{R}_{p \times 1}$.
+Let $\bm x \in \mathbb{R}_{p \times 1}$, and suppose that $1 \le k, k' \le K$.
+Then, $\xbar_k - \xbar_{k'} = \sum_{i=1}^N \alpha_i \bm x_i$, where $\alpha_i = (n_k n_{k'})^{-1} \{ I(y_i = k) n_{k'} - I(y_i = k') n_k \}$.
 Then, $\bm U_2' (\bm x - \xbar_k) = \bm U_2' (\bm x - \xbar_{k'})$ is equivalent to $\bm U_2'(\xbar_k - \xbar_{k'}) = \sum_{i=1}^N \alpha_i \bm U_2' \bm x_i$.
@@
 Now, because $\bm x_i \in \mathcal{C}(\widehat{\bm \Sigma})$, for $i = 1, \ldots, N$, we have that $\bm U_2' \bm x_i = \bm 0$, which implies that $\bm U_2'(\xbar_k - \xbar_{k'}) = \bm 0$.
+Therefore, $\bm U_2' (\bm x - \xbar_k) = \bm U_2' (\bm x - \xbar_{k'})$.
 \end{proof}
 
-Here, we present our main result. We partition the \emph{RDA} decision rule into a sum.
+Lemma 2 is essential to our main result below because we decompose \eqref{eq:grda} into two components, one of which involves the constant term given in Lemma 3.
+The remaining components involve $\bm U_1$, a much smaller matrix than $\bm U_2$.
+Now, we present our main result, where we partition our generalized \emph{RDA} decision rule and show that the component requiring the largest computational costs does not contribute to the classification performed.
+Hence, we are able to omit the term and apply the reduced decision rule as follows.
 
 \begin{thm}
-	The \emph{RDA} decision rule in \eqref{eq:rda} is equivalent to
+Let $\tilde{\bm \Sigma}_k$ be defined as in \eqref{eq:rda-matrix} and $\bm W_k$ be defined as in \eqref{eq:Wk}.
+Also, let $\bm U_1$ be defined as above.
+Then, the generalized \emph{RDA} decision rule in \eqref{eq:grda} is equivalent to the reduced form
 	\begin{align}
-		D_{RDA}(\bm x) &= \argmin_k (\bm x - \xbar_k)' \bm U_1 \bm W_k^{-1} \bm U_1' (\bm x - \xbar_k)' + \ln |\bm Q_k| + (p - q) \ln \gamma + q \ln \{ (1 - \gamma) \lambda \} \label{eq:rda-decomposed}
+		D_{RDA}(\bm x) &= \argmin_k  (\bm x - \xbar_k)' \bm U_1 \bm W_k^{-1} \bm U_1' (\bm x - \xbar_k)' + \ln | \bm W_k | + (p - q) \ln \gamma. \label{eq:rda-decomposed}
 	\end{align}
@@
 \begin{proof}
-	From \eqref{eq:rda-matrix}, we have that
-	\begin{align*}
-		\tilde{\bm \Sigma}_k^{-1} = \bm U \begin{bmatrix}
-			\bm W_k^{-1} & \bm 0\\
-			\bm 0 & \gamma^{-1} \bm I_{p - q}
-		\end{bmatrix} \bm U'
-	\end{align*}
-	and $|\tilde{\bm \Sigma}_k| = \gamma^{p-q} | \bm W_k |$. Therefore, for all $\bm x \in \mathbb{R}_{p \times 1}$, we have that
+	From \eqref{eq:rda-matrix}, we have that $\tilde{\bm \Sigma}_k^{-1} = \bm U(\bm W_k^{-1} \oplus\gamma^{-1} \bm I_{p - q})\bm U'$ and $|\tilde{\bm \Sigma}_k| = \gamma^{p-q} | \bm W_k |$.
+Therefore, for all $\bm x \in \mathbb{R}_{p \times 1}$, we have that
 	\begin{align*}
@@
 	\end{align*}
-Applying Lemma \ref{lemma:RDA-constant-term}, $\bm U_2' (\bm x - \xbar_k)$ is constant for all $k = 1, \ldots, K$.
-Also, from Lemma \ref{lemma:rda-W_k}, we have that $\ln |\bm W_k| = \ln |\bm \Gamma_k| + \ln |\bm Q_k|$, and because \eqref{eq:rda-decomposed} is independent of $\ln |\bm D_q|$, the proof follows.
+From Lemma \ref{lemma:RDA-constant-term}, $\bm U_2' (\bm x - \xbar_k)$ is constant for all $k = 1, \ldots, K$, and the proof follows.
 \end{proof}
@@
 
-In our following result we show that the inverse and determinants of $\bm W_k$ can be written so that they are performed on $n_k \times n_k$ matrices rather than $q \times q$ matrices.
-Given that $n_k$ is often quite small, say, between 15 and 50, this can reduce the amount of computation substantially.
-We apply the well-known Woodbury formula and a related determinant lemma.
+Thus far we have seen that the generalized \emph{RDA} classifier in \eqref{eq:rda-decomposed} is invariant to the $\bm U_2$ term, yielding a substantial reduction in the computational costs.
+While the classifier is much improved in terms of computational performance compared to the \emph{RDA} classifier originally proposed by \cite{Friedman:1989tm}, we provide two additional results that further improve the computational performance of our generalized \emph{RDA} classifier.
+First, we show that the inverse and determinants of $\bm W_k$ can be written so that they are performed on $n_k \times n_k$ matrices rather than $q \times q$ matrices.
+Given that $n_k$ is often quite small, say, between 15 and 50, the following result facilitates a further reduction in the computational costs.
 
@@
 
-Note that we can write the pooled sample covariance matrix in \eqref{eq:pooled-cov} as
-\begin{align*}
-	\widehat{\bm\Sigma} = \bm X' \bm X,
-\end{align*}
-where we define $\bm X = [\sqrt{c_{1k}(1)} \bm x_1', \ldots,  \sqrt{c_{Nk}(1)} \bm x_N']'$.
-Now, we can apply the Fast SVD to compute the eigenvalue decomposition of $\widehat{\bm\Sigma}$ efficiently.
-By writing the SVD of $\bm X = \bm U \bm D \bm V'$, we have the spectral decomposition $\widehat{\bm\Sigma} = \bm X' \bm X = \bm V \bm D^2 \bm V'$, where $\bm U \in \mathbb{R}_{N \times N}$ is orthogonal, $\bm D \in \mathbb{R}_{N \times N}$ is a diagonal matrix consisting of the singular values of $\bm X$, and $\bm V \in \mathbb{R}_{p \times N}$ is orthogonal.
-Notice that $\bm X \bm X' = \bm U \bm D^2 \bm U'$, so that the eigenvalue decomposition of $\bm X \bm X'$ yields $\bm U$ and $\bm D$.
-This is much faster when $p \gg N$ because $\bm X \bm X' \in \mathbb{R}_{N \times N}$.
-Also, notice that rank$(\widehat{\bm \Sigma}) = q < N$ so that $\bm D^{-1}$ does not exist.
-After obtaining $\bm U$ and $\bm D$, we compute the matrix consisting of the first $N$ eigenvectors of $\widehat{\bm\Sigma}$ by simply computing\footnote{TODO: Update this paragraph to a proposition along with proof.}
-\begin{align*}
-	\bm V = \bm X' \bm U \bm D^{-1}.
-\end{align*}
-
-We can determine $q$ efficiently by computing the number of nonzero eigenvalues.
+Thus far, we have construction each $\tilde{\bm \Sigma}_k$ in terms of the first $q$ eigenvectors of $\widehat{\bm \Sigma}$.
+In practice, we would usually require a costly eigendecomposition of $\widehat{\bm \Sigma}$, but by utilizing a technique similar to the so-called kernel trick, we can obtain $\bm U_1$ by computing the eigendecomposition of a much smaller $N \times N$ matrix when $p \gg N$ \citep{Hastie:2008dt}.
+We begin our derivation by writing the pooled sample covariance matrix in \eqref{eq:pooled-cov} as $\widehat{\bm\Sigma} = \bm X' \bm X$, where we define $\bm X = [\sqrt{c_{1k}(1)} \bm x_1', \ldots,  \sqrt{c_{Nk}(1)} \bm x_N']'$.
+Then, by writing the SVD of $\bm X = \bm M \bm D \bm U'$, we have the spectral decomposition $\widehat{\bm\Sigma} = \bm U \bm D^2 \bm U'$, where $\bm M \in \mathbb{R}_{N \times N}$ is orthogonal, $\bm D \in \mathbb{R}_{N \times N}$ is a diagonal matrix consisting of the singular values of $\bm X$, and $\bm U \in \mathbb{R}_{p \times N}$ is orthogonal.
+Now, we see that $\bm X \bm X' = \bm M \bm D^2 \bm M'$ is the eigenvalue decomposition of $\bm X \bm X'$, where $\bm D^2$ is the matrix of eigenvalues of $\bm X \bm X'$ and the columns of $\bm M$  are the corresponding eigenvectors.
+Hence, after a much faster matrix multiplication, we can obtain $\bm M$ and $\bm D$ from a quick eigendecomposition of $\bm X \bm X' \in \mathbb{R}_{N \times N}$.
+Next, we compute $\bm U = \bm X' \bm M \bm D^{+}$, where $\bm D^{+} = \bm D_q^{-1} \oplus \bm 0_{N-q}$ is the Moore-Penrose pseudoinverse of $\bm D$.
+We can determine $q$ efficiently by computing the number of nonzero eigenvalues present in $\bm D^2$.
 However, in our experience, we must take care because the $N-q$ smallest eigenvalues are numerically nonzero, albeit they are near zero.
-Thus, we select $q$ to be the number of eigenvalues that exceeds some pre specified tolerance value, say, $10^{-6}$.
-
-As noted by \cite{Hoyle:2011vt}, \cite{Raudys:1998dd}, and \cite{Ramey:2011ji}, the pseudoinverse is often inadvisable because the resulting classifier often yields a sharp decrease in classification accuracy if $p \approx N$.\footnote{TODO: Mention that the care recommended above is related to the pseudoinverse and the issue with the smallest eigenvalues.}
-
-Cross-Validation\footnote{TODO: Use the Sherman-Morrison formula for cross-validation.}
+Thus, we select $q$ to be the number of eigenvalues that exceeds some given tolerance value, say, $1 \times 10^{-6}$.
+After calculating $q$, we need only retain the first $q$ columns of $\bm U$ to extract $\bm U_1$.
 
@@
 
-Our\footnote{TODO: Look at Ye's (2005) JMLR paper on his choice of the $\bm M$ matrix.
-He has a nice way of saying that he considers special cases.} generalization of the \emph{RDA} classifier facilitates a large number of potential models based on the specification of the matrix functions to shrink the eigenvalues.
+Our generalization of the \emph{RDA} classifier facilitates a large number of potential models based on the specification of the matrix functions to shrink the eigenvalues.
 In our simulations below, we consider two such eigenvalue value functions.
@@
 
-TODO\footnote{TODO: Add results discussion and table here.}
-
 \section{Discussion}
@@
 
+\section{Next Actions}
+
+\begin{itemize}
+\item Add paragraph of other notation, including null space, orthogonal complement, and direct sum.
+\item Generalize the shrinkage.
+Define the diagonal matrix of adjusted eigenvalues as $\bm \Phi$ and partition it into two parts: $\bm \Phi_1$ corresponding to the first $q$ eigenvalues of $\widehat{\bm \Sigma}$, and $\bm \Phi_2$ corresponding to the remaining $p - q$ eigenvalues of $\widehat{\bm \Sigma}$.
+\item Look at Ye's (2005) JMLR paper on his choice of the $\bm M$ matrix. He has a nice way of saying that he considers special cases
+\item Implement classifier
+\item Run simulations
+\item Add simulation results
+\item Discussion simulation results
+\item Use the Sherman-Morrison formula for cross-validation
+\item Discuss choice of $\lambda$ and eigenvalue shrinkage, if necessary
+\item Question to answer: What do we do when $\lambda = 0$, corresponding to \emph{QDA}?
+\item Show a timing comparison between our implementation and that of the {\tt klaR} package.
+\end{itemize}
+
 \bibliographystyle{plainnat}

### 2013-02-19 reports/paper-rda.tex
Removed any mention of matrix functions in the paper. Then, updated the formulation of the generalized RDA classifier to consider a family of eigenvalue-adjustment methods.

@@
 \begin{abstract}
-\cite{Friedman:1989tm} proposed the popular regularized discriminant analysis (\emph{RDA}) classifier for the small-sample, high-dimensional setting by employing a regularized covariance-matrix estimator, which incorporates a ridge-like shrinkage factor by adding a positive constant to the zero-eigenvalue estimates to ensure positive definiteness.
-We show that this shrinkage is unnecessary because shrinking the zero eigenvalues corresponding to the null space of the pooled sample covariance matrix has no effect on the \emph{RDA} classifier.
+\cite{Friedman:1989tm} has proposed the popular regularized discriminant analysis (\emph{RDA}) classifier for the small-sample, high-dimensional setting by employing a regularized covariance-matrix estimator, which incorporates a ridge-like shrinkage factor by adding a positive constant to the zero-eigenvalue estimates to ensure positive definiteness.
+Under mild regularity conditions we show that this shrinkage is unnecessary because shrinking the zero eigenvalues corresponding to the null space of the pooled sample covariance matrix has no effect on the \emph{RDA} classifier.
 Accordingly, we propose a novel generalization of the \emph{RDA} classifier that is more computationally efficient than its original counterpart because the inverses and determinants computed in the decision rule can be reduced from the original feature space to a linear subspace with dimensions corresponding to the individual class sample sizes.
 Additionally, our formulation of the \emph{RDA} classifier allows us to estimate efficiently optimal tuning parameters from a large set of candidate values.
-Furthermore, our generalization simplifies the construction of the covariance-matrix estimator to improve its interpretation and allows for a more flexible classification model by viewing shrinkage applied to the nonzero eigenvalue estimates in terms of matrix functions.
-Fortuitously, this matrix-function formulation includes as special cases the original \emph{RDA} classifier and several other well-known regularized classifiers.
+Furthermore, our generalization simplifies the construction of the covariance-matrix estimator to improve its interpretation and allows for a more flexible classification model.
+Our proposed formulation includes as special cases the original \emph{RDA} classifier and several other well-known regularized classifiers.
 Moreover, we demonstrate with several high-dimensional microarray data sets that our proposed classifier is competitive and often superior to several recently proposed sparse and regularized classifiers.
@@
 This simplification is advantageous in that it improves the interpretation of the pooling operation of the \emph{RDA} classifier.
-Next, we consider covariance-matrix regularization in terms of matrix functions, which includes a large family of shrinkage methods, such as the methods from \cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and a wide variety of other shrinkage methods studied by \cite{Ramey:2011ji} and \cite*{Xu:2009fl}.
-While a large number of these methods can be viewed as variants of ridge regression, we consider the more general case, where a real monotonic function is applied to the eigenvalue estimates.
-We show that adjusting the zero eigenvalues with a general matrix function has no effect on the \emph{RDA} classifier.
-That is, we show that the \emph{RDA} classifier is invariant to any adjustments of the eigenvalues corresponding to the null space of the pooled sample covariance matrix.
-Moreover, there is no classificatory information available in the null space of the pooled sample covariance matrix.
+Next, we propose a general formulation of covariance-matrix regularization, which includes a large family of shrinkage methods, such as the methods from \cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and a wide variety of other shrinkage methods studied by \cite{Ramey:2011ji} and \cite*{Xu:2009fl}.
+We show that under mild regularity conditions the \emph{RDA} classifier is invariant to any adjustments of the eigenvalues corresponding to the null space of the pooled sample covariance matrix.
 Hence, no gain in classification performance can be obtained by adjusting these zero eigenvalues.
@@
 
+We also require the following additional notation throughout the document.
+Let $\bm A \oplus \bm B$ denote the direct sum of $\bm A \in \mathbb{R}_{r \times r}$ and $\bm B \in \mathbb{R}_{s \times s}$ \citep[Chapter 1]{Lutkepohl:1996uz}.
+Let $\bm A^+$ and $\mathcal{N}(\bm A) = \{\bm z \in \mathbb{R}_{p \times 1} : \bm A \bm z = \bm 0\}$ denote the Moore-Penrose pseudoinverse and null space of $\bm A \in \mathbb{R}_{m \times n}$, respectively.
+Let $V^{\perp} = \{\bm y \in \mathbb{R}_{p \times 1} : \bm y'\bm v = 0\ \forall \bm v \in V \}$ denote the orthogonal complement of a vector space $V \subset \mathbb{R}_{p \times 1}$.
+Let $\bm I_m$ be the $m \times m$ identity matrix, and let $\bm 0_{m \times n}$ be the $m \times n$ matrix of zeros, where $\bm 0_m$ is understood to be $\bm 0_{m \times m}$.
+
 \subsection{Covariance Matrix Regularization}
@@
 Several regularization methods, such as the methods considered by \citet*{Guo:2007te}, \cite{Mkhadri:1995jp}, and \citet*{Xu:2009fl}, have been proposed in the literature to adjust the eigenvalues of $\widehat{\bm \Sigma}_k$ so that \eqref{eq:qda} is calculable and has reduced variability in the standard bias-variance tradeoff.
-These methods often can be written concisely as a matrix function, which we define here using the notation of \cite{Izenman:2008gm}.
-Let $\bm A \in \mathbb{R}_{p \times p}$ be symmetric, and let $\phi:\mathbb{R} \rightarrow \mathbb{R}$ be monotonic.
-Then
+A common form of the covariance-matrix regularization applies a shrinkage factor $\gamma_k > 0$, so that
 \begin{align}
-	\phi(\bm A) = \sum_{j = 1}^p \phi(e_j) \bm v_j \bm v_j'\label{eq:spectral-decomp}
-\end{align}
-is a matrix function defined explicitly in terms of the eigenvalues of $\bm A$.
-For this reason, $\phi$ is often called a spectral transform \citep*{Zhu:2010vu} or a transfer function \citep{Ye:2009gd}.
-Special cases of \eqref{eq:spectral-decomp} include the spectral decomposition of $\bm A$ when $\phi(e_j) = e_j$, the matrix square-root of $\bm A$ when $\phi(e_j) = e_j^{1/2}$, and, if it exists, the inverse of $\bm A$ when $\phi(e_j) = e_j^{-1}$ \citep{Harville:2008wja}.
-
-Notice that if  $\phi(e_{jk}) = e_{jk}$, $k = 1, \ldots, K$, no eigenvalue adjustment is performed, and \eqref{eq:qda} remains incalculable.
-Similarly, if $\phi(e_{jk}) = e_{j}$, $k = 1, \ldots, K$, the \emph{LDA} classifier is employed without eigenvalue shrinkage.
-Additionally, setting $\phi(e_j) = e_j^{+}$ with $z^{+} = 1/z$ if $z > 0$ and $0$ otherwise, we ignore the zero eigenvalues, so that $\phi(\widehat{\bm \Sigma}_k) = \widehat{\bm \Sigma}_k^{+}$, the Moore-Penrose pseudoinverse of $\widehat{\bm \Sigma}_k$.
-
-Another common form of the covariance-matrix regularization applies a shrinkage factor $\gamma_k > 0$, so that
-\begin{align}
-	\phi(e_{jk}) = e_{jk} + \gamma_k, \label{eq:ridge-estimator}
+	\widehat{\bm \Sigma}_k(\gamma_k) =  \widehat{\bm \Sigma}_k + \gamma_k, \label{eq:ridge-estimator}
 \end{align}
 similar to a method employed in ridge regression.
-Equation \eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance matrix $\widehat{\bm\Sigma}_k$ towards the $p$-dimensional identity matrix $\bm I_p$, thereby increasing the eigenvalues of $\widehat{\bm\Sigma}_k$ by $\gamma_k$.
+Equation \eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance matrix $\widehat{\bm\Sigma}_k$ towards $\bm I_p$, thereby increasing the eigenvalues of $\widehat{\bm\Sigma}_k$ by $\gamma_k$.
 Specifically, the zero eigenvalues are replaced with $\gamma_k$, resulting in $\widehat{\bm \Sigma}_k + \gamma_k \bm I_p$ being positive definite.
@@
 \end{align}
-for class $\omega_k$, where $\gamma \in [0, 1]$ is a regularization parameter that controls the shrinkage of \eqref{eq:sig-rda}  towards the $p$-dimensional identity matrix $\bm I_p$, which is weighted by the the average of the eigenvalues of \eqref{eq:sig-lambda}.
+for class $\omega_k$, where $\gamma \in [0, 1]$ is a regularization parameter that controls the shrinkage of \eqref{eq:sig-rda}  towards $\bm I_p$, which is weighted by the the average of the eigenvalues of \eqref{eq:sig-lambda}.
 Thus, the \emph{pooling} parameter $\lambda$ controls the amount that we borrow from $\widehat{\bm\Sigma}$ to estimate $\bm \Sigma_k$, and the \emph{shrinkage} parameter $\gamma$ determines the amount of shrinkage applied.
@@
 \end{align}
-Notice that we can write the shrinkage term in \eqref{eq:sig-rda} as the matrix function
-\begin{align*}
-	\phi(e_{jk}(\lambda)) = (1 - \gamma) e_{jk}(\lambda) + \gamma \bar{e}_k,
-\end{align*}
-for class $\omega_k$, where $e_{jk}(\lambda)$ is the $j$th largest eigenvalue of $\widehat{\bm\Sigma}_k(\lambda)$ and $\bar{e}_k = p^{-1} \sum_{j=1}^p e_{jk}(\lambda)$ is the average of the eigenvalues of \eqref{eq:sig-lambda}.
 
@@
 Here, we define our generalization of the \emph{RDA} classifier by first simplifying the formulation of $\widehat{\bm \Sigma}_k(\lambda)$ in \eqref{eq:sig-lambda} and demonstrating its clear interpretation as a linear combination of the crossproducts of the training observations centered by their respective class sample means.
-Then, we generalize the adjustment of the eigenvalues of $\widehat{\bm \Sigma}_k(\lambda)$ by introducing a more general matrix-function formulation.
+Then, we generalize the adjustment of the eigenvalues of $\widehat{\bm \Sigma}_k(\lambda)$.
 We define
@@
 Furthermore, if $\lambda = 1$, the weights in \eqref{eq:sig-lambda-alternative2} reduce to $1/N$, corresponding to $\widehat{\bm\Sigma}$.
+For brevity, for $\lambda = 1$, we define $\bm X = [\sqrt{c_{1k}(1)} \bm x_1', \ldots,  \sqrt{c_{Nk}(1)} \bm x_N']'$ such that $\widehat{\bm \Sigma} = \bm X' \bm X$.
+Similarly, for $\lambda = 0$, we define $\bm X_k = [\sqrt{c_{1k}(0)} \bm x_1', \ldots,  \sqrt{c_{Nk}(0)} \bm x_N']'$ such that $\widehat{\bm \Sigma}_k = \bm X_k' \bm X_k$, $k = 1, \ldots, K$.
 
 As we have discussed, several eigenvalue adjustment methods have been proposed to increase eigenvalues (approximately) equal to 0.
-In this section we define a more general eigenvalue adjustment of \eqref{eq:sig-lambda-alternative} by applying the matrix function $\phi_k(\widehat{\bm \Sigma}_k(\lambda))$.
-To simplify the notation, we write $\tilde{\bm \Sigma}_k = \phi_k(\widehat{\bm \Sigma}_k(\lambda))$.
+Here, we define a more general eigenvalue adjustment of \eqref{eq:sig-lambda-alternative} as
+\begin{align}
+	\tilde{\bm \Sigma}_k = \alpha_k \widehat{\bm \Sigma}_k(\lambda) + \bm \Phi_k,\label{eq:grda-cov}
+\end{align}
+where $\alpha_k \ge 0$ and $\bm \Phi_k \in \mathbb{R}_{p \times p}^{\ge}$ is a diagonal matrix consisting of the eigenvalue shrinkage constants.
 Thus, we have the generalized \emph{RDA} (\emph{GRDA}) classifier
@@
 \end{align}
-The specification of $\phi_k$ in \eqref{eq:grda} allows for a large family of eigenvalue shrinkage methods to be examined and for their relationships to be examined in depth.
+The specification of $\alpha_k$ and $\bm \Phi_k$ in \eqref{eq:grda-cov} allows for a large family of eigenvalue shrinkage methods to be examined and for their relationships to be examined in depth.
 
@@
 Our approach is to decompose \eqref{eq:grda} into a sum of two components, where the first summand consists of matrix operations applied to low-dimensional matrices and the second summand corresponds to the null space of $\widehat{\bm \Sigma}$ in \eqref{eq:pooled-cov}.
-We show that the matrix operations performed on the null space of $\widehat{\bm \Sigma}$ yield constant quadratic forms across all classes and can be omitted.
+We show that the matrix operations performed on the null space of $\widehat{\bm \Sigma}$ yield constant quadratic forms across all classes and can be omitted under mild regularity conditions.
 For $p \gg N$, the constant component involves determinants and inverses of high-dimensional matrices, and by ignoring these calculations, a substantial reduction in computational costs is achieved relative to the \emph{RDA} classifier originally proposed by \cite{Friedman:1989tm}.
@@
 In Lemma \ref{lemma:rda-tilde-Sigma_k} below, we derive an alternative expression for $\tilde{\bm \Sigma}_k$ in terms of the eigenvectors of $\widehat{\bm \Sigma}$.
-Let $\widehat{\bm \Sigma} = \bm U \bm D \bm U'$ be the eigendecomposition of $\widehat{\bm \Sigma}$ such that $\bm D \in \mathbb{R}_{p \times p}$ is the diagonal matrix of eigenvalues of $\widehat{\bm \Sigma}$ with $\bm D = \bm D_q \oplus \bm 0_{p-q}$, the columns of $\bm U \in \mathbb{R}_{p \times p}$ are the corresponding eigenvectors of $\widehat{\bm \Sigma}$, and rank$(\widehat{\bm \Sigma}) = q$, where $\bm A \oplus \bm B$ denotes the direct sum of $\bm A \in \mathbb{R}_{r \times r}$ and $\bm B \in \mathbb{R}_{s \times s}$ \citep[Chapter 1]{Lutkepohl:1996uz}.
-Then, partitioning $\bm U = (\bm U_1, \bm U_2)$ such that $\bm U_1 \in \mathbb{R}_{p \times q}$ and $\bm U_2 \in \mathbb{R}_{p \times (p - q)}$, we see that the component of $\tilde{\bm \Sigma}_k$ corresponding to the null space of $\widehat{\bm \Sigma}$ is in terms of the matrix function $\phi_k$ alone.
+Let $\widehat{\bm \Sigma} = \bm U \bm D \bm U'$ be the eigendecomposition of $\widehat{\bm \Sigma}$ such that $\bm D \in \mathbb{R}_{p \times p}$ is the diagonal matrix of eigenvalues of $\widehat{\bm \Sigma}$ with $\bm D = \bm D_q \oplus \bm 0_{p-q}$, $\bm D_q \in \mathbb{R}_{q \times q}$ is the diagonal matrix consisting of the positive eigenvalues of $\widehat{\bm \Sigma}$,  the columns of $\bm U \in \mathbb{R}_{p \times p}$ are the corresponding eigenvectors of $\widehat{\bm \Sigma}$, and rank$(\widehat{\bm \Sigma}) = q$.
+Then, we partition $\bm U = (\bm U_1, \bm U_2)$ such that $\bm U_1 \in \mathbb{R}_{p \times q}$ and $\bm U_2 \in \mathbb{R}_{p \times (p - q)}$.
+We also partition $\bm \Phi_k = \bm \Phi_{k1} \oplus \bm \Phi_{k2}$ such that $\bm \Phi_{k1} \in \mathbb{R}_{q \times q}$ and $\bm \Phi_{k2} \in \mathbb{R}_{(p-q) \times (p-q)}$.
 
@@
 \begin{align}
-	\tilde{\bm \Sigma}_k &= \bm U(\bm W_k \oplus \gamma \bm I_{p - q})\bm U',\label{eq:rda-matrix}
+	\tilde{\bm \Sigma}_k &= \bm U(\bm W_k \oplus \bm \Phi_{k2})\bm U',\label{eq:rda-matrix}
 \intertext{where}
-\bm W_k &= (1-\gamma) \{(1 - \lambda) \bm U_1' \widehat{\bm \Sigma}_k \bm U_1 + \lambda \bm D_q\} + \gamma \bm I_q.\label{eq:Wk}
+\bm W_k &= \alpha_k \{(1 - \lambda) \bm U_1' \widehat{\bm \Sigma}_k \bm U_1 + \lambda \bm D_q\} + \bm \Phi_{k1}.\label{eq:Wk}
 \end{align}
@@
 \begin{align*}
-	\bm U' \tilde{\bm \Sigma}_k \bm U = (1-\gamma) \{(1 - \lambda) \bm U_1' \widehat{\bm \Sigma}_k \bm U_1 + \lambda \bm D_q\} + \gamma \bm U' \bm U.
+	\bm U' \tilde{\bm \Sigma}_k \bm U = \alpha_k \{(1 - \lambda) \bm U' \widehat{\bm \Sigma}_k \bm U + \lambda \bm D_q\} + \bm U' \bm \Phi_k \bm U.
 \end{align*}
-From Lemma \ref{lemma:null-spaces}, we have that the columns of $\bm U_2$ span the null space of $\widehat{\bm \Sigma}$, which implies that $\widehat{\bm \Sigma}_k \bm U_2 = \bm 0$.
+From Lemma \ref{lemma:null-spaces}, the columns of $\bm U_2$ span the null space of $\widehat{\bm \Sigma}$, which implies that $\widehat{\bm \Sigma}_k \bm U_2 = \bm 0_{p \times (p - q)}$.
 Hence, for $k = 1, \ldots, K$, $\bm U' \widehat{\bm \Sigma}_k \bm U = \bm U_1' \widehat{\bm \Sigma}_k \bm U_1 \oplus \bm 0_{p-q}$.
@@
 An immediate consequence of Lemma \ref{lemma:rda-tilde-Sigma_k} is rank$(\widehat{\bm \Sigma}_k(\lambda)) = q$, $k = 1, \ldots, K$ if the pooling parameter $\lambda \ne 0$.
-Thus, by incorporating each $\bm x_i$ into the estimation of $\bm \Sigma_k$, we increase the rank of $\widehat{\bm \Sigma}_k(\lambda)$ and, therefore, $\tilde{\bm \Sigma}_k$.
+Thus, by incorporating each $\bm x_i$ into the estimation of $\bm \Sigma_k$, we increase the rank of $\widehat{\bm \Sigma}_k(\lambda)$.
 We state this formally here.
@@
 \begin{proof}
-The proof follows by setting $\gamma = 0$ in Lemma \ref{lemma:rda-tilde-Sigma_k}.
+The proof follows by setting $\bm \Phi_k = \bm 0_p$ in \eqref{eq:Wk}.
 \end{proof}
 
-Also from Lemma 2, we see that $\tilde{\bm \Sigma}_k$ has a similar form, even if the matrix function $\phi_k$ is unique.
-Next, we provide an essential result that will allow us to prove that for any unlabeled observation $\bm x \in \mathbb{R}_{p \times 1}$ that \eqref{eq:grda} is invariant to adjustments to the eigenvalues of $\tilde{\bm \Sigma}_k$ corresponding to the null space of $\widehat{\bm \Sigma}$.
+Next, we provide an essential result that will allow us to prove that for any unlabeled observation $\bm x \in \mathbb{R}_{p \times 1}$, \eqref{eq:grda} is invariant to adjustments to the eigenvalues of $\tilde{\bm \Sigma}_k$ corresponding to the null space of $\widehat{\bm \Sigma}$.
 
@@
 Let $\bm x \in \mathbb{R}_{p \times 1}$, and suppose that $1 \le k, k' \le K$.
-Then, $\xbar_k - \xbar_{k'} = \sum_{i=1}^N \alpha_i \bm x_i$, where $\alpha_i = (n_k n_{k'})^{-1} \{ I(y_i = k) n_{k'} - I(y_i = k') n_k \}$.
-Then, $\bm U_2' (\bm x - \xbar_k) = \bm U_2' (\bm x - \xbar_{k'})$ is equivalent to $\bm U_2'(\xbar_k - \xbar_{k'}) = \sum_{i=1}^N \alpha_i \bm U_2' \bm x_i$.
+Then, $\xbar_k - \xbar_{k'} = \sum_{i=1}^N \beta_i \bm x_i$, where $\beta_i = (n_k n_{k'})^{-1} \{ I(y_i = k) n_{k'} - I(y_i = k') n_k \}$.
+Then, $\bm U_2' (\bm x - \xbar_k) = \bm U_2' (\bm x - \xbar_{k'})$ is equivalent to $\bm U_2'(\xbar_k - \xbar_{k'}) = \sum_{i=1}^N \beta_i \bm U_2' \bm x_i$.
 Recall that $\bm U_2 \in \mathcal{N}(\widehat{\bm \Sigma})$, which implies that $\bm U_2' \in \mathcal{C}(\widehat{\bm \Sigma})^{\perp}$ \citep[Lemma 1.2.5]{Kollo:2005vp}.
-Now, because $\bm x_i \in \mathcal{C}(\widehat{\bm \Sigma})$, for $i = 1, \ldots, N$, we have that $\bm U_2' \bm x_i = \bm 0$, which implies that $\bm U_2'(\xbar_k - \xbar_{k'}) = \bm 0$.
+Now, because $\bm x_i \in \mathcal{C}(\widehat{\bm \Sigma})$, for $i = 1, \ldots, N$, $\bm U_2' \bm x_i = \bm 0_{(p-q) \times 1}$, which implies that $\bm U_2'(\xbar_k - \xbar_{k'}) = \bm 0_{(p-q) \times 1}$.
 Therefore, $\bm U_2' (\bm x - \xbar_k) = \bm U_2' (\bm x - \xbar_{k'})$.
@@
 Now, we present our main result, where we partition our generalized \emph{RDA} decision rule and show that the component requiring the largest computational costs does not contribute to the classification performed.
+In order for this term to be constant to reduce the computational costs of the generalized \emph{RDA} classifier, we require that $\bm \Phi_{k2} = \bm \Phi_2$, $k = 1, \ldots, K$.
 Hence, we are able to omit the term and apply the reduced decision rule as follows.
@@
 Also, let $\bm U_1$ be defined as above.
+Additionally, let $\bm \Phi_{k2} = \bm \Phi_2$, $k = 1, \ldots, K$.
 Then, the generalized \emph{RDA} decision rule in \eqref{eq:grda} is equivalent to the reduced form
 	\begin{align}
-		D_{RDA}(\bm x) &= \argmin_k  (\bm x - \xbar_k)' \bm U_1 \bm W_k^{-1} \bm U_1' (\bm x - \xbar_k)' + \ln | \bm W_k | + (p - q) \ln \gamma. \label{eq:rda-decomposed}
+		D_{GRDA}(\bm x) &= \argmin_k  (\bm x - \xbar_k)' \bm U_1 \bm W_k^{-1} \bm U_1' (\bm x - \xbar_k) + \ln | \bm W_k |. \label{eq:rda-decomposed}
 	\end{align}
@@
 \begin{proof}
-	From \eqref{eq:rda-matrix}, we have that $\tilde{\bm \Sigma}_k^{-1} = \bm U(\bm W_k^{-1} \oplus\gamma^{-1} \bm I_{p - q})\bm U'$ and $|\tilde{\bm \Sigma}_k| = \gamma^{p-q} | \bm W_k |$.
+	From \eqref{eq:rda-matrix}, we have that $\tilde{\bm \Sigma}_k^{-1} = \bm U(\bm W_k^{-1} \oplus\bm \Phi_{k2}^{-1})\bm U'$ and $|\tilde{\bm \Sigma}_k| = | \bm W_k | |\bm \Phi_{k2}|$.
 Therefore, for all $\bm x \in \mathbb{R}_{p \times 1}$, we have that
 	\begin{align*}
-	(\bm x - \xbar_k)' \tilde{\bm\Sigma}_k^{-1}(\bm x - \xbar_k)  + \ln |\tilde{\bm\Sigma}_k| &= (\bm x - \xbar_k)' \bm U_1 \bm W_k^{-1} \bm U_1' (\bm x - \xbar_k)'\\
-	&+ \gamma^{-1} (\bm x - \xbar_k)' \bm U_2 \bm U_2' (\bm x - \xbar_k) + \ln | \bm W_k | + (p - q) \ln \gamma.
+	(\bm x - \xbar_k)' \tilde{\bm\Sigma}_k^{-1}(\bm x - \xbar_k)  + \ln |\tilde{\bm\Sigma}_k| &= (\bm x - \xbar_k)' \bm U_1 \bm W_k^{-1} \bm U_1' (\bm x - \xbar_k)\\
+	&+ (\bm x - \xbar_k)' \bm U_2 \bm \Phi_{k2}^{-1} \bm U_2' (\bm x - \xbar_k) + \ln | \bm W_k | + \ln |\bm \Phi_{k2}|.
 	\end{align*}
-From Lemma \ref{lemma:RDA-constant-term}, $\bm U_2' (\bm x - \xbar_k)$ is constant for all $k = 1, \ldots, K$, and the proof follows.
+From Lemma \ref{lemma:RDA-constant-term}, $\bm U_2' (\bm x - \xbar_k)$ is constant for all $k = 1, \ldots, K$. Therefore, the proof follows because $\bm \Phi_{k2} = \bm \Phi_2$.
 \end{proof}
@@
 	\begin{align}
-		\bm W_k^{-1} &= \bm \Gamma_k^{-1} - (1 - \gamma)(1 - \lambda) \bm \Gamma_k^{-1} \bm U_1' \bm X_k' \bm Q_k^{-1} \bm X_k \bm U_1 \bm \Gamma_k^{-1},
+		\bm W_k^{-1} &= \bm \Gamma_k^{-1} - \alpha_k(1 - \lambda) \bm \Gamma_k^{-1} \bm U_1' \bm X_k' \bm Q_k^{-1} \bm X_k \bm U_1 \bm \Gamma_k^{-1},
 	\end{align}
-where $\bm Q_k = \bm I_{n_k} + (1 - \gamma)(1- \lambda) \bm X_k \bm U_1 \bm \Gamma_k^{-1} \bm U_1' \bm X_k'$ and $\bm \Gamma_k = (1 - \gamma) \lambda \bm D_q + \gamma \bm I_q$.
+where $\bm Q_k = \bm I_{n_k} + \alpha_k(1- \lambda) \bm X_k \bm U_1 \bm \Gamma_k^{-1} \bm U_1' \bm X_k'$ and $\bm \Gamma_k = \alpha_k \lambda \bm D_q + \bm \Phi_{k1}$.
 \end{proposition}
 \begin{proof}
-	Set $\bm A = \bm \Gamma_k$, $\bm B = (1 - \gamma)(1 - \lambda) \bm U_1' \bm X_k'$, $\bm T = \bm I_{n_k}$, and $\bm C = \bm X_k \bm U_1$.
+	Set $\bm A = \bm \Gamma_k$, $\bm B = \alpha_k (1 - \lambda) \bm U_1' \bm X_k'$, $\bm T = \bm I_{n_k}$, and $\bm C = \bm X_k \bm U_1$.
 To calculate $|\bm W_k|$, from \cite{Harville:2008wja} we apply Theorem 18.1.1, which states that $|\bm A + \bm B \bm T \bm C| = |\bm A| |\bm T| |\bm T^{-1} + \bm C \bm A^{-1} \bm B|$.
 Thus, we have $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$.
-Also from Theorem 18.2.8 in \cite{Harville:2008wja}, we apply the Woodbury formula $(\bm A + \bm B \bm C)^{-1} = \bm A^{-1} - \bm A^{-1} \bm B(\bm I + \bm C \bm A^{-1} \bm B)^{-1} \bm C \bm A^{-1}$, which implies that $\bm W_k^{-1} = \bm \Gamma_k^{-1} - (1 - \gamma)(1 - \lambda) \bm \Gamma_k^{-1} \bm U_1' \bm X_k' \bm Q_k^{-1} \bm X_k \bm U_1 \bm \Gamma_k^{-1}$.
+Also from Theorem 18.2.8 in \cite{Harville:2008wja}, we apply the Woodbury formula $(\bm A + \bm B \bm C)^{-1} = \bm A^{-1} - \bm A^{-1} \bm B(\bm I + \bm C \bm A^{-1} \bm B)^{-1} \bm C \bm A^{-1}$, and the proof follows.
 \end{proof}
@@
 In practice, we would usually require a costly eigendecomposition of $\widehat{\bm \Sigma}$, but by utilizing a technique similar to the so-called kernel trick, we can obtain $\bm U_1$ by computing the eigendecomposition of a much smaller $N \times N$ matrix when $p \gg N$ \citep{Hastie:2008dt}.
-We begin our derivation by writing the pooled sample covariance matrix in \eqref{eq:pooled-cov} as $\widehat{\bm\Sigma} = \bm X' \bm X$, where we define $\bm X = [\sqrt{c_{1k}(1)} \bm x_1', \ldots,  \sqrt{c_{Nk}(1)} \bm x_N']'$.
-Then, by writing the SVD of $\bm X = \bm M \bm D \bm U'$, we have the spectral decomposition $\widehat{\bm\Sigma} = \bm U \bm D^2 \bm U'$, where $\bm M \in \mathbb{R}_{N \times N}$ is orthogonal, $\bm D \in \mathbb{R}_{N \times N}$ is a diagonal matrix consisting of the singular values of $\bm X$, and $\bm U \in \mathbb{R}_{p \times N}$ is orthogonal.
-Now, we see that $\bm X \bm X' = \bm M \bm D^2 \bm M'$ is the eigenvalue decomposition of $\bm X \bm X'$, where $\bm D^2$ is the matrix of eigenvalues of $\bm X \bm X'$ and the columns of $\bm M$  are the corresponding eigenvectors.
-Hence, after a much faster matrix multiplication, we can obtain $\bm M$ and $\bm D$ from a quick eigendecomposition of $\bm X \bm X' \in \mathbb{R}_{N \times N}$.
-Next, we compute $\bm U = \bm X' \bm M \bm D^{+}$, where $\bm D^{+} = \bm D_q^{-1} \oplus \bm 0_{N-q}$ is the Moore-Penrose pseudoinverse of $\bm D$.
-We can determine $q$ efficiently by computing the number of nonzero eigenvalues present in $\bm D^2$.
+Recall that the pooled sample covariance matrix $\widehat{\bm\Sigma} = \bm X' \bm X$.
+Then, by writing the SVD of $\bm X = \bm M \bm D^{1/2} \bm U'$, we have the spectral decomposition $\widehat{\bm\Sigma} = \bm U \bm D \bm U'$, where $\bm M \in \mathbb{R}_{N \times N}$ is orthogonal, $\bm D^{1/2} \in \mathbb{R}_{N \times N}$ is a diagonal matrix consisting of the singular values of $\bm X$, and $\bm U \in \mathbb{R}_{p \times N}$ is orthogonal.
+Now, we see that $\bm X \bm X' = \bm M \bm D \bm M'$ is the eigenvalue decomposition of $\bm X \bm X'$, where $\bm D$ is the matrix of eigenvalues of $\bm X \bm X'$ and the columns of $\bm M$  are the corresponding eigenvectors.
+Hence, we obtain $\bm M$ and $\bm D$ from an eigendecomposition of $\bm X \bm X' \in \mathbb{R}_{N \times N}$.
+Next, we compute $\bm U = \bm X' \bm M \bm D^{+/2}$, where $\bm D^{+/2} = \bm D_q^{-1/2} \oplus \bm 0_{N-q}$.
+We can determine $q$ efficiently by computing the number of nonzero eigenvalues present in $\bm D$.
 However, in our experience, we must take care because the $N-q$ smallest eigenvalues are numerically nonzero, albeit they are near zero.
@@
 
-Our generalization of the \emph{RDA} classifier facilitates a large number of potential models based on the specification of the matrix functions to shrink the eigenvalues.
-In our simulations below, we consider two such eigenvalue value functions.
+Our generalization of the \emph{RDA} classifier facilitates a large number of potential models based on the specification of $\alpha_k$ and $\bm \Phi_k$ in \eqref{eq:grda-cov} to shrink the eigenvalues of $\widehat{\bm \Sigma}_k(\lambda)$.
+In our simulations below, we consider two choices of $\alpha_k$ and $\bm \Phi_k$ based on variants of eigenvalue adjustment methods proposed in the literature.
 First, ...
-Second, we consider the matrix function from \cite{Srivastava:2007ww}, who derive an empirical-Bayes covariance-matrix estimator under the assumption that the data are independent and identically distributed multivariate normal observations.
-\cite{Srivastava:2007ww} then substitute the estimator into the \eqref{eq:lda} to attain the minimum distance empirical Bayes (\emph{MDEB}) classifier.
-To derive the empirical-Bayes covariance-matrix estimator, \cite{Srivastava:2007ww} first assume that $\bm \Sigma_k^{-1}$ follows a Wishart distribution \emph{a priori} with mean $\lambda^{-1} \bm I_p$, $\lambda > 0$, and degrees of freedom, $l \ge p$.
-The resulting posterior estimator for $\bm \Sigma_k^{-1}$ is $\frac{n_k + l}{n_k}(n_k^{-1}\lambda \bm I_p + \widehat{\bm \Sigma}_k)^{-1}$.
-Using an empirical Bayes argument to estimate $\lambda$ for $p > n_k$, \cite{Srivastava:2007ww} propose an empirical-Bayes covariance-matrix estimator with matrix function
-\begin{align}
-	\phi(e_{jk}) = e_{jk} + n_k^{-1} \sum_{j = 1}^p e_{jk},\label{eq:cov-mdeb}
-\end{align}
-where the shrinkage factor in \eqref{eq:cov-mdeb} is approximately equal to the average of the nonzero eigenvalues of $\widehat{\bm \Sigma}_k$.
-We note that \cite{Srivastava:2007ww} set assumed that $\bm \Sigma_k = \bm \Sigma$ so that $\phi(e_{jk}) = \phi(e_k)$, $k = 1, \ldots, K$.
 
-\cite{Srivastava:2007ww} claim that the \emph{MDEB} classifier is the best regularization method they have encountered in the literature.
-When $p \le n$, the shrinkage factors of the \emph{RDA} and \emph{MDEB} classifiers are equal, but for $p > n$ this shrinkage can differ substantially.
-In fact, \cite{Ramey:2011ji} show empirically that the \emph{MDEB} classifier can yield superior classification accuracy to the \emph{RDA} classifier.
+Second, \cite{Srivastava:2007ww} derive an empirical-Bayes covariance-matrix estimator under the assumption that the data are independent and identically distributed multivariate normal observations and then substitute the estimator into the \emph{LDA} classifier to attain the minimum distance empirical Bayes (\emph{MDEB}) classifier.
+To derive the empirical-Bayes covariance-matrix estimator, \cite{Srivastava:2007ww} first assume that $\bm \Sigma^{-1}$ follows a Wishart distribution \emph{a priori} with mean $\lambda^{-1} \bm I_p$, $\lambda > 0$, and degrees of freedom, $l \ge p$.
+The resulting posterior estimator for $\bm \Sigma^{-1}$ is $\frac{N + l}{N}(N^{-1}\lambda \bm I_p + \widehat{\bm \Sigma})^{-1}$.
+Using an empirical Bayes argument to estimate $\lambda$, \cite{Srivastava:2007ww} propose the covariance-matrix estimator
+\begin{align*}
+	\widehat{\bm \Sigma}_{MDEB} = \widehat{\bm \Sigma} + \frac{\tr \{\widehat{\bm \Sigma}\}}{\min\{N, p\}} \bm I_p,
+\end{align*}
+where the shrinkage factor is approximately equal to the average of the nonzero eigenvalues of $\widehat{\bm \Sigma}$.
+We consider a variant of the \emph{MDEB} covariance-matrix estimator because \cite{Srivastava:2007ww} claim that the \emph{MDEB} classifier is the best covariance-matrix regularization method applied to the \emph{LDA} classifier they have encountered in the literature.
 
@@
 \cite{Nakayama:2007fl} determined from their data that these 10 types fell into 4 broader groups: (1) SS; (2) MLS; (3) Lipoma, WDLS, and part of DDLS; (4) Spindle cell and pleomorophic sarcomas including DDLS, MFS, LMS, MPNST, FS, and MFH.
-Following \cite{Witten:2011kc}, we restrict our analysis to the five tumor types having at least 15 samples observed.
+Following \cite{Witten:2011kc}, we restrict our analysis to the five tumor types having at least 15 observations.
 
@@
 
-Question to answer: What do we do when $\lambda = 0$, corresponding to \emph{QDA}?
-
-It is interesting to note that our approach is similar to an aside made by Friedman (1989) in the case that the RDA covariance matrix estimator is singular.
-Friedman (1989) recommended that the zero eigenvalues be replaced with a small value such as 0.001.
-However, in Theorem 1, we saw that even small adjustments can be inappropriate because they result in ineffective shrinkage.
-
 \section{Next Actions}
@@
 \begin{itemize}
-\item Add paragraph of other notation, including null space, orthogonal complement, and direct sum.
-\item Generalize the shrinkage.
-Define the diagonal matrix of adjusted eigenvalues as $\bm \Phi$ and partition it into two parts: $\bm \Phi_1$ corresponding to the first $q$ eigenvalues of $\widehat{\bm \Sigma}$, and $\bm \Phi_2$ corresponding to the remaining $p - q$ eigenvalues of $\widehat{\bm \Sigma}$.
-\item Look at Ye's (2005) JMLR paper on his choice of the $\bm M$ matrix. He has a nice way of saying that he considers special cases
 \item Implement classifier
@@
 \item Discussion simulation results
-\item Use the Sherman-Morrison formula for cross-validation
+\item Show a timing comparison between our implementation and that of the {\tt klaR} package.
 \item Discuss choice of $\lambda$ and eigenvalue shrinkage, if necessary
 \item Question to answer: What do we do when $\lambda = 0$, corresponding to \emph{QDA}?
-\item Show a timing comparison between our implementation and that of the {\tt klaR} package.
 \end{itemize}

### 2013-03-21 lib/helpers.R
Modified the simulations to speed them up and make them run successfully.

@@ -169,5 +169,9 @@ Clemmensen <- function(train_x, train_y, test_x, normalize_data = FALSE, cv_vari
 Cao <- function(train_x, train_y, test_x) {
   plsda_out <- plsda(X = train_x, Y = train_y)
   plsda_predict <- predict(plsda_out, test_x, method = "max.dist")
-  factor(plsda_predict$class$max.dist[, 2], levels = c(1, 2), labels = levels(train_y))
+  factor(plsda_predict$class$max.dist[, 2], levels = seq_len(nlevels(train_y)),
+         labels = levels(train_y))
 }
+
+
+

### 2013-03-21 sim-microarray.r

@@ -14,7 +14,9 @@ results <- mclapply(data_sets, function(data_set) {
 
   cv_folds <- cv_partition(y = data$y, num_folds = 10)
 
-  cv_results <- lapply(cv_folds, function(cv_fold) {
+  cv_results <- lapply(seq_along(cv_folds), function(i) {
+    message("Dataset: ", data_set, " -- Fold ", i)
+    cv_fold <- cv_folds[[i]]
     train_x <- as.matrix(data$x[cv_fold$training, ])
     train_y <- data$y[cv_fold$training]
     test_x <- as.matrix(data$x[cv_fold$test, ])
@@ -24,18 +26,21 @@ results <- mclapply(data_sets, function(data_set) {
     num_classes <- nlevels(train_y)
     prior_probs <- rep(1, num_classes) / num_classes
 
+    # For now, I've commented out the Technometrics method because it's too slow.
+    # As in, it took about 12 hours for a single fold in my 10-fold CV simulations
+    # of several data sets.
     # Clemmensen, Hastie, Witten and Ersbøll (2012) - Technometrics
-    Clemmensen_out <- Clemmensen(train_x, train_y, test_x, cv_variables = TRUE,
-                                 normalize_data = TRUE)
-    Clemmensen_errors <- sum(Clemmensen_out != test_y)
+    # Clemmensen_out <- Clemmensen(train_x, train_y, test_x, cv_variables = TRUE,
+    #                             normalize_data = TRUE)
+    # Clemmensen_errors <- sum(Clemmensen_out != test_y)
 
     # Witten and Tibshirani (2011) - JRSS B
     Witten_out <- Witten_Tibshirani(train_x, train_y, test_x)
     Witten_errors <- sum(Witten_out$predictions != test_y)
 
     # Applies the variable selection from Witten and Tibshirani (2011)
-    train_x <- train_x[, Witten_out$variables]
-    test_x <- test_x[, Witten_out$variables]
+    # train_x <- train_x[, Witten_out$variables]
+    # test_x <- test_x[, Witten_out$variables]
 
     # GRDA
     cv_out <- grda_cv(x = train_x, y = train_y, prior = prior_probs)
@@ -55,7 +60,6 @@ results <- mclapply(data_sets, function(data_set) {
 
     list(
       Cao = Cao_errors,
-      Clemmensen = Clemmensen_errors,
       dlda = dlda_errors,
       dqda = dqda_errors,
       grda = grda_errors,

### 2013-03-21 lib/helpers.R
Added Guo et al. (2007) classifier to simulation.

@@ -174,4 +174,35 @@ Cao <- function(train_x, train_y, test_x) {
 }
 
 
+# Classifier from Guo, Hastie, and Tibshirani (2007) - Biostatistics
+scrda_train <- function(x, y) {
+  x <- t(x)
+  rda_out <- rda(x = x, y = y)
+  rda_cv_out <- rda.cv(rda_out, x = x, y = y)
+
+  # Which alpha and delta give min cv error?
+  min_cv_error <- with(rda_cv_out, which(cv.err == min(cv.err), arr.ind = TRUE))
+
+  # Out of the minima, what are the gene dimensions?
+  num_genes <- rda_cv_out$ngene[min_cv_error]
+
+  # Which alpha-delta has smallest dimension? If there's still a tie, choose one
+  # at random.
+  min_alpha_delta <- sample(which(num_genes == min(num_genes)), size = 1)
+
+  # Now we determine the corresponding alpha-delta pair.
+  alpha_delta <- min_cv_error[min_alpha_delta,]
+  alpha <- as.numeric(rownames(rda_cv_out$cv.err)[alpha_delta[1]])
+  delta <- as.numeric(colnames(rda_cv_out$cv.err)[alpha_delta[2]])
+
+  list(rda_out = rda_out, rda_cv_out = rda_cv_out, alpha = alpha, delta = delta)
+}
+
+# Classifier from Guo, Hastie, and Tibshirani (2007) - Biostatistics
+scrda_predict <- function(rda_out, train_x, train_y, test_x, alpha, delta) {
+  group_names <- levels(train_y)
+  factor(
+    predict(rda_out, x = t(train_x), y = train_y, xnew = t(test_x), alpha = alpha, delta = delta),
+    levels = seq_along(group_names), labels = group_names)
+}
 

### 2013-03-21 sim-microarray.r

@@ -42,6 +42,12 @@ results <- mclapply(data_sets, function(data_set) {
     # train_x <- train_x[, Witten_out$variables]
     # test_x <- test_x[, Witten_out$variables]
 
+    # Classifier from Guo, Hastie, and Tibshirani (2007) - Biostatistics
+    Guo_out <- scrda_train(x = train_x, y = train_y)
+    Guo_pred <- with(Guo_out, scrda_predict(rda_out, train_x, train_y, test_x,
+                                            alpha, delta))
+    Guo_errors <- sum(Guo_pred != test_y)
+
     # GRDA
     cv_out <- grda_cv(x = train_x, y = train_y, prior = prior_probs)
     grda_out <- with(cv_out, grda(x = train_x, y = train_y, lambda = lambda, gamma = gamma, prior = prior_probs))
@@ -63,6 +69,7 @@ results <- mclapply(data_sets, function(data_set) {
       dlda = dlda_errors,
       dqda = dqda_errors,
       grda = grda_errors,
+      Guo = Guo_errors,
       Tong = Tong_errors,
       Witten = Witten_errors
       )

### 2013-03-21 sim-microarray.r
Added Pang et al. (2009) SDLDA and SDQDA classifiers to simulation.

@@ -42,7 +42,7 @@ results <- mclapply(data_sets, function(data_set) {
     # train_x <- train_x[, Witten_out$variables]
     # test_x <- test_x[, Witten_out$variables]
 
-    # Classifier from Guo, Hastie, and Tibshirani (2007) - Biostatistics
+    # Guo, Hastie, and Tibshirani (2007) - Biostatistics
     Guo_out <- scrda_train(x = train_x, y = train_y)
     Guo_pred <- with(Guo_out, scrda_predict(rda_out, train_x, train_y, test_x,
                                             alpha, delta))
@@ -57,7 +57,11 @@ results <- mclapply(data_sets, function(data_set) {
     dlda_errors <- sum(predict(dlda(x = train_x, y = train_y, prior = prior_probs), test_x)$class != test_y)
     dqda_errors <- sum(predict(dqda(x = train_x, y = train_y, prior = prior_probs), test_x)$class != test_y)
 
-    # Tong et al. (2012) - Bioinformatics
+    # Pang, Tong, and Zhao (2009) - Biometrics
+    sdlda_errors <- sum(predict(sdlda(x = train_x, y = train_y, prior = prior_probs), test_x)$class != test_y)
+    sdqda_errors <- sum(predict(sdqda(x = train_x, y = train_y, prior = prior_probs), test_x)$class != test_y)
+
+    # Tong, Chen, and Zhao (2012) - Bioinformatics
     Tong_out <- dlda(x = train_x, y = train_y, est_mean = "tong", prior = prior_probs)
     Tong_errors <- sum(predict(Tong_out, test_x)$class != test_y)
 
@@ -70,6 +74,8 @@ results <- mclapply(data_sets, function(data_set) {
       dqda = dqda_errors,
       grda = grda_errors,
       Guo = Guo_errors,
+      sdlda = sdlda_errors,
+      sdqda = sdqda_errors,
       Tong = Tong_errors,
       Witten = Witten_errors
       )

### 2013-03-22 sim-microarray.r
Results in for classifiers considered before Guo's and Pang's methods were added. Updated the data sets to remove a problem data set. Also, added error handling to ensure simulations don't stop from errors.

@@ -4,13 +4,17 @@ load.project()
 set.seed(42)
 num_cores <- 12
 
+# Problem Data Sets
+# burczynski
+# chiaretti
 data_sets <- c("alon", "burczynski", "chiaretti", "christensen", "gordon",
-               "gravier", "khan", "nakayama", "shipp", "singh", "sorlie", "su",
-               "yeoh")
+               "gravier", "khan", "nakayama", "shipp", "singh", "sorlie", "yeoh")
 
 results <- mclapply(data_sets, function(data_set) {
   data <- load_microarray(data_set)
   data$x <- as.matrix(data$x)
+
+  set.seed(42)
 
   cv_folds <- cv_partition(y = data$y, num_folds = 10)
 
@@ -35,38 +39,34 @@ results <- mclapply(data_sets, function(data_set) {
     # Clemmensen_errors <- sum(Clemmensen_out != test_y)
 
     # Witten and Tibshirani (2011) - JRSS B
-    Witten_out <- Witten_Tibshirani(train_x, train_y, test_x)
+    Witten_out <- try_default(Witten_Tibshirani(train_x, train_y, test_x), NA)
     Witten_errors <- sum(Witten_out$predictions != test_y)
 
-    # Applies the variable selection from Witten and Tibshirani (2011)
-    # train_x <- train_x[, Witten_out$variables]
-    # test_x <- test_x[, Witten_out$variables]
-
     # Guo, Hastie, and Tibshirani (2007) - Biostatistics
-    Guo_out <- scrda_train(x = train_x, y = train_y)
-    Guo_pred <- with(Guo_out, scrda_predict(rda_out, train_x, train_y, test_x,
-                                            alpha, delta))
+    Guo_out <- try_default(scrda_train(x = train_x, y = train_y), NA)
+    Guo_pred <- try_default(with(Guo_out, scrda_predict(rda_out, train_x, train_y, test_x,
+                                            alpha, delta)), NA)
     Guo_errors <- sum(Guo_pred != test_y)
 
     # GRDA
-    cv_out <- grda_cv(x = train_x, y = train_y, prior = prior_probs)
-    grda_out <- with(cv_out, grda(x = train_x, y = train_y, lambda = lambda, gamma = gamma, prior = prior_probs))
-    grda_errors <- sum(predict(grda_out, test_x)$class != test_y)
[…]

### 2013-03-23 sim-microarray.r
Removed Clemmensen's method from the microarray simulations because it is too slow even for a moderate number of variables.

@@ -26,21 +26,9 @@ results <- mclapply(data_sets, function(data_set) {
     num_classes <- nlevels(train_y)
     prior_probs <- rep(1, num_classes) / num_classes
 
-    # For now, I've commented out the Technometrics method because it's too slow.
-    # As in, it took about 12 hours for a single fold in my 10-fold CV simulations
-    # of several data sets.
-    # Clemmensen, Hastie, Witten and Ersbøll (2012) - Technometrics
-    # Clemmensen_out <- Clemmensen(train_x, train_y, test_x, cv_variables = TRUE,
-    #                             normalize_data = TRUE)
-    # Clemmensen_errors <- sum(Clemmensen_out != test_y)
-
     # Witten and Tibshirani (2011) - JRSS B
     Witten_out <- Witten_Tibshirani(train_x, train_y, test_x)
     Witten_errors <- sum(Witten_out$predictions != test_y)
-
-    # Applies the variable selection from Witten and Tibshirani (2011)
-    # train_x <- train_x[, Witten_out$variables]
-    # test_x <- test_x[, Witten_out$variables]
 
     # Guo, Hastie, and Tibshirani (2007) - Biostatistics
     Guo_out <- scrda_train(x = train_x, y = train_y)

### 2013-03-23 sim-microarray.r
Updated simulations to a repeated split with various numbers of variables selected by Dudoit's method.

@@ -4,27 +4,36 @@ load.project()
 set.seed(42)
 num_cores <- 12
 
-# Problem Data Sets
-# burczynski
-# chiaretti
+train_pct <- 2/3
+num_iterations <- 50
 data_sets <- c("alon", "burczynski", "chiaretti", "christensen", "gordon",
                "gravier", "khan", "nakayama", "shipp", "singh", "sorlie", "yeoh")
+q <- c(100, 250, 500, 750, 1000)
 
-results <- mclapply(data_sets, function(data_set) {
+sim_config <- expand.grid(data_sets = data_sets, q = q, stringsAsFactors = FALSE)
+
+results <- mclapply(seq_len(nrow(sim_config)), function(i) {
+  data_set <- sim_config$data_set[i]
+  q <- sim_config$q[i]
+  
   data <- load_microarray(data_set)
   data$x <- as.matrix(data$x)
 
-  set.seed(42)
+  rand_splits <- rand_partition(y = data$y, num_partitions = num_iterations,
+                                train_pct = train_pct)
 
-  cv_folds <- cv_partition(y = data$y, num_folds = 10)
+  split_results <- lapply(seq_along(rand_splits), function(split) {
+    message("Dataset: ", data_set, " -- Split ", split)
+    rand_split <- rand_splits[[split]]
+    train_x <- as.matrix(data$x[rand_split$training, ])
+    train_y <- data$y[rand_split$training]
+    test_x <- as.matrix(data$x[rand_split$test, ])
+    test_y <- data$y[rand_split$test]
 
-  cv_results <- lapply(seq_along(cv_folds), function(i) {
-    message("Dataset: ", data_set, " -- Fold ", i)
-    cv_fold <- cv_folds[[i]]
-    train_x <- as.matrix(data$x[cv_fold$training, ])
-    train_y <- data$y[cv_fold$training]
-    test_x <- as.matrix(data$x[cv_fold$test, ])
-    test_y <- data$y[cv_fold$test]
+    # Apply Dudoit variable selection
+    dudoit_out <- dudoit(train_x, train_y, test_x, q = q)
+    train_x <- dudoit_out$train_x
+    test_x <- dudoit_out$test_x
 
     # Sets the prior probabilities equal
     num_classes <- nlevels(train_y)
@@ -32,33 +41,33 @@ results <- mclapply(data_sets, function(data_set) {
 
     # Witten and Tibshirani (2011) - JRSS B
[…]

### 2013-03-23 sim-microarray.r
Updated path of saved results.

@@ -89,5 +89,5 @@ results <- mclapply(seq_len(nrow(sim_config)), function(i) {
 }, mc.cores = num_cores)
 results <- do.call(rbind, results)
 
-save(results, file = 'results-microarray.RData')
+save(results, file = 'data/results-microarray.RData')
 

### 2013-03-24 reports/paper-rda.tex

@@
 
-\title{Another Look at Regularized Discriminant Analysis for High-Dimensional Classification}
+\title{A Generalization of Regularized Discriminant Analysis for High-Dimensional Classification}
 
@@
 \begin{abstract}
-\cite{Friedman:1989tm} has proposed the popular regularized discriminant analysis (\emph{RDA}) classifier for the small-sample, high-dimensional setting by employing a regularized covariance-matrix estimator, which incorporates a ridge-like shrinkage factor by adding a positive constant to the zero-eigenvalue estimates to ensure positive definiteness.
-Under mild regularity conditions we show that this shrinkage is unnecessary because shrinking the zero eigenvalues corresponding to the null space of the pooled sample covariance matrix has no effect on the \emph{RDA} classifier.
-Accordingly, we propose a novel generalization of the \emph{RDA} classifier that is more computationally efficient than its original counterpart because the inverses and determinants computed in the decision rule can be reduced from the original feature space to a linear subspace with dimensions corresponding to the individual class sample sizes.
-Additionally, our formulation of the \emph{RDA} classifier allows us to estimate efficiently optimal tuning parameters from a large set of candidate values.
-Furthermore, our generalization simplifies the construction of the covariance-matrix estimator to improve its interpretation and allows for a more flexible classification model.
-Our proposed formulation includes as special cases the original \emph{RDA} classifier and several other well-known regularized classifiers.
+\cite{Friedman:1989tm} has proposed the popular regularized discriminant analysis (\emph{RDA}) classifier for the small-sample, high-dimensional setting.
+This \emph{RDA} classifier employs a regularized covariance-matrix estimator, which incorporates a ridge-like shrinkage factor by adding a positive constant to the zero-eigenvalue estimates to ensure positive definiteness.
+Here, we propose a novel generalization of Friedman's \emph{RDA} classifier that is more computationally efficient than its original counterpart because the inverses and determinants computed in the decision rule can be reduced from the original feature space to a linear subspace with dimensions corresponding to the individual class sample sizes.
+Additionally, our generalization of the \emph{RDA} (\emph{GRDA}) classifier enables efficient estimation of optimal tuning parameters from a large set of candidate values, simplifies the construction of the regularized covariance-matrix estimator to improve its interpretation, and allows for a more flexible classification model.
 Moreover, we demonstrate with several high-dimensional microarray data sets that our proposed classifier is competitive and often superior to several recently proposed sparse and regularized classifiers.
-Finally, we provide implementations of the competing classifiers in the {\tt regdiscrim} R package, available on CRAN.
+Finally, we provide an implementation of the \emph{GRDA} classifier in the {\tt regdiscrim} R package, available on CRAN.
 \end{abstract}
@@
 
-Here, we consider specifically the popular regularized discriminant analysis (\emph{RDA}) classifier from \cite{Friedman:1989tm}, who has proposed a weighted average of the covariance-matrix estimators from the \emph{LDA} and \emph{QDA} classifiers.
+Here, we consider a generalization of the popular \emph{RDA} classifier of \cite{Friedman:1989tm}, who has proposed a weighted average of the covariance-matrix estimators from the \emph{LDA} and \emph{QDA} classifiers.
 Because the resulting biased covariance-matrix estimator can remain singular or nearly singular, \cite{Friedman:1989tm} has additionally incorporated applied covariance-matrix shrinkage, where the shrinkage constant is a function of the average eigenvalue of the original estimator.
@@
 
-Our improvements to the calculation and model selection of the \emph{RDA} classifier yields a natural generalization of the \emph{RDA} classifier.
-In our proposed generalization we first simplify the pooling matrix in the \emph{RDA} classifier by constructing instead a convex combination of the covariance-matrix estimators employed in the \emph{LDA} and \emph{QDA} classifiers.
+Also, for our \emph{GRDA} classifier we first simplify the pooling matrix in the \emph{RDA} classifier by constructing instead a convex combination of the covariance-matrix estimators employed in the \emph{LDA} and \emph{QDA} classifiers.
 This simplification is advantageous in that it improves the interpretation of the pooling operation of the \emph{RDA} classifier.
-Next, we propose a general formulation of covariance-matrix regularization, which includes a large family of shrinkage methods, such as the methods from \cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and a wide variety of other shrinkage methods studied by \cite{Ramey:2011ji} and \cite*{Xu:2009fl}.
-We show that under mild regularity conditions the \emph{RDA} classifier is invariant to any adjustments of the eigenvalues corresponding to the null space of the pooled sample covariance matrix.
+We show that under mild regularity conditions the \emph{GRDA} classifier is invariant to any adjustments of the eigenvalues corresponding to the null space of the pooled sample covariance matrix.
 Hence, no gain in classification performance can be obtained by adjusting these zero eigenvalues.
-Although \cite{Ji:2008wp} and \cite{Ye:2006jm} have provided a similar invariance property for the \emph{LDA} decision rule, the \emph{LDA} classifier involves the assumption that the population covariance matrices are equal, which is often too stringent \citep{Clemmensen:2011kr}.
-The \emph{RDA} classifier allows us to relax this assumption by including unrestricted covariance matrices in the model.
-The \emph{RDA} classifier is then a compromise between the two approaches, favoring the stronger linear assumption if a larger value of the pooling parameter is selected as optimal.
-Additionally, in the linear case  \cite{Ye:2009gd} have remarked that  a lack of a systematic study has been conduced to explore the relationships among various shrinkage methods.
+Although \cite{Ji:2008wp} and \cite{Ye:2006jm} have provided a similar invariance property for the \emph{LDA} decision rule, the \emph{LDA} classifier involves the assumption that the population covariance matrices are equal, which is often too stringent in practice \citep{Clemmensen:2011kr}.
+The \emph{GRDA} classifier allows us to relax this assumption by including unrestricted covariance matrices in the model.
+The \emph{GRDA} classifier is then a compromise between the two approaches, favoring the stronger linear assumption if a larger value of the pooling parameter is selected as optimal.
 Our formulation provides a general framework from which we can elucidate properties of covariance-regularization methods and the relationships among them in the linear case, the quadratic case, and the cases in between.
 
-In addition to the improvement in computational performance of classifiers with regularized covariance matrices, we demonstrate that our generalized \emph{RDA} classifier is competitive and often superior to several recent classifiers designed for small-sample, high-dimensional data.
-In particular, we compare our generalized RDA classifier with an alternatively-defined \emph{RDA} classifier from \cite*{Hastie:2008dt}, the regularized \emph{LDA} classifier from \cite{Guo:2007te}, a recently proposed variant of the \emph{DLDA} classifier from \cite*{Tong:2012hw}, and the penalized \emph{LDA} classifier from \cite{Witten:2011kc}.
+Many of the recently proposed classifiers have a variable-selection method integrated.
+While this is advantageous in potentially determining biomarkers that are correlated with the classification label, these methods often yield poor classification accuracy if a large number of variables are retained.
+Moreover, these methods can require large computational costs, making these methods undesirable in practice.
+To study these issues in depth, we consider several simulated data sets, where we show that the classification degrades as the number of features is increased.
+Contrarily, our proposed \emph{GRDA} classifier handles the large number of features well because our included dimension-reduction method retains only the classificatory information in the covariance-matrix estimators.
+Moreover, while the competing classifiers exhibit large computational costs as the number of features in the original data set is increased, the \emph{GRDA} classifier handles the larger number of features efficiently.
+
+In addition to the improvement in computational performance of classifiers with regularized covariance matrices, we demonstrate that our \emph{GRDA} classifier is competitive with and often superior to several recent classifiers designed for small-sample, high-dimensional data.
+In particular, we compare our \emph{GRDA} classifier with an alternatively-defined \emph{RDA} classifier from \cite*{Hastie:2008dt}, the regularized \emph{LDA} classifier from \cite{Guo:2007te}, a recently proposed variant of the \emph{DLDA} classifier from \cite*{Tong:2012hw}, and the penalized \emph{LDA} classifier from \cite{Witten:2011kc}.
 We conduct our comparison of the classification accuracy of the competing classifiers using three small-sample, high-dimensional microarray data sets and examine the classification accuracy of the competing classifiers using \textbf{three} simulation configurations.
@@
 Let $\bm A \oplus \bm B$ denote the direct sum of $\bm A \in \mathbb{R}_{r \times r}$ and $\bm B \in \mathbb{R}_{s \times s}$ \citep[Chapter 1]{Lutkepohl:1996uz}.
-Let $\bm A^+$ and $\mathcal{N}(\bm A) = \{\bm z \in \mathbb{R}_{p \times 1} : \bm A \bm z = \bm 0\}$ denote the Moore-Penrose pseudoinverse and null space of $\bm A \in \mathbb{R}_{m \times n}$, respectively.
+Let $\bm A^+$ and $\mathcal{N}(\bm A) = \{\bm z \in \mathbb{R}_{p \times 1} : \bm A \bm z = \bm 0\}$ denote the Moore-Penrose pseudoinverse and null space of $\bm A \in \mathbb{R}_{m \times p}$, respectively.
 Let $V^{\perp} = \{\bm y \in \mathbb{R}_{p \times 1} : \bm y'\bm v = 0\ \forall \bm v \in V \}$ denote the orthogonal complement of a vector space $V \subset \mathbb{R}_{p \times 1}$.
-Let $\bm I_m$ be the $m \times m$ identity matrix, and let $\bm 0_{m \times n}$ be the $m \times n$ matrix of zeros, where $\bm 0_m$ is understood to be $\bm 0_{m \times m}$.
+Let $\bm I_m$ be the $m \times m$ identity matrix, and let $\bm 0_{m \times p}$ be the $m \times p$ matrix of zeros, such that $\bm 0_m$ is understood to denote $\bm 0_{m \times m}$.
 
@@
 In fact, the eigenvalues of $\widehat{\bm \Sigma}_k$ are well-known to be biased if $p \ge n_k$ such that the smallest eigenvalues are underestimated \citep{Seber:2004uh}, Moreover, if $p > n_k$, then rank$(\widehat{\bm \Sigma}_k) \le n_k$, which implies that at least $p - n_k$ eigenvalues of $\widehat{\bm \Sigma}_k$ are 0.
-Hence, the bias of the eigenvalue estimators increases as $p$ increases relative to $n_k$.
 Furthermore, although more feature information is available to discriminate among the $K$ classes, if $p > n_k$, \eqref{eq:qda} is incalculable because $\widehat{\bm \Sigma}_k^{-1}$ does not exist.
@@
 \begin{align}
-	\widehat{\bm \Sigma}_k(\gamma_k) =  \widehat{\bm \Sigma}_k + \gamma_k, \label{eq:ridge-estimator}
+	\widehat{\bm \Sigma}_k(\gamma_k) =  \widehat{\bm \Sigma}_k + \gamma_k \bm I_p, \label{eq:ridge-estimator}
 \end{align}
-similar to a method employed in ridge regression.
+similar to a method employed in ridge regression \citep{Hoerl:1970cd}.
 Equation \eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance matrix $\widehat{\bm\Sigma}_k$ towards $\bm I_p$, thereby increasing the eigenvalues of $\widehat{\bm\Sigma}_k$ by $\gamma_k$.
-Specifically, the zero eigenvalues are replaced with $\gamma_k$, resulting in $\widehat{\bm \Sigma}_k + \gamma_k \bm I_p$ being positive definite.
+Specifically, the zero eigenvalues are replaced with $\gamma_k$, so that \eqref{eq:ridge-estimator} is positive definite.
 For additional covariance-matrix regularization methods, see \cite{Ramey:2011ji}, \cite{Xu:2009fl}, and \cite{Ye:2009gd}.
 
-\subsection{Original Definition of Regularized Discriminant Analysis}
+\subsection{Friedman's Regularized Discriminant Analysis Classifier}
 
@@
 Here, we define our generalization of the \emph{RDA} classifier by first simplifying the formulation of $\widehat{\bm \Sigma}_k(\lambda)$ in \eqref{eq:sig-lambda} and demonstrating its clear interpretation as a linear combination of the crossproducts of the training observations centered by their respective class sample means.
-Then, we generalize the adjustment of the eigenvalues of $\widehat{\bm \Sigma}_k(\lambda)$.
 We define
@@
 as a convex combination of the MLEs of $\bm \Sigma_k$ and $\bm \Sigma$, where $\lambda \in [0, 1]$ is the \emph{pooling} parameter.
-Thus, we no longer have need for the convex combination of $n_k$ and $N$ as defined by \cite{Friedman:1989tm}.
+Thus, we avoid the convex combination of $n_k$ and $N$ as defined by \cite{Friedman:1989tm}.
 By rewriting \eqref{eq:sig-lambda-alternative} in terms of the observations $\bm x_i$, $i = 1, \ldots, N$, each centered by its class sample mean, we attain a clear interpretation of $\widehat{\bm\Sigma}_k(\lambda)$.
@@
 That is, for $0 < \lambda \le 1$, we estimate $\bm \Sigma_k$ with $N$ rather than $n_k$ observations.
-Below, we will show that the pooling operation is advantageous in increasing the rank of each $\widehat{\bm\Sigma}_k(\lambda)$ from rank$(\widehat{\bm\Sigma}_k)$ to rank$(\widehat{\bm\Sigma}_k)$ for $0 < \lambda \le 1$.
+Below, we will show that the pooling operation is advantageous in increasing the rank of each $\widehat{\bm\Sigma}_k(\lambda)$ from rank$(\widehat{\bm\Sigma}_k)$ to rank$(\widehat{\bm\Sigma})$ for $0 < \lambda \le 1$.
 Notice that if $\lambda  = 0$, then the observations from the remaining $K - 1$ classes do not contribute to the estimation of $\bm \Sigma_k$, corresponding to $\widehat{\bm \Sigma}_k$.
@@
 \end{align}
-where $\alpha_k \ge 0$ and $\bm \Phi_k \in \mathbb{R}_{p \times p}^{\ge}$ is a diagonal matrix consisting of the eigenvalue shrinkage constants.
-Thus, we have the generalized \emph{RDA} (\emph{GRDA}) classifier
+where $\alpha_k \ge 0$ and $\bm \Phi_k \in \mathbb{R}_{p \times p}^{\ge}$ is a diagonal matrix consisting of the eigenvalue shrinkage constants. Thus, we have the \emph{GRDA} (\emph{GRDA}) classifier
 \begin{align}
@@
 \end{align}
-The specification of $\alpha_k$ and $\bm \Phi_k$ in \eqref{eq:grda-cov} allows for a large family of eigenvalue shrinkage methods to be examined and for their relationships to be examined in depth.
+The specification of $\alpha_k$ and $\bm \Phi_k$ in \eqref{eq:grda-cov} allows for a large family of eigenvalue-shrinkage methods and their relationships to be examined in depth.
 
-\section{The Generalized RDA Classifier}
-In this section, we demonstrate that our proposed generalization of the \emph{RDA} classifier facilitates an efficient calculation of \eqref{eq:grda} and a rapid model selection of the pooling parameter $\lambda$.
+\section{The GRDA Classifier}
+In this section, we demonstrate that the \emph{GRDA} classifier facilitates an efficient calculation of \eqref{eq:grda} and a rapid model selection of the pooling parameter $\lambda$.
 Our approach is to decompose \eqref{eq:grda} into a sum of two components, where the first summand consists of matrix operations applied to low-dimensional matrices and the second summand corresponds to the null space of $\widehat{\bm \Sigma}$ in \eqref{eq:pooled-cov}.
 We show that the matrix operations performed on the null space of $\widehat{\bm \Sigma}$ yield constant quadratic forms across all classes and can be omitted under mild regularity conditions.
-For $p \gg N$, the constant component involves determinants and inverses of high-dimensional matrices, and by ignoring these calculations, a substantial reduction in computational costs is achieved relative to the \emph{RDA} classifier originally proposed by \cite{Friedman:1989tm}.
+For $p \gg N$, the constant component involves determinants and inverses of high-dimensional matrices, and by ignoring these calculations, a substantial reduction in computational costs is achieved relative to Friedman's \emph{RDA} classifier in \eqref{eq:rda}.
 Furthermore, a byproduct of this omission is that adjustments to the associated eigenvalues have no effect on \eqref{eq:grda}.
@@
 Now, suppose $\bm z \in \mathcal{N}(\widehat{\bm \Sigma})$.
-Similarly, we have that $0 = \bm z' \widehat{\bm \Sigma} \bm z = N^{-1} \sum_{k = 1}^K n_k \bm z' \widehat{\bm \Sigma}_k \bm z$, which implies that $z \in \mathcal{N}(\widehat{\bm \Sigma}_k)$ because $\widehat{\bm \Sigma}_k \in \mathbb{R}_{p \times p}^{\ge}$ for $k = 1, \ldots, K$.
+Similarly, we have that $0 = \bm z' \widehat{\bm \Sigma} \bm z = N^{-1} \sum_{k = 1}^K n_k \bm z' \widehat{\bm \Sigma}_k \bm z$, which implies that $\bm z \in \mathcal{N}(\widehat{\bm \Sigma}_k)$ because $\widehat{\bm \Sigma}_k \in \mathbb{R}_{p \times p}^{\ge}$ for $k = 1, \ldots, K$.
 Therefore, $\mathcal{N}(\widehat{\bm \Sigma}) \subset \mathcal{N}(\widehat{\bm \Sigma}_k)$.
@@
 
-In Lemma \ref{lemma:rda-tilde-Sigma_k} below, we derive an alternative expression for $\tilde{\bm \Sigma}_k$ in terms of the eigenvectors of $\widehat{\bm \Sigma}$.
-Let $\widehat{\bm \Sigma} = \bm U \bm D \bm U'$ be the eigendecomposition of $\widehat{\bm \Sigma}$ such that $\bm D \in \mathbb{R}_{p \times p}$ is the diagonal matrix of eigenvalues of $\widehat{\bm \Sigma}$ with $\bm D = \bm D_q \oplus \bm 0_{p-q}$, $\bm D_q \in \mathbb{R}_{q \times q}$ is the diagonal matrix consisting of the positive eigenvalues of $\widehat{\bm \Sigma}$,  the columns of $\bm U \in \mathbb{R}_{p \times p}$ are the corresponding eigenvectors of $\widehat{\bm \Sigma}$, and rank$(\widehat{\bm \Sigma}) = q$.
+In Lemma \ref{lemma:rda-tilde-Sigma_k} below, we derive an alternative expression for $\tilde{\bm \Sigma}_k$ in terms of the matrix of eigenvectors of $\widehat{\bm \Sigma}$.
+Let $\widehat{\bm \Sigma} = \bm U \bm D \bm U'$ be the eigendecomposition of $\widehat{\bm \Sigma}$ such that $\bm D \in \mathbb{R}_{p \times p}$ is the diagonal matrix of eigenvalues of $\widehat{\bm \Sigma}$ with $\bm D = \bm D_q \oplus \bm 0_{p-q}$, $\bm D_q \in \mathbb{R}_{q \times q}^{>}$ is the diagonal matrix consisting of the positive eigenvalues of $\widehat{\bm \Sigma}$,  the columns of $\bm U \in \mathbb{R}_{p \times p}$ are the corresponding orthonormal eigenvectors of $\widehat{\bm \Sigma}$, and rank$(\widehat{\bm \Sigma}) = q$.
 Then, we partition $\bm U = (\bm U_1, \bm U_2)$ such that $\bm U_1 \in \mathbb{R}_{p \times q}$ and $\bm U_2 \in \mathbb{R}_{p \times (p - q)}$.
@@
 \begin{proof}
-First, notice that
-\begin{align*}
-	\bm U' \tilde{\bm \Sigma}_k \bm U = \alpha_k \{(1 - \lambda) \bm U' \widehat{\bm \Sigma}_k \bm U + \lambda \bm D_q\} + \bm U' \bm \Phi_k \bm U.
-\end{align*}
 From Lemma \ref{lemma:null-spaces}, the columns of $\bm U_2$ span the null space of $\widehat{\bm \Sigma}$, which implies that $\widehat{\bm \Sigma}_k \bm U_2 = \bm 0_{p \times (p - q)}$.
 Hence, for $k = 1, \ldots, K$, $\bm U' \widehat{\bm \Sigma}_k \bm U = \bm U_1' \widehat{\bm \Sigma}_k \bm U_1 \oplus \bm 0_{p-q}$.
+Then, notice that $\bm U' \tilde{\bm \Sigma}_k \bm U = \alpha_k \{(1 - \lambda) \bm U' \widehat{\bm \Sigma}_k \bm U + \lambda \bm D\} + \bm U' \bm \Phi_k \bm U$.
 Thus, \eqref{eq:rda-matrix} follows because $\bm U$ is orthogonal.
@@
 
-An immediate consequence of Lemma \ref{lemma:rda-tilde-Sigma_k} is rank$(\widehat{\bm \Sigma}_k(\lambda)) = q$, $k = 1, \ldots, K$ if the pooling parameter $\lambda \ne 0$.
-Thus, by incorporating each $\bm x_i$ into the estimation of $\bm \Sigma_k$, we increase the rank of $\widehat{\bm \Sigma}_k(\lambda)$.
-We state this formally here.
+As an immediate immediate consequence of Lemma \ref{lemma:rda-tilde-Sigma_k}, we have the following corollary.
 
@@
 \begin{proof}
-The proof follows by setting $\bm \Phi_k = \bm 0_p$ in \eqref{eq:Wk}.
+The proof follows by setting $\bm \Phi_k = \bm 0_p$ in Lemma \ref{lemma:rda-tilde-Sigma_k}.
 \end{proof}
 
-Next, we provide an essential result that will allow us to prove that for any unlabeled observation $\bm x \in \mathbb{R}_{p \times 1}$, \eqref{eq:grda} is invariant to adjustments to the eigenvalues of $\tilde{\bm \Sigma}_k$ corresponding to the null space of $\widehat{\bm \Sigma}$.
+Thus, by incorporating each $\bm x_i$ into the estimation of $\bm \Sigma_k$, we increase the rank of $\widehat{\bm \Sigma}_k(\lambda)$ to $q$, $k = 1, \ldots, K$ if the pooling parameter $\lambda \ne 0$.
+Next, we provide an essential result that enables us to prove that, under mild regularity conditions, \eqref{eq:grda} is invariant to adjustments to the eigenvalues of $\tilde{\bm \Sigma}_k$ corresponding to the null space of $\widehat{\bm \Sigma}$ for any unlabeled observation $\bm x \in \mathbb{R}_{p \times 1}$.
 
@@
 
-Lemma 2 is essential to our main result below because we decompose \eqref{eq:grda} into two components, one of which involves the constant term given in Lemma 3.
-The remaining components involve $\bm U_1$, a much smaller matrix than $\bm U_2$.
-Now, we present our main result, where we partition our generalized \emph{RDA} decision rule and show that the component requiring the largest computational costs does not contribute to the classification performed.
-In order for this term to be constant to reduce the computational costs of the generalized \emph{RDA} classifier, we require that $\bm \Phi_{k2} = \bm \Phi_2$, $k = 1, \ldots, K$.
-Hence, we are able to omit the term and apply the reduced decision rule as follows.
+Now, we present our main result, where we decompose the \emph{GRDA} decision rule and show that the component requiring the largest computational costs does not contribute to the classification performed using Lemma \ref{lemma:RDA-constant-term}.
+To reduce the computational costs of the \emph{GRDA} classifier, we require that $\bm \Phi_{k2} = \bm \Phi_2$, $k = 1, \ldots, K$.
+Here, we present our main result that reduces \eqref{eq:grda} to an equivalent, more computationally efficient decision rule.
 
 \begin{thm}
-Let $\tilde{\bm \Sigma}_k$ be defined as in \eqref{eq:rda-matrix} and $\bm W_k$ be defined as in \eqref{eq:Wk}.
-Also, let $\bm U_1$ be defined as above.
+Let $\tilde{\bm \Sigma}_k$ and $\bm W_k$ be defined as in \eqref{eq:rda-matrix} and \eqref{eq:Wk}, respectively, and let $\bm U_1$ be defined as above.
 Additionally, let $\bm \Phi_{k2} = \bm \Phi_2$, $k = 1, \ldots, K$.
-Then, the generalized \emph{RDA} decision rule in \eqref{eq:grda} is equivalent to the reduced form
+Then, the \emph{GRDA} decision rule in \eqref{eq:grda} is equivalent to the reduced form
 	\begin{align}
@@
 
-Thus far we have seen that the generalized \emph{RDA} classifier in \eqref{eq:rda-decomposed} is invariant to the $\bm U_2$ term, yielding a substantial reduction in the computational costs.
-While the classifier is much improved in terms of computational performance compared to the \emph{RDA} classifier originally proposed by \cite{Friedman:1989tm}, we provide two additional results that further improve the computational performance of our generalized \emph{RDA} classifier.
-First, we show that the inverse and determinants of $\bm W_k$ can be written so that they are performed on $n_k \times n_k$ matrices rather than $q \times q$ matrices.
-Given that $n_k$ is often quite small, say, between 15 and 50, the following result facilitates a further reduction in the computational costs.
+Thus far we have shown that the \emph{GRDA} classifier in \eqref{eq:rda-decomposed} is invariant to the $\bm U_2$ term, yielding a substantial reduction in the computational costs.
+While the classifier is much improved in terms of computational performance compared to the \emph{RDA} classifier originally proposed by \cite{Friedman:1989tm}, we provide two additional results that further improve the computational performance of our \emph{GRDA} classifier.
+First, we demonstrate that the inverse and determinants of $\bm W_k$ can be written so that they are performed on $n_k \times n_k$ matrices rather than the larger $q \times q$ matrices, resulting in further improvements to the computational efficiency of the \emph{GRDA} classifier.
 
@@
 	\begin{align}
-		\bm W_k^{-1} &= \bm \Gamma_k^{-1} - \alpha_k(1 - \lambda) \bm \Gamma_k^{-1} \bm U_1' \bm X_k' \bm Q_k^{-1} \bm X_k \bm U_1 \bm \Gamma_k^{-1},
+		\bm W_k^{-1} &= \bm \Gamma_k^{-1} - \alpha_k(1 - \lambda) \bm \Gamma_k^{-1} \bm U_1' \bm X_k' \bm Q_k^{-1} \bm X_k \bm U_1 \bm \Gamma_k^{-1},\label{eq:W_k_inv}
 	\end{align}
@@
 \begin{proof}
-	Set $\bm A = \bm \Gamma_k$, $\bm B = \alpha_k (1 - \lambda) \bm U_1' \bm X_k'$, $\bm T = \bm I_{n_k}$, and $\bm C = \bm X_k \bm U_1$.
 	To calculate $|\bm W_k|$, from \cite{Harville:2008wja} we apply Theorem 18.1.1, which states that $|\bm A + \bm B \bm T \bm C| = |\bm A| |\bm T| |\bm T^{-1} + \bm C \bm A^{-1} \bm B|$.
-Thus, we have $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$.
-Also from Theorem 18.2.8 in \cite{Harville:2008wja}, we apply the Woodbury formula $(\bm A + \bm B \bm C)^{-1} = \bm A^{-1} - \bm A^{-1} \bm B(\bm I + \bm C \bm A^{-1} \bm B)^{-1} \bm C \bm A^{-1}$, and the proof follows.
+Thus, setting $\bm A = \bm \Gamma_k$, $\bm B = \alpha_k (1 - \lambda) \bm U_1' \bm X_k'$, $\bm T = \bm I_{n_k}$, and $\bm C = \bm X_k \bm U_1$, we have $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$.
+Also,  \eqref{eq:W_k_inv} follows from the well-known Sherman-Woodbury formula \citep[Theorem 18.2.8]{Harville:2008wja}.
 \end{proof}
@@
 Next, we compute $\bm U = \bm X' \bm M \bm D^{+/2}$, where $\bm D^{+/2} = \bm D_q^{-1/2} \oplus \bm 0_{N-q}$.
-We can determine $q$ efficiently by computing the number of nonzero eigenvalues present in $\bm D$.
-However, in our experience, we must take care because the $N-q$ smallest eigenvalues are numerically nonzero, albeit they are near zero.
-Thus, we select $q$ to be the number of eigenvalues that exceeds some given tolerance value, say, $1 \times 10^{-6}$.
-After calculating $q$, we need only retain the first $q$ columns of $\bm U$ to extract $\bm U_1$.
+We determine $q$ efficiently by computing the number of numerically nonzero eigenvalues present in $\bm D$ such that $q$ is the number of eigenvalues that exceeds some tolerance value, say, $1 \times 10^{-6}$.
+After calculating $q$, we retain only the first $q$ columns of $\bm U$ to extract $\bm U_1$.
 
@@
 
+As we can see, our competitors attain decent classification accuracy only when the number of variables selected is small.
+As the number of variables is increased, the classification accuracy of our competitors degrades.
+
 \section{Discussion}
+
+Our formulation of covariance-matrix regularization employed in the \emph{GRDA} classifier includes a large family of shrinkage methods, such as the methods from \cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and a wide variety of other shrinkage methods studied by \cite{Ramey:2011ji} and \cite*{Xu:2009fl}.
 
@@
 \begin{itemize}
-\item Implement classifier
-\item Run simulations
 \item Add simulation results

### 2013-03-24 reports/paper-rda.tex
Added overview of simulation design.

@@
 Furthermore, well-known classifiers, including linear discriminant analysis \emph{(LDA)} and quadratic discriminant analysis \emph{(QDA)}, are incalculable when $p > N$ because the sample covariance matrices employed in both classifiers are singular.
-As a result, many researchers, including \cite*{Merchante:2012vk}, \cite{Witten:2011kc}, \cite*{Pang:2009ik}, \cite*{Huang:2010ju}, and \cite*{Clemmensen:2011kr}, have introduced restrictions on the population covariance matrices in the \emph{LDA} and \emph{QDA} models to reduce number of covariance parameters that must be estimated to ensure that the resulting classifiers are calculable.
+As a result, many researchers, including \cite*{Mai:2012bf}, \cite*{Merchante:2012vk}, \cite{Witten:2011kc}, \cite*{Pang:2009ik}, \cite*{Huang:2010ju}, and \cite*{Clemmensen:2011kr}, have introduced restrictions on the population covariance matrices in the \emph{LDA} and \emph{QDA} models to reduce number of covariance parameters that must be estimated to ensure that the resulting classifiers are calculable.
 For instance, \cite{Dudoit:2002ev} have proposed the diagonal \emph{LDA} (\emph{DLDA}) classifier, a modification of the \emph{LDA} classifier where the class covariance matrices are assumed to be diagonal.
@@
 
-Thus far, we have construction each $\tilde{\bm \Sigma}_k$ in terms of the first $q$ eigenvectors of $\widehat{\bm \Sigma}$.
-In practice, we would usually require a costly eigendecomposition of $\widehat{\bm \Sigma}$, but by utilizing a technique similar to the so-called kernel trick, we can obtain $\bm U_1$ by computing the eigendecomposition of a much smaller $N \times N$ matrix when $p \gg N$ \citep{Hastie:2008dt}.
+Thus far, we have constructed each $\tilde{\bm \Sigma}_k$ in terms of the first $q$ eigenvectors of $\widehat{\bm \Sigma}$.
+By utilizing a technique similar to the so-called kernel trick, we can obtain $\bm U_1$ by computing the eigendecomposition of a much smaller $N \times N$ matrix when $p \gg N$ \citep{Hastie:2008dt}.
 Recall that the pooled sample covariance matrix $\widehat{\bm\Sigma} = \bm X' \bm X$.
-Then, by writing the SVD of $\bm X = \bm M \bm D^{1/2} \bm U'$, we have the spectral decomposition $\widehat{\bm\Sigma} = \bm U \bm D \bm U'$, where $\bm M \in \mathbb{R}_{N \times N}$ is orthogonal, $\bm D^{1/2} \in \mathbb{R}_{N \times N}$ is a diagonal matrix consisting of the singular values of $\bm X$, and $\bm U \in \mathbb{R}_{p \times N}$ is orthogonal.
+Then, by writing the SVD of $\bm X = \bm M \bm D^{1/2} \bm U'$, we have the spectral decomposition $\widehat{\bm\Sigma} = \bm U \bm D \bm U'$, where $\bm M \in \mathbb{R}_{N \times N}$ is orthogonal, $\bm D^{1/2} \in \mathbb{R}_{N \times N}^{\ge}$ is a diagonal matrix consisting of the singular values of $\bm X$, and $\bm U \in \mathbb{R}_{p \times N}$ is orthogonal.
 Now, we see that $\bm X \bm X' = \bm M \bm D \bm M'$ is the eigenvalue decomposition of $\bm X \bm X'$, where $\bm D$ is the matrix of eigenvalues of $\bm X \bm X'$ and the columns of $\bm M$  are the corresponding eigenvectors.
@@
 
-\section{Monte Carlo Simulations}
-
-Our generalization of the \emph{RDA} classifier facilitates a large number of potential models based on the specification of $\alpha_k$ and $\bm \Phi_k$ in \eqref{eq:grda-cov} to shrink the eigenvalues of $\widehat{\bm \Sigma}_k(\lambda)$.
-In our simulations below, we consider two choices of $\alpha_k$ and $\bm \Phi_k$ based on variants of eigenvalue adjustment methods proposed in the literature.
-First, ...
-
-Second, \cite{Srivastava:2007ww} derive an empirical-Bayes covariance-matrix estimator under the assumption that the data are independent and identically distributed multivariate normal observations and then substitute the estimator into the \emph{LDA} classifier to attain the minimum distance empirical Bayes (\emph{MDEB}) classifier.
-To derive the empirical-Bayes covariance-matrix estimator, \cite{Srivastava:2007ww} first assume that $\bm \Sigma^{-1}$ follows a Wishart distribution \emph{a priori} with mean $\lambda^{-1} \bm I_p$, $\lambda > 0$, and degrees of freedom, $l \ge p$.
-The resulting posterior estimator for $\bm \Sigma^{-1}$ is $\frac{N + l}{N}(N^{-1}\lambda \bm I_p + \widehat{\bm \Sigma})^{-1}$.
-Using an empirical Bayes argument to estimate $\lambda$, \cite{Srivastava:2007ww} propose the covariance-matrix estimator
-\begin{align*}
-	\widehat{\bm \Sigma}_{MDEB} = \widehat{\bm \Sigma} + \frac{\tr \{\widehat{\bm \Sigma}\}}{\min\{N, p\}} \bm I_p,
-\end{align*}
-where the shrinkage factor is approximately equal to the average of the nonzero eigenvalues of $\widehat{\bm \Sigma}$.
-We consider a variant of the \emph{MDEB} covariance-matrix estimator because \cite{Srivastava:2007ww} claim that the \emph{MDEB} classifier is the best covariance-matrix regularization method applied to the \emph{LDA} classifier they have encountered in the literature.
-
 \section{High-Dimensional Microarray Data Sets}
 
-In this section, we describe three high-dimensional microarray data sets and compare our proposed \emph{GRDA} classifier with two recently proposed classifiers for small-sample, high-dimensional data: the penalized \emph{LDA} (\emph{PLDA}) classifier from \cite{Witten:2011kc}, the shrinkage-mean-based DLDA (\emph{SmDLDA}) classifier from \cite{Tong:2012hw}, the \emph{RDA} classifier as defined in \cite{Hastie:2008dt}, and the sparse discriminant analysis (\emph{SDA}) classifier from \cite{Clemmensen:2011kr}.
-We calculated 10-fold cross-validation error rates \citep{Hastie:2008dt} to evaluate the performance of each classifier considered.
-Because the \emph{PLDA} classifier has a built-in variable selection method, for fair comparison we compared each of the classifiers using the variables selected from each cross-validation fold.
+In this section, we describe \textbf{three} high-dimensional microarray data sets and compare our proposed \emph{GRDA} classifier with \textbf{three} recently proposed classifiers for small-sample, high-dimensional data: the penalized \emph{LDA} (\emph{PLDA}) classifier from \cite{Witten:2011kc}, the shrinkage-mean-based DLDA (\emph{SmDLDA}) classifier from \cite{Tong:2012hw}, and the sparse discriminant analysis (\emph{SDA}) classifier from \cite{Clemmensen:2011kr}.
+We evaluated the classification performance of each classifier by randomly partitioning the data such that $2/3$ of the observations were allocated as training data and the remaining $1/3$ of the observations were allocated as a test data set.
+For each random partition, we applied the feature-selection method from \cite{Dudoit:2002ev} to choose the $d$ features having the largest ratio of their between-group to within-group sums of squares from the training data.
+We then filtered the test data set with the corresponding $d$ features.
+After training each classifier from the filtered training data, we classified the filtered test data sets and computed the proportion of mislabeled test observations to estimate the classification error rate for each classifier.
+Repeating this process 50 times, we computed the average of the error-rate estimates for each classifier.
+We remark that our simulation design avoids the selection bias discussed in \cite{Ambroise:2002fa}.
+
+We used version 2.15.1 of the open-source statistical software {\tt R} for the classification study.
+For each classifier's {\tt R} implementation, we used the default settings with the exception that we explicitly set the prior probabilities as equal.
+In our {\tt R} implementation of the \emph{GRDA} classifier in the {\tt regdiscrim} package, we set $\alpha_k = 1$, $k = 1, \ldots, K$ for simplicity.
+We estimated $\lambda$ from a grid of 21 equidistant candidate values between 0 and 1, inclusively.
+Similarly, we selected $\gamma$ from a grid consisting of the values $10^{-1}, \ldots, 10^4$, and $10^5$.
+We selected optimal estimates of $\lambda$ and $\gamma$ using 10-fold cross-validation \citep{Hastie:2008dt}.
 
@@
 
-\section{Next Actions}
-
-\begin{itemize}
-\item Add simulation results
-\item Discussion simulation results
-\item Show a timing comparison between our implementation and that of the {\tt klaR} package.
-\item Discuss choice of $\lambda$ and eigenvalue shrinkage, if necessary
-\item Question to answer: What do we do when $\lambda = 0$, corresponding to \emph{QDA}?
-\end{itemize}
-
 \bibliographystyle{plainnat}

### 2013-03-25 summary.r
Added a script to summarize the classification simulations.

@@ -1,0 +1,52 @@
+library('ProjectTemplate')
+load.project()
+
+# Remove the following data sets because there is too much separation for a good comparison
+data_sets <- c("burczynski", "nakayama", "shipp", "singh")
+results <- subset(results, data_set %in% data_sets)
+
+results$q <- as.factor(as.numeric(results$q))
+results$Error <- as.numeric(results$Error)
+results$data_set <- as.character(results$data_set)
+
+
+# Removes the following classifiers from consideration
+results <- subset(results, !Classifier %in% c("Cao", "sdqda"))
+
+p <- ggplot(results, aes(x = q, y = Error))
+p <- p + geom_boxplot(aes(fill = Classifier))
+p + facet_wrap(~ data_set)
+
+
+results_summary <- ddply(results, .(data_set, q, Classifier), summarize,
+                         Avg_Error = mean(Error),
+                         SE_Error = sqrt(Avg_Error * (1 - Avg_Error) / length(Error)))
+
+# Table containing a summary of the data sets
+data_summary <- subset(describe_data(), author %in% data_sets)
+data_summary$K[data_summary$author == "nakayama"] <- 5
+rownames(data_summary) <- NULL
+colnames(data_summary)[1:3] <- c("Author", "Year", "N")
+data_summary$Author <- Hmisc:::capitalize(as.character(data_summary$Author))
+print(xtable(data_summary, caption = "Summary of high-dimensional microarray data sets.",
+             label = "tab:data-summary", digits = 0), include.rownames = FALSE)
+
+
+# Tables containing the classification results for each value of 'q'
+results_summary$data_set <- Hmisc:::capitalize(results_summary$data_set)
+results_summary$q <- as.factor(results_summary$q)
+results_summary$Avg_Error <- round_digits(results_summary$Avg_Error, 3)
+results_summary$SE_Error <- round_digits(results_summary$SE_Error, 3)
+results_summary$Error <- with(results_summary, paste0(Avg_Error, " (", SE_Error, ")"))
+results_summary <- subset(results_summary, select = -c(Avg_Error, SE_Error))
+
+results_tables <- tapply(seq_along(results_summary$q), results_summary$q, function(i) {
+  dcast(results_summary[i, ], data_set ~ Classifier, value.var = "Error")
+})
+
+
+results_xtables <- lapply(names(results_tables), function(q) {
+  tab_caption <- "The average of the test error rates for the microarray data sets with q = "
[…]

### 2013-03-25 sim-microarray.r
Updated simulations to increase the number of iterations from 50 to 500 and reduced the number of data sets to 4.

@@ -5,16 +5,15 @@ set.seed(42)
 num_cores <- 12
 
 train_pct <- 2/3
-num_iterations <- 50
-data_sets <- c("alon", "burczynski", "chiaretti", "christensen", "gordon",
-               "gravier", "khan", "nakayama", "shipp", "singh", "sorlie", "yeoh")
-q <- c(100, 250, 500, 750, 1000)
+num_iterations <- 500
+data_sets <- c("burczynski", "nakayama", "shipp", "singh")
+d <- c(100, 250, 500, 1000)
 
-sim_config <- expand.grid(data_sets = data_sets, q = q, stringsAsFactors = FALSE)
+sim_config <- expand.grid(data_sets = data_sets, d = d, stringsAsFactors = FALSE)
 
 results <- mclapply(seq_len(nrow(sim_config)), function(i) {
   data_set <- sim_config$data_set[i]
-  q <- sim_config$q[i]
+  d <- sim_config$d[i]
   
   data <- load_microarray(data_set)
   data$x <- as.matrix(data$x)
@@ -31,7 +30,7 @@ results <- mclapply(seq_len(nrow(sim_config)), function(i) {
     test_y <- data$y[rand_split$test]
 
     # Apply Dudoit variable selection
-    dudoit_out <- dudoit(train_x, train_y, test_x, q = q)
+    dudoit_out <- dudoit(train_x, train_y, test_x, q = d)
     train_x <- dudoit_out$train_x
     test_x <- dudoit_out$test_x
 
@@ -54,29 +53,21 @@ results <- mclapply(seq_len(nrow(sim_config)), function(i) {
     grda_out <- try_default(with(cv_out, grda(x = train_x, y = train_y, lambda = lambda, gamma = gamma, prior = prior_probs)), NA)
     grda_errors <- try_default(mean(predict(grda_out, test_x)$class != test_y), NA)
 
-    # DLDA and DQDA
+    # DLDA
     dlda_errors <- try_default(mean(predict(dlda(x = train_x, y = train_y, prior = prior_probs), test_x)$class != test_y), NA)
-    dqda_errors <- try_default(mean(predict(dqda(x = train_x, y = train_y, prior = prior_probs), test_x)$class != test_y), NA)
 
     # Pang, Tong, and Zhao (2009) - Biometrics
     sdlda_errors <- try_default(mean(predict(sdlda(x = train_x, y = train_y, prior = prior_probs), test_x)$class != test_y), NA)
[…]

### 2013-03-25 summary.r

@@ -28,6 +28,8 @@ data_summary$K[data_summary$author == "nakayama"] <- 5
 rownames(data_summary) <- NULL
 colnames(data_summary)[1:3] <- c("Author", "Year", "N")
 data_summary$Author <- Hmisc:::capitalize(as.character(data_summary$Author))
+data_summary$Author <- with(data_summary, paste0(Author, " et al. (", Year, ")"))
+data_summary <- subset(data_summary, select = -Year)
 print(xtable(data_summary, caption = "Summary of high-dimensional microarray data sets.",
              label = "tab:data-summary", digits = 0), include.rownames = FALSE)
 
@@ -46,7 +48,7 @@ results_tables <- tapply(seq_along(results_summary$q), results_summary$q, functi
 
 
 results_xtables <- lapply(names(results_tables), function(q) {
-  tab_caption <- "The average of the test error rates for the microarray data sets with q = "
-  tab_caption <- paste0(tab_caption, q, ". Standard errors are given in parentheses.")
-  xtable(results_tables[[q]], caption = tab_caption, label = paste0("tab:results-q", q))
+  tab_caption <- "The average of the test error rates for the microarray data sets with $d = "
+  tab_caption <- paste0(tab_caption, q, "$. Approximate standard errors are given in parentheses.")
+  xtable(results_tables[[q]], caption = tab_caption, label = paste0("tab:results-d", q))
 })

### 2013-03-25 reports/paper-rda.tex
Added simulation results, a discussion of the results, and a discussion to the paper.

@@
 
-Alternatively, several researchers, including \cite*{Zhang:2010va}, \cite{Ye:2009gd}, \cite{Ji:2008wp}, \cite*{Guo:2007te}, \cite{Srivastava:2007ud}, \cite{Ye:2006vx}, \cite*{Thomaz:2006ef}, \cite{Ye:2005uu}, \cite{Mkhadri:1995jp}, and \cite{Peck:1982tt}, have proposed regularization methods to improve the estimation of the population covariance matrices before applying the \emph{LDA} and \emph{QDA} classifiers.
-Covariance-matrix regularization methods often are based on the well-known ridge-regression approach from \cite{Hoerl:1970cd} and adjust the eigenvalues of the sample covariance matrices to ensure that the estimator is positive definite.
-Typically, a positive shrinkage value is added to each eigenvalue estimate, replacing the zero eigenvalues with the shrinkage constant and stabilizing the eigenvalues near zero to ensure that the inverse of the the sample covariance matrix is well-posed.
-This common form of covariance-matrix regularization is equivalent to summing the original covariance-matrix estimator with the identity matrix scaled by the shrinkage constant.
-For an excellent overview of the regularization problem in discriminant analysis, see \cite{Murphy:2012uq} and \cite{Mkhadri:1997gy}.
+Alternatively, several researchers, including \cite*{Zhang:2010va}, \cite{Ye:2009gd}, \cite{Ji:2008wp}, \cite*{Guo:2007te}, \cite{Srivastava:2007ud}, \cite{Ye:2006vx}, and \cite{Ye:2005uu}, have proposed regularization methods to improve the estimation of the population covariance matrices before applying the \emph{LDA} and \emph{QDA} classifiers.
+Covariance-matrix regularization methods often are based on the well-known ridge-regression approach from \cite{Hoerl:1970cd} and adjust the eigenvalues of the sample covariance matrices to ensure positive definiteness.
+For excellent overviews of the regularization problem in discriminant analysis, see \cite{Murphy:2012uq} and \cite{Mkhadri:1997gy}.
 
 Here, we consider a generalization of the popular \emph{RDA} classifier of \cite{Friedman:1989tm}, who has proposed a weighted average of the covariance-matrix estimators from the \emph{LDA} and \emph{QDA} classifiers.
-Because the resulting biased covariance-matrix estimator can remain singular or nearly singular, \cite{Friedman:1989tm} has additionally incorporated applied covariance-matrix shrinkage, where the shrinkage constant is a function of the average eigenvalue of the original estimator.
+\cite{Friedman:1989tm} has additionally incorporated a covariance-matrix shrinkage term to ensure that the \emph{RDA} classifier is well-posed.
 \cite{Friedman:1989tm} has shown that the \emph{RDA} classifier can attain improved classification accuracy, compared to the \emph{LDA} and \emph{QDA} classifiers when $p > N$.
@@
 Furthermore, the \emph{RDA} classifier has two tuning parameters that are selected via cross-validation, which adds a substantial computational burden to the model-selection process.
-Despite the introduction of a computational shortcut by \cite{Friedman:1989tm} for model selection to alleviate some of the computation, the \emph{RDA} classifier is ill-advised on modern data sets because the model selection requires on the order of $Kp^3$ calculations for each tuning-parameter pair for each fold in the cross-validation scheme.
+The \emph{RDA} classifier is ill-advised on modern data sets because the model selection requires $O(Kp^3)$ calculations for each tuning-parameter pair for each fold in the cross-validation scheme.
 
@@
 In particular, we compare our \emph{GRDA} classifier with an alternatively-defined \emph{RDA} classifier from \cite*{Hastie:2008dt}, the regularized \emph{LDA} classifier from \cite{Guo:2007te}, a recently proposed variant of the \emph{DLDA} classifier from \cite*{Tong:2012hw}, and the penalized \emph{LDA} classifier from \cite{Witten:2011kc}.
-We conduct our comparison of the classification accuracy of the competing classifiers using three small-sample, high-dimensional microarray data sets and examine the classification accuracy of the competing classifiers using \textbf{three} simulation configurations.
+We conduct our comparison of the classification accuracy of the competing classifiers using four small-sample, high-dimensional microarray data sets.
 
@@
 In Section 3 we show that our proposed classifier includes several well-known regularized classifiers as special cases and briefly discuss some commonalities among them.
-In Section 4 we describe the \textbf{three} simulation configurations and three small-sample, high-dimensional microarray data sets and then demonstrate that our two proposed classifiers are superior to the competing classifiers in terms of classification accuracy.
+In Section 4 we describe four small-sample, high-dimensional microarray data sets and then demonstrate that our two proposed classifiers are superior to the competing classifiers in terms of classification accuracy.
 We then conclude with a brief discussion in Section 5.
@@
 In discriminant analysis we wish to assign correctly an unlabeled $p$-dimensional observation vector $\bm x$ to one of $K$ unique, known classes (or populations) by constructing a classifier from $n$ training observations that can accurately predict the class membership of $\bm x$.
-Let $\bm x_i = (x_{i1}, x_{i2}, \ldots, x_{ip}) \in \mathbb{R}_{p \times 1}$ be the $i$th observation $(i = 1, \ldots, n)$ with true, unique membership $y_i \in \{\omega_1, \ldots, \omega_K\}$, where $\mathbb{R}_{a \times b}$ denotes the matrix space of all $a \times b$ matrices over the real field $\mathbb{R}$.
+Let $\bm x_i = (x_{i1}, \ldots, x_{ip}) \in \mathbb{R}_{p \times 1}$ be the $i$th observation $(i = 1, \ldots, N)$ with true, unique membership $y_i \in \{\omega_1, \ldots, \omega_K\}$, where $\mathbb{R}_{a \times b}$ denotes the matrix space of all $a \times b$ matrices over the real field $\mathbb{R}$.
 We assume that $(\bm x_i, y_i)$ is a realization from a mixture distribution $p(\bm x) = \sum_{k=1}^K p(\bm x | \omega_k) p(\omega_k)$, where $p(\bm x | \omega_k)$ is the probability density function (PDF) of the $k$th class, $p(\omega_k)$ is prior probability of class membership of the $k$th class.
@@
 \begin{align}
-	\widehat{\bm\Sigma} = \frac{1}{n} \sum_{k=1}^K n_k \widehat{\bm\Sigma}_k. \label{eq:pooled-cov}
+	\widehat{\bm\Sigma} = \frac{1}{N} \sum_{k=1}^K n_k \widehat{\bm\Sigma}_k. \label{eq:pooled-cov}
 \end{align}
@@
 Specifically, the zero eigenvalues are replaced with $\gamma_k$, so that \eqref{eq:ridge-estimator} is positive definite.
-For additional covariance-matrix regularization methods, see \cite{Ramey:2011ji}, \cite{Xu:2009fl}, and \cite{Ye:2009gd}.
+For additional covariance-matrix regularization methods, see \cite{Ramey:2013ji}, \cite{Xu:2009fl}, and \cite{Ye:2009gd}.
 
@@
 \end{align}
-where $\lambda \in [0, 1]$, $n_k(\lambda) = (1 - \lambda) n_k + \lambda n$, $\bm S_k = n_k \widehat{\bm\Sigma}_k$, $\bm S = \sum_{k=1}^K \bm S_k$, and $\bm S_k(\lambda) = (1 - \lambda) \bm S_k + \lambda \bm S$.
+where $\lambda \in [0, 1]$, $n_k(\lambda) = (1 - \lambda) n_k + \lambda N$, $\bm S_k = n_k \widehat{\bm\Sigma}_k$, $\bm S = \sum_{k=1}^K \bm S_k$, and $\bm S_k(\lambda) = (1 - \lambda) \bm S_k + \lambda \bm S$.
 We can interpret \eqref{eq:sig-lambda} as a covariance matrix estimator for class $\omega_k$ that borrows from $\widehat{\bm\Sigma}$ in \eqref{eq:pooled-cov} to better estimate $\bm \Sigma_k$.
@@
 
-In this section, we describe \textbf{three} high-dimensional microarray data sets and compare our proposed \emph{GRDA} classifier with \textbf{three} recently proposed classifiers for small-sample, high-dimensional data: the penalized \emph{LDA} (\emph{PLDA}) classifier from \cite{Witten:2011kc}, the shrinkage-mean-based DLDA (\emph{SmDLDA}) classifier from \cite{Tong:2012hw}, and the sparse discriminant analysis (\emph{SDA}) classifier from \cite{Clemmensen:2011kr}.
+In this section, we describe four high-dimensional microarray data sets and compare our proposed \emph{GRDA} classifier with five recently proposed classifiers designed for small-sample, high-dimensional data: the \emph{DLDA} classifier from \cite{Dudoit:2002ev}, the shrunken centroids \emph{RDA} (\emph{SCRDA}) classifier from \cite{Guo:2007te}, the shrinkage-based \emph{DLDA} (\emph{SDLDA}) classifier from \cite{Pang:2009ik}, and the shrinkage-mean-based DLDA (\emph{SmDLDA}) classifier from \cite{Tong:2012hw}, and the penalized \emph{LDA} (\emph{PLDA}) classifier from \cite{Witten:2011kc}.
 We evaluated the classification performance of each classifier by randomly partitioning the data such that $2/3$ of the observations were allocated as training data and the remaining $1/3$ of the observations were allocated as a test data set.
@@
 After training each classifier from the filtered training data, we classified the filtered test data sets and computed the proportion of mislabeled test observations to estimate the classification error rate for each classifier.
-Repeating this process 50 times, we computed the average of the error-rate estimates for each classifier.
+Repeating this process 500 times, we computed the average of the error-rate estimates for each classifier.
 We remark that our simulation design avoids the selection bias discussed in \cite{Ambroise:2002fa}.
+In Table \ref{tab:data-summary}, we summarize the attributes of each microarray data set.
 
@@
 We estimated $\lambda$ from a grid of 21 equidistant candidate values between 0 and 1, inclusively.
-Similarly, we selected $\gamma$ from a grid consisting of the values $10^{-1}, \ldots, 10^4$, and $10^5$.
+Similarly, we estimated $\gamma$ from a grid consisting of the values $10^{-1}, \ldots, 10^4$, and $10^5$.
 We selected optimal estimates of $\lambda$ and $\gamma$ using 10-fold cross-validation \citep{Hastie:2008dt}.
+
+ % latex table generated in R 2.15.1 by xtable 1.7-0 package
+ % Sun Mar 24 19:54:37 2013
+ \begin{table}[ht]
+ \begin{center}
+ \begin{tabular}{lrrrl}
+   \hline
+ Author & $N$ & $p$ & $K$ & Disease \\
+   \hline
+ Burczynski et al. (2006) & 127 & 22283 & 3 & Crohn's Disease \\
+   Nakayama et al. (2001) & 105 & 22283 & 5 & Sarcoma \\
+   Shipp et al. (2002) & 58 & 6817 & 2 & Lymphoma \\
+   Singh et al. (2002) & 102 & 12600 & 2 & Prostate Cancer \\
+    \hline
+ \end{tabular}
+ \caption{Summary of high-dimensional microarray data sets.}
+ \label{tab:data-summary}
+ \end{center}
+ \end{table}
 
@@
 
+\subsection{\cite{Shipp:2002ka} Data Set}
+
+Approximately 30-40\% of adult non-Hodgkin lymphomas are diffuse large B-cell lymphomas (DLBCLs) \cite{Shipp:2002ka}.
+However, only a small proportion of DLBCL patients are cured with modern chemotherapeutic regimens.
+Several models have been proposed, such as the International Prognostic Index (IPI), to determine a patient's curability.
+These models rely on clinical covariates, such as age, to determine if the patient can be cured and are often ineffective.
+\cite{Shipp:2002ka} have argued that more effective means are desired to determine a patient's curability.
+The authors measured 6,817 gene expression levels from 58 DLBCL patient samples with customized cDNA (‘lymphochip’) microarrays to investigate the curability of patients treated with cyclophosphamide, adriamycin, vincristine, and prednisone (CHOP)-based chemotherapy.
+Among the 58 DLBCL patient samples, 32 are from cured patients while 26 are from patients with fatal or refractory disease.
+
 \subsection{\cite{Singh:2002fh} Data Set}
@@
 The authors used oligonucleotide microarrays containing probes for approximately 12,600 genes and expressed sequence tags.
-They have reported that 102 of the radical prostatectomy specimens are of high quality: 52 prostate tumour samples and 50 non-tumour prostate samples.
+They have reported that 102 of the radical prostatectomy specimens are of high quality: 52 prostate tumor samples and 50 non-tumor prostate samples.
 
@@
 
-As we can see, our competitors attain decent classification accuracy only when the number of variables selected is small.
-As the number of variables is increased, the classification accuracy of our competitors degrades.
+For the microarray data sets, we computed error-rate estimates for the competing classifiers after reducing the original feature space dimension with variable selection to $d=100,250,500$, and $1000$.
+Our choices for the values of $d$ are similar to those of \cite{Xu:2009fl}.
+We report the error-rate estimates along with approximate standard errors in parentheses in Tables \ref{tab:results-d100} -- \ref{tab:results-d1000} along with the chosen reduced dimension $d$.
+
+The \emph{GRDA} classifier was superior in classification performance for the majority of the simulations.
+While the error-rate estimates for the majority of the classifiers increased as $d$ was increased, the classification accuracy for the \emph{GRDA} classifier improved as $d$ increased for the Burczynski and Nakayama data sets.
+For the Nakayama data set, the \emph{GRDA} classifier was outperformed by the Dudoit, Pang, Tong, and Witten classifiers for small values of $d$ but attained classification accuracy comparable to its competitors for larger values of $d$.
+Moreover, the only situation where the other classifiers considered were competitive with the \emph{GRDA} classifier was in the case of the Nakayama data set.
+
+The error-rate estimates for the Dudoit, Pang, Tong, and Witten classifiers were comparable regardless of the choice of $d$.
+In particular, the error-rate estimates for Dudoit's and Tong's methods were nearly identical, which suggests that the improved mean estimator employed by Tong provides little improvement to the classification performance in the \emph{DLDA} classifier.
+The popular classifier from Guo is competitive only for large values of $d$ for the Shipp and Singh data sets.
+For the Burczynski and Nakayama data sets, the implementation of Guo's method in the {\tt rda} package often classified all of the test observations into a single class, thus yielding larger error rates than its competitors.
+
+\[ \left[\text{Table 2 goes about here }\right]\]
+
+\[ \left[\text{Table 3 goes about here }\right]\]
+
+\[ \left[\text{Table 4 goes about here }\right]\]
+
+\[ \left[\text{Table 5 goes about here }\right]\]
 
@@
 
-Our formulation of covariance-matrix regularization employed in the \emph{GRDA} classifier includes a large family of shrinkage methods, such as the methods from \cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and a wide variety of other shrinkage methods studied by \cite{Ramey:2011ji} and \cite*{Xu:2009fl}.
+The analysis of high-dimensional microarray data sets is important to modern biomedical studies and presents challenges with supervised classification because the number of gene expressions greatly outnumbers the number of collected responses.
+Consequently, regularization techniques are often employed to stabilize popular classifiers for improved classification performance.
+We have provided a generalization of the well-known \emph{RDA} classifier from \cite{Friedman:1989tm}, aimed for small-sample, high-dimensional classification.
+The covariance-matrix regularization employed in the \emph{GRDA} classifier includes a large family of shrinkage methods.
+Our formulation of the \emph{GRDA} classifier in terms of the nonzero eigenvalues of the pooled sample covariance matrix indicates that eigenvalue-adjustment methods applied to the \emph{LDA} classifier explains the similiarities in classification accuracy of several regularization methods applied to the \emph{LDA} classifier reported by \cite{Ramey:2013ji} in a high-dimensional classification study.
+
+We have provided an implementation of the \emph{GRDA} classifier in the {\tt regdiscrim} R package, available on CRAN.
+Our implementation facilitates a rapid classification of high-dimensional data by employing the novel results provided in this paper.
+Moreover, our 10-fold cross-validation procedure typically requires a couple of seconds for data sets having large $p$.
+When preliminary variable selection is applied to the data as we performed in our classification study, the time to perform model selection is nearly instantaneous, confirming the computational complexity reported regarding the operations employed in the \emph{GRDA} classifier.
 
@@
 
+\clearpage
+
+% latex table generated in R 2.15.1 by xtable 1.7-0 package
+% Mon Mar 25 07:04:20 2013
+\begin{table}[ht]
+\begin{center}
+\begin{tabular}{lllllll}
+  \hline
+Dataset & Dudoit & GRDA & Guo & Pang & Tong & Witten \\ 
+  \hline
+Burczynski & 0.232 (0.019) & 0.179 (0.017) & 0.513 (0.022) & 0.232 (0.019) & 0.231 (0.019) & 0.232 (0.019) \\ 
+  Nakayama & 0.262 (0.020) & 0.308 (0.021) & 0.586 (0.022) & 0.262 (0.020) & 0.262 (0.020) & 0.278 (0.020) \\ 
+  Shipp & 0.136 (0.015) & 0.106 (0.014) & 0.310 (0.021) & 0.136 (0.015) & 0.136 (0.015) & 0.136 (0.015) \\ 
+  Singh & 0.111 (0.014) & 0.097 (0.013) & 0.338 (0.021) & 0.111 (0.014) & 0.111 (0.014) & 0.111 (0.014) \\ 
+   \hline
+\end{tabular}
+\caption{The average of the test error rates for the microarray data sets with $d = 100$. Approximate standard errors are given in parentheses.}
+\label{tab:results-d100}
+\end{center}
+\end{table}
+
+% latex table generated in R 2.15.1 by xtable 1.7-0 package
+% Mon Mar 25 07:04:37 2013
+\begin{table}[ht]
+\begin{center}
+\begin{tabular}{lllllll}
+  \hline
+Dataset & Dudoit & GRDA & Guo & Pang & Tong & Witten \\ 
+  \hline
+Burczynski & 0.244 (0.019) & 0.163 (0.017) & 0.459 (0.022) & 0.244 (0.019) & 0.244 (0.019) & 0.243 (0.019) \\ 
+  Nakayama & 0.244 (0.019) & 0.267 (0.020) & 0.498 (0.022) & 0.244 (0.019) & 0.244 (0.019) & 0.259 (0.020) \\ 
+  Shipp & 0.153 (0.016) & 0.104 (0.014) & 0.188 (0.017) & 0.153 (0.016) & 0.153 (0.016) & 0.154 (0.016) \\ 
+  Singh & 0.135 (0.015) & 0.117 (0.014) & 0.188 (0.017) & 0.135 (0.015) & 0.135 (0.015) & 0.135 (0.015) \\ 
+   \hline
+\end{tabular}
+\caption{The average of the test error rates for the microarray data sets with $d = 250$. Approximate standard errors are given in parentheses.}
+\label{tab:results-d250}
+\end{center}
+\end{table}
+
+% latex table generated in R 2.15.1 by xtable 1.7-0 package
+% Mon Mar 25 07:04:48 2013
+\begin{table}[ht]
+\begin{center}
+\begin{tabular}{lllllll}
+  \hline
+Dataset & Dudoit & GRDA & Guo & Pang & Tong & Witten \\ 
+  \hline
+Burczynski & 0.243 (0.019) & 0.159 (0.016) & 0.429 (0.022) & 0.243 (0.019) & 0.243 (0.019) & 0.243 (0.019) \\ 
+  Nakayama & 0.247 (0.019) & 0.248 (0.019) & 0.465 (0.022) & 0.247 (0.019) & 0.248 (0.019) & 0.261 (0.020) \\ 
+  Shipp & 0.169 (0.017) & 0.097 (0.013) & 0.119 (0.015) & 0.169 (0.017) & 0.169 (0.017) & 0.170 (0.017) \\ 
+  Singh & 0.174 (0.017) & 0.125 (0.015) & 0.155 (0.016) & 0.174 (0.017) & 0.174 (0.017) & 0.174 (0.017) \\ 
+   \hline
+\end{tabular}
+\caption{The average of the test error rates for the microarray data sets with $d = 500$. Approximate standard errors are given in parentheses.}
+\label{tab:results-d500}
+\end{center}
+\end{table}
+
+% latex table generated in R 2.15.1 by xtable 1.7-0 package
+% Mon Mar 25 07:05:01 2013
+\begin{table}[ht]
+\begin{center}
+\begin{tabular}{lllllll}
+  \hline
+Dataset & Dudoit & GRDA & Guo & Pang & Tong & Witten \\ 
+  \hline
+Burczynski & 0.249 (0.019) & 0.168 (0.017) & 0.381 (0.022) & 0.249 (0.019) & 0.249 (0.019) & 0.249 (0.019) \\ 
+  Nakayama & 0.237 (0.019) & 0.238 (0.019) & 0.460 (0.022) & 0.237 (0.019) & 0.237 (0.019) & 0.246 (0.019) \\ 
+  Shipp & 0.196 (0.018) & 0.106 (0.014) & 0.095 (0.013) & 0.196 (0.018) & 0.195 (0.018) & 0.198 (0.018) \\ 
+  Singh & 0.224 (0.019) & 0.140 (0.016) & 0.142 (0.016) & 0.224 (0.019) & 0.224 (0.019) & 0.224 (0.019) \\ 
+   \hline
+\end{tabular}
+\caption{The average of the test error rates for the microarray data sets with $d = 1000$. Approximate standard errors are given in parentheses.}
+\label{tab:results-d1000}
+\end{center}
+\end{table}
+
 \end{document}

### 2013-03-25 summary.r

@@ -2,28 +2,21 @@ library('ProjectTemplate')
 load.project()
 
 # Remove the following data sets because there is too much separation for a good comparison
-data_sets <- c("burczynski", "nakayama", "shipp", "singh")
-results <- subset(results, data_set %in% data_sets)
+results <- na.omit(results)
 
-results$q <- as.factor(as.numeric(results$q))
-results$Error <- as.numeric(results$Error)
+results$d <- as.factor(results$d)
 results$data_set <- as.character(results$data_set)
 
-
-# Removes the following classifiers from consideration
-results <- subset(results, !Classifier %in% c("Cao", "sdqda"))
-
-p <- ggplot(results, aes(x = q, y = Error))
+p <- ggplot(results, aes(x = d, y = Error))
 p <- p + geom_boxplot(aes(fill = Classifier))
 p + facet_wrap(~ data_set)
 
-
-results_summary <- ddply(results, .(data_set, q, Classifier), summarize,
+results_summary <- ddply(results, .(data_set, d, Classifier), summarize,
                          Avg_Error = mean(Error),
-                         SE_Error = sqrt(Avg_Error * (1 - Avg_Error) / length(Error)))
+                         SE_Error = sqrt(mean(Error) * (1 - mean(Error)) / length(Error)))
 
 # Table containing a summary of the data sets
-data_summary <- subset(describe_data(), author %in% data_sets)
+data_summary <- subset(describe_data(), author %in% unique(results$data_set))
 data_summary$K[data_summary$author == "nakayama"] <- 5
 rownames(data_summary) <- NULL
 colnames(data_summary)[1:3] <- c("Author", "Year", "N")
@@ -36,19 +29,19 @@ print(xtable(data_summary, caption = "Summary of high-dimensional microarray dat
 
 # Tables containing the classification results for each value of 'q'
 results_summary$data_set <- Hmisc:::capitalize(results_summary$data_set)
-results_summary$q <- as.factor(results_summary$q)
+results_summary$d <- as.factor(results_summary$d)
 results_summary$Avg_Error <- round_digits(results_summary$Avg_Error, 3)
 results_summary$SE_Error <- round_digits(results_summary$SE_Error, 3)
 results_summary$Error <- with(results_summary, paste0(Avg_Error, " (", SE_Error, ")"))
 results_summary <- subset(results_summary, select = -c(Avg_Error, SE_Error))
+colnames(results_summary) <- c("Dataset", "d", "Classifier", "Error")
 
[…]

### 2013-03-26 reports/paper-rda.Rnw
Finished first draft of paper.

--- 2013-03-25 reports/paper-rda.tex
@@
 \begin{abstract}
-\cite{Friedman:1989tm} has proposed the popular regularized discriminant analysis (\emph{RDA}) classifier for the small-sample, high-dimensional setting.
-This \emph{RDA} classifier employs a regularized covariance-matrix estimator, which incorporates a ridge-like shrinkage factor by adding a positive constant to the zero-eigenvalue estimates to ensure positive definiteness.
-Here, we propose a novel generalization of Friedman's \emph{RDA} classifier that is more computationally efficient than its original counterpart because the inverses and determinants computed in the decision rule can be reduced from the original feature space to a linear subspace with dimensions corresponding to the individual class sample sizes.
-Additionally, our generalization of the \emph{RDA} (\emph{GRDA}) classifier enables efficient estimation of optimal tuning parameters from a large set of candidate values, simplifies the construction of the regularized covariance-matrix estimator to improve its interpretation, and allows for a more flexible classification model.
-Moreover, we demonstrate with several high-dimensional microarray data sets that our proposed classifier is competitive and often superior to several recently proposed sparse and regularized classifiers.
+
+A reliable classification of tumors from gene expression microarray data is essential for successful diagnosis and treatment of diseases such as cancer, but the analysis of microarray data often presents difficult challenges because the number of gene expression levels is typically much larger than the sample size.
+Consequently, several sparse and regularized classifiers have been proposed often with strict covariance assumptions or eigenvalue adjustments to ensure that the classifiers are well-posed.
+From \cite{Friedman:1989tm} we consider specifically the popular regularized discriminant analysis (\emph{RDA}) classifier, which requires computationally expensive matrix operations and involves a time-consuming model-selection procedure, making the \emph{RDA} classifier impractical for high-dimensional data.
+We propose a novel generalization of the \emph{RDA} (\emph{GRDA}) classifier that improves upon the interpretation of the \emph{RDA} classifier and incorporates a more general covariance-matrix regularization approach, which includes as special cases several proposed sparse and regularization methods.
+Furthermore, we address the computational shortcomings of the \emph{RDA} classifier by integrating a novel dimension-reduction technique that enables a computationally efficient model-selection procedure.
+Moreover, we demonstrate with several high-dimensional microarray data sets that our proposed classifier is competitive and often superior to several recently proposed sparse and regularized classifiers in terms of classification accuracy.
 Finally, we provide an implementation of the \emph{GRDA} classifier in the {\tt regdiscrim} R package, available on CRAN.
+
 \end{abstract}
@@
 Furthermore, well-known classifiers, including linear discriminant analysis \emph{(LDA)} and quadratic discriminant analysis \emph{(QDA)}, are incalculable when $p > N$ because the sample covariance matrices employed in both classifiers are singular.
-As a result, many researchers, including \cite*{Mai:2012bf}, \cite*{Merchante:2012vk}, \cite{Witten:2011kc}, \cite*{Pang:2009ik}, \cite*{Huang:2010ju}, and \cite*{Clemmensen:2011kr}, have introduced restrictions on the population covariance matrices in the \emph{LDA} and \emph{QDA} models to reduce number of covariance parameters that must be estimated to ensure that the resulting classifiers are calculable.
+As a result, many researchers, including \cite*{Mai:2012bf}, \cite*{Merchante:2012vk}, \cite*{Clemmensen:2011kr}, and \cite*{Huang:2010ju}, have introduced restrictions on the population covariance matrices in the \emph{LDA} and \emph{QDA} models to reduce number of covariance parameters that must be estimated to ensure that the resulting classifiers are calculable.
 For instance, \cite{Dudoit:2002ev} have proposed the diagonal \emph{LDA} (\emph{DLDA}) classifier, a modification of the \emph{LDA} classifier where the class covariance matrices are assumed to be diagonal.
@@
 
-Here, we consider a generalization of the popular \emph{RDA} classifier of \cite{Friedman:1989tm}, who has proposed a weighted average of the covariance-matrix estimators from the \emph{LDA} and \emph{QDA} classifiers.
+Here, we consider a generalization of the popular \emph{RDA} classifier of \cite{Friedman:1989tm}, who has proposed pooling of the covariance-matrix estimators from the \emph{LDA} and \emph{QDA} classifiers.
 \cite{Friedman:1989tm} has additionally incorporated a covariance-matrix shrinkage term to ensure that the \emph{RDA} classifier is well-posed.
@@
 Furthermore, the \emph{RDA} classifier has two tuning parameters that are selected via cross-validation, which adds a substantial computational burden to the model-selection process.
-The \emph{RDA} classifier is ill-advised on modern data sets because the model selection requires $O(Kp^3)$ calculations for each tuning-parameter pair for each fold in the cross-validation scheme.
+The \emph{RDA} classifier is ill-advised on modern data sets because the model selection requires $O(Kp^3)$ calculations for each tuning-parameter pair.
 
-In this paper we address the computational shortcomings of the \emph{RDA} classifier.
+In this paper we provide a generalization of the \emph{RDA} classifier by first simplifying the pooling matrix employed in the \emph{RDA} classifier.
+We construct instead a convex combination of the covariance-matrix estimators employed in the \emph{LDA} and \emph{QDA} classifiers.
+We demonstrate that this simplification improves the interpretation of the pooling operation of the \emph{RDA} classifier.
+The \emph{LDA} classifier involves the assumption that the population covariance matrices are equal, which is often too stringent in practice \citep{Clemmensen:2011kr}.
+The \emph{GRDA} classifier allows us to relax the linearity assumption by including unequal covariance matrices in the model.
+The \emph{GRDA} classifier is then a compromise between the \emph{LDA} and \emph{QDA} classifiers,  so that the \emph{GRDA} classifier includes both the \emph{LDA} and \emph{QDA} classifiers as special cases.
+We then introduce a more general shrinkage term to the \emph{RDA} classifier to establish our \emph{GRDA} classifier.
+Our formulation of covariance-matrix regularization employed in the \emph{GRDA} classifier includes a large family of shrinkage methods, such as the methods from \cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and a wide variety of other shrinkage methods studied by \cite{Ramey:2013ji} and \cite*{Xu:2009fl}.
+
+We show that the \emph{GRDA} classifier is invariant to any adjustments of the eigenvalues corresponding to the null space of the pooled sample covariance matrix.
 Applying a technique similar to \cite{Ye:2006tq}, we decompose the \emph{RDA} classifier and show that the sum related to the null space of the pooled sample covariance matrix can be discarded.
-In turn, this omission provides a substantial increase in the computational efficiency of the \emph{RDA} classifier and its model-selection.
-In fact, we show rigorously that the inverse and determinants can be reduced to calculations of $n_k \times n_k$ matrices, where $n_k$ is the sample size of the $k$th class, $k = 1, \ldots, K$.
+Hence, no gain in classification performance can be obtained by adjusting these zero eigenvalues.
+This invariance property addresses the computational shortcomings of the \emph{RDA} classifier and provides a substantial increase in the computational efficiency of the \emph{RDA} classifier and its model selection.
+In fact, we show that the inverse and determinants can be reduced to calculations of $n_k \times n_k$ matrices, where $n_k$ is the sample size of the $k$th class, $k = 1, \ldots, K$.
 This reduction in calculation is clearly advantageous because $n_k$ is often quite small, say, 15--20, in microarray data sets.
-Thus, we demonstrate that the calculation of the \emph{RDA} decision rule involving inverses and determinants of extremely large covariance matrices is equivalent a decision rule involving inverses and determinants on much smaller matrices.
+Thus, through a dimension-reduction technique we demonstrate that the calculation of the \emph{RDA} decision rule involving inverses and determinants of extremely large covariance matrices is equivalent a decision rule involving inverses and determinants on much smaller matrices.
 
-Also, for our \emph{GRDA} classifier we first simplify the pooling matrix in the \emph{RDA} classifier by constructing instead a convex combination of the covariance-matrix estimators employed in the \emph{LDA} and \emph{QDA} classifiers.
-This simplification is advantageous in that it improves the interpretation of the pooling operation of the \emph{RDA} classifier.
-We show that under mild regularity conditions the \emph{GRDA} classifier is invariant to any adjustments of the eigenvalues corresponding to the null space of the pooled sample covariance matrix.
-Hence, no gain in classification performance can be obtained by adjusting these zero eigenvalues.
-Although \cite{Ji:2008wp} and \cite{Ye:2006jm} have provided a similar invariance property for the \emph{LDA} decision rule, the \emph{LDA} classifier involves the assumption that the population covariance matrices are equal, which is often too stringent in practice \citep{Clemmensen:2011kr}.
-The \emph{GRDA} classifier allows us to relax this assumption by including unrestricted covariance matrices in the model.
-The \emph{GRDA} classifier is then a compromise between the two approaches, favoring the stronger linear assumption if a larger value of the pooling parameter is selected as optimal.
-Our formulation provides a general framework from which we can elucidate properties of covariance-regularization methods and the relationships among them in the linear case, the quadratic case, and the cases in between.
-
-Many of the recently proposed classifiers have a variable-selection method integrated.
-While this is advantageous in potentially determining biomarkers that are correlated with the classification label, these methods often yield poor classification accuracy if a large number of variables are retained.
-Moreover, these methods can require large computational costs, making these methods undesirable in practice.
-To study these issues in depth, we consider several simulated data sets, where we show that the classification degrades as the number of features is increased.
-Contrarily, our proposed \emph{GRDA} classifier handles the large number of features well because our included dimension-reduction method retains only the classificatory information in the covariance-matrix estimators.
-Moreover, while the competing classifiers exhibit large computational costs as the number of features in the original data set is increased, the \emph{GRDA} classifier handles the larger number of features efficiently.
-
-In addition to the improvement in computational performance of classifiers with regularized covariance matrices, we demonstrate that our \emph{GRDA} classifier is competitive with and often superior to several recent classifiers designed for small-sample, high-dimensional data.
-In particular, we compare our \emph{GRDA} classifier with an alternatively-defined \emph{RDA} classifier from \cite*{Hastie:2008dt}, the regularized \emph{LDA} classifier from \cite{Guo:2007te}, a recently proposed variant of the \emph{DLDA} classifier from \cite*{Tong:2012hw}, and the penalized \emph{LDA} classifier from \cite{Witten:2011kc}.
+In addition to the improvement in computational performance, we demonstrate that our \emph{GRDA} classifier is competitive with and often superior to several recent classifiers designed for small-sample, high-dimensional data.
+In particular, we compare our \emph{GRDA} classifier with the \emph{DLDA} classifier from \cite{Dudoit:2002ev} and the recently proposed classifiers from \cite{Guo:2007te}, \cite{Pang:2009ik}, \cite{Tong:2012hw}, and \cite{Witten:2011kc}.
 We conduct our comparison of the classification accuracy of the competing classifiers using four small-sample, high-dimensional microarray data sets.
@@
 We have organized the remainder of the paper as follows.
-In Section 2 we present our generalization of the \emph{RDA} classifier after reviewing its original counterpart from \cite{Friedman:1989tm}.
-In Section 3 we show that our proposed classifier includes several well-known regularized classifiers as special cases and briefly discuss some commonalities among them.
-In Section 4 we describe four small-sample, high-dimensional microarray data sets and then demonstrate that our two proposed classifiers are superior to the competing classifiers in terms of classification accuracy.
+In Section 2 we review the \emph{RDA} classifier from \cite{Friedman:1989tm} and discuss our generalization in Section 3.
+In Section 4 we describe four small-sample, high-dimensional microarray data sets and then demonstrate that our proposed \emph{GRDA} classifier is often superior to the competing classifiers in terms of classification accuracy.
 We then conclude with a brief discussion in Section 5.
@@
 Let $\bm I_m$ be the $m \times m$ identity matrix, and let $\bm 0_{m \times p}$ be the $m \times p$ matrix of zeros, such that $\bm 0_m$ is understood to denote $\bm 0_{m \times m}$.
+Lastly, for $c \in \mathbb{R}$, let $c^+ = 1/c$ if $c \ne 0$ and $0$ otherwise.
 
@@
 \begin{align}
-	\tilde{\bm \Sigma}_k = \alpha_k \widehat{\bm \Sigma}_k(\lambda) + \bm \Phi_k,\label{eq:grda-cov}
+	\tilde{\bm \Sigma}_k = \alpha_k \widehat{\bm \Sigma}_k(\lambda) + \gamma \bm I_p,\label{eq:grda-cov}
 \end{align}
-where $\alpha_k \ge 0$ and $\bm \Phi_k \in \mathbb{R}_{p \times p}^{\ge}$ is a diagonal matrix consisting of the eigenvalue shrinkage constants. Thus, we have the \emph{GRDA} (\emph{GRDA}) classifier
+where $\alpha_k \ge 0$ and $\gamma \ge 0$ is an eigenvalue-shrinkage constant. Thus, we have the \emph{GRDA} classifier
 \begin{align}
-	D_{GRDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\tilde{\bm\Sigma}_k^{-1}(\bm x - \xbar_k)  + \ln |\tilde{\bm\Sigma}_k|, \label{eq:grda}
+	D_{GRDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\tilde{\bm\Sigma}_k^{+}(\bm x - \xbar_k)  + \ln |\tilde{\bm\Sigma}_k|. \label{eq:grda}
 \end{align}
-The specification of $\alpha_k$ and $\bm \Phi_k$ in \eqref{eq:grda-cov} allows for a large family of eigenvalue-shrinkage methods and their relationships to be examined in depth.
+For $\gamma > 0$, $\tilde{\bm\Sigma}_k$ is nonsingular such that $\tilde{\bm\Sigma}_k^{-1}$ can be substituted for $\tilde{\bm\Sigma}_k^{+}$ in \eqref{eq:grda}.
+If $\gamma = 0$, we explicitly set $|\tilde{\bm\Sigma}_k|$ equal to the product of the positive eigenvalues of $\tilde{\bm\Sigma}_k$.
 
@@
 Then, we partition $\bm U = (\bm U_1, \bm U_2)$ such that $\bm U_1 \in \mathbb{R}_{p \times q}$ and $\bm U_2 \in \mathbb{R}_{p \times (p - q)}$.
-We also partition $\bm \Phi_k = \bm \Phi_{k1} \oplus \bm \Phi_{k2}$ such that $\bm \Phi_{k1} \in \mathbb{R}_{q \times q}$ and $\bm \Phi_{k2} \in \mathbb{R}_{(p-q) \times (p-q)}$.
 
@@
 \begin{align}
-	\tilde{\bm \Sigma}_k &= \bm U(\bm W_k \oplus \bm \Phi_{k2})\bm U',\label{eq:rda-matrix}
+	\tilde{\bm \Sigma}_k &= \bm U(\bm W_k \oplus \gamma \bm I_{p-q})\bm U',\label{eq:rda-matrix}
 \intertext{where}
-\bm W_k &= \alpha_k \{(1 - \lambda) \bm U_1' \widehat{\bm \Sigma}_k \bm U_1 + \lambda \bm D_q\} + \bm \Phi_{k1}.\label{eq:Wk}
+\bm W_k &= \alpha_k \{(1 - \lambda) \bm U_1' \widehat{\bm \Sigma}_k \bm U_1 + \lambda \bm D_q\} + \gamma \bm I_{q}.\label{eq:Wk}
 \end{align}
@@
 Hence, for $k = 1, \ldots, K$, $\bm U' \widehat{\bm \Sigma}_k \bm U = \bm U_1' \widehat{\bm \Sigma}_k \bm U_1 \oplus \bm 0_{p-q}$.
-Then, notice that $\bm U' \tilde{\bm \Sigma}_k \bm U = \alpha_k \{(1 - \lambda) \bm U' \widehat{\bm \Sigma}_k \bm U + \lambda \bm D\} + \bm U' \bm \Phi_k \bm U$.
+Then, notice that $\bm U' \tilde{\bm \Sigma}_k \bm U = \alpha_k \{(1 - \lambda) \bm U' \widehat{\bm \Sigma}_k \bm U + \lambda \bm D\} + \gamma \bm I_p$.
 Thus, \eqref{eq:rda-matrix} follows because $\bm U$ is orthogonal.
@@
 \begin{proof}
-The proof follows by setting $\bm \Phi_k = \bm 0_p$ in Lemma \ref{lemma:rda-tilde-Sigma_k}.
+The proof follows by setting $\gamma = 0$ in Lemma \ref{lemma:rda-tilde-Sigma_k}.
 \end{proof}
@@
 Now, we present our main result, where we decompose the \emph{GRDA} decision rule and show that the component requiring the largest computational costs does not contribute to the classification performed using Lemma \ref{lemma:RDA-constant-term}.
-To reduce the computational costs of the \emph{GRDA} classifier, we require that $\bm \Phi_{k2} = \bm \Phi_2$, $k = 1, \ldots, K$.
 Here, we present our main result that reduces \eqref{eq:grda} to an equivalent, more computationally efficient decision rule.
@@
 Let $\tilde{\bm \Sigma}_k$ and $\bm W_k$ be defined as in \eqref{eq:rda-matrix} and \eqref{eq:Wk}, respectively, and let $\bm U_1$ be defined as above.
-Additionally, let $\bm \Phi_{k2} = \bm \Phi_2$, $k = 1, \ldots, K$.
 Then, the \emph{GRDA} decision rule in \eqref{eq:grda} is equivalent to the reduced form
@@
 \begin{proof}
-	From \eqref{eq:rda-matrix}, we have that $\tilde{\bm \Sigma}_k^{-1} = \bm U(\bm W_k^{-1} \oplus\bm \Phi_{k2}^{-1})\bm U'$ and $|\tilde{\bm \Sigma}_k| = | \bm W_k | |\bm \Phi_{k2}|$.
+	From \eqref{eq:rda-matrix}, we have that $\tilde{\bm \Sigma}_k^{+} = \bm U(\bm W_k^{-1} \oplus \gamma^{+} \bm I_{p-q} )\bm U'$ and $|\tilde{\bm \Sigma}_k| = \gamma^{p-q}| \bm W_k |$.
 Therefore, for all $\bm x \in \mathbb{R}_{p \times 1}$, we have that
@@
 	(\bm x - \xbar_k)' \tilde{\bm\Sigma}_k^{-1}(\bm x - \xbar_k)  + \ln |\tilde{\bm\Sigma}_k| &= (\bm x - \xbar_k)' \bm U_1 \bm W_k^{-1} \bm U_1' (\bm x - \xbar_k)\\
-	&+ (\bm x - \xbar_k)' \bm U_2 \bm \Phi_{k2}^{-1} \bm U_2' (\bm x - \xbar_k) + \ln | \bm W_k | + \ln |\bm \Phi_{k2}|.
+	&+ \gamma^{+} (\bm x - \xbar_k)' \bm U_2 \bm U_2' (\bm x - \xbar_k) + \ln | \bm W_k | + (p - q) \ln \gamma.
 	\end{align*}
-From Lemma \ref{lemma:RDA-constant-term}, $\bm U_2' (\bm x - \xbar_k)$ is constant for all $k = 1, \ldots, K$. Therefore, the proof follows because $\bm \Phi_{k2} = \bm \Phi_2$.
+Therefore, the proof follows from Lemma \ref{lemma:RDA-constant-term} because $\bm U_2' (\bm x - \xbar_k)$ is constant for all $k = 1, \ldots, K$.
 \end{proof}
@@
 	\end{align}
-where $\bm Q_k = \bm I_{n_k} + \alpha_k(1- \lambda) \bm X_k \bm U_1 \bm \Gamma_k^{-1} \bm U_1' \bm X_k'$ and $\bm \Gamma_k = \alpha_k \lambda \bm D_q + \bm \Phi_{k1}$.
+where $\bm Q_k = \bm I_{n_k} + \alpha_k(1- \lambda) \bm X_k \bm U_1 \bm \Gamma_k^{-1} \bm U_1' \bm X_k'$ and $\bm \Gamma_k = \alpha_k \lambda \bm D_q + \gamma \bm I_q$.
 \end{proposition}
@@
 
-In this section, we describe four high-dimensional microarray data sets and compare our proposed \emph{GRDA} classifier with five recently proposed classifiers designed for small-sample, high-dimensional data: the \emph{DLDA} classifier from \cite{Dudoit:2002ev}, the shrunken centroids \emph{RDA} (\emph{SCRDA}) classifier from \cite{Guo:2007te}, the shrinkage-based \emph{DLDA} (\emph{SDLDA}) classifier from \cite{Pang:2009ik}, and the shrinkage-mean-based DLDA (\emph{SmDLDA}) classifier from \cite{Tong:2012hw}, and the penalized \emph{LDA} (\emph{PLDA}) classifier from \cite{Witten:2011kc}.
+In this section, we describe four high-dimensional microarray data sets and compare our proposed \emph{GRDA} classifier with five recently proposed classifiers designed for small-sample, high-dimensional data: the \emph{DLDA} classifier from \cite{Dudoit:2002ev}, the shrunken centroids \emph{RDA} (\emph{SCRDA}) classifier from \cite{Guo:2007te}, the shrinkage-based \emph{DLDA} (\emph{SDLDA}) classifier from \cite{Pang:2009ik}, the shrinkage-mean-based DLDA (\emph{SmDLDA}) classifier from \cite{Tong:2012hw}, and the penalized \emph{LDA} (\emph{PLDA}) classifier from \cite{Witten:2011kc}.
+We did not include the \emph{RDA} classifier from \cite{Friedman:1989tm} in our studies to avoid its excessive computations.
 We evaluated the classification performance of each classifier by randomly partitioning the data such that $2/3$ of the observations were allocated as training data and the remaining $1/3$ of the observations were allocated as a test data set.
@@
 
- % latex table generated in R 2.15.1 by xtable 1.7-0 package
- % Sun Mar 24 19:54:37 2013
- \begin{table}[ht]
- \begin{center}
- \begin{tabular}{lrrrl}
-   \hline
- Author & $N$ & $p$ & $K$ & Disease \\
-   \hline
- Burczynski et al. (2006) & 127 & 22283 & 3 & Crohn's Disease \\
-   Nakayama et al. (2001) & 105 & 22283 & 5 & Sarcoma \\
-   Shipp et al. (2002) & 58 & 6817 & 2 & Lymphoma \\
-   Singh et al. (2002) & 102 & 12600 & 2 & Prostate Cancer \\
-    \hline
- \end{tabular}
- \caption{Summary of high-dimensional microarray data sets.}
- \label{tab:data-summary}
- \end{center}
- \end{table}
+\[ \left[\text{Table 1 goes about here }\right]\]
 
@@
 \cite{Shipp:2002ka} have argued that more effective means are desired to determine a patient's curability.
-The authors measured 6,817 gene expression levels from 58 DLBCL patient samples with customized cDNA (‘lymphochip’) microarrays to investigate the curability of patients treated with cyclophosphamide, adriamycin, vincristine, and prednisone (CHOP)-based chemotherapy.
+The authors measured 6,817 gene expression levels from 58 DLBCL patient samples with customized cDNA (lymphochip) microarrays to investigate the curability of patients treated with cyclophosphamide, adriamycin, vincristine, and prednisone (CHOP)-based chemotherapy.
 Among the 58 DLBCL patient samples, 32 are from cured patients while 26 are from patients with fatal or refractory disease.
@@
 Our formulation of the \emph{GRDA} classifier in terms of the nonzero eigenvalues of the pooled sample covariance matrix indicates that eigenvalue-adjustment methods applied to the \emph{LDA} classifier explains the similiarities in classification accuracy of several regularization methods applied to the \emph{LDA} classifier reported by \cite{Ramey:2013ji} in a high-dimensional classification study.
+We have provided an invariance property that extends results of \cite{Ji:2008wp} and \cite{Ye:2006jm}, who have provided a similar invariance property for the \emph{LDA} decision rule.
 
@@
 \clearpage
+
+ % latex table generated in R 2.15.1 by xtable 1.7-0 package
+ % Sun Mar 24 19:54:37 2013
+ \begin{table}[ht]
+ \begin{center}
+ \begin{tabular}{lrrrl}
+   \hline
+ Author & $N$ & $p$ & $K$ & Disease \\
+   \hline
+ Burczynski et al. (2006) & 127 & 22283 & 3 & Crohn's Disease \\
+   Nakayama et al. (2001) & 105 & 22283 & 5 & Sarcoma \\
+   Shipp et al. (2002) & 58 & 6817 & 2 & Lymphoma \\
+   Singh et al. (2002) & 102 & 12600 & 2 & Prostate Cancer \\
+    \hline
+ \end{tabular}
+ \caption{Summary of high-dimensional microarray data sets.}
+ \label{tab:data-summary}
+ \end{center}
+ \end{table}
 

### 2013-03-26 sim-microarray.r
Updated the number of iterations from 500 to 1000 in microarray data simulations.

@@ -5,7 +5,7 @@ set.seed(42)
 num_cores <- 12
 
 train_pct <- 2/3
-num_iterations <- 500
+num_iterations <- 1000
 data_sets <- c("burczynski", "nakayama", "shipp", "singh")
 d <- c(100, 250, 500, 1000)
 

### 2013-03-27 reports/paper-rda.Rnw
Updated Rnw file for paper to summarize automatically the classification results in tables.

@@
 \author{John A. Ramey, Caleb K. Stein, and Dean M. Young}
+
+<<setup, include=FALSE, cache=FALSE, echo=FALSE, warning=FALSE>>=
+opts_chunk$set(fig.align = 'default', dev = 'png', message = FALSE, warning = FALSE,
+               cache = FALSE, echo = FALSE)
+
+setwd("..")
+library(ProjectTemplate)
+load.project()
+
+# Clean up results for display
+results <- na.omit(results)
+results$d <- as.factor(results$d)
+results$data_set <- as.character(results$data_set)
+
+@
 
@@
 After training each classifier from the filtered training data, we classified the filtered test data sets and computed the proportion of mislabeled test observations to estimate the classification error rate for each classifier.
-Repeating this process 500 times, we computed the average of the error-rate estimates for each classifier.
+Repeating this process 1000 times, we computed the average of the error-rate estimates for each classifier.
 We remark that our simulation design avoids the selection bias discussed in \cite{Ambroise:2002fa}.
@@
 
- % latex table generated in R 2.15.1 by xtable 1.7-0 package
- % Sun Mar 24 19:54:37 2013
- \begin{table}[ht]
- \begin{center}
- \begin{tabular}{lrrrl}
-   \hline
- Author & $N$ & $p$ & $K$ & Disease \\
-   \hline
- Burczynski et al. (2006) & 127 & 22283 & 3 & Crohn's Disease \\
-   Nakayama et al. (2001) & 105 & 22283 & 5 & Sarcoma \\
-   Shipp et al. (2002) & 58 & 6817 & 2 & Lymphoma \\
-   Singh et al. (2002) & 102 & 12600 & 2 & Prostate Cancer \\
-    \hline
- \end{tabular}
- \caption{Summary of high-dimensional microarray data sets.}
- \label{tab:data-summary}
- \end{center}
- \end{table}
+<<data_summary, results="asis">>=
+# Table containing a summary of the data sets
+data_summary <- subset(describe_data(), author %in% unique(results$data_set))
+data_summary$K[data_summary$author == "nakayama"] <- 5
+rownames(data_summary) <- NULL
+colnames(data_summary)[1:3] <- c("Author", "Year", "N")
+data_summary$Author <- Hmisc:::capitalize(as.character(data_summary$Author))
+data_summary$Author <- with(data_summary, paste0(Author, " et al. (", Year, ")"))
+data_summary <- subset(data_summary, select = -Year)
+print(xtable(data_summary, caption = "Summary of high-dimensional microarray data sets.",
+             label = "tab:data-summary", digits = 0), include.rownames = FALSE)
 
-% latex table generated in R 2.15.1 by xtable 1.7-0 package
-% Mon Mar 25 07:04:20 2013
-\begin{table}[ht]
-\begin{center}
-\begin{tabular}{lllllll}
-  \hline
-Dataset & Dudoit & GRDA & Guo & Pang & Tong & Witten \\ 
-  \hline
-Burczynski & 0.232 (0.019) & 0.179 (0.017) & 0.513 (0.022) & 0.232 (0.019) & 0.231 (0.019) & 0.232 (0.019) \\ 
-  Nakayama & 0.262 (0.020) & 0.308 (0.021) & 0.586 (0.022) & 0.262 (0.020) & 0.262 (0.020) & 0.278 (0.020) \\ 
-  Shipp & 0.136 (0.015) & 0.106 (0.014) & 0.310 (0.021) & 0.136 (0.015) & 0.136 (0.015) & 0.136 (0.015) \\ 
-  Singh & 0.111 (0.014) & 0.097 (0.013) & 0.338 (0.021) & 0.111 (0.014) & 0.111 (0.014) & 0.111 (0.014) \\ 
-   \hline
-\end{tabular}
-\caption{The average of the test error rates for the microarray data sets with $d = 100$. Approximate standard errors are given in parentheses.}
-\label{tab:results-d100}
-\end{center}
-\end{table}
+@
 
-% latex table generated in R 2.15.1 by xtable 1.7-0 package
-% Mon Mar 25 07:04:37 2013
-\begin{table}[ht]
-\begin{center}
-\begin{tabular}{lllllll}
-  \hline
-Dataset & Dudoit & GRDA & Guo & Pang & Tong & Witten \\ 
-  \hline
-Burczynski & 0.244 (0.019) & 0.163 (0.017) & 0.459 (0.022) & 0.244 (0.019) & 0.244 (0.019) & 0.243 (0.019) \\ 
-  Nakayama & 0.244 (0.019) & 0.267 (0.020) & 0.498 (0.022) & 0.244 (0.019) & 0.244 (0.019) & 0.259 (0.020) \\ 
-  Shipp & 0.153 (0.016) & 0.104 (0.014) & 0.188 (0.017) & 0.153 (0.016) & 0.153 (0.016) & 0.154 (0.016) \\ 
-  Singh & 0.135 (0.015) & 0.117 (0.014) & 0.188 (0.017) & 0.135 (0.015) & 0.135 (0.015) & 0.135 (0.015) \\ 
-   \hline
-\end{tabular}
-\caption{The average of the test error rates for the microarray data sets with $d = 250$. Approximate standard errors are given in parentheses.}
-\label{tab:results-d250}
-\end{center}
-\end{table}
+<<microarray_results, results="asis">>=
+results_summary <- ddply(results, .(data_set, d, Classifier), summarize,
+                         Avg_Error = mean(Error),
+                         SE_Error = sqrt(mean(Error) * (1 - mean(Error)) / length(Error)))
 
-% latex table generated in R 2.15.1 by xtable 1.7-0 package
-% Mon Mar 25 07:04:48 2013
-\begin{table}[ht]
-\begin{center}
-\begin{tabular}{lllllll}
-  \hline
-Dataset & Dudoit & GRDA & Guo & Pang & Tong & Witten \\ 
-  \hline
-Burczynski & 0.243 (0.019) & 0.159 (0.016) & 0.429 (0.022) & 0.243 (0.019) & 0.243 (0.019) & 0.243 (0.019) \\ 
-  Nakayama & 0.247 (0.019) & 0.248 (0.019) & 0.465 (0.022) & 0.247 (0.019) & 0.248 (0.019) & 0.261 (0.020) \\ 
-  Shipp & 0.169 (0.017) & 0.097 (0.013) & 0.119 (0.015) & 0.169 (0.017) & 0.169 (0.017) & 0.170 (0.017) \\ 
-  Singh & 0.174 (0.017) & 0.125 (0.015) & 0.155 (0.016) & 0.174 (0.017) & 0.174 (0.017) & 0.174 (0.017) \\ 
-   \hline
-\end{tabular}
-\caption{The average of the test error rates for the microarray data sets with $d = 500$. Approximate standard errors are given in parentheses.}
-\label{tab:results-d500}
-\end{center}
-\end{table}
+# Tables containing the classification results for each value of 'q'
+results_summary$data_set <- Hmisc:::capitalize(results_summary$data_set)
+results_summary$d <- as.factor(results_summary$d)
+results_summary$Avg_Error <- round_digits(results_summary$Avg_Error, 3)
+results_summary$SE_Error <- round_digits(results_summary$SE_Error, 3)
+results_summary$Error <- with(results_summary, paste0(Avg_Error, " (", SE_Error, ")"))
+results_summary <- subset(results_summary, select = -c(Avg_Error, SE_Error))
+colnames(results_summary) <- c("Dataset", "d", "Classifier", "Error")
 
-% latex table generated in R 2.15.1 by xtable 1.7-0 package
-% Mon Mar 25 07:05:01 2013
-\begin{table}[ht]
-\begin{center}
-\begin{tabular}{lllllll}
-  \hline
-Dataset & Dudoit & GRDA & Guo & Pang & Tong & Witten \\ 
-  \hline
-Burczynski & 0.249 (0.019) & 0.168 (0.017) & 0.381 (0.022) & 0.249 (0.019) & 0.249 (0.019) & 0.249 (0.019) \\ 
-  Nakayama & 0.237 (0.019) & 0.238 (0.019) & 0.460 (0.022) & 0.237 (0.019) & 0.237 (0.019) & 0.246 (0.019) \\ 
-  Shipp & 0.196 (0.018) & 0.106 (0.014) & 0.095 (0.013) & 0.196 (0.018) & 0.195 (0.018) & 0.198 (0.018) \\ 
-  Singh & 0.224 (0.019) & 0.140 (0.016) & 0.142 (0.016) & 0.224 (0.019) & 0.224 (0.019) & 0.224 (0.019) \\ 
-   \hline
-\end{tabular}
-\caption{The average of the test error rates for the microarray data sets with $d = 1000$. Approximate standard errors are given in parentheses.}
-\label{tab:results-d1000}
-\end{center}
-\end{table}
+results_tables <- tapply(seq_along(results_summary$d), results_summary$d, function(i) {
+  dcast(results_summary[i, ], Dataset ~ Classifier, value.var = "Error")
+})
+
+results_xtables <- lapply(names(results_tables), function(d) {
+  tab_caption <- "The average of the test error rates for the microarray data sets with $d = "
+  tab_caption <- paste0(tab_caption, d, "$. Approximate standard errors are given in parentheses.")
+  xtable(results_tables[[d]], caption = tab_caption, label = paste0("tab:results-d", d))
+})
+
+lapply(results_xtables, print, include.rownames = FALSE)
+
+@
 

### 2013-03-27 summary.r

@@ -1,47 +1,0 @@
-library('ProjectTemplate')
-load.project()
-
-# Remove the following data sets because there is too much separation for a good comparison
-results <- na.omit(results)
-
-results$d <- as.factor(results$d)
-results$data_set <- as.character(results$data_set)
-
-p <- ggplot(results, aes(x = d, y = Error))
-p <- p + geom_boxplot(aes(fill = Classifier))
-p + facet_wrap(~ data_set)
-
-results_summary <- ddply(results, .(data_set, d, Classifier), summarize,
-                         Avg_Error = mean(Error),
-                         SE_Error = sqrt(mean(Error) * (1 - mean(Error)) / length(Error)))
-
-# Table containing a summary of the data sets
-data_summary <- subset(describe_data(), author %in% unique(results$data_set))
-data_summary$K[data_summary$author == "nakayama"] <- 5
-rownames(data_summary) <- NULL
-colnames(data_summary)[1:3] <- c("Author", "Year", "N")
-data_summary$Author <- Hmisc:::capitalize(as.character(data_summary$Author))
-data_summary$Author <- with(data_summary, paste0(Author, " et al. (", Year, ")"))
-data_summary <- subset(data_summary, select = -Year)
-print(xtable(data_summary, caption = "Summary of high-dimensional microarray data sets.",
-             label = "tab:data-summary", digits = 0), include.rownames = FALSE)
-
-
-# Tables containing the classification results for each value of 'q'
-results_summary$data_set <- Hmisc:::capitalize(results_summary$data_set)
-results_summary$d <- as.factor(results_summary$d)
-results_summary$Avg_Error <- round_digits(results_summary$Avg_Error, 3)
-results_summary$SE_Error <- round_digits(results_summary$SE_Error, 3)
-results_summary$Error <- with(results_summary, paste0(Avg_Error, " (", SE_Error, ")"))
-results_summary <- subset(results_summary, select = -c(Avg_Error, SE_Error))
-colnames(results_summary) <- c("Dataset", "d", "Classifier", "Error")
-
-results_tables <- tapply(seq_along(results_summary$d), results_summary$d, function(i) {
-  dcast(results_summary[i, ], Dataset ~ Classifier, value.var = "Error")
-})
-
-results_xtables <- lapply(names(results_tables), function(d) {
-  tab_caption <- "The average of the test error rates for the microarray data sets with $d = "
[…]

### 2013-03-27 reports/paper-rda.Rnw
Applied edits from Dean.

@@
 
-A reliable classification of tumors from gene expression microarray data is essential for successful diagnosis and treatment of diseases such as cancer, but the analysis of microarray data often presents difficult challenges because the number of gene expression levels is typically much larger than the sample size.
-Consequently, several sparse and regularized classifiers have been proposed often with strict covariance assumptions or eigenvalue adjustments to ensure that the classifiers are well-posed.
-From \cite{Friedman:1989tm} we consider specifically the popular regularized discriminant analysis (\emph{RDA}) classifier, which requires computationally expensive matrix operations and involves a time-consuming model-selection procedure, making the \emph{RDA} classifier impractical for high-dimensional data.
-We propose a novel generalization of the \emph{RDA} (\emph{GRDA}) classifier that improves upon the interpretation of the \emph{RDA} classifier and incorporates a more general covariance-matrix regularization approach, which includes as special cases several proposed sparse and regularization methods.
-Furthermore, we address the computational shortcomings of the \emph{RDA} classifier by integrating a novel dimension-reduction technique that enables a computationally efficient model-selection procedure.
+A reliable supervised classification of tumors from gene expression microarray data is essential for successful diagnosis and treatment of diseases such as cancer, but the analysis of microarray data often presents difficult challenges because the number of gene expression levels is typically much greater than the sample size.
+Consequently, several sparse and regularized classifiers have been proposed, often with strict covariance assumptions or eigenvalue adjustments to ensure that the classifiers are well-posed.
+Here, we consider an improved version of the regularized discriminant analysis (\emph{RDA}) classifier popularized by \cite{Friedman:1989tm}, which requires computationally expensive matrix operations and involves a time-consuming model-selection procedure, making the \emph{RDA} classifier impractical for high-dimensional data.
+We propose a novel modification of the \emph{RDA} (\emph{GRDA}) classifier that incorporates a more general covariance-matrix regularization approach that includes several proposed sparse and regularization methods as special cases.
+Furthermore, we overcome the computational shortcomings of the \emph{RDA} classifier by integrating a novel dimension-reduction technique that enables a computationally efficient model-selection procedure.
 Moreover, we demonstrate with several high-dimensional microarray data sets that our proposed classifier is competitive and often superior to several recently proposed sparse and regularized classifiers in terms of classification accuracy.
@@
 We focus on the classification of tissues and other biological samples into one of $K$ disjoint, predefined cancer types, where our goal is to ascertain the class membership of a new tissue sample based on gene-expression microarray data.
-However, microarray data typically consist of a large number of gene expressions $p$ relative to the sample size $N$.
-Moreover, standard benchmark microarray data sets usually consist of between $p = 10,000$ and $55,000$ probes and fewer than $N = 100$ observations, in which case standard supervised-learning methods exhibit poor classification performance, due to the \emph{curse of dimensionality} \citep{Bellman:1961tn}.
+However, microarray data typically consist of a large number of gene expressions $p$ relative to the total sample size $N$.
+Moreover, standard benchmark microarray data sets usually consist of between $p = 10,000$ and $55,000$ probes and fewer than $N = 100$ observations, in which case standard supervised-learning methods exhibit poor classification performance because of the \emph{curse of dimensionality} \citep{Bellman:1961tn}.
 Furthermore, well-known classifiers, including linear discriminant analysis \emph{(LDA)} and quadratic discriminant analysis \emph{(QDA)}, are incalculable when $p > N$ because the sample covariance matrices employed in both classifiers are singular.
@@
 
-Alternatively, several researchers, including \cite*{Zhang:2010va}, \cite{Ye:2009gd}, \cite{Ji:2008wp}, \cite*{Guo:2007te}, \cite{Srivastava:2007ud}, \cite{Ye:2006vx}, and \cite{Ye:2005uu}, have proposed regularization methods to improve the estimation of the population covariance matrices before applying the \emph{LDA} and \emph{QDA} classifiers.
-Covariance-matrix regularization methods often are based on the well-known ridge-regression approach from \cite{Hoerl:1970cd} and adjust the eigenvalues of the sample covariance matrices to ensure positive definiteness.
+Alternatively, several researchers, including \cite*{Zhang:2010va}, \cite{Ye:2009gd}, \cite{Ji:2008wp}, \cite*{Guo:2007te}, \cite*{Srivastava:2007ud}, \cite{Ye:2006vx}, and \cite{Ye:2005uu}, have proposed regularization methods to improve the estimation of the population covariance matrices before applying the \emph{LDA} and \emph{QDA} classifiers.
+Covariance-matrix regularization methods often are based on the well-known ridge-regression approach of \cite{Hoerl:1970cd} and adjust the eigenvalues of the sample covariance matrices to ensure positive definiteness.
 For excellent overviews of the regularization problem in discriminant analysis, see \cite{Murphy:2012uq} and \cite{Mkhadri:1997gy}.
 
-Here, we consider a generalization of the popular \emph{RDA} classifier of \cite{Friedman:1989tm}, who has proposed pooling of the covariance-matrix estimators from the \emph{LDA} and \emph{QDA} classifiers.
-\cite{Friedman:1989tm} has additionally incorporated a covariance-matrix shrinkage term to ensure that the \emph{RDA} classifier is well-posed.
+Here, we propose a generalization of the popular \emph{RDA} classifier of \cite{Friedman:1989tm}, who has proposed pooling of the covariance-matrix estimators from the \emph{LDA} and \emph{QDA} classifiers and has additionally incorporated a covariance-matrix shrinkage term to ensure that the \emph{RDA} classifier is well-posed.
 \cite{Friedman:1989tm} has shown that the \emph{RDA} classifier can attain improved classification accuracy, compared to the \emph{LDA} and \emph{QDA} classifiers when $p > N$.
@@
 
-In this paper we provide a generalization of the \emph{RDA} classifier by first simplifying the pooling matrix employed in the \emph{RDA} classifier.
-We construct instead a convex combination of the covariance-matrix estimators employed in the \emph{LDA} and \emph{QDA} classifiers.
-We demonstrate that this simplification improves the interpretation of the pooling operation of the \emph{RDA} classifier.
+In this paper we modify the \emph{RDA} classifier by first simplifying the pooling matrix employed in the \emph{RDA} classifier.
+We construct a convex combination of the covariance-matrix estimators employed in the \emph{LDA} and \emph{QDA} classifiers and demonstrate that this simplification improves the interpretation of the pooling operation of the \emph{RDA} classifier.
 The \emph{LDA} classifier involves the assumption that the population covariance matrices are equal, which is often too stringent in practice \citep{Clemmensen:2011kr}.
@@
 We then introduce a more general shrinkage term to the \emph{RDA} classifier to establish our \emph{GRDA} classifier.
-Our formulation of covariance-matrix regularization employed in the \emph{GRDA} classifier includes a large family of shrinkage methods, such as the methods from \cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and a wide variety of other shrinkage methods studied by \cite{Ramey:2013ji} and \cite*{Xu:2009fl}.
 
 We show that the \emph{GRDA} classifier is invariant to any adjustments of the eigenvalues corresponding to the null space of the pooled sample covariance matrix.
-Applying a technique similar to \cite{Ye:2006tq}, we decompose the \emph{RDA} classifier and show that the sum related to the null space of the pooled sample covariance matrix can be discarded.
+Applying a reasoning similar to \cite{Ye:2006tq}, we decompose the \emph{RDA} classifier and show that the sum related to the null space of the pooled sample covariance matrix can be discarded.
 Hence, no gain in classification performance can be obtained by adjusting these zero eigenvalues.
@@
 This reduction in calculation is clearly advantageous because $n_k$ is often quite small, say, 15--20, in microarray data sets.
-Thus, through a dimension-reduction technique we demonstrate that the calculation of the \emph{RDA} decision rule involving inverses and determinants of extremely large covariance matrices is equivalent a decision rule involving inverses and determinants on much smaller matrices.
+Thus, through a dimension-reduction technique we demonstrate that the calculation of the \emph{RDA} decision rule involving inverses and determinants of extremely large covariance matrices is equivalent to a decision rule involving inverses and determinants for much lower-dimensional matrices.
 
@@
 
-\subsection{A Generalization of Regularized Discriminant Analysis}
+\section{The GRDA Classifier}
 
@@
 \end{align}
-where $\alpha_k \ge 0$ and $\gamma \ge 0$ is an eigenvalue-shrinkage constant. Thus, we have the \emph{GRDA} classifier
+where $\alpha_k \ge 0$ and $\gamma \ge 0$ is an eigenvalue-shrinkage constant. Thus, the \emph{GRDA} classifier is
 \begin{align}
@@
 
-\section{The GRDA Classifier}
-In this section, we demonstrate that the \emph{GRDA} classifier facilitates an efficient calculation of \eqref{eq:grda} and a rapid model selection of the pooling parameter $\lambda$.
-Our approach is to decompose \eqref{eq:grda} into a sum of two components, where the first summand consists of matrix operations applied to low-dimensional matrices and the second summand corresponds to the null space of $\widehat{\bm \Sigma}$ in \eqref{eq:pooled-cov}.
+Next, we demonstrate that the \emph{GRDA} classifier lends itself to an efficient calculation of \eqref{eq:grda} and a rapid model selection of the pooling parameter $\lambda$.
+We decompose \eqref{eq:grda} into a sum of two components, where the first summand consists of matrix operations applied to low-dimensional matrices and the second summand corresponds to the null space of $\widehat{\bm \Sigma}$ in \eqref{eq:pooled-cov}.
 We show that the matrix operations performed on the null space of $\widehat{\bm \Sigma}$ yield constant quadratic forms across all classes and can be omitted under mild regularity conditions.
 For $p \gg N$, the constant component involves determinants and inverses of high-dimensional matrices, and by ignoring these calculations, a substantial reduction in computational costs is achieved relative to Friedman's \emph{RDA} classifier in \eqref{eq:rda}.
-Furthermore, a byproduct of this omission is that adjustments to the associated eigenvalues have no effect on \eqref{eq:grda}.
-Lastly, we exploit the so-called Fast singular value decomposition (SVD) to construct the eigenvalue decomposition of $\widehat{\bm \Sigma}$, further reducing the computational costs of our proposed \emph{RDA} classifier.
+Furthermore, a byproduct is that adjustments to the associated eigenvalues have no effect on \eqref{eq:grda}.
+Lastly, we exploit the so-called fast singular value decomposition (SVD) to construct the eigenvalue decomposition of $\widehat{\bm \Sigma}$, further reducing the computational costs of our proposed \emph{RDA} classifier.
 
@@
 Hence, for $k = 1, \ldots, K$, $\bm U' \widehat{\bm \Sigma}_k \bm U = \bm U_1' \widehat{\bm \Sigma}_k \bm U_1 \oplus \bm 0_{p-q}$.
-Then, notice that $\bm U' \tilde{\bm \Sigma}_k \bm U = \alpha_k \{(1 - \lambda) \bm U' \widehat{\bm \Sigma}_k \bm U + \lambda \bm D\} + \gamma \bm I_p$.
-Thus, \eqref{eq:rda-matrix} follows because $\bm U$ is orthogonal.
+Thus, $\bm U' \tilde{\bm \Sigma}_k \bm U = \alpha_k \{(1 - \lambda) \bm U' \widehat{\bm \Sigma}_k \bm U + \lambda \bm D\} + \gamma \bm I_p$, and \eqref{eq:rda-matrix} follows because $\bm U$ is orthogonal.
 \end{proof}
@@
 Recall that $\bm U_2 \in \mathcal{N}(\widehat{\bm \Sigma})$, which implies that $\bm U_2' \in \mathcal{C}(\widehat{\bm \Sigma})^{\perp}$ \citep[Lemma 1.2.5]{Kollo:2005vp}.
-Now, because $\bm x_i \in \mathcal{C}(\widehat{\bm \Sigma})$, for $i = 1, \ldots, N$, $\bm U_2' \bm x_i = \bm 0_{(p-q) \times 1}$, which implies that $\bm U_2'(\xbar_k - \xbar_{k'}) = \bm 0_{(p-q) \times 1}$.
-Therefore, $\bm U_2' (\bm x - \xbar_k) = \bm U_2' (\bm x - \xbar_{k'})$.
+Now, because $\bm x_i \in \mathcal{C}(\widehat{\bm \Sigma})$, $i = 1, \ldots, N$, $\bm U_2' \bm x_i = \bm 0_{(p-q) \times 1}$, which implies that $\bm U_2' (\bm x - \xbar_k) = \bm U_2' (\bm x - \xbar_{k'})$.
 \end{proof}
 
-Now, we present our main result, where we decompose the \emph{GRDA} decision rule and show that the component requiring the largest computational costs does not contribute to the classification performed using Lemma \ref{lemma:RDA-constant-term}.
-Here, we present our main result that reduces \eqref{eq:grda} to an equivalent, more computationally efficient decision rule.
+We now present our main result, where we decompose the \emph{GRDA} decision rule and show that the component requiring the largest computational costs does not contribute to the classification of an unlabeled observation performed using Lemma \ref{lemma:RDA-constant-term}.
+Our main result reduces \eqref{eq:grda} to an equivalent, more computationally efficient decision rule.
 
@@
 Let $\tilde{\bm \Sigma}_k$ and $\bm W_k$ be defined as in \eqref{eq:rda-matrix} and \eqref{eq:Wk}, respectively, and let $\bm U_1$ be defined as above.
-Then, the \emph{GRDA} decision rule in \eqref{eq:grda} is equivalent to the reduced form
+Then, the \emph{GRDA} decision rule in \eqref{eq:grda} is equivalent to
 	\begin{align}
@@
 	\end{align*}
-Therefore, the proof follows from Lemma \ref{lemma:RDA-constant-term} because $\bm U_2' (\bm x - \xbar_k)$ is constant for all $k = 1, \ldots, K$.
+The proof follows from Lemma \ref{lemma:RDA-constant-term} because $\bm U_2' (\bm x - \xbar_k)$ is constant for all $k = 1, \ldots, K$.
 \end{proof}
@@
 
-Thus far we have shown that the \emph{GRDA} classifier in \eqref{eq:rda-decomposed} is invariant to the $\bm U_2$ term, yielding a substantial reduction in the computational costs.
+Thus far we have shown that the \emph{GRDA} classifier in \eqref{eq:rda-decomposed} is invariant to the term $\bm U_2$, thus yielding a substantial reduction in the computational costs.
 While the classifier is much improved in terms of computational performance compared to the \emph{RDA} classifier originally proposed by \cite{Friedman:1989tm}, we provide two additional results that further improve the computational performance of our \emph{GRDA} classifier.
@@
 \begin{proof}
-	To calculate $|\bm W_k|$, from \cite{Harville:2008wja} we apply Theorem 18.1.1, which states that $|\bm A + \bm B \bm T \bm C| = |\bm A| |\bm T| |\bm T^{-1} + \bm C \bm A^{-1} \bm B|$.
+	To calculate $|\bm W_k|$, we apply Theorem 18.1.1 from \cite{Harville:2008wja}, which states that $|\bm A + \bm B \bm T \bm C| = |\bm A| |\bm T| |\bm T^{-1} + \bm C \bm A^{-1} \bm B|$.
 Thus, setting $\bm A = \bm \Gamma_k$, $\bm B = \alpha_k (1 - \lambda) \bm U_1' \bm X_k'$, $\bm T = \bm I_{n_k}$, and $\bm C = \bm X_k \bm U_1$, we have $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$.
@@
 Recall that the pooled sample covariance matrix $\widehat{\bm\Sigma} = \bm X' \bm X$.
-Then, by writing the SVD of $\bm X = \bm M \bm D^{1/2} \bm U'$, we have the spectral decomposition $\widehat{\bm\Sigma} = \bm U \bm D \bm U'$, where $\bm M \in \mathbb{R}_{N \times N}$ is orthogonal, $\bm D^{1/2} \in \mathbb{R}_{N \times N}^{\ge}$ is a diagonal matrix consisting of the singular values of $\bm X$, and $\bm U \in \mathbb{R}_{p \times N}$ is orthogonal.
-Now, we see that $\bm X \bm X' = \bm M \bm D \bm M'$ is the eigenvalue decomposition of $\bm X \bm X'$, where $\bm D$ is the matrix of eigenvalues of $\bm X \bm X'$ and the columns of $\bm M$  are the corresponding eigenvectors.
+Then, because $\bm X = \bm M \bm D^{1/2} \bm U'$ by the SVD, we have the spectral decomposition $\widehat{\bm\Sigma} = \bm U \bm D \bm U'$, where $\bm M \in \mathbb{R}_{N \times N}$ is orthogonal, $\bm D^{1/2} \in \mathbb{R}_{N \times N}^{\ge}$ is a diagonal matrix consisting of the singular values of $\bm X$, and $\bm U \in \mathbb{R}_{p \times N}$ is orthogonal.
+Now, $\bm X \bm X' = \bm M \bm D \bm M'$ is the eigenvalue decomposition of $\bm X \bm X'$, where $\bm D$ is the matrix of eigenvalues of $\bm X \bm X'$ and the columns of $\bm M$  are the corresponding eigenvectors.
 Hence, we obtain $\bm M$ and $\bm D$ from an eigendecomposition of $\bm X \bm X' \in \mathbb{R}_{N \times N}$.
 Next, we compute $\bm U = \bm X' \bm M \bm D^{+/2}$, where $\bm D^{+/2} = \bm D_q^{-1/2} \oplus \bm 0_{N-q}$.
-We determine $q$ efficiently by computing the number of numerically nonzero eigenvalues present in $\bm D$ such that $q$ is the number of eigenvalues that exceeds some tolerance value, say, $1 \times 10^{-6}$.
-After calculating $q$, we retain only the first $q$ columns of $\bm U$ to extract $\bm U_1$.
+We then determine $q$ by computing the number of numerically nonzero eigenvalues present in $\bm D$, where $q$ is the number of eigenvalues that exceeds some tolerance value, say, $\epsilon = 1 \times 10^{-6}$.
+After calculating $q$, we extract $\bm U_1$ by retaining only the first $q$ columns of $\bm U$.
 
@@
 
+Although several variants of the \emph{LDA} classifier have been proposed for high-dimensional, classification, the \emph{QDA} classifier has largely been ignored, in part due to the large number of parameters that must be estimated.
+Our proposed classifier ...
+
+Our formulation of covariance-matrix regularization employed in the \emph{GRDA} classifier includes a large family of shrinkage methods, such as the methods from \cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and a wide variety of other shrinkage methods studied by \cite{Ramey:2013ji} and \cite*{Xu:2009fl}.
+
 \bibliographystyle{plainnat}

### 2013-04-13 reports/paper-rda.Rnw
Added and polished Caleb's writeup of simulation results. Also, reformatted the paragraphs through the paper using the magnificent power of emacs.

@@
 
-A reliable supervised classification of tumors from gene expression microarray data is essential for successful diagnosis and treatment of diseases such as cancer, but the analysis of microarray data often presents difficult challenges because the number of gene expression levels is typically much greater than the sample size.
-Consequently, several sparse and regularized classifiers have been proposed, often with strict covariance assumptions or eigenvalue adjustments to ensure that the classifiers are well-posed.
-Here, we consider an improved version of the regularized discriminant analysis (\emph{RDA}) classifier popularized by \cite{Friedman:1989tm}, which requires computationally expensive matrix operations and involves a time-consuming model-selection procedure, making the \emph{RDA} classifier impractical for high-dimensional data.
-We propose a novel modification of the \emph{RDA} (\emph{GRDA}) classifier that incorporates a more general covariance-matrix regularization approach that includes several proposed sparse and regularization methods as special cases.
-Furthermore, we overcome the computational shortcomings of the \emph{RDA} classifier by integrating a novel dimension-reduction technique that enables a computationally efficient model-selection procedure.
-Moreover, we demonstrate with several high-dimensional microarray data sets that our proposed classifier is competitive and often superior to several recently proposed sparse and regularized classifiers in terms of classification accuracy.
-Finally, we provide an implementation of the \emph{GRDA} classifier in the {\tt regdiscrim} R package, available on CRAN.
+A reliable supervised classification of tumors from gene expression microarray
+data is essential for successful diagnosis and treatment of diseases such as
+cancer, but the analysis of microarray data often presents difficult challenges
+because the number of gene expression levels is typically much greater than the
+sample size. Consequently, several sparse and regularized classifiers have been
+proposed, often with strict covariance assumptions or eigenvalue adjustments to
+ensure that the classifiers are well-posed. Here, we consider an improved
+version of the regularized discriminant analysis (\emph{RDA}) classifier
+popularized by \cite{Friedman:1989tm}, which requires computationally expensive
+matrix operations and involves a time-consuming model-selection procedure,
+making the \emph{RDA} classifier impractical for high-dimensional data. We
+propose a novel modification of the \emph{RDA} (\emph{GRDA}) classifier that
+incorporates a more general covariance-matrix regularization approach that
+includes several proposed sparse and regularization methods as special
+cases. Furthermore, we overcome the computational shortcomings of the \emph{RDA}
+classifier by integrating a novel dimension-reduction technique that enables a
+computationally efficient model-selection procedure. Moreover, we demonstrate
+with several high-dimensional microarray data sets that our proposed classifier
+is competitive and often superior to several recently proposed sparse and
+regularized classifiers in terms of classification accuracy. Finally, we provide
+an implementation of the \emph{GRDA} classifier in the {\tt regdiscrim} R
+package, available on CRAN.
 
@@
 
-Microarrays have enabled researchers to measure gene-expression levels of thousands of genes in a single experiment to identify a subset of genes and their expression patterns that either characterize a particular cell state or facilitate an accurate prediction of future cell state.
-We focus on the classification of tissues and other biological samples into one of $K$ disjoint, predefined cancer types, where our goal is to ascertain the class membership of a new tissue sample based on gene-expression microarray data.
-However, microarray data typically consist of a large number of gene expressions $p$ relative to the total sample size $N$.
-Moreover, standard benchmark microarray data sets usually consist of between $p = 10,000$ and $55,000$ probes and fewer than $N = 100$ observations, in which case standard supervised-learning methods exhibit poor classification performance because of the \emph{curse of dimensionality} \citep{Bellman:1961tn}.
-Furthermore, well-known classifiers, including linear discriminant analysis \emph{(LDA)} and quadratic discriminant analysis \emph{(QDA)}, are incalculable when $p > N$ because the sample covariance matrices employed in both classifiers are singular.
-As a result, many researchers, including \cite*{Mai:2012bf}, \cite*{Merchante:2012vk}, \cite*{Clemmensen:2011kr}, and \cite*{Huang:2010ju}, have introduced restrictions on the population covariance matrices in the \emph{LDA} and \emph{QDA} models to reduce number of covariance parameters that must be estimated to ensure that the resulting classifiers are calculable.
-For instance, \cite{Dudoit:2002ev} have proposed the diagonal \emph{LDA} (\emph{DLDA}) classifier, a modification of the \emph{LDA} classifier where the class covariance matrices are assumed to be diagonal.
+Microarrays have enabled researchers to measure gene-expression levels of
+thousands of genes in a single experiment to identify a subset of genes and
+their expression patterns that either characterize a particular cell state or
+facilitate an accurate prediction of future cell state.  We focus on the
+classification of tissues and other biological samples into one of $K$ disjoint,
+predefined cancer types, where our goal is to ascertain the class membership of
+a new tissue sample based on gene-expression microarray data. However,
+microarray data typically consist of a large number of gene expressions $p$
+relative to the total sample size $N$. Moreover, standard benchmark microarray
+data sets usually consist of between $p = 10,000$ and $55,000$ probes and fewer
+than $N = 100$ observations, in which case standard supervised-learning methods
+exhibit poor classification performance because of the \emph{curse of
+  dimensionality} \citep{Bellman:1961tn}. Furthermore, well-known classifiers,
+including linear discriminant analysis \emph{(LDA)} and quadratic discriminant
+analysis \emph{(QDA)}, are incalculable when $p > N$ because the sample
+covariance matrices employed in both classifiers are singular. As a result, many
+researchers, including \cite*{Mai:2012bf}, \cite*{Merchante:2012vk},
+\cite*{Clemmensen:2011kr}, and \cite*{Huang:2010ju}, have introduced
+restrictions on the population covariance matrices in the \emph{LDA} and
+\emph{QDA} models to reduce number of covariance parameters that must be
+estimated to ensure that the resulting classifiers are calculable.  For
+instance, \cite{Dudoit:2002ev} have proposed the diagonal \emph{LDA}
+(\emph{DLDA}) classifier, a modification of the \emph{LDA} classifier where the
+class covariance matrices are assumed to be diagonal.
 
-Alternatively, several researchers, including \cite*{Zhang:2010va}, \cite{Ye:2009gd}, \cite{Ji:2008wp}, \cite*{Guo:2007te}, \cite*{Srivastava:2007ud}, \cite{Ye:2006vx}, and \cite{Ye:2005uu}, have proposed regularization methods to improve the estimation of the population covariance matrices before applying the \emph{LDA} and \emph{QDA} classifiers.
-Covariance-matrix regularization methods often are based on the well-known ridge-regression approach of \cite{Hoerl:1970cd} and adjust the eigenvalues of the sample covariance matrices to ensure positive definiteness.
-For excellent overviews of the regularization problem in discriminant analysis, see \cite{Murphy:2012uq} and \cite{Mkhadri:1997gy}.
+Alternatively, several researchers, including \cite*{Zhang:2010va},
+\cite{Ye:2009gd}, \cite{Ji:2008wp}, \cite*{Guo:2007te},
+\cite*{Srivastava:2007ud}, \cite{Ye:2006vx}, and \cite{Ye:2005uu}, have proposed
+regularization methods to improve the estimation of the population covariance
+matrices before applying the \emph{LDA} and \emph{QDA}
+classifiers. Covariance-matrix regularization methods often are based on the
+well-known ridge-regression approach of \cite{Hoerl:1970cd} and adjust the
+eigenvalues of the sample covariance matrices to ensure positive
+definiteness. For excellent overviews of the regularization problem in
+discriminant analysis, see \cite{Murphy:2012uq} and \cite{Mkhadri:1997gy}.
 
-Here, we propose a generalization of the popular \emph{RDA} classifier of \cite{Friedman:1989tm}, who has proposed pooling of the covariance-matrix estimators from the \emph{LDA} and \emph{QDA} classifiers and has additionally incorporated a covariance-matrix shrinkage term to ensure that the \emph{RDA} classifier is well-posed.
-\cite{Friedman:1989tm} has shown that the \emph{RDA} classifier can attain improved classification accuracy, compared to the \emph{LDA} and \emph{QDA} classifiers when $p > N$.
-However, in its original form, the \emph{RDA} classifier is impractical when $p \gg N$ because the \emph{RDA} decision rule requires the calculation of the inverses and determinants of the $p \times p$ covariance-matrix estimators for each class.
-Furthermore, the \emph{RDA} classifier has two tuning parameters that are selected via cross-validation, which adds a substantial computational burden to the model-selection process.
-The \emph{RDA} classifier is ill-advised on modern data sets because the model selection requires $O(Kp^3)$ calculations for each tuning-parameter pair.
+Here, we propose a generalization of the popular \emph{RDA} classifier of
+\cite{Friedman:1989tm}, who has proposed pooling of the covariance-matrix
+estimators from the \emph{LDA} and \emph{QDA} classifiers and has additionally
+incorporated a covariance-matrix shrinkage term to ensure that the \emph{RDA}
+classifier is well-posed. \cite{Friedman:1989tm} has shown that the \emph{RDA}
+classifier can attain improved classification accuracy, compared to the
+\emph{LDA} and \emph{QDA} classifiers when $p > N$. However, in its original
+form, the \emph{RDA} classifier is impractical when $p \gg N$ because the
+\emph{RDA} decision rule requires the calculation of the inverses and
+determinants of the $p \times p$ covariance-matrix estimators for each
+class. Furthermore, the \emph{RDA} classifier has two tuning parameters that are
+selected via cross-validation, which adds a substantial computational burden to
+the model-selection process. The \emph{RDA} classifier is ill-advised on modern
+data sets because the model selection requires $O(Kp^3)$ calculations for each
+tuning-parameter pair.
 
-In this paper we modify the \emph{RDA} classifier by first simplifying the pooling matrix employed in the \emph{RDA} classifier.
-We construct a convex combination of the covariance-matrix estimators employed in the \emph{LDA} and \emph{QDA} classifiers and demonstrate that this simplification improves the interpretation of the pooling operation of the \emph{RDA} classifier.
-The \emph{LDA} classifier involves the assumption that the population covariance matrices are equal, which is often too stringent in practice \citep{Clemmensen:2011kr}.
-The \emph{GRDA} classifier allows us to relax the linearity assumption by including unequal covariance matrices in the model.
-The \emph{GRDA} classifier is then a compromise between the \emph{LDA} and \emph{QDA} classifiers,  so that the \emph{GRDA} classifier includes both the \emph{LDA} and \emph{QDA} classifiers as special cases.
-We then introduce a more general shrinkage term to the \emph{RDA} classifier to establish our \emph{GRDA} classifier.
+In this paper we modify the \emph{RDA} classifier by first simplifying the
+pooling matrix employed in the \emph{RDA} classifier. We construct a convex
+combination of the covariance-matrix estimators employed in the \emph{LDA} and
+\emph{QDA} classifiers and demonstrate that this simplification improves the
+interpretation of the pooling operation of the \emph{RDA} classifier. The
+\emph{LDA} classifier involves the assumption that the population covariance
+matrices are equal, which is often too stringent in practice
+\citep{Clemmensen:2011kr}. The \emph{GRDA} classifier allows us to relax the
+linearity assumption by including unequal covariance matrices in the model. The
+\emph{GRDA} classifier is then a compromise between the \emph{LDA} and
+\emph{QDA} classifiers, so that the \emph{GRDA} classifier includes both the
+\emph{LDA} and \emph{QDA} classifiers as special cases. We then introduce a more
+general shrinkage term to the \emph{RDA} classifier to establish our \emph{GRDA}
+classifier.
 
-We show that the \emph{GRDA} classifier is invariant to any adjustments of the eigenvalues corresponding to the null space of the pooled sample covariance matrix.
-Applying a reasoning similar to \cite{Ye:2006tq}, we decompose the \emph{RDA} classifier and show that the sum related to the null space of the pooled sample covariance matrix can be discarded.
-Hence, no gain in classification performance can be obtained by adjusting these zero eigenvalues.
-This invariance property addresses the computational shortcomings of the \emph{RDA} classifier and provides a substantial increase in the computational efficiency of the \emph{RDA} classifier and its model selection.
-In fact, we show that the inverse and determinants can be reduced to calculations of $n_k \times n_k$ matrices, where $n_k$ is the sample size of the $k$th class, $k = 1, \ldots, K$.
-This reduction in calculation is clearly advantageous because $n_k$ is often quite small, say, 15--20, in microarray data sets.
-Thus, through a dimension-reduction technique we demonstrate that the calculation of the \emph{RDA} decision rule involving inverses and determinants of extremely large covariance matrices is equivalent to a decision rule involving inverses and determinants for much lower-dimensional matrices.
+We show that the \emph{GRDA} classifier is invariant to any adjustments of the
+eigenvalues corresponding to the null space of the pooled sample covariance
+matrix. Applying a reasoning similar to \cite{Ye:2006tq}, we decompose the
+\emph{RDA} classifier and show that the sum related to the null space of the
+pooled sample covariance matrix can be discarded. Hence, no gain in
+classification performance can be obtained by adjusting these zero
+eigenvalues. This invariance property addresses the computational shortcomings
+of the \emph{RDA} classifier and provides a substantial increase in the
+computational efficiency of the \emph{RDA} classifier and its model
+selection. In fact, we show that the inverse and determinants can be reduced to
+calculations of $n_k \times n_k$ matrices, where $n_k$ is the sample size of the
+$k$th class, $k = 1, \ldots, K$. This reduction in calculation is clearly
+advantageous because $n_k$ is often quite small, say, 15--20, in microarray data
+sets. Thus, through a dimension-reduction technique we demonstrate that the
+calculation of the \emph{RDA} decision rule involving inverses and determinants
+of extremely large covariance matrices is equivalent to a decision rule
+involving inverses and determinants for much lower-dimensional matrices.
 
-In addition to the improvement in computational performance, we demonstrate that our \emph{GRDA} classifier is competitive with and often superior to several recent classifiers designed for small-sample, high-dimensional data.
-In particular, we compare our \emph{GRDA} classifier with the \emph{DLDA} classifier from \cite{Dudoit:2002ev} and the recently proposed classifiers from \cite{Guo:2007te}, \cite{Pang:2009ik}, \cite{Tong:2012hw}, and \cite{Witten:2011kc}.
-We conduct our comparison of the classification accuracy of the competing classifiers using four small-sample, high-dimensional microarray data sets.
+In addition to the improvement in computational performance, we demonstrate that
+our \emph{GRDA} classifier is competitive with and often superior to several
+recent classifiers designed for small-sample, high-dimensional data. In
+particular, we compare our \emph{GRDA} classifier with the \emph{DLDA}
+classifier from \cite{Dudoit:2002ev} and the recently proposed classifiers from
+\cite{Guo:2007te}, \cite{Pang:2009ik}, \cite{Tong:2012hw}, and
+\cite{Witten:2011kc}. We conduct our comparison of the classification accuracy
+of the competing classifiers using four small-sample, high-dimensional
+microarray data sets.
 
-We have organized the remainder of the paper as follows.
-In Section 2 we review the \emph{RDA} classifier from \cite{Friedman:1989tm} and discuss our generalization in Section 3.
-In Section 4 we describe four small-sample, high-dimensional microarray data sets and then demonstrate that our proposed \emph{GRDA} classifier is often superior to the competing classifiers in terms of classification accuracy.
-We then conclude with a brief discussion in Section 5.
+We have organized the remainder of the paper as follows. In Section 2 we review
+the \emph{RDA} classifier from \cite{Friedman:1989tm} and discuss our
+generalization in Section 3. In Section 4 we describe four small-sample,
+high-dimensional microarray data sets and then demonstrate that our proposed
+\emph{GRDA} classifier is often superior to the competing classifiers in terms
+of classification accuracy. We then conclude with a brief discussion in Section
+5.
 
@@
 
-In discriminant analysis we wish to assign correctly an unlabeled $p$-dimensional observation vector $\bm x$ to one of $K$ unique, known classes (or populations) by constructing a classifier from $n$ training observations that can accurately predict the class membership of $\bm x$.
-Let $\bm x_i = (x_{i1}, \ldots, x_{ip}) \in \mathbb{R}_{p \times 1}$ be the $i$th observation $(i = 1, \ldots, N)$ with true, unique membership $y_i \in \{\omega_1, \ldots, \omega_K\}$, where $\mathbb{R}_{a \times b}$ denotes the matrix space of all $a \times b$ matrices over the real field $\mathbb{R}$.
-We assume that $(\bm x_i, y_i)$ is a realization from a mixture distribution $p(\bm x) = \sum_{k=1}^K p(\bm x | \omega_k) p(\omega_k)$, where $p(\bm x | \omega_k)$ is the probability density function (PDF) of the $k$th class, $p(\omega_k)$ is prior probability of class membership of the $k$th class.
+In discriminant analysis we wish to assign correctly an unlabeled
+$p$-dimensional observation vector $\bm x$ to one of $K$ unique, known classes
+(or populations) by constructing a classifier from $n$ training observations
+that can accurately predict the class membership of $\bm x$. Let $\bm x_i =
+(x_{i1}, \ldots, x_{ip}) \in \mathbb{R}_{p \times 1}$ be the $i$th observation
+$(i = 1, \ldots, N)$ with true, unique membership $y_i \in \{\omega_1, \ldots,
+\omega_K\}$, where $\mathbb{R}_{a \times b}$ denotes the matrix space of all $a
+\times b$ matrices over the real field $\mathbb{R}$. We assume that $(\bm x_i,
+y_i)$ is a realization from a mixture distribution $p(\bm x) = \sum_{k=1}^K
+p(\bm x | \omega_k) p(\omega_k)$, where $p(\bm x | \omega_k)$ is the probability
+density function (PDF) of the $k$th class, $p(\omega_k)$ is prior probability of
+class membership of the $k$th class.
 
-The \emph{QDA} classifier is the optimal Bayesian decision rule with respect to a $0-1$ loss function when $p(\bm x | \omega_k)$ is the PDF of the multivariate normal distribution with known mean vector  $\bm\mu_k \in \mathbb{R}_{p \times 1}$ and known covariance matrix $\bm\Sigma_k \in \mathbb{R}_{p \times p}^{>}$, $k = 1, 2, \ldots, K$, where $\mathbb{R}_{p \times p}^{>}$ denotes the cone of real $p \times p$ positive-definite matrices.
-Because the parameters are typically unknown, we estimate the unknown parameters $\bm \mu_k$ and $\bm\Sigma_k$ with their maximum likelihood estimators (MLEs) and substitute the MLEs into the \emph{QDA} classifier.
-Assuming that the prior probabilities of class membership $p(\omega_k)$ are equal for $k = 1, \ldots, K$, we assign an unlabeled observation $\bm x$ to class $\omega_k$ with the sample \emph{QDA} classifier
+The \emph{QDA} classifier is the optimal Bayesian decision rule with respect to
+a $0-1$ loss function when $p(\bm x | \omega_k)$ is the PDF of the multivariate
+normal distribution with known mean vector $\bm\mu_k \in \mathbb{R}_{p \times
+  1}$ and known covariance matrix $\bm\Sigma_k \in \mathbb{R}_{p \times p}^{>}$,
+$k = 1, 2, \ldots, K$, where $\mathbb{R}_{p \times p}^{>}$ denotes the cone of
+real $p \times p$ positive-definite matrices. Because the parameters are
+typically unknown, we estimate the unknown parameters $\bm \mu_k$ and
+$\bm\Sigma_k$ with their maximum likelihood estimators (MLEs) and substitute the
+MLEs into the \emph{QDA} classifier. Assuming that the prior probabilities of
+class membership $p(\omega_k)$ are equal for $k = 1, \ldots, K$, we assign an
+unlabeled observation $\bm x$ to class $\omega_k$ with the sample \emph{QDA}
+classifier
 \begin{align}
@@
 \end{align}
-where $\xbar_k$ and $\widehat{\bm\Sigma}_k$ are the MLEs for $\bm \mu_k$ and $\bm \Sigma_k$, respectively.
-If we assume further that the population covariance matrices are equal for each class (i.e., $\bm\Sigma_k = \bm\Sigma$ for all $k$), then the pooled sample covariance matrix $\widehat{\bm\Sigma}$ is substituted for $\widehat{\bm \Sigma}_k$ in \eqref{eq:qda}, where
+where $\xbar_k$ and $\widehat{\bm\Sigma}_k$ are the MLEs for $\bm \mu_k$ and
+$\bm \Sigma_k$, respectively. If we assume further that the population
+covariance matrices are equal for each class (i.e., $\bm\Sigma_k = \bm\Sigma$
+for all $k$), then the pooled sample covariance matrix $\widehat{\bm\Sigma}$ is
+substituted for $\widehat{\bm \Sigma}_k$ in \eqref{eq:qda}, where
 \begin{align}
@@
 \end{align}
-is the MLE for $\bm \Sigma$. In this case \eqref{eq:qda} reduces to Fisher's \emph{LDA} classifier, and the log-determinant is omitted because it is constant across all classes.
+is the MLE for $\bm \Sigma$. In this case \eqref{eq:qda} reduces to Fisher's
+\emph{LDA} classifier, and the log-determinant is omitted because it is constant
+across all classes.
 
-We also require the following additional notation throughout the document.
-Let $\bm A \oplus \bm B$ denote the direct sum of $\bm A \in \mathbb{R}_{r \times r}$ and $\bm B \in \mathbb{R}_{s \times s}$ \citep[Chapter 1]{Lutkepohl:1996uz}.
-Let $\bm A^+$ and $\mathcal{N}(\bm A) = \{\bm z \in \mathbb{R}_{p \times 1} : \bm A \bm z = \bm 0\}$ denote the Moore-Penrose pseudoinverse and null space of $\bm A \in \mathbb{R}_{m \times p}$, respectively.
-Let $V^{\perp} = \{\bm y \in \mathbb{R}_{p \times 1} : \bm y'\bm v = 0\ \forall \bm v \in V \}$ denote the orthogonal complement of a vector space $V \subset \mathbb{R}_{p \times 1}$.
-Let $\bm I_m$ be the $m \times m$ identity matrix, and let $\bm 0_{m \times p}$ be the $m \times p$ matrix of zeros, such that $\bm 0_m$ is understood to denote $\bm 0_{m \times m}$.
-Lastly, for $c \in \mathbb{R}$, let $c^+ = 1/c$ if $c \ne 0$ and $0$ otherwise.
+We also require the following additional notation throughout the document. Let
+$\bm A \oplus \bm B$ denote the direct sum of $\bm A \in \mathbb{R}_{r \times
+  r}$ and $\bm B \in \mathbb{R}_{s \times s}$ \citep[Chapter
+  1]{Lutkepohl:1996uz}. Let $\bm A^+$ and $\mathcal{N}(\bm A) = \{\bm z \in
+\mathbb{R}_{p \times 1} : \bm A \bm z = \bm 0\}$ denote the Moore-Penrose
+pseudoinverse and null space of $\bm A \in \mathbb{R}_{m \times p}$,
+respectively. Let $V^{\perp} = \{\bm y \in \mathbb{R}_{p \times 1} : \bm y'\bm v
+= 0\ \forall \bm v \in V \}$ denote the orthogonal complement of a vector space
+$V \subset \mathbb{R}_{p \times 1}$. Let $\bm I_m$ be the $m \times m$ identity
+matrix, and let $\bm 0_{m \times p}$ be the $m \times p$ matrix of zeros, such
+that $\bm 0_m$ is understood to denote $\bm 0_{m \times m}$. Lastly, for $c \in
+\mathbb{R}$, let $c^+ = 1/c$ if $c \ne 0$ and $0$ otherwise.
 
@@
 
-The smallest eigenvalues and the directions associated with their eigenvectors highly influence \eqref{eq:qda}.
-In fact, the eigenvalues of $\widehat{\bm \Sigma}_k$ are well-known to be biased if $p \ge n_k$ such that the smallest eigenvalues are underestimated \citep{Seber:2004uh}, Moreover, if $p > n_k$, then rank$(\widehat{\bm \Sigma}_k) \le n_k$, which implies that at least $p - n_k$ eigenvalues of $\widehat{\bm \Sigma}_k$ are 0.
-Furthermore, although more feature information is available to discriminate among the $K$ classes, if $p > n_k$, \eqref{eq:qda} is incalculable because $\widehat{\bm \Sigma}_k^{-1}$ does not exist.
+The smallest eigenvalues and the directions associated with their eigenvectors
+highly influence \eqref{eq:qda}. In fact, the eigenvalues of $\widehat{\bm
+  \Sigma}_k$ are well-known to be biased if $p \ge n_k$ such that the smallest
+eigenvalues are underestimated \citep{Seber:2004uh}, Moreover, if $p > n_k$,
+then rank$(\widehat{\bm \Sigma}_k) \le n_k$, which implies that at least $p -
+n_k$ eigenvalues of $\widehat{\bm \Sigma}_k$ are 0. Furthermore, although more
+feature information is available to discriminate among the $K$ classes, if $p >
+n_k$, \eqref{eq:qda} is incalculable because $\widehat{\bm \Sigma}_k^{-1}$ does
+not exist.
 
-Several regularization methods, such as the methods considered by \citet*{Guo:2007te}, \cite{Mkhadri:1995jp}, and \citet*{Xu:2009fl}, have been proposed in the literature to adjust the eigenvalues of $\widehat{\bm \Sigma}_k$ so that \eqref{eq:qda} is calculable and has reduced variability in the standard bias-variance tradeoff.
-A common form of the covariance-matrix regularization applies a shrinkage factor $\gamma_k > 0$, so that
+Several regularization methods, such as the methods considered by
+\citet*{Guo:2007te}, \cite{Mkhadri:1995jp}, and \citet*{Xu:2009fl}, have been
+proposed in the literature to adjust the eigenvalues of $\widehat{\bm \Sigma}_k$
+so that \eqref{eq:qda} is calculable and has reduced variability in the standard
+bias-variance tradeoff.  A common form of the covariance-matrix regularization
+applies a shrinkage factor $\gamma_k > 0$, so that
 \begin{align}
@@
 \end{align}
-similar to a method employed in ridge regression \citep{Hoerl:1970cd}.
-Equation \eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance matrix $\widehat{\bm\Sigma}_k$ towards $\bm I_p$, thereby increasing the eigenvalues of $\widehat{\bm\Sigma}_k$ by $\gamma_k$.
-Specifically, the zero eigenvalues are replaced with $\gamma_k$, so that \eqref{eq:ridge-estimator} is positive definite.
-For additional covariance-matrix regularization methods, see \cite{Ramey:2013ji}, \cite{Xu:2009fl}, and \cite{Ye:2009gd}.
+similar to a method employed in ridge regression \citep{Hoerl:1970cd}. Equation
+\eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance
+matrix $\widehat{\bm\Sigma}_k$ towards $\bm I_p$, thereby increasing the
+eigenvalues of $\widehat{\bm\Sigma}_k$ by $\gamma_k$. Specifically, the zero
+eigenvalues are replaced with $\gamma_k$, so that \eqref{eq:ridge-estimator} is
+positive definite. For additional covariance-matrix regularization methods, see
+\cite{Ramey:2013ji}, \cite{Xu:2009fl}, and \cite{Ye:2009gd}.
 
@@
 
-\cite{Friedman:1989tm} has proposed the \emph{RDA} classifier by incorporating a weighted average of the sample covariance matrix $\widehat{\bm \Sigma}_k$ for class $\omega_k$ and the pooled sample covariance matrix $\widehat{\bm\Sigma}$ to estimate the covariance matrix for class $\omega_k$ with
+\cite{Friedman:1989tm} has proposed the \emph{RDA} classifier by incorporating a
+weighted average of the sample covariance matrix $\widehat{\bm \Sigma}_k$ for
+class $\omega_k$ and the pooled sample covariance matrix $\widehat{\bm\Sigma}$
+to estimate the covariance matrix for class $\omega_k$ with
 \begin{align}
@@
 \end{align}
-where $\lambda \in [0, 1]$, $n_k(\lambda) = (1 - \lambda) n_k + \lambda N$, $\bm S_k = n_k \widehat{\bm\Sigma}_k$, $\bm S = \sum_{k=1}^K \bm S_k$, and $\bm S_k(\lambda) = (1 - \lambda) \bm S_k + \lambda \bm S$.
-We can interpret \eqref{eq:sig-lambda} as a covariance matrix estimator for class $\omega_k$ that borrows from $\widehat{\bm\Sigma}$ in \eqref{eq:pooled-cov} to better estimate $\bm \Sigma_k$.
-If $\lambda = 0$, \eqref{eq:sig-lambda} is equal to $\widehat{\bm\Sigma}_k$ used in the \emph{QDA} classifier in \eqref{eq:qda}.
-Alternatively, if $\lambda = 1$, \eqref{eq:sig-lambda} reduces to \eqref{eq:pooled-cov} used in the \emph{LDA} classifier.
+where $\lambda \in [0, 1]$, $n_k(\lambda) = (1 - \lambda) n_k + \lambda N$, $\bm
+S_k = n_k \widehat{\bm\Sigma}_k$, $\bm S = \sum_{k=1}^K \bm S_k$, and $\bm
+S_k(\lambda) = (1 - \lambda) \bm S_k + \lambda \bm S$. We can interpret
+\eqref{eq:sig-lambda} as a covariance matrix estimator for class $\omega_k$ that
+borrows from $\widehat{\bm\Sigma}$ in \eqref{eq:pooled-cov} to better estimate
+$\bm \Sigma_k$. If $\lambda = 0$, \eqref{eq:sig-lambda} is equal to
+$\widehat{\bm\Sigma}_k$ used in the \emph{QDA} classifier in
+\eqref{eq:qda}. Alternatively, if $\lambda = 1$, \eqref{eq:sig-lambda} reduces
+to \eqref{eq:pooled-cov} used in the \emph{LDA} classifier.
 
-To further improve the estimation of $\bm \Sigma_k$ and to stabilize the inverse of \eqref{eq:sig-lambda}, \cite{Friedman:1989tm} has proposed the biased \emph{RDA} covariance-matrix estimator
+To further improve the estimation of $\bm \Sigma_k$ and to stabilize the inverse
+of \eqref{eq:sig-lambda}, \cite{Friedman:1989tm} has proposed the biased
+\emph{RDA} covariance-matrix estimator
 \begin{align}
@@
 \end{align}
-for class $\omega_k$, where $\gamma \in [0, 1]$ is a regularization parameter that controls the shrinkage of \eqref{eq:sig-rda}  towards $\bm I_p$, which is weighted by the the average of the eigenvalues of \eqref{eq:sig-lambda}.
-Thus, the \emph{pooling} parameter $\lambda$ controls the amount that we borrow from $\widehat{\bm\Sigma}$ to estimate $\bm \Sigma_k$, and the \emph{shrinkage} parameter $\gamma$ determines the amount of shrinkage applied.
+for class $\omega_k$, where $\gamma \in [0, 1]$ is a regularization parameter
+that controls the shrinkage of \eqref{eq:sig-rda} towards $\bm I_p$, which is
+weighted by the the average of the eigenvalues of \eqref{eq:sig-lambda}. Thus,
+the \emph{pooling} parameter $\lambda$ controls the amount that we borrow from
+$\widehat{\bm\Sigma}$ to estimate $\bm \Sigma_k$, and the \emph{shrinkage}
+parameter $\gamma$ determines the amount of shrinkage applied.
 
-\cite{Friedman:1989tm} substituted \eqref{eq:sig-rda} into \eqref{eq:qda} to obtain the \emph{RDA} classifier
+\cite{Friedman:1989tm} substituted \eqref{eq:sig-rda} into \eqref{eq:qda} to
+obtain the \emph{RDA} classifier
 \begin{align}
@@
 
-Here, we define our generalization of the \emph{RDA} classifier by first simplifying the formulation of $\widehat{\bm \Sigma}_k(\lambda)$ in \eqref{eq:sig-lambda} and demonstrating its clear interpretation as a linear combination of the crossproducts of the training observations centered by their respective class sample means.
-We define
+Here, we define our generalization of the \emph{RDA} classifier by first
+simplifying the formulation of $\widehat{\bm \Sigma}_k(\lambda)$ in
+\eqref{eq:sig-lambda} and demonstrating its clear interpretation as a linear
+combination of the crossproducts of the training observations centered by their
+respective class sample means. We define
 \begin{align}
@@
 \end{align}
-as a convex combination of the MLEs of $\bm \Sigma_k$ and $\bm \Sigma$, where $\lambda \in [0, 1]$ is the \emph{pooling} parameter.
-Thus, we avoid the convex combination of $n_k$ and $N$ as defined by \cite{Friedman:1989tm}.
-By rewriting \eqref{eq:sig-lambda-alternative} in terms of the observations $\bm x_i$, $i = 1, \ldots, N$, each centered by its class sample mean, we attain a clear interpretation of $\widehat{\bm\Sigma}_k(\lambda)$.
-We have that
+as a convex combination of the MLEs of $\bm \Sigma_k$ and $\bm \Sigma$, where
+$\lambda \in [0, 1]$ is the \emph{pooling} parameter. Thus, we avoid the convex
+combination of $n_k$ and $N$ as defined by \cite{Friedman:1989tm}. By rewriting
+\eqref{eq:sig-lambda-alternative} in terms of the observations $\bm x_i$, $i =
+1, \ldots, N$, each centered by its class sample mean, we attain a clear
+interpretation of $\widehat{\bm\Sigma}_k(\lambda)$. We have that
 \begin{align}
@@
 where $c_{ik}(\lambda) = \lambda N^{-1} + (1 - \lambda)n_k^{-1}I(y_i = k)$.
-From \eqref{eq:sig-lambda-alternative2}, $\lambda$ weights the contribution of each of the $N$ observations to estimating $\bm \Sigma_k$.
-That is, for $0 < \lambda \le 1$, we estimate $\bm \Sigma_k$ with $N$ rather than $n_k$ observations.
-Below, we will show that the pooling operation is advantageous in increasing the rank of each $\widehat{\bm\Sigma}_k(\lambda)$ from rank$(\widehat{\bm\Sigma}_k)$ to rank$(\widehat{\bm\Sigma})$ for $0 < \lambda \le 1$.
-Notice that if $\lambda  = 0$, then the observations from the remaining $K - 1$ classes do not contribute to the estimation of $\bm \Sigma_k$, corresponding to $\widehat{\bm \Sigma}_k$.
-Furthermore, if $\lambda = 1$, the weights in \eqref{eq:sig-lambda-alternative2} reduce to $1/N$, corresponding to $\widehat{\bm\Sigma}$.
-For brevity, for $\lambda = 1$, we define $\bm X = [\sqrt{c_{1k}(1)} \bm x_1', \ldots,  \sqrt{c_{Nk}(1)} \bm x_N']'$ such that $\widehat{\bm \Sigma} = \bm X' \bm X$.
-Similarly, for $\lambda = 0$, we define $\bm X_k = [\sqrt{c_{1k}(0)} \bm x_1', \ldots,  \sqrt{c_{Nk}(0)} \bm x_N']'$ such that $\widehat{\bm \Sigma}_k = \bm X_k' \bm X_k$, $k = 1, \ldots, K$.
+From \eqref{eq:sig-lambda-alternative2}, $\lambda$ weights the contribution of
+each of the $N$ observations to estimating $\bm \Sigma_k$. That is, for $0 <
+\lambda \le 1$, we estimate $\bm \Sigma_k$ with $N$ rather than $n_k$
+observations. Below, we will show that the pooling operation is advantageous in
+increasing the rank of each $\widehat{\bm\Sigma}_k(\lambda)$ from
+rank$(\widehat{\bm\Sigma}_k)$ to rank$(\widehat{\bm\Sigma})$ for $0 < \lambda
+\le 1$. Notice that if $\lambda = 0$, then the observations from the remaining
+$K - 1$ classes do not contribute to the estimation of $\bm \Sigma_k$,
+corresponding to $\widehat{\bm \Sigma}_k$. Furthermore, if $\lambda = 1$, the
+weights in \eqref{eq:sig-lambda-alternative2} reduce to $1/N$, corresponding to
+$\widehat{\bm\Sigma}$. For brevity, for $\lambda = 1$, we define $\bm X =
+[\sqrt{c_{1k}(1)} \bm x_1', \ldots, \sqrt{c_{Nk}(1)} \bm x_N']'$ such that
+$\widehat{\bm \Sigma} = \bm X' \bm X$. Similarly, for $\lambda = 0$, we define
+$\bm X_k = [\sqrt{c_{1k}(0)} \bm x_1', \ldots, \sqrt{c_{Nk}(0)} \bm x_N']'$ such
+that $\widehat{\bm \Sigma}_k = \bm X_k' \bm X_k$, $k = 1, \ldots, K$.
 
-As we have discussed, several eigenvalue adjustment methods have been proposed to increase eigenvalues (approximately) equal to 0.
-Here, we define a more general eigenvalue adjustment of \eqref{eq:sig-lambda-alternative} as
+As we have discussed, several eigenvalue adjustment methods have been proposed
+to increase eigenvalues (approximately) equal to 0. Here, we define a more
+general eigenvalue adjustment of \eqref{eq:sig-lambda-alternative} as
 \begin{align}
@@
 \end{align}
-where $\alpha_k \ge 0$ and $\gamma \ge 0$ is an eigenvalue-shrinkage constant. Thus, the \emph{GRDA} classifier is
+where $\alpha_k \ge 0$ and $\gamma \ge 0$ is an eigenvalue-shrinkage
+constant. Thus, the \emph{GRDA} classifier is
 \begin{align}
@@
 \end{align}
-For $\gamma > 0$, $\tilde{\bm\Sigma}_k$ is nonsingular such that $\tilde{\bm\Sigma}_k^{-1}$ can be substituted for $\tilde{\bm\Sigma}_k^{+}$ in \eqref{eq:grda}.
-If $\gamma = 0$, we explicitly set $|\tilde{\bm\Sigma}_k|$ equal to the product of the positive eigenvalues of $\tilde{\bm\Sigma}_k$.
+For $\gamma > 0$, $\tilde{\bm\Sigma}_k$ is nonsingular such that
+$\tilde{\bm\Sigma}_k^{-1}$ can be substituted for $\tilde{\bm\Sigma}_k^{+}$ in
+\eqref{eq:grda}. If $\gamma = 0$, we explicitly set $|\tilde{\bm\Sigma}_k|$
+equal to the product of the positive eigenvalues of $\tilde{\bm\Sigma}_k$.
 
-Next, we demonstrate that the \emph{GRDA} classifier lends itself to an efficient calculation of \eqref{eq:grda} and a rapid model selection of the pooling parameter $\lambda$.
-We decompose \eqref{eq:grda} into a sum of two components, where the first summand consists of matrix operations applied to low-dimensional matrices and the second summand corresponds to the null space of $\widehat{\bm \Sigma}$ in \eqref{eq:pooled-cov}.
-We show that the matrix operations performed on the null space of $\widehat{\bm \Sigma}$ yield constant quadratic forms across all classes and can be omitted under mild regularity conditions.
-For $p \gg N$, the constant component involves determinants and inverses of high-dimensional matrices, and by ignoring these calculations, a substantial reduction in computational costs is achieved relative to Friedman's \emph{RDA} classifier in \eqref{eq:rda}.
-Furthermore, a byproduct is that adjustments to the associated eigenvalues have no effect on \eqref{eq:grda}.
-Lastly, we exploit the so-called fast singular value decomposition (SVD) to construct the eigenvalue decomposition of $\widehat{\bm \Sigma}$, further reducing the computational costs of our proposed \emph{RDA} classifier.
+Next, we demonstrate that the \emph{GRDA} classifier lends itself to an
+efficient calculation of \eqref{eq:grda} and a rapid model selection of the
+pooling parameter $\lambda$. We decompose \eqref{eq:grda} into a sum of two
+components, where the first summand consists of matrix operations applied to
+low-dimensional matrices and the second summand corresponds to the null space of
+$\widehat{\bm \Sigma}$ in \eqref{eq:pooled-cov}. We show that the matrix
+operations performed on the null space of $\widehat{\bm \Sigma}$ yield constant
+quadratic forms across all classes and can be omitted under mild regularity
+conditions. For $p \gg N$, the constant component involves determinants and
+inverses of high-dimensional matrices, and by ignoring these calculations, a
+substantial reduction in computational costs is achieved relative to Friedman's
+\emph{RDA} classifier in \eqref{eq:rda}. Furthermore, a byproduct is that
+adjustments to the associated eigenvalues have no effect on
+\eqref{eq:grda}. Lastly, we exploit the so-called fast singular value
+decomposition (SVD) to construct the eigenvalue decomposition of $\widehat{\bm
+  \Sigma}$, further reducing the computational costs of our proposed \emph{RDA}
+classifier.
 
-First, we require the following relationship regarding the null spaces of $\widehat{\bm \Sigma}_k(\lambda)$, $\widehat{\bm \Sigma}$, and $\widehat{\bm \Sigma}_k$.
+First, we require the following relationship regarding the null spaces of
+$\widehat{\bm \Sigma}_k(\lambda)$, $\widehat{\bm \Sigma}$, and $\widehat{\bm
+  \Sigma}_k$.
 
 \begin{lemma}\label{lemma:null-spaces}
-Let $\widehat{\bm \Sigma}_k$ and $\widehat{\bm \Sigma}$ be the MLEs of $\bm \Sigma_k$ and $\bm \Sigma$ as discussed in Section \ref{sec:rda}, and let $\widehat{\bm \Sigma}_k(\lambda)$ be defined as in \eqref{eq:sig-lambda-alternative}.
-Then, for $k = 1, \ldots, K$, $\mathcal{N}(\widehat{\bm \Sigma}_k(\lambda)) \subset \mathcal{N}(\widehat{\bm \Sigma}) \subset \mathcal{N}(\widehat{\bm \Sigma}_k)$.
+Let $\widehat{\bm \Sigma}_k$ and $\widehat{\bm \Sigma}$ be the MLEs of $\bm
+\Sigma_k$ and $\bm \Sigma$ as discussed in Section \ref{sec:rda}, and let
+$\widehat{\bm \Sigma}_k(\lambda)$ be defined as in
+\eqref{eq:sig-lambda-alternative}. Then, for $k = 1, \ldots, K$,
+$\mathcal{N}(\widehat{\bm \Sigma}_k(\lambda)) \subset \mathcal{N}(\widehat{\bm
+  \Sigma}) \subset \mathcal{N}(\widehat{\bm \Sigma}_k)$.
 \end{lemma}
 \begin{proof}
-Let $\bm z \in \mathcal{N}(\widehat{\bm \Sigma}_k(\lambda))$ for some $k = 1, \ldots, K$.
-Hence, $0 = \bm z' \widehat{\bm \Sigma}_k(\lambda) \bm z = (1 - \lambda) \bm z' \widehat{\bm \Sigma}_k \bm z + \lambda \bm z' \widehat{\bm \Sigma} \bm z$.
-Because $\widehat{\bm \Sigma}_k, \widehat{\bm \Sigma}\in \mathbb{R}_{p \times p}^{\ge}$, we have $\bm z \in \mathcal{N}(\widehat{\bm \Sigma})$ and $\bm z \in \mathcal{N}(\widehat{\bm \Sigma}_k)$.
-In particular, we have that $\mathcal{N}(\widehat{\bm \Sigma}_k(\lambda)) \subset \mathcal{N}(\widehat{\bm \Sigma})$.
-Now, suppose $\bm z \in \mathcal{N}(\widehat{\bm \Sigma})$.
-Similarly, we have that $0 = \bm z' \widehat{\bm \Sigma} \bm z = N^{-1} \sum_{k = 1}^K n_k \bm z' \widehat{\bm \Sigma}_k \bm z$, which implies that $\bm z \in \mathcal{N}(\widehat{\bm \Sigma}_k)$ because $\widehat{\bm \Sigma}_k \in \mathbb{R}_{p \times p}^{\ge}$ for $k = 1, \ldots, K$.
-Therefore, $\mathcal{N}(\widehat{\bm \Sigma}) \subset \mathcal{N}(\widehat{\bm \Sigma}_k)$.
+Let $\bm z \in \mathcal{N}(\widehat{\bm \Sigma}_k(\lambda))$ for some $k = 1,
+\ldots, K$. Hence, $0 = \bm z' \widehat{\bm \Sigma}_k(\lambda) \bm z = (1 -
+\lambda) \bm z' \widehat{\bm \Sigma}_k \bm z + \lambda \bm z' \widehat{\bm
+  \Sigma} \bm z$. Because $\widehat{\bm \Sigma}_k, \widehat{\bm \Sigma}\in
+\mathbb{R}_{p \times p}^{\ge}$, we have $\bm z \in \mathcal{N}(\widehat{\bm
+  \Sigma})$ and $\bm z \in \mathcal{N}(\widehat{\bm \Sigma}_k)$. In particular,
+we have that $\mathcal{N}(\widehat{\bm \Sigma}_k(\lambda)) \subset
+\mathcal{N}(\widehat{\bm \Sigma})$. Now, suppose $\bm z \in
+\mathcal{N}(\widehat{\bm \Sigma})$. Similarly, we have that $0 = \bm z'
+\widehat{\bm \Sigma} \bm z = N^{-1} \sum_{k = 1}^K n_k \bm z' \widehat{\bm
+  \Sigma}_k \bm z$, which implies that $\bm z \in \mathcal{N}(\widehat{\bm
+  \Sigma}_k)$ because $\widehat{\bm \Sigma}_k \in \mathbb{R}_{p \times p}^{\ge}$
+for $k = 1, \ldots, K$. Therefore, $\mathcal{N}(\widehat{\bm \Sigma}) \subset
+\mathcal{N}(\widehat{\bm \Sigma}_k)$.
 \end{proof}
 
-In Lemma \ref{lemma:rda-tilde-Sigma_k} below, we derive an alternative expression for $\tilde{\bm \Sigma}_k$ in terms of the matrix of eigenvectors of $\widehat{\bm \Sigma}$.
-Let $\widehat{\bm \Sigma} = \bm U \bm D \bm U'$ be the eigendecomposition of $\widehat{\bm \Sigma}$ such that $\bm D \in \mathbb{R}_{p \times p}$ is the diagonal matrix of eigenvalues of $\widehat{\bm \Sigma}$ with $\bm D = \bm D_q \oplus \bm 0_{p-q}$, $\bm D_q \in \mathbb{R}_{q \times q}^{>}$ is the diagonal matrix consisting of the positive eigenvalues of $\widehat{\bm \Sigma}$,  the columns of $\bm U \in \mathbb{R}_{p \times p}$ are the corresponding orthonormal eigenvectors of $\widehat{\bm \Sigma}$, and rank$(\widehat{\bm \Sigma}) = q$.
-Then, we partition $\bm U = (\bm U_1, \bm U_2)$ such that $\bm U_1 \in \mathbb{R}_{p \times q}$ and $\bm U_2 \in \mathbb{R}_{p \times (p - q)}$.
+In Lemma \ref{lemma:rda-tilde-Sigma_k} below, we derive an alternative
+expression for $\tilde{\bm \Sigma}_k$ in terms of the matrix of eigenvectors of
+$\widehat{\bm \Sigma}$. Let $\widehat{\bm \Sigma} = \bm U \bm D \bm U'$ be the
+eigendecomposition of $\widehat{\bm \Sigma}$ such that $\bm D \in \mathbb{R}_{p
+  \times p}$ is the diagonal matrix of eigenvalues of $\widehat{\bm \Sigma}$
+with $\bm D = \bm D_q \oplus \bm 0_{p-q}$, $\bm D_q \in \mathbb{R}_{q \times
+  q}^{>}$ is the diagonal matrix consisting of the positive eigenvalues of
+$\widehat{\bm \Sigma}$, the columns of $\bm U \in \mathbb{R}_{p \times p}$ are
+the corresponding orthonormal eigenvectors of $\widehat{\bm \Sigma}$, and
+rank$(\widehat{\bm \Sigma}) = q$. Then, we partition $\bm U = (\bm U_1, \bm
+U_2)$ such that $\bm U_1 \in \mathbb{R}_{p \times q}$ and $\bm U_2 \in
+\mathbb{R}_{p \times (p - q)}$.
 
 \begin{lemma}\label{lemma:rda-tilde-Sigma_k}
-Let $\widehat{\bm \Sigma} = \bm U \bm D \bm U'$ be the eigendecomposition of $\widehat{\bm \Sigma}$ as above, and suppose that rank$(\widehat{\bm \Sigma}) = q \le p$. Then, we have
+Let $\widehat{\bm \Sigma} = \bm U \bm D \bm U'$ be the eigendecomposition of
+$\widehat{\bm \Sigma}$ as above, and suppose that rank$(\widehat{\bm \Sigma}) =
+q \le p$. Then, we have
 \begin{align}
@@
 \begin{proof}
-From Lemma \ref{lemma:null-spaces}, the columns of $\bm U_2$ span the null space of $\widehat{\bm \Sigma}$, which implies that $\widehat{\bm \Sigma}_k \bm U_2 = \bm 0_{p \times (p - q)}$.
-Hence, for $k = 1, \ldots, K$, $\bm U' \widehat{\bm \Sigma}_k \bm U = \bm U_1' \widehat{\bm \Sigma}_k \bm U_1 \oplus \bm 0_{p-q}$.
-Thus, $\bm U' \tilde{\bm \Sigma}_k \bm U = \alpha_k \{(1 - \lambda) \bm U' \widehat{\bm \Sigma}_k \bm U + \lambda \bm D\} + \gamma \bm I_p$, and \eqref{eq:rda-matrix} follows because $\bm U$ is orthogonal.
+From Lemma \ref{lemma:null-spaces}, the columns of $\bm U_2$ span the null space
+of $\widehat{\bm \Sigma}$, which implies that $\widehat{\bm \Sigma}_k \bm U_2 =
+\bm 0_{p \times (p - q)}$. Hence, for $k = 1, \ldots, K$, $\bm U' \widehat{\bm
+  \Sigma}_k \bm U = \bm U_1' \widehat{\bm \Sigma}_k \bm U_1 \oplus \bm
+0_{p-q}$. Thus, $\bm U' \tilde{\bm \Sigma}_k \bm U = \alpha_k \{(1 - \lambda)
+\bm U' \widehat{\bm \Sigma}_k \bm U + \lambda \bm D\} + \gamma \bm I_p$, and
+\eqref{eq:rda-matrix} follows because $\bm U$ is orthogonal.
 \end{proof}
 
-As an immediate immediate consequence of Lemma \ref{lemma:rda-tilde-Sigma_k}, we have the following corollary.
+As an immediate immediate consequence of Lemma \ref{lemma:rda-tilde-Sigma_k}, we
+have the following corollary.
 
 \begin{cor}
-Let $\widehat{\bm \Sigma}_k(\lambda)$ be defined as in \eqref{eq:sig-lambda-alternative}. Then, for $\lambda \in (0, 1]$, rank$(\widehat{\bm \Sigma}_k(\lambda)) = q$ for all $k = 1, \ldots, K$.
+Let $\widehat{\bm \Sigma}_k(\lambda)$ be defined as in
+\eqref{eq:sig-lambda-alternative}. Then, for $\lambda \in (0, 1]$,
+  rank$(\widehat{\bm \Sigma}_k(\lambda)) = q$ for all $k = 1, \ldots, K$.
 \end{cor}
@@
 
-Thus, by incorporating each $\bm x_i$ into the estimation of $\bm \Sigma_k$, we increase the rank of $\widehat{\bm \Sigma}_k(\lambda)$ to $q$, $k = 1, \ldots, K$ if the pooling parameter $\lambda \ne 0$.
-Next, we provide an essential result that enables us to prove that, under mild regularity conditions, \eqref{eq:grda} is invariant to adjustments to the eigenvalues of $\tilde{\bm \Sigma}_k$ corresponding to the null space of $\widehat{\bm \Sigma}$ for any unlabeled observation $\bm x \in \mathbb{R}_{p \times 1}$.
+Thus, by incorporating each $\bm x_i$ into the estimation of $\bm \Sigma_k$, we
+increase the rank of $\widehat{\bm \Sigma}_k(\lambda)$ to $q$, $k = 1, \ldots,
+K$ if the pooling parameter $\lambda \ne 0$. Next, we provide an essential
+result that enables us to prove that, under mild regularity conditions,
+\eqref{eq:grda} is invariant to adjustments to the eigenvalues of $\tilde{\bm
+  \Sigma}_k$ corresponding to the null space of $\widehat{\bm \Sigma}$ for any
+unlabeled observation $\bm x \in \mathbb{R}_{p \times 1}$.
 
 \begin{lemma}\label{lemma:RDA-constant-term}
-Let $\bm U_2$ be defined as above. Then, for all $\bm x \in \mathbb{R}_{p \times 1}$, $\bm U_2' (\bm x - \xbar_k) = \bm U_2' (\bm x - \xbar_{k'})$, $1 \le k, k' \le K$.
+Let $\bm U_2$ be defined as above. Then, for all $\bm x \in \mathbb{R}_{p \times
+  1}$, $\bm U_2' (\bm x - \xbar_k) = \bm U_2' (\bm x - \xbar_{k'})$, $1 \le k,
+k' \le K$.
 \end{lemma}
 \begin{proof}
-Let $\bm x \in \mathbb{R}_{p \times 1}$, and suppose that $1 \le k, k' \le K$.
-Then, $\xbar_k - \xbar_{k'} = \sum_{i=1}^N \beta_i \bm x_i$, where $\beta_i = (n_k n_{k'})^{-1} \{ I(y_i = k) n_{k'} - I(y_i = k') n_k \}$.
-Then, $\bm U_2' (\bm x - \xbar_k) = \bm U_2' (\bm x - \xbar_{k'})$ is equivalent to $\bm U_2'(\xbar_k - \xbar_{k'}) = \sum_{i=1}^N \beta_i \bm U_2' \bm x_i$.
-Recall that $\bm U_2 \in \mathcal{N}(\widehat{\bm \Sigma})$, which implies that $\bm U_2' \in \mathcal{C}(\widehat{\bm \Sigma})^{\perp}$ \citep[Lemma 1.2.5]{Kollo:2005vp}.
-Now, because $\bm x_i \in \mathcal{C}(\widehat{\bm \Sigma})$, $i = 1, \ldots, N$, $\bm U_2' \bm x_i = \bm 0_{(p-q) \times 1}$, which implies that $\bm U_2' (\bm x - \xbar_k) = \bm U_2' (\bm x - \xbar_{k'})$.
+Let $\bm x \in \mathbb{R}_{p \times 1}$, and suppose that $1 \le k, k' \le
+K$. Then, $\xbar_k - \xbar_{k'} = \sum_{i=1}^N \beta_i \bm x_i$, where $\beta_i
+= (n_k n_{k'})^{-1} \{ I(y_i = k) n_{k'} - I(y_i = k') n_k \}$.  Then, $\bm U_2'
+(\bm x - \xbar_k) = \bm U_2' (\bm x - \xbar_{k'})$ is equivalent to $\bm
+U_2'(\xbar_k - \xbar_{k'}) = \sum_{i=1}^N \beta_i \bm U_2' \bm x_i$. Recall that
+$\bm U_2 \in \mathcal{N}(\widehat{\bm \Sigma})$, which implies that $\bm U_2'
+\in \mathcal{C}(\widehat{\bm \Sigma})^{\perp}$ \citep[Lemma
+  1.2.5]{Kollo:2005vp}. Now, because $\bm x_i \in \mathcal{C}(\widehat{\bm
+  \Sigma})$, $i = 1, \ldots, N$, $\bm U_2' \bm x_i = \bm 0_{(p-q) \times 1}$,
+which implies that $\bm U_2' (\bm x - \xbar_k) = \bm U_2' (\bm x - \xbar_{k'})$.
 \end{proof}
 
-We now present our main result, where we decompose the \emph{GRDA} decision rule and show that the component requiring the largest computational costs does not contribute to the classification of an unlabeled observation performed using Lemma \ref{lemma:RDA-constant-term}.
-Our main result reduces \eqref{eq:grda} to an equivalent, more computationally efficient decision rule.
+We now present our main result, where we decompose the \emph{GRDA} decision rule
+and show that the component requiring the largest computational costs does not
+contribute to the classification of an unlabeled observation performed using
+Lemma \ref{lemma:RDA-constant-term}. Our main result reduces \eqref{eq:grda} to
+an equivalent, more computationally efficient decision rule.
 
 \begin{thm}
-Let $\tilde{\bm \Sigma}_k$ and $\bm W_k$ be defined as in \eqref{eq:rda-matrix} and \eqref{eq:Wk}, respectively, and let $\bm U_1$ be defined as above.
-Then, the \emph{GRDA} decision rule in \eqref{eq:grda} is equivalent to
+Let $\tilde{\bm \Sigma}_k$ and $\bm W_k$ be defined as in \eqref{eq:rda-matrix}
+and \eqref{eq:Wk}, respectively, and let $\bm U_1$ be defined as above. Then,
+the \emph{GRDA} decision rule in \eqref{eq:grda} is equivalent to
 	\begin{align}
@@
 \begin{proof}
-	From \eqref{eq:rda-matrix}, we have that $\tilde{\bm \Sigma}_k^{+} = \bm U(\bm W_k^{-1} \oplus \gamma^{+} \bm I_{p-q} )\bm U'$ and $|\tilde{\bm \Sigma}_k| = \gamma^{p-q}| \bm W_k |$.
-Therefore, for all $\bm x \in \mathbb{R}_{p \times 1}$, we have that
+From \eqref{eq:rda-matrix}, we have that $\tilde{\bm \Sigma}_k^{+} = \bm U(\bm
+W_k^{-1} \oplus \gamma^{+} \bm I_{p-q} )\bm U'$ and $|\tilde{\bm \Sigma}_k| =
+\gamma^{p-q}| \bm W_k |$. Therefore, for all $\bm x \in \mathbb{R}_{p \times 1}$,
+we have that
 	\begin{align*}
@@
 	\end{align*}
-The proof follows from Lemma \ref{lemma:RDA-constant-term} because $\bm U_2' (\bm x - \xbar_k)$ is constant for all $k = 1, \ldots, K$.
+The proof follows from Lemma \ref{lemma:RDA-constant-term} because $\bm U_2'
+(\bm x - \xbar_k)$ is constant for all $k = 1, \ldots, K$.
 \end{proof}
@@
 
-Thus far we have shown that the \emph{GRDA} classifier in \eqref{eq:rda-decomposed} is invariant to the term $\bm U_2$, thus yielding a substantial reduction in the computational costs.
-While the classifier is much improved in terms of computational performance compared to the \emph{RDA} classifier originally proposed by \cite{Friedman:1989tm}, we provide two additional results that further improve the computational performance of our \emph{GRDA} classifier.
-First, we demonstrate that the inverse and determinants of $\bm W_k$ can be written so that they are performed on $n_k \times n_k$ matrices rather than the larger $q \times q$ matrices, resulting in further improvements to the computational efficiency of the \emph{GRDA} classifier.
+Thus far we have shown that the \emph{GRDA} classifier in
+\eqref{eq:rda-decomposed} is invariant to the term $\bm U_2$, thus yielding a
+substantial reduction in the computational costs. While the classifier is much
+improved in terms of computational performance compared to the \emph{RDA}
+classifier originally proposed by \cite{Friedman:1989tm}, we provide two
+additional results that further improve the computational performance of our
+\emph{GRDA} classifier. First, we demonstrate that the inverse and determinants
+of $\bm W_k$ can be written so that they are performed on $n_k \times n_k$
+matrices rather than the larger $q \times q$ matrices, resulting in further
+improvements to the computational efficiency of the \emph{GRDA} classifier.
 
@@
 	\end{align}
-where $\bm Q_k = \bm I_{n_k} + \alpha_k(1- \lambda) \bm X_k \bm U_1 \bm \Gamma_k^{-1} \bm U_1' \bm X_k'$ and $\bm \Gamma_k = \alpha_k \lambda \bm D_q + \gamma \bm I_q$.
+where $\bm Q_k = \bm I_{n_k} + \alpha_k(1- \lambda) \bm X_k \bm U_1 \bm
+\Gamma_k^{-1} \bm U_1' \bm X_k'$ and $\bm \Gamma_k = \alpha_k \lambda \bm D_q +
+\gamma \bm I_q$.
 \end{proposition}
 \begin{proof}
-	To calculate $|\bm W_k|$, we apply Theorem 18.1.1 from \cite{Harville:2008wja}, which states that $|\bm A + \bm B \bm T \bm C| = |\bm A| |\bm T| |\bm T^{-1} + \bm C \bm A^{-1} \bm B|$.
-Thus, setting $\bm A = \bm \Gamma_k$, $\bm B = \alpha_k (1 - \lambda) \bm U_1' \bm X_k'$, $\bm T = \bm I_{n_k}$, and $\bm C = \bm X_k \bm U_1$, we have $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$.
-Also, \eqref{eq:W_k_inv} follows from the well-known Sherman-Woodbury formula \citep[Theorem 18.2.8]{Harville:2008wja}.
+To calculate $|\bm W_k|$, we apply Theorem 18.1.1 from \cite{Harville:2008wja},
+which states that $|\bm A + \bm B \bm T \bm C| = |\bm A| |\bm T| |\bm T^{-1} +
+\bm C \bm A^{-1} \bm B|$. Thus, setting $\bm A = \bm \Gamma_k$, $\bm B =
+\alpha_k (1 - \lambda) \bm U_1' \bm X_k'$, $\bm T = \bm I_{n_k}$, and $\bm C =
+\bm X_k \bm U_1$, we have $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$. Also,
+\eqref{eq:W_k_inv} follows from the well-known Sherman-Woodbury formula
+\citep[Theorem 18.2.8]{Harville:2008wja}.
 \end{proof}
 
-Thus far, we have constructed each $\tilde{\bm \Sigma}_k$ in terms of the first $q$ eigenvectors of $\widehat{\bm \Sigma}$.
-By utilizing a technique similar to the so-called kernel trick, we can obtain $\bm U_1$ by computing the eigendecomposition of a much smaller $N \times N$ matrix when $p \gg N$ \citep{Hastie:2008dt}.
-Recall that the pooled sample covariance matrix $\widehat{\bm\Sigma} = \bm X' \bm X$.
-Then, because $\bm X = \bm M \bm D^{1/2} \bm U'$ by the SVD, we have the spectral decomposition $\widehat{\bm\Sigma} = \bm U \bm D \bm U'$, where $\bm M \in \mathbb{R}_{N \times N}$ is orthogonal, $\bm D^{1/2} \in \mathbb{R}_{N \times N}^{\ge}$ is a diagonal matrix consisting of the singular values of $\bm X$, and $\bm U \in \mathbb{R}_{p \times N}$ is orthogonal.
-Now, $\bm X \bm X' = \bm M \bm D \bm M'$ is the eigenvalue decomposition of $\bm X \bm X'$, where $\bm D$ is the matrix of eigenvalues of $\bm X \bm X'$ and the columns of $\bm M$  are the corresponding eigenvectors.
-Hence, we obtain $\bm M$ and $\bm D$ from an eigendecomposition of $\bm X \bm X' \in \mathbb{R}_{N \times N}$.
-Next, we compute $\bm U = \bm X' \bm M \bm D^{+/2}$, where $\bm D^{+/2} = \bm D_q^{-1/2} \oplus \bm 0_{N-q}$.
-We then determine $q$ by computing the number of numerically nonzero eigenvalues present in $\bm D$, where $q$ is the number of eigenvalues that exceeds some tolerance value, say, $\epsilon = 1 \times 10^{-6}$.
-After calculating $q$, we extract $\bm U_1$ by retaining only the first $q$ columns of $\bm U$.
+Thus far, we have constructed each $\tilde{\bm \Sigma}_k$ in terms of the first
+$q$ eigenvectors of $\widehat{\bm \Sigma}$. By utilizing a technique similar to
+the so-called kernel trick, we can obtain $\bm U_1$ by computing the
+eigendecomposition of a much smaller $N \times N$ matrix when $p \gg N$
+\citep{Hastie:2008dt}. Recall that the pooled sample covariance matrix
+$\widehat{\bm\Sigma} = \bm X' \bm X$. Then, because $\bm X = \bm M \bm D^{1/2}
+\bm U'$ by the SVD, we have the spectral decomposition $\widehat{\bm\Sigma} =
+\bm U \bm D \bm U'$, where $\bm M \in \mathbb{R}_{N \times N}$ is orthogonal,
+$\bm D^{1/2} \in \mathbb{R}_{N \times N}^{\ge}$ is a diagonal matrix consisting
+of the singular values of $\bm X$, and $\bm U \in \mathbb{R}_{p \times N}$ is
+orthogonal. Now, $\bm X \bm X' = \bm M \bm D \bm M'$ is the eigenvalue
+decomposition of $\bm X \bm X'$, where $\bm D$ is the matrix of eigenvalues of
+$\bm X \bm X'$ and the columns of $\bm M$ are the corresponding
+eigenvectors. Hence, we obtain $\bm M$ and $\bm D$ from an eigendecomposition of
+$\bm X \bm X' \in \mathbb{R}_{N \times N}$. Next, we compute $\bm U = \bm X' \bm
+M \bm D^{+/2}$, where $\bm D^{+/2} = \bm D_q^{-1/2} \oplus \bm 0_{N-q}$. We then
+determine $q$ by computing the number of numerically nonzero eigenvalues present
+in $\bm D$, where $q$ is the number of eigenvalues that exceeds some tolerance
+value, say, $\epsilon = 1 \times 10^{-6}$. After calculating $q$, we extract
+$\bm U_1$ by retaining only the first $q$ columns of $\bm U$.
 
@@
 
-In this section, we describe four high-dimensional microarray data sets and compare our proposed \emph{GRDA} classifier with five recently proposed classifiers designed for small-sample, high-dimensional data: the \emph{DLDA} classifier from \cite{Dudoit:2002ev}, the shrunken centroids \emph{RDA} (\emph{SCRDA}) classifier from \cite{Guo:2007te}, the shrinkage-based \emph{DLDA} (\emph{SDLDA}) classifier from \cite{Pang:2009ik}, the shrinkage-mean-based DLDA (\emph{SmDLDA}) classifier from \cite{Tong:2012hw}, and the penalized \emph{LDA} (\emph{PLDA}) classifier from \cite{Witten:2011kc}.
-We did not include the \emph{RDA} classifier from \cite{Friedman:1989tm} in our studies to avoid its excessive computations.
-We evaluated the classification performance of each classifier by randomly partitioning the data such that $2/3$ of the observations were allocated as training data and the remaining $1/3$ of the observations were allocated as a test data set.
-For each random partition, we applied the feature-selection method from \cite{Dudoit:2002ev} to choose the $d$ features having the largest ratio of their between-group to within-group sums of squares from the training data.
-We then filtered the test data set with the corresponding $d$ features.
-After training each classifier from the filtered training data, we classified the filtered test data sets and computed the proportion of mislabeled test observations to estimate the classification error rate for each classifier.
-Repeating this process 1000 times, we computed the average of the error-rate estimates for each classifier.
-We remark that our simulation design avoids the selection bias discussed in \cite{Ambroise:2002fa}.
-In Table \ref{tab:data-summary}, we summarize the attributes of each microarray data set.
+In this section, we describe four high-dimensional microarray data sets and
+compare our proposed \emph{GRDA} classifier with five recently proposed
+classifiers designed for small-sample, high-dimensional data: the \emph{DLDA}
+classifier from \cite{Dudoit:2002ev}, the shrunken centroids \emph{RDA}
+classifier from \cite{Guo:2007te}, the shrinkage-based \emph{DLDA} classifier
+from \cite{Pang:2009ik}, the shrinkage-mean-based DLDA classifier from
+\cite{Tong:2012hw}, and the penalized \emph{LDA} classifier from
+\cite{Witten:2011kc}. Below, we refer to each classifier by the first author's
+surname. We evaluated the classification performance of each classifier by
+randomly partitioning the data such that $2/3$ of the observations were
+allocated as training data and the remaining $1/3$ of the observations were
+allocated as a test data set. For each random partition, we applied the
+feature-selection method from \cite{Dudoit:2002ev} to choose the $d$ features
+having the largest ratio of their between-group to within-group sums of squares
+from the training data. We then filtered the test data set with the
+corresponding $d$ features. After training each classifier from the filtered
+training data, we classified the filtered test data sets and computed the
+proportion of mislabeled test observations to estimate the classification error
+rate for each classifier. Repeating this process 1000 times, we computed the
+average of the error-rate estimates for each classifier. We remark that our
+simulation design avoids the selection bias discussed in
+\cite{Ambroise:2002fa}. We did not include the \emph{RDA} classifier from
+\cite{Friedman:1989tm} in our studies to avoid its excessive computations. In
+Table \ref{tab:data-summary}, we summarize the attributes of each microarray
+data set.
 
-We used version 2.15.1 of the open-source statistical software {\tt R} for the classification study.
-For each classifier's {\tt R} implementation, we used the default settings with the exception that we explicitly set the prior probabilities as equal.
-In our {\tt R} implementation of the \emph{GRDA} classifier in the {\tt regdiscrim} package, we set $\alpha_k = 1$, $k = 1, \ldots, K$ for simplicity.
-We estimated $\lambda$ from a grid of 21 equidistant candidate values between 0 and 1, inclusively.
-Similarly, we estimated $\gamma$ from a grid consisting of the values $10^{-1}, \ldots, 10^4$, and $10^5$.
-We selected optimal estimates of $\lambda$ and $\gamma$ using 10-fold cross-validation \citep{Hastie:2008dt}.
+We used version 3.0.0 of the open-source statistical software {\tt R} for the
+classification study. For each classifier's {\tt R} implementation, we used the
+default settings with the exception that we explicitly set the prior
+probabilities as equal. In our {\tt R} implementation of the \emph{GRDA}
+classifier in the {\tt regdiscrim} package, we set $\alpha_k = 1$, $k = 1,
+\ldots, K$ for simplicity. We estimated $\lambda$ from a grid of 21 equidistant
+candidate values between 0 and 1, inclusively. Similarly, we estimated $\gamma$
+from a grid consisting of the values $10^{-1}, \ldots, 10^4$, and $10^5$. We
+selected optimal estimates of $\lambda$ and $\gamma$ using 10-fold
+cross-validation \citep{Hastie:2008dt}.
 
@@
 
-\cite{Burczynski:2006ik} acquired the peripheral blood mononuclear cells (PBMC) through hybridization to microarrays from 127 individuals resulting in 22,283 sequences.
-Of the 127 individuals, 42 were healthy, 59 had Crohn's disease (CD), and 26 had ulcerative colitis (UC).
-The goal of \cite{Burczynski:2006ik} was to improve accuracy in the discrimination of Inflammatory Bowel Disease (IBD) using the PBMC-based gene expression signature of a patient.
+\cite{Burczynski:2006ik} acquired the peripheral blood mononuclear cells (PBMC)
+through hybridization to microarrays from 127 individuals resulting in 22,283
+sequences. Of the 127 individuals, 42 were healthy, 59 had Crohn's disease (CD),
+and 26 had ulcerative colitis (UC). The goal of \cite{Burczynski:2006ik} was to
+improve accuracy in the discrimination of Inflammatory Bowel Disease (IBD) using
+the PBMC-based gene expression signature of a patient.
 
@@
 
-\cite{Nakayama:2007fl} acquired the gene expression through an oligonucleotide microarray from 105 samples of 10 types of soft tissue tumors.
-This included 16 samples of synovial sarcoma (SS), 19 samples of myxoid/round cell liposarcoma (MLS), 3 samples of lipoma, 3 samples of well-differentiated liposarcoma (WDLS), 15 samples of dedifferentiated liposarcoma (DDLS), 15 samples of myxofibrosarcoma (MFS), 6 samples of leiomyosarcoma (LMS), 3 samples of malignant nerve sheathe tumor (MPNST), 4 samples of fibrosarcoma (FS), and 21 samples of malignant fibrous histiocytoma (MFH).
-\cite{Nakayama:2007fl} determined from their data that these 10 types fell into 4 broader groups: (1) SS; (2) MLS; (3) Lipoma, WDLS, and part of DDLS; (4) Spindle cell and pleomorophic sarcomas including DDLS, MFS, LMS, MPNST, FS, and MFH.
-Following \cite{Witten:2011kc}, we restrict our analysis to the five tumor types having at least 15 observations.
+\cite{Nakayama:2007fl} acquired the gene expression through an oligonucleotide
+microarray from 105 samples of 10 types of soft tissue tumors. This included 16
+samples of synovial sarcoma (SS), 19 samples of myxoid/round cell liposarcoma
+(MLS), 3 samples of lipoma, 3 samples of well-differentiated liposarcoma (WDLS),
+15 samples of dedifferentiated liposarcoma (DDLS), 15 samples of
+myxofibrosarcoma (MFS), 6 samples of leiomyosarcoma (LMS), 3 samples of
+malignant nerve sheathe tumor (MPNST), 4 samples of fibrosarcoma (FS), and 21
+samples of malignant fibrous histiocytoma (MFH). \cite{Nakayama:2007fl}
+determined from their data that these 10 types fell into 4 broader groups: (1)
+SS; (2) MLS; (3) Lipoma, WDLS, and part of DDLS; (4) Spindle cell and
+pleomorophic sarcomas including DDLS, MFS, LMS, MPNST, FS, and MFH. Following
+\cite{Witten:2011kc}, we restrict our analysis to the five tumor types having at
+least 15 observations.
 
@@
 
-Approximately 30-40\% of adult non-Hodgkin lymphomas are diffuse large B-cell lymphomas (DLBCLs) \cite{Shipp:2002ka}.
-However, only a small proportion of DLBCL patients are cured with modern chemotherapeutic regimens.
-Several models have been proposed, such as the International Prognostic Index (IPI), to determine a patient's curability.
-These models rely on clinical covariates, such as age, to determine if the patient can be cured and are often ineffective.
-\cite{Shipp:2002ka} have argued that more effective means are desired to determine a patient's curability.
-The authors measured 6,817 gene expression levels from 58 DLBCL patient samples with customized cDNA (lymphochip) microarrays to investigate the curability of patients treated with cyclophosphamide, adriamycin, vincristine, and prednisone (CHOP)-based chemotherapy.
-Among the 58 DLBCL patient samples, 32 are from cured patients while 26 are from patients with fatal or refractory disease.
+Approximately 30-40\% of adult non-Hodgkin lymphomas are diffuse large B-cell
+lymphomas (DLBCLs) \cite{Shipp:2002ka}.  However, only a small proportion of
+DLBCL patients are cured with modern chemotherapeutic regimens.  Several models
+have been proposed, such as the International Prognostic Index (IPI), to
+determine a patient's curability.  These models rely on clinical covariates,
+such as age, to determine if the patient can be cured and are often
+ineffective. \cite{Shipp:2002ka} have argued that more effective means are
+desired to determine a patient's curability. The authors measured 6,817 gene
+expression levels from 58 DLBCL patient samples with customized cDNA
+(lymphochip) microarrays to investigate the curability of patients treated with
+cyclophosphamide, adriamycin, vincristine, and prednisone (CHOP)-based
+chemotherapy.  Among the 58 DLBCL patient samples, 32 are from cured patients
+while 26 are from patients with fatal or refractory disease.
 
@@
 
-\cite{Singh:2002fh} have examined 235 radical prostatectomy specimens from surgery patients between 1995 and 1997.
-The authors used oligonucleotide microarrays containing probes for approximately 12,600 genes and expressed sequence tags.
-They have reported that 102 of the radical prostatectomy specimens are of high quality: 52 prostate tumor samples and 50 non-tumor prostate samples.
+\cite{Singh:2002fh} have examined 235 radical prostatectomy specimens from
+surgery patients between 1995 and 1997. The authors used oligonucleotide
+microarrays containing probes for approximately 12,600 genes and expressed
+sequence tags. They have reported that 102 of the radical prostatectomy
+specimens are of high quality: 52 prostate tumor samples and 50 non-tumor
+prostate samples.
 
@@
 
-For the microarray data sets, we computed error-rate estimates for the competing classifiers after reducing the original feature space dimension with variable selection to $d=100,250,500$, and $1000$.
-Our choices for the values of $d$ are similar to those of \cite{Xu:2009fl}.
-We report the error-rate estimates along with approximate standard errors in parentheses in Tables \ref{tab:results-d100} -- \ref{tab:results-d1000} along with the chosen reduced dimension $d$.
+For the microarray data sets, we computed error-rate estimates for the competing
+classifiers after reducing the dimension of the data sets with feature selection
+to $d = 100, \ldots, 1000$, similar to the values of $d$ considered by
+\cite{Xu:2009fl}. To compare misclassification rates between different values of
+$d$, we applied each value of $d$ considered to the same random allocation of
+training and test data. Hence, for a given random partition, the set of features
+selected for a small value of $d$ is a proper subset of the features selected
+for a larger value of $d$. We report the error-rate estimates along with
+approximate standard errors in parentheses in Tables \ref{tab:results-d100} --
+\ref{tab:results-d1000} along with the reduced dimension $d$.
 
-The \emph{GRDA} classifier was superior in classification performance for the majority of the simulations.
-While the error-rate estimates for the majority of the classifiers increased as $d$ was increased, the classification accuracy for the \emph{GRDA} classifier improved as $d$ increased for the Burczynski and Nakayama data sets.
-For the Nakayama data set, the \emph{GRDA} classifier was outperformed by the Dudoit, Pang, Tong, and Witten classifiers for small values of $d$ but attained classification accuracy comparable to its competitors for larger values of $d$.
-Moreover, the only situation where the other classifiers considered were competitive with the \emph{GRDA} classifier was in the case of the Nakayama data set.
+The \emph{GRDA} classifier was superior in classification performance for the
+majority of the simulations. For the Burczysnki, Shipp, and Singh data sets, the
+\emph{GRDA} classifier outperformed its competitors for all values of $d$
+examined. For the Nakayama data set, the \emph{GRDA} classifier was outperformed
+by the Dudoit, Pang, and Tong classifiers for small values of $d$ but attained
+classification rates superior to its competitors for larger values of
+$d$. Furthermore, Figure \textbf{TODO} emphasizes two strengths of the
+\emph{GRDA} classifier: reliability of the classifier for small $d$ and
+stability of the classifier for large $d$. In contrast to the other classifiers,
+the \emph{GRDA} classifier exhibited less variability its misclassification
+estimates for small values of $d$. Also, the \emph{GRDA} classifier yielded
+superior classification accuracy as $d$ increased and was hardly effected by the
+inclusion of more variables in the data sets. We attribute the apparent
+stability of the error-rate estimates of the \emph{GRDA} classifier for large
+values of $d$ to its novel dimension-reduction technique.
 
-The error-rate estimates for the Dudoit, Pang, Tong, and Witten classifiers were comparable regardless of the choice of $d$.
-In particular, the error-rate estimates for Dudoit's and Tong's methods were nearly identical, which suggests that the improved mean estimator employed by Tong provides little improvement to the classification performance in the \emph{DLDA} classifier.
-The popular classifier from Guo is competitive only for large values of $d$ for the Shipp and Singh data sets.
-For the Burczynski and Nakayama data sets, the implementation of Guo's method in the {\tt rda} package often classified all of the test observations into a single class, thus yielding larger error rates than its competitors.
+The error-rate estimates for the Dudoit, Pang, Tong, and Witten classifiers were
+comparable for all values of $d$ examined. Furthermore, the error-rate estimates
+of the Dudoit, Pang, and Tong classifiers were approximately equal throughout
+our study, and the Witten classifier yielded approximately equal error-rate
+estimates for the Burczysnki, Shipp, and Singh data sets. Furthermore, the
+error-rate estimates for the Dudoit and Tong classifiers were nearly identical,
+which suggests that the improved mean estimator employed by \cite{Tong:2012hw}
+provided little improvement to the classification performance of the \emph{DLDA}
+classifier.
+
+The classification performance of the the Dudoit, Pang, Tong, and Witten
+classifiers for the Shipp and Singh data sets revealed the inability of these
+methods to handle larger values of $d$. After closer examination of the
+classification studies performed in the authors' original papers, we noticed
+that classification accuracy estimates were reported only for small values of
+$d$, thereby ignoring the inclusion of a larger number of variables. Hence, we
+advise that practioners be aware of the necessity to filter a large number of
+features before employing these methods, lest their classification performance
+can degrade.
+
+The popular classifier from Guo is most competitive over large values of $d$ for
+the Shipp and Singh data sets. However, for the Burczynski and Nakayama data
+sets, the implementation of Guo's method in the {\tt rda} package often
+classified every test observation into a single class, thus yielding larger
+error rates than its competitors. Furthermore, the Guo classifier exhibited
+unreliable classification performance for smaller values of $d$. The Guo
+classifier required a larger subset of the features to perform competitively in
+our study.
+
+As seen in Figure ?, the \emph{GRDA} classifier exhibits an interesting bathtub
+shape in selection of the lambda tuning parameter through cross-validation. (Not
+sure what else to write here.)
 
@@
 
-The analysis of high-dimensional microarray data sets is important to modern biomedical studies and presents challenges with supervised classification because the number of gene expressions greatly outnumbers the number of collected responses.
-Consequently, regularization techniques are often employed to stabilize popular classifiers for improved classification performance.
-We have provided a generalization of the well-known \emph{RDA} classifier from \cite{Friedman:1989tm}, aimed for small-sample, high-dimensional classification.
-The covariance-matrix regularization employed in the \emph{GRDA} classifier includes a large family of shrinkage methods.
-Our formulation of the \emph{GRDA} classifier in terms of the nonzero eigenvalues of the pooled sample covariance matrix indicates that eigenvalue-adjustment methods applied to the \emph{LDA} classifier explains the similiarities in classification accuracy of several regularization methods applied to the \emph{LDA} classifier reported by \cite{Ramey:2013ji} in a high-dimensional classification study.
-We have provided an invariance property that extends results of \cite{Ji:2008wp} and \cite{Ye:2006jm}, who have provided a similar invariance property for the \emph{LDA} decision rule.
+The analysis of high-dimensional microarray data sets is important to modern
+biomedical studies and presents challenges with supervised classification
+because the number of gene expressions greatly outnumbers the number of
+collected responses. Consequently, regularization techniques are often employed
+to stabilize popular classifiers for improved classification performance. We
+have provided a generalization of the well-known \emph{RDA} classifier from
+\cite{Friedman:1989tm}, aimed for small-sample, high-dimensional
+classification. The covariance-matrix regularization employed in the \emph{GRDA}
+classifier includes a large family of shrinkage methods. Our formulation of the
+\emph{GRDA} classifier in terms of the nonzero eigenvalues of the pooled sample
+covariance matrix indicates that eigenvalue-adjustment methods applied to the
+\emph{LDA} classifier explains the similiarities in classification accuracy of
+several regularization methods applied to the \emph{LDA} classifier reported by
+\cite{Ramey:2013ji} in a high-dimensional classification study. We have provided
+an invariance property that extends results of \cite{Ji:2008wp} and
+\cite{Ye:2006jm}, who have provided a similar invariance property for the
+\emph{LDA} decision rule.
 
-We have provided an implementation of the \emph{GRDA} classifier in the {\tt regdiscrim} R package, available on CRAN.
-Our implementation facilitates a rapid classification of high-dimensional data by employing the novel results provided in this paper.
-Moreover, our 10-fold cross-validation procedure typically requires a couple of seconds for data sets having large $p$.
-When preliminary variable selection is applied to the data as we performed in our classification study, the time to perform model selection is nearly instantaneous, confirming the computational complexity reported regarding the operations employed in the \emph{GRDA} classifier.
+We have provided an implementation of the \emph{GRDA} classifier in the {\tt
+  regdiscrim} R package, available on CRAN. Our implementation facilitates a
+rapid classification of high-dimensional data by employing the novel results
+provided in this paper. Moreover, our 10-fold cross-validation procedure
+typically requires a couple of seconds for data sets having large $p$. When
+preliminary variable selection is applied to the data as we performed in our
+classification study, the time to perform model selection is nearly
+instantaneous, confirming the computational complexity reported regarding the
+operations employed in the \emph{GRDA} classifier.
 
-Although several variants of the \emph{LDA} classifier have been proposed for high-dimensional, classification, the \emph{QDA} classifier has largely been ignored, in part due to the large number of parameters that must be estimated.
-Our proposed classifier ...
+Although several variants of the \emph{LDA} classifier have been proposed for
+high-dimensional, classification, the \emph{QDA} classifier has largely been
+ignored, in part due to the large number of parameters that must be
+estimated. Our proposed classifier ...
 
-Our formulation of covariance-matrix regularization employed in the \emph{GRDA} classifier includes a large family of shrinkage methods, such as the methods from \cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and a wide variety of other shrinkage methods studied by \cite{Ramey:2013ji} and \cite*{Xu:2009fl}.
+Our formulation of covariance-matrix regularization employed in the \emph{GRDA}
+classifier includes a large family of shrinkage methods, such as the methods
+from \cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and a wide variety of other
+shrinkage methods studied by \cite{Ramey:2013ji} and \cite*{Xu:2009fl}.
 

### 2013-04-15 reports/paper-rda.Rnw
Major update to simulation results summary. Closes #3, #4, #6, #7, and #8.

@@
 
-\title{A Generalization of Regularized Discriminant Analysis for High-Dimensional Classification}
+\title{Another Look at Regularized Discriminant Analysis for High-Dimensional Classification}
 
@@
 
-<<setup, include=FALSE, cache=FALSE, echo=FALSE, warning=FALSE>>=
-opts_chunk$set(fig.align = 'default', dev = 'png', message = FALSE, warning = FALSE,
-               cache = FALSE, echo = FALSE)
+<<knitr_setup, include=FALSE, echo=FALSE, cache=FALSE>>=
+# Round digits throughout document including \Sexpr statements.
+# See Yihui's response here:
+# http://stackoverflow.com/questions/11062497/how-to-avoid-using-round-in-every-sexpr
+options(digits = 3)
 
+opts_chunk$set(fig.align = 'default', dev = 'png', message = FALSE,
+               warning = FALSE, cache = TRUE, echo = FALSE, autodep = TRUE,
+               fig.width = 10, fig.height = 10, out.width = "\\linewidth")
+@
+
+<<setup>>=
 setwd("..")
@@
 
-# Clean up results for display
-results <- na.omit(results)
-results$d <- as.factor(results$d)
-results$data_set <- as.character(results$data_set)
+results <- lapply(results, function(res) {
+  res <- lapply(res, function(res_d) {
+    errors <- do.call(cbind, res_d$errors)
+    res_d[["errors"]] <- NULL
+    cbind(do.call(cbind, res_d), errors)
+  })
+  do.call(rbind, res)
+})
+results <- data.frame(do.call(rbind, results), stringsAsFactors = FALSE)
+
+results$lambda <- as.numeric(results$lambda)
+results$gamma <- as.numeric(results$gamma)
+results$Dudoit <- as.numeric(results$Dudoit)
+results$GRDA <- as.numeric(results$GRDA)
+results$Guo <- as.numeric(results$Guo)
+results$Pang <- as.numeric(results$Pang)
+results$Tong <- as.numeric(results$Tong)
+results$Witten <- as.numeric(results$Witten)
+
+results$d <- factor(results$d, labels = sort(unique(as.numeric(results$d))))
+
+# Summary of the classification error rates for each data set and number of
+# variables selected, d.
+error_rates <- subset(results, select = -c(lambda, gamma))
+m_error <- melt(error_rates, variable.name = "Classifier", value.name = "Error")
+error_rates <- ddply(m_error, .(data_set, d, Classifier), plyr:::summarize,
+                     Avg_Error = mean(Error), Num_Reps = length(Error))
+error_rates <- ddply(error_rates, .(data_set, d, Classifier), transform,
+                     SE_Error = sqrt(Avg_Error * (1 - Avg_Error) / Num_Reps))
+error_rates <- subset(error_rates, select = -Num_Reps)
+
+# Candidate values of tuning parameters.
+seq_lambda <- seq(0, 1, by = 0.05)
+seq_gamma <- 10^seq.int(-1, 5)
+
+# Summary of the optimal tuning-parameter estimates for each data set and number
+# of variables selected, d.
+grda_optimal <- subset(results, select = c(data_set, d, lambda, gamma))
+grda_optimal$lambda <- factor(grda_optimal$lambda, levels = seq_lambda)
+grda_optimal$gamma <- factor(grda_optimal$gamma, levels = seq_gamma)
+
+grda_optimal <- ddply(grda_optimal, .(data_set, d), function(x) {
+  x <- subset(x, select = c(lambda, gamma))
+  melt(table(x))
+})
 
@@
 popularized by \cite{Friedman:1989tm}, which requires computationally expensive
-matrix operations and involves a time-consuming model-selection procedure,
-making the \emph{RDA} classifier impractical for high-dimensional data. We
+matrix operations and involves a time-consuming model-selection procedure. We
 propose a novel modification of the \emph{RDA} (\emph{GRDA}) classifier that
-incorporates a more general covariance-matrix regularization approach that
-includes several proposed sparse and regularization methods as special
-cases. Furthermore, we overcome the computational shortcomings of the \emph{RDA}
-classifier by integrating a novel dimension-reduction technique that enables a
-computationally efficient model-selection procedure. Moreover, we demonstrate
-with several high-dimensional microarray data sets that our proposed classifier
-is competitive and often superior to several recently proposed sparse and
-regularized classifiers in terms of classification accuracy. Finally, we provide
-an implementation of the \emph{GRDA} classifier in the {\tt regdiscrim} R
-package, available on CRAN.
+provides a straightforward interpration and employs a novel dimension-reduction
+technique to enable a computationally efficient model-selection procedure, thus
+overcoming the computational shortcomings of the \emph{RDA}
+classifier. Moreover, we demonstrate with several high-dimensional microarray
+data sets that our proposed classifier is competitive and often superior to
+several recently proposed sparse and regularized classifiers in terms of
+classification accuracy. Finally, we provide an implementation of the
+\emph{GRDA} classifier in the {\tt sparsediscrim} R package, available on CRAN.
 
@@
 classifier can attain improved classification accuracy, compared to the
-\emph{LDA} and \emph{QDA} classifiers when $p > N$. However, in its original
+\emph{LDA} and \emph{QDA} classifiers when $p < N$. However, in its original
 form, the \emph{RDA} classifier is impractical when $p \gg N$ because the
@@
 advantageous because $n_k$ is often quite small, say, 15--20, in microarray data
-sets. Thus, through a dimension-reduction technique we demonstrate that the
-calculation of the \emph{RDA} decision rule involving inverses and determinants
-of extremely large covariance matrices is equivalent to a decision rule
-involving inverses and determinants for much lower-dimensional matrices.
+sets.
 
@@
 weights in \eqref{eq:sig-lambda-alternative2} reduce to $1/N$, corresponding to
-$\widehat{\bm\Sigma}$. For brevity, for $\lambda = 1$, we define $\bm X =
+$\widehat{\bm\Sigma}$. In Figure \ref{fig:grda-contours} we demonstrate the
+effect of $\lambda$ on the contours of five multivariate normal populations with
+unequal covariance matrices. For brevity, when $\lambda = 1$, we define $\bm X =
 [\sqrt{c_{1k}(1)} \bm x_1', \ldots, \sqrt{c_{Nk}(1)} \bm x_N']'$ such that
@@
 that $\widehat{\bm \Sigma}_k = \bm X_k' \bm X_k$, $k = 1, \ldots, K$.
+
+\[ \left[\text{Figure \ref{fig:grda-contours} goes about here }\right]\]
 
@@
 \begin{proof}
-Let $\bm x \in \mathbb{R}_{p \times 1}$, and suppose that $1 \le k, k' \le
-K$. Then, $\xbar_k - \xbar_{k'} = \sum_{i=1}^N \beta_i \bm x_i$, where $\beta_i
-= (n_k n_{k'})^{-1} \{ I(y_i = k) n_{k'} - I(y_i = k') n_k \}$.  Then, $\bm U_2'
-(\bm x - \xbar_k) = \bm U_2' (\bm x - \xbar_{k'})$ is equivalent to $\bm
-U_2'(\xbar_k - \xbar_{k'}) = \sum_{i=1}^N \beta_i \bm U_2' \bm x_i$. Recall that
-$\bm U_2 \in \mathcal{N}(\widehat{\bm \Sigma})$, which implies that $\bm U_2'
-\in \mathcal{C}(\widehat{\bm \Sigma})^{\perp}$ \citep[Lemma
+Let $\bm x \in \mathbb{R}_{p \times 1}$, and suppose that $1 \le k, k' \le K$.
+Recall that $\bm U_2 \in \mathcal{N}(\widehat{\bm \Sigma})$, which implies that
+$\bm U_2' \in \mathcal{C}(\widehat{\bm \Sigma})^{\perp}$ \citep[Lemma
   1.2.5]{Kollo:2005vp}. Now, because $\bm x_i \in \mathcal{C}(\widehat{\bm
-  \Sigma})$, $i = 1, \ldots, N$, $\bm U_2' \bm x_i = \bm 0_{(p-q) \times 1}$,
-which implies that $\bm U_2' (\bm x - \xbar_k) = \bm U_2' (\bm x - \xbar_{k'})$.
+  \Sigma})$, $i = 1, \ldots, N$, $\bm U_2' \bm x_i = \bm 0_{(p-q) \times 1}$.
+Hence, $\bm 0_{(p-q) \times 1} = \sum_{i=1}^N \beta_i \bm U_2'\bm x_i = \bm U_2'(\xbar_k -
+\xbar_{k'})$, where $\beta_i = (n_k n_{k'})^{-1} \{ I(y_i = k) n_{k'} - I(y_i =
+k') n_k \}$. Therefore, $\bm U_2' (\bm x - \xbar_k) = \bm U_2' (\bm x - \xbar_{k'})$.
 \end{proof}
@@
 probabilities as equal. In our {\tt R} implementation of the \emph{GRDA}
-classifier in the {\tt regdiscrim} package, we set $\alpha_k = 1$, $k = 1,
+classifier in the {\tt sparsediscrim} package, we set $\alpha_k = 1$, $k = 1,
 \ldots, K$ for simplicity. We estimated $\lambda$ from a grid of 21 equidistant
@@
 selected for a small value of $d$ is a proper subset of the features selected
-for a larger value of $d$. We report the error-rate estimates along with
-approximate standard errors in parentheses in Tables \ref{tab:results-d100} --
-\ref{tab:results-d1000} along with the reduced dimension $d$.
+for a larger value of $d$. We report the average of the error-rate estimates for
+each value of $d$ in Figure \ref{fig:error-rates}. Approximate standard errors
+were no greater than \Sexpr{max(error_rates[["SE_Error"]])}.
 
@@
 classification rates superior to its competitors for larger values of
-$d$. Furthermore, Figure \textbf{TODO} emphasizes two strengths of the
+$d$. Furthermore, Figure \ref{fig:error-rates} emphasizes two strengths of the
 \emph{GRDA} classifier: reliability of the classifier for small $d$ and
@@
 
-The popular classifier from Guo is most competitive over large values of $d$ for
-the Shipp and Singh data sets. However, for the Burczynski and Nakayama data
-sets, the implementation of Guo's method in the {\tt rda} package often
-classified every test observation into a single class, thus yielding larger
-error rates than its competitors. Furthermore, the Guo classifier exhibited
-unreliable classification performance for smaller values of $d$. The Guo
-classifier required a larger subset of the features to perform competitively in
-our study.
+The popular Guo classifier is most competitive over large values of $d$ for the
+Shipp and Singh data sets. However, for the Burczynski and Nakayama data sets,
+the implementation of the Guo method in the {\tt rda} package often classified
+every test observation into a single class, thus yielding larger error rates
+than its competitors. Furthermore, the Guo classifier exhibited unreliable
+classification performance for smaller values of $d$. The Guo classifier
+required a larger subset of the features to perform competitively in our study.
 
-As seen in Figure ?, the \emph{GRDA} classifier exhibits an interesting bathtub
-shape in selection of the lambda tuning parameter through cross-validation. (Not
-sure what else to write here.)
+\[ \left[\text{Figure \ref{fig:error-rates} goes about here }\right]\]
 
-\[ \left[\text{Table 2 goes about here }\right]\]
+Next, we examine the optimal estimates of the tuning parameters for the
+\emph{GRDA} classifier. For brevity, we include only a graphical summary of the
+optimal estimates of $\lambda$ for the Shipp data set in Figure
+\ref{fig:optimal-shipp}. Similar figures for the other three data sets can be
+found in the Supplementary Information. In Figure \ref{fig:optimal-shipp}, the
+distribution of the optimal values of $\lambda$ selected through
+cross-validation has a notable interesting bathtub shape, resembling a Beta
+distribution with shape parameters $1/2$ and $1/2$. Specifically, the optimal
+values of $\lambda$ are typically near 0 or near 1, corresponding to the
+\emph{QDA} and \emph{LDA} classifiers, respectively. We noticed a similar
+distribution of optimal values of $\lambda$ for the other three data sets.
 
-\[ \left[\text{Table 3 goes about here }\right]\]
+For the Shipp data set with smaller values of $d$, the range of optimal values
+of $\lambda$ was typically no larger than 0.05, indicating there were a large
+number of training data sets for which we can safely assume the covariance
+matrices are unequal or nearly unequal. As $d$ increased, the majority of the
+optimal values of $\lambda$ were selected as unity, thereby an increased number
+of variables resulted in the \emph{GRDA} classifier preferring the more
+parsimonious assumption of equal covariance matrices. In this instance we can
+see the flexibility of the \emph{GRDA} classifier to employ the more realistic
+assumption that the class covariance matrices are unequal, but if optimal, we
+can employ the more parsimonious \emph{LDA} classifier as a special case.
 
-\[ \left[\text{Table 4 goes about here }\right]\]
-
-\[ \left[\text{Table 5 goes about here }\right]\]
+\[ \left[\text{Figure \ref{fig:optimal-shipp} goes about here }\right]\]
 
@@
 because the number of gene expressions greatly outnumbers the number of
-collected responses. Consequently, regularization techniques are often employed
-to stabilize popular classifiers for improved classification performance. We
-have provided a generalization of the well-known \emph{RDA} classifier from
-\cite{Friedman:1989tm}, aimed for small-sample, high-dimensional
-classification. The covariance-matrix regularization employed in the \emph{GRDA}
-classifier includes a large family of shrinkage methods. Our formulation of the
-\emph{GRDA} classifier in terms of the nonzero eigenvalues of the pooled sample
-covariance matrix indicates that eigenvalue-adjustment methods applied to the
-\emph{LDA} classifier explains the similiarities in classification accuracy of
-several regularization methods applied to the \emph{LDA} classifier reported by
-\cite{Ramey:2013ji} in a high-dimensional classification study. We have provided
-an invariance property that extends results of \cite{Ji:2008wp} and
-\cite{Ye:2006jm}, who have provided a similar invariance property for the
-\emph{LDA} decision rule.
+collected responses. Consequently, several variants of the well-known \emph{LDA}
+classifier have been proposed by employing for improved classification
+performance. However, the \emph{LDA} classifier includes the strong assumption
+that the covariance matrices of each class are equal. Ideally, we could model
+the class covariance matrices as unrestricted and unequal in the \emph{QDA}
+classifier, but it is well known to have a large number of parameters that must
+be estimated and require a large number of samples, which are often unavailable
+in the high-dimensional setting. With this mind, \cite{Friedman:1989tm} proposed
+the \emph{RDA} classifier as a compromise between the \emph{LDA} and \emph{QDA}
+classifiers, controlling the amount of pooling of the individual class
+covariance-matrix estimators using the tuning parameter $\lambda$. However,
+despite its success for data sets where $p > N$, the \emph{RDA} classifier
+requires excessive calculations and a time-consuming model selection.
 
-We have provided an implementation of the \emph{GRDA} classifier in the {\tt
-  regdiscrim} R package, available on CRAN. Our implementation facilitates a
-rapid classification of high-dimensional data by employing the novel results
-provided in this paper. Moreover, our 10-fold cross-validation procedure
-typically requires a couple of seconds for data sets having large $p$. When
+With this in mind, we have adapted the \emph{RDA} classifier for usage with
+modern small-sample, high-dimensional data sets. Our model formulation
+facilitates a rapid classification of high-dimensional data by employing a novel
+dimension reduction technique. Moreover, our 10-fold cross-validation procedure
+typically requires a small runtime for data sets having large $p$. When
 preliminary variable selection is applied to the data as we performed in our
@@
 instantaneous, confirming the computational complexity reported regarding the
-operations employed in the \emph{GRDA} classifier.
+operations employed in the \emph{GRDA} classifier. Additionally, we have
+provided an implementation of the \emph{GRDA} classifier in the {\tt
+  sparsediscrim} R package that is easy to use and available on CRAN.
 
-Although several variants of the \emph{LDA} classifier have been proposed for
-high-dimensional, classification, the \emph{QDA} classifier has largely been
-ignored, in part due to the large number of parameters that must be
-estimated. Our proposed classifier ...
+We have parameterized the covariance-matrix regularization employed in the
+\emph{GRDA} classifier to include a large family of shrinkage methods, such as
+the methods from \cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and a wide variety
+of other shrinkage methods studied by \cite{Ramey:2013ji} and
+\cite{Xu:2009fl}. Our formulation of the \emph{GRDA} classifier in terms of the
+nonzero eigenvalues of the pooled sample covariance matrix indicates that
+eigenvalue-adjustment methods applied to the \emph{LDA} classifier explains the
+similiarities in classification accuracy of several regularization methods
+applied to the \emph{LDA} classifier reported by \cite{Ramey:2013ji} in a
+high-dimensional classification study. We have provided an invariance property
+that extends results of \cite{Ji:2008wp} and \cite{Ye:2006jm}, who have provided
+a similar invariance property for the \emph{LDA} decision rule. Additionally,
+our model formulation improves the interpretation of the \emph{RDA} classifier
+by explicitly defining the covariance-matrix estimator so that each of the $N$
+observations contributes to the estimation of the individual class covariance
+matrices.
 
-Our formulation of covariance-matrix regularization employed in the \emph{GRDA}
-classifier includes a large family of shrinkage methods, such as the methods
-from \cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and a wide variety of other
-shrinkage methods studied by \cite{Ramey:2013ji} and \cite*{Xu:2009fl}.
+In our simulation study, we compared the \emph{GRDA} classifier to several
+recently proposed classifiers, including the Dudoit, Pang, Tong, and Guo
+classifiers, each of which assumes independence of the features conditional on
+class membership. Despite the rapid computational performance of diagonal
+classifiers and the reduction in the number of parameters to estimate,
+\cite{Mai:2012bf} have noted that diagonal classifiers can often yield inferior
+classification performance to other classification methods. We have shown that
+our proposed \emph{GRDA} classifier often yields superior classification
+accuracy to the diagonal classifiers considered, confirming the assertion of
+\cite{Mai:2012bf}. Furthermore, the \emph{GRDA} classifier is able to compete
+with the diagonal classifiers in terms of speed by incorporating a novel
+dimension-reduction technique. Although the diagonal classifiers are clearly
+faster to compute, the differences in runtime for small-sample, high-dimensional
+data sets is hardly noticeable.
 
@@
 
-<<microarray_results, results="asis">>=
-results_summary <- ddply(results, .(data_set, d, Classifier), summarize,
-                         Avg_Error = mean(Error),
-                         SE_Error = sqrt(mean(Error) * (1 - mean(Error)) / length(Error)))
+\begin{figure}
+<<grda_contours, fig.width=8, fig.height=8, out.width='.49\\linewidth', fig.show='hold'>>=
+rda_contours(lambda = 0)
+rda_contours(lambda = 1/4)
+rda_contours(lambda = 3/4)
+rda_contours(lambda = 1)
+@ 
+\caption{Contours of five multivariate normal populations as a function of the
+  pooling parameter $\lambda$.}
+\label{fig:grda-contours}
+\end{figure}
 
-# Tables containing the classification results for each value of 'q'
-results_summary$data_set <- Hmisc:::capitalize(results_summary$data_set)
-results_summary$d <- as.factor(results_summary$d)
-results_summary$Avg_Error <- round_digits(results_summary$Avg_Error, 3)
-results_summary$SE_Error <- round_digits(results_summary$SE_Error, 3)
-results_summary$Error <- with(results_summary, paste0(Avg_Error, " (", SE_Error, ")"))
-results_summary <- subset(results_summary, select = -c(Avg_Error, SE_Error))
-colnames(results_summary) <- c("Dataset", "d", "Classifier", "Error")
+\begin{figure}
+<<error_rates>>=
+error_rates$data_set <- Hmisc:::capitalize(error_rates$data_set)
+p <- ggplot(error_rates, aes(x = d, y = Avg_Error, group = Classifier))
+p <- p + geom_line(aes(color = Classifier, linetype = Classifier), size = 1.5)
+p <- p + facet_wrap(~ data_set) + ylim(c(0, max(error_rates$Avg_Error))) + theme_bw()
+p <- p + xlab("Number of Variables Selected") + ylab("Average Error Rate")
+p
+@ 
+\caption{Average error-rate estimates of the competing classifiers as a function
+  of $d$. Approximate standard errors are no greater than \Sexpr{max(error_rates[["SE_Error"]])}.}
+\label{fig:error-rates}
+\end{figure}
 
-results_tables <- tapply(seq_along(results_summary$d), results_summary$d, function(i) {
-  dcast(results_summary[i, ], Dataset ~ Classifier, value.var = "Error")
-})
-
-results_xtables <- lapply(names(results_tables), function(d) {
-  tab_caption <- "The average of the test error rates for the microarray data sets with $d = "
-  tab_caption <- paste0(tab_caption, d, "$. Approximate standard errors are given in parentheses.")
-  xtable(results_tables[[d]], caption = tab_caption, label = paste0("tab:results-d", d))
-})
-
-lapply(results_xtables, print, include.rownames = FALSE)
-
+\begin{figure}
+<<optimal_estimates>>=
+grda_optimal$lambda <- factor(grda_optimal$lambda)
+grda_optimal$gamma <- factor(grda_optimal$gamma)
+p <- ggplot(subset(grda_optimal, data_set == "shipp" & d %in% c(100, 500, 1000)), aes(lambda))
+p <- p + geom_bar(aes(fill = gamma, weight = value))
+p <- p + facet_grid(d ~ .) + theme_bw()
+p
 @ 
+\caption{Counts of the optimal tuning-parameter estimates for the \emph{GRDA}
+  classifier for the Shipp data set for $d = 100$, $500$, and $1000$.}
+\label{fig:optimal-shipp}
+\end{figure}
 

### 2013-04-15 lib/helpers.R

@@ -65,6 +65,69 @@ round_digits <- function(x, num_digits = 3) {
   sprintf(paste("%.", num_digits, "f", sep = ""), round(x, num_digits))
 }
 
+#' Generates a 2x2 rotation matrix using the given angle
+#' #'
+#' #' @param theta numeric value giving the angle of rotation
+#' #' @return 2x2 rotation matrix
+rotation_matrix <- function(theta) {
+  matrix(c(cos(theta), -sin(theta), sin(theta), cos(theta)), nrow = 2, byrow = TRUE)
+}
+
+#' Creates a ggplot2 object containing the contours of the GRDA classifier as a
+#' function of the tuning parameter lambda
+#'
+#' @param lambda numeric value between 0 and 1, inclusively.
+#' @return a \code{ggplot2} object
+rda_contours <- function(lambda) {
+  mu1 <- c(-2, -2)
+  Sigma1 <- 3 * diag(2)
+
+  mu2 <- c(3, 3)
+  Sigma2 <- diag(2)
+  Sigma2 <- 3.5 * rotation_matrix(-pi/2) %*% (0.9^abs(row(Sigma2) - col(Sigma2))) %*% t(rotation_matrix(-pi/2))
+
+  mu3 <- c(4, -2)
+  Sigma3 <- diag(2)
+  Sigma3 <- 1 * (-0.5)^abs(row(Sigma3) - col(Sigma3))
+
+  mu4 <- c(-3, 3)
+  Sigma4 <- diag(2)
+  Sigma4 <- 0.5 * rotation_matrix(-pi/8) %*% (0.9)^abs(row(Sigma4) - col(Sigma4)) %*% t(rotation_matrix(-pi/8))
+
+  mu5 <- c(1, 1)
+  Sigma5 <- 0.1 * diag(2)
+
+  Sigma_pool <- 0.2 * (Sigma1 + Sigma2 + Sigma3 + Sigma4 + Sigma5)
+
+  Sigma1 <- (1 - lambda) * Sigma1 + lambda * Sigma_pool
+  Sigma2 <- (1 - lambda) * Sigma2 + lambda * Sigma_pool
+  Sigma3 <- (1 - lambda) * Sigma3 + lambda * Sigma_pool
+  Sigma4 <- (1 - lambda) * Sigma4 + lambda * Sigma_pool
+  Sigma5 <- (1 - lambda) * Sigma5 + lambda * Sigma_pool
+
[…]

### 2013-04-17 reports/paper-rda.Rnw
Renamed GRDA classifier to HDRDA. Also, applied edits from Dean.

@@
 results$Dudoit <- as.numeric(results$Dudoit)
-results$GRDA <- as.numeric(results$GRDA)
+results$HDRDA <- as.numeric(results$HDRDA)
 results$Guo <- as.numeric(results$Guo)
@@
 # of variables selected, d.
-grda_optimal <- subset(results, select = c(data_set, d, lambda, gamma))
-grda_optimal$lambda <- factor(grda_optimal$lambda, levels = seq_lambda)
-grda_optimal$gamma <- factor(grda_optimal$gamma, levels = seq_gamma)
+hdrda_optimal <- subset(results, select = c(data_set, d, lambda, gamma))
+hdrda_optimal$lambda <- factor(hdrda_optimal$lambda, levels = seq_lambda)
+hdrda_optimal$gamma <- factor(hdrda_optimal$gamma, levels = seq_gamma)
 
-grda_optimal <- ddply(grda_optimal, .(data_set, d), function(x) {
+hdrda_optimal <- ddply(hdrda_optimal, .(data_set, d), function(x) {
   x <- subset(x, select = c(lambda, gamma))
@@
 data is essential for successful diagnosis and treatment of diseases such as
-cancer, but the analysis of microarray data often presents difficult challenges
-because the number of gene expression levels is typically much greater than the
-sample size. Consequently, several sparse and regularized classifiers have been
-proposed, often with strict covariance assumptions or eigenvalue adjustments to
-ensure that the classifiers are well-posed. Here, we consider an improved
-version of the regularized discriminant analysis (\emph{RDA}) classifier
-popularized by \cite{Friedman:1989tm}, which requires computationally expensive
-matrix operations and involves a time-consuming model-selection procedure. We
-propose a novel modification of the \emph{RDA} (\emph{GRDA}) classifier that
-provides a straightforward interpration and employs a novel dimension-reduction
-technique to enable a computationally efficient model-selection procedure, thus
-overcoming the computational shortcomings of the \emph{RDA}
-classifier. Moreover, we demonstrate with several high-dimensional microarray
-data sets that our proposed classifier is competitive and often superior to
-several recently proposed sparse and regularized classifiers in terms of
-classification accuracy. Finally, we provide an implementation of the
-\emph{GRDA} classifier in the {\tt sparsediscrim} R package, available on CRAN.
+cancer; however, the analysis of microarray data often presents difficult
+challenges because the number of gene expression levels is typically much
+greater than the sample size. Consequently, several sparse and regularized
+classifiers have been proposed, often with strict covariance assumptions or
+eigenvalue adjustments to ensure that the classifiers are well-posed. Here, we
+consider an improved version of the regularized discriminant analysis
+(\emph{RDA}) classifier popularized by \cite{Friedman:1989tm}. That is, we
+propose the high-dimensional \emph{RDA} (\emph{HDRDA}) classifier, a novel
+modification of the \emph{RDA} classifier, that provides a straightforward
+interpretation and employs a new dimension-reduction technique to enable a
+computationally efficient model-selection procedure, thus overcoming the
+computational shortcomings of the \emph{RDA} classifier for high-dimensional
+data. Moreover, we demonstrate with several high-dimensional microarray data
+sets that our proposed classifier is competitive and often superior to several
+recently proposed sparse and regularized classifiers in terms of classification
+accuracy. Finally, we provide an implementation of the \emph{HDRDA} classifier in
+the {\tt sparsediscrim} R package, available on CRAN.
 
@@
 matrices are equal, which is often too stringent in practice
-\citep{Clemmensen:2011kr}. The \emph{GRDA} classifier allows us to relax the
+\citep{Clemmensen:2011kr}. The \emph{HDRDA} classifier allows us to relax the
 linearity assumption by including unequal covariance matrices in the model. The
-\emph{GRDA} classifier is then a compromise between the \emph{LDA} and
-\emph{QDA} classifiers, so that the \emph{GRDA} classifier includes both the
+\emph{HDRDA} classifier is then a compromise between the \emph{LDA} and
+\emph{QDA} classifiers, so that the \emph{HDRDA} classifier includes both the
 \emph{LDA} and \emph{QDA} classifiers as special cases. We then introduce a more
-general shrinkage term to the \emph{RDA} classifier to establish our \emph{GRDA}
+general shrinkage term to the \emph{RDA} classifier to establish our \emph{HDRDA}
 classifier.
 
-We show that the \emph{GRDA} classifier is invariant to any adjustments of the
+We show that the \emph{HDRDA} classifier is invariant to any adjustments of the
 eigenvalues corresponding to the null space of the pooled sample covariance
@@
 In addition to the improvement in computational performance, we demonstrate that
-our \emph{GRDA} classifier is competitive with and often superior to several
+our \emph{HDRDA} classifier is competitive with and often superior to several
 recent classifiers designed for small-sample, high-dimensional data. In
-particular, we compare our \emph{GRDA} classifier with the \emph{DLDA}
+particular, we compare our \emph{HDRDA} classifier with the \emph{DLDA}
 classifier from \cite{Dudoit:2002ev} and the recently proposed classifiers from
@@
 high-dimensional microarray data sets and then demonstrate that our proposed
-\emph{GRDA} classifier is often superior to the competing classifiers in terms
+\emph{HDRDA} classifier is often superior to the competing classifiers in terms
 of classification accuracy. We then conclude with a brief discussion in Section
@@
 
-\section{The GRDA Classifier}
+\section{The HDRDA Classifier}
 
@@
 weights in \eqref{eq:sig-lambda-alternative2} reduce to $1/N$, corresponding to
-$\widehat{\bm\Sigma}$. In Figure \ref{fig:grda-contours} we demonstrate the
-effect of $\lambda$ on the contours of five multivariate normal populations with
-unequal covariance matrices. For brevity, when $\lambda = 1$, we define $\bm X =
+$\widehat{\bm\Sigma}$. For brevity, when $\lambda = 1$, we define $\bm X =
 [\sqrt{c_{1k}(1)} \bm x_1', \ldots, \sqrt{c_{Nk}(1)} \bm x_N']'$ such that
@@
 $\bm X_k = [\sqrt{c_{1k}(0)} \bm x_1', \ldots, \sqrt{c_{Nk}(0)} \bm x_N']'$ such
-that $\widehat{\bm \Sigma}_k = \bm X_k' \bm X_k$, $k = 1, \ldots, K$.
+that $\widehat{\bm \Sigma}_k = \bm X_k' \bm X_k$, $k = 1, \ldots, K$. In Figure
+\ref{fig:hdrda-contours} we demonstrate the effect of $\lambda$ on the contours
+of five multivariate normal populations with unequal covariance matrices.
 
-\[ \left[\text{Figure \ref{fig:grda-contours} goes about here }\right]\]
+\[ \left[\text{Figure \ref{fig:hdrda-contours} goes about here }\right]\]
 
 As we have discussed, several eigenvalue adjustment methods have been proposed
-to increase eigenvalues (approximately) equal to 0. Here, we define a more
-general eigenvalue adjustment of \eqref{eq:sig-lambda-alternative} as
+to increase eigenvalues (approximately) equal to 0. Here, we define the
+eigenvalue adjustment of \eqref{eq:sig-lambda-alternative} as
 \begin{align}
-	\tilde{\bm \Sigma}_k = \alpha_k \widehat{\bm \Sigma}_k(\lambda) + \gamma \bm I_p,\label{eq:grda-cov}
+	\tilde{\bm \Sigma}_k = \alpha_k \widehat{\bm \Sigma}_k(\lambda) + \gamma \bm I_p,\label{eq:hdrda-cov}
 \end{align}
 where $\alpha_k \ge 0$ and $\gamma \ge 0$ is an eigenvalue-shrinkage
-constant. Thus, the \emph{GRDA} classifier is
+constant. The choice of $\alpha_k$ allows for a flexible formulation of
+covariance-matrix estimators. For instance, if $\alpha_k = 1$, $k = 1, \ldots,
+K$, then \eqref{eq:hdrda-cov} resembles \eqref{eq:ridge-estimator}. Similarly,
+if $\alpha_k = 1 - \gamma$, then \eqref{eq:hdrda-cov} has a form similar to
+\eqref{eq:sig-rda}. Now, substituting \eqref{eq:hdrda-cov} into \eqref{eq:qda},
+we define the \emph{HDRDA} classifier as
 \begin{align}
-	D_{GRDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\tilde{\bm\Sigma}_k^{+}(\bm x - \xbar_k)  + \ln |\tilde{\bm\Sigma}_k|. \label{eq:grda}
+	D_{HDRDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\tilde{\bm\Sigma}_k^{+}(\bm x - \xbar_k)  + \ln |\tilde{\bm\Sigma}_k|. \label{eq:hdrda}
 \end{align}
@@
 $\tilde{\bm\Sigma}_k^{-1}$ can be substituted for $\tilde{\bm\Sigma}_k^{+}$ in
-\eqref{eq:grda}. If $\gamma = 0$, we explicitly set $|\tilde{\bm\Sigma}_k|$
+\eqref{eq:hdrda}. If $\gamma = 0$, we explicitly set $|\tilde{\bm\Sigma}_k|$
 equal to the product of the positive eigenvalues of $\tilde{\bm\Sigma}_k$.
 
-Next, we demonstrate that the \emph{GRDA} classifier lends itself to an
-efficient calculation of \eqref{eq:grda} and a rapid model selection of the
-pooling parameter $\lambda$. We decompose \eqref{eq:grda} into a sum of two
+Next, we demonstrate that the \emph{HDRDA} classifier lends itself to an
+efficient calculation of \eqref{eq:hdrda} and a rapid model selection of the
+pooling parameter $\lambda$. We decompose \eqref{eq:hdrda} into a sum of two
 components, where the first summand consists of matrix operations applied to
@@
 adjustments to the associated eigenvalues have no effect on
-\eqref{eq:grda}. Lastly, we exploit the so-called fast singular value
+\eqref{eq:hdrda}. Lastly, we exploit the so-called fast singular value
 decomposition (SVD) to construct the eigenvalue decomposition of $\widehat{\bm
@@
 Thus, by incorporating each $\bm x_i$ into the estimation of $\bm \Sigma_k$, we
-increase the rank of $\widehat{\bm \Sigma}_k(\lambda)$ to $q$, $k = 1, \ldots,
-K$ if the pooling parameter $\lambda \ne 0$. Next, we provide an essential
-result that enables us to prove that, under mild regularity conditions,
-\eqref{eq:grda} is invariant to adjustments to the eigenvalues of $\tilde{\bm
-  \Sigma}_k$ corresponding to the null space of $\widehat{\bm \Sigma}$ for any
-unlabeled observation $\bm x \in \mathbb{R}_{p \times 1}$.
+increase the rank of $\widehat{\bm \Sigma}_k(\lambda)$ to $q$ if $\lambda \ne
+0$. Next, we provide an essential result that enables us to prove that, under
+mild regularity conditions, \eqref{eq:hdrda} is invariant to adjustments to the
+eigenvalues of $\tilde{\bm \Sigma}_k$ corresponding to the null space of
+$\widehat{\bm \Sigma}$ for any unlabeled observation $\bm x \in \mathbb{R}_{p
+  \times 1}$.
 
@@
 
-We now present our main result, where we decompose the \emph{GRDA} decision rule
+We now present our main result, where we decompose the \emph{HDRDA} decision rule
 and show that the component requiring the largest computational costs does not
 contribute to the classification of an unlabeled observation performed using
-Lemma \ref{lemma:RDA-constant-term}. Our main result reduces \eqref{eq:grda} to
-an equivalent, more computationally efficient decision rule.
+Lemma \ref{lemma:RDA-constant-term}. Our main result reduces \eqref{eq:hdrda} to
+an equivalent, more computationally efficient and more stable decision rule.
 
@@
 and \eqref{eq:Wk}, respectively, and let $\bm U_1$ be defined as above. Then,
-the \emph{GRDA} decision rule in \eqref{eq:grda} is equivalent to
+the \emph{HDRDA} decision rule in \eqref{eq:hdrda} is equivalent to
 	\begin{align}
-		D_{GRDA}(\bm x) &= \argmin_k  (\bm x - \xbar_k)' \bm U_1 \bm W_k^{-1} \bm U_1' (\bm x - \xbar_k) + \ln | \bm W_k |. \label{eq:rda-decomposed}
+		D_{HDRDA}(\bm x) &= \argmin_k  (\bm x - \xbar_k)' \bm U_1 \bm W_k^{-1} \bm U_1' (\bm x - \xbar_k) + \ln | \bm W_k |. \label{eq:rda-decomposed}
 	\end{align}
@@
 
-Thus far we have shown that the \emph{GRDA} classifier in
+Thus far we have shown that the \emph{HDRDA} classifier in
 \eqref{eq:rda-decomposed} is invariant to the term $\bm U_2$, thus yielding a
@@
 additional results that further improve the computational performance of our
-\emph{GRDA} classifier. First, we demonstrate that the inverse and determinants
+\emph{HDRDA} classifier. First, we demonstrate that the inverse and determinants
 of $\bm W_k$ can be written so that they are performed on $n_k \times n_k$
 matrices rather than the larger $q \times q$ matrices, resulting in further
-improvements to the computational efficiency of the \emph{GRDA} classifier.
+improvements to the computational efficiency of the \emph{HDRDA} classifier.
 
@@
 In this section, we describe four high-dimensional microarray data sets and
-compare our proposed \emph{GRDA} classifier with five recently proposed
+compare our proposed \emph{HDRDA} classifier with five recently proposed
 classifiers designed for small-sample, high-dimensional data: the \emph{DLDA}
@@
 default settings with the exception that we explicitly set the prior
-probabilities as equal. In our {\tt R} implementation of the \emph{GRDA}
+probabilities as equal. In our {\tt R} implementation of the \emph{HDRDA}
 classifier in the {\tt sparsediscrim} package, we set $\alpha_k = 1$, $k = 1,
-\ldots, K$ for simplicity. We estimated $\lambda$ from a grid of 21 equidistant
+\ldots, K$, so that the covariance-matrix estimator \eqref{eq:hdrda-cov}
+employed in the \emph{HDRDA} classifier has a form resembling
+\eqref{eq:ridge-estimator}. We estimated $\lambda$ from a grid of 21 equidistant
 candidate values between 0 and 1, inclusively. Similarly, we estimated $\gamma$
@@
 
-The \emph{GRDA} classifier was superior in classification performance for the
+The \emph{HDRDA} classifier was superior in classification performance for the
 majority of the simulations. For the Burczysnki, Shipp, and Singh data sets, the
-\emph{GRDA} classifier outperformed its competitors for all values of $d$
-examined. For the Nakayama data set, the \emph{GRDA} classifier was outperformed
-by the Dudoit, Pang, and Tong classifiers for small values of $d$ but attained
-classification rates superior to its competitors for larger values of
-$d$. Furthermore, Figure \ref{fig:error-rates} emphasizes two strengths of the
-\emph{GRDA} classifier: reliability of the classifier for small $d$ and
+\emph{HDRDA} classifier outperformed its competitors for all values of $d$
+examined. For the Nakayama data set, the \emph{HDRDA} classifier was
+outperformed by the Dudoit, Pang, and Tong classifiers for small values of $d$
+but attained classification rates superior to its competitors for larger values
+of $d$. Furthermore, Figure \ref{fig:error-rates} emphasizes two strengths of
+the \emph{HDRDA} classifier: reliability of the classifier for small $d$ and
 stability of the classifier for large $d$. In contrast to the other classifiers,
-the \emph{GRDA} classifier exhibited less variability its misclassification
-estimates for small values of $d$. Also, the \emph{GRDA} classifier yielded
+the \emph{HDRDA} classifier exhibited less variability in its misclassification
+estimates for small values of $d$. Also, the \emph{HDRDA} classifier yielded
 superior classification accuracy as $d$ increased and was hardly effected by the
 inclusion of more variables in the data sets. We attribute the apparent
-stability of the error-rate estimates of the \emph{GRDA} classifier for large
+stability of the error-rate estimates of the \emph{HDRDA} classifier for large
 values of $d$ to its novel dimension-reduction technique.
@@
 Next, we examine the optimal estimates of the tuning parameters for the
-\emph{GRDA} classifier. For brevity, we include only a graphical summary of the
+\emph{HDRDA} classifier. For brevity, we include only a graphical summary of the
 optimal estimates of $\lambda$ for the Shipp data set in Figure
@@
 distribution of the optimal values of $\lambda$ selected through
-cross-validation has a notable interesting bathtub shape, resembling a Beta
-distribution with shape parameters $1/2$ and $1/2$. Specifically, the optimal
-values of $\lambda$ are typically near 0 or near 1, corresponding to the
-\emph{QDA} and \emph{LDA} classifiers, respectively. We noticed a similar
-distribution of optimal values of $\lambda$ for the other three data sets.
+cross-validation has a notable bathtub shape, resembling a Beta distribution
+with both shape parameters less than unity. Specifically, the optimal values of
+$\lambda$ are typically near 0 or near 1, corresponding to the \emph{QDA} and
+\emph{LDA} classifiers, respectively. We noticed a similar distribution of
+optimal values of $\lambda$ for the other three data sets.
 
@@
 optimal values of $\lambda$ were selected as unity, thereby an increased number
-of variables resulted in the \emph{GRDA} classifier preferring the more
+of variables resulted in the \emph{HDRDA} classifier preferring the more
 parsimonious assumption of equal covariance matrices. In this instance we can
-see the flexibility of the \emph{GRDA} classifier to employ the more realistic
+see the flexibility of the \emph{HDRDA} classifier to employ the more realistic
 assumption that the class covariance matrices are unequal, but if optimal, we
@@
 collected responses. Consequently, several variants of the well-known \emph{LDA}
-classifier have been proposed by employing for improved classification
-performance. However, the \emph{LDA} classifier includes the strong assumption
-that the covariance matrices of each class are equal. Ideally, we could model
-the class covariance matrices as unrestricted and unequal in the \emph{QDA}
-classifier, but it is well known to have a large number of parameters that must
-be estimated and require a large number of samples, which are often unavailable
-in the high-dimensional setting. With this mind, \cite{Friedman:1989tm} proposed
-the \emph{RDA} classifier as a compromise between the \emph{LDA} and \emph{QDA}
+classifier have been proposed for improved classification performance. However,
+the \emph{LDA} classifier includes the strong assumption that the covariance
+matrices of each class are equal. Ideally, we could model the class covariance
+matrices as unrestricted and unequal in the \emph{QDA} classifier, but it is
+well known to have a large number of parameters that must be estimated and
+require a large number of samples, which are often unavailable in the
+high-dimensional setting. With this mind, \cite{Friedman:1989tm} proposed the
+\emph{RDA} classifier as a compromise between the \emph{LDA} and \emph{QDA}
 classifiers, controlling the amount of pooling of the individual class
@@
 
-With this in mind, we have adapted the \emph{RDA} classifier for usage with
-modern small-sample, high-dimensional data sets. Our model formulation
-facilitates a rapid classification of high-dimensional data by employing a novel
-dimension reduction technique. Moreover, our 10-fold cross-validation procedure
-typically requires a small runtime for data sets having large $p$. When
-preliminary variable selection is applied to the data as we performed in our
-classification study, the time to perform model selection is nearly
-instantaneous, confirming the computational complexity reported regarding the
-operations employed in the \emph{GRDA} classifier. Additionally, we have
-provided an implementation of the \emph{GRDA} classifier in the {\tt
+We have adapted the \emph{RDA} classifier for usage with modern small-sample,
+high-dimensional data sets. Our model formulation facilitates a rapid
+classification of high-dimensional data by employing a novel dimension reduction
+technique. Moreover, our 10-fold cross-validation procedure typically requires a
+small runtime for data sets having large $p$. When preliminary variable
+selection is applied to the data as we performed in our classification study,
+the time to perform model selection is nearly instantaneous. Additionally, we
+have provided an implementation of the \emph{HDRDA} classifier in the {\tt
   sparsediscrim} R package that is easy to use and available on CRAN.
@@
 We have parameterized the covariance-matrix regularization employed in the
-\emph{GRDA} classifier to include a large family of shrinkage methods, such as
+\emph{HDRDA} classifier to include a large family of shrinkage methods, such as
 the methods from \cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and a wide variety
 of other shrinkage methods studied by \cite{Ramey:2013ji} and
-\cite{Xu:2009fl}. Our formulation of the \emph{GRDA} classifier in terms of the
+\cite{Xu:2009fl}. Our formulation of the \emph{HDRDA} classifier in terms of the
 nonzero eigenvalues of the pooled sample covariance matrix indicates that
@@
 
-In our simulation study, we compared the \emph{GRDA} classifier to several
+In our simulation study, we compared the \emph{HDRDA} classifier to several
 recently proposed classifiers, including the Dudoit, Pang, Tong, and Guo
@@
 classification performance to other classification methods. We have shown that
-our proposed \emph{GRDA} classifier often yields superior classification
+our proposed \emph{HDRDA} classifier often yields superior classification
 accuracy to the diagonal classifiers considered, confirming the assertion of
-\cite{Mai:2012bf}. Furthermore, the \emph{GRDA} classifier is able to compete
-with the diagonal classifiers in terms of speed by incorporating a novel
-dimension-reduction technique. Although the diagonal classifiers are clearly
-faster to compute, the differences in runtime for small-sample, high-dimensional
-data sets is hardly noticeable.
+\cite{Mai:2012bf}. Furthermore, the \emph{HDRDA} classifier is able to compete
+with the diagonal classifiers in terms of speed by incorporating a new
+dimension-reduction technique.
 
@@
 \begin{figure}
-<<grda_contours, fig.width=8, fig.height=8, out.width='.49\\linewidth', fig.show='hold'>>=
+<<hdrda_contours, fig.width=8, fig.height=8, out.width='.49\\linewidth', fig.show='hold'>>=
 rda_contours(lambda = 0)
@@
   pooling parameter $\lambda$.}
-\label{fig:grda-contours}
+\label{fig:hdrda-contours}
 \end{figure}
@@
 p <- p + xlab("Number of Variables Selected") + ylab("Average Error Rate")
-p
+p  + theme(strip.text.x = element_text(size = 12))
 @ 
@@
 <<optimal_estimates>>=
-grda_optimal$lambda <- factor(grda_optimal$lambda)
-grda_optimal$gamma <- factor(grda_optimal$gamma)
-p <- ggplot(subset(grda_optimal, data_set == "shipp" & d %in% c(100, 500, 1000)), aes(lambda))
+hdrda_optimal$lambda <- factor(hdrda_optimal$lambda)
+hdrda_optimal$gamma <- factor(hdrda_optimal$gamma)
+p <- ggplot(subset(hdrda_optimal, data_set == "shipp" & d %in% c(100, 500, 1000)), aes(lambda))
 p <- p + geom_bar(aes(fill = gamma, weight = value))
-p <- p + facet_grid(d ~ .) + theme_bw()
-p
+p <- p + facet_grid(d ~ ., labeller = label_bquote(d == .(x))) + theme_bw()
+p + theme(strip.text.y = element_text(size = 16))
 @ 
-\caption{Counts of the optimal tuning-parameter estimates for the \emph{GRDA}
+\caption{Counts of the optimal tuning-parameter estimates for the \emph{HDRDA}
   classifier for the Shipp data set for $d = 100$, $500$, and $1000$.}

### 2013-04-17 lib/helpers.R

@@ -125,7 +125,7 @@ rda_contours <- function(lambda) {
   p <- p + facet_grid(. ~ lambda, labeller = label_bquote(lambda == .(x)))
   p <- p + theme_bw() + xlab("") + ylab("") + theme(legend.position = "none")
   p <- p + scale_x_continuous(breaks = NULL) + scale_y_continuous(breaks = NULL)
-  p  + theme(strip.text.x = element_text(size = 18))
+  p + theme(strip.text.x = element_text(size = 18))
 }
 
 # Classifier from Witten and Tibshirani (2011) - JRSS B

### 2013-04-17 sim-microarray.r

@@ -52,12 +52,12 @@ results <- mclapply(seq_len(nrow(sim_config)), function(i) {
                                              alpha, delta)), NA)
     Guo_errors <- mean(Guo_pred != test_y)
 
-    # GRDA
-    cv_out <- try_default(grda_cv(x = trn_x, y = train_y, prior = prior_probs), NA)
-    grda_lambda <- cv_out$lambda
-    grda_gamma <- cv_out$gamma
-    grda_out <- try_default(with(cv_out, grda(x = trn_x, y = train_y, lambda = lambda, gamma = gamma, prior = prior_probs)), NA)
-    grda_errors <- try_default(mean(predict(grda_out, tst_x)$class != test_y), NA)
+    # HDRDA
+    cv_out <- try_default(hdrda_cv(x = trn_x, y = train_y, prior = prior_probs), NA)
+    hdrda_lambda <- cv_out$lambda
+    hdrda_gamma <- cv_out$gamma
+    hdrda_out <- try_default(with(cv_out, hdrda(x = trn_x, y = train_y, lambda = lambda, gamma = gamma, prior = prior_probs)), NA)
+    hdrda_errors <- try_default(mean(predict(hdrda_out, tst_x)$class != test_y), NA)
 
     # DLDA
     dlda_errors <- try_default(mean(predict(dlda(x = trn_x, y = train_y, prior = prior_probs), tst_x)$class != test_y), NA)
@@ -71,12 +71,12 @@ results <- mclapply(seq_len(nrow(sim_config)), function(i) {
 
     list(errors = list(
            Dudoit = dlda_errors,
-           GRDA = grda_errors,
+           HDRDA = hdrda_errors,
            Guo = Guo_errors,
            Pang = sdlda_errors,
            Tong = Tong_errors,
            Witten = Witten_errors
-      ), data_set = data_set, d = d, lambda = grda_lambda, gamma = grda_gamma)
+      ), data_set = data_set, d = d, lambda = hdrda_lambda, gamma = hdrda_gamma)
   })
 }, mc.cores = num_cores)
 

### 2013-04-20 reports/paper-rda.Rnw
Updated the abstract and made some minor edits to the paper.

@@
 covariance matrices employed in both classifiers are singular. As a result, many
-researchers, including \cite*{Mai:2012bf}, \cite*{Merchante:2012vk},
-\cite*{Clemmensen:2011kr}, and \cite*{Huang:2010ju}, have introduced
-restrictions on the population covariance matrices in the \emph{LDA} and
-\emph{QDA} models to reduce number of covariance parameters that must be
-estimated to ensure that the resulting classifiers are calculable.  For
+researchers, including \cite*{Mai:2012bf}, \cite*{Fan:2012iq},
+\cite*{Merchante:2012vk}, \cite*{Clemmensen:2011kr}, and \cite*{Huang:2010ju},
+have introduced restrictions on the population covariance matrices in the
+\emph{LDA} and \emph{QDA} models to reduce number of covariance parameters that
+must be estimated to ensure that the resulting classifiers are calculable.  For
 instance, \cite{Dudoit:2002ev} have proposed the diagonal \emph{LDA}
@@
 \emph{QDA} classifiers and demonstrate that this simplification improves the
-interpretation of the pooling operation of the \emph{RDA} classifier. The
-\emph{LDA} classifier involves the assumption that the population covariance
-matrices are equal, which is often too stringent in practice
-\citep{Clemmensen:2011kr}. The \emph{HDRDA} classifier allows us to relax the
-linearity assumption by including unequal covariance matrices in the model. The
+interpretation of the pooling operation of the \emph{RDA} classifier.
+\cite{Clemmensen:2011kr} have argued that the assumption of equal population
+covariance matrices within the \emph{LDA} classifier is often too stringent in
+practice. The \emph{HDRDA} classifier allows us to relax the linearity
+assumption by including unequal covariance matrices in the model. The
 \emph{HDRDA} classifier is then a compromise between the \emph{LDA} and
 \emph{QDA} classifiers, so that the \emph{HDRDA} classifier includes both the
-\emph{LDA} and \emph{QDA} classifiers as special cases. We then introduce a more
-general shrinkage term to the \emph{RDA} classifier to establish our \emph{HDRDA}
-classifier.
+\emph{LDA} and \emph{QDA} classifiers as special cases. We then establish the
+\emph{HDRDA} classifier by applying covariance-matrix regularization with a
+parameterization that includes a large family of shrinkage methods, such as the
+methods from \cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and a wide variety of
+other shrinkage methods studied by \cite{Ramey:2013ji} and \cite{Xu:2009fl}.
 
@@
 eigenvalues. This invariance property addresses the computational shortcomings
-of the \emph{RDA} classifier and provides a substantial increase in the
-computational efficiency of the \emph{RDA} classifier and its model
-selection. In fact, we show that the inverse and determinants can be reduced to
-calculations of $n_k \times n_k$ matrices, where $n_k$ is the sample size of the
-$k$th class, $k = 1, \ldots, K$. This reduction in calculation is clearly
-advantageous because $n_k$ is often quite small, say, 15--20, in microarray data
-sets.
+of the \emph{RDA} classifier and provides a substantial increase in
+computational efficiency and model selection. In fact, we show that the inverse
+and determinants can be reduced to calculations of $n_k \times n_k$ matrices,
+where $n_k$ is the sample size of the $k$th class, $k = 1, \ldots, K$. This
+reduction in calculation is clearly advantageous because $n_k$ is often quite
+small, say, 15--20, in microarray data sets.
 
-In addition to the improvement in computational performance, we demonstrate that
-our \emph{HDRDA} classifier is competitive with and often superior to several
-recent classifiers designed for small-sample, high-dimensional data. In
-particular, we compare our \emph{HDRDA} classifier with the \emph{DLDA}
-classifier from \cite{Dudoit:2002ev} and the recently proposed classifiers from
-\cite{Guo:2007te}, \cite{Pang:2009ik}, \cite{Tong:2012hw}, and
-\cite{Witten:2011kc}. We conduct our comparison of the classification accuracy
-of the competing classifiers using four small-sample, high-dimensional
-microarray data sets.
+We also demonstrate that our \emph{HDRDA} classifier is competitive with and
+often superior to several recent classifiers designed for small-sample,
+high-dimensional data. In particular, we compare our \emph{HDRDA} classifier
+with the \emph{DLDA} classifier from \cite{Dudoit:2002ev} and the recently
+proposed classifiers from \cite{Guo:2007te}, \cite{Pang:2009ik},
+\cite{Tong:2012hw}, and \cite{Witten:2011kc}. We conduct our comparison of the
+classification accuracy of the competing classifiers using four small-sample,
+high-dimensional microarray data sets.
 
@@
 
-Here, we define our generalization of the \emph{RDA} classifier by first
-simplifying the formulation of $\widehat{\bm \Sigma}_k(\lambda)$ in
-\eqref{eq:sig-lambda} and demonstrating its clear interpretation as a linear
-combination of the crossproducts of the training observations centered by their
-respective class sample means. We define
+Here, we define our adaptation of the \emph{RDA} classifier by first simplifying
+the formulation of $\widehat{\bm \Sigma}_k(\lambda)$ in \eqref{eq:sig-lambda}
+and demonstrating its clear interpretation as a linear combination of the
+crossproducts of the training observations centered by their respective class
+sample means. We define
 \begin{align}
@@
 From \eqref{eq:sig-lambda-alternative2}, $\lambda$ weights the contribution of
-each of the $N$ observations to estimating $\bm \Sigma_k$. That is, for $0 <
+each of the $N$ observations to estimate $\bm \Sigma_k$. That is, for $0 <
 \lambda \le 1$, we estimate $\bm \Sigma_k$ with $N$ rather than $n_k$
@@
 that $\widehat{\bm \Sigma}_k = \bm X_k' \bm X_k$, $k = 1, \ldots, K$. In Figure
-\ref{fig:hdrda-contours} we demonstrate the effect of $\lambda$ on the contours
-of five multivariate normal populations with unequal covariance matrices.
+\ref{fig:hdrda-contours} we plot the contours of five multivariate normal
+populations with unequal covariance matrices and $\lambda = 0$. As $\lambda$ is
+increased towards 1, the contours become more similar, ultimately resulting in
+equal covariance matrices and contours for the parsimonious case of $\lambda =
+1$.
 
@@
 adjustments to the associated eigenvalues have no effect on
-\eqref{eq:hdrda}. Lastly, we exploit the so-called fast singular value
-decomposition (SVD) to construct the eigenvalue decomposition of $\widehat{\bm
-  \Sigma}$, further reducing the computational costs of our proposed \emph{RDA}
-classifier.
+\eqref{eq:hdrda}. Lastly, we utilize the singular value decomposition (SVD) to
+calculate efficiently the eigenvalue decomposition of $\widehat{\bm \Sigma}$,
+further reducing the computational costs of our proposed \emph{RDA} classifier.
 
@@
 
-We now present our main result, where we decompose the \emph{HDRDA} decision rule
-and show that the component requiring the largest computational costs does not
-contribute to the classification of an unlabeled observation performed using
-Lemma \ref{lemma:RDA-constant-term}. Our main result reduces \eqref{eq:hdrda} to
-an equivalent, more computationally efficient and more stable decision rule.
+We now present our main result, where we decompose the \emph{HDRDA} decision
+rule and show that the component requiring the largest computational costs does
+not contribute to the classification of an unlabeled observation performed using
+Lemma \ref{lemma:RDA-constant-term}. Hence, we reduce \eqref{eq:hdrda} to an
+equivalent, more computationally efficient decision rule.
 
@@
 	\begin{align}
-		D_{HDRDA}(\bm x) &= \argmin_k  (\bm x - \xbar_k)' \bm U_1 \bm W_k^{-1} \bm U_1' (\bm x - \xbar_k) + \ln | \bm W_k |. \label{eq:rda-decomposed}
+		D_{HDRDA}(\bm x) &= \argmin_k  (\bm x - \xbar_k)' \bm U_1 \bm W_k^{-1} \bm U_1' (\bm x - \xbar_k) + \ln | \bm W_k |. \label{eq:hdrda-decomposed}
 	\end{align}
@@
 
-Thus far we have shown that the \emph{HDRDA} classifier in
-\eqref{eq:rda-decomposed} is invariant to the term $\bm U_2$, thus yielding a
-substantial reduction in the computational costs. While the classifier is much
-improved in terms of computational performance compared to the \emph{RDA}
-classifier originally proposed by \cite{Friedman:1989tm}, we provide two
-additional results that further improve the computational performance of our
-\emph{HDRDA} classifier. First, we demonstrate that the inverse and determinants
-of $\bm W_k$ can be written so that they are performed on $n_k \times n_k$
-matrices rather than the larger $q \times q$ matrices, resulting in further
-improvements to the computational efficiency of the \emph{HDRDA} classifier.
+Thus far, we have shown that the \emph{HDRDA} classifier in \eqref{eq:hdrda} is
+invariant to the term $\bm U_2$, thus yielding an equivalent classifier in
+\eqref{eq:hdrda-decomposed} with a substantial reduction in the computational
+complexity. While the classifier is much improved in terms of computational
+performance compared to the \emph{RDA} classifier originally proposed by
+\cite{Friedman:1989tm}, we provide two additional results that further improve
+the computational performance of our \emph{HDRDA} classifier. First, we
+demonstrate that the inverse and determinants of $\bm W_k$ can be written so
+that they are performed on $n_k \times n_k$ matrices rather than the larger $q
+\times q$ matrices, resulting in further improvements to the computational
+efficiency.
 
@@
 \cite{Tong:2012hw}, and the penalized \emph{LDA} classifier from
-\cite{Witten:2011kc}. Below, we refer to each classifier by the first author's
-surname. We evaluated the classification performance of each classifier by
-randomly partitioning the data such that $2/3$ of the observations were
-allocated as training data and the remaining $1/3$ of the observations were
-allocated as a test data set. For each random partition, we applied the
-feature-selection method from \cite{Dudoit:2002ev} to choose the $d$ features
-having the largest ratio of their between-group to within-group sums of squares
-from the training data. We then filtered the test data set with the
-corresponding $d$ features. After training each classifier from the filtered
-training data, we classified the filtered test data sets and computed the
-proportion of mislabeled test observations to estimate the classification error
-rate for each classifier. Repeating this process 1000 times, we computed the
-average of the error-rate estimates for each classifier. We remark that our
-simulation design avoids the selection bias discussed in
-\cite{Ambroise:2002fa}. We did not include the \emph{RDA} classifier from
-\cite{Friedman:1989tm} in our studies to avoid its excessive computations. In
-Table \ref{tab:data-summary}, we summarize the attributes of each microarray
-data set.
+\cite{Witten:2011kc}. The classifiers from \cite{Tong:2012hw} and
+\cite{Pang:2009ik} are both modifications of the \emph{DLDA} classifier from
+\cite{Dudoit:2002ev}, where the former employs an improved mean estimator and
+the latter utilizes an improved variance estimator. Below, we refer to each
+classifier by the first author's surname.
+
+We evaluated the classification performance of each classifier by randomly
+partitioning the data such that $2/3$ of the observations were allocated as
+training data and the remaining $1/3$ of the observations were allocated as a
+test data set. For each random partition, we applied the feature-selection
+method from \cite{Dudoit:2002ev} to the training data to choose the $d$ features
+having the largest ratio of their between-group to within-group sums of squares.
+We then filtered the test data set with the corresponding $d$ features. After
+training each classifier from the filtered training data, we classified the
+filtered test data sets and computed the proportion of mislabeled test
+observations to estimate the classification error rate for each
+classifier. Repeating this process 1000 times, we computed the average of the
+error-rate estimates for each classifier. We remark that our simulation design
+avoids the selection bias discussed in \cite{Ambroise:2002fa}. We did not
+include the \emph{RDA} classifier from \cite{Friedman:1989tm} in our studies to
+avoid its excessive computations. In Table \ref{tab:data-summary}, we summarize
+the attributes of each microarray data set.
 
@@
 \ref{fig:optimal-shipp}. Similar figures for the other three data sets can be
-found in the Supplementary Information. In Figure \ref{fig:optimal-shipp}, the
+found in the Supplementary Material. In Figure \ref{fig:optimal-shipp}, the
 distribution of the optimal values of $\lambda$ selected through
@@
 classifiers and the reduction in the number of parameters to estimate,
-\cite{Mai:2012bf} have noted that diagonal classifiers can often yield inferior
-classification performance to other classification methods. We have shown that
-our proposed \emph{HDRDA} classifier often yields superior classification
-accuracy to the diagonal classifiers considered, confirming the assertion of
-\cite{Mai:2012bf}. Furthermore, the \emph{HDRDA} classifier is able to compete
+\cite{Mai:2012bf} and \cite{Fan:2012iq} have noted that diagonal classifiers
+can often yield inferior classification performance to other classification
+methods. We have shown that our proposed \emph{HDRDA} classifier often yields
+superior classification accuracy to the diagonal classifiers considered,
+confirming the assertions of \cite{Mai:2012bf} and
+\cite{Fan:2012iq}. Furthermore, the \emph{HDRDA} classifier is able to compete
 with the diagonal classifiers in terms of speed by incorporating a new

### 2013-04-21 reports/paper-rda.Rnw
Updated paper to match the majority of Biometrika's requirements. Still remaining: tweak notation and update figures.

@@
 
+# To abide by Biometrika's requirements of no abbreviations, we rename the
+# results of our HDRDA classifier as simply "Ours."
+results$Ours <- results$HDRDA
+results <- subset(results, select = -HDRDA)
+
 results$d <- factor(results$d, labels = sort(unique(as.numeric(results$d))))
@@
 m_error <- melt(error_rates, variable.name = "Classifier", value.name = "Error")
+m_error$Classifier <- factor(m_error$Classifier)
 error_rates <- ddply(m_error, .(data_set, d, Classifier), plyr:::summarize,
@@
 eigenvalue adjustments to ensure that the classifiers are well-posed. Here, we
-consider an improved version of the regularized discriminant analysis
-(\emph{RDA}) classifier popularized by \cite{Friedman:1989tm}. That is, we
-propose the high-dimensional \emph{RDA} (\emph{HDRDA}) classifier, a novel
-modification of the \emph{RDA} classifier, that provides a straightforward
-interpretation and employs a new dimension-reduction technique to enable a
-computationally efficient model-selection procedure, thus overcoming the
-computational shortcomings of the \emph{RDA} classifier for high-dimensional
-data. Moreover, we demonstrate with several high-dimensional microarray data
-sets that our proposed classifier is competitive and often superior to several
-recently proposed sparse and regularized classifiers in terms of classification
-accuracy. Finally, we provide an implementation of the \emph{HDRDA} classifier in
-the {\tt sparsediscrim} R package, available on CRAN.
+propose a novel modification of the regularized discriminant analysis classifier
+popularized by \cite{Friedman:1989tm}. Our proposed classifier provides a
+straightforward interpretation and employs a new dimension-reduction technique
+to enable a computationally efficient model-selection procedure, thus overcoming
+the computational shortcomings of \citeauthor{Friedman:1989tm}'s classifier for
+high-dimensional data. Moreover, we demonstrate with several high-dimensional
+microarray data sets that our proposed classifier is competitive and often
+superior to several recently proposed sparse and regularized classifiers in
+terms of classification accuracy. Finally, we provide an implementation of our
+proposed classifier in the {\tt sparsediscrim} R package, available on CRAN.
 
@@
 exhibit poor classification performance because of the \emph{curse of
-  dimensionality} \citep{Bellman:1961tn}. Furthermore, well-known classifiers,
-including linear discriminant analysis \emph{(LDA)} and quadratic discriminant
-analysis \emph{(QDA)}, are incalculable when $p > N$ because the sample
-covariance matrices employed in both classifiers are singular. As a result, many
-researchers, including \cite*{Mai:2012bf}, \cite*{Fan:2012iq},
-\cite*{Merchante:2012vk}, \cite*{Clemmensen:2011kr}, and \cite*{Huang:2010ju},
-have introduced restrictions on the population covariance matrices in the
-\emph{LDA} and \emph{QDA} models to reduce number of covariance parameters that
-must be estimated to ensure that the resulting classifiers are calculable.  For
-instance, \cite{Dudoit:2002ev} have proposed the diagonal \emph{LDA}
-(\emph{DLDA}) classifier, a modification of the \emph{LDA} classifier where the
-class covariance matrices are assumed to be diagonal.
+  dimensionality} \citep{Bellman:1961tn}. Furthermore, some well-known
+classifiers, such as linear and quadratic discriminant analysis, are
+incalculable when $p > N$ because the sample covariance matrices employed are
+singular. As a result, many researchers, including \cite{Mai:2012bf},
+\cite{Fan:2012iq}, \cite{Merchante:2012vk}, \cite{Clemmensen:2011kr}, and
+\cite{Huang:2010ju}, have introduced restrictions on the population covariance
+matrices in linear and quadratic discriminant analysis to reduce number of
+covariance parameters that must be estimated to ensure that the resulting
+classifiers are calculable. For instance, \cite{Dudoit:2002ev} have modified the
+linear and quadratic discriminant analysis classifiers by assuming that the
+population covariance matrices are diagonal, resulting in classifiers that are
+special cases of the Naive Bayes classifier \citep[Chapter 4]{Hastie:2008dt}.
 
@@
 regularization methods to improve the estimation of the population covariance
-matrices before applying the \emph{LDA} and \emph{QDA}
-classifiers. Covariance-matrix regularization methods often are based on the
+matrices before applying either linear or quadratic discriminant
+analysis. Covariance-matrix regularization methods often are based on the
 well-known ridge-regression approach of \cite{Hoerl:1970cd} and adjust the
@@
 
-Here, we propose a generalization of the popular \emph{RDA} classifier of
-\cite{Friedman:1989tm}, who has proposed pooling of the covariance-matrix
-estimators from the \emph{LDA} and \emph{QDA} classifiers and has additionally
-incorporated a covariance-matrix shrinkage term to ensure that the \emph{RDA}
-classifier is well-posed. \cite{Friedman:1989tm} has shown that the \emph{RDA}
-classifier can attain improved classification accuracy, compared to the
-\emph{LDA} and \emph{QDA} classifiers when $p < N$. However, in its original
-form, the \emph{RDA} classifier is impractical when $p \gg N$ because the
-\emph{RDA} decision rule requires the calculation of the inverses and
-determinants of the $p \times p$ covariance-matrix estimators for each
-class. Furthermore, the \emph{RDA} classifier has two tuning parameters that are
+Here, we propose an improvement to the popular regularized discriminant
+analysis classifier of \cite{Friedman:1989tm}, who has proposed pooling of the
+covariance-matrix estimators from linear and quadratic discriminant analysis and
+has additionally incorporated a covariance-matrix shrinkage term to ensure that
+the classifier is well-posed. \citeauthor{Friedman:1989tm}'s classifier can
+attain improved classification accuracy, compared to linear and quadratic
+discriminant analysis when $p < N$. However, in its original form,
+\citeauthor{Friedman:1989tm}'s classifier is impractical when $p \gg N$ because
+the decision rule requires the calculation of the inverses and determinants of
+large $p \times p$ covariance-matrix estimators for each class. Furthermore,
+\citeauthor{Friedman:1989tm}'s classifier has two tuning parameters that are
 selected via cross-validation, which adds a substantial computational burden to
-the model-selection process. The \emph{RDA} classifier is ill-advised on modern
-data sets because the model selection requires $O(Kp^3)$ calculations for each
-tuning-parameter pair.
+the model-selection process. \citeauthor{Friedman:1989tm}'s classifier is
+ill-advised on modern data sets because the model selection requires $O(Kp^3)$
+calculations for each tuning-parameter pair.
 
-In this paper we modify the \emph{RDA} classifier by first simplifying the
-pooling matrix employed in the \emph{RDA} classifier. We construct a convex
-combination of the covariance-matrix estimators employed in the \emph{LDA} and
-\emph{QDA} classifiers and demonstrate that this simplification improves the
-interpretation of the pooling operation of the \emph{RDA} classifier.
-\cite{Clemmensen:2011kr} have argued that the assumption of equal population
-covariance matrices within the \emph{LDA} classifier is often too stringent in
-practice. The \emph{HDRDA} classifier allows us to relax the linearity
-assumption by including unequal covariance matrices in the model. The
-\emph{HDRDA} classifier is then a compromise between the \emph{LDA} and
-\emph{QDA} classifiers, so that the \emph{HDRDA} classifier includes both the
-\emph{LDA} and \emph{QDA} classifiers as special cases. We then establish the
-\emph{HDRDA} classifier by applying covariance-matrix regularization with a
-parameterization that includes a large family of shrinkage methods, such as the
-methods from \cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and a wide variety of
-other shrinkage methods studied by \cite{Ramey:2013ji} and \cite{Xu:2009fl}.
+In this paper we modify \citeauthor{Friedman:1989tm}'s classifier by first
+constructing the pooling matrix as a convex combination of the covariance-matrix
+estimators employed in linear and quadratic discriminant analysis and
+demonstrate that this simplification improves the interpretation of the pooling
+operation of \citeauthor{Friedman:1989tm}'s classifier. Given that the
+assumption of equal population covariance matrices within linear discriminant
+analysis is often too stringent in practice \citep{Clemmensen:2011kr}, our
+proposed classifier can relax the linearity assumption by instead incorporating
+unequal covariance matrices while retaining linear and quadratic discriminant
+analysis as special cases. We then establish our classifier by applying
+covariance-matrix regularization with a parameterization that includes a large
+family of shrinkage methods, such as those from \cite{Srivastava:2007ww},
+\cite{Rao:1971ul}, and other methods studied by \cite{Ramey:2013ji} and
+\cite{Xu:2009fl}.
 
-We show that the \emph{HDRDA} classifier is invariant to any adjustments of the
+We show that our proposed classifier is invariant to any adjustments of the
 eigenvalues corresponding to the null space of the pooled sample covariance
 matrix. Applying a reasoning similar to \cite{Ye:2006tq}, we decompose the
-\emph{RDA} classifier and show that the sum related to the null space of the
-pooled sample covariance matrix can be discarded. Hence, no gain in
+decision rule of our classifier and show that the sum related to the null space
+of the pooled sample covariance matrix can be discarded. Hence, no gain in
 classification performance can be obtained by adjusting these zero
 eigenvalues. This invariance property addresses the computational shortcomings
-of the \emph{RDA} classifier and provides a substantial increase in
-computational efficiency and model selection. In fact, we show that the inverse
-and determinants can be reduced to calculations of $n_k \times n_k$ matrices,
-where $n_k$ is the sample size of the $k$th class, $k = 1, \ldots, K$. This
-reduction in calculation is clearly advantageous because $n_k$ is often quite
-small, say, 15--20, in microarray data sets.
+of \citeauthor{Friedman:1989tm}'s classifier and provides a substantial increase
+in computational efficiency and model selection. In fact, we show that the
+inverse and determinants can be reduced to calculations of $n_k \times n_k$
+matrices, where $n_k$ is the sample size of the $k$th class $(k = 1, \ldots,
+K)$. This reduction in calculation is clearly advantageous because $n_k$ is
+often quite small in microarray data sets (e.g., 15--20 observations).
 
-We also demonstrate that our \emph{HDRDA} classifier is competitive with and
-often superior to several recent classifiers designed for small-sample,
-high-dimensional data. In particular, we compare our \emph{HDRDA} classifier
-with the \emph{DLDA} classifier from \cite{Dudoit:2002ev} and the recently
-proposed classifiers from \cite{Guo:2007te}, \cite{Pang:2009ik},
-\cite{Tong:2012hw}, and \cite{Witten:2011kc}. We conduct our comparison of the
-classification accuracy of the competing classifiers using four small-sample,
-high-dimensional microarray data sets.
-
-We have organized the remainder of the paper as follows. In Section 2 we review
-the \emph{RDA} classifier from \cite{Friedman:1989tm} and discuss our
-generalization in Section 3. In Section 4 we describe four small-sample,
-high-dimensional microarray data sets and then demonstrate that our proposed
-\emph{HDRDA} classifier is often superior to the competing classifiers in terms
-of classification accuracy. We then conclude with a brief discussion in Section
-5.
+We also demonstrate that our proposed classifier is competitive with and often
+superior to several recent classifiers designed for small-sample,
+high-dimensional data. In particular, we compare our classifier with diagonal
+linear discriminant analysis from \cite{Dudoit:2002ev} and the recently proposed
+classifiers from \cite{Guo:2007te}, \cite{Pang:2009ik}, \cite{Tong:2012hw}, and
+\cite{Witten:2011kc}. We conduct our comparison of the classification accuracy
+of the competing classifiers using four small-sample, high-dimensional
+benchmark microarray data sets.
 
@@
 
+\subsection{Preliminaries}
 In discriminant analysis we wish to assign correctly an unlabeled
@@
 p(\bm x | \omega_k) p(\omega_k)$, where $p(\bm x | \omega_k)$ is the probability
-density function (PDF) of the $k$th class, $p(\omega_k)$ is prior probability of
-class membership of the $k$th class.
+density function of the $k$th class and $p(\omega_k)$ is the prior probability
+of class membership of the $k$th class. We further assume that each of
+$p(\omega_k)$ is equal $(k = 1, \ldots, K)$.
 
-The \emph{QDA} classifier is the optimal Bayesian decision rule with respect to
-a $0-1$ loss function when $p(\bm x | \omega_k)$ is the PDF of the multivariate
-normal distribution with known mean vector $\bm\mu_k \in \mathbb{R}_{p \times
-  1}$ and known covariance matrix $\bm\Sigma_k \in \mathbb{R}_{p \times p}^{>}$,
-$k = 1, 2, \ldots, K$, where $\mathbb{R}_{p \times p}^{>}$ denotes the cone of
-real $p \times p$ positive-definite matrices. Because the parameters are
-typically unknown, we estimate the unknown parameters $\bm \mu_k$ and
-$\bm\Sigma_k$ with their maximum likelihood estimators (MLEs) and substitute the
-MLEs into the \emph{QDA} classifier. Assuming that the prior probabilities of
-class membership $p(\omega_k)$ are equal for $k = 1, \ldots, K$, we assign an
-unlabeled observation $\bm x$ to class $\omega_k$ with the sample \emph{QDA}
-classifier
+Quadratic discriminant analysis is the optimal Bayesian decision rule with
+respect to a $0-1$ loss function when $p(\bm x | \omega_k)$ is the probability
+density function of the multivariate normal distribution with known mean vector
+$\bm\mu_k \in \mathbb{R}_{p \times 1}$ and known covariance matrix $\bm\Sigma_k
+\in \mathbb{R}_{p \times p}^{>}$, $k = 1, 2, \ldots, K$, where $\mathbb{R}_{p
+  \times p}^{>}$ denotes the cone of real $p \times p$ positive-definite
+matrices. Because $\bm \mu_k$ and $\bm\Sigma_k$ are typically unknown, we assign
+an unlabeled observation $\bm x$ to class $\omega_k$ with the sample quadratic
+discriminant analysis classifier
 \begin{align}
-	D_{QDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\widehat{\bm\Sigma}_k^{-1}(\bm x - \xbar_k)  + \ln |\widehat{\bm\Sigma}_k|, \label{eq:qda}
+	D_{QDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\widehat{\bm\Sigma}_k^{-1}(\bm x - \xbar_k)  + \log |\widehat{\bm\Sigma}_k|, \label{eq:qda}
 \end{align}
-where $\xbar_k$ and $\widehat{\bm\Sigma}_k$ are the MLEs for $\bm \mu_k$ and
-$\bm \Sigma_k$, respectively. If we assume further that the population
-covariance matrices are equal for each class (i.e., $\bm\Sigma_k = \bm\Sigma$
-for all $k$), then the pooled sample covariance matrix $\widehat{\bm\Sigma}$ is
-substituted for $\widehat{\bm \Sigma}_k$ in \eqref{eq:qda}, where
+where $\xbar_k$ and $\widehat{\bm\Sigma}_k$ are the maximum likelihood
+estimators for $\bm \mu_k$ and $\bm \Sigma_k$, respectively. If we assume
+further that $\bm\Sigma_k = \bm\Sigma$ $(k = 1, \ldots, K)$, then the pooled
+sample covariance matrix $\widehat{\bm\Sigma}$ is substituted for $\widehat{\bm
+  \Sigma}_k$ in \eqref{eq:qda}, where
 \begin{align}
-	\widehat{\bm\Sigma} = \frac{1}{N} \sum_{k=1}^K n_k \widehat{\bm\Sigma}_k. \label{eq:pooled-cov}
+	\widehat{\bm\Sigma} = N^{-1} \sum_{k=1}^K n_k \widehat{\bm\Sigma}_k \label{eq:pooled-cov}
 \end{align}
-is the MLE for $\bm \Sigma$. In this case \eqref{eq:qda} reduces to Fisher's
-\emph{LDA} classifier, and the log-determinant is omitted because it is constant
-across all classes.
+is the maximum likelihood estimator for $\bm \Sigma$. In this case
+\eqref{eq:qda} reduces to the sample linear discriminant analysis classifier,
+and the log-determinant is omitted because it is constant across all classes.
 
-We also require the following additional notation throughout the document. Let
+Throughout the document we also require the following additional notation. Let
 $\bm A \oplus \bm B$ denote the direct sum of $\bm A \in \mathbb{R}_{r \times
@@
 bias-variance tradeoff. A common form of the covariance-matrix regularization
-applies a shrinkage factor $\gamma_k > 0$, so that
+applies a shrinkage factor $\gamma > 0$, so that
 \begin{align}
-	\widehat{\bm \Sigma}_k(\gamma_k) =  \widehat{\bm \Sigma}_k + \gamma_k \bm I_p, \label{eq:ridge-estimator}
+	\widehat{\bm \Sigma}_k(\gamma) =  \widehat{\bm \Sigma}_k + \gamma \bm I_p, \label{eq:ridge-estimator}
 \end{align}
@@
 matrix $\widehat{\bm\Sigma}_k$ towards $\bm I_p$, thereby increasing the
-eigenvalues of $\widehat{\bm\Sigma}_k$ by $\gamma_k$. Specifically, the zero
-eigenvalues are replaced with $\gamma_k$, so that \eqref{eq:ridge-estimator} is
+eigenvalues of $\widehat{\bm\Sigma}_k$ by $\gamma$. Specifically, the zero
+eigenvalues are replaced with $\gamma$, so that \eqref{eq:ridge-estimator} is
 positive definite. For additional covariance-matrix regularization methods, see
@@
 
-\cite{Friedman:1989tm} has proposed the \emph{RDA} classifier by incorporating a
-weighted average of the sample covariance matrix $\widehat{\bm \Sigma}_k$ for
-class $\omega_k$ and the pooled sample covariance matrix $\widehat{\bm\Sigma}$
-to estimate the covariance matrix for class $\omega_k$ with
+\cite{Friedman:1989tm} has proposed regularized discriminant analysis by
+incorporating a weighted average of the sample covariance matrix $\widehat{\bm
+  \Sigma}_k$ and the pooled sample covariance matrix $\widehat{\bm\Sigma}$ to
+estimate $\bm \Sigma_k$ with
 \begin{align}
-  \widehat{\bm\Sigma}_k(\lambda) = n_k^{-1}(\lambda) \bm S_k(\lambda),\label{eq:sig-lambda}
+  \widehat{\bm\Sigma}_k(\lambda) = n_k^{-1}(\lambda) \bm S_k(\lambda) \quad (k =
+  1, \ldots, K),\label{eq:sig-lambda}
 \end{align}
@@
 S_k(\lambda) = (1 - \lambda) \bm S_k + \lambda \bm S$. We can interpret
-\eqref{eq:sig-lambda} as a covariance matrix estimator for class $\omega_k$ that
-borrows from $\widehat{\bm\Sigma}$ in \eqref{eq:pooled-cov} to better estimate
-$\bm \Sigma_k$. If $\lambda = 0$, \eqref{eq:sig-lambda} is equal to
-$\widehat{\bm\Sigma}_k$ used in the \emph{QDA} classifier in
-\eqref{eq:qda}. Alternatively, if $\lambda = 1$, \eqref{eq:sig-lambda} reduces
-to \eqref{eq:pooled-cov} used in the \emph{LDA} classifier.
+\eqref{eq:sig-lambda} as a covariance-matrix estimator that borrows from
+$\widehat{\bm\Sigma}$ in \eqref{eq:pooled-cov} to better estimate $\bm
+\Sigma_k$. If $\lambda = 0$, \eqref{eq:sig-lambda} is equal to
+$\widehat{\bm\Sigma}_k$ used in \eqref{eq:qda}. Alternatively, if $\lambda = 1$,
+\eqref{eq:sig-lambda} reduces to \eqref{eq:pooled-cov}.
 
@@
 of \eqref{eq:sig-lambda}, \cite{Friedman:1989tm} has proposed the biased
-\emph{RDA} covariance-matrix estimator
+covariance-matrix estimator
 \begin{align}
@@
 \end{align}
-for class $\omega_k$, where $\gamma \in [0, 1]$ is a regularization parameter
-that controls the shrinkage of \eqref{eq:sig-rda} towards $\bm I_p$, which is
-weighted by the the average of the eigenvalues of \eqref{eq:sig-lambda}. Thus,
-the \emph{pooling} parameter $\lambda$ controls the amount that we borrow from
-$\widehat{\bm\Sigma}$ to estimate $\bm \Sigma_k$, and the \emph{shrinkage}
-parameter $\gamma$ determines the amount of shrinkage applied.
+where $\gamma \in [0, 1]$ is a regularization parameter that controls the
+shrinkage of \eqref{eq:sig-rda} towards $\bm I_p$ and is weighted by the average
+of the eigenvalues of \eqref{eq:sig-lambda}. Thus, the \emph{pooling} parameter
+$\lambda$ controls the amount borrowed from $\widehat{\bm\Sigma}$ to estimate
+$\bm \Sigma_k$, and the \emph{shrinkage} parameter $\gamma$ determines the
+amount of eigenvalue shrinkage applied. Finally, by substituting
+\eqref{eq:sig-rda} into \eqref{eq:qda}, we have the regularized discriminant
+analysis classifier
+\begin{align}
+	D_{RDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\widehat{\bm\Sigma}_k(\lambda, \gamma)^{-1}(\bm x - \xbar_k)  + \log |\widehat{\bm\Sigma}_k(\lambda, \gamma)|. \label{eq:rda}
+\end{align}
+If $\gamma$ is sufficiently small, \eqref{eq:sig-rda} can be singular or
+near-singular. For this case \cite{Friedman:1989tm} has suggested replacing the
+zero eigenvalues of \eqref{eq:sig-rda} with a small positive constant that
+permits a numerically stable matrix inversion.
 
-\cite{Friedman:1989tm} substituted \eqref{eq:sig-rda} into \eqref{eq:qda} to
-obtain the \emph{RDA} classifier
+\section{High-Dimensional Regularized Discriminant Analysis}
+
+\subsection{Improvements to \citeauthor{Friedman:1989tm}'s Classifier}
+
+Here, we define our adaptation of \citeauthor{Friedman:1989tm}'s classifier by
+first simplifying the formulation of $\widehat{\bm \Sigma}_k(\lambda)$ in
+\eqref{eq:sig-lambda} and demonstrating its clear interpretation as a linear
+combination of the crossproducts of the training observations centered by their
+respective class sample means. We define the convex combination
 \begin{align}
-	D_{RDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\widehat{\bm\Sigma}_k(\lambda, \gamma)^{-1}(\bm x - \xbar_k)  + \ln |\widehat{\bm\Sigma}_k(\lambda, \gamma)|. \label{eq:rda}
+  \widehat{\bm\Sigma}_k(\lambda) = (1 - \lambda) \widehat{\bm\Sigma}_k + \lambda
+  \widehat{\bm\Sigma}, \quad (k = 1, \ldots, K)\label{eq:sig-lambda-alternative}
 \end{align}
-
-\section{The HDRDA Classifier}
-
-Here, we define our adaptation of the \emph{RDA} classifier by first simplifying
-the formulation of $\widehat{\bm \Sigma}_k(\lambda)$ in \eqref{eq:sig-lambda}
-and demonstrating its clear interpretation as a linear combination of the
-crossproducts of the training observations centered by their respective class
-sample means. We define
-\begin{align}
-  \widehat{\bm\Sigma}_k(\lambda) = (1 - \lambda) \widehat{\bm\Sigma}_k + \lambda \widehat{\bm\Sigma}.\label{eq:sig-lambda-alternative}
-\end{align}
-as a convex combination of the MLEs of $\bm \Sigma_k$ and $\bm \Sigma$, where
-$\lambda \in [0, 1]$ is the \emph{pooling} parameter. Thus, we avoid the convex
-combination of $n_k$ and $N$ as defined by \cite{Friedman:1989tm}. By rewriting
-\eqref{eq:sig-lambda-alternative} in terms of the observations $\bm x_i$, $i =
-1, \ldots, N$, each centered by its class sample mean, we attain a clear
-interpretation of $\widehat{\bm\Sigma}_k(\lambda)$. We have that
+where $\lambda \in [0, 1]$ is the \emph{pooling} parameter. Thus, we avoid the
+convex combination of $n_k$ and $N$ as defined by \cite{Friedman:1989tm}. By
+rewriting \eqref{eq:sig-lambda-alternative} in terms of the observations $\bm
+x_i$ $(i = 1, \ldots, N)$, each centered by its class sample mean, we attain a
+clear interpretation of $\widehat{\bm\Sigma}_k(\lambda)$. That is,
 \begin{align}
@@
 From \eqref{eq:sig-lambda-alternative2}, $\lambda$ weights the contribution of
-each of the $N$ observations to estimate $\bm \Sigma_k$. That is, for $0 <
-\lambda \le 1$, we estimate $\bm \Sigma_k$ with $N$ rather than $n_k$
+each of the $N$ observations to estimate $\bm \Sigma_k$. Hence, for $0 < \lambda
+\le 1$, we estimate $\bm \Sigma_k$ with $N$ rather than $n_k$
 observations. Below, we will show that the pooling operation is advantageous in
@@
 $\bm X_k = [\sqrt{c_{1k}(0)} \bm x_1', \ldots, \sqrt{c_{Nk}(0)} \bm x_N']'$ such
-that $\widehat{\bm \Sigma}_k = \bm X_k' \bm X_k$, $k = 1, \ldots, K$. In Figure
+that $\widehat{\bm \Sigma}_k = \bm X_k' \bm X_k$. In Figure
 \ref{fig:hdrda-contours} we plot the contours of five multivariate normal
-populations with unequal covariance matrices and $\lambda = 0$. As $\lambda$ is
-increased towards 1, the contours become more similar, ultimately resulting in
-equal covariance matrices and contours for the parsimonious case of $\lambda =
-1$.
+populations for $\lambda = 0$ with unequal covariance matrices. As $\lambda$
+approaches 1, the contours become more similar, resulting in identical contours
+for $\lambda = 1$.
 
@@
 
-As we have discussed, several eigenvalue adjustment methods have been proposed
-to increase eigenvalues (approximately) equal to 0. Here, we define the
+As we have discussed above, several eigenvalue adjustment methods have been
+proposed to increase eigenvalues (approximately) equal to 0. Here, we define the
 eigenvalue adjustment of \eqref{eq:sig-lambda-alternative} as
@@
 \eqref{eq:sig-rda}. Now, substituting \eqref{eq:hdrda-cov} into \eqref{eq:qda},
-we define the \emph{HDRDA} classifier as
+we define the high-dimensional regularized discriminant analysis classifier
 \begin{align}
-	D_{HDRDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\tilde{\bm\Sigma}_k^{+}(\bm x - \xbar_k)  + \ln |\tilde{\bm\Sigma}_k|. \label{eq:hdrda}
+	D_{HDRDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\tilde{\bm\Sigma}_k^{+}(\bm x - \xbar_k)  + \log |\tilde{\bm\Sigma}_k|. \label{eq:hdrda}
 \end{align}
@@
 
-Next, we demonstrate that the \emph{HDRDA} classifier lends itself to an
-efficient calculation of \eqref{eq:hdrda} and a rapid model selection of the
-pooling parameter $\lambda$. We decompose \eqref{eq:hdrda} into a sum of two
-components, where the first summand consists of matrix operations applied to
+\subsection{Enhanced Computational Efficiency}
+
+Next, we demonstrate that our proposed classifier in \eqref{eq:hdrda} lends
+itself to a more efficient calculation and a rapid model selection of the
+parameters $\lambda$ and $\gamma$\footnote{TODO: Briefly mention the updating
+  trick employed in the R package}. We decompose \eqref{eq:hdrda} into a sum of
+two components, where the first summand consists of matrix operations applied to
 low-dimensional matrices and the second summand corresponds to the null space of
@@
 operations performed on the null space of $\widehat{\bm \Sigma}$ yield constant
-quadratic forms across all classes and can be omitted under mild regularity
-conditions. For $p \gg N$, the constant component involves determinants and
-inverses of high-dimensional matrices, and by ignoring these calculations, a
-substantial reduction in computational costs is achieved relative to Friedman's
-\emph{RDA} classifier in \eqref{eq:rda}. Furthermore, a byproduct is that
-adjustments to the associated eigenvalues have no effect on
-\eqref{eq:hdrda}. Lastly, we utilize the singular value decomposition (SVD) to
+quadratic forms across all classes and can be omitted. For $p \gg N$, the
+constant component involves determinants and inverses of high-dimensional
+matrices, and by ignoring these calculations, a substantial reduction in
+computational costs is achieved relative to Friedman's classifier in
+\eqref{eq:rda}\footnote{TODO: Add a brief comparison of the computations of
+  \emph{RDA} and \emph{HDRDA} in terms of Big Oh notation}. Furthermore, a
+byproduct is that adjustments to the associated eigenvalues have no effect on
+\eqref{eq:hdrda}. Lastly, we utilize the singular value decomposition to
 calculate efficiently the eigenvalue decomposition of $\widehat{\bm \Sigma}$,
-further reducing the computational costs of our proposed \emph{RDA} classifier.
+further reducing the computational costs of our proposed classifier.
 
@@
 \begin{lemma}\label{lemma:null-spaces}
-Let $\widehat{\bm \Sigma}_k$ and $\widehat{\bm \Sigma}$ be the MLEs of $\bm
-\Sigma_k$ and $\bm \Sigma$ as discussed in Section \ref{sec:rda}, and let
+Let $\widehat{\bm \Sigma}_k$ and $\widehat{\bm \Sigma}$ be the maximum
+likelihood estimators of $\bm \Sigma_k$ and $\bm \Sigma$, respectively. Let
 $\widehat{\bm \Sigma}_k(\lambda)$ be defined as in
-\eqref{eq:sig-lambda-alternative}. Then, for $k = 1, \ldots, K$,
-$\mathcal{N}(\widehat{\bm \Sigma}_k(\lambda)) \subset \mathcal{N}(\widehat{\bm
-  \Sigma}) \subset \mathcal{N}(\widehat{\bm \Sigma}_k)$.
+\eqref{eq:sig-lambda-alternative}. Then, $\mathcal{N}\{\widehat{\bm \Sigma}_k(\lambda)\} \subset \mathcal{N}(\widehat{\bm
+  \Sigma}) \subset \mathcal{N}(\widehat{\bm \Sigma}_k)$ $(k = 1, \ldots, K)$.
 \end{lemma}
 \begin{proof}
-Let $\bm z \in \mathcal{N}(\widehat{\bm \Sigma}_k(\lambda))$ for some $k = 1,
+Let $\bm z \in \mathcal{N}\{\widehat{\bm \Sigma}_k(\lambda)\}$ for some $k = 1,
 \ldots, K$. Hence, $0 = \bm z' \widehat{\bm \Sigma}_k(\lambda) \bm z = (1 -
@@
   \Sigma})$ and $\bm z \in \mathcal{N}(\widehat{\bm \Sigma}_k)$. In particular,
-we have that $\mathcal{N}(\widehat{\bm \Sigma}_k(\lambda)) \subset
+we have that $\mathcal{N}\{\widehat{\bm \Sigma}_k(\lambda)\} \subset
 \mathcal{N}(\widehat{\bm \Sigma})$. Now, suppose $\bm z \in
@@
   \Sigma}_k \bm z$, which implies that $\bm z \in \mathcal{N}(\widehat{\bm
-  \Sigma}_k)$ because $\widehat{\bm \Sigma}_k \in \mathbb{R}_{p \times p}^{\ge}$
-for $k = 1, \ldots, K$. Therefore, $\mathcal{N}(\widehat{\bm \Sigma}) \subset
+  \Sigma}_k)$ because $\widehat{\bm \Sigma}_k \in \mathbb{R}_{p \times
+  p}^{\ge}$. Therefore, $\mathcal{N}(\widehat{\bm \Sigma}) \subset
 \mathcal{N}(\widehat{\bm \Sigma}_k)$.
@@
 eigendecomposition of $\widehat{\bm \Sigma}$ such that $\bm D \in \mathbb{R}_{p
-  \times p}$ is the diagonal matrix of eigenvalues of $\widehat{\bm \Sigma}$
+  \times p}^{\ge}$ is the diagonal matrix of eigenvalues of $\widehat{\bm \Sigma}$
 with $\bm D = \bm D_q \oplus \bm 0_{p-q}$, $\bm D_q \in \mathbb{R}_{q \times
@@
 \begin{align}
-	\tilde{\bm \Sigma}_k &= \bm U(\bm W_k \oplus \gamma \bm I_{p-q})\bm U',\label{eq:rda-matrix}
+	\tilde{\bm \Sigma}_k &= \bm U(\bm W_k \oplus \gamma \bm I_{p-q})\bm U' \quad
+  (k = 1, \ldots, K),\label{eq:rda-matrix}
 \intertext{where}
@@
 of $\widehat{\bm \Sigma}$, which implies that $\widehat{\bm \Sigma}_k \bm U_2 =
-\bm 0_{p \times (p - q)}$. Hence, for $k = 1, \ldots, K$, $\bm U' \widehat{\bm
-  \Sigma}_k \bm U = \bm U_1' \widehat{\bm \Sigma}_k \bm U_1 \oplus \bm
-0_{p-q}$. Thus, $\bm U' \tilde{\bm \Sigma}_k \bm U = \alpha_k \{(1 - \lambda)
-\bm U' \widehat{\bm \Sigma}_k \bm U + \lambda \bm D\} + \gamma \bm I_p$, and
+\bm 0_{p \times (p - q)}$. Hence, $\bm U' \widehat{\bm \Sigma}_k \bm U = \bm
+U_1' \widehat{\bm \Sigma}_k \bm U_1 \oplus \bm 0_{p-q}$ $(k = 1, \ldots,
+K)$. Thus, $\bm U' \tilde{\bm \Sigma}_k \bm U = \alpha_k \{(1 - \lambda) \bm U'
+\widehat{\bm \Sigma}_k \bm U + \lambda \bm D\} + \gamma \bm I_p$, and
 \eqref{eq:rda-matrix} follows because $\bm U$ is orthogonal.
@@
 \eqref{eq:sig-lambda-alternative}. Then, for $\lambda \in (0, 1]$,
-  rank$(\widehat{\bm \Sigma}_k(\lambda)) = q$ for all $k = 1, \ldots, K$.
+  rank$\{\widehat{\bm \Sigma}_k(\lambda)\} = q$ $(k = 1, \ldots, K)$.
 \end{cor}
@@
 increase the rank of $\widehat{\bm \Sigma}_k(\lambda)$ to $q$ if $\lambda \ne
-0$. Next, we provide an essential result that enables us to prove that, under
-mild regularity conditions, \eqref{eq:hdrda} is invariant to adjustments to the
-eigenvalues of $\tilde{\bm \Sigma}_k$ corresponding to the null space of
-$\widehat{\bm \Sigma}$ for any unlabeled observation $\bm x \in \mathbb{R}_{p
-  \times 1}$.
+0$. Next, we provide an essential result that enables us to prove that
+\eqref{eq:hdrda} is invariant to adjustments to the eigenvalues of $\tilde{\bm
+  \Sigma}_k$ corresponding to the null space of $\widehat{\bm \Sigma}$ for any
+unlabeled observation $\bm x \in \mathbb{R}_{p \times 1}$.
 
@@
 Let $\bm U_2$ be defined as above. Then, for all $\bm x \in \mathbb{R}_{p \times
-  1}$, $\bm U_2' (\bm x - \xbar_k) = \bm U_2' (\bm x - \xbar_{k'})$, $1 \le k,
-k' \le K$.
+  1}$, $\bm U_2' (\bm x - \xbar_k) = \bm U_2' (\bm x - \xbar_{k'})$ $(1 \le k,
+k' \le K)$.
 \end{lemma}
@@
   1.2.5]{Kollo:2005vp}. Now, because $\bm x_i \in \mathcal{C}(\widehat{\bm
-  \Sigma})$, $i = 1, \ldots, N$, $\bm U_2' \bm x_i = \bm 0_{(p-q) \times 1}$.
+  \Sigma})$ $(i = 1, \ldots, N)$, $\bm U_2' \bm x_i = \bm 0_{(p-q) \times 1}$.
 Hence, $\bm 0_{(p-q) \times 1} = \sum_{i=1}^N \beta_i \bm U_2'\bm x_i = \bm U_2'(\xbar_k -
@@
 
-We now present our main result, where we decompose the \emph{HDRDA} decision
-rule and show that the component requiring the largest computational costs does
-not contribute to the classification of an unlabeled observation performed using
-Lemma \ref{lemma:RDA-constant-term}. Hence, we reduce \eqref{eq:hdrda} to an
-equivalent, more computationally efficient decision rule.
+We now present our main result, where we decompose the decision rule of our
+proposed classifier and show that the term requiring the largest computational
+costs does not contribute to the classification of an unlabeled observation
+performed using Lemma \ref{lemma:RDA-constant-term}. Hence, we reduce
+\eqref{eq:hdrda} to an equivalent, more computationally efficient decision rule.
 
@@
 and \eqref{eq:Wk}, respectively, and let $\bm U_1$ be defined as above. Then,
-the \emph{HDRDA} decision rule in \eqref{eq:hdrda} is equivalent to
+the decision rule in \eqref{eq:hdrda} is equivalent to
 	\begin{align}
-		D_{HDRDA}(\bm x) &= \argmin_k  (\bm x - \xbar_k)' \bm U_1 \bm W_k^{-1} \bm U_1' (\bm x - \xbar_k) + \ln | \bm W_k |. \label{eq:hdrda-decomposed}
+		D_{HDRDA}(\bm x) &= \argmin_k  (\bm x - \xbar_k)' \bm U_1 \bm W_k^{-1} \bm U_1' (\bm x - \xbar_k) + \log | \bm W_k |. \label{eq:hdrda-decomposed}
 	\end{align}
-
 \end{thm}
@@
 W_k^{-1} \oplus \gamma^{+} \bm I_{p-q} )\bm U'$ and $|\tilde{\bm \Sigma}_k| =
-\gamma^{p-q}| \bm W_k |$. Therefore, for all $\bm x \in \mathbb{R}_{p \times 1}$,
-we have that
+\gamma^{p-q}| \bm W_k |$ $(k = 1, \ldots, K)$. Therefore, for all $\bm x \in
+\mathbb{R}_{p \times 1}$, we have that
 	\begin{align*}
-	(\bm x - \xbar_k)' \tilde{\bm\Sigma}_k^{-1}(\bm x - \xbar_k)  + \ln |\tilde{\bm\Sigma}_k| &= (\bm x - \xbar_k)' \bm U_1 \bm W_k^{-1} \bm U_1' (\bm x - \xbar_k)\\
-	&+ \gamma^{+} (\bm x - \xbar_k)' \bm U_2 \bm U_2' (\bm x - \xbar_k) + \ln | \bm W_k | + (p - q) \ln \gamma.
+	(\bm x - \xbar_k)' \tilde{\bm\Sigma}_k^{+}(\bm x - \xbar_k)  + \log |\tilde{\bm\Sigma}_k| &= (\bm x - \xbar_k)' \bm U_1 \bm W_k^{-1} \bm U_1' (\bm x - \xbar_k)\\
+	&+ \gamma^{+} (\bm x - \xbar_k)' \bm U_2 \bm U_2' (\bm x - \xbar_k) + \log | \bm W_k | + (p - q) \log \gamma.
 	\end{align*}
-The proof follows from Lemma \ref{lemma:RDA-constant-term} because $\bm U_2'
-(\bm x - \xbar_k)$ is constant for all $k = 1, \ldots, K$.
+Because $\gamma$ is constant for $k = 1, \ldots, K$, we can omit the $(p - q)
+\log \gamma$ term and particularly avoid the calculation of $\log 0$ for $\gamma
+= 0$. Then, the proof follows from Lemma \ref{lemma:RDA-constant-term} because
+$\bm U_2' (\bm x - \xbar_k)$ is constant for $k = 1, \ldots, K$.
 \end{proof}
@@
 
-Thus far, we have shown that the \emph{HDRDA} classifier in \eqref{eq:hdrda} is
-invariant to the term $\bm U_2$, thus yielding an equivalent classifier in
+Thus far, we have shown that our classifier in \eqref{eq:hdrda} is invariant to
+the term $\bm U_2$, thus yielding an equivalent classifier in
 \eqref{eq:hdrda-decomposed} with a substantial reduction in the computational
-complexity. While the classifier is much improved in terms of computational
-performance compared to the \emph{RDA} classifier originally proposed by
-\cite{Friedman:1989tm}, we provide two additional results that further improve
-the computational performance of our \emph{HDRDA} classifier. First, we
-demonstrate that the inverse and determinants of $\bm W_k$ can be written so
-that they are performed on $n_k \times n_k$ matrices rather than the larger $q
-\times q$ matrices, resulting in further improvements to the computational
-efficiency.
+complexity. Here, we provide two\footnote{TODO: Two? Perhaps, I mean one and
+  then the Fast SVD below. If so, clarify this.} additional results that further
+improve the computational performance of our classifier. First, we demonstrate
+that the inverse and determinants of $\bm W_k$ can be written so that they are
+performed on $n_k \times n_k$ matrices rather than the larger $q \times q$
+matrices, resulting in further improvements to the computational efficiency.
 
 \begin{proposition}\label{proposition:rda-W_k}
-Let $\bm W_k$ be defined as above. Then, $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$ and
+Let $\bm W_k$ be defined as above $(k = 1, \ldots, K)$. Then, $|\bm W_k| = |\bm
+\Gamma_k| |\bm Q_k|$ and
 	\begin{align}
@@
 $\widehat{\bm\Sigma} = \bm X' \bm X$. Then, because $\bm X = \bm M \bm D^{1/2}
-\bm U'$ by the SVD, we have the spectral decomposition $\widehat{\bm\Sigma} =
-\bm U \bm D \bm U'$, where $\bm M \in \mathbb{R}_{N \times N}$ is orthogonal,
-$\bm D^{1/2} \in \mathbb{R}_{N \times N}^{\ge}$ is a diagonal matrix consisting
-of the singular values of $\bm X$, and $\bm U \in \mathbb{R}_{p \times N}$ is
-orthogonal. Now, $\bm X \bm X' = \bm M \bm D \bm M'$ is the eigenvalue
-decomposition of $\bm X \bm X'$, where $\bm D$ is the matrix of eigenvalues of
-$\bm X \bm X'$ and the columns of $\bm M$ are the corresponding
-eigenvectors. Hence, we obtain $\bm M$ and $\bm D$ from an eigendecomposition of
-$\bm X \bm X' \in \mathbb{R}_{N \times N}$. Next, we compute $\bm U = \bm X' \bm
-M \bm D^{+/2}$, where $\bm D^{+/2} = \bm D_q^{-1/2} \oplus \bm 0_{N-q}$. We then
-determine $q$ by computing the number of numerically nonzero eigenvalues present
-in $\bm D$, where $q$ is the number of eigenvalues that exceeds some tolerance
-value, say, $\epsilon = 1 \times 10^{-6}$. After calculating $q$, we extract
-$\bm U_1$ by retaining only the first $q$ columns of $\bm U$.
+\bm U'$ by the singular value decomposition, we have the eigendecomposition
+$\widehat{\bm\Sigma} = \bm U \bm D \bm U'$, where $\bm M \in \mathbb{R}_{N
+  \times N}$ is orthogonal, $\bm D^{1/2} \in \mathbb{R}_{N \times N}^{\ge}$ is a
+diagonal matrix consisting of the singular values of $\bm X$, and $\bm U \in
+\mathbb{R}_{p \times N}$ is orthogonal. Now, $\bm X \bm X' = \bm M \bm D \bm M'$
+is the eigenvalue decomposition of $\bm X \bm X'$, where $\bm D$ is the matrix
+of eigenvalues of $\bm X \bm X'$ and the columns of $\bm M$ are the
+corresponding eigenvectors. Hence, we obtain $\bm M$ and $\bm D$ from an
+eigendecomposition of $\bm X \bm X' \in \mathbb{R}_{N \times N}$. Next, we
+compute $\bm U = \bm X' \bm M \bm D^{+/2}$, where $\bm D^{+/2} = \bm D_q^{-1/2}
+\oplus \bm 0_{N-q}$. We then determine $q$ by computing the number of
+numerically nonzero eigenvalues present in $\bm D$, where $q$ is the number of
+eigenvalues that exceeds some tolerance value, say, $\epsilon = 1 \times
+10^{-6}$. After calculating $q$, we extract $\bm U_1$ by retaining only the
+first $q$ columns of $\bm U$.
 
@@
 In this section, we describe four high-dimensional microarray data sets and
-compare our proposed \emph{HDRDA} classifier with five recently proposed
-classifiers designed for small-sample, high-dimensional data: the \emph{DLDA}
-classifier from \cite{Dudoit:2002ev}, the shrunken centroids \emph{RDA}
-classifier from \cite{Guo:2007te}, the shrinkage-based \emph{DLDA} classifier
-from \cite{Pang:2009ik}, the shrinkage-mean-based DLDA classifier from
-\cite{Tong:2012hw}, and the penalized \emph{LDA} classifier from
-\cite{Witten:2011kc}. The classifiers from \cite{Tong:2012hw} and
-\cite{Pang:2009ik} are both modifications of the \emph{DLDA} classifier from
-\cite{Dudoit:2002ev}, where the former employs an improved mean estimator and
-the latter utilizes an improved variance estimator. Below, we refer to each
+compare our proposed classifier with five classifiers recently proposed for
+small-sample, high-dimensional data. Our competitors include diagonal linear
+discriminant analysis from \cite{Dudoit:2002ev}, shrunken centroids regularized
+discriminant analysis from \cite{Guo:2007te}, and penalized linear discriminant
+analysis from \cite{Witten:2011kc}. We also consider two modifications of
+diagonal linear discriminant analysis from \cite{Tong:2012hw} and
+\cite{Pang:2009ik}, where the former employs an improved mean estimator and the
+latter utilizes an improved variance estimator. Below, we refer to each
 classifier by the first author's surname.
@@
 avoids the selection bias discussed in \cite{Ambroise:2002fa}. We did not
-include the \emph{RDA} classifier from \cite{Friedman:1989tm} in our studies to
-avoid its excessive computations. In Table \ref{tab:data-summary}, we summarize
-the attributes of each microarray data set.
+include \citeauthor{Friedman:1989tm}'s classifier in our studies to avoid its
+excessive computations. In Table \ref{tab:data-summary}, we summarize the
+attributes of each microarray data set.
 
@@
 classification study. For each classifier's {\tt R} implementation, we used the
-default settings with the exception that we explicitly set the prior
-probabilities as equal. In our {\tt R} implementation of the \emph{HDRDA}
-classifier in the {\tt sparsediscrim} package, we set $\alpha_k = 1$, $k = 1,
-\ldots, K$, so that the covariance-matrix estimator \eqref{eq:hdrda-cov}
-employed in the \emph{HDRDA} classifier has a form resembling
-\eqref{eq:ridge-estimator}. We estimated $\lambda$ from a grid of 21 equidistant
-candidate values between 0 and 1, inclusively. Similarly, we estimated $\gamma$
-from a grid consisting of the values $10^{-1}, \ldots, 10^4$, and $10^5$. We
-selected optimal estimates of $\lambda$ and $\gamma$ using 10-fold
-cross-validation \citep{Hastie:2008dt}.
+default settings with the exception that we explicitly set prior probabilities
+as equal. For our classifier, we set $\alpha_k = 1$ $(k = 1, \ldots, K)$, so
+that the covariance-matrix estimator \eqref{eq:hdrda-cov} employed had a form
+resembling \eqref{eq:ridge-estimator}. We estimated $\lambda$ from a grid of 21
+equidistant candidate values between 0 and 1, inclusively. Similarly, we
+estimated $\gamma$ from a grid consisting of the values $10^{-1}, \ldots, 10^4$,
+and $10^5$. We selected optimal estimates of $\lambda$ and $\gamma$ using
+10-fold cross-validation \citep{Hastie:2008dt}.
 
@@
 
-The \emph{HDRDA} classifier was superior in classification performance for the
-majority of the simulations. For the Burczysnki, Shipp, and Singh data sets, the
-\emph{HDRDA} classifier outperformed its competitors for all values of $d$
-examined. For the Nakayama data set, the \emph{HDRDA} classifier was
-outperformed by the Dudoit, Pang, and Tong classifiers for small values of $d$
-but attained classification rates superior to its competitors for larger values
-of $d$. Furthermore, Figure \ref{fig:error-rates} emphasizes two strengths of
-the \emph{HDRDA} classifier: reliability of the classifier for small $d$ and
-stability of the classifier for large $d$. In contrast to the other classifiers,
-the \emph{HDRDA} classifier exhibited less variability in its misclassification
-estimates for small values of $d$. Also, the \emph{HDRDA} classifier yielded
-superior classification accuracy as $d$ increased and was hardly effected by the
-inclusion of more variables in the data sets. We attribute the apparent
-stability of the error-rate estimates of the \emph{HDRDA} classifier for large
+Our classifier was superior in classification performance for the majority of
+the simulations. For the Burczysnki, Shipp, and Singh data sets, our classifier
+outperformed its competitors for all values of $d$ examined. For the Nakayama
+data set, our classifier was outperformed by the Dudoit, Pang, and Tong
+classifiers for small values of $d$ but attained classification rates superior
+to its competitors for larger values of $d$. Furthermore, Figure
+\ref{fig:error-rates} emphasizes two strengths of our classifier: reliability of
+the classifier for small $d$ and stability of the classifier for large $d$. In
+contrast to the other classifiers, our classifier exhibited less variability in
+its misclassification estimates for small values of $d$. Also, our classifier
+yielded superior classification accuracy as $d$ increased and was hardly
+effected by the inclusion of more variables in the data sets. We attribute the
+apparent stability of the error-rate estimates of our classifier for large
 values of $d$ to its novel dimension-reduction technique.
@@
 which suggests that the improved mean estimator employed by \cite{Tong:2012hw}
-provided little improvement to the classification performance of the \emph{DLDA}
-classifier.
+provided little improvement to the classification performance of diagonal linear
+discriminant analysis.
 
@@
 
-Next, we examine the optimal estimates of the tuning parameters for the
-\emph{HDRDA} classifier. For brevity, we include only a graphical summary of the
-optimal estimates of $\lambda$ for the Shipp data set in Figure
-\ref{fig:optimal-shipp}. Similar figures for the other three data sets can be
-found in the Supplementary Material. In Figure \ref{fig:optimal-shipp}, the
-distribution of the optimal values of $\lambda$ selected through
-cross-validation has a notable bathtub shape, resembling a Beta distribution
-with both shape parameters less than unity. Specifically, the optimal values of
-$\lambda$ are typically near 0 or near 1, corresponding to the \emph{QDA} and
-\emph{LDA} classifiers, respectively. We noticed a similar distribution of
-optimal values of $\lambda$ for the other three data sets.
-
-For the Shipp data set with smaller values of $d$, the range of optimal values
+Next, we examine the optimal estimates of the tuning parameters for our
+classifier. For brevity, we provide graphical summaries of the optimal estimates
+of $\lambda$ and $\gamma$ in the Supplementary Material. For each of the four
+data sets, the distribution of the optimal values of $\lambda$ selected through
+cross-validation often had a notable bathtub shape, resembling a Beta
+distribution with both shape parameters less than unity. Specifically, the
+optimal values of $\lambda$ were typically near 0 or near 1, corresponding to
+quadratic and linear discriminant analysis, respectively. For instance, with
+smaller values of $d$ applied to the Shipp data set, the range of optimal values
 of $\lambda$ was typically no larger than 0.05, indicating there were a large
@@
 optimal values of $\lambda$ were selected as unity, thereby an increased number
-of variables resulted in the \emph{HDRDA} classifier preferring the more
-parsimonious assumption of equal covariance matrices. In this instance we can
-see the flexibility of the \emph{HDRDA} classifier to employ the more realistic
-assumption that the class covariance matrices are unequal, but if optimal, we
-can employ the more parsimonious \emph{LDA} classifier as a special case.
-
-\[ \left[\text{Figure \ref{fig:optimal-shipp} goes about here }\right]\]
+of variables resulted in our classifier preferring the more parsimonious
+assumption of equal covariance matrices. In this instance we can see the
+flexibility of our classifier to employ the more realistic assumption that the
+class covariance matrices are unequal, but if optimal, we can employ the more
+parsimonious assumption of equal covariance matrices.
 
@@
 
-The analysis of high-dimensional microarray data sets is important to modern
-biomedical studies and presents challenges with supervised classification
-because the number of gene expressions greatly outnumbers the number of
-collected responses. Consequently, several variants of the well-known \emph{LDA}
-classifier have been proposed for improved classification performance. However,
-the \emph{LDA} classifier includes the strong assumption that the covariance
-matrices of each class are equal. Ideally, we could model the class covariance
-matrices as unrestricted and unequal in the \emph{QDA} classifier, but it is
-well known to have a large number of parameters that must be estimated and
-require a large number of samples, which are often unavailable in the
-high-dimensional setting. With this mind, \cite{Friedman:1989tm} proposed the
-\emph{RDA} classifier as a compromise between the \emph{LDA} and \emph{QDA}
-classifiers, controlling the amount of pooling of the individual class
-covariance-matrix estimators using the tuning parameter $\lambda$. However,
-despite its success for data sets where $p > N$, the \emph{RDA} classifier
-requires excessive calculations and a time-consuming model selection.
-
-We have adapted the \emph{RDA} classifier for usage with modern small-sample,
-high-dimensional data sets. Our model formulation facilitates a rapid
-classification of high-dimensional data by employing a novel dimension reduction
-technique. Moreover, our 10-fold cross-validation procedure typically requires a
-small runtime for data sets having large $p$. When preliminary variable
-selection is applied to the data as we performed in our classification study,
-the time to perform model selection is nearly instantaneous. Additionally, we
-have provided an implementation of the \emph{HDRDA} classifier in the {\tt
-  sparsediscrim} R package that is easy to use and available on CRAN.
-
-We have parameterized the covariance-matrix regularization employed in the
-\emph{HDRDA} classifier to include a large family of shrinkage methods, such as
-the methods from \cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and a wide variety
-of other shrinkage methods studied by \cite{Ramey:2013ji} and
-\cite{Xu:2009fl}. Our formulation of the \emph{HDRDA} classifier in terms of the
-nonzero eigenvalues of the pooled sample covariance matrix indicates that
-eigenvalue-adjustment methods applied to the \emph{LDA} classifier explains the
-similiarities in classification accuracy of several regularization methods
-applied to the \emph{LDA} classifier reported by \cite{Ramey:2013ji} in a
-high-dimensional classification study. We have provided an invariance property
-that extends results of \cite{Ji:2008wp} and \cite{Ye:2006jm}, who have provided
-a similar invariance property for the \emph{LDA} decision rule. Additionally,
-our model formulation improves the interpretation of the \emph{RDA} classifier
-by explicitly defining the covariance-matrix estimator so that each of the $N$
-observations contributes to the estimation of the individual class covariance
-matrices.
-
-In our simulation study, we compared the \emph{HDRDA} classifier to several
-recently proposed classifiers, including the Dudoit, Pang, Tong, and Guo
-classifiers, each of which assumes independence of the features conditional on
-class membership. Despite the rapid computational performance of diagonal
-classifiers and the reduction in the number of parameters to estimate,
-\cite{Mai:2012bf} and \cite{Fan:2012iq} have noted that diagonal classifiers
-can often yield inferior classification performance to other classification
-methods. We have shown that our proposed \emph{HDRDA} classifier often yields
-superior classification accuracy to the diagonal classifiers considered,
-confirming the assertions of \cite{Mai:2012bf} and
-\cite{Fan:2012iq}. Furthermore, the \emph{HDRDA} classifier is able to compete
-with the diagonal classifiers in terms of speed by incorporating a new
-dimension-reduction technique.
+In our simulation study, we compared our classifier to several recently proposed
+classifiers, including the Dudoit, Pang, Tong, and Guo classifiers, each of
+which assumes independence of the features conditional on class
+membership. Despite the rapid computational performance of diagonal classifiers
+and the reduction in the number of parameters to estimate, \cite{Mai:2012bf} and
+\cite{Fan:2012iq} have noted that diagonal classifiers can often yield inferior
+classification performance to other classification methods. We have shown that
+our proposed classifier often yields superior classification accuracy to the
+diagonal classifiers considered, confirming the assertions of \cite{Mai:2012bf}
+and \cite{Fan:2012iq}. Furthermore, the implementation of our classifier in the
+{\tt sparsediscrim} R package available on CRAN was comparable to that of the
+diagonal classifiers in terms of computational runtime. Moreover, when
+preliminary variable selection was applied to the data in our classification
+study, the time to perform model selection with our proposed classifier was
+nearly instantaneous.
 

### 2013-04-21 reports/paper-rda.Rnw
Removed bolding of vectors and matrices to match Biometrika's requirements.

@@
 
-\newcommand{\xbar}{\bar{\bm x}}
+\newcommand{\xbar}{\bar{x}}
 \newcommand{\tr}{\text{tr}}
@@
 In discriminant analysis we wish to assign correctly an unlabeled
-$p$-dimensional observation vector $\bm x$ to one of $K$ unique, known classes
-(or populations) by constructing a classifier from $n$ training observations
-that can accurately predict the class membership of $\bm x$. Let $\bm x_i =
-(x_{i1}, \ldots, x_{ip}) \in \mathbb{R}_{p \times 1}$ be the $i$th observation
-$(i = 1, \ldots, N)$ with true, unique membership $y_i \in \{\omega_1, \ldots,
-\omega_K\}$, where $\mathbb{R}_{a \times b}$ denotes the matrix space of all $a
-\times b$ matrices over the real field $\mathbb{R}$. We assume that $(\bm x_i,
-y_i)$ is a realization from a mixture distribution $p(\bm x) = \sum_{k=1}^K
-p(\bm x | \omega_k) p(\omega_k)$, where $p(\bm x | \omega_k)$ is the probability
-density function of the $k$th class and $p(\omega_k)$ is the prior probability
-of class membership of the $k$th class. We further assume that each of
-$p(\omega_k)$ is equal $(k = 1, \ldots, K)$.
+$p$-dimensional observation vector $x$ to one of $K$ unique, known classes
+(or populations) by constructing a classifier from $N$ training
+observations. Let $x_i = (x_{i1}, \ldots, x_{ip}) \in \mathbb{R}_{p \times
+  1}$ be the $i$th observation $(i = 1, \ldots, N)$ with true, unique membership
+$y_i \in \{\omega_1, \ldots, \omega_K\}$, where $\mathbb{R}_{a \times b}$
+denotes the matrix space of all $a \times b$ matrices over the real field
+$\mathbb{R}$. We assume that $(x_i, y_i)$ is a realization from a mixture
+distribution $p(x) = \sum_{k=1}^K p(x | \omega_k) p(\omega_k)$, where
+$p(x | \omega_k)$ is the probability density function of the $k$th class and
+$p(\omega_k)$ is the prior probability of class membership of the $k$th
+class. We further assume that each of $p(\omega_k)$ is equal $(k = 1, \ldots,
+K)$.
 
 Quadratic discriminant analysis is the optimal Bayesian decision rule with
-respect to a $0-1$ loss function when $p(\bm x | \omega_k)$ is the probability
+respect to a $0-1$ loss function when $p(x | \omega_k)$ is the probability
 density function of the multivariate normal distribution with known mean vector
-$\bm\mu_k \in \mathbb{R}_{p \times 1}$ and known covariance matrix $\bm\Sigma_k
+$\mu_k \in \mathbb{R}_{p \times 1}$ and known covariance matrix $\Sigma_k
 \in \mathbb{R}_{p \times p}^{>}$, $k = 1, 2, \ldots, K$, where $\mathbb{R}_{p
   \times p}^{>}$ denotes the cone of real $p \times p$ positive-definite
-matrices. Because $\bm \mu_k$ and $\bm\Sigma_k$ are typically unknown, we assign
-an unlabeled observation $\bm x$ to class $\omega_k$ with the sample quadratic
+matrices. Because $\mu_k$ and $\Sigma_k$ are typically unknown, we assign
+an unlabeled observation $x$ to class $\omega_k$ with the sample quadratic
 discriminant analysis classifier
 \begin{align}
-	D_{QDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\widehat{\bm\Sigma}_k^{-1}(\bm x - \xbar_k)  + \log |\widehat{\bm\Sigma}_k|, \label{eq:qda}
+	D_{QDA}(x) = \argmin_{k}  (x - \xbar_k)'\widehat{\Sigma}_k^{-1}(x - \xbar_k)  + \log |\widehat{\Sigma}_k|, \label{eq:qda}
 \end{align}
-where $\xbar_k$ and $\widehat{\bm\Sigma}_k$ are the maximum likelihood
-estimators for $\bm \mu_k$ and $\bm \Sigma_k$, respectively. If we assume
-further that $\bm\Sigma_k = \bm\Sigma$ $(k = 1, \ldots, K)$, then the pooled
-sample covariance matrix $\widehat{\bm\Sigma}$ is substituted for $\widehat{\bm
+where $\xbar_k$ and $\widehat{\Sigma}_k$ are the maximum likelihood
+estimators for $\mu_k$ and $\Sigma_k$, respectively. If we assume
+further that $\Sigma_k = \Sigma$ $(k = 1, \ldots, K)$, then the pooled
+sample covariance matrix $\widehat{\Sigma}$ is substituted for $\widehat{
   \Sigma}_k$ in \eqref{eq:qda}, where
 \begin{align}
-	\widehat{\bm\Sigma} = N^{-1} \sum_{k=1}^K n_k \widehat{\bm\Sigma}_k \label{eq:pooled-cov}
+	\widehat{\Sigma} = N^{-1} \sum_{k=1}^K n_k \widehat{\Sigma}_k \label{eq:pooled-cov}
 \end{align}
-is the maximum likelihood estimator for $\bm \Sigma$. In this case
-\eqref{eq:qda} reduces to the sample linear discriminant analysis classifier,
-and the log-determinant is omitted because it is constant across all classes.
+is the maximum likelihood estimator for $\Sigma$. In this case \eqref{eq:qda}
+reduces to the sample linear discriminant analysis classifier, and the
+log-determinant is omitted because it is constant across all classes.
 
 Throughout the document we also require the following additional notation. Let
-$\bm A \oplus \bm B$ denote the direct sum of $\bm A \in \mathbb{R}_{r \times
-  r}$ and $\bm B \in \mathbb{R}_{s \times s}$ \citep[Chapter
-  1]{Lutkepohl:1996uz}. Let $\bm A^+$ and $\mathcal{N}(\bm A) = \{\bm z \in
-\mathbb{R}_{p \times 1} : \bm A \bm z = \bm 0\}$ denote the Moore-Penrose
-pseudoinverse and null space of $\bm A \in \mathbb{R}_{m \times p}$,
-respectively. Let $V^{\perp} = \{\bm y \in \mathbb{R}_{p \times 1} : \bm y'\bm v
-= 0\ \forall \bm v \in V \}$ denote the orthogonal complement of a vector space
-$V \subset \mathbb{R}_{p \times 1}$. Let $\bm I_m$ be the $m \times m$ identity
-matrix, and let $\bm 0_{m \times p}$ be the $m \times p$ matrix of zeros, such
-that $\bm 0_m$ is understood to denote $\bm 0_{m \times m}$. Lastly, for $c \in
-\mathbb{R}$, let $c^+ = 1/c$ if $c \ne 0$ and $0$ otherwise.
+$I_m$ be the $m \times m$ identity matrix, and let $0_{m \times p}$ be
+the $m \times p$ matrix of zeros, such that $0_m$ is understood to denote
+$0_{m \times m}$. Lastly, for $c \in \mathbb{R}$, let $c^+ = 1/c$ if $c \ne
+0$ and $0$ otherwise. Let $A \oplus B$ denote the direct sum of $A \in
+\mathbb{R}_{r \times r}$ and $B \in \mathbb{R}_{s \times s}$ \citep[Chapter
+  1]{Lutkepohl:1996uz}. Let $A^+$ and $\mathcal{N}(A) = \{z \in \mathbb{R}_{p
+  \times 1} : A z = 0_{m \times 1}\}$ denote the Moore-Penrose pseudoinverse and null space
+of $A \in \mathbb{R}_{m \times p}$, respectively. Let $V^{\perp} = \{y
+\in \mathbb{R}_{p \times 1} : y'v = 0\ \forall v \in V \}$ denote
+the orthogonal complement of a vector space $V \subset \mathbb{R}_{p \times 1}$.
 
@@
 The smallest eigenvalues and the directions associated with their eigenvectors
-highly influence \eqref{eq:qda}. In fact, the eigenvalues of $\widehat{\bm
+highly influence \eqref{eq:qda}. In fact, the eigenvalues of $\widehat{
   \Sigma}_k$ are well-known to be biased if $p \ge n_k$ such that the smallest
 eigenvalues are underestimated \citep{Seber:2004uh}, Moreover, if $p > n_k$,
-then rank$(\widehat{\bm \Sigma}_k) \le n_k$, which implies that at least $p -
-n_k$ eigenvalues of $\widehat{\bm \Sigma}_k$ are 0. Furthermore, although more
+then rank$(\widehat{\Sigma}_k) \le n_k$, which implies that at least $p -
+n_k$ eigenvalues of $\widehat{\Sigma}_k$ are 0. Furthermore, although more
 feature information is available to discriminate among the $K$ classes, if $p >
-n_k$, \eqref{eq:qda} is incalculable because $\widehat{\bm \Sigma}_k^{-1}$ does
+n_k$, \eqref{eq:qda} is incalculable because $\widehat{\Sigma}_k^{-1}$ does
 not exist.
@@
 \citet*{Guo:2007te}, \cite{Mkhadri:1995jp}, and \citet*{Xu:2009fl}, have been
-proposed in the literature to adjust the eigenvalues of $\widehat{\bm \Sigma}_k$
+proposed in the literature to adjust the eigenvalues of $\widehat{\Sigma}_k$
 so that \eqref{eq:qda} is calculable and has reduced variability in the standard
@@
 \begin{align}
-	\widehat{\bm \Sigma}_k(\gamma) =  \widehat{\bm \Sigma}_k + \gamma \bm I_p, \label{eq:ridge-estimator}
+	\widehat{\Sigma}_k(\gamma) =  \widehat{\Sigma}_k + \gamma I_p, \label{eq:ridge-estimator}
 \end{align}
@@
 \eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance
-matrix $\widehat{\bm\Sigma}_k$ towards $\bm I_p$, thereby increasing the
-eigenvalues of $\widehat{\bm\Sigma}_k$ by $\gamma$. Specifically, the zero
+matrix $\widehat{\Sigma}_k$ towards $I_p$, thereby increasing the
+eigenvalues of $\widehat{\Sigma}_k$ by $\gamma$. Specifically, the zero
 eigenvalues are replaced with $\gamma$, so that \eqref{eq:ridge-estimator} is
@@
 \cite{Friedman:1989tm} has proposed regularized discriminant analysis by
-incorporating a weighted average of the sample covariance matrix $\widehat{\bm
-  \Sigma}_k$ and the pooled sample covariance matrix $\widehat{\bm\Sigma}$ to
-estimate $\bm \Sigma_k$ with
+incorporating a weighted average of the sample covariance matrix $\widehat{
+  \Sigma}_k$ and the pooled sample covariance matrix $\widehat{\Sigma}$ to
+estimate $\Sigma_k$ with
 \begin{align}
-  \widehat{\bm\Sigma}_k(\lambda) = n_k^{-1}(\lambda) \bm S_k(\lambda) \quad (k =
+  \widehat{\Sigma}_k(\lambda) = n_k^{-1}(\lambda) S_k(\lambda) \quad (k =
   1, \ldots, K),\label{eq:sig-lambda}
 \end{align}
-where $\lambda \in [0, 1]$, $n_k(\lambda) = (1 - \lambda) n_k + \lambda N$, $\bm
-S_k = n_k \widehat{\bm\Sigma}_k$, $\bm S = \sum_{k=1}^K \bm S_k$, and $\bm
-S_k(\lambda) = (1 - \lambda) \bm S_k + \lambda \bm S$. We can interpret
+where $\lambda \in [0, 1]$, $n_k(\lambda) = (1 - \lambda) n_k + \lambda N$, $
+S_k = n_k \widehat{\Sigma}_k$, $S = \sum_{k=1}^K S_k$, and $
+S_k(\lambda) = (1 - \lambda) S_k + \lambda S$. We can interpret
 \eqref{eq:sig-lambda} as a covariance-matrix estimator that borrows from
-$\widehat{\bm\Sigma}$ in \eqref{eq:pooled-cov} to better estimate $\bm
+$\widehat{\Sigma}$ in \eqref{eq:pooled-cov} to better estimate $
 \Sigma_k$. If $\lambda = 0$, \eqref{eq:sig-lambda} is equal to
-$\widehat{\bm\Sigma}_k$ used in \eqref{eq:qda}. Alternatively, if $\lambda = 1$,
+$\widehat{\Sigma}_k$ used in \eqref{eq:qda}. Alternatively, if $\lambda = 1$,
 \eqref{eq:sig-lambda} reduces to \eqref{eq:pooled-cov}.
 
-To further improve the estimation of $\bm \Sigma_k$ and to stabilize the inverse
+To further improve the estimation of $\Sigma_k$ and to stabilize the inverse
 of \eqref{eq:sig-lambda}, \cite{Friedman:1989tm} has proposed the biased
@@
 \begin{align}
-	\widehat{\bm\Sigma}_k(\lambda, \gamma) = (1 - \gamma) \widehat{\bm\Sigma}_k(\lambda) + \gamma \frac{\tr\{\widehat{\bm\Sigma}_k(\lambda)\}}{p} \bm I_p,\label{eq:sig-rda}
+	\widehat{\Sigma}_k(\lambda, \gamma) = (1 - \gamma) \widehat{\Sigma}_k(\lambda) + \gamma \frac{\tr\{\widehat{\Sigma}_k(\lambda)\}}{p} I_p,\label{eq:sig-rda}
 \end{align}
 where $\gamma \in [0, 1]$ is a regularization parameter that controls the
-shrinkage of \eqref{eq:sig-rda} towards $\bm I_p$ and is weighted by the average
+shrinkage of \eqref{eq:sig-rda} towards $I_p$ and is weighted by the average
 of the eigenvalues of \eqref{eq:sig-lambda}. Thus, the \emph{pooling} parameter
-$\lambda$ controls the amount borrowed from $\widehat{\bm\Sigma}$ to estimate
-$\bm \Sigma_k$, and the \emph{shrinkage} parameter $\gamma$ determines the
+$\lambda$ controls the amount borrowed from $\widehat{\Sigma}$ to estimate
+$\Sigma_k$, and the \emph{shrinkage} parameter $\gamma$ determines the
 amount of eigenvalue shrinkage applied. Finally, by substituting
@@
 \begin{align}
-	D_{RDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\widehat{\bm\Sigma}_k(\lambda, \gamma)^{-1}(\bm x - \xbar_k)  + \log |\widehat{\bm\Sigma}_k(\lambda, \gamma)|. \label{eq:rda}
+	D_{RDA}(x) = \argmin_{k}  (x - \xbar_k)'\widehat{\Sigma}_k(\lambda, \gamma)^{-1}(x - \xbar_k)  + \log |\widehat{\Sigma}_k(\lambda, \gamma)|. \label{eq:rda}
 \end{align}
@@
 Here, we define our adaptation of \citeauthor{Friedman:1989tm}'s classifier by
-first simplifying the formulation of $\widehat{\bm \Sigma}_k(\lambda)$ in
+first simplifying the formulation of $\widehat{\Sigma}_k(\lambda)$ in
 \eqref{eq:sig-lambda} and demonstrating its clear interpretation as a linear
@@
 \begin{align}
-  \widehat{\bm\Sigma}_k(\lambda) = (1 - \lambda) \widehat{\bm\Sigma}_k + \lambda
-  \widehat{\bm\Sigma}, \quad (k = 1, \ldots, K)\label{eq:sig-lambda-alternative}
+  \widehat{\Sigma}_k(\lambda) = (1 - \lambda) \widehat{\Sigma}_k + \lambda
+  \widehat{\Sigma}, \quad (k = 1, \ldots, K)\label{eq:sig-lambda-alternative}
 \end{align}
@@
 convex combination of $n_k$ and $N$ as defined by \cite{Friedman:1989tm}. By
-rewriting \eqref{eq:sig-lambda-alternative} in terms of the observations $\bm
+rewriting \eqref{eq:sig-lambda-alternative} in terms of the observations $
 x_i$ $(i = 1, \ldots, N)$, each centered by its class sample mean, we attain a
-clear interpretation of $\widehat{\bm\Sigma}_k(\lambda)$. That is,
+clear interpretation of $\widehat{\Sigma}_k(\lambda)$. That is,
 \begin{align}
-	\widehat{\bm\Sigma}_k(\lambda) &= \widehat{\bm\Sigma}_k + \lambda (\widehat{\bm\Sigma} - \widehat{\bm\Sigma}_k) \nonumber\\
-%	&= \widehat{\bm\Sigma}_k + \lambda \left(\sum_{k' = 1}^K \frac{n_k}{N} \widehat{\bm\Sigma}_{k'}  - \widehat{\bm\Sigma}_k \right) \nonumber\\
-	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right) \widehat{\bm\Sigma}_k +  \frac{\lambda}{N} \sum_{\substack{k' = 1\\k' \ne k}} n_{k'} \widehat{\bm\Sigma}_{k'} \nonumber \\
-%	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N I(y_i = k) \bm x_i \bm x_i' +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k) \bm x_i \bm x_i' \nonumber \\
-	&= \sum_{i=1}^N c_{ik}(\lambda) \bm x_i \bm x_i',\label{eq:sig-lambda-alternative2}
+	\widehat{\Sigma}_k(\lambda) &= \widehat{\Sigma}_k + \lambda (\widehat{\Sigma} - \widehat{\Sigma}_k) \nonumber\\
+%	&= \widehat{\Sigma}_k + \lambda \left(\sum_{k' = 1}^K \frac{n_k}{N} \widehat{\Sigma}_{k'}  - \widehat{\Sigma}_k \right) \nonumber\\
+	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right) \widehat{\Sigma}_k +  \frac{\lambda}{N} \sum_{\substack{k' = 1\\k' \ne k}} n_{k'} \widehat{\Sigma}_{k'} \nonumber \\
+%	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N I(y_i = k) x_i x_i' +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k) x_i x_i' \nonumber \\
+	&= \sum_{i=1}^N c_{ik}(\lambda) x_i x_i',\label{eq:sig-lambda-alternative2}
 \end{align}
@@
 From \eqref{eq:sig-lambda-alternative2}, $\lambda$ weights the contribution of
-each of the $N$ observations to estimate $\bm \Sigma_k$. Hence, for $0 < \lambda
-\le 1$, we estimate $\bm \Sigma_k$ with $N$ rather than $n_k$
+each of the $N$ observations to estimate $\Sigma_k$. Hence, for $0 < \lambda
+\le 1$, we estimate $\Sigma_k$ with $N$ rather than $n_k$
 observations. Below, we will show that the pooling operation is advantageous in
-increasing the rank of each $\widehat{\bm\Sigma}_k(\lambda)$ from
-rank$(\widehat{\bm\Sigma}_k)$ to rank$(\widehat{\bm\Sigma})$ for $0 < \lambda
+increasing the rank of each $\widehat{\Sigma}_k(\lambda)$ from
+rank$(\widehat{\Sigma}_k)$ to rank$(\widehat{\Sigma})$ for $0 < \lambda
 \le 1$. Notice that if $\lambda = 0$, then the observations from the remaining
-$K - 1$ classes do not contribute to the estimation of $\bm \Sigma_k$,
-corresponding to $\widehat{\bm \Sigma}_k$. Furthermore, if $\lambda = 1$, the
+$K - 1$ classes do not contribute to the estimation of $\Sigma_k$,
+corresponding to $\widehat{\Sigma}_k$. Furthermore, if $\lambda = 1$, the
 weights in \eqref{eq:sig-lambda-alternative2} reduce to $1/N$, corresponding to
-$\widehat{\bm\Sigma}$. For brevity, when $\lambda = 1$, we define $\bm X =
-[\sqrt{c_{1k}(1)} \bm x_1', \ldots, \sqrt{c_{Nk}(1)} \bm x_N']'$ such that
-$\widehat{\bm \Sigma} = \bm X' \bm X$. Similarly, for $\lambda = 0$, we define
-$\bm X_k = [\sqrt{c_{1k}(0)} \bm x_1', \ldots, \sqrt{c_{Nk}(0)} \bm x_N']'$ such
-that $\widehat{\bm \Sigma}_k = \bm X_k' \bm X_k$. In Figure
+$\widehat{\Sigma}$. For brevity, when $\lambda = 1$, we define $X =
+[\sqrt{c_{1k}(1)} x_1', \ldots, \sqrt{c_{Nk}(1)} x_N']'$ such that
+$\widehat{\Sigma} = X' X$. Similarly, for $\lambda = 0$, we define
+$X_k = [\sqrt{c_{1k}(0)} x_1', \ldots, \sqrt{c_{Nk}(0)} x_N']'$ such
+that $\widehat{\Sigma}_k = X_k' X_k$. In Figure
 \ref{fig:hdrda-contours} we plot the contours of five multivariate normal
@@
 \begin{align}
-	\tilde{\bm \Sigma}_k = \alpha_k \widehat{\bm \Sigma}_k(\lambda) + \gamma \bm I_p,\label{eq:hdrda-cov}
+	\tilde{\Sigma}_k = \alpha_k \widehat{\Sigma}_k(\lambda) + \gamma I_p,\label{eq:hdrda-cov}
 \end{align}
@@
 \begin{align}
-	D_{HDRDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\tilde{\bm\Sigma}_k^{+}(\bm x - \xbar_k)  + \log |\tilde{\bm\Sigma}_k|. \label{eq:hdrda}
+	D_{HDRDA}(x) = \argmin_{k}  (x - \xbar_k)'\tilde{\Sigma}_k^{+}(x - \xbar_k)  + \log |\tilde{\Sigma}_k|. \label{eq:hdrda}
 \end{align}
-For $\gamma > 0$, $\tilde{\bm\Sigma}_k$ is nonsingular such that
-$\tilde{\bm\Sigma}_k^{-1}$ can be substituted for $\tilde{\bm\Sigma}_k^{+}$ in
-\eqref{eq:hdrda}. If $\gamma = 0$, we explicitly set $|\tilde{\bm\Sigma}_k|$
-equal to the product of the positive eigenvalues of $\tilde{\bm\Sigma}_k$.
+For $\gamma > 0$, $\tilde{\Sigma}_k$ is nonsingular such that
+$\tilde{\Sigma}_k^{-1}$ can be substituted for $\tilde{\Sigma}_k^{+}$ in
+\eqref{eq:hdrda}. If $\gamma = 0$, we explicitly set $|\tilde{\Sigma}_k|$
+equal to the product of the positive eigenvalues of $\tilde{\Sigma}_k$.
 
@@
 low-dimensional matrices and the second summand corresponds to the null space of
-$\widehat{\bm \Sigma}$ in \eqref{eq:pooled-cov}. We show that the matrix
-operations performed on the null space of $\widehat{\bm \Sigma}$ yield constant
+$\widehat{\Sigma}$ in \eqref{eq:pooled-cov}. We show that the matrix
+operations performed on the null space of $\widehat{\Sigma}$ yield constant
 quadratic forms across all classes and can be omitted. For $p \gg N$, the
@@
 \eqref{eq:hdrda}. Lastly, we utilize the singular value decomposition to
-calculate efficiently the eigenvalue decomposition of $\widehat{\bm \Sigma}$,
+calculate efficiently the eigenvalue decomposition of $\widehat{\Sigma}$,
 further reducing the computational costs of our proposed classifier.
@@
 First, we require the following relationship regarding the null spaces of
-$\widehat{\bm \Sigma}_k(\lambda)$, $\widehat{\bm \Sigma}$, and $\widehat{\bm
+$\widehat{\Sigma}_k(\lambda)$, $\widehat{\Sigma}$, and $\widehat{
   \Sigma}_k$.
@@
 \begin{lemma}\label{lemma:null-spaces}
-Let $\widehat{\bm \Sigma}_k$ and $\widehat{\bm \Sigma}$ be the maximum
-likelihood estimators of $\bm \Sigma_k$ and $\bm \Sigma$, respectively. Let
-$\widehat{\bm \Sigma}_k(\lambda)$ be defined as in
-\eqref{eq:sig-lambda-alternative}. Then, $\mathcal{N}\{\widehat{\bm \Sigma}_k(\lambda)\} \subset \mathcal{N}(\widehat{\bm
-  \Sigma}) \subset \mathcal{N}(\widehat{\bm \Sigma}_k)$ $(k = 1, \ldots, K)$.
+Let $\widehat{\Sigma}_k$ and $\widehat{\Sigma}$ be the maximum
+likelihood estimators of $\Sigma_k$ and $\Sigma$, respectively. Let
+$\widehat{\Sigma}_k(\lambda)$ be defined as in
+\eqref{eq:sig-lambda-alternative}. Then, $\mathcal{N}\{\widehat{\Sigma}_k(\lambda)\} \subset \mathcal{N}(\widehat{
+  \Sigma}) \subset \mathcal{N}(\widehat{\Sigma}_k)$ $(k = 1, \ldots, K)$.
 \end{lemma}
 \begin{proof}
-Let $\bm z \in \mathcal{N}\{\widehat{\bm \Sigma}_k(\lambda)\}$ for some $k = 1,
-\ldots, K$. Hence, $0 = \bm z' \widehat{\bm \Sigma}_k(\lambda) \bm z = (1 -
-\lambda) \bm z' \widehat{\bm \Sigma}_k \bm z + \lambda \bm z' \widehat{\bm
-  \Sigma} \bm z$. Because $\widehat{\bm \Sigma}_k, \widehat{\bm \Sigma}\in
-\mathbb{R}_{p \times p}^{\ge}$, we have $\bm z \in \mathcal{N}(\widehat{\bm
-  \Sigma})$ and $\bm z \in \mathcal{N}(\widehat{\bm \Sigma}_k)$. In particular,
-we have that $\mathcal{N}\{\widehat{\bm \Sigma}_k(\lambda)\} \subset
-\mathcal{N}(\widehat{\bm \Sigma})$. Now, suppose $\bm z \in
-\mathcal{N}(\widehat{\bm \Sigma})$. Similarly, we have that $0 = \bm z'
-\widehat{\bm \Sigma} \bm z = N^{-1} \sum_{k = 1}^K n_k \bm z' \widehat{\bm
-  \Sigma}_k \bm z$, which implies that $\bm z \in \mathcal{N}(\widehat{\bm
-  \Sigma}_k)$ because $\widehat{\bm \Sigma}_k \in \mathbb{R}_{p \times
-  p}^{\ge}$. Therefore, $\mathcal{N}(\widehat{\bm \Sigma}) \subset
-\mathcal{N}(\widehat{\bm \Sigma}_k)$.
+Let $z \in \mathcal{N}\{\widehat{\Sigma}_k(\lambda)\}$ for some $k = 1,
+\ldots, K$. Hence, $0 = z' \widehat{\Sigma}_k(\lambda) z = (1 -
+\lambda) z' \widehat{\Sigma}_k z + \lambda z' \widehat{
+  \Sigma} z$. Because $\widehat{\Sigma}_k, \widehat{\Sigma}\in
+\mathbb{R}_{p \times p}^{\ge}$, we have $z \in \mathcal{N}(\widehat{
+  \Sigma})$ and $z \in \mathcal{N}(\widehat{\Sigma}_k)$. In particular,
+we have that $\mathcal{N}\{\widehat{\Sigma}_k(\lambda)\} \subset
+\mathcal{N}(\widehat{\Sigma})$. Now, suppose $z \in
+\mathcal{N}(\widehat{\Sigma})$. Similarly, we have that $0 = z'
+\widehat{\Sigma} z = N^{-1} \sum_{k = 1}^K n_k z' \widehat{
+  \Sigma}_k z$, which implies that $z \in \mathcal{N}(\widehat{
+  \Sigma}_k)$ because $\widehat{\Sigma}_k \in \mathbb{R}_{p \times
+  p}^{\ge}$. Therefore, $\mathcal{N}(\widehat{\Sigma}) \subset
+\mathcal{N}(\widehat{\Sigma}_k)$.
 \end{proof}
@@
 In Lemma \ref{lemma:rda-tilde-Sigma_k} below, we derive an alternative
-expression for $\tilde{\bm \Sigma}_k$ in terms of the matrix of eigenvectors of
-$\widehat{\bm \Sigma}$. Let $\widehat{\bm \Sigma} = \bm U \bm D \bm U'$ be the
-eigendecomposition of $\widehat{\bm \Sigma}$ such that $\bm D \in \mathbb{R}_{p
-  \times p}^{\ge}$ is the diagonal matrix of eigenvalues of $\widehat{\bm \Sigma}$
-with $\bm D = \bm D_q \oplus \bm 0_{p-q}$, $\bm D_q \in \mathbb{R}_{q \times
+expression for $\tilde{\Sigma}_k$ in terms of the matrix of eigenvectors of
+$\widehat{\Sigma}$. Let $\widehat{\Sigma} = U D U'$ be the
+eigendecomposition of $\widehat{\Sigma}$ such that $D \in \mathbb{R}_{p
+  \times p}^{\ge}$ is the diagonal matrix of eigenvalues of $\widehat{\Sigma}$
+with $D = D_q \oplus 0_{p-q}$, $D_q \in \mathbb{R}_{q \times
   q}^{>}$ is the diagonal matrix consisting of the positive eigenvalues of
-$\widehat{\bm \Sigma}$, the columns of $\bm U \in \mathbb{R}_{p \times p}$ are
-the corresponding orthonormal eigenvectors of $\widehat{\bm \Sigma}$, and
-rank$(\widehat{\bm \Sigma}) = q$. Then, we partition $\bm U = (\bm U_1, \bm
-U_2)$ such that $\bm U_1 \in \mathbb{R}_{p \times q}$ and $\bm U_2 \in
+$\widehat{\Sigma}$, the columns of $U \in \mathbb{R}_{p \times p}$ are
+the corresponding orthonormal eigenvectors of $\widehat{\Sigma}$, and
+rank$(\widehat{\Sigma}) = q$. Then, we partition $U = (U_1, 
+U_2)$ such that $U_1 \in \mathbb{R}_{p \times q}$ and $U_2 \in
 \mathbb{R}_{p \times (p - q)}$.
@@
 \begin{lemma}\label{lemma:rda-tilde-Sigma_k}
-Let $\widehat{\bm \Sigma} = \bm U \bm D \bm U'$ be the eigendecomposition of
-$\widehat{\bm \Sigma}$ as above, and suppose that rank$(\widehat{\bm \Sigma}) =
+Let $\widehat{\Sigma} = U D U'$ be the eigendecomposition of
+$\widehat{\Sigma}$ as above, and suppose that rank$(\widehat{\Sigma}) =
 q \le p$. Then, we have
 \begin{align}
-	\tilde{\bm \Sigma}_k &= \bm U(\bm W_k \oplus \gamma \bm I_{p-q})\bm U' \quad
+	\tilde{\Sigma}_k &= U(W_k \oplus \gamma I_{p-q})U' \quad
   (k = 1, \ldots, K),\label{eq:rda-matrix}
 \intertext{where}
-\bm W_k &= \alpha_k \{(1 - \lambda) \bm U_1' \widehat{\bm \Sigma}_k \bm U_1 + \lambda \bm D_q\} + \gamma \bm I_{q}.\label{eq:Wk}
+W_k &= \alpha_k \{(1 - \lambda) U_1' \widehat{\Sigma}_k U_1 + \lambda D_q\} + \gamma I_{q}.\label{eq:Wk}
 \end{align}
@@
 \begin{proof}
-From Lemma \ref{lemma:null-spaces}, the columns of $\bm U_2$ span the null space
-of $\widehat{\bm \Sigma}$, which implies that $\widehat{\bm \Sigma}_k \bm U_2 =
-\bm 0_{p \times (p - q)}$. Hence, $\bm U' \widehat{\bm \Sigma}_k \bm U = \bm
-U_1' \widehat{\bm \Sigma}_k \bm U_1 \oplus \bm 0_{p-q}$ $(k = 1, \ldots,
-K)$. Thus, $\bm U' \tilde{\bm \Sigma}_k \bm U = \alpha_k \{(1 - \lambda) \bm U'
-\widehat{\bm \Sigma}_k \bm U + \lambda \bm D\} + \gamma \bm I_p$, and
-\eqref{eq:rda-matrix} follows because $\bm U$ is orthogonal.
+From Lemma \ref{lemma:null-spaces}, the columns of $U_2$ span the null space
+of $\widehat{\Sigma}$, which implies that $\widehat{\Sigma}_k U_2 =
+0_{p \times (p - q)}$. Hence, $U' \widehat{\Sigma}_k U = 
+U_1' \widehat{\Sigma}_k U_1 \oplus 0_{p-q}$ $(k = 1, \ldots,
+K)$. Thus, $U' \tilde{\Sigma}_k U = \alpha_k \{(1 - \lambda) U'
+\widehat{\Sigma}_k U + \lambda D\} + \gamma I_p$, and
+\eqref{eq:rda-matrix} follows because $U$ is orthogonal.
 \end{proof}
@@
 \begin{cor}
-Let $\widehat{\bm \Sigma}_k(\lambda)$ be defined as in
+Let $\widehat{\Sigma}_k(\lambda)$ be defined as in
 \eqref{eq:sig-lambda-alternative}. Then, for $\lambda \in (0, 1]$,
-  rank$\{\widehat{\bm \Sigma}_k(\lambda)\} = q$ $(k = 1, \ldots, K)$.
+  rank$\{\widehat{\Sigma}_k(\lambda)\} = q$ $(k = 1, \ldots, K)$.
 \end{cor}
@@
 
-Thus, by incorporating each $\bm x_i$ into the estimation of $\bm \Sigma_k$, we
-increase the rank of $\widehat{\bm \Sigma}_k(\lambda)$ to $q$ if $\lambda \ne
+Thus, by incorporating each $x_i$ into the estimation of $\Sigma_k$, we
+increase the rank of $\widehat{\Sigma}_k(\lambda)$ to $q$ if $\lambda \ne
 0$. Next, we provide an essential result that enables us to prove that
-\eqref{eq:hdrda} is invariant to adjustments to the eigenvalues of $\tilde{\bm
-  \Sigma}_k$ corresponding to the null space of $\widehat{\bm \Sigma}$ for any
-unlabeled observation $\bm x \in \mathbb{R}_{p \times 1}$.
+\eqref{eq:hdrda} is invariant to adjustments to the eigenvalues of $\tilde{
+  \Sigma}_k$ corresponding to the null space of $\widehat{\Sigma}$ for any
+unlabeled observation $x \in \mathbb{R}_{p \times 1}$.
 
 \begin{lemma}\label{lemma:RDA-constant-term}
-Let $\bm U_2$ be defined as above. Then, for all $\bm x \in \mathbb{R}_{p \times
-  1}$, $\bm U_2' (\bm x - \xbar_k) = \bm U_2' (\bm x - \xbar_{k'})$ $(1 \le k,
+Let $U_2$ be defined as above. Then, for all $x \in \mathbb{R}_{p \times
+  1}$, $U_2' (x - \xbar_k) = U_2' (x - \xbar_{k'})$ $(1 \le k,
 k' \le K)$.
@@
 \begin{proof}
-Let $\bm x \in \mathbb{R}_{p \times 1}$, and suppose that $1 \le k, k' \le K$.
-Recall that $\bm U_2 \in \mathcal{N}(\widehat{\bm \Sigma})$, which implies that
-$\bm U_2' \in \mathcal{C}(\widehat{\bm \Sigma})^{\perp}$ \citep[Lemma
-  1.2.5]{Kollo:2005vp}. Now, because $\bm x_i \in \mathcal{C}(\widehat{\bm
-  \Sigma})$ $(i = 1, \ldots, N)$, $\bm U_2' \bm x_i = \bm 0_{(p-q) \times 1}$.
-Hence, $\bm 0_{(p-q) \times 1} = \sum_{i=1}^N \beta_i \bm U_2'\bm x_i = \bm U_2'(\xbar_k -
+Let $x \in \mathbb{R}_{p \times 1}$, and suppose that $1 \le k, k' \le K$.
+Recall that $U_2 \in \mathcal{N}(\widehat{\Sigma})$, which implies that
+$U_2' \in \mathcal{C}(\widehat{\Sigma})^{\perp}$ \citep[Lemma
+  1.2.5]{Kollo:2005vp}. Now, because $x_i \in \mathcal{C}(\widehat{
+  \Sigma})$ $(i = 1, \ldots, N)$, $U_2' x_i = 0_{(p-q) \times 1}$.
+Hence, $0_{(p-q) \times 1} = \sum_{i=1}^N \beta_i U_2'x_i = U_2'(\xbar_k -
 \xbar_{k'})$, where $\beta_i = (n_k n_{k'})^{-1} \{ I(y_i = k) n_{k'} - I(y_i =
-k') n_k \}$. Therefore, $\bm U_2' (\bm x - \xbar_k) = \bm U_2' (\bm x - \xbar_{k'})$.
+k') n_k \}$. Therefore, $U_2' (x - \xbar_k) = U_2' (x - \xbar_{k'})$.
 \end{proof}
@@
 \begin{thm}
-Let $\tilde{\bm \Sigma}_k$ and $\bm W_k$ be defined as in \eqref{eq:rda-matrix}
-and \eqref{eq:Wk}, respectively, and let $\bm U_1$ be defined as above. Then,
+Let $\tilde{\Sigma}_k$ and $W_k$ be defined as in \eqref{eq:rda-matrix}
+and \eqref{eq:Wk}, respectively, and let $U_1$ be defined as above. Then,
 the decision rule in \eqref{eq:hdrda} is equivalent to
 	\begin{align}
-		D_{HDRDA}(\bm x) &= \argmin_k  (\bm x - \xbar_k)' \bm U_1 \bm W_k^{-1} \bm U_1' (\bm x - \xbar_k) + \log | \bm W_k |. \label{eq:hdrda-decomposed}
+		D_{HDRDA}(x) &= \argmin_k  (x - \xbar_k)' U_1 W_k^{-1} U_1' (x - \xbar_k) + \log | W_k |. \label{eq:hdrda-decomposed}
 	\end{align}
@@
 \begin{proof}
-From \eqref{eq:rda-matrix}, we have that $\tilde{\bm \Sigma}_k^{+} = \bm U(\bm
-W_k^{-1} \oplus \gamma^{+} \bm I_{p-q} )\bm U'$ and $|\tilde{\bm \Sigma}_k| =
-\gamma^{p-q}| \bm W_k |$ $(k = 1, \ldots, K)$. Therefore, for all $\bm x \in
+From \eqref{eq:rda-matrix}, we have that $\tilde{\Sigma}_k^{+} = U(
+W_k^{-1} \oplus \gamma^{+} I_{p-q} )U'$ and $|\tilde{\Sigma}_k| =
+\gamma^{p-q}| W_k |$ $(k = 1, \ldots, K)$. Therefore, for all $x \in
 \mathbb{R}_{p \times 1}$, we have that
 	\begin{align*}
-	(\bm x - \xbar_k)' \tilde{\bm\Sigma}_k^{+}(\bm x - \xbar_k)  + \log |\tilde{\bm\Sigma}_k| &= (\bm x - \xbar_k)' \bm U_1 \bm W_k^{-1} \bm U_1' (\bm x - \xbar_k)\\
-	&+ \gamma^{+} (\bm x - \xbar_k)' \bm U_2 \bm U_2' (\bm x - \xbar_k) + \log | \bm W_k | + (p - q) \log \gamma.
+	(x - \xbar_k)' \tilde{\Sigma}_k^{+}(x - \xbar_k)  + \log |\tilde{\Sigma}_k| &= (x - \xbar_k)' U_1 W_k^{-1} U_1' (x - \xbar_k)\\
+	&+ \gamma^{+} (x - \xbar_k)' U_2 U_2' (x - \xbar_k) + \log | W_k | + (p - q) \log \gamma.
 	\end{align*}
@@
 = 0$. Then, the proof follows from Lemma \ref{lemma:RDA-constant-term} because
-$\bm U_2' (\bm x - \xbar_k)$ is constant for $k = 1, \ldots, K$.
+$U_2' (x - \xbar_k)$ is constant for $k = 1, \ldots, K$.
 \end{proof}
@@
 Thus far, we have shown that our classifier in \eqref{eq:hdrda} is invariant to
-the term $\bm U_2$, thus yielding an equivalent classifier in
+the term $U_2$, thus yielding an equivalent classifier in
 \eqref{eq:hdrda-decomposed} with a substantial reduction in the computational
@@
 improve the computational performance of our classifier. First, we demonstrate
-that the inverse and determinants of $\bm W_k$ can be written so that they are
+that the inverse and determinants of $W_k$ can be written so that they are
 performed on $n_k \times n_k$ matrices rather than the larger $q \times q$
@@
 \begin{proposition}\label{proposition:rda-W_k}
-Let $\bm W_k$ be defined as above $(k = 1, \ldots, K)$. Then, $|\bm W_k| = |\bm
-\Gamma_k| |\bm Q_k|$ and
+Let $W_k$ be defined as above $(k = 1, \ldots, K)$. Then, $|W_k| = |
+\Gamma_k| |Q_k|$ and
 	\begin{align}
-		\bm W_k^{-1} &= \bm \Gamma_k^{-1} - \alpha_k(1 - \lambda) \bm \Gamma_k^{-1} \bm U_1' \bm X_k' \bm Q_k^{-1} \bm X_k \bm U_1 \bm \Gamma_k^{-1},\label{eq:W_k_inv}
+		W_k^{-1} &= \Gamma_k^{-1} - \alpha_k(1 - \lambda) \Gamma_k^{-1} U_1' X_k' Q_k^{-1} X_k U_1 \Gamma_k^{-1},\label{eq:W_k_inv}
 	\end{align}
-where $\bm Q_k = \bm I_{n_k} + \alpha_k(1- \lambda) \bm X_k \bm U_1 \bm
-\Gamma_k^{-1} \bm U_1' \bm X_k'$ and $\bm \Gamma_k = \alpha_k \lambda \bm D_q +
-\gamma \bm I_q$.
+where $Q_k = I_{n_k} + \alpha_k(1- \lambda) X_k U_1 
+\Gamma_k^{-1} U_1' X_k'$ and $\Gamma_k = \alpha_k \lambda D_q +
+\gamma I_q$.
 \end{proposition}
 \begin{proof}
-To calculate $|\bm W_k|$, we apply Theorem 18.1.1 from \cite{Harville:2008wja},
-which states that $|\bm A + \bm B \bm T \bm C| = |\bm A| |\bm T| |\bm T^{-1} +
-\bm C \bm A^{-1} \bm B|$. Thus, setting $\bm A = \bm \Gamma_k$, $\bm B =
-\alpha_k (1 - \lambda) \bm U_1' \bm X_k'$, $\bm T = \bm I_{n_k}$, and $\bm C =
-\bm X_k \bm U_1$, we have $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$. Also,
+To calculate $|W_k|$, we apply Theorem 18.1.1 from \cite{Harville:2008wja},
+which states that $|A + B T C| = |A| |T| |T^{-1} +
+C A^{-1} B|$. Thus, setting $A = \Gamma_k$, $B =
+\alpha_k (1 - \lambda) U_1' X_k'$, $T = I_{n_k}$, and $C =
+X_k U_1$, we have $|W_k| = |\Gamma_k| |Q_k|$. Also,
 \eqref{eq:W_k_inv} follows from the well-known Sherman-Woodbury formula
@@
 
-Thus far, we have constructed each $\tilde{\bm \Sigma}_k$ in terms of the first
-$q$ eigenvectors of $\widehat{\bm \Sigma}$. By utilizing a technique similar to
-the so-called kernel trick, we can obtain $\bm U_1$ by computing the
+Thus far, we have constructed each $\tilde{\Sigma}_k$ in terms of the first
+$q$ eigenvectors of $\widehat{\Sigma}$. By utilizing a technique similar to
+the so-called kernel trick, we can obtain $U_1$ by computing the
 eigendecomposition of a much smaller $N \times N$ matrix when $p \gg N$
 \citep{Hastie:2008dt}. Recall that the pooled sample covariance matrix
-$\widehat{\bm\Sigma} = \bm X' \bm X$. Then, because $\bm X = \bm M \bm D^{1/2}
-\bm U'$ by the singular value decomposition, we have the eigendecomposition
-$\widehat{\bm\Sigma} = \bm U \bm D \bm U'$, where $\bm M \in \mathbb{R}_{N
-  \times N}$ is orthogonal, $\bm D^{1/2} \in \mathbb{R}_{N \times N}^{\ge}$ is a
-diagonal matrix consisting of the singular values of $\bm X$, and $\bm U \in
-\mathbb{R}_{p \times N}$ is orthogonal. Now, $\bm X \bm X' = \bm M \bm D \bm M'$
-is the eigenvalue decomposition of $\bm X \bm X'$, where $\bm D$ is the matrix
-of eigenvalues of $\bm X \bm X'$ and the columns of $\bm M$ are the
-corresponding eigenvectors. Hence, we obtain $\bm M$ and $\bm D$ from an
-eigendecomposition of $\bm X \bm X' \in \mathbb{R}_{N \times N}$. Next, we
-compute $\bm U = \bm X' \bm M \bm D^{+/2}$, where $\bm D^{+/2} = \bm D_q^{-1/2}
-\oplus \bm 0_{N-q}$. We then determine $q$ by computing the number of
-numerically nonzero eigenvalues present in $\bm D$, where $q$ is the number of
+$\widehat{\Sigma} = X' X$. Then, because $X = M D^{1/2}
+U'$ by the singular value decomposition, we have the eigendecomposition
+$\widehat{\Sigma} = U D U'$, where $M \in \mathbb{R}_{N
+  \times N}$ is orthogonal, $D^{1/2} \in \mathbb{R}_{N \times N}^{\ge}$ is a
+diagonal matrix consisting of the singular values of $X$, and $U \in
+\mathbb{R}_{p \times N}$ is orthogonal. Now, $X X' = M D M'$
+is the eigenvalue decomposition of $X X'$, where $D$ is the matrix
+of eigenvalues of $X X'$ and the columns of $M$ are the
+corresponding eigenvectors. Hence, we obtain $M$ and $D$ from an
+eigendecomposition of $X X' \in \mathbb{R}_{N \times N}$. Next, we
+compute $U = X' M D^{+/2}$, where $D^{+/2} = D_q^{-1/2}
+\oplus 0_{N-q}$. We then determine $q$ by computing the number of
+numerically nonzero eigenvalues present in $D$, where $q$ is the number of
 eigenvalues that exceeds some tolerance value, say, $\epsilon = 1 \times
-10^{-6}$. After calculating $q$, we extract $\bm U_1$ by retaining only the
-first $q$ columns of $\bm U$.
+10^{-6}$. After calculating $q$, we extract $U_1$ by retaining only the
+first $q$ columns of $U$.
 

### 2013-04-21 reports/paper-rda.Rnw
Updated vector and matrix transposes to use T rather than an apostrophe.

@@
 \begin{align}
-	D_{QDA}(x) = \argmin_{k}  (x - \xbar_k)'\widehat{\Sigma}_k^{-1}(x - \xbar_k)  + \log |\widehat{\Sigma}_k|, \label{eq:qda}
+	D_{QDA}(x) = \argmin_{k}  (x - \xbar_k)^{T}\widehat{\Sigma}_k^{-1}(x - \xbar_k)  + \log |\widehat{\Sigma}_k|, \label{eq:qda}
 \end{align}
 where $\xbar_k$ and $\widehat{\Sigma}_k$ are the maximum likelihood
-estimators for $\mu_k$ and $\Sigma_k$, respectively. If we assume
+estimators for $\mu_k$ and $\Sigma_k$, respectively, and $A^{T}$ denotes the
+matrix transpose of $A \in \mathbb{R}_{m \times p}$. If we assume
 further that $\Sigma_k = \Sigma$ $(k = 1, \ldots, K)$, then the pooled
@@
 of $A \in \mathbb{R}_{m \times p}$, respectively. Let $V^{\perp} = \{y
-\in \mathbb{R}_{p \times 1} : y'v = 0\ \forall v \in V \}$ denote
+\in \mathbb{R}_{p \times 1} : y^{T}v = 0\ \forall v \in V \}$ denote
 the orthogonal complement of a vector space $V \subset \mathbb{R}_{p \times 1}$.
@@
 \begin{align}
-	D_{RDA}(x) = \argmin_{k}  (x - \xbar_k)'\widehat{\Sigma}_k(\lambda, \gamma)^{-1}(x - \xbar_k)  + \log |\widehat{\Sigma}_k(\lambda, \gamma)|. \label{eq:rda}
+	D_{RDA}(x) = \argmin_{k}  (x - \xbar_k)^{T}\widehat{\Sigma}_k(\lambda, \gamma)^{-1}(x - \xbar_k)  + \log |\widehat{\Sigma}_k(\lambda, \gamma)|. \label{eq:rda}
 \end{align}
@@
 	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right) \widehat{\Sigma}_k +  \frac{\lambda}{N} \sum_{\substack{k' = 1\\k' \ne k}} n_{k'} \widehat{\Sigma}_{k'} \nonumber \\
-%	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N I(y_i = k) x_i x_i' +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k) x_i x_i' \nonumber \\
-	&= \sum_{i=1}^N c_{ik}(\lambda) x_i x_i',\label{eq:sig-lambda-alternative2}
+%	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N I(y_i = k) x_i x_i^{T} +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k) x_i x_i^{T} \nonumber \\
+	&= \sum_{i=1}^N c_{ik}(\lambda) x_i x_i^{T},\label{eq:sig-lambda-alternative2}
 \end{align}
@@
 $\widehat{\Sigma}$. For brevity, when $\lambda = 1$, we define $X =
-[\sqrt{c_{1k}(1)} x_1', \ldots, \sqrt{c_{Nk}(1)} x_N']'$ such that
-$\widehat{\Sigma} = X' X$. Similarly, for $\lambda = 0$, we define
-$X_k = [\sqrt{c_{1k}(0)} x_1', \ldots, \sqrt{c_{Nk}(0)} x_N']'$ such
-that $\widehat{\Sigma}_k = X_k' X_k$. In Figure
+[\sqrt{c_{1k}(1)} x_1^{T}, \ldots, \sqrt{c_{Nk}(1)} x_N^{T}]^{T}$ such that
+$\widehat{\Sigma} = X^{T} X$. Similarly, for $\lambda = 0$, we define
+$X_k = [\sqrt{c_{1k}(0)} x_1^{T}, \ldots, \sqrt{c_{Nk}(0)} x_N^{T}]^{T}$ such
+that $\widehat{\Sigma}_k = X_k^{T} X_k$. In Figure
 \ref{fig:hdrda-contours} we plot the contours of five multivariate normal
@@
 \begin{align}
-	D_{HDRDA}(x) = \argmin_{k}  (x - \xbar_k)'\tilde{\Sigma}_k^{+}(x - \xbar_k)  + \log |\tilde{\Sigma}_k|. \label{eq:hdrda}
+	D_{HDRDA}(x) = \argmin_{k}  (x - \xbar_k)^{T}\tilde{\Sigma}_k^{+}(x - \xbar_k)  + \log |\tilde{\Sigma}_k|. \label{eq:hdrda}
 \end{align}
@@
 Let $z \in \mathcal{N}\{\widehat{\Sigma}_k(\lambda)\}$ for some $k = 1,
-\ldots, K$. Hence, $0 = z' \widehat{\Sigma}_k(\lambda) z = (1 -
-\lambda) z' \widehat{\Sigma}_k z + \lambda z' \widehat{
+\ldots, K$. Hence, $0 = z^{T} \widehat{\Sigma}_k(\lambda) z = (1 -
+\lambda) z^{T} \widehat{\Sigma}_k z + \lambda z^{T} \widehat{
   \Sigma} z$. Because $\widehat{\Sigma}_k, \widehat{\Sigma}\in
@@
 \mathcal{N}(\widehat{\Sigma})$. Now, suppose $z \in
-\mathcal{N}(\widehat{\Sigma})$. Similarly, we have that $0 = z'
-\widehat{\Sigma} z = N^{-1} \sum_{k = 1}^K n_k z' \widehat{
+\mathcal{N}(\widehat{\Sigma})$. Similarly, we have that $0 = z^{T}
+\widehat{\Sigma} z = N^{-1} \sum_{k = 1}^K n_k z^{T} \widehat{
   \Sigma}_k z$, which implies that $z \in \mathcal{N}(\widehat{
@@
 expression for $\tilde{\Sigma}_k$ in terms of the matrix of eigenvectors of
-$\widehat{\Sigma}$. Let $\widehat{\Sigma} = U D U'$ be the
+$\widehat{\Sigma}$. Let $\widehat{\Sigma} = U D U^{T}$ be the
 eigendecomposition of $\widehat{\Sigma}$ such that $D \in \mathbb{R}_{p
@@
 \begin{lemma}\label{lemma:rda-tilde-Sigma_k}
-Let $\widehat{\Sigma} = U D U'$ be the eigendecomposition of
+Let $\widehat{\Sigma} = U D U^{T}$ be the eigendecomposition of
 $\widehat{\Sigma}$ as above, and suppose that rank$(\widehat{\Sigma}) =
@@
 \begin{align}
-	\tilde{\Sigma}_k &= U(W_k \oplus \gamma I_{p-q})U' \quad
+	\tilde{\Sigma}_k &= U(W_k \oplus \gamma I_{p-q})U^{T} \quad
   (k = 1, \ldots, K),\label{eq:rda-matrix}
 \intertext{where}
-W_k &= \alpha_k \{(1 - \lambda) U_1' \widehat{\Sigma}_k U_1 + \lambda D_q\} + \gamma I_{q}.\label{eq:Wk}
+W_k &= \alpha_k \{(1 - \lambda) U_1^{T} \widehat{\Sigma}_k U_1 + \lambda D_q\} + \gamma I_{q}.\label{eq:Wk}
 \end{align}
@@
 of $\widehat{\Sigma}$, which implies that $\widehat{\Sigma}_k U_2 =
-0_{p \times (p - q)}$. Hence, $U' \widehat{\Sigma}_k U = 
-U_1' \widehat{\Sigma}_k U_1 \oplus 0_{p-q}$ $(k = 1, \ldots,
-K)$. Thus, $U' \tilde{\Sigma}_k U = \alpha_k \{(1 - \lambda) U'
+0_{p \times (p - q)}$. Hence, $U^{T} \widehat{\Sigma}_k U = 
+U_1^{T} \widehat{\Sigma}_k U_1 \oplus 0_{p-q}$ $(k = 1, \ldots,
+K)$. Thus, $U^{T} \tilde{\Sigma}_k U = \alpha_k \{(1 - \lambda) U^{T}
 \widehat{\Sigma}_k U + \lambda D\} + \gamma I_p$, and
@@
 Let $U_2$ be defined as above. Then, for all $x \in \mathbb{R}_{p \times
-  1}$, $U_2' (x - \xbar_k) = U_2' (x - \xbar_{k'})$ $(1 \le k,
+  1}$, $U_2^{T} (x - \xbar_k) = U_2^{T} (x - \xbar_{k'})$ $(1 \le k,
 k' \le K)$.
@@
 Recall that $U_2 \in \mathcal{N}(\widehat{\Sigma})$, which implies that
-$U_2' \in \mathcal{C}(\widehat{\Sigma})^{\perp}$ \citep[Lemma
+$U_2^{T} \in \mathcal{C}(\widehat{\Sigma})^{\perp}$ \citep[Lemma
   1.2.5]{Kollo:2005vp}. Now, because $x_i \in \mathcal{C}(\widehat{
-  \Sigma})$ $(i = 1, \ldots, N)$, $U_2' x_i = 0_{(p-q) \times 1}$.
-Hence, $0_{(p-q) \times 1} = \sum_{i=1}^N \beta_i U_2'x_i = U_2'(\xbar_k -
+  \Sigma})$ $(i = 1, \ldots, N)$, $U_2^{T} x_i = 0_{(p-q) \times 1}$.
+Hence, $0_{(p-q) \times 1} = \sum_{i=1}^N \beta_i U_2^{T}x_i = U_2^{T}(\xbar_k -
 \xbar_{k'})$, where $\beta_i = (n_k n_{k'})^{-1} \{ I(y_i = k) n_{k'} - I(y_i =
-k') n_k \}$. Therefore, $U_2' (x - \xbar_k) = U_2' (x - \xbar_{k'})$.
+k') n_k \}$. Therefore, $U_2^{T} (x - \xbar_k) = U_2^{T} (x - \xbar_{k'})$.
 \end{proof}
@@
 	\begin{align}
-		D_{HDRDA}(x) &= \argmin_k  (x - \xbar_k)' U_1 W_k^{-1} U_1' (x - \xbar_k) + \log | W_k |. \label{eq:hdrda-decomposed}
+		D_{HDRDA}(x) &= \argmin_k  (x - \xbar_k)^{T} U_1 W_k^{-1} U_1^{T} (x - \xbar_k) + \log | W_k |. \label{eq:hdrda-decomposed}
 	\end{align}
@@
 From \eqref{eq:rda-matrix}, we have that $\tilde{\Sigma}_k^{+} = U(
-W_k^{-1} \oplus \gamma^{+} I_{p-q} )U'$ and $|\tilde{\Sigma}_k| =
+W_k^{-1} \oplus \gamma^{+} I_{p-q} )U^{T}$ and $|\tilde{\Sigma}_k| =
 \gamma^{p-q}| W_k |$ $(k = 1, \ldots, K)$. Therefore, for all $x \in
@@
 	\begin{align*}
-	(x - \xbar_k)' \tilde{\Sigma}_k^{+}(x - \xbar_k)  + \log |\tilde{\Sigma}_k| &= (x - \xbar_k)' U_1 W_k^{-1} U_1' (x - \xbar_k)\\
-	&+ \gamma^{+} (x - \xbar_k)' U_2 U_2' (x - \xbar_k) + \log | W_k | + (p - q) \log \gamma.
+	(x - \xbar_k)^{T} \tilde{\Sigma}_k^{+}(x - \xbar_k)  + \log |\tilde{\Sigma}_k| &= (x - \xbar_k)^{T} U_1 W_k^{-1} U_1^{T} (x - \xbar_k)\\
+	&+ \gamma^{+} (x - \xbar_k)^{T} U_2 U_2^{T} (x - \xbar_k) + \log | W_k | + (p - q) \log \gamma.
 	\end{align*}
@@
 = 0$. Then, the proof follows from Lemma \ref{lemma:RDA-constant-term} because
-$U_2' (x - \xbar_k)$ is constant for $k = 1, \ldots, K$.
+$U_2^{T} (x - \xbar_k)$ is constant for $k = 1, \ldots, K$.
 \end{proof}
@@
 	\begin{align}
-		W_k^{-1} &= \Gamma_k^{-1} - \alpha_k(1 - \lambda) \Gamma_k^{-1} U_1' X_k' Q_k^{-1} X_k U_1 \Gamma_k^{-1},\label{eq:W_k_inv}
+		W_k^{-1} &= \Gamma_k^{-1} - \alpha_k(1 - \lambda) \Gamma_k^{-1} U_1^{T} X_k^{T} Q_k^{-1} X_k U_1 \Gamma_k^{-1},\label{eq:W_k_inv}
 	\end{align}
 where $Q_k = I_{n_k} + \alpha_k(1- \lambda) X_k U_1 
-\Gamma_k^{-1} U_1' X_k'$ and $\Gamma_k = \alpha_k \lambda D_q +
+\Gamma_k^{-1} U_1^{T} X_k^{T}$ and $\Gamma_k = \alpha_k \lambda D_q +
 \gamma I_q$.
@@
 C A^{-1} B|$. Thus, setting $A = \Gamma_k$, $B =
-\alpha_k (1 - \lambda) U_1' X_k'$, $T = I_{n_k}$, and $C =
+\alpha_k (1 - \lambda) U_1^{T} X_k^{T}$, $T = I_{n_k}$, and $C =
 X_k U_1$, we have $|W_k| = |\Gamma_k| |Q_k|$. Also,
@@
 \citep{Hastie:2008dt}. Recall that the pooled sample covariance matrix
-$\widehat{\Sigma} = X' X$. Then, because $X = M D^{1/2}
-U'$ by the singular value decomposition, we have the eigendecomposition
-$\widehat{\Sigma} = U D U'$, where $M \in \mathbb{R}_{N
+$\widehat{\Sigma} = X^{T} X$. Then, because $X = M D^{1/2}
+U^{T}$ by the singular value decomposition, we have the eigendecomposition
+$\widehat{\Sigma} = U D U^{T}$, where $M \in \mathbb{R}_{N
   \times N}$ is orthogonal, $D^{1/2} \in \mathbb{R}_{N \times N}^{\ge}$ is a
 diagonal matrix consisting of the singular values of $X$, and $U \in
-\mathbb{R}_{p \times N}$ is orthogonal. Now, $X X' = M D M'$
-is the eigenvalue decomposition of $X X'$, where $D$ is the matrix
-of eigenvalues of $X X'$ and the columns of $M$ are the
+\mathbb{R}_{p \times N}$ is orthogonal. Now, $X X^{T} = M D M^{T}$
+is the eigenvalue decomposition of $X X^{T}$, where $D$ is the matrix
+of eigenvalues of $X X^{T}$ and the columns of $M$ are the
 corresponding eigenvectors. Hence, we obtain $M$ and $D$ from an
-eigendecomposition of $X X' \in \mathbb{R}_{N \times N}$. Next, we
-compute $U = X' M D^{+/2}$, where $D^{+/2} = D_q^{-1/2}
+eigendecomposition of $X X^{T} \in \mathbb{R}_{N \times N}$. Next, we
+compute $U = X^{T} M D^{+/2}$, where $D^{+/2} = D_q^{-1/2}
 \oplus 0_{N-q}$. We then determine $q$ by computing the number of

### 2013-04-21 reports/paper-rda.Rnw
Moved figures of HDRDA optimal estimates to Supplementary Material document.

@@
 
-\begin{figure}
-<<optimal_estimates>>=
-hdrda_optimal$lambda <- factor(hdrda_optimal$lambda)
-hdrda_optimal$gamma <- factor(hdrda_optimal$gamma)
-p <- ggplot(subset(hdrda_optimal, data_set == "shipp" & d %in% c(100, 500, 1000)), aes(lambda))
-p <- p + geom_bar(aes(fill = gamma, weight = value))
-p <- p + facet_grid(d ~ ., labeller = label_bquote(d == .(x))) + theme_bw()
-p + theme(strip.text.y = element_text(size = 16))
-@ 
-\caption{Counts of the optimal tuning-parameter estimates for the \emph{HDRDA}
-  classifier for the Shipp data set for $d = 100$, $500$, and $1000$.}
-\label{fig:optimal-shipp}
-\end{figure}
-
 \end{document}

### 2013-04-21 reports/supplementary-material.Rnw
Moved figures of HDRDA optimal estimates to Supplementary Material document.

--- 2013-04-21 reports/paper-rda.Rnw
@@
 
-\title{Another Look at Regularized Discriminant Analysis for High-Dimensional Classification}
+\title{Supplementary Material for ``Another Look at Regularized Discriminant Analysis for High-Dimensional Classification''}
 
@@
                warning = FALSE, cache = TRUE, echo = FALSE, autodep = TRUE,
-               fig.width = 10, fig.height = 10, out.width = "\\linewidth")
+               fig.width = 12, fig.height = 12, out.width = "\\linewidth")
 @
@@
 
-\newtheorem{cor}{Corollary}
-\newtheorem{lemma}{Lemma}
-\newtheorem{proposition}{Proposition}
-\newtheorem{thm}{Theorem}
-
-\bibpunct{(}{)}{;}{a}{,}{,}
-
 \doublespacing
@@
 
-\begin{abstract}
-
-A reliable supervised classification of tumors from gene expression microarray
-data is essential for successful diagnosis and treatment of diseases such as
-cancer; however, the analysis of microarray data often presents difficult
-challenges because the number of gene expression levels is typically much
-greater than the sample size. Consequently, several sparse and regularized
-classifiers have been proposed, often with strict covariance assumptions or
-eigenvalue adjustments to ensure that the classifiers are well-posed. Here, we
-propose a novel modification of the regularized discriminant analysis classifier
-popularized by \cite{Friedman:1989tm}. Our proposed classifier provides a
-straightforward interpretation and employs a new dimension-reduction technique
-to enable a computationally efficient model-selection procedure, thus overcoming
-the computational shortcomings of \citeauthor{Friedman:1989tm}'s classifier for
-high-dimensional data. Moreover, we demonstrate with several high-dimensional
-microarray data sets that our proposed classifier is competitive and often
-superior to several recently proposed sparse and regularized classifiers in
-terms of classification accuracy. Finally, we provide an implementation of our
-proposed classifier in the {\tt sparsediscrim} R package, available on CRAN.
-
-\end{abstract}
-
-\section{Introduction}
-
-Microarrays have enabled researchers to measure gene-expression levels of
-thousands of genes in a single experiment to identify a subset of genes and
-their expression patterns that either characterize a particular cell state or
-facilitate an accurate prediction of future cell state.  We focus on the
-classification of tissues and other biological samples into one of $K$ disjoint,
-predefined cancer types, where our goal is to ascertain the class membership of
-a new tissue sample based on gene-expression microarray data. However,
-microarray data typically consist of a large number of gene expressions $p$
-relative to the total sample size $N$. Moreover, standard benchmark microarray
-data sets usually consist of between $p = 10,000$ and $55,000$ probes and fewer
-than $N = 100$ observations, in which case standard supervised-learning methods
-exhibit poor classification performance because of the \emph{curse of
-  dimensionality} \citep{Bellman:1961tn}. Furthermore, some well-known
-classifiers, such as linear and quadratic discriminant analysis, are
-incalculable when $p > N$ because the sample covariance matrices employed are
-singular. As a result, many researchers, including \cite{Mai:2012bf},
-\cite{Fan:2012iq}, \cite{Merchante:2012vk}, \cite{Clemmensen:2011kr}, and
-\cite{Huang:2010ju}, have introduced restrictions on the population covariance
-matrices in linear and quadratic discriminant analysis to reduce number of
-covariance parameters that must be estimated to ensure that the resulting
-classifiers are calculable. For instance, \cite{Dudoit:2002ev} have modified the
-linear and quadratic discriminant analysis classifiers by assuming that the
-population covariance matrices are diagonal, resulting in classifiers that are
-special cases of the Naive Bayes classifier \citep[Chapter 4]{Hastie:2008dt}.
-
-Alternatively, several researchers, including \cite*{Zhang:2010va},
-\cite{Ye:2009gd}, \cite{Ji:2008wp}, \cite*{Guo:2007te},
-\cite*{Srivastava:2007ud}, \cite{Ye:2006vx}, and \cite{Ye:2005uu}, have proposed
-regularization methods to improve the estimation of the population covariance
-matrices before applying either linear or quadratic discriminant
-analysis. Covariance-matrix regularization methods often are based on the
-well-known ridge-regression approach of \cite{Hoerl:1970cd} and adjust the
-eigenvalues of the sample covariance matrices to ensure positive
-definiteness. For excellent overviews of the regularization problem in
-discriminant analysis, see \cite{Murphy:2012uq} and \cite{Mkhadri:1997gy}.
-
-Here, we propose an improvement to the popular regularized discriminant
-analysis classifier of \cite{Friedman:1989tm}, who has proposed pooling of the
-covariance-matrix estimators from linear and quadratic discriminant analysis and
-has additionally incorporated a covariance-matrix shrinkage term to ensure that
-the classifier is well-posed. \citeauthor{Friedman:1989tm}'s classifier can
-attain improved classification accuracy, compared to linear and quadratic
-discriminant analysis when $p < N$. However, in its original form,
-\citeauthor{Friedman:1989tm}'s classifier is impractical when $p \gg N$ because
-the decision rule requires the calculation of the inverses and determinants of
-large $p \times p$ covariance-matrix estimators for each class. Furthermore,
-\citeauthor{Friedman:1989tm}'s classifier has two tuning parameters that are
-selected via cross-validation, which adds a substantial computational burden to
-the model-selection process. \citeauthor{Friedman:1989tm}'s classifier is
-ill-advised on modern data sets because the model selection requires $O(Kp^3)$
-calculations for each tuning-parameter pair.
-
-In this paper we modify \citeauthor{Friedman:1989tm}'s classifier by first
-constructing the pooling matrix as a convex combination of the covariance-matrix
-estimators employed in linear and quadratic discriminant analysis and
-demonstrate that this simplification improves the interpretation of the pooling
-operation of \citeauthor{Friedman:1989tm}'s classifier. Given that the
-assumption of equal population covariance matrices within linear discriminant
-analysis is often too stringent in practice \citep{Clemmensen:2011kr}, our
-proposed classifier can relax the linearity assumption by instead incorporating
-unequal covariance matrices while retaining linear and quadratic discriminant
-analysis as special cases. We then establish our classifier by applying
-covariance-matrix regularization with a parameterization that includes a large
-family of shrinkage methods, such as those from \cite{Srivastava:2007ww},
-\cite{Rao:1971ul}, and other methods studied by \cite{Ramey:2013ji} and
-\cite{Xu:2009fl}.
-
-We show that our proposed classifier is invariant to any adjustments of the
-eigenvalues corresponding to the null space of the pooled sample covariance
-matrix. Applying a reasoning similar to \cite{Ye:2006tq}, we decompose the
-decision rule of our classifier and show that the sum related to the null space
-of the pooled sample covariance matrix can be discarded. Hence, no gain in
-classification performance can be obtained by adjusting these zero
-eigenvalues. This invariance property addresses the computational shortcomings
-of \citeauthor{Friedman:1989tm}'s classifier and provides a substantial increase
-in computational efficiency and model selection. In fact, we show that the
-inverse and determinants can be reduced to calculations of $n_k \times n_k$
-matrices, where $n_k$ is the sample size of the $k$th class $(k = 1, \ldots,
-K)$. This reduction in calculation is clearly advantageous because $n_k$ is
-often quite small in microarray data sets (e.g., 15--20 observations).
-
-We also demonstrate that our proposed classifier is competitive with and often
-superior to several recent classifiers designed for small-sample,
-high-dimensional data. In particular, we compare our classifier with diagonal
-linear discriminant analysis from \cite{Dudoit:2002ev} and the recently proposed
-classifiers from \cite{Guo:2007te}, \cite{Pang:2009ik}, \cite{Tong:2012hw}, and
-\cite{Witten:2011kc}. We conduct our comparison of the classification accuracy
-of the competing classifiers using four small-sample, high-dimensional
-benchmark microarray data sets.
-
-\section{Regularized Discriminant Analysis}
-\label{sec:rda}
-
-\subsection{Preliminaries}
-In discriminant analysis we wish to assign correctly an unlabeled
-$p$-dimensional observation vector $x$ to one of $K$ unique, known classes
-(or populations) by constructing a classifier from $N$ training
-observations. Let $x_i = (x_{i1}, \ldots, x_{ip}) \in \mathbb{R}_{p \times
-  1}$ be the $i$th observation $(i = 1, \ldots, N)$ with true, unique membership
-$y_i \in \{\omega_1, \ldots, \omega_K\}$, where $\mathbb{R}_{a \times b}$
-denotes the matrix space of all $a \times b$ matrices over the real field
-$\mathbb{R}$. We assume that $(x_i, y_i)$ is a realization from a mixture
-distribution $p(x) = \sum_{k=1}^K p(x | \omega_k) p(\omega_k)$, where
-$p(x | \omega_k)$ is the probability density function of the $k$th class and
-$p(\omega_k)$ is the prior probability of class membership of the $k$th
-class. We further assume that each of $p(\omega_k)$ is equal $(k = 1, \ldots,
-K)$.
-
-Quadratic discriminant analysis is the optimal Bayesian decision rule with
-respect to a $0-1$ loss function when $p(x | \omega_k)$ is the probability
-density function of the multivariate normal distribution with known mean vector
-$\mu_k \in \mathbb{R}_{p \times 1}$ and known covariance matrix $\Sigma_k
-\in \mathbb{R}_{p \times p}^{>}$, $k = 1, 2, \ldots, K$, where $\mathbb{R}_{p
-  \times p}^{>}$ denotes the cone of real $p \times p$ positive-definite
-matrices. Because $\mu_k$ and $\Sigma_k$ are typically unknown, we assign
-an unlabeled observation $x$ to class $\omega_k$ with the sample quadratic
-discriminant analysis classifier
-\begin{align}
-	D_{QDA}(x) = \argmin_{k}  (x - \xbar_k)'\widehat{\Sigma}_k^{-1}(x - \xbar_k)  + \log |\widehat{\Sigma}_k|, \label{eq:qda}
-\end{align}
-where $\xbar_k$ and $\widehat{\Sigma}_k$ are the maximum likelihood
-estimators for $\mu_k$ and $\Sigma_k$, respectively. If we assume
-further that $\Sigma_k = \Sigma$ $(k = 1, \ldots, K)$, then the pooled
-sample covariance matrix $\widehat{\Sigma}$ is substituted for $\widehat{
-  \Sigma}_k$ in \eqref{eq:qda}, where
-\begin{align}
-	\widehat{\Sigma} = N^{-1} \sum_{k=1}^K n_k \widehat{\Sigma}_k \label{eq:pooled-cov}
-\end{align}
-is the maximum likelihood estimator for $\Sigma$. In this case \eqref{eq:qda}
-reduces to the sample linear discriminant analysis classifier, and the
-log-determinant is omitted because it is constant across all classes.
-
-Throughout the document we also require the following additional notation. Let
-$I_m$ be the $m \times m$ identity matrix, and let $0_{m \times p}$ be
-the $m \times p$ matrix of zeros, such that $0_m$ is understood to denote
-$0_{m \times m}$. Lastly, for $c \in \mathbb{R}$, let $c^+ = 1/c$ if $c \ne
-0$ and $0$ otherwise. Let $A \oplus B$ denote the direct sum of $A \in
-\mathbb{R}_{r \times r}$ and $B \in \mathbb{R}_{s \times s}$ \citep[Chapter
-  1]{Lutkepohl:1996uz}. Let $A^+$ and $\mathcal{N}(A) = \{z \in \mathbb{R}_{p
-  \times 1} : A z = 0_{m \times 1}\}$ denote the Moore-Penrose pseudoinverse and null space
-of $A \in \mathbb{R}_{m \times p}$, respectively. Let $V^{\perp} = \{y
-\in \mathbb{R}_{p \times 1} : y'v = 0\ \forall v \in V \}$ denote
-the orthogonal complement of a vector space $V \subset \mathbb{R}_{p \times 1}$.
-
-\subsection{Covariance Matrix Regularization}
-
-The smallest eigenvalues and the directions associated with their eigenvectors
-highly influence \eqref{eq:qda}. In fact, the eigenvalues of $\widehat{
-  \Sigma}_k$ are well-known to be biased if $p \ge n_k$ such that the smallest
-eigenvalues are underestimated \citep{Seber:2004uh}, Moreover, if $p > n_k$,
-then rank$(\widehat{\Sigma}_k) \le n_k$, which implies that at least $p -
-n_k$ eigenvalues of $\widehat{\Sigma}_k$ are 0. Furthermore, although more
-feature information is available to discriminate among the $K$ classes, if $p >
-n_k$, \eqref{eq:qda} is incalculable because $\widehat{\Sigma}_k^{-1}$ does
-not exist.
-
-Several regularization methods, such as the methods considered by
-\citet*{Guo:2007te}, \cite{Mkhadri:1995jp}, and \citet*{Xu:2009fl}, have been
-proposed in the literature to adjust the eigenvalues of $\widehat{\Sigma}_k$
-so that \eqref{eq:qda} is calculable and has reduced variability in the standard
-bias-variance tradeoff. A common form of the covariance-matrix regularization
-applies a shrinkage factor $\gamma > 0$, so that
-\begin{align}
-	\widehat{\Sigma}_k(\gamma) =  \widehat{\Sigma}_k + \gamma I_p, \label{eq:ridge-estimator}
-\end{align}
-similar to a method employed in ridge regression \citep{Hoerl:1970cd}. Equation
-\eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance
-matrix $\widehat{\Sigma}_k$ towards $I_p$, thereby increasing the
-eigenvalues of $\widehat{\Sigma}_k$ by $\gamma$. Specifically, the zero
-eigenvalues are replaced with $\gamma$, so that \eqref{eq:ridge-estimator} is
-positive definite. For additional covariance-matrix regularization methods, see
-\cite{Ramey:2013ji}, \cite{Xu:2009fl}, and \cite{Ye:2009gd}.
-
-\subsection{Friedman's Regularized Discriminant Analysis Classifier}
-
-\cite{Friedman:1989tm} has proposed regularized discriminant analysis by
-incorporating a weighted average of the sample covariance matrix $\widehat{
-  \Sigma}_k$ and the pooled sample covariance matrix $\widehat{\Sigma}$ to
-estimate $\Sigma_k$ with
-\begin{align}
-  \widehat{\Sigma}_k(\lambda) = n_k^{-1}(\lambda) S_k(\lambda) \quad (k =
-  1, \ldots, K),\label{eq:sig-lambda}
-\end{align}
-where $\lambda \in [0, 1]$, $n_k(\lambda) = (1 - \lambda) n_k + \lambda N$, $
-S_k = n_k \widehat{\Sigma}_k$, $S = \sum_{k=1}^K S_k$, and $
-S_k(\lambda) = (1 - \lambda) S_k + \lambda S$. We can interpret
-\eqref{eq:sig-lambda} as a covariance-matrix estimator that borrows from
-$\widehat{\Sigma}$ in \eqref{eq:pooled-cov} to better estimate $
-\Sigma_k$. If $\lambda = 0$, \eqref{eq:sig-lambda} is equal to
-$\widehat{\Sigma}_k$ used in \eqref{eq:qda}. Alternatively, if $\lambda = 1$,
-\eqref{eq:sig-lambda} reduces to \eqref{eq:pooled-cov}.
-
-To further improve the estimation of $\Sigma_k$ and to stabilize the inverse
-of \eqref{eq:sig-lambda}, \cite{Friedman:1989tm} has proposed the biased
-covariance-matrix estimator
-\begin{align}
-	\widehat{\Sigma}_k(\lambda, \gamma) = (1 - \gamma) \widehat{\Sigma}_k(\lambda) + \gamma \frac{\tr\{\widehat{\Sigma}_k(\lambda)\}}{p} I_p,\label{eq:sig-rda}
-\end{align}
-where $\gamma \in [0, 1]$ is a regularization parameter that controls the
-shrinkage of \eqref{eq:sig-rda} towards $I_p$ and is weighted by the average
-of the eigenvalues of \eqref{eq:sig-lambda}. Thus, the \emph{pooling} parameter
-$\lambda$ controls the amount borrowed from $\widehat{\Sigma}$ to estimate
-$\Sigma_k$, and the \emph{shrinkage} parameter $\gamma$ determines the
-amount of eigenvalue shrinkage applied. Finally, by substituting
-\eqref{eq:sig-rda} into \eqref{eq:qda}, we have the regularized discriminant
-analysis classifier
-\begin{align}
-	D_{RDA}(x) = \argmin_{k}  (x - \xbar_k)'\widehat{\Sigma}_k(\lambda, \gamma)^{-1}(x - \xbar_k)  + \log |\widehat{\Sigma}_k(\lambda, \gamma)|. \label{eq:rda}
-\end{align}
-If $\gamma$ is sufficiently small, \eqref{eq:sig-rda} can be singular or
-near-singular. For this case \cite{Friedman:1989tm} has suggested replacing the
-zero eigenvalues of \eqref{eq:sig-rda} with a small positive constant that
-permits a numerically stable matrix inversion.
-
-\section{High-Dimensional Regularized Discriminant Analysis}
-
-\subsection{Improvements to \citeauthor{Friedman:1989tm}'s Classifier}
-
-Here, we define our adaptation of \citeauthor{Friedman:1989tm}'s classifier by
-first simplifying the formulation of $\widehat{\Sigma}_k(\lambda)$ in
-\eqref{eq:sig-lambda} and demonstrating its clear interpretation as a linear
-combination of the crossproducts of the training observations centered by their
-respective class sample means. We define the convex combination
-\begin{align}
-  \widehat{\Sigma}_k(\lambda) = (1 - \lambda) \widehat{\Sigma}_k + \lambda
-  \widehat{\Sigma}, \quad (k = 1, \ldots, K)\label{eq:sig-lambda-alternative}
-\end{align}
-where $\lambda \in [0, 1]$ is the \emph{pooling} parameter. Thus, we avoid the
-convex combination of $n_k$ and $N$ as defined by \cite{Friedman:1989tm}. By
-rewriting \eqref{eq:sig-lambda-alternative} in terms of the observations $
-x_i$ $(i = 1, \ldots, N)$, each centered by its class sample mean, we attain a
-clear interpretation of $\widehat{\Sigma}_k(\lambda)$. That is,
-\begin{align}
-	\widehat{\Sigma}_k(\lambda) &= \widehat{\Sigma}_k + \lambda (\widehat{\Sigma} - \widehat{\Sigma}_k) \nonumber\\
-%	&= \widehat{\Sigma}_k + \lambda \left(\sum_{k' = 1}^K \frac{n_k}{N} \widehat{\Sigma}_{k'}  - \widehat{\Sigma}_k \right) \nonumber\\
-	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right) \widehat{\Sigma}_k +  \frac{\lambda}{N} \sum_{\substack{k' = 1\\k' \ne k}} n_{k'} \widehat{\Sigma}_{k'} \nonumber \\
-%	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N I(y_i = k) x_i x_i' +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k) x_i x_i' \nonumber \\
-	&= \sum_{i=1}^N c_{ik}(\lambda) x_i x_i',\label{eq:sig-lambda-alternative2}
-\end{align}
-where $c_{ik}(\lambda) = \lambda N^{-1} + (1 - \lambda)n_k^{-1}I(y_i = k)$.
-From \eqref{eq:sig-lambda-alternative2}, $\lambda$ weights the contribution of
-each of the $N$ observations to estimate $\Sigma_k$. Hence, for $0 < \lambda
-\le 1$, we estimate $\Sigma_k$ with $N$ rather than $n_k$
-observations. Below, we will show that the pooling operation is advantageous in
-increasing the rank of each $\widehat{\Sigma}_k(\lambda)$ from
-rank$(\widehat{\Sigma}_k)$ to rank$(\widehat{\Sigma})$ for $0 < \lambda
-\le 1$. Notice that if $\lambda = 0$, then the observations from the remaining
-$K - 1$ classes do not contribute to the estimation of $\Sigma_k$,
-corresponding to $\widehat{\Sigma}_k$. Furthermore, if $\lambda = 1$, the
-weights in \eqref{eq:sig-lambda-alternative2} reduce to $1/N$, corresponding to
-$\widehat{\Sigma}$. For brevity, when $\lambda = 1$, we define $X =
-[\sqrt{c_{1k}(1)} x_1', \ldots, \sqrt{c_{Nk}(1)} x_N']'$ such that
-$\widehat{\Sigma} = X' X$. Similarly, for $\lambda = 0$, we define
-$X_k = [\sqrt{c_{1k}(0)} x_1', \ldots, \sqrt{c_{Nk}(0)} x_N']'$ such
-that $\widehat{\Sigma}_k = X_k' X_k$. In Figure
-\ref{fig:hdrda-contours} we plot the contours of five multivariate normal
-populations for $\lambda = 0$ with unequal covariance matrices. As $\lambda$
-approaches 1, the contours become more similar, resulting in identical contours
-for $\lambda = 1$.
-
-\[ \left[\text{Figure \ref{fig:hdrda-contours} goes about here }\right]\]
-
-As we have discussed above, several eigenvalue adjustment methods have been
-proposed to increase eigenvalues (approximately) equal to 0. Here, we define the
-eigenvalue adjustment of \eqref{eq:sig-lambda-alternative} as
-\begin{align}
-	\tilde{\Sigma}_k = \alpha_k \widehat{\Sigma}_k(\lambda) + \gamma I_p,\label{eq:hdrda-cov}
-\end{align}
-where $\alpha_k \ge 0$ and $\gamma \ge 0$ is an eigenvalue-shrinkage
-constant. The choice of $\alpha_k$ allows for a flexible formulation of
-covariance-matrix estimators. For instance, if $\alpha_k = 1$, $k = 1, \ldots,
-K$, then \eqref{eq:hdrda-cov} resembles \eqref{eq:ridge-estimator}. Similarly,
-if $\alpha_k = 1 - \gamma$, then \eqref{eq:hdrda-cov} has a form similar to
-\eqref{eq:sig-rda}. Now, substituting \eqref{eq:hdrda-cov} into \eqref{eq:qda},
-we define the high-dimensional regularized discriminant analysis classifier
-\begin{align}
-	D_{HDRDA}(x) = \argmin_{k}  (x - \xbar_k)'\tilde{\Sigma}_k^{+}(x - \xbar_k)  + \log |\tilde{\Sigma}_k|. \label{eq:hdrda}
-\end{align}
-For $\gamma > 0$, $\tilde{\Sigma}_k$ is nonsingular such that
-$\tilde{\Sigma}_k^{-1}$ can be substituted for $\tilde{\Sigma}_k^{+}$ in
-\eqref{eq:hdrda}. If $\gamma = 0$, we explicitly set $|\tilde{\Sigma}_k|$
-equal to the product of the positive eigenvalues of $\tilde{\Sigma}_k$.
-
-\subsection{Enhanced Computational Efficiency}
-
-Next, we demonstrate that our proposed classifier in \eqref{eq:hdrda} lends
-itself to a more efficient calculation and a rapid model selection of the
-parameters $\lambda$ and $\gamma$\footnote{TODO: Briefly mention the updating
-  trick employed in the R package}. We decompose \eqref{eq:hdrda} into a sum of
-two components, where the first summand consists of matrix operations applied to
-low-dimensional matrices and the second summand corresponds to the null space of
-$\widehat{\Sigma}$ in \eqref{eq:pooled-cov}. We show that the matrix
-operations performed on the null space of $\widehat{\Sigma}$ yield constant
-quadratic forms across all classes and can be omitted. For $p \gg N$, the
-constant component involves determinants and inverses of high-dimensional
-matrices, and by ignoring these calculations, a substantial reduction in
-computational costs is achieved relative to Friedman's classifier in
-\eqref{eq:rda}\footnote{TODO: Add a brief comparison of the computations of
-  \emph{RDA} and \emph{HDRDA} in terms of Big Oh notation}. Furthermore, a
-byproduct is that adjustments to the associated eigenvalues have no effect on
-\eqref{eq:hdrda}. Lastly, we utilize the singular value decomposition to
-calculate efficiently the eigenvalue decomposition of $\widehat{\Sigma}$,
-further reducing the computational costs of our proposed classifier.
-
-First, we require the following relationship regarding the null spaces of
-$\widehat{\Sigma}_k(\lambda)$, $\widehat{\Sigma}$, and $\widehat{
-  \Sigma}_k$.
-
-\begin{lemma}\label{lemma:null-spaces}
-Let $\widehat{\Sigma}_k$ and $\widehat{\Sigma}$ be the maximum
-likelihood estimators of $\Sigma_k$ and $\Sigma$, respectively. Let
-$\widehat{\Sigma}_k(\lambda)$ be defined as in
-\eqref{eq:sig-lambda-alternative}. Then, $\mathcal{N}\{\widehat{\Sigma}_k(\lambda)\} \subset \mathcal{N}(\widehat{
-  \Sigma}) \subset \mathcal{N}(\widehat{\Sigma}_k)$ $(k = 1, \ldots, K)$.
-\end{lemma}
-\begin{proof}
-Let $z \in \mathcal{N}\{\widehat{\Sigma}_k(\lambda)\}$ for some $k = 1,
-\ldots, K$. Hence, $0 = z' \widehat{\Sigma}_k(\lambda) z = (1 -
-\lambda) z' \widehat{\Sigma}_k z + \lambda z' \widehat{
-  \Sigma} z$. Because $\widehat{\Sigma}_k, \widehat{\Sigma}\in
-\mathbb{R}_{p \times p}^{\ge}$, we have $z \in \mathcal{N}(\widehat{
-  \Sigma})$ and $z \in \mathcal{N}(\widehat{\Sigma}_k)$. In particular,
-we have that $\mathcal{N}\{\widehat{\Sigma}_k(\lambda)\} \subset
-\mathcal{N}(\widehat{\Sigma})$. Now, suppose $z \in
-\mathcal{N}(\widehat{\Sigma})$. Similarly, we have that $0 = z'
-\widehat{\Sigma} z = N^{-1} \sum_{k = 1}^K n_k z' \widehat{
-  \Sigma}_k z$, which implies that $z \in \mathcal{N}(\widehat{
-  \Sigma}_k)$ because $\widehat{\Sigma}_k \in \mathbb{R}_{p \times
-  p}^{\ge}$. Therefore, $\mathcal{N}(\widehat{\Sigma}) \subset
-\mathcal{N}(\widehat{\Sigma}_k)$.
-\end{proof}
-
-In Lemma \ref{lemma:rda-tilde-Sigma_k} below, we derive an alternative
-expression for $\tilde{\Sigma}_k$ in terms of the matrix of eigenvectors of
-$\widehat{\Sigma}$. Let $\widehat{\Sigma} = U D U'$ be the
-eigendecomposition of $\widehat{\Sigma}$ such that $D \in \mathbb{R}_{p
-  \times p}^{\ge}$ is the diagonal matrix of eigenvalues of $\widehat{\Sigma}$
-with $D = D_q \oplus 0_{p-q}$, $D_q \in \mathbb{R}_{q \times
-  q}^{>}$ is the diagonal matrix consisting of the positive eigenvalues of
-$\widehat{\Sigma}$, the columns of $U \in \mathbb{R}_{p \times p}$ are
-the corresponding orthonormal eigenvectors of $\widehat{\Sigma}$, and
-rank$(\widehat{\Sigma}) = q$. Then, we partition $U = (U_1, 
-U_2)$ such that $U_1 \in \mathbb{R}_{p \times q}$ and $U_2 \in
-\mathbb{R}_{p \times (p - q)}$.
-
-\begin{lemma}\label{lemma:rda-tilde-Sigma_k}
-Let $\widehat{\Sigma} = U D U'$ be the eigendecomposition of
-$\widehat{\Sigma}$ as above, and suppose that rank$(\widehat{\Sigma}) =
-q \le p$. Then, we have
-\begin{align}
-	\tilde{\Sigma}_k &= U(W_k \oplus \gamma I_{p-q})U' \quad
-  (k = 1, \ldots, K),\label{eq:rda-matrix}
-\intertext{where}
-W_k &= \alpha_k \{(1 - \lambda) U_1' \widehat{\Sigma}_k U_1 + \lambda D_q\} + \gamma I_{q}.\label{eq:Wk}
-\end{align}
-\end{lemma}
-\begin{proof}
-From Lemma \ref{lemma:null-spaces}, the columns of $U_2$ span the null space
-of $\widehat{\Sigma}$, which implies that $\widehat{\Sigma}_k U_2 =
-0_{p \times (p - q)}$. Hence, $U' \widehat{\Sigma}_k U = 
-U_1' \widehat{\Sigma}_k U_1 \oplus 0_{p-q}$ $(k = 1, \ldots,
-K)$. Thus, $U' \tilde{\Sigma}_k U = \alpha_k \{(1 - \lambda) U'
-\widehat{\Sigma}_k U + \lambda D\} + \gamma I_p$, and
-\eqref{eq:rda-matrix} follows because $U$ is orthogonal.
-\end{proof}
-
-As an immediate immediate consequence of Lemma \ref{lemma:rda-tilde-Sigma_k}, we
-have the following corollary.
-
-\begin{cor}
-Let $\widehat{\Sigma}_k(\lambda)$ be defined as in
-\eqref{eq:sig-lambda-alternative}. Then, for $\lambda \in (0, 1]$,
-  rank$\{\widehat{\Sigma}_k(\lambda)\} = q$ $(k = 1, \ldots, K)$.
-\end{cor}
-\begin{proof}
-The proof follows by setting $\gamma = 0$ in Lemma \ref{lemma:rda-tilde-Sigma_k}.
-\end{proof}
-
-Thus, by incorporating each $x_i$ into the estimation of $\Sigma_k$, we
-increase the rank of $\widehat{\Sigma}_k(\lambda)$ to $q$ if $\lambda \ne
-0$. Next, we provide an essential result that enables us to prove that
-\eqref{eq:hdrda} is invariant to adjustments to the eigenvalues of $\tilde{
-  \Sigma}_k$ corresponding to the null space of $\widehat{\Sigma}$ for any
-unlabeled observation $x \in \mathbb{R}_{p \times 1}$.
-
-\begin{lemma}\label{lemma:RDA-constant-term}
-Let $U_2$ be defined as above. Then, for all $x \in \mathbb{R}_{p \times
-  1}$, $U_2' (x - \xbar_k) = U_2' (x - \xbar_{k'})$ $(1 \le k,
-k' \le K)$.
-\end{lemma}
-\begin{proof}
-Let $x \in \mathbb{R}_{p \times 1}$, and suppose that $1 \le k, k' \le K$.
-Recall that $U_2 \in \mathcal{N}(\widehat{\Sigma})$, which implies that
-$U_2' \in \mathcal{C}(\widehat{\Sigma})^{\perp}$ \citep[Lemma
-  1.2.5]{Kollo:2005vp}. Now, because $x_i \in \mathcal{C}(\widehat{
-  \Sigma})$ $(i = 1, \ldots, N)$, $U_2' x_i = 0_{(p-q) \times 1}$.
-Hence, $0_{(p-q) \times 1} = \sum_{i=1}^N \beta_i U_2'x_i = U_2'(\xbar_k -
-\xbar_{k'})$, where $\beta_i = (n_k n_{k'})^{-1} \{ I(y_i = k) n_{k'} - I(y_i =
-k') n_k \}$. Therefore, $U_2' (x - \xbar_k) = U_2' (x - \xbar_{k'})$.
-\end{proof}
-
-We now present our main result, where we decompose the decision rule of our
-proposed classifier and show that the term requiring the largest computational
-costs does not contribute to the classification of an unlabeled observation
-performed using Lemma \ref{lemma:RDA-constant-term}. Hence, we reduce
-\eqref{eq:hdrda} to an equivalent, more computationally efficient decision rule.
-
-\begin{thm}
-Let $\tilde{\Sigma}_k$ and $W_k$ be defined as in \eqref{eq:rda-matrix}
-and \eqref{eq:Wk}, respectively, and let $U_1$ be defined as above. Then,
-the decision rule in \eqref{eq:hdrda} is equivalent to
-	\begin{align}
-		D_{HDRDA}(x) &= \argmin_k  (x - \xbar_k)' U_1 W_k^{-1} U_1' (x - \xbar_k) + \log | W_k |. \label{eq:hdrda-decomposed}
-	\end{align}
-\end{thm}
-\begin{proof}
-From \eqref{eq:rda-matrix}, we have that $\tilde{\Sigma}_k^{+} = U(
-W_k^{-1} \oplus \gamma^{+} I_{p-q} )U'$ and $|\tilde{\Sigma}_k| =
-\gamma^{p-q}| W_k |$ $(k = 1, \ldots, K)$. Therefore, for all $x \in
-\mathbb{R}_{p \times 1}$, we have that
-	\begin{align*}
-	(x - \xbar_k)' \tilde{\Sigma}_k^{+}(x - \xbar_k)  + \log |\tilde{\Sigma}_k| &= (x - \xbar_k)' U_1 W_k^{-1} U_1' (x - \xbar_k)\\
-	&+ \gamma^{+} (x - \xbar_k)' U_2 U_2' (x - \xbar_k) + \log | W_k | + (p - q) \log \gamma.
-	\end{align*}
-Because $\gamma$ is constant for $k = 1, \ldots, K$, we can omit the $(p - q)
-\log \gamma$ term and particularly avoid the calculation of $\log 0$ for $\gamma
-= 0$. Then, the proof follows from Lemma \ref{lemma:RDA-constant-term} because
-$U_2' (x - \xbar_k)$ is constant for $k = 1, \ldots, K$.
-\end{proof}
-
-\subsection{Model Selection}
-
-Thus far, we have shown that our classifier in \eqref{eq:hdrda} is invariant to
-the term $U_2$, thus yielding an equivalent classifier in
-\eqref{eq:hdrda-decomposed} with a substantial reduction in the computational
-complexity. Here, we provide two\footnote{TODO: Two? Perhaps, I mean one and
-  then the Fast SVD below. If so, clarify this.} additional results that further
-improve the computational performance of our classifier. First, we demonstrate
-that the inverse and determinants of $W_k$ can be written so that they are
-performed on $n_k \times n_k$ matrices rather than the larger $q \times q$
-matrices, resulting in further improvements to the computational efficiency.
-
-\begin{proposition}\label{proposition:rda-W_k}
-Let $W_k$ be defined as above $(k = 1, \ldots, K)$. Then, $|W_k| = |
-\Gamma_k| |Q_k|$ and
-	\begin{align}
-		W_k^{-1} &= \Gamma_k^{-1} - \alpha_k(1 - \lambda) \Gamma_k^{-1} U_1' X_k' Q_k^{-1} X_k U_1 \Gamma_k^{-1},\label{eq:W_k_inv}
-	\end{align}
-where $Q_k = I_{n_k} + \alpha_k(1- \lambda) X_k U_1 
-\Gamma_k^{-1} U_1' X_k'$ and $\Gamma_k = \alpha_k \lambda D_q +
-\gamma I_q$.
-\end{proposition}
-\begin{proof}
-To calculate $|W_k|$, we apply Theorem 18.1.1 from \cite{Harville:2008wja},
-which states that $|A + B T C| = |A| |T| |T^{-1} +
-C A^{-1} B|$. Thus, setting $A = \Gamma_k$, $B =
-\alpha_k (1 - \lambda) U_1' X_k'$, $T = I_{n_k}$, and $C =
-X_k U_1$, we have $|W_k| = |\Gamma_k| |Q_k|$. Also,
-\eqref{eq:W_k_inv} follows from the well-known Sherman-Woodbury formula
-\citep[Theorem 18.2.8]{Harville:2008wja}.
-\end{proof}
-
-Thus far, we have constructed each $\tilde{\Sigma}_k$ in terms of the first
-$q$ eigenvectors of $\widehat{\Sigma}$. By utilizing a technique similar to
-the so-called kernel trick, we can obtain $U_1$ by computing the
-eigendecomposition of a much smaller $N \times N$ matrix when $p \gg N$
-\citep{Hastie:2008dt}. Recall that the pooled sample covariance matrix
-$\widehat{\Sigma} = X' X$. Then, because $X = M D^{1/2}
-U'$ by the singular value decomposition, we have the eigendecomposition
-$\widehat{\Sigma} = U D U'$, where $M \in \mathbb{R}_{N
-  \times N}$ is orthogonal, $D^{1/2} \in \mathbb{R}_{N \times N}^{\ge}$ is a
-diagonal matrix consisting of the singular values of $X$, and $U \in
-\mathbb{R}_{p \times N}$ is orthogonal. Now, $X X' = M D M'$
-is the eigenvalue decomposition of $X X'$, where $D$ is the matrix
-of eigenvalues of $X X'$ and the columns of $M$ are the
-corresponding eigenvectors. Hence, we obtain $M$ and $D$ from an
-eigendecomposition of $X X' \in \mathbb{R}_{N \times N}$. Next, we
-compute $U = X' M D^{+/2}$, where $D^{+/2} = D_q^{-1/2}
-\oplus 0_{N-q}$. We then determine $q$ by computing the number of
-numerically nonzero eigenvalues present in $D$, where $q$ is the number of
-eigenvalues that exceeds some tolerance value, say, $\epsilon = 1 \times
-10^{-6}$. After calculating $q$, we extract $U_1$ by retaining only the
-first $q$ columns of $U$.
-
-\section{High-Dimensional Microarray Data Sets}
-
-In this section, we describe four high-dimensional microarray data sets and
-compare our proposed classifier with five classifiers recently proposed for
-small-sample, high-dimensional data. Our competitors include diagonal linear
-discriminant analysis from \cite{Dudoit:2002ev}, shrunken centroids regularized
-discriminant analysis from \cite{Guo:2007te}, and penalized linear discriminant
-analysis from \cite{Witten:2011kc}. We also consider two modifications of
-diagonal linear discriminant analysis from \cite{Tong:2012hw} and
-\cite{Pang:2009ik}, where the former employs an improved mean estimator and the
-latter utilizes an improved variance estimator. Below, we refer to each
-classifier by the first author's surname.
-
-We evaluated the classification performance of each classifier by randomly
-partitioning the data such that $2/3$ of the observations were allocated as
-training data and the remaining $1/3$ of the observations were allocated as a
-test data set. For each random partition, we applied the feature-selection
-method from \cite{Dudoit:2002ev} to the training data to choose the $d$ features
-having the largest ratio of their between-group to within-group sums of squares.
-We then filtered the test data set with the corresponding $d$ features. After
-training each classifier from the filtered training data, we classified the
-filtered test data sets and computed the proportion of mislabeled test
-observations to estimate the classification error rate for each
-classifier. Repeating this process 1000 times, we computed the average of the
-error-rate estimates for each classifier. We remark that our simulation design
-avoids the selection bias discussed in \cite{Ambroise:2002fa}. We did not
-include \citeauthor{Friedman:1989tm}'s classifier in our studies to avoid its
-excessive computations. In Table \ref{tab:data-summary}, we summarize the
-attributes of each microarray data set.
-
-We used version 3.0.0 of the open-source statistical software {\tt R} for the
-classification study. For each classifier's {\tt R} implementation, we used the
-default settings with the exception that we explicitly set prior probabilities
-as equal. For our classifier, we set $\alpha_k = 1$ $(k = 1, \ldots, K)$, so
-that the covariance-matrix estimator \eqref{eq:hdrda-cov} employed had a form
-resembling \eqref{eq:ridge-estimator}. We estimated $\lambda$ from a grid of 21
-equidistant candidate values between 0 and 1, inclusively. Similarly, we
-estimated $\gamma$ from a grid consisting of the values $10^{-1}, \ldots, 10^4$,
-and $10^5$. We selected optimal estimates of $\lambda$ and $\gamma$ using
-10-fold cross-validation \citep{Hastie:2008dt}.
-
-\[ \left[\text{Table 1 goes about here }\right]\]
-
-\subsection{\cite{Burczynski:2006ik} Data Set}
-
-\cite{Burczynski:2006ik} acquired the peripheral blood mononuclear cells (PBMC)
-through hybridization to microarrays from 127 individuals resulting in 22,283
-sequences. Of the 127 individuals, 42 were healthy, 59 had Crohn's disease (CD),
-and 26 had ulcerative colitis (UC). The goal of \cite{Burczynski:2006ik} was to
-improve accuracy in the discrimination of Inflammatory Bowel Disease (IBD) using
-the PBMC-based gene expression signature of a patient.
-
-\subsection{\cite{Nakayama:2007fl} Data Set}
-
-\cite{Nakayama:2007fl} acquired the gene expression through an oligonucleotide
-microarray from 105 samples of 10 types of soft tissue tumors. This included 16
-samples of synovial sarcoma (SS), 19 samples of myxoid/round cell liposarcoma
-(MLS), 3 samples of lipoma, 3 samples of well-differentiated liposarcoma (WDLS),
-15 samples of dedifferentiated liposarcoma (DDLS), 15 samples of
-myxofibrosarcoma (MFS), 6 samples of leiomyosarcoma (LMS), 3 samples of
-malignant nerve sheathe tumor (MPNST), 4 samples of fibrosarcoma (FS), and 21
-samples of malignant fibrous histiocytoma (MFH). \cite{Nakayama:2007fl}
-determined from their data that these 10 types fell into 4 broader groups: (1)
-SS; (2) MLS; (3) Lipoma, WDLS, and part of DDLS; (4) Spindle cell and
-pleomorophic sarcomas including DDLS, MFS, LMS, MPNST, FS, and MFH. Following
-\cite{Witten:2011kc}, we restrict our analysis to the five tumor types having at
-least 15 observations.
-
-\subsection{\cite{Shipp:2002ka} Data Set}
-
-Approximately 30-40\% of adult non-Hodgkin lymphomas are diffuse large B-cell
-lymphomas (DLBCLs) \cite{Shipp:2002ka}.  However, only a small proportion of
-DLBCL patients are cured with modern chemotherapeutic regimens.  Several models
-have been proposed, such as the International Prognostic Index (IPI), to
-determine a patient's curability.  These models rely on clinical covariates,
-such as age, to determine if the patient can be cured and are often
-ineffective. \cite{Shipp:2002ka} have argued that more effective means are
-desired to determine a patient's curability. The authors measured 6,817 gene
-expression levels from 58 DLBCL patient samples with customized cDNA
-(lymphochip) microarrays to investigate the curability of patients treated with
-cyclophosphamide, adriamycin, vincristine, and prednisone (CHOP)-based
-chemotherapy.  Among the 58 DLBCL patient samples, 32 are from cured patients
-while 26 are from patients with fatal or refractory disease.
-
-\subsection{\cite{Singh:2002fh} Data Set}
-
-\cite{Singh:2002fh} have examined 235 radical prostatectomy specimens from
-surgery patients between 1995 and 1997. The authors used oligonucleotide
-microarrays containing probes for approximately 12,600 genes and expressed
-sequence tags. They have reported that 102 of the radical prostatectomy
-specimens are of high quality: 52 prostate tumor samples and 50 non-tumor
-prostate samples.
-
-\subsection{Classification Results}
-
-For the microarray data sets, we computed error-rate estimates for the competing
-classifiers after reducing the dimension of the data sets with feature selection
-to $d = 100, \ldots, 1000$, similar to the values of $d$ considered by
-\cite{Xu:2009fl}. To compare misclassification rates between different values of
-$d$, we applied each value of $d$ considered to the same random allocation of
-training and test data. Hence, for a given random partition, the set of features
-selected for a small value of $d$ is a proper subset of the features selected
-for a larger value of $d$. We report the average of the error-rate estimates for
-each value of $d$ in Figure \ref{fig:error-rates}. Approximate standard errors
-were no greater than \Sexpr{max(error_rates[["SE_Error"]])}.
-
-Our classifier was superior in classification performance for the majority of
-the simulations. For the Burczysnki, Shipp, and Singh data sets, our classifier
-outperformed its competitors for all values of $d$ examined. For the Nakayama
-data set, our classifier was outperformed by the Dudoit, Pang, and Tong
-classifiers for small values of $d$ but attained classification rates superior
-to its competitors for larger values of $d$. Furthermore, Figure
-\ref{fig:error-rates} emphasizes two strengths of our classifier: reliability of
-the classifier for small $d$ and stability of the classifier for large $d$. In
-contrast to the other classifiers, our classifier exhibited less variability in
-its misclassification estimates for small values of $d$. Also, our classifier
-yielded superior classification accuracy as $d$ increased and was hardly
-effected by the inclusion of more variables in the data sets. We attribute the
-apparent stability of the error-rate estimates of our classifier for large
-values of $d$ to its novel dimension-reduction technique.
-
-The error-rate estimates for the Dudoit, Pang, Tong, and Witten classifiers were
-comparable for all values of $d$ examined. Furthermore, the error-rate estimates
-of the Dudoit, Pang, and Tong classifiers were approximately equal throughout
-our study, and the Witten classifier yielded approximately equal error-rate
-estimates for the Burczysnki, Shipp, and Singh data sets. Furthermore, the
-error-rate estimates for the Dudoit and Tong classifiers were nearly identical,
-which suggests that the improved mean estimator employed by \cite{Tong:2012hw}
-provided little improvement to the classification performance of diagonal linear
-discriminant analysis.
-
-The classification performance of the the Dudoit, Pang, Tong, and Witten
-classifiers for the Shipp and Singh data sets revealed the inability of these
-methods to handle larger values of $d$. After closer examination of the
-classification studies performed in the authors' original papers, we noticed
-that classification accuracy estimates were reported only for small values of
-$d$, thereby ignoring the inclusion of a larger number of variables. Hence, we
-advise that practioners be aware of the necessity to filter a large number of
-features before employing these methods, lest their classification performance
-can degrade.
-
-The popular Guo classifier is most competitive over large values of $d$ for the
-Shipp and Singh data sets. However, for the Burczynski and Nakayama data sets,
-the implementation of the Guo method in the {\tt rda} package often classified
-every test observation into a single class, thus yielding larger error rates
-than its competitors. Furthermore, the Guo classifier exhibited unreliable
-classification performance for smaller values of $d$. The Guo classifier
-required a larger subset of the features to perform competitively in our study.
-
-\[ \left[\text{Figure \ref{fig:error-rates} goes about here }\right]\]
-
-Next, we examine the optimal estimates of the tuning parameters for our
-classifier. For brevity, we provide graphical summaries of the optimal estimates
-of $\lambda$ and $\gamma$ in the Supplementary Material. For each of the four
-data sets, the distribution of the optimal values of $\lambda$ selected through
-cross-validation often had a notable bathtub shape, resembling a Beta
-distribution with both shape parameters less than unity. Specifically, the
-optimal values of $\lambda$ were typically near 0 or near 1, corresponding to
-quadratic and linear discriminant analysis, respectively. For instance, with
-smaller values of $d$ applied to the Shipp data set, the range of optimal values
-of $\lambda$ was typically no larger than 0.05, indicating there were a large
-number of training data sets for which we can safely assume the covariance
-matrices are unequal or nearly unequal. As $d$ increased, the majority of the
-optimal values of $\lambda$ were selected as unity, thereby an increased number
-of variables resulted in our classifier preferring the more parsimonious
-assumption of equal covariance matrices. In this instance we can see the
-flexibility of our classifier to employ the more realistic assumption that the
-class covariance matrices are unequal, but if optimal, we can employ the more
-parsimonious assumption of equal covariance matrices.
-
-\section{Discussion}
-
-In our simulation study, we compared our classifier to several recently proposed
-classifiers, including the Dudoit, Pang, Tong, and Guo classifiers, each of
-which assumes independence of the features conditional on class
-membership. Despite the rapid computational performance of diagonal classifiers
-and the reduction in the number of parameters to estimate, \cite{Mai:2012bf} and
-\cite{Fan:2012iq} have noted that diagonal classifiers can often yield inferior
-classification performance to other classification methods. We have shown that
-our proposed classifier often yields superior classification accuracy to the
-diagonal classifiers considered, confirming the assertions of \cite{Mai:2012bf}
-and \cite{Fan:2012iq}. Furthermore, the implementation of our classifier in the
-{\tt sparsediscrim} R package available on CRAN was comparable to that of the
-diagonal classifiers in terms of computational runtime. Moreover, when
-preliminary variable selection was applied to the data in our classification
-study, the time to perform model selection with our proposed classifier was
-nearly instantaneous.
-
-\bibliographystyle{plainnat}
-\bibliography{rda}
-
-\clearpage
-
-<<data_summary, results="asis">>=
-# Table containing a summary of the data sets
-data_summary <- subset(describe_data(), author %in% unique(results$data_set))
-data_summary$K[data_summary$author == "nakayama"] <- 5
-rownames(data_summary) <- NULL
-colnames(data_summary)[1:3] <- c("Author", "Year", "N")
-data_summary$Author <- Hmisc:::capitalize(as.character(data_summary$Author))
-data_summary$Author <- with(data_summary, paste0(Author, " et al. (", Year, ")"))
-data_summary <- subset(data_summary, select = -Year)
-print(xtable(data_summary, caption = "Summary of high-dimensional microarray data sets.",
-             label = "tab:data-summary", digits = 0), include.rownames = FALSE)
-
+\begin{figure}
+<<optimal_estimates_burczynski>>=
+hdrda_optimal$lambda <- factor(hdrda_optimal$lambda)
+hdrda_optimal$gamma <- factor(hdrda_optimal$gamma)
+p <- ggplot(subset(hdrda_optimal, data_set == "burczynski" & d %in% c(100, 500, 1000)), aes(lambda))
+p <- p + geom_bar(aes(fill = gamma, weight = value))
+p <- p + facet_grid(d ~ ., labeller = label_bquote(d == .(x))) + theme_bw()
+p <- p + theme(strip.text.y = element_text(size = 16))
+p <- p + theme(axis.text = element_text(size = 16))
+p <- p + theme(axis.title.x = element_text(size = 24))
+p <- p + theme(axis.title.y = element_text(size = 20))
+p <- p + scale_fill_discrete(name = expression(gamma))
+p <- p + theme(legend.title = element_text(size = 28))
+p <- p + theme(legend.key.size = unit(1.25, "cm"))
+p <- p + theme(legend.text = element_text(size = 18))
+p + xlab(expression(lambda)) + ylab("Counts")
 @ 
-
-\begin{figure}
-<<hdrda_contours, fig.width=8, fig.height=8, out.width='.49\\linewidth', fig.show='hold'>>=
-rda_contours(lambda = 0)
-rda_contours(lambda = 1/4)
-rda_contours(lambda = 3/4)
-rda_contours(lambda = 1)
-@ 
-\caption{Contours of five multivariate normal populations as a function of the
-  pooling parameter $\lambda$.}
-\label{fig:hdrda-contours}
+\caption{Counts of the optimal tuning-parameter estimates for the \emph{HDRDA}
+  classifier for the Burczynski data set for $d = 100$, $500$, and $1000$.}
+\label{fig:optimal-burczynski}
 \end{figure}
@@
 \begin{figure}
-<<error_rates>>=
-error_rates$data_set <- Hmisc:::capitalize(error_rates$data_set)
-p <- ggplot(error_rates, aes(x = d, y = Avg_Error, group = Classifier))
-p <- p + geom_line(aes(color = Classifier, linetype = Classifier), size = 1.5)
-p <- p + facet_wrap(~ data_set) + ylim(c(0, max(error_rates$Avg_Error))) + theme_bw()
-p <- p + xlab("Number of Variables Selected") + ylab("Average Error Rate")
-p  + theme(strip.text.x = element_text(size = 12))
+<<optimal_estimates_nakayama>>=
+hdrda_optimal$lambda <- factor(hdrda_optimal$lambda)
+hdrda_optimal$gamma <- factor(hdrda_optimal$gamma)
+p <- ggplot(subset(hdrda_optimal, data_set == "nakayama" & d %in% c(100, 500, 1000)), aes(lambda))
+p <- p + geom_bar(aes(fill = gamma, weight = value))
+p <- p + facet_grid(d ~ ., labeller = label_bquote(d == .(x))) + theme_bw()
+p <- p + theme(strip.text.y = element_text(size = 16))
+p <- p + theme(axis.text = element_text(size = 16))
+p <- p + theme(axis.title.x = element_text(size = 24))
+p <- p + theme(axis.title.y = element_text(size = 20))
+p <- p + scale_fill_discrete(name = expression(gamma))
+p <- p + theme(legend.title = element_text(size = 28))
+p <- p + theme(legend.key.size = unit(1.25, "cm"))
+p <- p + theme(legend.text = element_text(size = 18))
+p + xlab(expression(lambda)) + ylab("Counts")
 @ 
-\caption{Average error-rate estimates of the competing classifiers as a function
-  of $d$. Approximate standard errors are no greater than \Sexpr{max(error_rates[["SE_Error"]])}.}
-\label{fig:error-rates}
+\caption{Counts of the optimal tuning-parameter estimates for the \emph{HDRDA}
+  classifier for the Nakayama data set for $d = 100$, $500$, and $1000$.}
+\label{fig:optimal-nakayama}
 \end{figure}
@@
 \begin{figure}
-<<optimal_estimates>>=
+<<optimal_estimates_shipp>>=
 hdrda_optimal$lambda <- factor(hdrda_optimal$lambda)
@@
 p <- p + facet_grid(d ~ ., labeller = label_bquote(d == .(x))) + theme_bw()
-p + theme(strip.text.y = element_text(size = 16))
+p <- p + theme(strip.text.y = element_text(size = 16))
+p <- p + theme(axis.text = element_text(size = 16))
+p <- p + theme(axis.title.x = element_text(size = 24))
+p <- p + theme(axis.title.y = element_text(size = 20))
+p <- p + scale_fill_discrete(name = expression(gamma))
+p <- p + theme(legend.title = element_text(size = 28))
+p <- p + theme(legend.key.size = unit(1.25, "cm"))
+p <- p + theme(legend.text = element_text(size = 18))
+p + xlab(expression(lambda)) + ylab("Counts")
 @ 
@@
 
+\begin{figure}
+<<optimal_estimates_singh>>=
+hdrda_optimal$lambda <- factor(hdrda_optimal$lambda)
+hdrda_optimal$gamma <- factor(hdrda_optimal$gamma)
+p <- ggplot(subset(hdrda_optimal, data_set == "singh" & d %in% c(100, 500, 1000)), aes(lambda))
+p <- p + geom_bar(aes(fill = gamma, weight = value))
+p <- p + facet_grid(d ~ ., labeller = label_bquote(d == .(x))) + theme_bw()
+p <- p + theme(strip.text.y = element_text(size = 16))
+p <- p + theme(axis.text = element_text(size = 16))
+p <- p + theme(axis.title.x = element_text(size = 24))
+p <- p + theme(axis.title.y = element_text(size = 20))
+p <- p + scale_fill_discrete(name = expression(gamma))
+p <- p + theme(legend.title = element_text(size = 28))
+p <- p + theme(legend.key.size = unit(1.25, "cm"))
+p <- p + theme(legend.text = element_text(size = 18))
+p + xlab(expression(lambda)) + ylab("Counts")
+@ 
+\caption{Counts of the optimal tuning-parameter estimates for the \emph{HDRDA}
+  classifier for the Singh data set for $d = 100$, $500$, and $1000$.}
+\label{fig:optimal-singh}
+\end{figure}
+
 \end{document}

### 2013-04-21 reports/paper-rda.Rnw
Updated figures to abide by Biometrika rules.

@@
 
-# To abide by Biometrika's requirements of no abbreviations, we rename the
-# results of our HDRDA classifier as simply "Ours."
-results$Ours <- results$HDRDA
-results <- subset(results, select = -HDRDA)
-
 results$d <- factor(results$d, labels = sort(unique(as.numeric(results$d))))
@@
 m_error <- melt(error_rates, variable.name = "Classifier", value.name = "Error")
-m_error$Classifier <- factor(m_error$Classifier)
 error_rates <- ddply(m_error, .(data_set, d, Classifier), plyr:::summarize,
@@
 error_rates <- subset(error_rates, select = -Num_Reps)
-
-# Candidate values of tuning parameters.
-seq_lambda <- seq(0, 1, by = 0.05)
-seq_gamma <- 10^seq.int(-1, 5)
-
-# Summary of the optimal tuning-parameter estimates for each data set and number
-# of variables selected, d.
-hdrda_optimal <- subset(results, select = c(data_set, d, lambda, gamma))
-hdrda_optimal$lambda <- factor(hdrda_optimal$lambda, levels = seq_lambda)
-hdrda_optimal$gamma <- factor(hdrda_optimal$gamma, levels = seq_gamma)
-
-hdrda_optimal <- ddply(hdrda_optimal, .(data_set, d), function(x) {
-  x <- subset(x, select = c(lambda, gamma))
-  melt(table(x))
-})
-
 @
@@
 p <- ggplot(error_rates, aes(x = d, y = Avg_Error, group = Classifier))
-p <- p + geom_line(aes(color = Classifier, linetype = Classifier), size = 1.5)
+p <- p + geom_line(aes(color = Classifier), size = 1.5, linetype = 2)
+p <- p + geom_point(aes(color = Classifier, shape = Classifier), size = 4)
 p <- p + facet_wrap(~ data_set) + ylim(c(0, max(error_rates$Avg_Error))) + theme_bw()
-p <- p + xlab("Number of Variables Selected") + ylab("Average Error Rate")
-p  + theme(strip.text.x = element_text(size = 12))
+p <- p + xlab("Number of Variables Selected, d") + ylab("Average Error Rate")
+p <- p + theme(axis.text = element_text(size = 16))
+p <- p + theme(axis.title.x = element_text(size = 20))
+p <- p + theme(axis.title.y = element_text(size = 20))
+p <- p + theme(strip.text.x = element_text(size = 18))
+p + theme(legend.position = "none")
 @ 
-\caption{Average error-rate estimates of the competing classifiers as a function
-  of $d$. Approximate standard errors are no greater than \Sexpr{max(error_rates[["SE_Error"]])}.}
+\caption{Averages of the 1000 error-rate estimates of the competing classifiers
+  for each data set as a function of $d$. The classifiers are marked as follows:
+  Dudoit (red circle), Guo (gold triangle), Pang (green square), Tong = (cyan
+  cross), Witten (blue square with X), and ours (pink asterisk). Approximate
+  standard errors are no greater than \Sexpr{max(error_rates[["SE_Error"]])}.}
 \label{fig:error-rates}

### 2013-04-21 lib/helpers.R

@@ -120,12 +120,12 @@ rda_contours <- function(lambda) {
   contours$Population <- factor(contours$Population)
   contours$lambda <- lambda
 
-  p <- ggplot(contours, aes(x = X1, y = X2, colour = Population))
-  p <- p + geom_path(size = 1.5, linetype = 2)
+  p <- ggplot(contours, aes(x = X1, y = X2, colour = Population, linetype = Population))
+  p <- p + geom_path(size = 1.5)
   p <- p + facet_grid(. ~ lambda, labeller = label_bquote(lambda == .(x)))
   p <- p + theme_bw() + xlab("") + ylab("") + theme(legend.position = "none")
   p <- p + scale_x_continuous(breaks = NULL) + scale_y_continuous(breaks = NULL)
-  p + theme(strip.text.x = element_text(size = 18))
+  p + theme(strip.text.x = element_text(size = 28))
 }
 
 # Classifier from Witten and Tibshirani (2011) - JRSS B

### 2013-04-21 reports/paper-rda.Rnw
Addressed TODO items in footnotes.

@@
 selected via cross-validation, which adds a substantial computational burden to
-the model-selection process. \citeauthor{Friedman:1989tm}'s classifier is
-ill-advised on modern data sets because the model selection requires $O(Kp^3)$
-calculations for each tuning-parameter pair.
+the model-selection process.
 
@@
 
-\subsection{Improvements to \citeauthor{Friedman:1989tm}'s Classifier}
+\subsection{Model Specification}
 
@@
 $\tilde{\Sigma}_k^{-1}$ can be substituted for $\tilde{\Sigma}_k^{+}$ in
-\eqref{eq:hdrda}. If $\gamma = 0$, we explicitly set $|\tilde{\Sigma}_k|$
-equal to the product of the positive eigenvalues of $\tilde{\Sigma}_k$.
+\eqref{eq:hdrda}. If $\gamma = 0$, we explicitly set $|\tilde{\Sigma}_k|$ equal
+to the product of the positive eigenvalues of $\tilde{\Sigma}_k$. Following
+\cite{Friedman:1989tm}, we select $\lambda$ and $\gamma$ from a grid of
+candidate models via cross-validation \citep{Hastie:2008dt}. We provide an
+implementation of \eqref{eq:hdrda} in the {\tt hdrda} function in the {\tt
+  sparsediscrim} R package, available on CRAN.
 
@@
 Next, we demonstrate that our proposed classifier in \eqref{eq:hdrda} lends
-itself to a more efficient calculation and a rapid model selection of the
-parameters $\lambda$ and $\gamma$\footnote{TODO: Briefly mention the updating
-  trick employed in the R package}. We decompose \eqref{eq:hdrda} into a sum of
-two components, where the first summand consists of matrix operations applied to
-low-dimensional matrices and the second summand corresponds to the null space of
-$\widehat{\Sigma}$ in \eqref{eq:pooled-cov}. We show that the matrix
+itself to a more efficient calculation. We decompose \eqref{eq:hdrda} into a sum
+of two components, where the first summand consists of matrix operations applied
+to low-dimensional matrices and the second summand corresponds to the null space
+of $\widehat{\Sigma}$ in \eqref{eq:pooled-cov}. We show that the matrix
 operations performed on the null space of $\widehat{\Sigma}$ yield constant
@@
 computational costs is achieved relative to Friedman's classifier in
-\eqref{eq:rda}\footnote{TODO: Add a brief comparison of the computations of
-  \emph{RDA} and \emph{HDRDA} in terms of Big Oh notation}. Furthermore, a
-byproduct is that adjustments to the associated eigenvalues have no effect on
-\eqref{eq:hdrda}. Lastly, we utilize the singular value decomposition to
-calculate efficiently the eigenvalue decomposition of $\widehat{\Sigma}$,
-further reducing the computational costs of our proposed classifier.
+\eqref{eq:rda}. Furthermore, a byproduct is that adjustments to the associated
+eigenvalues have no effect on \eqref{eq:hdrda}. Lastly, we utilize the singular
+value decomposition to calculate efficiently the eigenvalue decomposition of
+$\widehat{\Sigma}$, further reducing the computational costs of our proposed
+classifier.
 
@@
 
-\subsection{Model Selection}
-
 Thus far, we have shown that our classifier in \eqref{eq:hdrda} is invariant to
@@
 \eqref{eq:hdrda-decomposed} with a substantial reduction in the computational
-complexity. Here, we provide two\footnote{TODO: Two? Perhaps, I mean one and
-  then the Fast SVD below. If so, clarify this.} additional results that further
-improve the computational performance of our classifier. First, we demonstrate
-that the inverse and determinants of $W_k$ can be written so that they are
-performed on $n_k \times n_k$ matrices rather than the larger $q \times q$
-matrices, resulting in further improvements to the computational efficiency.
+complexity. Here, we demonstrate that the computational efficiency in
+calculating the inverse and determinants of $W_k$ can be further improved via
+standard matrix operations.
 
-\begin{proposition}\label{proposition:rda-W_k}
+\begin{proposition}\label{proposition:hdrda-W_k}
 Let $W_k$ be defined as above $(k = 1, \ldots, K)$. Then, $|W_k| = |
@@
 	\end{align}
-where $Q_k = I_{n_k} + \alpha_k(1- \lambda) X_k U_1 
+where $Q_k = I_{n_k} + \alpha_k(1 - \lambda) X_k U_1 
 \Gamma_k^{-1} U_1^{T} X_k^{T}$ and $\Gamma_k = \alpha_k \lambda D_q +
@@
 
-Thus far, we have constructed each $\tilde{\Sigma}_k$ in terms of the first
-$q$ eigenvectors of $\widehat{\Sigma}$. By utilizing a technique similar to
-the so-called kernel trick, we can obtain $U_1$ by computing the
-eigendecomposition of a much smaller $N \times N$ matrix when $p \gg N$
+Conveniently, we see that the inverse and determinant of $W_k$ are each in terms
+of the inverse and determinant of $Q_k \in \mathbb{R}_{n_k \times n_k}$,
+respectively, as well as the diagonal matrix $\Gamma_k \in \mathbb{R}_{q \times
+  q}$. This formulation is especially advantageous when the number of classes
+$K$ is large because $n_k$ is then potentially much smaller than $q$. The
+expressions given in Proposition \ref{proposition:hdrda-W_k} also expedite the
+selection of $\lambda$ and $\gamma$ via cross-validation. The most
+time-consuming matrix operations involved in computing $W_k^{-1}$ and $|W_k|$ is
+$X_k U_1 \in \mathbb{R}_{n_k \times q}$, which is independent of $\lambda$ and
+$\gamma$. Hence, we need calculate $X_k U_1$ only once for each cross-validation
+fold. The subsequent operations in calculating $W_k^{-1}$ and $|W_K|$ can be
+simply updated for different pairs of $\lambda$ and $\gamma$.
+
+Thus far, we have constructed each $\tilde{\Sigma}_k$ in terms of the first $q$
+eigenvectors of $\widehat{\Sigma}$. By utilizing a technique similar to the
+so-called kernel trick, we can obtain $U_1$ by computing the eigendecomposition
+of a much smaller $N \times N$ matrix when $p \gg N$
 \citep{Hastie:2008dt}. Recall that the pooled sample covariance matrix
-$\widehat{\Sigma} = X^{T} X$. Then, because $X = M D^{1/2}
-U^{T}$ by the singular value decomposition, we have the eigendecomposition
-$\widehat{\Sigma} = U D U^{T}$, where $M \in \mathbb{R}_{N
-  \times N}$ is orthogonal, $D^{1/2} \in \mathbb{R}_{N \times N}^{\ge}$ is a
-diagonal matrix consisting of the singular values of $X$, and $U \in
-\mathbb{R}_{p \times N}$ is orthogonal. Now, $X X^{T} = M D M^{T}$
-is the eigenvalue decomposition of $X X^{T}$, where $D$ is the matrix
-of eigenvalues of $X X^{T}$ and the columns of $M$ are the
-corresponding eigenvectors. Hence, we obtain $M$ and $D$ from an
-eigendecomposition of $X X^{T} \in \mathbb{R}_{N \times N}$. Next, we
-compute $U = X^{T} M D^{+/2}$, where $D^{+/2} = D_q^{-1/2}
-\oplus 0_{N-q}$. We then determine $q$ by computing the number of
-numerically nonzero eigenvalues present in $D$, where $q$ is the number of
-eigenvalues that exceeds some tolerance value, say, $\epsilon = 1 \times
-10^{-6}$. After calculating $q$, we extract $U_1$ by retaining only the
+$\widehat{\Sigma} = X^{T} X$. Then, because $X = M D^{1/2} U^{T}$ by the
+singular value decomposition, we have the eigendecomposition $\widehat{\Sigma} =
+U D U^{T}$, where $M \in \mathbb{R}_{N \times N}$ is orthogonal, $D^{1/2} \in
+\mathbb{R}_{N \times N}^{\ge}$ is a diagonal matrix consisting of the singular
+values of $X$, and $U \in \mathbb{R}_{p \times N}$ is orthogonal. Now, $X X^{T}
+= M D M^{T}$ is the eigenvalue decomposition of $X X^{T}$, where $D$ is the
+matrix of eigenvalues of $X X^{T}$ and the columns of $M$ are the corresponding
+eigenvectors. Hence, we obtain $M$ and $D$ from an eigendecomposition of $X
+X^{T} \in \mathbb{R}_{N \times N}$. Next, we compute $U = X^{T} M D^{+/2}$,
+where $D^{+/2} = D_q^{-1/2} \oplus 0_{N-q}$. We then determine $q$ by computing
+the number of numerically nonzero eigenvalues present in $D$, where $q$ is the
+number of eigenvalues that exceeds some tolerance value, say, $\epsilon = 1
+\times 10^{-6}$. After calculating $q$, we extract $U_1$ by retaining only the
 first $q$ columns of $U$.
@@
 compare our proposed classifier with five classifiers recently proposed for
-small-sample, high-dimensional data. Our competitors include diagonal linear
+small-sample, high-dimensional data. In Table \ref{tab:data-summary}, we
+summarize the four microarray data sets. Our competitors include diagonal linear
 discriminant analysis from \cite{Dudoit:2002ev}, shrunken centroids regularized
@@
 include \citeauthor{Friedman:1989tm}'s classifier in our studies to avoid its
-excessive computations. In Table \ref{tab:data-summary}, we summarize the
-attributes of each microarray data set.
+excessive computations.
 
 We used version 3.0.0 of the open-source statistical software {\tt R} for the
-classification study. For each classifier's {\tt R} implementation, we used the
-default settings with the exception that we explicitly set prior probabilities
-as equal. For our classifier, we set $\alpha_k = 1$ $(k = 1, \ldots, K)$, so
-that the covariance-matrix estimator \eqref{eq:hdrda-cov} employed had a form
-resembling \eqref{eq:ridge-estimator}. We estimated $\lambda$ from a grid of 21
-equidistant candidate values between 0 and 1, inclusively. Similarly, we
+classification study. We used the implementations of the Witten and Guo
+classifiers from the {\tt penalizedLDA} and {\tt rda} packages, respectively, on
+CRAN. Because the {\tt rda} package does not perform the authors' ``Min-Min''
+rule automatically, we applied this rule in our {\tt R} code. We also utilized
+our implementations of the Dudoit, Pang, and Tong classifiers available in the
+{\tt sparsediscrim} package. We used the default settings for each classifier
+with the exception that we explicitly set prior probabilities as equal, if the
+option was available. For our classifier, we set $\alpha_k = 1$ $(k = 1, \ldots,
+K)$, so that the covariance-matrix estimator \eqref{eq:hdrda-cov} employed had a
+form resembling \eqref{eq:ridge-estimator}. We estimated $\lambda$ from a grid
+of 21 equidistant candidate values between 0 and 1, inclusively. Similarly, we
 estimated $\gamma$ from a grid consisting of the values $10^{-1}, \ldots, 10^4$,
@@
 and \cite{Fan:2012iq}. Furthermore, the implementation of our classifier in the
-{\tt sparsediscrim} R package available on CRAN was comparable to that of the
-diagonal classifiers in terms of computational runtime. Moreover, when
-preliminary variable selection was applied to the data in our classification
-study, the time to perform model selection with our proposed classifier was
-nearly instantaneous.
+{\tt sparsediscrim} R package was comparable to that of the diagonal classifiers
+in terms of computational runtime. Moreover, when preliminary variable selection
+was applied to the data in our classification study, the time to perform model
+selection with our proposed classifier was nearly instantaneous.
 

### 2013-04-22 reports/paper-rda.Rnw
Updated LaTeX documentstyle to match Biometrika's.

@@
-\documentclass[11pt]{article}
-\usepackage{graphicx, amsmath, amssymb, bm, url, mathtools, natbib, amsthm, setspace}
+\documentclass[lineno]{biometrika}
 
-\pagestyle{plain}
-%----------------Page dimensions ----------------
-\oddsidemargin 0.0in
-\evensidemargin 0.0in
-\topmargin -0.75in
-\leftmargin 0in
-\headheight 0.0in
-\headsep 0.5in
-%\footheight 0.0in
-\footskip 0.5in
-\footnotesep 0.0in
-\textwidth 6.7in
-\textheight 9.5in
-%-----------Define Pictures---------------------
-\def\picture #1 by #2 (#3){
- \vbox to #2{
-   \hrule width #1 height 0pt depth 0pt
-   \vfill
-   \special{picture #3} % this is the low-level interface
-   }
- }
-\def\scaledpicture #1 by #2 (#3 scaled #4){{
- \dimen0=#1 \dimen1=#2
- \divide\dimen0 by 1000 \multiply\dimen0 by #4
- \divide\dimen1 by 1000 \multiply\dimen1 by #4
- \picture \dimen0 by \dimen1 (#3 scaled #4)}
- }
-
+\usepackage{graphicx, amsmath, amssymb, bm, natbib}
 \newcommand{\xbar}{\bar{x}}
@@
 \DeclareMathOperator*{\argmin}{arg\,min}
-\DeclareMathOperator*{\argmax}{arg\,max}
 
+\newtheorem{cor}{Corollary}
+%\newtheorem{lemma}{Lemma}
+%\newtheorem{proposition}{Proposition}
+\newtheorem{thm}{Theorem}
+
+\bibpunct{(}{)}{;}{a}{,}{,}
+
+%% Please use the following statements for
+%% managing the text and math fonts for your papers:
+\usepackage{times}
+%\usepackage[cmbold]{mathtime}
+
+\usepackage[plain,noend]{algorithm2e}
+
+\makeatletter
+\renewcommand{\algocf@captiontext}[2]{#1\algocf@typo. \AlCapFnt{}#2} % text of caption
+\renewcommand{\AlTitleFnt}[1]{#1\unskip}% default definition
+\def\@algocf@capt@plain{top}
+\renewcommand{\algocf@makecaption}[2]{%
+  \addtolength{\hsize}{\algomargin}%
+  \sbox\@tempboxa{\algocf@captiontext{#1}{#2}}%
+  \ifdim\wd\@tempboxa >\hsize%     % if caption is longer than a line
+    \hskip .5\algomargin%
+    \parbox[t]{\hsize}{\algocf@captiontext{#1}{#2}}% then caption is not centered
+  \else%
+    \global\@minipagefalse%
+    \hbox to\hsize{\box\@tempboxa}% else caption is centered
+  \fi%
+  \addtolength{\hsize}{-\algomargin}%
+}
+\makeatother
+
+%%% User-defined macros should be placed here, but keep them to a minimum.
+\def\Bka{{\it Biometrika}}
+\def\AIC{\textsc{aic}}
+\def\T{{ \mathrm{\scriptscriptstyle T} }}
+\def\v{{\varepsilon}}
+
+\begin{document}
+
+\jname{Biometrika}
+%% The year, volume, and number are determined on publication
+\jyear{2013}
+\jvol{99}
+\jnum{1}
+%% The \doi{...} and \accessdate commands are used by the production team
+%\doi{10.1093/biomet/asm023}
+\accessdate{Advance Access publication on 31 July 2012}
+\copyrightinfo{\Copyright\ 2013 Biometrika Trust\goodbreak {\em Printed in Great Britain}}
+
+%% These dates are usually set by the production team
+\received{April 2013}
+\revised{April 2013}
+
+%% The left and right page headers are defined here:
+\markboth{J. A. Ramey, C. K. Stein \and D. M. Young}{High-Dimensional
+  Regularized Discriminant Analysis}
+
+%% Here are the title, author names and addresses
 \title{Another Look at Regularized Discriminant Analysis for High-Dimensional Classification}
 
-\author{John A. Ramey, Caleb K. Stein, and Dean M. Young}
+% \author{John A. Ramey, Caleb K. Stein, and Dean M. Young}
+\author{J. A. RAMEY}
+\affil{Vaccine and Infectious Disease Division, Fred Hutchinson Cancer Research
+  Center,\\ Seattle, Washington 98115, U.S.A. \email{johnramey@gmail.com}}
+
+\author{C. K. STEIN \and D. M. YOUNG}
+\affil{Department of Statistical Science, Baylor University, Waco, Texas 76706, U.S.A. \email{Caleb\_Stein@baylor.edu} \email{Dean\_Young@baylor.edu}}
 
@@
 error_rates <- subset(error_rates, select = -Num_Reps)
+
+max_se <- max(error_rates[["SE_Error"]])
 @
-
-\begin{document}
-
-\newtheorem{cor}{Corollary}
-\newtheorem{lemma}{Lemma}
-\newtheorem{proposition}{Proposition}
-\newtheorem{thm}{Theorem}
-
-\bibpunct{(}{)}{;}{a}{,}{,}
-
-\doublespacing
 
@@
 for $\lambda = 1$.
-
 \[ \left[\text{Figure \ref{fig:hdrda-contours} goes about here }\right]\]
-
 As we have discussed above, several eigenvalue adjustment methods have been
@@
 compare our proposed classifier with five classifiers recently proposed for
-small-sample, high-dimensional data. In Table \ref{tab:data-summary}, we
-summarize the four microarray data sets. Our competitors include diagonal linear
+small-sample, high-dimensional data. Our competitors include diagonal linear
 discriminant analysis from \cite{Dudoit:2002ev}, shrunken centroids regularized
@@
 10-fold cross-validation \citep{Hastie:2008dt}.
-
-\[ \left[\text{Table 1 goes about here }\right]\]
 
@@
 each value of $d$ in Figure \ref{fig:error-rates}. Approximate standard errors
-were no greater than \Sexpr{max(error_rates[["SE_Error"]])}.
+were no greater than $\Sexpr{gsub('.', '\\cdot', round(max_se, 3), fixed = TRUE)}$.
 
@@
 required a larger subset of the features to perform competitively in our study.
-
 \[ \left[\text{Figure \ref{fig:error-rates} goes about here }\right]\]
-
 Next, we examine the optimal estimates of the tuning parameters for our
@@
 
-\bibliographystyle{plainnat}
+%\bibliographystyle{plainnat}
+\bibliographystyle{biometrika}
 \bibliography{rda}
@@
 \clearpage
-
-<<data_summary, results="asis">>=
-# Table containing a summary of the data sets
-data_summary <- subset(describe_data(), author %in% unique(results$data_set))
-data_summary$K[data_summary$author == "nakayama"] <- 5
-rownames(data_summary) <- NULL
-colnames(data_summary)[1:3] <- c("Author", "Year", "N")
-data_summary$Author <- Hmisc:::capitalize(as.character(data_summary$Author))
-data_summary$Author <- with(data_summary, paste0(Author, " et al. (", Year, ")"))
-data_summary <- subset(data_summary, select = -Year)
-print(xtable(data_summary, caption = "Summary of high-dimensional microarray data sets.",
-             label = "tab:data-summary", digits = 0), include.rownames = FALSE)
-
-@
 
@@
   cross), Witten (blue square with X), and ours (pink asterisk). Approximate
-  standard errors are no greater than \Sexpr{max(error_rates[["SE_Error"]])}.}
+  standard errors are no greater than $\Sexpr{gsub('.', '\\cdot', round(max_se, 3), fixed = TRUE)}$.}
 \label{fig:error-rates}

### 2013-04-23 reports/paper-rda.Rnw
Added a section entitled Supplementary Material
Updated paper's title at DMY's recommendation. Also, manually added biometrika bibstyle file.

@@
 %% Here are the title, author names and addresses
-\title{Another Look at Regularized Discriminant Analysis for High-Dimensional Classification}
+\title{High-Dimensional Regularized Discriminant Analysis}
 
@@
 From \eqref{eq:sig-lambda-alternative2}, $\lambda$ weights the contribution of
-each of the $N$ observations to estimate $\Sigma_k$. Hence, for $0 < \lambda
+each of the $N$ observations in estimating $\Sigma_k$. Hence, for $0 < \lambda
 \le 1$, we estimate $\Sigma_k$ with $N$ rather than $n_k$
@@
 
-%\bibliographystyle{plainnat}
+\section*{Supplementary material}
+\label{SM}
+Supplementary material available at \Bka\ online includes the graphical summary
+of the optimal estimates of our classifier from the classification study.
+
 \bibliographystyle{biometrika}

### 2013-04-26 reports/paper-rda.Rnw
Applied final edits from DMY. Also, changed figures to black and white.

@@
 \usepackage{graphicx, amsmath, amssymb, bm, natbib}
-\newcommand{\xbar}{\bar{x}}
-\newcommand{\tr}{\text{tr}}
-\DeclareMathOperator*{\argmin}{arg\,min}
 
 \newtheorem{cor}{Corollary}
-%\newtheorem{lemma}{Lemma}
-%\newtheorem{proposition}{Proposition}
 \newtheorem{thm}{Theorem}
@@
 \def\v{{\varepsilon}}
+
+\newcommand{\xbar}{\bar{x}}
+\newcommand{\tr}{\text{tr}}
+\DeclareMathOperator*{\argmin}{arg\,min}
 
@@
 the computational shortcomings of \citeauthor{Friedman:1989tm}'s classifier for
-high-dimensional data. Moreover, we demonstrate with several high-dimensional
-microarray data sets that our proposed classifier is competitive and often
+high-dimensional data. Moreover, we demonstrate with four high-dimensional
+microarray data sets that our proposed classifier is competitive with and often
 superior to several recently proposed sparse and regularized classifiers in
@@
 operation of \citeauthor{Friedman:1989tm}'s classifier. Given that the
-assumption of equal population covariance matrices within linear discriminant
+assumption of equal population covariance matrices for linear discriminant
 analysis is often too stringent in practice \citep{Clemmensen:2011kr}, our
@@
 high-dimensional data. In particular, we compare our classifier with diagonal
-linear discriminant analysis from \cite{Dudoit:2002ev} and the recently proposed
-classifiers from \cite{Guo:2007te}, \cite{Pang:2009ik}, \cite{Tong:2012hw}, and
-\cite{Witten:2011kc}. We conduct our comparison of the classification accuracy
-of the competing classifiers using four small-sample, high-dimensional
-benchmark microarray data sets.
+linear discriminant analysis proposed by \cite{Dudoit:2002ev} and the recently
+proposed classifiers from \cite{Guo:2007te}, \cite{Pang:2009ik},
+\cite{Tong:2012hw}, and \cite{Witten:2011kc}. We conduct our comparison of the
+classification accuracy of the competing classifiers using four small-sample,
+high-dimensional benchmark microarray data sets.
 
@@
 Throughout the document we also require the following additional notation. Let
-$I_m$ be the $m \times m$ identity matrix, and let $0_{m \times p}$ be
-the $m \times p$ matrix of zeros, such that $0_m$ is understood to denote
-$0_{m \times m}$. Lastly, for $c \in \mathbb{R}$, let $c^+ = 1/c$ if $c \ne
-0$ and $0$ otherwise. Let $A \oplus B$ denote the direct sum of $A \in
-\mathbb{R}_{r \times r}$ and $B \in \mathbb{R}_{s \times s}$ \citep[Chapter
+$I_m$ be the $m \times m$ identity matrix, and let $0_{m \times p}$ be the $m
+\times p$ matrix of zeros, such that $0_m$ is understood to denote $0_{m \times
+  m}$. For $c \in \mathbb{R}$, let $c^+ = 1/c$ if $c \ne 0$ and $0$
+otherwise. Let $A \oplus B$ denote the direct sum of $A \in \mathbb{R}_{r \times
+  r}$ and $B \in \mathbb{R}_{s \times s}$ \citep[Chapter
   1]{Lutkepohl:1996uz}. Let $A^+$ and $\mathcal{N}(A) = \{z \in \mathbb{R}_{p
-  \times 1} : A z = 0_{m \times 1}\}$ denote the Moore-Penrose pseudoinverse and null space
-of $A \in \mathbb{R}_{m \times p}$, respectively. Let $V^{\perp} = \{y
-\in \mathbb{R}_{p \times 1} : y^{T}v = 0\ \forall v \in V \}$ denote
-the orthogonal complement of a vector space $V \subset \mathbb{R}_{p \times 1}$.
+  \times 1} : A z = 0_{m \times 1}\}$ denote the Moore-Penrose pseudoinverse and
+null space of $A \in \mathbb{R}_{m \times p}$, respectively. Let $V^{\perp} =
+\{y \in \mathbb{R}_{p \times 1} : y^{T}v = 0\ \forall v \in V \}$ denote the
+orthogonal complement of a vector space $V \subset \mathbb{R}_{p \times 1}$.
 
@@
 for $\lambda = 1$.
-\[ \left[\text{Figure \ref{fig:hdrda-contours} goes about here }\right]\]
+\begin{figure}
+<<hdrda_contours, fig.width=8, fig.height=8, out.width='.49\\linewidth', fig.show='hold'>>=
+rda_contours(lambda = 0)
+rda_contours(lambda = 1/4)
+rda_contours(lambda = 3/4)
+rda_contours(lambda = 1)
+@ 
+\caption{Contours of five multivariate normal populations as a function of the
+  pooling parameter $\lambda$.}
+\label{fig:hdrda-contours}
+\end{figure}
 As we have discussed above, several eigenvalue adjustment methods have been
@@
 \widehat{\Sigma}_k U + \lambda D\} + \gamma I_p$, and
-\eqref{eq:rda-matrix} follows because $U$ is orthogonal.
+\eqref{eq:rda-matrix} holds because $U$ is orthogonal.
 \end{proof}
@@
 of a much smaller $N \times N$ matrix when $p \gg N$
-\citep{Hastie:2008dt}. Recall that the pooled sample covariance matrix
+\citep{Hastie:2008dt}. Recall that the pooled sample covariance matrix is
 $\widehat{\Sigma} = X^{T} X$. Then, because $X = M D^{1/2} U^{T}$ by the
@@
 X^{T} \in \mathbb{R}_{N \times N}$. Next, we compute $U = X^{T} M D^{+/2}$,
-where $D^{+/2} = D_q^{-1/2} \oplus 0_{N-q}$. We then determine $q$ by computing
-the number of numerically nonzero eigenvalues present in $D$, where $q$ is the
-number of eigenvalues that exceeds some tolerance value, say, $\epsilon = 1
-\times 10^{-6}$. After calculating $q$, we extract $U_1$ by retaining only the
-first $q$ columns of $U$.
+where $D^{+/2} = D_q^{-1/2} \oplus 0_{N-q}$. We then determine $q$, the number
+of numerically nonzero eigenvalues present in $D$, by calculating the number of
+eigenvalues that exceeds some tolerance value, say, $\epsilon = 1 \times
+10^{-6}$. After calculating $q$, we extract $U_1$ by retaining only the first
+$q$ columns of $U$.
 
-\section{High-Dimensional Microarray Data Sets}
+\section{Classification Study}
 
@@
 option was available. For our classifier, we set $\alpha_k = 1$ $(k = 1, \ldots,
-K)$, so that the covariance-matrix estimator \eqref{eq:hdrda-cov} employed had a
-form resembling \eqref{eq:ridge-estimator}. We estimated $\lambda$ from a grid
-of 21 equidistant candidate values between 0 and 1, inclusively. Similarly, we
+K)$, so that the covariance-matrix estimator \eqref{eq:hdrda-cov} employed
+resembled \eqref{eq:ridge-estimator}. We estimated $\lambda$ from a grid of 21
+equidistant candidate values between 0 and 1, inclusively. Similarly, we
 estimated $\gamma$ from a grid consisting of the values $10^{-1}, \ldots, 10^4$,
@@
 contrast to the other classifiers, our classifier exhibited less variability in
-its misclassification estimates for small values of $d$. Also, our classifier
-yielded superior classification accuracy as $d$ increased and was hardly
-effected by the inclusion of more variables in the data sets. We attribute the
-apparent stability of the error-rate estimates of our classifier for large
-values of $d$ to its novel dimension-reduction technique.
+its misclassification estimates for small values of $d$ and yielded superior
+classification accuracy as $d$ increased and was hardly effected by the
+inclusion of more variables in the data sets. We attribute the apparent
+stability of the error-rate estimates of our classifier for large values of $d$
+to its novel dimension-reduction technique.
 
@@
 features before employing these methods, lest their classification performance
-can degrade.
+be degraded.
 
@@
 than its competitors. Furthermore, the Guo classifier exhibited unreliable
-classification performance for smaller values of $d$. The Guo classifier
-required a larger subset of the features to perform competitively in our study.
-\[ \left[\text{Figure \ref{fig:error-rates} goes about here }\right]\]
+classification performance for smaller values of $d$ and required a larger
+subset of the features to perform competitively in our study.
+\begin{figure}
+<<error_rates>>=
+error_rates$data_set <- Hmisc:::capitalize(error_rates$data_set)
+p <- ggplot(error_rates, aes(x = d, y = Avg_Error, group = Classifier))
+p <- p + geom_line(aes(), size = 1, linetype = 2)
+p <- p + geom_point(aes(shape = Classifier), size = 6)
+p <- p + facet_wrap(~ data_set) + ylim(c(0, max(error_rates$Avg_Error))) + theme_bw()
+p <- p + xlab("Number of Variables Selected, d") + ylab("Average Error Rate")
+p <- p + theme(axis.text = element_text(size = 16))
+p <- p + theme(axis.title.x = element_text(size = 20))
+p <- p + theme(axis.title.y = element_text(size = 20))
+p <- p + theme(strip.text.x = element_text(size = 18))
+p + theme(legend.position = "none")
+@ 
+\caption{Averages of the 1000 error-rate estimates of the competing classifiers
+  for each data set as a function of $d$. The classifiers are marked as follows:
+  Dudoit (circle), Guo (triangle), Pang (square), Tong = (cross), Witten (
+  square with X), and ours (asterisk). Approximate standard errors are no
+  greater than $\Sexpr{gsub('.', '\\cdot', round(max_se, 3), fixed = TRUE)}$.}
+\label{fig:error-rates}
+\end{figure}
 Next, we examine the optimal estimates of the tuning parameters for our
@@
 
-\clearpage
-
-\begin{figure}
-<<hdrda_contours, fig.width=8, fig.height=8, out.width='.49\\linewidth', fig.show='hold'>>=
-rda_contours(lambda = 0)
-rda_contours(lambda = 1/4)
-rda_contours(lambda = 3/4)
-rda_contours(lambda = 1)
-@ 
-\caption{Contours of five multivariate normal populations as a function of the
-  pooling parameter $\lambda$.}
-\label{fig:hdrda-contours}
-\end{figure}
-
-\begin{figure}
-<<error_rates>>=
-error_rates$data_set <- Hmisc:::capitalize(error_rates$data_set)
-p <- ggplot(error_rates, aes(x = d, y = Avg_Error, group = Classifier))
-p <- p + geom_line(aes(color = Classifier), size = 1.5, linetype = 2)
-p <- p + geom_point(aes(color = Classifier, shape = Classifier), size = 4)
-p <- p + facet_wrap(~ data_set) + ylim(c(0, max(error_rates$Avg_Error))) + theme_bw()
-p <- p + xlab("Number of Variables Selected, d") + ylab("Average Error Rate")
-p <- p + theme(axis.text = element_text(size = 16))
-p <- p + theme(axis.title.x = element_text(size = 20))
-p <- p + theme(axis.title.y = element_text(size = 20))
-p <- p + theme(strip.text.x = element_text(size = 18))
-p + theme(legend.position = "none")
-@ 
-\caption{Averages of the 1000 error-rate estimates of the competing classifiers
-  for each data set as a function of $d$. The classifiers are marked as follows:
-  Dudoit (red circle), Guo (gold triangle), Pang (green square), Tong = (cyan
-  cross), Witten (blue square with X), and ours (pink asterisk). Approximate
-  standard errors are no greater than $\Sexpr{gsub('.', '\\cdot', round(max_se, 3), fixed = TRUE)}$.}
-\label{fig:error-rates}
-\end{figure}
-
 \end{document}

### 2013-04-26 lib/helpers.R

@@ -120,7 +120,7 @@ rda_contours <- function(lambda) {
   contours$Population <- factor(contours$Population)
   contours$lambda <- lambda
 
-  p <- ggplot(contours, aes(x = X1, y = X2, colour = Population, linetype = Population))
+  p <- ggplot(contours, aes(x = X1, y = X2, linetype = Population))
   p <- p + geom_path(size = 1.5)
   p <- p + facet_grid(. ~ lambda, labeller = label_bquote(lambda == .(x)))
   p <- p + theme_bw() + xlab("") + ylab("") + theme(legend.position = "none")

### 2013-05-07 reports/paper-rda.Rnw
Last minute tweaks. Submitted
Removed Biometrika LaTeX template from paper. However, I have not yet added the bolding back.

@@
-\documentclass[lineno]{biometrika}
+\documentclass[11pt]{article}
+\usepackage{graphicx, amsmath, amssymb, bm, url, mathtools, natbib, amsthm, setspace}
 
-\usepackage{graphicx, amsmath, amssymb, bm, natbib}
+\pagestyle{plain}
+%----------------Page dimensions ----------------
+\oddsidemargin 0.0in
+\evensidemargin 0.0in
+\topmargin -0.75in
+\leftmargin 0in
+\headheight 0.0in
+\headsep 0.5in
+%\footheight 0.0in
+\footskip 0.5in
+\footnotesep 0.0in
+\textwidth 6.7in
+\textheight 9.5in
+%-----------Define Pictures---------------------
+\def\picture #1 by #2 (#3){
+   \vbox to #2{
+        \hrule width #1 height 0pt depth 0pt
+           \vfill
+              \special{picture #3} % this is the low-level interface
+                 }
+    }
+\def\scaledpicture #1 by #2 (#3 scaled #4){{
+     \dimen0=#1 \dimen1=#2
+      \divide\dimen0 by 1000 \multiply\dimen0 by #4
+       \divide\dimen1 by 1000 \multiply\dimen1 by #4
+        \picture \dimen0 by \dimen1 (#3 scaled #4)}
+   }
+
+\title{High-Dimensional Regularized Discriminant Analysis}
+
+\author{John A. Ramey, Caleb K. Stein, and Dean M. Young}
+%Vaccine and Infectious Disease Division, Fred Hutchinson Cancer Research
+%  Center,\\ Seattle, Washington 98115, U.S.A. \email{johnramey@gmail.com}}
+%\affil{Department of Statistical Science, Baylor University, Waco, Texas 76706, U.S.A. \email{Caleb\_Stein@baylor.edu} \email{Dean\_Young@baylor.edu}}
+
+\newcommand{\xbar}{\bar{\bm x}}
+\newcommand{\tr}{\text{tr}}
+\DeclareMathOperator*{\argmin}{arg\,min}
+\DeclareMathOperator*{\argmax}{arg\,max}
 
 \newtheorem{cor}{Corollary}
+\newtheorem{lemma}{Lemma}
+\newtheorem{proposition}{Proposition}
 \newtheorem{thm}{Theorem}
@@
 
-%% Please use the following statements for
-%% managing the text and math fonts for your papers:
-\usepackage{times}
-%\usepackage[cmbold]{mathtime}
-
-\usepackage[plain,noend]{algorithm2e}
-
-\makeatletter
-\renewcommand{\algocf@captiontext}[2]{#1\algocf@typo. \AlCapFnt{}#2} % text of caption
-\renewcommand{\AlTitleFnt}[1]{#1\unskip}% default definition
-\def\@algocf@capt@plain{top}
-\renewcommand{\algocf@makecaption}[2]{%
-  \addtolength{\hsize}{\algomargin}%
-  \sbox\@tempboxa{\algocf@captiontext{#1}{#2}}%
-  \ifdim\wd\@tempboxa >\hsize%     % if caption is longer than a line
-    \hskip .5\algomargin%
-    \parbox[t]{\hsize}{\algocf@captiontext{#1}{#2}}% then caption is not centered
-  \else%
-    \global\@minipagefalse%
-    \hbox to\hsize{\box\@tempboxa}% else caption is centered
-  \fi%
-  \addtolength{\hsize}{-\algomargin}%
-}
-\makeatother
-
-%%% User-defined macros should be placed here, but keep them to a minimum.
-\def\Bka{{\it Biometrika}}
-\def\AIC{\textsc{aic}}
-\def\T{{ \mathrm{\scriptscriptstyle T} }}
-\def\v{{\varepsilon}}
-
-\newcommand{\xbar}{\bar{x}}
-\newcommand{\tr}{\text{tr}}
-\DeclareMathOperator*{\argmin}{arg\,min}
-
 \begin{document}
 
-\jname{Biometrika}
-%% The year, volume, and number are determined on publication
-\jyear{2013}
-\jvol{99}
-\jnum{1}
-%% The \doi{...} and \accessdate commands are used by the production team
-%\doi{10.1093/biomet/asm023}
-\accessdate{Advance Access publication on 31 July 2012}
-\copyrightinfo{\Copyright\ 2013 Biometrika Trust\goodbreak {\em Printed in Great Britain}}
+\doublespacing
 
-%% These dates are usually set by the production team
-\received{April 2013}
-\revised{April 2013}
-
-%% The left and right page headers are defined here:
-\markboth{J. A. Ramey, C. K. Stein \and D. M. Young}{High-Dimensional
-  Regularized Discriminant Analysis}
-
-%% Here are the title, author names and addresses
-\title{High-Dimensional Regularized Discriminant Analysis}
-
-% \author{John A. Ramey, Caleb K. Stein, and Dean M. Young}
-\author{J. A. RAMEY}
-\affil{Vaccine and Infectious Disease Division, Fred Hutchinson Cancer Research
-  Center,\\ Seattle, Washington 98115, U.S.A. \email{johnramey@gmail.com}}
-
-\author{C. K. STEIN \and D. M. YOUNG}
-\affil{Department of Statistical Science, Baylor University, Waco, Texas 76706, U.S.A. \email{Caleb\_Stein@baylor.edu} \email{Dean\_Young@baylor.edu}}
+\maketitle
 
@@
   for each data set as a function of $d$. The classifiers are marked as follows:
-  Dudoit (circle), Guo (triangle), Pang (square), Tong = (cross), Witten (
-  square with X), and ours (asterisk). Approximate standard errors are no
-  greater than $\Sexpr{gsub('.', '\\cdot', round(max_se, 3), fixed = TRUE)}$.}
+  Dudoit (circle), Guo (triangle), Pang (square), Tong = (cross), Witten (square
+  with X), and ours (asterisk). Approximate standard errors are no greater than
+  $\Sexpr{gsub('.', '\\cdot', round(max_se, 3), fixed = TRUE)}$.}
 \label{fig:error-rates}
@@
 \label{SM}
-Supplementary material available at \Bka\ online includes the graphical summary
+Supplementary material available online includes the graphical summary
 of the optimal estimates of our classifier from the classification study.
 
-\bibliographystyle{biometrika}
+\bibliographystyle{plainnat}
 \bibliography{rda}

### 2013-05-10 reports/paper-rda.Rnw
Added bolding back to vectors and matrices.

@@
 In discriminant analysis we wish to assign correctly an unlabeled
-$p$-dimensional observation vector $x$ to one of $K$ unique, known classes
+$p$-dimensional observation vector $\bm x$ to one of $K$ unique, known classes
 (or populations) by constructing a classifier from $N$ training
-observations. Let $x_i = (x_{i1}, \ldots, x_{ip}) \in \mathbb{R}_{p \times
+observations. Let $\bm x_i = (x_{i1}, \ldots, x_{ip}) \in \mathbb{R}_{p \times
   1}$ be the $i$th observation $(i = 1, \ldots, N)$ with true, unique membership
@@
 denotes the matrix space of all $a \times b$ matrices over the real field
-$\mathbb{R}$. We assume that $(x_i, y_i)$ is a realization from a mixture
-distribution $p(x) = \sum_{k=1}^K p(x | \omega_k) p(\omega_k)$, where
-$p(x | \omega_k)$ is the probability density function of the $k$th class and
+$\mathbb{R}$. We assume that $(\bm x_i, y_i)$ is a realization from a mixture
+distribution $p(\bm x) = \sum_{k=1}^K p(\bm x | \omega_k) p(\omega_k)$, where
+$p(\bm x | \omega_k)$ is the probability density function of the $k$th class and
 $p(\omega_k)$ is the prior probability of class membership of the $k$th
@@
 Quadratic discriminant analysis is the optimal Bayesian decision rule with
-respect to a $0-1$ loss function when $p(x | \omega_k)$ is the probability
+respect to a $0-1$ loss function when $p(\bm x | \omega_k)$ is the probability
 density function of the multivariate normal distribution with known mean vector
-$\mu_k \in \mathbb{R}_{p \times 1}$ and known covariance matrix $\Sigma_k
-\in \mathbb{R}_{p \times p}^{>}$, $k = 1, 2, \ldots, K$, where $\mathbb{R}_{p
-  \times p}^{>}$ denotes the cone of real $p \times p$ positive-definite
-matrices. Because $\mu_k$ and $\Sigma_k$ are typically unknown, we assign
-an unlabeled observation $x$ to class $\omega_k$ with the sample quadratic
-discriminant analysis classifier
+$\bm \mu_k \in \mathbb{R}_{p \times 1}$ and known covariance matrix $\bm
+\Sigma_k \in \mathbb{R}_{p \times p}^{>}$, $k = 1, 2, \ldots, K$, where
+$\mathbb{R}_{p \times p}^{>}$ denotes the cone of real $p \times p$
+positive-definite matrices. Because $\bm \mu_k$ and $\bm \Sigma_k$ are typically
+unknown, we assign an unlabeled observation $\bm x$ to class $\omega_k$ with the
+sample quadratic discriminant analysis classifier
 \begin{align}
-	D_{QDA}(x) = \argmin_{k}  (x - \xbar_k)^{T}\widehat{\Sigma}_k^{-1}(x - \xbar_k)  + \log |\widehat{\Sigma}_k|, \label{eq:qda}
+	D_{QDA}(\bm x) = \argmin_{k} (\bm x - \xbar_k)^{T}\widehat{\bm
+    \Sigma}_k^{-1}(\bm x - \xbar_k) + \log |\widehat{\bm \Sigma}_k|, \label{eq:qda}
 \end{align}
-where $\xbar_k$ and $\widehat{\Sigma}_k$ are the maximum likelihood
-estimators for $\mu_k$ and $\Sigma_k$, respectively, and $A^{T}$ denotes the
-matrix transpose of $A \in \mathbb{R}_{m \times p}$. If we assume
-further that $\Sigma_k = \Sigma$ $(k = 1, \ldots, K)$, then the pooled
-sample covariance matrix $\widehat{\Sigma}$ is substituted for $\widehat{
-  \Sigma}_k$ in \eqref{eq:qda}, where
+where $\xbar_k$ and $\widehat{\bm \Sigma}_k$ are the maximum likelihood
+estimators for $\bm \mu_k$ and $\bm \Sigma_k$, respectively, and $\bm A^{T}$
+denotes the matrix transpose of $\bm A \in \mathbb{R}_{m \times p}$. If we
+assume further that $\bm \Sigma_k = \bm \Sigma$ $(k = 1, \ldots, K)$, then the
+pooled sample covariance matrix $\widehat{\bm \Sigma}$ is substituted for
+$\widehat{ \bm \Sigma}_k$ in \eqref{eq:qda}, where
 \begin{align}
-	\widehat{\Sigma} = N^{-1} \sum_{k=1}^K n_k \widehat{\Sigma}_k \label{eq:pooled-cov}
+	\widehat{\bm \Sigma} = N^{-1} \sum_{k=1}^K n_k \widehat{\bm \Sigma}_k \label{eq:pooled-cov}
 \end{align}
-is the maximum likelihood estimator for $\Sigma$. In this case \eqref{eq:qda}
+is the maximum likelihood estimator for $\bm \Sigma$. In this case \eqref{eq:qda}
 reduces to the sample linear discriminant analysis classifier, and the
@@
 Throughout the document we also require the following additional notation. Let
-$I_m$ be the $m \times m$ identity matrix, and let $0_{m \times p}$ be the $m
-\times p$ matrix of zeros, such that $0_m$ is understood to denote $0_{m \times
-  m}$. For $c \in \mathbb{R}$, let $c^+ = 1/c$ if $c \ne 0$ and $0$
-otherwise. Let $A \oplus B$ denote the direct sum of $A \in \mathbb{R}_{r \times
-  r}$ and $B \in \mathbb{R}_{s \times s}$ \citep[Chapter
-  1]{Lutkepohl:1996uz}. Let $A^+$ and $\mathcal{N}(A) = \{z \in \mathbb{R}_{p
-  \times 1} : A z = 0_{m \times 1}\}$ denote the Moore-Penrose pseudoinverse and
-null space of $A \in \mathbb{R}_{m \times p}$, respectively. Let $V^{\perp} =
-\{y \in \mathbb{R}_{p \times 1} : y^{T}v = 0\ \forall v \in V \}$ denote the
-orthogonal complement of a vector space $V \subset \mathbb{R}_{p \times 1}$.
+$\bm I_m$ be the $m \times m$ identity matrix, and let $\bm 0_{m \times p}$ be
+the $m \times p$ matrix of zeros, such that $\bm 0_m$ is understood to denote
+$\bm 0_{m \times m}$. For $c \in \mathbb{R}$, let $c^+ = 1/c$ if $c \ne 0$ and
+$0$ otherwise. Let $\bm A \oplus \bm B$ denote the direct sum of $\bm A \in
+\mathbb{R}_{r \times r}$ and $\bm B \in \mathbb{R}_{s \times s}$ \citep[Chapter
+  1]{Lutkepohl:1996uz}. Let $\bm A^+$ and $\mathcal{N}(\bm A) = \{\bm z \in
+\mathbb{R}_{p \times 1} : \bm A \bm z = \bm 0_{m \times 1}\}$ denote the
+Moore-Penrose pseudoinverse and null space of $\bm A \in \mathbb{R}_{m \times
+  p}$, respectively. Let $V^{\perp} = \{\bm y \in \mathbb{R}_{p \times 1} : \bm
+y^{T} \bm v = 0\ \forall \bm v \in V \}$ denote the orthogonal complement of a
+vector space $V \subset \mathbb{R}_{p \times 1}$.
 
@@
 The smallest eigenvalues and the directions associated with their eigenvectors
-highly influence \eqref{eq:qda}. In fact, the eigenvalues of $\widehat{
+highly influence \eqref{eq:qda}. In fact, the eigenvalues of $\widehat{ \bm
   \Sigma}_k$ are well-known to be biased if $p \ge n_k$ such that the smallest
 eigenvalues are underestimated \citep{Seber:2004uh}, Moreover, if $p > n_k$,
-then rank$(\widehat{\Sigma}_k) \le n_k$, which implies that at least $p -
-n_k$ eigenvalues of $\widehat{\Sigma}_k$ are 0. Furthermore, although more
+then rank$(\widehat{\bm \Sigma}_k) \le n_k$, which implies that at least $p -
+n_k$ eigenvalues of $\widehat{\bm \Sigma}_k$ are 0. Furthermore, although more
 feature information is available to discriminate among the $K$ classes, if $p >
-n_k$, \eqref{eq:qda} is incalculable because $\widehat{\Sigma}_k^{-1}$ does
+n_k$, \eqref{eq:qda} is incalculable because $\widehat{\bm \Sigma}_k^{-1}$ does
 not exist.
@@
 \citet*{Guo:2007te}, \cite{Mkhadri:1995jp}, and \citet*{Xu:2009fl}, have been
-proposed in the literature to adjust the eigenvalues of $\widehat{\Sigma}_k$
+proposed in the literature to adjust the eigenvalues of $\widehat{\bm \Sigma}_k$
 so that \eqref{eq:qda} is calculable and has reduced variability in the standard
@@
 \begin{align}
-	\widehat{\Sigma}_k(\gamma) =  \widehat{\Sigma}_k + \gamma I_p, \label{eq:ridge-estimator}
+	\widehat{\bm \Sigma}_k(\gamma) =  \widehat{\bm \Sigma}_k + \gamma \bm I_p, \label{eq:ridge-estimator}
 \end{align}
@@
 \eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance
-matrix $\widehat{\Sigma}_k$ towards $I_p$, thereby increasing the
-eigenvalues of $\widehat{\Sigma}_k$ by $\gamma$. Specifically, the zero
+matrix $\widehat{\bm \Sigma}_k$ towards $\bm I_p$, thereby increasing the
+eigenvalues of $\widehat{\bm \Sigma}_k$ by $\gamma$. Specifically, the zero
 eigenvalues are replaced with $\gamma$, so that \eqref{eq:ridge-estimator} is
@@
 \cite{Friedman:1989tm} has proposed regularized discriminant analysis by
-incorporating a weighted average of the sample covariance matrix $\widehat{
-  \Sigma}_k$ and the pooled sample covariance matrix $\widehat{\Sigma}$ to
-estimate $\Sigma_k$ with
+incorporating a weighted average of the sample covariance matrix $\widehat{ \bm
+  \Sigma}_k$ and the pooled sample covariance matrix $\widehat{\bm \Sigma}$ to
+estimate $\bm \Sigma_k$ with
 \begin{align}
-  \widehat{\Sigma}_k(\lambda) = n_k^{-1}(\lambda) S_k(\lambda) \quad (k =
+  \widehat{\bm \Sigma}_k(\lambda) = n_k^{-1}(\lambda) \bm S_k(\lambda) \quad (k =
   1, \ldots, K),\label{eq:sig-lambda}
@@
 where $\lambda \in [0, 1]$, $n_k(\lambda) = (1 - \lambda) n_k + \lambda N$, $
-S_k = n_k \widehat{\Sigma}_k$, $S = \sum_{k=1}^K S_k$, and $
-S_k(\lambda) = (1 - \lambda) S_k + \lambda S$. We can interpret
+\bm S_k = n_k \widehat{\bm \Sigma}_k$, $\bm S = \sum_{k=1}^K \bm S_k$, and $ \bm
+S_k(\lambda) = (1 - \lambda) \bm S_k + \lambda \bm S$. We can interpret
 \eqref{eq:sig-lambda} as a covariance-matrix estimator that borrows from
-$\widehat{\Sigma}$ in \eqref{eq:pooled-cov} to better estimate $
-\Sigma_k$. If $\lambda = 0$, \eqref{eq:sig-lambda} is equal to
-$\widehat{\Sigma}_k$ used in \eqref{eq:qda}. Alternatively, if $\lambda = 1$,
+$\widehat{\bm \Sigma}$ in \eqref{eq:pooled-cov} to better estimate $ \bm
+\Sigma_k$. If $\lambda = 0$, \eqref{eq:sig-lambda} is equal to $\widehat{\bm
+  \Sigma}_k$ used in \eqref{eq:qda}. Alternatively, if $\lambda = 1$,
 \eqref{eq:sig-lambda} reduces to \eqref{eq:pooled-cov}.
 
-To further improve the estimation of $\Sigma_k$ and to stabilize the inverse
+To further improve the estimation of $\bm \Sigma_k$ and to stabilize the inverse
 of \eqref{eq:sig-lambda}, \cite{Friedman:1989tm} has proposed the biased
@@
 \begin{align}
-	\widehat{\Sigma}_k(\lambda, \gamma) = (1 - \gamma) \widehat{\Sigma}_k(\lambda) + \gamma \frac{\tr\{\widehat{\Sigma}_k(\lambda)\}}{p} I_p,\label{eq:sig-rda}
+	\widehat{\bm \Sigma}_k(\lambda, \gamma) = (1 - \gamma) \widehat{\bm \Sigma}_k(\lambda) + \gamma \frac{\tr\{\widehat{\bm \Sigma}_k(\lambda)\}}{p} \bm I_p,\label{eq:sig-rda}
 \end{align}
 where $\gamma \in [0, 1]$ is a regularization parameter that controls the
-shrinkage of \eqref{eq:sig-rda} towards $I_p$ and is weighted by the average
+shrinkage of \eqref{eq:sig-rda} towards $\bm I_p$ and is weighted by the average
 of the eigenvalues of \eqref{eq:sig-lambda}. Thus, the \emph{pooling} parameter
-$\lambda$ controls the amount borrowed from $\widehat{\Sigma}$ to estimate
-$\Sigma_k$, and the \emph{shrinkage} parameter $\gamma$ determines the
+$\lambda$ controls the amount borrowed from $\widehat{\bm \Sigma}$ to estimate
+$\bm \Sigma_k$, and the \emph{shrinkage} parameter $\gamma$ determines the
 amount of eigenvalue shrinkage applied. Finally, by substituting
@@
 \begin{align}
-	D_{RDA}(x) = \argmin_{k}  (x - \xbar_k)^{T}\widehat{\Sigma}_k(\lambda, \gamma)^{-1}(x - \xbar_k)  + \log |\widehat{\Sigma}_k(\lambda, \gamma)|. \label{eq:rda}
+	D_{RDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)^{T}\widehat{\bm \Sigma}_k(\lambda, \gamma)^{-1}(\bm x - \xbar_k)  + \log |\widehat{\bm \Sigma}_k(\lambda, \gamma)|. \label{eq:rda}
 \end{align}
@@
 Here, we define our adaptation of \citeauthor{Friedman:1989tm}'s classifier by
-first simplifying the formulation of $\widehat{\Sigma}_k(\lambda)$ in
+first simplifying the formulation of $\widehat{\bm \Sigma}_k(\lambda)$ in
 \eqref{eq:sig-lambda} and demonstrating its clear interpretation as a linear
@@
 \begin{align}
-  \widehat{\Sigma}_k(\lambda) = (1 - \lambda) \widehat{\Sigma}_k + \lambda
-  \widehat{\Sigma}, \quad (k = 1, \ldots, K)\label{eq:sig-lambda-alternative}
+  \widehat{\bm \Sigma}_k(\lambda) = (1 - \lambda) \widehat{\bm \Sigma}_k + \lambda
+  \widehat{\bm \Sigma}, \quad (k = 1, \ldots, K)\label{eq:sig-lambda-alternative}
 \end{align}
@@
 rewriting \eqref{eq:sig-lambda-alternative} in terms of the observations $
-x_i$ $(i = 1, \ldots, N)$, each centered by its class sample mean, we attain a
-clear interpretation of $\widehat{\Sigma}_k(\lambda)$. That is,
+\bm x_i$ $(i = 1, \ldots, N)$, each centered by its class sample mean, we attain a
+clear interpretation of $\widehat{\bm \Sigma}_k(\lambda)$. That is,
 \begin{align}
-	\widehat{\Sigma}_k(\lambda) &= \widehat{\Sigma}_k + \lambda (\widehat{\Sigma} - \widehat{\Sigma}_k) \nonumber\\
-%	&= \widehat{\Sigma}_k + \lambda \left(\sum_{k' = 1}^K \frac{n_k}{N} \widehat{\Sigma}_{k'}  - \widehat{\Sigma}_k \right) \nonumber\\
-	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right) \widehat{\Sigma}_k +  \frac{\lambda}{N} \sum_{\substack{k' = 1\\k' \ne k}} n_{k'} \widehat{\Sigma}_{k'} \nonumber \\
+	\widehat{\bm \Sigma}_k(\lambda) &= \widehat{\bm \Sigma}_k + \lambda (\widehat{\bm \Sigma} - \widehat{\bm \Sigma}_k) \nonumber\\
+%	&= \widehat{\bm \Sigma}_k + \lambda \left(\sum_{k' = 1}^K \frac{n_k}{N} \widehat{\bm \Sigma}_{k'}  - \widehat{\bm \Sigma}_k \right) \nonumber\\
+	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right) \widehat{\bm \Sigma}_k +  \frac{\lambda}{N} \sum_{\substack{k' = 1\\k' \ne k}} n_{k'} \widehat{\bm \Sigma}_{k'} \nonumber \\
 %	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N I(y_i = k) x_i x_i^{T} +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k) x_i x_i^{T} \nonumber \\
-	&= \sum_{i=1}^N c_{ik}(\lambda) x_i x_i^{T},\label{eq:sig-lambda-alternative2}
+	&= \sum_{i=1}^N c_{ik}(\lambda) \bm x_i \bm x_i^{T},\label{eq:sig-lambda-alternative2}
 \end{align}
@@
 From \eqref{eq:sig-lambda-alternative2}, $\lambda$ weights the contribution of
-each of the $N$ observations in estimating $\Sigma_k$. Hence, for $0 < \lambda
-\le 1$, we estimate $\Sigma_k$ with $N$ rather than $n_k$
+each of the $N$ observations in estimating $\bm \Sigma_k$. Hence, for $0 <
+\lambda \le 1$, we estimate $\bm \Sigma_k$ with $N$ rather than $n_k$
 observations. Below, we will show that the pooling operation is advantageous in
-increasing the rank of each $\widehat{\Sigma}_k(\lambda)$ from
-rank$(\widehat{\Sigma}_k)$ to rank$(\widehat{\Sigma})$ for $0 < \lambda
+increasing the rank of each $\widehat{\bm \Sigma}_k(\lambda)$ from
+rank$(\widehat{\bm \Sigma}_k)$ to rank$(\widehat{\bm \Sigma})$ for $0 < \lambda
 \le 1$. Notice that if $\lambda = 0$, then the observations from the remaining
-$K - 1$ classes do not contribute to the estimation of $\Sigma_k$,
-corresponding to $\widehat{\Sigma}_k$. Furthermore, if $\lambda = 1$, the
+$K - 1$ classes do not contribute to the estimation of $\bm \Sigma_k$,
+corresponding to $\widehat{\bm \Sigma}_k$. Furthermore, if $\lambda = 1$, the
 weights in \eqref{eq:sig-lambda-alternative2} reduce to $1/N$, corresponding to
-$\widehat{\Sigma}$. For brevity, when $\lambda = 1$, we define $X =
-[\sqrt{c_{1k}(1)} x_1^{T}, \ldots, \sqrt{c_{Nk}(1)} x_N^{T}]^{T}$ such that
-$\widehat{\Sigma} = X^{T} X$. Similarly, for $\lambda = 0$, we define
-$X_k = [\sqrt{c_{1k}(0)} x_1^{T}, \ldots, \sqrt{c_{Nk}(0)} x_N^{T}]^{T}$ such
-that $\widehat{\Sigma}_k = X_k^{T} X_k$. In Figure
-\ref{fig:hdrda-contours} we plot the contours of five multivariate normal
+$\widehat{\bm \Sigma}$. For brevity, when $\lambda = 1$, we define $\bm X =
+[\sqrt{c_{1k}(1)} \bm x_1^{T}, \ldots, \sqrt{c_{Nk}(1)} \bm x_N^{T}]^{T}$ such
+that $\widehat{\bm \Sigma} = \bm X^{T} \bm X$. Similarly, for $\lambda = 0$, we
+define $\bm X_k = [\sqrt{c_{1k}(0)} \bm x_1^{T}, \ldots, \sqrt{c_{Nk}(0)} \bm
+  x_N^{T}]^{T}$ such that $\widehat{\bm \Sigma}_k = \bm X_k^{T} \bm X_k$. In
+Figure \ref{fig:hdrda-contours} we plot the contours of five multivariate normal
 populations for $\lambda = 0$ with unequal covariance matrices. As $\lambda$
@@
 \begin{align}
-	\tilde{\Sigma}_k = \alpha_k \widehat{\Sigma}_k(\lambda) + \gamma I_p,\label{eq:hdrda-cov}
+	\tilde{\bm \Sigma}_k = \alpha_k \widehat{\bm \Sigma}_k(\lambda) + \gamma \bm I_p,\label{eq:hdrda-cov}
 \end{align}
@@
 \begin{align}
-	D_{HDRDA}(x) = \argmin_{k}  (x - \xbar_k)^{T}\tilde{\Sigma}_k^{+}(x - \xbar_k)  + \log |\tilde{\Sigma}_k|. \label{eq:hdrda}
+	D_{HDRDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)^{T}\tilde{\bm \Sigma}_k^{+}(\bm x - \xbar_k)  + \log |\tilde{\bm \Sigma}_k|. \label{eq:hdrda}
 \end{align}
-For $\gamma > 0$, $\tilde{\Sigma}_k$ is nonsingular such that
-$\tilde{\Sigma}_k^{-1}$ can be substituted for $\tilde{\Sigma}_k^{+}$ in
-\eqref{eq:hdrda}. If $\gamma = 0$, we explicitly set $|\tilde{\Sigma}_k|$ equal
-to the product of the positive eigenvalues of $\tilde{\Sigma}_k$. Following
-\cite{Friedman:1989tm}, we select $\lambda$ and $\gamma$ from a grid of
-candidate models via cross-validation \citep{Hastie:2008dt}. We provide an
-implementation of \eqref{eq:hdrda} in the {\tt hdrda} function in the {\tt
-  sparsediscrim} R package, available on CRAN.
+For $\gamma > 0$, $\tilde{\bm \Sigma}_k$ is nonsingular such that $\tilde{\bm
+  \Sigma}_k^{-1}$ can be substituted for $\tilde{\bm \Sigma}_k^{+}$ in
+\eqref{eq:hdrda}. If $\gamma = 0$, we explicitly set $|\tilde{\bm \Sigma}_k|$
+equal to the product of the positive eigenvalues of $\tilde{\bm
+  \Sigma}_k$. Following \cite{Friedman:1989tm}, we select $\lambda$ and $\gamma$
+from a grid of candidate models via cross-validation \citep{Hastie:2008dt}. We
+provide an implementation of \eqref{eq:hdrda} in the {\tt hdrda} function in the
+{\tt sparsediscrim} R package, available on CRAN.
 
@@
 to low-dimensional matrices and the second summand corresponds to the null space
-of $\widehat{\Sigma}$ in \eqref{eq:pooled-cov}. We show that the matrix
-operations performed on the null space of $\widehat{\Sigma}$ yield constant
+of $\widehat{\bm \Sigma}$ in \eqref{eq:pooled-cov}. We show that the matrix
+operations performed on the null space of $\widehat{\bm \Sigma}$ yield constant
 quadratic forms across all classes and can be omitted. For $p \gg N$, the
@@
 value decomposition to calculate efficiently the eigenvalue decomposition of
-$\widehat{\Sigma}$, further reducing the computational costs of our proposed
+$\widehat{\bm \Sigma}$, further reducing the computational costs of our proposed
 classifier.
@@
 First, we require the following relationship regarding the null spaces of
-$\widehat{\Sigma}_k(\lambda)$, $\widehat{\Sigma}$, and $\widehat{
+$\widehat{\bm \Sigma}_k(\lambda)$, $\widehat{\bm \Sigma}$, and $\widehat{ \bm
   \Sigma}_k$.
@@
 \begin{lemma}\label{lemma:null-spaces}
-Let $\widehat{\Sigma}_k$ and $\widehat{\Sigma}$ be the maximum
-likelihood estimators of $\Sigma_k$ and $\Sigma$, respectively. Let
-$\widehat{\Sigma}_k(\lambda)$ be defined as in
-\eqref{eq:sig-lambda-alternative}. Then, $\mathcal{N}\{\widehat{\Sigma}_k(\lambda)\} \subset \mathcal{N}(\widehat{
-  \Sigma}) \subset \mathcal{N}(\widehat{\Sigma}_k)$ $(k = 1, \ldots, K)$.
+Let $\widehat{\bm \Sigma}_k$ and $\widehat{\bm \Sigma}$ be the maximum
+likelihood estimators of $\bm \Sigma_k$ and $\bm \Sigma$, respectively. Let
+$\widehat{\bm \Sigma}_k(\lambda)$ be defined as in
+\eqref{eq:sig-lambda-alternative}. Then, $\mathcal{N}\{\widehat{\bm \Sigma}_k(\lambda)\} \subset \mathcal{N}(\widehat{
+  \bm \Sigma}) \subset \mathcal{N}(\widehat{\bm \Sigma}_k)$ $(k = 1, \ldots, K)$.
 \end{lemma}
 \begin{proof}
-Let $z \in \mathcal{N}\{\widehat{\Sigma}_k(\lambda)\}$ for some $k = 1,
-\ldots, K$. Hence, $0 = z^{T} \widehat{\Sigma}_k(\lambda) z = (1 -
-\lambda) z^{T} \widehat{\Sigma}_k z + \lambda z^{T} \widehat{
-  \Sigma} z$. Because $\widehat{\Sigma}_k, \widehat{\Sigma}\in
-\mathbb{R}_{p \times p}^{\ge}$, we have $z \in \mathcal{N}(\widehat{
-  \Sigma})$ and $z \in \mathcal{N}(\widehat{\Sigma}_k)$. In particular,
-we have that $\mathcal{N}\{\widehat{\Sigma}_k(\lambda)\} \subset
-\mathcal{N}(\widehat{\Sigma})$. Now, suppose $z \in
-\mathcal{N}(\widehat{\Sigma})$. Similarly, we have that $0 = z^{T}
-\widehat{\Sigma} z = N^{-1} \sum_{k = 1}^K n_k z^{T} \widehat{
-  \Sigma}_k z$, which implies that $z \in \mathcal{N}(\widehat{
-  \Sigma}_k)$ because $\widehat{\Sigma}_k \in \mathbb{R}_{p \times
-  p}^{\ge}$. Therefore, $\mathcal{N}(\widehat{\Sigma}) \subset
-\mathcal{N}(\widehat{\Sigma}_k)$.
+Let $\bm z \in \mathcal{N}\{\widehat{\bm \Sigma}_k(\lambda)\}$ for some $k = 1,
+\ldots, K$. Hence, $0 = \bm z^{T} \widehat{\bm \Sigma}_k(\lambda) \bm z = (1 -
+\lambda) \bm z^{T} \widehat{\bm \Sigma}_k \bm z + \lambda \bm z^{T} \widehat{
+  \bm \Sigma} \bm z$. Because $\widehat{\bm \Sigma}_k, \widehat{\bm \Sigma}\in
+\mathbb{R}_{p \times p}^{\ge}$, we have $\bm z \in \mathcal{N}(\widehat{
+  \bm \Sigma})$ and $\bm z \in \mathcal{N}(\widehat{\bm \Sigma}_k)$. In particular,
+we have that $\mathcal{N}\{\widehat{\bm \Sigma}_k(\lambda)\} \subset
+\mathcal{N}(\widehat{\bm \Sigma})$. Now, suppose $\bm z \in
+\mathcal{N}(\widehat{\bm \Sigma})$. Similarly, we have that $0 = \bm z^{T}
+\widehat{\bm \Sigma} \bm z = N^{-1} \sum_{k = 1}^K n_k \bm z^{T} \widehat{
+  \bm \Sigma}_k \bm z$, which implies that $\bm z \in \mathcal{N}(\widehat{
+  \bm \Sigma}_k)$ because $\widehat{\bm \Sigma}_k \in \mathbb{R}_{p \times
+  p}^{\ge}$. Therefore, $\mathcal{N}(\widehat{\bm \Sigma}) \subset
+\mathcal{N}(\widehat{\bm \Sigma}_k)$.
 \end{proof}
@@
 In Lemma \ref{lemma:rda-tilde-Sigma_k} below, we derive an alternative
-expression for $\tilde{\Sigma}_k$ in terms of the matrix of eigenvectors of
-$\widehat{\Sigma}$. Let $\widehat{\Sigma} = U D U^{T}$ be the
-eigendecomposition of $\widehat{\Sigma}$ such that $D \in \mathbb{R}_{p
-  \times p}^{\ge}$ is the diagonal matrix of eigenvalues of $\widehat{\Sigma}$
-with $D = D_q \oplus 0_{p-q}$, $D_q \in \mathbb{R}_{q \times
+expression for $\tilde{\bm \Sigma}_k$ in terms of the matrix of eigenvectors of
+$\widehat{\bm \Sigma}$. Let $\widehat{\bm \Sigma} = \bm U \bm D \bm U^{T}$ be the
+eigendecomposition of $\widehat{\bm \Sigma}$ such that $\bm D \in \mathbb{R}_{p
+  \times p}^{\ge}$ is the diagonal matrix of eigenvalues of $\widehat{\bm \Sigma}$
+with $\bm D = \bm D_q \oplus \bm 0_{p-q}$, $\bm D_q \in \mathbb{R}_{q \times
   q}^{>}$ is the diagonal matrix consisting of the positive eigenvalues of
-$\widehat{\Sigma}$, the columns of $U \in \mathbb{R}_{p \times p}$ are
-the corresponding orthonormal eigenvectors of $\widehat{\Sigma}$, and
-rank$(\widehat{\Sigma}) = q$. Then, we partition $U = (U_1, 
-U_2)$ such that $U_1 \in \mathbb{R}_{p \times q}$ and $U_2 \in
+$\widehat{\bm \Sigma}$, the columns of $\bm U \in \mathbb{R}_{p \times p}$ are
+the corresponding orthonormal eigenvectors of $\widehat{\bm \Sigma}$, and
+rank$(\widehat{\bm \Sigma}) = q$. Then, we partition $\bm U = (\bm U_1, 
+\bm U_2)$ such that $\bm U_1 \in \mathbb{R}_{p \times q}$ and $\bm U_2 \in
 \mathbb{R}_{p \times (p - q)}$.
@@
 \begin{lemma}\label{lemma:rda-tilde-Sigma_k}
-Let $\widehat{\Sigma} = U D U^{T}$ be the eigendecomposition of
-$\widehat{\Sigma}$ as above, and suppose that rank$(\widehat{\Sigma}) =
+Let $\widehat{\bm \Sigma} = \bm U \bm D \bm U^{T}$ be the eigendecomposition of
+$\widehat{\bm \Sigma}$ as above, and suppose that rank$(\widehat{\bm \Sigma}) =
 q \le p$. Then, we have
 \begin{align}
-	\tilde{\Sigma}_k &= U(W_k \oplus \gamma I_{p-q})U^{T} \quad
+	\tilde{\bm \Sigma}_k &= \bm U(\bm W_k \oplus \gamma \bm I_{p-q})\bm U^{T} \quad
   (k = 1, \ldots, K),\label{eq:rda-matrix}
 \intertext{where}
-W_k &= \alpha_k \{(1 - \lambda) U_1^{T} \widehat{\Sigma}_k U_1 + \lambda D_q\} + \gamma I_{q}.\label{eq:Wk}
+\bm W_k &= \alpha_k \{(1 - \lambda) \bm U_1^{T} \widehat{\bm \Sigma}_k \bm U_1 + \lambda \bm D_q\} + \gamma \bm I_{q}.\label{eq:Wk}
 \end{align}
@@
 \begin{proof}
-From Lemma \ref{lemma:null-spaces}, the columns of $U_2$ span the null space
-of $\widehat{\Sigma}$, which implies that $\widehat{\Sigma}_k U_2 =
-0_{p \times (p - q)}$. Hence, $U^{T} \widehat{\Sigma}_k U = 
-U_1^{T} \widehat{\Sigma}_k U_1 \oplus 0_{p-q}$ $(k = 1, \ldots,
-K)$. Thus, $U^{T} \tilde{\Sigma}_k U = \alpha_k \{(1 - \lambda) U^{T}
-\widehat{\Sigma}_k U + \lambda D\} + \gamma I_p$, and
-\eqref{eq:rda-matrix} holds because $U$ is orthogonal.
+From Lemma \ref{lemma:null-spaces}, the columns of $\bm U_2$ span the null space
+of $\widehat{\bm \Sigma}$, which implies that $\widehat{\bm \Sigma}_k \bm U_2 =
+\bm 0_{p \times (p - q)}$. Hence, $\bm U^{T} \widehat{\bm \Sigma}_k \bm U = \bm
+U_1^{T} \widehat{\bm \Sigma}_k \bm U_1 \oplus \bm 0_{p-q}$ $(k = 1, \ldots,
+K)$. Thus, $\bm U^{T} \tilde{\bm \Sigma}_k \bm U = \alpha_k \{(1 - \lambda) \bm
+U^{T} \widehat{\bm \Sigma}_k \bm U + \lambda \bm D\} + \gamma \bm I_p$, and
+\eqref{eq:rda-matrix} holds because $\bm U$ is orthogonal.
 \end{proof}
@@
 \begin{cor}
-Let $\widehat{\Sigma}_k(\lambda)$ be defined as in
+Let $\widehat{\bm \Sigma}_k(\lambda)$ be defined as in
 \eqref{eq:sig-lambda-alternative}. Then, for $\lambda \in (0, 1]$,
-  rank$\{\widehat{\Sigma}_k(\lambda)\} = q$ $(k = 1, \ldots, K)$.
+  rank$\{\widehat{\bm \Sigma}_k(\lambda)\} = q$ $(k = 1, \ldots, K)$.
 \end{cor}
@@
 
-Thus, by incorporating each $x_i$ into the estimation of $\Sigma_k$, we
-increase the rank of $\widehat{\Sigma}_k(\lambda)$ to $q$ if $\lambda \ne
+Thus, by incorporating each $\bm x_i$ into the estimation of $\bm \Sigma_k$, we
+increase the rank of $\widehat{\bm \Sigma}_k(\lambda)$ to $q$ if $\lambda \ne
 0$. Next, we provide an essential result that enables us to prove that
-\eqref{eq:hdrda} is invariant to adjustments to the eigenvalues of $\tilde{
-  \Sigma}_k$ corresponding to the null space of $\widehat{\Sigma}$ for any
-unlabeled observation $x \in \mathbb{R}_{p \times 1}$.
+\eqref{eq:hdrda} is invariant to adjustments to the eigenvalues of $\tilde{ \bm
+  \Sigma}_k$ corresponding to the null space of $\widehat{\bm \Sigma}$ for any
+unlabeled observation $\bm x \in \mathbb{R}_{p \times 1}$.
 
 \begin{lemma}\label{lemma:RDA-constant-term}
-Let $U_2$ be defined as above. Then, for all $x \in \mathbb{R}_{p \times
-  1}$, $U_2^{T} (x - \xbar_k) = U_2^{T} (x - \xbar_{k'})$ $(1 \le k,
+Let $\bm U_2$ be defined as above. Then, for all $\bm x \in \mathbb{R}_{p \times
+  1}$, $\bm U_2^{T} (\bm x - \xbar_k) = \bm U_2^{T} (\bm x - \xbar_{k'})$ $(1 \le k,
 k' \le K)$.
@@
 \begin{proof}
-Let $x \in \mathbb{R}_{p \times 1}$, and suppose that $1 \le k, k' \le K$.
-Recall that $U_2 \in \mathcal{N}(\widehat{\Sigma})$, which implies that
-$U_2^{T} \in \mathcal{C}(\widehat{\Sigma})^{\perp}$ \citep[Lemma
-  1.2.5]{Kollo:2005vp}. Now, because $x_i \in \mathcal{C}(\widehat{
-  \Sigma})$ $(i = 1, \ldots, N)$, $U_2^{T} x_i = 0_{(p-q) \times 1}$.
-Hence, $0_{(p-q) \times 1} = \sum_{i=1}^N \beta_i U_2^{T}x_i = U_2^{T}(\xbar_k -
-\xbar_{k'})$, where $\beta_i = (n_k n_{k'})^{-1} \{ I(y_i = k) n_{k'} - I(y_i =
-k') n_k \}$. Therefore, $U_2^{T} (x - \xbar_k) = U_2^{T} (x - \xbar_{k'})$.
+Let $\bm x \in \mathbb{R}_{p \times 1}$, and suppose that $1 \le k, k' \le K$.
+Recall that $\bm U_2 \in \mathcal{N}(\widehat{\bm \Sigma})$, which implies that
+$\bm U_2^{T} \in \mathcal{C}(\widehat{\bm \Sigma})^{\perp}$ \citep[Lemma
+  1.2.5]{Kollo:2005vp}. Now, because $\bm x_i \in \mathcal{C}(\widehat{ \bm
+  \Sigma})$ $(i = 1, \ldots, N)$, $\bm U_2^{T} \bm x_i = \bm 0_{(p-q) \times
+  1}$.  Hence, $\bm 0_{(p-q) \times 1} = \sum_{i=1}^N \beta_i \bm U_2^{T}\bm x_i
+= \bm U_2^{T}(\xbar_k - \xbar_{k'})$, where $\beta_i = (n_k n_{k'})^{-1} \{
+I(y_i = k) n_{k'} - I(y_i = k') n_k \}$. Therefore, $\bm U_2^{T} (\bm x -
+\xbar_k) = \bm U_2^{T} (\bm x - \xbar_{k'})$.
 \end{proof}
@@
 \begin{thm}
-Let $\tilde{\Sigma}_k$ and $W_k$ be defined as in \eqref{eq:rda-matrix}
-and \eqref{eq:Wk}, respectively, and let $U_1$ be defined as above. Then,
+Let $\tilde{\bm \Sigma}_k$ and $\bm W_k$ be defined as in \eqref{eq:rda-matrix}
+and \eqref{eq:Wk}, respectively, and let $\bm U_1$ be defined as above. Then,
 the decision rule in \eqref{eq:hdrda} is equivalent to
 	\begin{align}
-		D_{HDRDA}(x) &= \argmin_k  (x - \xbar_k)^{T} U_1 W_k^{-1} U_1^{T} (x - \xbar_k) + \log | W_k |. \label{eq:hdrda-decomposed}
+		D_{HDRDA}(\bm x) &= \argmin_k  (\bm x - \xbar_k)^{T} \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm x - \xbar_k) + \log | \bm W_k |. \label{eq:hdrda-decomposed}
 	\end{align}
@@
 \begin{proof}
-From \eqref{eq:rda-matrix}, we have that $\tilde{\Sigma}_k^{+} = U(
-W_k^{-1} \oplus \gamma^{+} I_{p-q} )U^{T}$ and $|\tilde{\Sigma}_k| =
-\gamma^{p-q}| W_k |$ $(k = 1, \ldots, K)$. Therefore, for all $x \in
+From \eqref{eq:rda-matrix}, we have that $\tilde{\bm \Sigma}_k^{+} = \bm U(
+\bm W_k^{-1} \oplus \gamma^{+} \bm I_{p-q} )\bm U^{T}$ and $|\tilde{\bm \Sigma}_k| =
+\gamma^{p-q}| \bm W_k |$ $(k = 1, \ldots, K)$. Therefore, for all $\bm x \in
 \mathbb{R}_{p \times 1}$, we have that
 	\begin{align*}
-	(x - \xbar_k)^{T} \tilde{\Sigma}_k^{+}(x - \xbar_k)  + \log |\tilde{\Sigma}_k| &= (x - \xbar_k)^{T} U_1 W_k^{-1} U_1^{T} (x - \xbar_k)\\
-	&+ \gamma^{+} (x - \xbar_k)^{T} U_2 U_2^{T} (x - \xbar_k) + \log | W_k | + (p - q) \log \gamma.
+	(\bm x - \xbar_k)^{T} \tilde{\bm \Sigma}_k^{+}(\bm x - \xbar_k)  + \log |\tilde{\bm \Sigma}_k| &= (\bm x - \xbar_k)^{T} \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm x - \xbar_k)\\
+	&+ \gamma^{+} (\bm x - \xbar_k)^{T} \bm U_2 \bm U_2^{T} (\bm x - \xbar_k) + \log | \bm W_k | + (p - q) \log \gamma.
 	\end{align*}
@@
 = 0$. Then, the proof follows from Lemma \ref{lemma:RDA-constant-term} because
-$U_2^{T} (x - \xbar_k)$ is constant for $k = 1, \ldots, K$.
+$\bm U_2^{T} (\bm x - \xbar_k)$ is constant for $k = 1, \ldots, K$.
 \end{proof}
@@
 Thus far, we have shown that our classifier in \eqref{eq:hdrda} is invariant to
-the term $U_2$, thus yielding an equivalent classifier in
+the term $\bm U_2$, thus yielding an equivalent classifier in
 \eqref{eq:hdrda-decomposed} with a substantial reduction in the computational
 complexity. Here, we demonstrate that the computational efficiency in
-calculating the inverse and determinants of $W_k$ can be further improved via
-standard matrix operations.
+calculating the inverse and determinants of $\bm W_k$ can be further improved
+via standard matrix operations.
 
 \begin{proposition}\label{proposition:hdrda-W_k}
-Let $W_k$ be defined as above $(k = 1, \ldots, K)$. Then, $|W_k| = |
-\Gamma_k| |Q_k|$ and
+Let $\bm W_k$ be defined as above $(k = 1, \ldots, K)$. Then, $|\bm W_k| = |
+\bm \Gamma_k| |\bm Q_k|$ and
 	\begin{align}
-		W_k^{-1} &= \Gamma_k^{-1} - \alpha_k(1 - \lambda) \Gamma_k^{-1} U_1^{T} X_k^{T} Q_k^{-1} X_k U_1 \Gamma_k^{-1},\label{eq:W_k_inv}
+		\bm W_k^{-1} &= \bm \Gamma_k^{-1} - \alpha_k(1 - \lambda) \bm \Gamma_k^{-1} \bm U_1^{T} \bm X_k^{T} \bm Q_k^{-1} \bm X_k \bm U_1 \bm \Gamma_k^{-1},\label{eq:W_k_inv}
 	\end{align}
-where $Q_k = I_{n_k} + \alpha_k(1 - \lambda) X_k U_1 
-\Gamma_k^{-1} U_1^{T} X_k^{T}$ and $\Gamma_k = \alpha_k \lambda D_q +
-\gamma I_q$.
+where $\bm Q_k = \bm I_{n_k} + \alpha_k(1 - \lambda) \bm X_k \bm U_1 
+\bm \Gamma_k^{-1} \bm U_1^{T} \bm X_k^{T}$ and $\bm \Gamma_k = \alpha_k \lambda \bm D_q +
+\gamma \bm I_q$.
 \end{proposition}
 \begin{proof}
-To calculate $|W_k|$, we apply Theorem 18.1.1 from \cite{Harville:2008wja},
-which states that $|A + B T C| = |A| |T| |T^{-1} +
-C A^{-1} B|$. Thus, setting $A = \Gamma_k$, $B =
-\alpha_k (1 - \lambda) U_1^{T} X_k^{T}$, $T = I_{n_k}$, and $C =
-X_k U_1$, we have $|W_k| = |\Gamma_k| |Q_k|$. Also,
+To calculate $|\bm W_k|$, we apply Theorem 18.1.1 from \cite{Harville:2008wja},
+which states that $|\bm A + \bm B \bm T \bm C| = |\bm A| |\bm T| |\bm T^{-1} +
+\bm C \bm A^{-1} \bm B|$. Thus, setting $\bm A = \bm \Gamma_k$, $\bm B =
+\alpha_k (1 - \lambda) \bm U_1^{T} \bm X_k^{T}$, $\bm T = \bm I_{n_k}$, and $\bm
+C = \bm X_k \bm U_1$, we have $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$. Also,
 \eqref{eq:W_k_inv} follows from the well-known Sherman-Woodbury formula
@@
 
-Conveniently, we see that the inverse and determinant of $W_k$ are each in terms
-of the inverse and determinant of $Q_k \in \mathbb{R}_{n_k \times n_k}$,
-respectively, as well as the diagonal matrix $\Gamma_k \in \mathbb{R}_{q \times
-  q}$. This formulation is especially advantageous when the number of classes
-$K$ is large because $n_k$ is then potentially much smaller than $q$. The
-expressions given in Proposition \ref{proposition:hdrda-W_k} also expedite the
-selection of $\lambda$ and $\gamma$ via cross-validation. The most
-time-consuming matrix operations involved in computing $W_k^{-1}$ and $|W_k|$ is
-$X_k U_1 \in \mathbb{R}_{n_k \times q}$, which is independent of $\lambda$ and
-$\gamma$. Hence, we need calculate $X_k U_1$ only once for each cross-validation
-fold. The subsequent operations in calculating $W_k^{-1}$ and $|W_K|$ can be
-simply updated for different pairs of $\lambda$ and $\gamma$.
+Conveniently, we see that the inverse and determinant of $\bm W_k$ are each in
+terms of the inverse and determinant of $\bm Q_k \in \mathbb{R}_{n_k \times
+  n_k}$, respectively, as well as the diagonal matrix $\bm \Gamma_k \in
+\mathbb{R}_{q \times q}$. This formulation is especially advantageous when the
+number of classes $K$ is large because $n_k$ is then potentially much smaller
+than $q$. The expressions given in Proposition \ref{proposition:hdrda-W_k} also
+expedite the selection of $\lambda$ and $\gamma$ via cross-validation. The most
+time-consuming matrix operations involved in computing $\bm W_k^{-1}$ and $|\bm
+W_k|$ is $\bm X_k \bm U_1 \in \mathbb{R}_{n_k \times q}$, which is independent
+of $\lambda$ and $\gamma$. Hence, we need calculate $\bm X_k \bm U_1$ only once
+for each cross-validation fold. The subsequent operations in calculating $\bm
+W_k^{-1}$ and $|\bm W_K|$ can be simply updated for different pairs of $\lambda$
+and $\gamma$.
 
-Thus far, we have constructed each $\tilde{\Sigma}_k$ in terms of the first $q$
-eigenvectors of $\widehat{\Sigma}$. By utilizing a technique similar to the
-so-called kernel trick, we can obtain $U_1$ by computing the eigendecomposition
+Thus far, we have constructed each $\tilde{\bm \Sigma}_k$ in terms of the first $q$
+eigenvectors of $\widehat{\bm \Sigma}$. By utilizing a technique similar to the
+so-called kernel trick, we can obtain $\bm U_1$ by computing the eigendecomposition
 of a much smaller $N \times N$ matrix when $p \gg N$
 \citep{Hastie:2008dt}. Recall that the pooled sample covariance matrix is
-$\widehat{\Sigma} = X^{T} X$. Then, because $X = M D^{1/2} U^{T}$ by the
-singular value decomposition, we have the eigendecomposition $\widehat{\Sigma} =
-U D U^{T}$, where $M \in \mathbb{R}_{N \times N}$ is orthogonal, $D^{1/2} \in
+$\widehat{\bm \Sigma} = \bm X^{T} \bm X$. Then, because $\bm X = \bm M \bm D^{1/2} \bm U^{T}$ by the
+singular value decomposition, we have the eigendecomposition $\widehat{\bm \Sigma} =
+\bm U \bm D \bm U^{T}$, where $\bm M \in \mathbb{R}_{N \times N}$ is orthogonal, $\bm D^{1/2} \in
 \mathbb{R}_{N \times N}^{\ge}$ is a diagonal matrix consisting of the singular
-values of $X$, and $U \in \mathbb{R}_{p \times N}$ is orthogonal. Now, $X X^{T}
-= M D M^{T}$ is the eigenvalue decomposition of $X X^{T}$, where $D$ is the
-matrix of eigenvalues of $X X^{T}$ and the columns of $M$ are the corresponding
-eigenvectors. Hence, we obtain $M$ and $D$ from an eigendecomposition of $X
-X^{T} \in \mathbb{R}_{N \times N}$. Next, we compute $U = X^{T} M D^{+/2}$,
-where $D^{+/2} = D_q^{-1/2} \oplus 0_{N-q}$. We then determine $q$, the number
-of numerically nonzero eigenvalues present in $D$, by calculating the number of
+values of $\bm X$, and $\bm U \in \mathbb{R}_{p \times N}$ is orthogonal. Now, $\bm X \bm X^{T}
+= \bm M \bm D \bm M^{T}$ is the eigenvalue decomposition of $\bm X \bm X^{T}$, where $\bm D$ is the
+matrix of eigenvalues of $\bm X \bm X^{T}$ and the columns of $\bm M$ are the corresponding
+eigenvectors. Hence, we obtain $\bm M$ and $\bm D$ from an eigendecomposition of $\bm X
+\bm X^{T} \in \mathbb{R}_{N \times N}$. Next, we compute $\bm U = \bm X^{T} \bm M \bm D^{+/2}$,
+where $\bm D^{+/2} = \bm D_q^{-1/2} \oplus \bm 0_{N-q}$. We then determine $q$, the number
+of numerically nonzero eigenvalues present in $\bm D$, by calculating the number of
 eigenvalues that exceeds some tolerance value, say, $\epsilon = 1 \times
-10^{-6}$. After calculating $q$, we extract $U_1$ by retaining only the first
-$q$ columns of $U$.
+10^{-6}$. After calculating $q$, we extract $\bm U_1$ by retaining only the first
+$q$ columns of $\bm U$.
 

### 2013-05-10 reports/paper-rda.Rnw
Added back the HDRDA acronym in various places.

@@
 \cite{Huang:2010ju}, have introduced restrictions on the population covariance
-matrices in linear and quadratic discriminant analysis to reduce number of
+matrices in linear and quadratic discriminant analysis to reduce the number of
 covariance parameters that must be estimated to ensure that the resulting
@@
 unequal covariance matrices while retaining linear and quadratic discriminant
-analysis as special cases. We then establish our classifier by applying
+analysis as special cases. We then establish the high-dimensional regularized
+discriminant analysis (\emph{HDRDA}) classifier by applying
 covariance-matrix regularization with a parameterization that includes a large
@@
 matrix. Applying a reasoning similar to \cite{Ye:2006tq}, we decompose the
-decision rule of our classifier and show that the sum related to the null space
-of the pooled sample covariance matrix can be discarded. Hence, no gain in
-classification performance can be obtained by adjusting these zero
+decision rule of the \emph{HDRDA} classifier and show that the sum related to
+the null space of the pooled sample covariance matrix can be discarded. Hence,
+no gain in classification performance can be obtained by adjusting these zero
 eigenvalues. This invariance property addresses the computational shortcomings
@@
 
-We also demonstrate that our proposed classifier is competitive with and often
-superior to several recent classifiers designed for small-sample,
-high-dimensional data. In particular, we compare our classifier with diagonal
-linear discriminant analysis proposed by \cite{Dudoit:2002ev} and the recently
-proposed classifiers from \cite{Guo:2007te}, \cite{Pang:2009ik},
+We also demonstrate that the \emph{HDRDA} classifier is competitive with and
+often superior to several recent classifiers designed for small-sample,
+high-dimensional data. In particular, we compare our proposed classifier with
+diagonal linear discriminant analysis proposed by \cite{Dudoit:2002ev} and the
+recently proposed classifiers from \cite{Guo:2007te}, \cite{Pang:2009ik},
 \cite{Tong:2012hw}, and \cite{Witten:2011kc}. We conduct our comparison of the
@@
 value decomposition to calculate efficiently the eigenvalue decomposition of
-$\widehat{\bm \Sigma}$, further reducing the computational costs of our proposed
-classifier.
+$\widehat{\bm \Sigma}$, further reducing the computational costs of the
+\emph{HDRDA} classifier.
 
@@
 
-We now present our main result, where we decompose the decision rule of our
-proposed classifier and show that the term requiring the largest computational
-costs does not contribute to the classification of an unlabeled observation
-performed using Lemma \ref{lemma:RDA-constant-term}. Hence, we reduce
-\eqref{eq:hdrda} to an equivalent, more computationally efficient decision rule.
+We now present our main result, where we decompose the decision rule of the
+\emph{HDRDA} classifier and show that the term requiring the largest
+computational costs does not contribute to the classification of an unlabeled
+observation performed using Lemma \ref{lemma:RDA-constant-term}. Hence, we
+reduce \eqref{eq:hdrda} to an equivalent, more computationally efficient
+decision rule.
 
@@
 
-Thus far, we have shown that our classifier in \eqref{eq:hdrda} is invariant to
-the term $\bm U_2$, thus yielding an equivalent classifier in
+Thus far, we have shown that our proposed classifier in \eqref{eq:hdrda} is
+invariant to the term $\bm U_2$, thus yielding an equivalent classifier in
 \eqref{eq:hdrda-decomposed} with a substantial reduction in the computational
@@
 with the exception that we explicitly set prior probabilities as equal, if the
-option was available. For our classifier, we set $\alpha_k = 1$ $(k = 1, \ldots,
-K)$, so that the covariance-matrix estimator \eqref{eq:hdrda-cov} employed
-resembled \eqref{eq:ridge-estimator}. We estimated $\lambda$ from a grid of 21
-equidistant candidate values between 0 and 1, inclusively. Similarly, we
-estimated $\gamma$ from a grid consisting of the values $10^{-1}, \ldots, 10^4$,
-and $10^5$. We selected optimal estimates of $\lambda$ and $\gamma$ using
+option was available. For the \emph{HDRDA} classifier, we set $\alpha_k = 1$ $(k
+= 1, \ldots, K)$, so that the covariance-matrix estimator \eqref{eq:hdrda-cov}
+employed resembled \eqref{eq:ridge-estimator}. We estimated $\lambda$ from a
+grid of 21 equidistant candidate values between 0 and 1, inclusively. Similarly,
+we estimated $\gamma$ from a grid consisting of the values $10^{-1}, \ldots,
+10^4$, and $10^5$. We selected optimal estimates of $\lambda$ and $\gamma$ using
 10-fold cross-validation \citep{Hastie:2008dt}.
@@
 
-Our classifier was superior in classification performance for the majority of
-the simulations. For the Burczysnki, Shipp, and Singh data sets, our classifier
-outperformed its competitors for all values of $d$ examined. For the Nakayama
-data set, our classifier was outperformed by the Dudoit, Pang, and Tong
-classifiers for small values of $d$ but attained classification rates superior
-to its competitors for larger values of $d$. Furthermore, Figure
+The \emph{HDRDA} classifier was superior in classification performance for the
+majority of the simulations. For the Burczysnki, Shipp, and Singh data sets, our
+classifier outperformed its competitors for all values of $d$ examined. For the
+Nakayama data set, the \emph{HDRDA} classifier was outperformed by the Dudoit,
+Pang, and Tong classifiers for small values of $d$ but attained classification
+rates superior to its competitors for larger values of $d$. Furthermore, Figure
 \ref{fig:error-rates} emphasizes two strengths of our classifier: reliability of
 the classifier for small $d$ and stability of the classifier for large $d$. In
-contrast to the other classifiers, our classifier exhibited less variability in
-its misclassification estimates for small values of $d$ and yielded superior
-classification accuracy as $d$ increased and was hardly effected by the
-inclusion of more variables in the data sets. We attribute the apparent
-stability of the error-rate estimates of our classifier for large values of $d$
-to its novel dimension-reduction technique.
+contrast to the other classifiers, the \emph{HDRDA} classifier exhibited less
+variability in its misclassification estimates for small values of $d$ and
+yielded superior classification accuracy as $d$ increased and was hardly
+effected by the inclusion of more variables in the data sets. We attribute the
+apparent stability of the error-rate estimates of our classifier for large
+values of $d$ to its novel dimension-reduction technique.
 
@@
 \end{figure}
-Next, we examine the optimal estimates of the tuning parameters for our
-classifier. For brevity, we provide graphical summaries of the optimal estimates
-of $\lambda$ and $\gamma$ in the Supplementary Material. For each of the four
-data sets, the distribution of the optimal values of $\lambda$ selected through
-cross-validation often had a notable bathtub shape, resembling a Beta
-distribution with both shape parameters less than unity. Specifically, the
-optimal values of $\lambda$ were typically near 0 or near 1, corresponding to
-quadratic and linear discriminant analysis, respectively. For instance, with
+Next, we examine the optimal estimates of the tuning parameters for the
+\emph{HDRDA} classifier. For brevity, we provide graphical summaries of the
+optimal estimates of $\lambda$ and $\gamma$ in the Supplementary Material. For
+each of the four data sets, the distribution of the optimal values of $\lambda$
+selected through cross-validation often had a notable bathtub shape, resembling
+a Beta distribution with both shape parameters less than unity. Specifically,
+the optimal values of $\lambda$ were typically near 0 or near 1, corresponding
+to quadratic and linear discriminant analysis, respectively. For instance, with
 smaller values of $d$ applied to the Shipp data set, the range of optimal values
@@
 optimal values of $\lambda$ were selected as unity, thereby an increased number
-of variables resulted in our classifier preferring the more parsimonious
-assumption of equal covariance matrices. In this instance we can see the
-flexibility of our classifier to employ the more realistic assumption that the
-class covariance matrices are unequal, but if optimal, we can employ the more
-parsimonious assumption of equal covariance matrices.
+of variables resulted in the \emph{HDRDA} classifier preferring the more
+parsimonious assumption of equal covariance matrices. In this instance we can
+see the flexibility of our proposed classifier to employ the more realistic
+assumption that the class covariance matrices are unequal, but if optimal, we
+can employ the more parsimonious assumption of equal covariance matrices.
 
@@
 
-In our simulation study, we compared our classifier to several recently proposed
-classifiers, including the Dudoit, Pang, Tong, and Guo classifiers, each of
-which assumes independence of the features conditional on class
-membership. Despite the rapid computational performance of diagonal classifiers
-and the reduction in the number of parameters to estimate, \cite{Mai:2012bf} and
-\cite{Fan:2012iq} have noted that diagonal classifiers can often yield inferior
-classification performance to other classification methods. We have shown that
-our proposed classifier often yields superior classification accuracy to the
-diagonal classifiers considered, confirming the assertions of \cite{Mai:2012bf}
-and \cite{Fan:2012iq}. Furthermore, the implementation of our classifier in the
-{\tt sparsediscrim} R package was comparable to that of the diagonal classifiers
-in terms of computational runtime. Moreover, when preliminary variable selection
-was applied to the data in our classification study, the time to perform model
-selection with our proposed classifier was nearly instantaneous.
+In our simulation study, we compared our proposed \emph{HDRDA} classifier to
+several recently proposed classifiers, including the Dudoit, Pang, Tong, and Guo
+classifiers, each of which assumes independence of the features conditional on
+class membership. Despite the rapid computational performance of diagonal
+classifiers and the reduction in the number of parameters to estimate,
+\cite{Mai:2012bf} and \cite{Fan:2012iq} have noted that diagonal classifiers can
+often yield inferior classification performance to other classification
+methods. We have shown that the \emph{HDRDA} classifier often yields superior
+classification accuracy to the diagonal classifiers considered, confirming the
+assertions of \cite{Mai:2012bf} and \cite{Fan:2012iq}. Furthermore, the
+implementation of our classifier in the {\tt sparsediscrim} R package was
+comparable to that of the diagonal classifiers in terms of computational
+runtime. Moreover, when preliminary variable selection was applied to the data
+in our classification study, the time to perform model selection with our
+proposed classifier was nearly instantaneous.
 

### 2013-05-10 reports/paper-hdrda.Rnw
It helps to spell HDRDA right in the paper's filename.

--- 2013-05-10 reports/paper-rda.Rnw
@@
 classifiers after reducing the dimension of the data sets with feature selection
-to $d = 100, \ldots, 1000$, similar to the values of $d$ considered by
+to $d = 100, 200, \ldots, 1000$, similar to the values of $d$ considered by
 \cite{Xu:2009fl}. To compare misclassification rates between different values of
@@
 each value of $d$ in Figure \ref{fig:error-rates}. Approximate standard errors
-were no greater than $\Sexpr{gsub('.', '\\cdot', round(max_se, 3), fixed = TRUE)}$.
+were no greater than \Sexpr{max_se}.
 
@@
   with X), and ours (asterisk). Approximate standard errors are no greater than
-  $\Sexpr{gsub('.', '\\cdot', round(max_se, 3), fixed = TRUE)}$.}
+  \Sexpr{max_se}.}
 \label{fig:error-rates}

### 2013-05-10 reports/paper-hdrda.Rnw
Removed most of the references to Friedman's RDA paper. Also, added the classifier's computational complexity.

@@
 data is essential for successful diagnosis and treatment of diseases such as
-cancer; however, the analysis of microarray data often presents difficult
-challenges because the number of gene expression levels is typically much
-greater than the sample size. Consequently, several sparse and regularized
-classifiers have been proposed, often with strict covariance assumptions or
-eigenvalue adjustments to ensure that the classifiers are well-posed. Here, we
-propose a novel modification of the regularized discriminant analysis classifier
-popularized by \cite{Friedman:1989tm}. Our proposed classifier provides a
-straightforward interpretation and employs a new dimension-reduction technique
-to enable a computationally efficient model-selection procedure, thus overcoming
-the computational shortcomings of \citeauthor{Friedman:1989tm}'s classifier for
-high-dimensional data. Moreover, we demonstrate with four high-dimensional
-microarray data sets that our proposed classifier is competitive with and often
+cancer; however, because the number of gene expression levels measured is
+typically much greater than the sample size, the analysis of microarray data
+often presents difficult statistical and computational challenges. Here, we
+propose a new regularized discriminant analysis classifier that incorporates
+dimension reduction and covariance-matrix shrinkage to enable a computationally
+efficient classifier and model-selection procedure. We demonstrate with four
+high-dimensional microarray data sets that our proposed classifier is often
 superior to several recently proposed sparse and regularized classifiers in
@@
 definiteness. For excellent overviews of the regularization problem in
-discriminant analysis, see \cite{Murphy:2012uq} and \cite{Mkhadri:1997gy}.
+discriminant analysis, see \cite{Murphy:2012uq}, \cite{Mkhadri:1997gy}, and
+\cite{Friedman:1989tm}.
 
-Here, we propose an improvement to the popular regularized discriminant
-analysis classifier of \cite{Friedman:1989tm}, who has proposed pooling of the
-covariance-matrix estimators from linear and quadratic discriminant analysis and
-has additionally incorporated a covariance-matrix shrinkage term to ensure that
-the classifier is well-posed. \citeauthor{Friedman:1989tm}'s classifier can
-attain improved classification accuracy, compared to linear and quadratic
-discriminant analysis when $p < N$. However, in its original form,
-\citeauthor{Friedman:1989tm}'s classifier is impractical when $p \gg N$ because
-the decision rule requires the calculation of the inverses and determinants of
-large $p \times p$ covariance-matrix estimators for each class. Furthermore,
-\citeauthor{Friedman:1989tm}'s classifier has two tuning parameters that are
-selected via cross-validation, which adds a substantial computational burden to
-the model-selection process.
-
-In this paper we modify \citeauthor{Friedman:1989tm}'s classifier by first
-constructing the pooling matrix as a convex combination of the covariance-matrix
-estimators employed in linear and quadratic discriminant analysis and
-demonstrate that this simplification improves the interpretation of the pooling
-operation of \citeauthor{Friedman:1989tm}'s classifier. Given that the
-assumption of equal population covariance matrices for linear discriminant
-analysis is often too stringent in practice \citep{Clemmensen:2011kr}, our
-proposed classifier can relax the linearity assumption by instead incorporating
-unequal covariance matrices while retaining linear and quadratic discriminant
-analysis as special cases. We then establish the high-dimensional regularized
-discriminant analysis (\emph{HDRDA}) classifier by applying
-covariance-matrix regularization with a parameterization that includes a large
-family of shrinkage methods, such as those from \cite{Srivastava:2007ww},
-\cite{Rao:1971ul}, and other methods studied by \cite{Ramey:2013ji} and
-\cite{Xu:2009fl}.
+Here, we propose the high-dimensional regularized discriminant analysis
+(\emph{HDRDA}) by first constructing a pooling matrix as a convex combination of
+the covariance-matrix estimators employed in linear and quadratic discriminant
+analysis. Given that the assumption of equal population covariance matrices for
+linear discriminant analysis is often too stringent in practice
+\citep{Clemmensen:2011kr}, our proposed classifier can relax the linearity
+assumption by instead incorporating unequal covariance matrices while retaining
+linear and quadratic discriminant analysis as special cases. We then establish
+\emph{HDRDA} classifier by applying covariance-matrix regularization with a
+parameterization that includes a large family of shrinkage methods, such as
+those from \cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and other methods
+studied by \cite{Ramey:2013ji} and \cite{Xu:2009fl}. This regularization has the
+effect of stabilizing the eigenvalues of the covariance-matrix estimators and
+ensuring that the classifier is well-posed.
 
@@
 no gain in classification performance can be obtained by adjusting these zero
-eigenvalues. This invariance property addresses the computational shortcomings
-of \citeauthor{Friedman:1989tm}'s classifier and provides a substantial increase
-in computational efficiency and model selection. In fact, we show that the
-inverse and determinants can be reduced to calculations of $n_k \times n_k$
+eigenvalues. This invariance property provides a substantial increase in
+computational efficiency of model training and selection. In fact, we show that
+the inverse and determinants can be reduced to calculations of $n_k \times n_k$
 matrices, where $n_k$ is the sample size of the $k$th class $(k = 1, \ldots,
@@
 where $\xbar_k$ and $\widehat{\bm \Sigma}_k$ are the maximum likelihood
-estimators for $\bm \mu_k$ and $\bm \Sigma_k$, respectively, and $\bm A^{T}$
+estimators of $\bm \mu_k$ and $\bm \Sigma_k$, respectively, and $\bm A^{T}$
 denotes the matrix transpose of $\bm A \in \mathbb{R}_{m \times p}$. If we
@@
 
-\subsection{Friedman's Regularized Discriminant Analysis Classifier}
-
-\cite{Friedman:1989tm} has proposed regularized discriminant analysis by
-incorporating a weighted average of the sample covariance matrix $\widehat{ \bm
-  \Sigma}_k$ and the pooled sample covariance matrix $\widehat{\bm \Sigma}$ to
-estimate $\bm \Sigma_k$ with
-\begin{align}
-  \widehat{\bm \Sigma}_k(\lambda) = n_k^{-1}(\lambda) \bm S_k(\lambda) \quad (k =
-  1, \ldots, K),\label{eq:sig-lambda}
-\end{align}
-where $\lambda \in [0, 1]$, $n_k(\lambda) = (1 - \lambda) n_k + \lambda N$, $
-\bm S_k = n_k \widehat{\bm \Sigma}_k$, $\bm S = \sum_{k=1}^K \bm S_k$, and $ \bm
-S_k(\lambda) = (1 - \lambda) \bm S_k + \lambda \bm S$. We can interpret
-\eqref{eq:sig-lambda} as a covariance-matrix estimator that borrows from
-$\widehat{\bm \Sigma}$ in \eqref{eq:pooled-cov} to better estimate $ \bm
-\Sigma_k$. If $\lambda = 0$, \eqref{eq:sig-lambda} is equal to $\widehat{\bm
-  \Sigma}_k$ used in \eqref{eq:qda}. Alternatively, if $\lambda = 1$,
-\eqref{eq:sig-lambda} reduces to \eqref{eq:pooled-cov}.
-
-To further improve the estimation of $\bm \Sigma_k$ and to stabilize the inverse
-of \eqref{eq:sig-lambda}, \cite{Friedman:1989tm} has proposed the biased
-covariance-matrix estimator
-\begin{align}
-	\widehat{\bm \Sigma}_k(\lambda, \gamma) = (1 - \gamma) \widehat{\bm \Sigma}_k(\lambda) + \gamma \frac{\tr\{\widehat{\bm \Sigma}_k(\lambda)\}}{p} \bm I_p,\label{eq:sig-rda}
-\end{align}
-where $\gamma \in [0, 1]$ is a regularization parameter that controls the
-shrinkage of \eqref{eq:sig-rda} towards $\bm I_p$ and is weighted by the average
-of the eigenvalues of \eqref{eq:sig-lambda}. Thus, the \emph{pooling} parameter
-$\lambda$ controls the amount borrowed from $\widehat{\bm \Sigma}$ to estimate
-$\bm \Sigma_k$, and the \emph{shrinkage} parameter $\gamma$ determines the
-amount of eigenvalue shrinkage applied. Finally, by substituting
-\eqref{eq:sig-rda} into \eqref{eq:qda}, we have the regularized discriminant
-analysis classifier
-\begin{align}
-	D_{RDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)^{T}\widehat{\bm \Sigma}_k(\lambda, \gamma)^{-1}(\bm x - \xbar_k)  + \log |\widehat{\bm \Sigma}_k(\lambda, \gamma)|. \label{eq:rda}
-\end{align}
-If $\gamma$ is sufficiently small, \eqref{eq:sig-rda} can be singular or
-near-singular. For this case \cite{Friedman:1989tm} has suggested replacing the
-zero eigenvalues of \eqref{eq:sig-rda} with a small positive constant that
-permits a numerically stable matrix inversion.
-
 \section{High-Dimensional Regularized Discriminant Analysis}
@@
 
-Here, we define our adaptation of \citeauthor{Friedman:1989tm}'s classifier by
-first simplifying the formulation of $\widehat{\bm \Sigma}_k(\lambda)$ in
-\eqref{eq:sig-lambda} and demonstrating its clear interpretation as a linear
-combination of the crossproducts of the training observations centered by their
-respective class sample means. We define the convex combination
+Here, we define the \emph{HDRDA} classifier by first formulating the
+covariance-matrix estimator $\widehat{\bm \Sigma}_k(\lambda)$ and demonstrating
+its clear interpretation as a linear combination of the crossproducts of the
+training observations centered by their respective class sample means. We define
+the convex combination
 \begin{align}
@@
 \end{align}
-where $\lambda \in [0, 1]$ is the \emph{pooling} parameter. Thus, we avoid the
-convex combination of $n_k$ and $N$ as defined by \cite{Friedman:1989tm}. By
-rewriting \eqref{eq:sig-lambda-alternative} in terms of the observations $
-\bm x_i$ $(i = 1, \ldots, N)$, each centered by its class sample mean, we attain a
-clear interpretation of $\widehat{\bm \Sigma}_k(\lambda)$. That is,
+where $\lambda \in [0, 1]$ is the \emph{pooling} parameter. By rewriting
+\eqref{eq:sig-lambda-alternative} in terms of the observations $ \bm x_i$ $(i =
+1, \ldots, N)$, each centered by its class sample mean, we attain a clear
+interpretation of $\widehat{\bm \Sigma}_k(\lambda)$. That is,
 \begin{align}
@@
 \end{align}
-where $c_{ik}(\lambda) = \lambda N^{-1} + (1 - \lambda)n_k^{-1}I(y_i = k)$.
-From \eqref{eq:sig-lambda-alternative2}, $\lambda$ weights the contribution of
-each of the $N$ observations in estimating $\bm \Sigma_k$. Hence, for $0 <
-\lambda \le 1$, we estimate $\bm \Sigma_k$ with $N$ rather than $n_k$
-observations. Below, we will show that the pooling operation is advantageous in
-increasing the rank of each $\widehat{\bm \Sigma}_k(\lambda)$ from
-rank$(\widehat{\bm \Sigma}_k)$ to rank$(\widehat{\bm \Sigma})$ for $0 < \lambda
-\le 1$. Notice that if $\lambda = 0$, then the observations from the remaining
-$K - 1$ classes do not contribute to the estimation of $\bm \Sigma_k$,
-corresponding to $\widehat{\bm \Sigma}_k$. Furthermore, if $\lambda = 1$, the
-weights in \eqref{eq:sig-lambda-alternative2} reduce to $1/N$, corresponding to
+where $c_{ik}(\lambda) = \lambda N^{-1} + (1 - \lambda)n_k^{-1}I(y_i = k)$. We
+can interpret \eqref{eq:sig-lambda-alternative2} as a covariance-matrix
+estimator that borrows from $\widehat{\bm \Sigma}$ in \eqref{eq:pooled-cov} to
+better estimate $ \bm \Sigma_k$. From \eqref{eq:sig-lambda-alternative2},
+$\lambda$ weights the contribution of each of the $N$ observations in estimating
+$\bm \Sigma_k$. Hence, for $0 < \lambda \le 1$, we estimate $\bm \Sigma_k$ with
+$N$ rather than $n_k$ observations. Below, we will show that the pooling
+operation is advantageous in increasing the rank of each $\widehat{\bm
+  \Sigma}_k(\lambda)$ from rank$(\widehat{\bm \Sigma}_k)$ to rank$(\widehat{\bm
+  \Sigma})$ for $0 < \lambda \le 1$. Notice that if $\lambda = 0$, then the
+observations from the remaining $K - 1$ classes do not contribute to the
+estimation of $\bm \Sigma_k$, corresponding to $\widehat{\bm
+  \Sigma}_k$. Furthermore, if $\lambda = 1$, the weights in
+\eqref{eq:sig-lambda-alternative2} reduce to $1/N$, corresponding to
 $\widehat{\bm \Sigma}$. For brevity, when $\lambda = 1$, we define $\bm X =
-[\sqrt{c_{1k}(1)} \bm x_1^{T}, \ldots, \sqrt{c_{Nk}(1)} \bm x_N^{T}]^{T}$ such
-that $\widehat{\bm \Sigma} = \bm X^{T} \bm X$. Similarly, for $\lambda = 0$, we
-define $\bm X_k = [\sqrt{c_{1k}(0)} \bm x_1^{T}, \ldots, \sqrt{c_{Nk}(0)} \bm
-  x_N^{T}]^{T}$ such that $\widehat{\bm \Sigma}_k = \bm X_k^{T} \bm X_k$. In
-Figure \ref{fig:hdrda-contours} we plot the contours of five multivariate normal
-populations for $\lambda = 0$ with unequal covariance matrices. As $\lambda$
-approaches 1, the contours become more similar, resulting in identical contours
-for $\lambda = 1$.
+      [\sqrt{c_{1k}(1)} \bm x_1^{T}, \ldots, \sqrt{c_{Nk}(1)} \bm x_N^{T}]^{T}$
+      such that $\widehat{\bm \Sigma} = \bm X^{T} \bm X$. Similarly, for
+      $\lambda = 0$, we define $\bm X_k = [\sqrt{c_{1k}(0)} \bm x_1^{T}, \ldots,
+        \sqrt{c_{Nk}(0)} \bm x_N^{T}]^{T}$ such that $\widehat{\bm \Sigma}_k =
+      \bm X_k^{T} \bm X_k$. In Figure \ref{fig:hdrda-contours} we plot the
+      contours of five multivariate normal populations for $\lambda = 0$ with
+      unequal covariance matrices. As $\lambda$ approaches 1, the contours
+      become more similar, resulting in identical contours for $\lambda = 1$.
 \begin{figure}
@@
 As we have discussed above, several eigenvalue adjustment methods have been
-proposed to increase eigenvalues (approximately) equal to 0. Here, we define the
-eigenvalue adjustment of \eqref{eq:sig-lambda-alternative} as
+proposed to increase eigenvalues (approximately) equal to 0. To further improve
+the estimation of $\bm \Sigma_k$ and to stabilize the estimator's inverse, we
+define the eigenvalue adjustment of \eqref{eq:sig-lambda-alternative} as
 \begin{align}
@@
 where $\alpha_k \ge 0$ and $\gamma \ge 0$ is an eigenvalue-shrinkage
-constant. The choice of $\alpha_k$ allows for a flexible formulation of
+constant. Thus, the \emph{pooling} parameter $\lambda$ controls the amount
+borrowed from $\widehat{\bm \Sigma}$ to estimate $\bm \Sigma_k$, and the
+\emph{shrinkage} parameter $\gamma$ determines the amount of eigenvalue
+shrinkage applied. The choice of $\alpha_k$ allows for a flexible formulation of
 covariance-matrix estimators. For instance, if $\alpha_k = 1$, $k = 1, \ldots,
 K$, then \eqref{eq:hdrda-cov} resembles \eqref{eq:ridge-estimator}. Similarly,
-if $\alpha_k = 1 - \gamma$, then \eqref{eq:hdrda-cov} has a form similar to
-\eqref{eq:sig-rda}. Now, substituting \eqref{eq:hdrda-cov} into \eqref{eq:qda},
-we define the high-dimensional regularized discriminant analysis classifier
+if $\alpha_k = 1 - \gamma$, then \eqref{eq:hdrda-cov} has a form similar to the
+\emph{RDA} classifier from \cite{Friedman:1989tm}. Now, substituting
+\eqref{eq:hdrda-cov} into \eqref{eq:qda}, we define the high-dimensional
+regularized discriminant analysis classifier
 \begin{align}
@@
 \subsection{Enhanced Computational Efficiency}
+\label{sec:computational-efficiency}
 
@@
 matrices, and by ignoring these calculations, a substantial reduction in
-computational costs is achieved relative to Friedman's classifier in
-\eqref{eq:rda}. Furthermore, a byproduct is that adjustments to the associated
-eigenvalues have no effect on \eqref{eq:hdrda}. Lastly, we utilize the singular
-value decomposition to calculate efficiently the eigenvalue decomposition of
-$\widehat{\bm \Sigma}$, further reducing the computational costs of the
-\emph{HDRDA} classifier.
+computational costs is achieved. Furthermore, a byproduct is that adjustments to
+the associated eigenvalues have no effect on \eqref{eq:hdrda}. Lastly, we
+utilize the singular value decomposition to calculate efficiently the eigenvalue
+decomposition of $\widehat{\bm \Sigma}$, further reducing the computational
+costs of the \emph{HDRDA} classifier.
 
@@
   n_k}$, respectively, as well as the diagonal matrix $\bm \Gamma_k \in
-\mathbb{R}_{q \times q}$. This formulation is especially advantageous when the
-number of classes $K$ is large because $n_k$ is then potentially much smaller
-than $q$. The expressions given in Proposition \ref{proposition:hdrda-W_k} also
-expedite the selection of $\lambda$ and $\gamma$ via cross-validation. The most
-time-consuming matrix operations involved in computing $\bm W_k^{-1}$ and $|\bm
-W_k|$ is $\bm X_k \bm U_1 \in \mathbb{R}_{n_k \times q}$, which is independent
-of $\lambda$ and $\gamma$. Hence, we need calculate $\bm X_k \bm U_1$ only once
-for each cross-validation fold. The subsequent operations in calculating $\bm
-W_k^{-1}$ and $|\bm W_K|$ can be simply updated for different pairs of $\lambda$
-and $\gamma$.
+\mathbb{R}_{q \times q}$. Hence, the resulting formulation requires $O(n_k p q)$
+calculations to compute both $\bm W_k^{-1}$ and $| \bm W_k |$ and is especially
+advantageous when the number of classes $K$ is large because $n_k$ is then
+potentially much smaller than $q$. The expressions given in Proposition
+\ref{proposition:hdrda-W_k} also expedite the selection of $\lambda$ and
+$\gamma$ via cross-validation because the most time-consuming matrix operation
+involved in computing $\bm W_k^{-1}$ and $|\bm W_k|$ is $\bm X_k \bm U_1 \in
+\mathbb{R}_{n_k \times q}$, which is independent of $\lambda$ and
+$\gamma$. Therefore, we need calculate $\bm X_k \bm U_1$ only once for each
+cross-validation fold. The subsequent operations in calculating $\bm W_k^{-1}$
+and $|\bm W_K|$ can be simply updated for different pairs of $\lambda$ and
+$\gamma$.
 
-Thus far, we have constructed each $\tilde{\bm \Sigma}_k$ in terms of the first $q$
-eigenvectors of $\widehat{\bm \Sigma}$. By utilizing a technique similar to the
-so-called kernel trick, we can obtain $\bm U_1$ by computing the eigendecomposition
-of a much smaller $N \times N$ matrix when $p \gg N$
+Thus far, we have constructed each $\tilde{\bm \Sigma}_k$ in terms of the first
+$q$ eigenvectors of $\widehat{\bm \Sigma}$. By utilizing a technique similar to
+the so-called kernel trick, we can obtain $\bm U_1$ by computing the
+eigendecomposition of a much smaller $N \times N$ matrix when $p \gg N$
 \citep{Hastie:2008dt}. Recall that the pooled sample covariance matrix is
-$\widehat{\bm \Sigma} = \bm X^{T} \bm X$. Then, because $\bm X = \bm M \bm D^{1/2} \bm U^{T}$ by the
-singular value decomposition, we have the eigendecomposition $\widehat{\bm \Sigma} =
-\bm U \bm D \bm U^{T}$, where $\bm M \in \mathbb{R}_{N \times N}$ is orthogonal, $\bm D^{1/2} \in
-\mathbb{R}_{N \times N}^{\ge}$ is a diagonal matrix consisting of the singular
-values of $\bm X$, and $\bm U \in \mathbb{R}_{p \times N}$ is orthogonal. Now, $\bm X \bm X^{T}
-= \bm M \bm D \bm M^{T}$ is the eigenvalue decomposition of $\bm X \bm X^{T}$, where $\bm D$ is the
-matrix of eigenvalues of $\bm X \bm X^{T}$ and the columns of $\bm M$ are the corresponding
-eigenvectors. Hence, we obtain $\bm M$ and $\bm D$ from an eigendecomposition of $\bm X
-\bm X^{T} \in \mathbb{R}_{N \times N}$. Next, we compute $\bm U = \bm X^{T} \bm M \bm D^{+/2}$,
-where $\bm D^{+/2} = \bm D_q^{-1/2} \oplus \bm 0_{N-q}$. We then determine $q$, the number
-of numerically nonzero eigenvalues present in $\bm D$, by calculating the number of
-eigenvalues that exceeds some tolerance value, say, $\epsilon = 1 \times
-10^{-6}$. After calculating $q$, we extract $\bm U_1$ by retaining only the first
-$q$ columns of $\bm U$.
+$\widehat{\bm \Sigma} = \bm X^{T} \bm X$. Then, because $\bm X = \bm M \bm
+D^{1/2} \bm U^{T}$ by the singular value decomposition, we have the
+eigendecomposition $\widehat{\bm \Sigma} = \bm U \bm D \bm U^{T}$, where $\bm M
+\in \mathbb{R}_{N \times N}$ is orthogonal, $\bm D^{1/2} \in \mathbb{R}_{N
+  \times N}^{\ge}$ is a diagonal matrix consisting of the singular values of
+$\bm X$, and $\bm U \in \mathbb{R}_{p \times N}$ is orthogonal. Now, $\bm X \bm
+X^{T} = \bm M \bm D \bm M^{T}$ is the eigenvalue decomposition of $\bm X \bm
+X^{T}$, where $\bm D$ is the matrix of eigenvalues of $\bm X \bm X^{T}$ and the
+columns of $\bm M$ are the corresponding eigenvectors. Hence, we obtain $\bm M$
+and $\bm D$ from an eigendecomposition of $\bm X \bm X^{T} \in \mathbb{R}_{N
+  \times N}$. Next, we compute $\bm U = \bm X^{T} \bm M \bm D^{+/2}$, where $\bm
+D^{+/2} = \bm D_q^{-1/2} \oplus \bm 0_{N-q}$. We then determine $q$, the number
+of numerically nonzero eigenvalues present in $\bm D$, by calculating the number
+of eigenvalues that exceeds some tolerance value, say, $\epsilon = 1 \times
+10^{-6}$. After calculating $q$, we extract $\bm U_1$ by retaining only the
+first $q$ columns of $\bm U$.
 
@@
 error-rate estimates for each classifier. We remark that our simulation design
-avoids the selection bias discussed in \cite{Ambroise:2002fa}. We did not
-include \citeauthor{Friedman:1989tm}'s classifier in our studies to avoid its
-excessive computations.
+avoids the selection bias discussed in \cite{Ambroise:2002fa}.
 
@@
 in our classification study, the time to perform model selection with our
-proposed classifier was nearly instantaneous.
+proposed classifier was nearly instantaneous. Hence, our proposed \emph{HDRDA}
+classifier is competitive in terms of classification accuracy for modern
+small-sample, high-dimensional data sets and requires substantially less
+computing time than standard regularized classifiers, such as regularized
+discriminant analysis proposed by \cite{Friedman:1989tm}.
 

### 2013-05-11 reports/paper-hdrda.Rnw
Improved the abstract.

@@
 cancer; however, because the number of gene expression levels measured is
-typically much greater than the sample size, the analysis of microarray data
-often presents difficult statistical and computational challenges. Here, we
-propose a new regularized discriminant analysis classifier that incorporates
-dimension reduction and covariance-matrix shrinkage to enable a computationally
-efficient classifier and model-selection procedure. We demonstrate with four
-high-dimensional microarray data sets that our proposed classifier is often
-superior to several recently proposed sparse and regularized classifiers in
-terms of classification accuracy. Finally, we provide an implementation of our
-proposed classifier in the {\tt sparsediscrim} R package, available on CRAN.
+typically much greater than the sample size, the analysis of microarray data can
+be difficult because classic classifiers are often ill-posed in this setting and
+require a large number of calculations. To address these challenges, we propose
+a new regularized discriminant analysis classifier that incorporates dimension
+reduction and covariance-matrix shrinkage to enable an efficacious,
+computationally efficient method. We demonstrate with four high-dimensional
+microarray data sets that our proposed classifier is often superior to several
+recently proposed sparse and regularized classifiers in terms of classification
+accuracy. Finally, we provide an implementation of our proposed classifier in
+the {\tt sparsediscrim} R package, available on CRAN.
 
@@
 \eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance
-matrix $\widehat{\bm \Sigma}_k$ towards $\bm I_p$, thereby increasing the
+matrix $\widehat{\bm \Sigma}_k$ toward $\bm I_p$, thereby increasing the
 eigenvalues of $\widehat{\bm \Sigma}_k$ by $\gamma$. Specifically, the zero

### 2013-05-11 reports/paper-hdrda.Rnw
Minor edits before customizing paper for AoAS.

@@
 classification of tissues and other biological samples into one of $K$ disjoint,
-predefined cancer types, where our goal is to ascertain the class membership of
+predefined types, where our goal is to ascertain the class membership of
 a new tissue sample based on gene-expression microarray data. However,
@@
 linear and quadratic discriminant analysis as special cases. We then establish
-\emph{HDRDA} classifier by applying covariance-matrix regularization with a
+the \emph{HDRDA} classifier by applying covariance-matrix regularization with a
 parameterization that includes a large family of shrinkage methods, such as
@@
 computational efficiency of model training and selection. In fact, we show that
-the inverse and determinants can be reduced to calculations of $n_k \times n_k$
+the inverse and determinant can be reduced to calculations of $n_k \times n_k$
 matrices, where $n_k$ is the sample size of the $k$th class $(k = 1, \ldots,
 K)$. This reduction in calculation is clearly advantageous because $n_k$ is
-often quite small in microarray data sets (e.g., 15--20 observations).
+often quite small in microarray data sets (e.g., 10--20 observations).
 
@@
 
-The smallest eigenvalues and the directions associated with their eigenvectors
-highly influence \eqref{eq:qda}. In fact, the eigenvalues of $\widehat{ \bm
-  \Sigma}_k$ are well-known to be biased if $p \ge n_k$ such that the smallest
-eigenvalues are underestimated \citep{Seber:2004uh}, Moreover, if $p > n_k$,
-then rank$(\widehat{\bm \Sigma}_k) \le n_k$, which implies that at least $p -
-n_k$ eigenvalues of $\widehat{\bm \Sigma}_k$ are 0. Furthermore, although more
-feature information is available to discriminate among the $K$ classes, if $p >
-n_k$, \eqref{eq:qda} is incalculable because $\widehat{\bm \Sigma}_k^{-1}$ does
-not exist.
+The smallest eigenvalues of $\widehat{\bm \Sigma}_k$ and the directions
+associated with their eigenvectors highly influence \eqref{eq:qda}. In fact, the
+eigenvalues of $\widehat{ \bm \Sigma}_k$ are well-known to be biased if $p \ge
+n_k$ such that the smallest eigenvalues are underestimated \citep{Seber:2004uh},
+Moreover, if $p > n_k$, then rank$(\widehat{\bm \Sigma}_k) \le n_k$, which
+implies that at least $p - n_k$ eigenvalues of $\widehat{\bm \Sigma}_k$ are
+0. Furthermore, although more feature information is available to discriminate
+among the $K$ classes, if $p > n_k$, \eqref{eq:qda} is incalculable because
+$\widehat{\bm \Sigma}_k^{-1}$ does not exist.
 
@@
 \emph{RDA} classifier from \cite{Friedman:1989tm}. Now, substituting
-\eqref{eq:hdrda-cov} into \eqref{eq:qda}, we define the high-dimensional
-regularized discriminant analysis classifier
+\eqref{eq:hdrda-cov} into \eqref{eq:qda}, we define the \emph{HDRDA} classifier
 \begin{align}
@@
 complexity. Here, we demonstrate that the computational efficiency in
-calculating the inverse and determinants of $\bm W_k$ can be further improved
+calculating the inverse and determinant of $\bm W_k$ can be further improved
 via standard matrix operations.
@@
 We evaluated the classification performance of each classifier by randomly
-partitioning the data such that $2/3$ of the observations were allocated as
-training data and the remaining $1/3$ of the observations were allocated as a
-test data set. For each random partition, we applied the feature-selection
-method from \cite{Dudoit:2002ev} to the training data to choose the $d$ features
-having the largest ratio of their between-group to within-group sums of squares.
-We then filtered the test data set with the corresponding $d$ features. After
-training each classifier from the filtered training data, we classified the
-filtered test data sets and computed the proportion of mislabeled test
-observations to estimate the classification error rate for each
-classifier. Repeating this process 1000 times, we computed the average of the
-error-rate estimates for each classifier. We remark that our simulation design
-avoids the selection bias discussed in \cite{Ambroise:2002fa}.
+partitioning the data set under consideration such that $2/3$ of the
+observations were allocated as training data and the remaining $1/3$ of the
+observations were allocated as a test data set. For each random partition, we
+applied the feature-selection method from \cite{Dudoit:2002ev} to the training
+data to choose the $d$ features having the largest ratio of their between-group
+to within-group sums of squares.  We then filtered the test data set with the
+corresponding $d$ features. After training each classifier from the filtered
+training data, we classified the filtered test data sets and computed the
+proportion of mislabeled test observations to estimate the classification error
+rate for each classifier. Repeating this process 1000 times, we computed the
+average of the error-rate estimates for each classifier. We remark that our
+simulation design avoids the selection bias discussed in \cite{Ambroise:2002fa}.
 
@@
 training and test data. Hence, for a given random partition, the set of features
-selected for a small value of $d$ is a proper subset of the features selected
+selected for a smaller value of $d$ is a proper subset of the features selected
 for a larger value of $d$. We report the average of the error-rate estimates for
@@
 $d$, thereby ignoring the inclusion of a larger number of variables. Hence, we
-advise that practioners be aware of the necessity to filter a large number of
-features before employing these methods, lest their classification performance
-be degraded.
+advise practioners to consider filtering a large number of features before
+employing these methods.
 
@@
 small-sample, high-dimensional data sets and requires substantially less
-computing time than standard regularized classifiers, such as regularized
-discriminant analysis proposed by \cite{Friedman:1989tm}.
+computing time than standard regularized classifiers, such as the \emph{RDA}
+classifier proposed by \cite{Friedman:1989tm}.
 

### 2013-05-11 reports/paper-hdrda.Rnw
Updated front matter to match AoAS format.

--- 2013-05-10 reports/paper-hdrda.Rnw
@@
-\documentclass[11pt]{article}
-\usepackage{graphicx, amsmath, amssymb, bm, url, mathtools, natbib, amsthm, setspace}
+\documentclass[aoas,preprint]{imsart}
 
-\pagestyle{plain}
-%----------------Page dimensions ----------------
-\oddsidemargin 0.0in
-\evensidemargin 0.0in
-\topmargin -0.75in
-\leftmargin 0in
-\headheight 0.0in
-\headsep 0.5in
-%\footheight 0.0in
-\footskip 0.5in
-\footnotesep 0.0in
-\textwidth 6.7in
-\textheight 9.5in
-%-----------Define Pictures---------------------
-\def\picture #1 by #2 (#3){
-   \vbox to #2{
-        \hrule width #1 height 0pt depth 0pt
-           \vfill
-              \special{picture #3} % this is the low-level interface
-                 }
-    }
-\def\scaledpicture #1 by #2 (#3 scaled #4){{
-     \dimen0=#1 \dimen1=#2
-      \divide\dimen0 by 1000 \multiply\dimen0 by #4
-       \divide\dimen1 by 1000 \multiply\dimen1 by #4
-        \picture \dimen0 by \dimen1 (#3 scaled #4)}
-   }
+\usepackage{graphicx, amsmath, amssymb, bm, mathtools, amsthm}
+\RequirePackage[OT1]{fontenc}
+\RequirePackage{natbib}
+\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
 
-\title{High-Dimensional Regularized Discriminant Analysis}
+% provide arXiv number if available:
+%\arxiv{arXiv:0000.0000}
 
-\author{John A. Ramey, Caleb K. Stein, and Dean M. Young}
-%Vaccine and Infectious Disease Division, Fred Hutchinson Cancer Research
-%  Center,\\ Seattle, Washington 98115, U.S.A. \email{johnramey@gmail.com}}
-%\affil{Department of Statistical Science, Baylor University, Waco, Texas 76706, U.S.A. \email{Caleb\_Stein@baylor.edu} \email{Dean\_Young@baylor.edu}}
-
+\startlocaldefs
 \newcommand{\xbar}{\bar{\bm x}}
@@
 \DeclareMathOperator*{\argmax}{arg\,max}
-
 \newtheorem{cor}{Corollary}
@@
 \newtheorem{thm}{Theorem}
-
-\bibpunct{(}{)}{;}{a}{,}{,}
+\endlocaldefs
 
@@
 
-\doublespacing
+\begin{frontmatter}
 
-\maketitle
+\title{High-Dimensional Regularized Discriminant Analysis}
+\runtitle{High-Dimensional Regularized Discriminant Analysis}
+
+\begin{aug}
+\author{\fnms{John A.} \snm{Ramey}\thanksref{m1}\ead[label=e1]{johnramey@gmail.com}}
+\author{\fnms{Caleb K.} \snm{Stein}\thanksref{m2}\ead[label=e2]{Caleb\_Stein@baylor.edu}}
+\and
+\author{\fnms{Dean M.} \snm{Young}\thanksref{m2}\ead[label=e3]{Dean\_Young@baylor.edu}}
+
+\runauthor{J. A. Ramey et al.}
+
+\affiliation{Fred Hutchinson Cancer Research Center\thanksmark{m1} and Baylor University\thanksmark{m2}}
+
+\address{Vaccine and Infectious Disease Division\\
+  Fred Hutchinson Cancer Research Center\\
+  1100 Fairview Ave. N., M2-C200\\
+  Seattle, Washington 98115, U.S.A.\\
+  \printead{e1}}
+
+\address{Department of Statistical Science\\
+  Baylor University\\
+  P.O. Box 97140\\
+  Waco, Texas 76706, U.S.A.\\
+  \printead{e2}\\
+  \phantom{E-mail:\ }\printead*{e3}}
+\end{aug}
+
+\begin{abstract}
+
+A reliable supervised classification of tumors from gene expression microarray
+data is essential for successful diagnosis and treatment of diseases such as
+cancer; however, because the number of gene expression levels measured is
+typically much greater than the sample size, the analysis of microarray data
+often presents difficult statistical and computational challenges. Here, we
+propose a new regularized discriminant analysis classifier that incorporates
+dimension reduction and covariance-matrix shrinkage to enable a computationally
+efficient classifier and model-selection procedure. We demonstrate with four
+high-dimensional microarray data sets that our proposed classifier is often
+superior to several recently proposed sparse and regularized classifiers in
+terms of classification accuracy. Finally, we provide an implementation of our
+proposed classifier in the {\tt sparsediscrim} R package, available on CRAN.
+
+\end{abstract}
+
+%\begin{keyword}[class=AMS]
+%\kwd[Primary ]{}
+%\kwd{}
+%\kwd[; secondary ]{}
+%\end{keyword}
+
+%\begin{keyword}
+%\kwd{}
+%\kwd{}
+%\end{keyword}
+
+\end{frontmatter}
+
+% AOS,AOAS: If there are supplements please fill:
+%\begin{supplement}[id=suppA]
+%  \sname{Supplement A}
+%  \stitle{Title}
+%  \slink[doi]{10.1214/00-AOASXXXXSUPP}
+%  \sdatatype{.pdf}" 
+%  \sdescription{Some text}
+%\end{supplement}
 
@@
 @
-
-\maketitle
-
-\begin{abstract}
-
-A reliable supervised classification of tumors from gene expression microarray
-data is essential for successful diagnosis and treatment of diseases such as
-cancer; however, because the number of gene expression levels measured is
-typically much greater than the sample size, the analysis of microarray data
-often presents difficult statistical and computational challenges. Here, we
-propose a new regularized discriminant analysis classifier that incorporates
-dimension reduction and covariance-matrix shrinkage to enable a computationally
-efficient classifier and model-selection procedure. We demonstrate with four
-high-dimensional microarray data sets that our proposed classifier is often
-superior to several recently proposed sparse and regularized classifiers in
-terms of classification accuracy. Finally, we provide an implementation of our
-proposed classifier in the {\tt sparsediscrim} R package, available on CRAN.
-
-\end{abstract}
 
@@
 
-\bibliographystyle{plainnat}
+\bibliographystyle{imsart-nameyear}
 \bibliography{rda}

### 2013-05-12 reports/paper-hdrda.Rnw
Updated figures to match AoAS requirements.

@@
 \begin{figure}
-<<hdrda_contours, fig.width=8, fig.height=8, out.width='.49\\linewidth', fig.show='hold'>>=
+<<hdrda_contours, fig.width=8, fig.height=8, out.width='.33\\linewidth', fig.show='hold'>>=
 rda_contours(lambda = 0)
@@
 \begin{figure}
-<<error_rates>>=
+<<error_rates, fig.width=8, fig.height=8, out.width='.65\\linewidth'>>=
 error_rates$data_set <- Hmisc:::capitalize(error_rates$data_set)
 p <- ggplot(error_rates, aes(x = d, y = Avg_Error, group = Classifier))
-p <- p + geom_line(aes(), size = 1, linetype = 2)
-p <- p + geom_point(aes(shape = Classifier), size = 6)
+p <- p + geom_line(aes(color = Classifier), size = 1, linetype = 2)
+p <- p + geom_point(aes(color = Classifier, shape = Classifier), size = 4)
 p <- p + facet_wrap(~ data_set) + ylim(c(0, max(error_rates$Avg_Error))) + theme_bw()
 p <- p + xlab("Number of Variables Selected, d") + ylab("Average Error Rate")
-p <- p + theme(axis.text = element_text(size = 16))
-p <- p + theme(axis.title.x = element_text(size = 20))
-p <- p + theme(axis.title.y = element_text(size = 20))
-p <- p + theme(strip.text.x = element_text(size = 18))
-p + theme(legend.position = "none")
+p <- p + theme(axis.text = element_text(size = 10))
+p <- p + theme(axis.title.x = element_text(size = 17))
+p <- p + theme(axis.title.y = element_text(size = 17))
+p <- p + theme(strip.text.x = element_text(size = 16))
+p <- p + theme(legend.title = element_text(size = 18, face = "bold"))
+p <- p + theme(legend.text = element_text(size = 16))
+p <- p + theme(legend.key.size = unit(1, "cm"))
+p
 @ 
 \caption{Averages of the 1000 error-rate estimates of the competing classifiers
-  for each data set as a function of $d$. The classifiers are marked as follows:
-  Dudoit (circle), Guo (triangle), Pang (square), Tong = (cross), Witten (square
-  with X), and ours (asterisk). Approximate standard errors are no greater than
-  \Sexpr{max_se}.}
+  for each data set as a function of $d$. Approximate standard errors are no
+  greater than \Sexpr{max_se}.}
 \label{fig:error-rates}

### 2013-05-12 lib/helpers.R

@@ -120,8 +120,8 @@ rda_contours <- function(lambda) {
   contours$Population <- factor(contours$Population)
   contours$lambda <- lambda
 
-  p <- ggplot(contours, aes(x = X1, y = X2, linetype = Population))
-  p <- p + geom_path(size = 1.5)
+  p <- ggplot(contours, aes(x = X1, y = X2, color = Population, linetype = Population))
+  p <- p + geom_path(size = 2)
   p <- p + facet_grid(. ~ lambda, labeller = label_bquote(lambda == .(x)))
   p <- p + theme_bw() + xlab("") + ylab("") + theme(legend.position = "none")
   p <- p + scale_x_continuous(breaks = NULL) + scale_y_continuous(breaks = NULL)

### 2013-05-12 reports/paper-hdrda.Rnw
Added supplementary material and keywords.

@@
 \begin{aug}
-\author{\fnms{John A.} \snm{Ramey}\thanksref{m1}\ead[label=e1]{johnramey@gmail.com}}
+\author{\fnms{John A.} \snm{Ramey}\thanksref{m1}
+  \ead[label=e1]{johnramey@gmail.com}
+  \ead[label=u1,url]{http://ramhiser.com}}
 \author{\fnms{Caleb K.} \snm{Stein}\thanksref{m2}\ead[label=e2]{Caleb\_Stein@baylor.edu}}
@@
   Seattle, Washington 98115, U.S.A.\\
-  \printead{e1}}
+  \printead{e1}\\
+  \printead{u1}}
 
@@
 
-%\begin{keyword}[class=AMS]
-%\kwd[Primary ]{}
-%\kwd{}
-%\kwd[; secondary ]{}
-%\end{keyword}
-
-%\begin{keyword}
-%\kwd{}
-%\kwd{}
-%\end{keyword}
+\begin{keyword}
+\kwd{classification}
+\kwd{regularized discriminant analysis}
+\kwd{covariance-matrix regularization}
+\kwd{high-dimensional microarray data}
+\kwd{dimension reduction}
+\end{keyword}
 
 \end{frontmatter}
-
-% AOS,AOAS: If there are supplements please fill:
-%\begin{supplement}[id=suppA]
-%  \sname{Supplement A}
-%  \stitle{Title}
-%  \slink[doi]{10.1214/00-AOASXXXXSUPP}
-%  \sdatatype{.pdf}" 
-%  \sdescription{Some text}
-%\end{supplement}
 
@@
 error_rates <- subset(error_rates, select = -Num_Reps)
+error_rates$Classifier <- factor(as.character(error_rates$Classifier))
 
@@
 \cite{Ramey:2013ji}, \cite{Xu:2009fl}, and \cite{Ye:2009gd}.
+
+\clearpage
 
@@
 10^4$, and $10^5$. We selected optimal estimates of $\lambda$ and $\gamma$ using
-10-fold cross-validation \citep{Hastie:2008dt}.
+10-fold cross-validation \citep{Hastie:2008dt}. Each of the microarray data sets
+listed below can be found in \ref{suppA}.
 
@@
 p <- p + theme(strip.text.x = element_text(size = 16))
-p <- p + theme(legend.title = element_text(size = 18, face = "bold"))
-p <- p + theme(legend.text = element_text(size = 16))
+p <- p + theme(legend.title = element_text(size = 16, face = "bold"))
+p <- p + theme(legend.text = element_text(size = 15))
 p <- p + theme(legend.key.size = unit(1, "cm"))
@@
 \emph{HDRDA} classifier. For brevity, we provide graphical summaries of the
-optimal estimates of $\lambda$ and $\gamma$ in the Supplementary Material. For
+optimal estimates of $\lambda$ and $\gamma$ in \ref{suppB}. For
 each of the four data sets, the distribution of the optimal values of $\lambda$
@@
 
-\section*{Supplementary material}
-\label{SM}
-Supplementary material available online includes the graphical summary
-of the optimal estimates of our classifier from the classification study.
+\begin{supplement}
+\sname{Supplement A}\label{suppA}
+\stitle{R Package Containing Microarray Data Sets}
+\slink[url]{https://github.com/ramey/datamicroarray}
+\sdescription{An R package containing all of the microarray data sets used in
+  the classification study.}
+\end{supplement}
+
+\begin{supplement}
+\sname{Supplement B}\label{suppB}
+\stitle{Optimal Estimates of the HDRDA Classifier}
+\slink[url]{https://github.com/ramey/paper-hdrda}
+\sdescription{The graphical summary of the optimal estimates of the \emph{HDRDA} classifier from the classification study.}
+\end{supplement}
 

### 2013-05-12 lib/helpers.R
Updated Guo classifier in simulation code to use equal prior probabilities.

@@ -259,10 +259,10 @@ Cao <- function(train_x, train_y, test_x) {
 
 
 # Classifier from Guo, Hastie, and Tibshirani (2007) - Biostatistics
-scrda_train <- function(x, y) {
+scrda_train <- function(x, y, prior) {
   x <- t(x)
-  rda_out <- rda(x = x, y = y)
-  rda_cv_out <- rda.cv(rda_out, x = x, y = y)
+  rda_out <- rda(x = x, y = y, prior = prior)
+  rda_cv_out <- rda.cv(rda_out, x = x, y = y, prior = prior)
 
   # Which alpha and delta give min cv error?
   min_cv_error <- with(rda_cv_out, which(cv.err == min(cv.err), arr.ind = TRUE))

### 2013-05-13 reports/paper-hdrda.Rnw
Fixed an alignment issue.

@@
 	(\bm x - \xbar_k)^{T} \tilde{\bm \Sigma}_k^{+}(\bm x - \xbar_k)  + \log |\tilde{\bm \Sigma}_k| &= (\bm x - \xbar_k)^{T} \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm x - \xbar_k)\\
-	&+ \gamma^{+} (\bm x - \xbar_k)^{T} \bm U_2 \bm U_2^{T} (\bm x - \xbar_k) + \log | \bm W_k | + (p - q) \log \gamma.
+	&+ \gamma^{+} (\bm x - \xbar_k)^{T} \bm U_2 \bm U_2^{T} (\bm x - \xbar_k) + \log | \bm W_k |\\
+  &+ (p - q) \log \gamma.
 	\end{align*}
@@
 p <- p + geom_line(aes(color = Classifier), size = 1, linetype = 2)
-p <- p + geom_point(aes(color = Classifier, shape = Classifier), size = 4)
+p <- p + geom_point(aes(color = Classifier, shape = Classifier), size = 3)
 p <- p + facet_wrap(~ data_set) + ylim(c(0, max(error_rates$Avg_Error))) + theme_bw()

### 2013-05-13 reports/paper-hdrda.Rnw
Applied DMY's edits.

@@
 
-A reliable supervised classification of tumors from gene expression microarray
+Reliable supervised classification of tumors from gene expression microarray
 data is essential for successful diagnosis and treatment of diseases such as
@@
 dimension reduction and covariance-matrix shrinkage to enable a computationally
-efficient classifier and model-selection procedure. We demonstrate with four
-high-dimensional microarray data sets that our proposed classifier is often
-superior to several recently proposed sparse and regularized classifiers in
-terms of classification accuracy. Finally, we provide an implementation of our
-proposed classifier in the {\tt sparsediscrim} R package, available on CRAN.
+efficient classifier. We demonstrate with four high-dimensional microarray data
+sets that our proposed classifier is often superior to several recently proposed
+sparse and regularized classifiers in terms of classification accuracy. Finally,
+we provide an implementation of our proposed classifier in the {\tt
+  sparsediscrim} R package, available on CRAN.
 
@@
 Here, we propose the high-dimensional regularized discriminant analysis
-(\emph{HDRDA}) by first constructing a pooling matrix as a convex combination of
-the covariance-matrix estimators employed in linear and quadratic discriminant
-analysis. Given that the assumption of equal population covariance matrices for
-linear discriminant analysis is often too stringent in practice
+(\emph{HDRDA}) classifier by first constructing a pooling matrix as a convex
+combination of the covariance-matrix estimators employed in linear and quadratic
+discriminant analysis. Given that the assumption of equal population covariance
+matrices for linear discriminant analysis is often too stringent in practice
 \citep{Clemmensen:2011kr}, our proposed classifier can relax the linearity
@@
 the inverse and determinants can be reduced to calculations of $n_k \times n_k$
-matrices, where $n_k$ is the sample size of the $k$th class $(k = 1, \ldots,
-K)$. This reduction in calculation is clearly advantageous because $n_k$ is
-often quite small in microarray data sets (e.g., 15--20 observations).
+matrices, where $n_k$ is the sample size of the $k$th class, $k = 1, \ldots, K$.
+This reduction in calculation is clearly advantageous because $n_k$ is often
+quite small in microarray data sets (e.g., 10--20 observations).
 
-We also demonstrate that the \emph{HDRDA} classifier is competitive with and
-often superior to several recent classifiers designed for small-sample,
-high-dimensional data. In particular, we compare our proposed classifier with
-diagonal linear discriminant analysis proposed by \cite{Dudoit:2002ev} and the
-recently proposed classifiers from \cite{Guo:2007te}, \cite{Pang:2009ik},
-\cite{Tong:2012hw}, and \cite{Witten:2011kc}. We conduct our comparison of the
-classification accuracy of the competing classifiers using four small-sample,
-high-dimensional benchmark microarray data sets.
+We also demonstrate that the \emph{HDRDA} classifier is often superior to
+several recent classifiers designed for small-sample, high-dimensional data. In
+particular, we compare our proposed classifier with diagonal linear discriminant
+analysis proposed by \cite{Dudoit:2002ev} and the recently proposed classifiers
+from \cite{Guo:2007te}, \cite{Pang:2009ik}, \cite{Tong:2012hw}, and
+\cite{Witten:2011kc}. We conduct our comparison of the classification accuracy
+of the competing classifiers using four small-sample, high-dimensional benchmark
+microarray data sets.
 
-\section{Regularized Discriminant Analysis}
+\section{High-Dimensional Regularized Discriminant Analysis}
 \label{sec:rda}
@@
 $p(\omega_k)$ is the prior probability of class membership of the $k$th
-class. We further assume that each of $p(\omega_k)$ is equal $(k = 1, \ldots,
-K)$.
+class. We further assume that each of $p(\omega_k)$ is equal, $k = 1, \ldots, K$.
 
@@
 respect to a $0-1$ loss function when $p(\bm x | \omega_k)$ is the probability
-density function of the multivariate normal distribution with known mean vector
-$\bm \mu_k \in \mathbb{R}_{p \times 1}$ and known covariance matrix $\bm
+density function of the multivariate normal distribution with known mean vectors
+$\bm \mu_k \in \mathbb{R}_{p \times 1}$ and known covariance matrices $\bm
 \Sigma_k \in \mathbb{R}_{p \times p}^{>}$, $k = 1, 2, \ldots, K$, where
@@
 denotes the matrix transpose of $\bm A \in \mathbb{R}_{m \times p}$. If we
-assume further that $\bm \Sigma_k = \bm \Sigma$ $(k = 1, \ldots, K)$, then the
+assume further that $\bm \Sigma_k = \bm \Sigma$, $k = 1, \ldots, K$, then the
 pooled sample covariance matrix $\widehat{\bm \Sigma}$ is substituted for
@@
 
-Throughout the document we also require the following additional notation. Let
-$\bm I_m$ be the $m \times m$ identity matrix, and let $\bm 0_{m \times p}$ be
-the $m \times p$ matrix of zeros, such that $\bm 0_m$ is understood to denote
-$\bm 0_{m \times m}$. For $c \in \mathbb{R}$, let $c^+ = 1/c$ if $c \ne 0$ and
-$0$ otherwise. Let $\bm A \oplus \bm B$ denote the direct sum of $\bm A \in
-\mathbb{R}_{r \times r}$ and $\bm B \in \mathbb{R}_{s \times s}$ \citep[Chapter
-  1]{Lutkepohl:1996uz}. Let $\bm A^+$ and $\mathcal{N}(\bm A) = \{\bm z \in
-\mathbb{R}_{p \times 1} : \bm A \bm z = \bm 0_{m \times 1}\}$ denote the
-Moore-Penrose pseudoinverse and null space of $\bm A \in \mathbb{R}_{m \times
-  p}$, respectively. Let $V^{\perp} = \{\bm y \in \mathbb{R}_{p \times 1} : \bm
-y^{T} \bm v = 0\ \forall \bm v \in V \}$ denote the orthogonal complement of a
-vector space $V \subset \mathbb{R}_{p \times 1}$.
-
-\subsection{Covariance Matrix Regularization}
-
 The smallest eigenvalues and the directions associated with their eigenvectors
@@
 then rank$(\widehat{\bm \Sigma}_k) \le n_k$, which implies that at least $p -
-n_k$ eigenvalues of $\widehat{\bm \Sigma}_k$ are 0. Furthermore, although more
-feature information is available to discriminate among the $K$ classes, if $p >
-n_k$, \eqref{eq:qda} is incalculable because $\widehat{\bm \Sigma}_k^{-1}$ does
-not exist.
+n_k$ eigenvalues of $\widehat{\bm \Sigma}_k$ are zero. Furthermore, although
+more feature information is available to discriminate among the $K$ classes, if
+$p > n_k$, \eqref{eq:qda} is incalculable because $\widehat{\bm \Sigma}_k^{-1}$
+does not exist.
 
@@
 
-\clearpage
-
-\section{High-Dimensional Regularized Discriminant Analysis}
+Throughout the remainder of this document we also require the following
+additional notation. Let $\bm I_m$ be the $m \times m$ identity matrix, and let
+$\bm 0_{m \times p}$ be the $m \times p$ matrix of zeros, such that $\bm 0_m$ is
+understood to denote $\bm 0_{m \times m}$. For $c \in \mathbb{R}$, let $c^+ =
+1/c$ if $c \ne 0$ and $0$ otherwise. Let $\bm A \oplus \bm B$ denote the direct
+sum of $\bm A \in \mathbb{R}_{r \times r}$ and $\bm B \in \mathbb{R}_{s \times
+  s}$ \citep[Chapter 1]{Lutkepohl:1996uz}. Let $\bm A^+$ and $\mathcal{N}(\bm A)
+= \{\bm z \in \mathbb{R}_{p \times 1} : \bm A \bm z = \bm 0_{m \times 1}\}$
+denote the Moore-Penrose pseudoinverse and null space of $\bm A \in
+\mathbb{R}_{m \times p}$, respectively. Let $V^{\perp} = \{\bm y \in
+\mathbb{R}_{p \times 1} : \bm y^{T} \bm v = 0$ for all $\bm v \in V \}$ denote
+the orthogonal complement of a vector space $V \subset \mathbb{R}_{p \times 1}$.
 
@@
   \widehat{\bm \Sigma}_k(\lambda) = (1 - \lambda) \widehat{\bm \Sigma}_k + \lambda
-  \widehat{\bm \Sigma}, \quad (k = 1, \ldots, K)\label{eq:sig-lambda-alternative}
+  \widehat{\bm \Sigma}, \quad k = 1, \ldots, K,\label{eq:sig-lambda-alternative}
 \end{align}
@@
 borrowed from $\widehat{\bm \Sigma}$ to estimate $\bm \Sigma_k$, and the
-\emph{shrinkage} parameter $\gamma$ determines the amount of eigenvalue
-shrinkage applied. The choice of $\alpha_k$ allows for a flexible formulation of
+\emph{shrinkage} parameter $\gamma$ determines the degree of eigenvalue
+shrinkage. The choice of $\alpha_k$ allows for a flexible formulation of
 covariance-matrix estimators. For instance, if $\alpha_k = 1$, $k = 1, \ldots,
@@
 \eqref{eq:sig-lambda-alternative}. Then, $\mathcal{N}\{\widehat{\bm \Sigma}_k(\lambda)\} \subset \mathcal{N}(\widehat{
-  \bm \Sigma}) \subset \mathcal{N}(\widehat{\bm \Sigma}_k)$ $(k = 1, \ldots, K)$.
+  \bm \Sigma}) \subset \mathcal{N}(\widehat{\bm \Sigma}_k)$, $k = 1, \ldots, K$.
 \end{lemma}
@@
 \begin{align}
-	\tilde{\bm \Sigma}_k &= \bm U(\bm W_k \oplus \gamma \bm I_{p-q})\bm U^{T} \quad
-  (k = 1, \ldots, K),\label{eq:rda-matrix}
+	\tilde{\bm \Sigma}_k &= \bm U(\bm W_k \oplus \gamma \bm I_{p-q})\bm U^{T}, \quad
+  k = 1, \ldots, K,\label{eq:rda-matrix}
 \intertext{where}
@@
 \bm 0_{p \times (p - q)}$. Hence, $\bm U^{T} \widehat{\bm \Sigma}_k \bm U = \bm
-U_1^{T} \widehat{\bm \Sigma}_k \bm U_1 \oplus \bm 0_{p-q}$ $(k = 1, \ldots,
-K)$. Thus, $\bm U^{T} \tilde{\bm \Sigma}_k \bm U = \alpha_k \{(1 - \lambda) \bm
+U_1^{T} \widehat{\bm \Sigma}_k \bm U_1 \oplus \bm 0_{p-q}$, $k = 1, \ldots,
+K$. Thus, $\bm U^{T} \tilde{\bm \Sigma}_k \bm U = \alpha_k \{(1 - \lambda) \bm
 U^{T} \widehat{\bm \Sigma}_k \bm U + \lambda \bm D\} + \gamma \bm I_p$, and
@@
 \eqref{eq:sig-lambda-alternative}. Then, for $\lambda \in (0, 1]$,
-  rank$\{\widehat{\bm \Sigma}_k(\lambda)\} = q$ $(k = 1, \ldots, K)$.
+  rank$\{\widehat{\bm \Sigma}_k(\lambda)\} = q$, $k = 1, \ldots, K$.
 \end{cor}
@@
 
-We now present our main result, where we decompose the decision rule of the
-\emph{HDRDA} classifier and show that the term requiring the largest
-computational costs does not contribute to the classification of an unlabeled
-observation performed using Lemma \ref{lemma:RDA-constant-term}. Hence, we
-reduce \eqref{eq:hdrda} to an equivalent, more computationally efficient
-decision rule.
+We now present our main result where we decompose \eqref{eq:hdrda} and show that
+the term requiring the largest computational costs does not contribute to the
+classification of an unlabeled observation performed using Lemma
+\ref{lemma:RDA-constant-term}. Hence, we reduce \eqref{eq:hdrda} to an
+equivalent, more computationally efficient decision rule.
 
@@
 \bm W_k^{-1} \oplus \gamma^{+} \bm I_{p-q} )\bm U^{T}$ and $|\tilde{\bm \Sigma}_k| =
-\gamma^{p-q}| \bm W_k |$ $(k = 1, \ldots, K)$. Therefore, for all $\bm x \in
+\gamma^{p-q}| \bm W_k |$, $k = 1, \ldots, K$. Therefore, for all $\bm x \in
 \mathbb{R}_{p \times 1}$, we have that
@@
 \begin{proposition}\label{proposition:hdrda-W_k}
-Let $\bm W_k$ be defined as above $(k = 1, \ldots, K)$. Then, $|\bm W_k| = |
+Let $\bm W_k$ be defined as above, $k = 1, \ldots, K$. Then, $|\bm W_k| = |
 \bm \Gamma_k| |\bm Q_k|$ and
@@
 with the exception that we explicitly set prior probabilities as equal, if the
-option was available. For the \emph{HDRDA} classifier, we set $\alpha_k = 1$ $(k
-= 1, \ldots, K)$, so that the covariance-matrix estimator \eqref{eq:hdrda-cov}
-employed resembled \eqref{eq:ridge-estimator}. We estimated $\lambda$ from a
-grid of 21 equidistant candidate values between 0 and 1, inclusively. Similarly,
-we estimated $\gamma$ from a grid consisting of the values $10^{-1}, \ldots,
-10^4$, and $10^5$. We selected optimal estimates of $\lambda$ and $\gamma$ using
-10-fold cross-validation \citep{Hastie:2008dt}. Each of the microarray data sets
-listed below can be found in \ref{suppA}.
+option was available. For the \emph{HDRDA} classifier
+\eqref{eq:hdrda-decomposed}, we set $\alpha_k = 1$, $k = 1, \ldots, K$, so that
+the covariance-matrix estimator \eqref{eq:hdrda-cov} employed resembled
+\eqref{eq:ridge-estimator}. We estimated $\lambda$ from a grid of 21 equidistant
+candidate values between 0 and 1, inclusively. Similarly, we estimated $\gamma$
+from a grid consisting of the values $10^{-1}, \ldots, 10^4$, and $10^5$. We
+selected optimal estimates of $\lambda$ and $\gamma$ using 10-fold
+cross-validation \citep{Hastie:2008dt}. Each of the microarray data sets listed
+below can be found in \ref{suppA}.
 
@@
 classifier outperformed its competitors for all values of $d$ examined. For the
-Nakayama data set, the \emph{HDRDA} classifier was outperformed by the Dudoit,
-Pang, and Tong classifiers for small values of $d$ but attained classification
-rates superior to its competitors for larger values of $d$. Furthermore, Figure
-\ref{fig:error-rates} emphasizes two strengths of our classifier: reliability of
-the classifier for small $d$ and stability of the classifier for large $d$. In
-contrast to the other classifiers, the \emph{HDRDA} classifier exhibited less
-variability in its misclassification estimates for small values of $d$ and
-yielded superior classification accuracy as $d$ increased and was hardly
-effected by the inclusion of more variables in the data sets. We attribute the
-apparent stability of the error-rate estimates of our classifier for large
-values of $d$ to its novel dimension-reduction technique.
+Nakayama data set, the \emph{HDRDA} classifier \eqref{eq:hdrda-decomposed} was
+slightly outperformed by the Dudoit, Pang, and Tong classifiers for small values
+of $d$ but attained classification rates superior to its competitors for larger
+values of $d$. Furthermore, Figure \ref{fig:error-rates} emphasizes two
+strengths of our classifier: reliability of the classifier for relatively small
+$d$ and stability of the classifier for large $d$. In contrast to the other
+classifiers, the \emph{HDRDA} classifier exhibited less variability in its
+misclassification estimates for small values of $d$ and yielded superior
+classification accuracy as $d$ increased and was hardly effected by the
+inclusion of more variables in the data sets. We attribute the apparent
+stability of the error-rate estimates of our classifier for large values of $d$
+to its novel dimension-reduction technique.
 
-The error-rate estimates for the Dudoit, Pang, Tong, and Witten classifiers were
-comparable for all values of $d$ examined. Furthermore, the error-rate estimates
-of the Dudoit, Pang, and Tong classifiers were approximately equal throughout
-our study, and the Witten classifier yielded approximately equal error-rate
-estimates for the Burczysnki, Shipp, and Singh data sets. Furthermore, the
-error-rate estimates for the Dudoit and Tong classifiers were nearly identical,
-which suggests that the improved mean estimator employed by \cite{Tong:2012hw}
-provided little improvement to the classification performance of diagonal linear
-discriminant analysis.
+The average error-rate estimates for the Dudoit, Pang, Tong, and Witten
+classifiers were comparable for all values of $d$ examined. Furthermore, the
+error-rate estimates of the Dudoit, Pang, and Tong classifiers were
+approximately equal throughout our study, and the Witten classifier yielded
+approximately equal error-rate estimates for the Burczysnki, Shipp, and Singh
+data sets. Furthermore, the error-rate estimates for the Dudoit and Tong
+classifiers were nearly identical, which suggests that the improved mean
+estimator employed by \cite{Tong:2012hw} provided little improvement to the
+classification performance of diagonal linear discriminant analysis.
 
@@
 \emph{HDRDA} classifier. For brevity, we provide graphical summaries of the
-optimal estimates of $\lambda$ and $\gamma$ in \ref{suppB}. For
-each of the four data sets, the distribution of the optimal values of $\lambda$
-selected through cross-validation often had a notable bathtub shape, resembling
-a Beta distribution with both shape parameters less than unity. Specifically,
-the optimal values of $\lambda$ were typically near 0 or near 1, corresponding
-to quadratic and linear discriminant analysis, respectively. For instance, with
+optimal estimates of $\lambda$ and $\gamma$ in \ref{suppB}. For each of the four
+data sets, the distribution of the optimal values of $\lambda$ selected through
+cross-validation often had a notable bathtub shape, resembling a Beta
+distribution with both shape parameters less than unity. Specifically, the
+optimal values of $\lambda$ were typically near 0 or near 1, corresponding to
+quadratic and linear discriminant analysis, respectively. For instance, with
 smaller values of $d$ applied to the Shipp data set, the range of optimal values
@@
 number of training data sets for which we can safely assume the covariance
-matrices are unequal or nearly unequal. As $d$ increased, the majority of the
-optimal values of $\lambda$ were selected as unity, thereby an increased number
-of variables resulted in the \emph{HDRDA} classifier preferring the more
-parsimonious assumption of equal covariance matrices. In this instance we can
-see the flexibility of our proposed classifier to employ the more realistic
-assumption that the class covariance matrices are unequal, but if optimal, we
-can employ the more parsimonious assumption of equal covariance matrices.
+matrices are unequal. As $d$ increased, the majority of the optimal values of
+$\lambda$ were selected as unity, thereby an increased number of variables
+resulted in the \emph{HDRDA} classifier preferring the more parsimonious
+assumption of equal covariance matrices. In this instance we can see the
+flexibility of our proposed classifier to employ the more realistic assumption
+of unequal class covariance matrices, but if optimal, the \emph{HDRDA}
+classifier \eqref{eq:hdrda-decomposed} can utilize the more parsimonious
+assumption of equal covariance matrices.
 
@@
 
-In our simulation study, we compared our proposed \emph{HDRDA} classifier to
-several recently proposed classifiers, including the Dudoit, Pang, Tong, and Guo
-classifiers, each of which assumes independence of the features conditional on
-class membership. Despite the rapid computational performance of diagonal
+In our simulation study, we have compared our proposed \emph{HDRDA} classifier
+to several recently proposed classifiers, including the Dudoit, Pang, Tong, and
+Guo classifiers, each of which assumes independence of the features conditional
+on class membership. Despite the rapid computational performance of diagonal
 classifiers and the reduction in the number of parameters to estimate,
@@
 assertions of \cite{Mai:2012bf} and \cite{Fan:2012iq}. Furthermore, the
-implementation of our classifier in the {\tt sparsediscrim} R package was
-comparable to that of the diagonal classifiers in terms of computational
-runtime. Moreover, when preliminary variable selection was applied to the data
-in our classification study, the time to perform model selection with our
-proposed classifier was nearly instantaneous. Hence, our proposed \emph{HDRDA}
-classifier is competitive in terms of classification accuracy for modern
-small-sample, high-dimensional data sets and requires substantially less
+implementation of \eqref{eq:hdrda-decomposed} in the {\tt sparsediscrim} R
+package was comparable to that of the diagonal classifiers in terms of
+computational runtime. Moreover, when preliminary variable selection was applied
+to the data in our classification study, the time to perform model selection
+with our proposed classifier was nearly instantaneous. Hence, our proposed
+\emph{HDRDA} classifier is competitive in terms of classification accuracy for
+modern small-sample, high-dimensional data sets and requires substantially less
 computing time than standard regularized classifiers, such as regularized

### 2013-12-11 reports/paper-hdrda.Rnw
Updated link to supplementary material in paper.
Updated the keywords.

@@
 \kwd{covariance-matrix regularization}
-\kwd{high-dimensional microarray data}
-\kwd{dimension reduction}
+\kwd{high-dimensional multivariate analysis}
+\kwd{microarray data}
 \end{keyword}
@@
 \stitle{Optimal Estimates of the HDRDA Classifier}
-\slink[url]{https://github.com/ramey/paper-hdrda}
+\slink[url]{https://github.com/ramey/paper-hdrda/tree/annals-of-applied-statistics/reports}
 \sdescription{The graphical summary of the optimal estimates of the \emph{HDRDA} classifier from the classification study.}

### 2013-12-11 reports/paper-hdrda.Rnw
Updated paper to use generic LaTeX template with doublespacing. Removed all AoAS formatting templates. Closes #9

@@
-\documentclass[aoas,preprint]{imsart}
+\documentclass{article}
 
 \usepackage{graphicx, amsmath, amssymb, bm, mathtools, amsthm}
-\RequirePackage[OT1]{fontenc}
-\RequirePackage{natbib}
-\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
+\usepackage{natbib, setspace}
 
-% provide arXiv number if available:
-%\arxiv{arXiv:0000.0000}
-
-\startlocaldefs
 \newcommand{\xbar}{\bar{\bm x}}
@@
 \newtheorem{thm}{Theorem}
-\endlocaldefs
 
@@
 
-\begin{frontmatter}
+\title{High-Dimensional Regularized Discriminant Analysis}
 
-\title{High-Dimensional Regularized Discriminant Analysis}
-\runtitle{High-Dimensional Regularized Discriminant Analysis}
+\author{John A. Ramey, Caleb K. Stein, \and Dean M. Young}
 
-\begin{aug}
-\author{\fnms{John A.} \snm{Ramey}\thanksref{m1}
-  \ead[label=e1]{johnramey@gmail.com}
-  \ead[label=u1,url]{http://ramhiser.com}}
-\author{\fnms{Caleb K.} \snm{Stein}\thanksref{m2}\ead[label=e2]{Caleb\_Stein@baylor.edu}}
-\and
-\author{\fnms{Dean M.} \snm{Young}\thanksref{m2}\ead[label=e3]{Dean\_Young@baylor.edu}}
+\maketitle
 
-\runauthor{J. A. Ramey et al.}
-
-\affiliation{Fred Hutchinson Cancer Research Center\thanksmark{m1} and Baylor University\thanksmark{m2}}
-
-\address{Vaccine and Infectious Disease Division\\
-  Fred Hutchinson Cancer Research Center\\
-  1100 Fairview Ave. N., M2-C200\\
-  Seattle, Washington 98115, U.S.A.\\
-  \printead{e1}\\
-  \printead{u1}}
-
-\address{Department of Statistical Science\\
-  Baylor University\\
-  P.O. Box 97140\\
-  Waco, Texas 76706, U.S.A.\\
-  \printead{e2}\\
-  \phantom{E-mail:\ }\printead*{e3}}
-\end{aug}
+\doublespacing
 
@@
 \end{abstract}
-
-\begin{keyword}
-\kwd{classification}
-\kwd{regularized discriminant analysis}
-\kwd{covariance-matrix regularization}
-\kwd{high-dimensional multivariate analysis}
-\kwd{microarray data}
-\end{keyword}
-
-\end{frontmatter}
 
@@
 Several regularization methods, such as the methods considered by
-\citet*{Guo:2007te}, \cite{Mkhadri:1995jp}, and \citet*{Xu:2009fl}, have been
+\citet*{Xu:2009fl}, \citet*{Guo:2007te}, and \cite{Mkhadri:1995jp} have been
 proposed in the literature to adjust the eigenvalues of $\widehat{\bm \Sigma}_k$
@@
 increase the rank of $\widehat{\bm \Sigma}_k(\lambda)$ to $q$ if $\lambda \ne
-0$. Next, we provide an essential result that enables us to prove that
-\eqref{eq:hdrda} is invariant to adjustments to the eigenvalues of $\tilde{ \bm
-  \Sigma}_k$ corresponding to the null space of $\widehat{\bm \Sigma}$ for any
-unlabeled observation $\bm x \in \mathbb{R}_{p \times 1}$.
+0$. The results is that the \emph{HDRDA} classifier has the advantage over
+\eqref{eq:qda} in that the rank$(\widehat{\bm \Sigma}_k) \le n_k$. Next, we
+provide an essential result that enables us to prove that \eqref{eq:hdrda} is
+invariant to adjustments to the eigenvalues of $\tilde{ \bm \Sigma}_k$
+corresponding to the null space of $\widehat{\bm \Sigma}$ for any unlabeled
+observation $\bm x \in \mathbb{R}_{p \times 1}$.
 
@@
 
-\begin{supplement}
-\sname{Supplement A}\label{suppA}
-\stitle{R Package Containing Microarray Data Sets}
-\slink[url]{https://github.com/ramey/datamicroarray}
-\sdescription{An R package containing all of the microarray data sets used in
-  the classification study.}
-\end{supplement}
-
-\begin{supplement}
-\sname{Supplement B}\label{suppB}
-\stitle{Optimal Estimates of the HDRDA Classifier}
-\slink[url]{https://github.com/ramey/paper-hdrda/tree/annals-of-applied-statistics/reports}
-\sdescription{The graphical summary of the optimal estimates of the \emph{HDRDA} classifier from the classification study.}
-\end{supplement}
-
-\bibliographystyle{imsart-nameyear}
+\bibliographystyle{plainnat}
 \bibliography{rda}

### 2013-12-11 reports/paper-hdrda.Rnw
Updated paper's title.

@@
 
-\title{High-Dimensional Regularized Discriminant Analysis}
+\title{Efficient Calculation of Regularized Discriminant Analysis for
+  High-Dimensional Classification}
 
@@
   \Sigma}_k$ are well-known to be biased if $p \ge n_k$ such that the smallest
-eigenvalues are underestimated \citep{Seber:2004uh}, Moreover, if $p > n_k$,
+eigenvalues are underestimated \citep{Seber:2004uh}. Moreover, if $p > n_k$,
 then rank$(\widehat{\bm \Sigma}_k) \le n_k$, which implies that at least $p -

### 2013-12-12 reports/paper-hdrda.Rnw
Updated LaTeX page dimensions.
Reorganized so that all the notation is at the beginning of Section 2.

@@
 \newtheorem{thm}{Theorem}
+
+\pagestyle{plain}
+%----------------Page dimensions ----------------
+\oddsidemargin 0.0in
+\evensidemargin 0.0in
+\topmargin -0.75in
+\leftmargin 0in
+\headheight 0.0in
+\headsep 0.5in
+\footskip 0.5in
+\footnotesep 0.0in
+\textwidth 6.5in
+\textheight 9.5in
 
@@
 \subsection{Preliminaries}
+To facilitate our discussion of covariance-matrix regularization and regularized
+discriminant analysis, we require the following notation. Let $\mathbb{R}_{a
+  \times b}$ denotes the matrix space of all $a \times b$ matrices over the real
+field $\mathbb{R}$. Denote by $\bm I_m$ the $m \times m$ identity matrix, and
+let $\bm 0_{m \times p}$ be the $m \times p$ matrix of zeros, such that $\bm
+0_m$ is understood to denote $\bm 0_{m \times m}$. Let $\bm A \oplus \bm B$
+denote the direct sum of $\bm A \in \mathbb{R}_{r \times r}$ and $\bm B \in
+\mathbb{R}_{s \times s}$ \citep[Chapter 1]{Lutkepohl:1996uz}. Let $\bm A^{T}$,
+$\bm A^+$, and $\mathcal{N}(\bm A) = \{\bm z \in \mathbb{R}_{p \times 1} : \bm A
+\bm z = \bm 0_{m \times 1}\}$ denote the transpose, the Moore-Penrose
+pseudoinverse, and the null space of $\bm A \in \mathbb{R}_{m \times p}$,
+respectively. Let $\mathbb{R}_{p \times p}^{>}$ denote the cone of real $p
+\times p$ positive-definite matrices. Similarly, let $\mathbb{R}_{p \times
+  p}^{\ge}$ denote the cone of real $p \times p$ positive-semidefinite
+matrices. Let $V^{\perp} = \{\bm y \in \mathbb{R}_{p \times 1} : \bm y^{T} \bm v
+= 0$ for all $\bm v \in V \}$ denote the orthogonal complement of a vector space
+$V \subset \mathbb{R}_{p \times 1}$. For $c \in \mathbb{R}$, let $c^+ = 1/c$ if
+$c \ne 0$ and $0$ otherwise.
+
 In discriminant analysis we wish to assign correctly an unlabeled
@@
   1}$ be the $i$th observation $(i = 1, \ldots, N)$ with true, unique membership
-$y_i \in \{\omega_1, \ldots, \omega_K\}$, where $\mathbb{R}_{a \times b}$
-denotes the matrix space of all $a \times b$ matrices over the real field
-$\mathbb{R}$. We assume that $(\bm x_i, y_i)$ is a realization from a mixture
-distribution $p(\bm x) = \sum_{k=1}^K p(\bm x | \omega_k) p(\omega_k)$, where
-$p(\bm x | \omega_k)$ is the probability density function of the $k$th class and
-$p(\omega_k)$ is the prior probability of class membership of the $k$th
-class. We further assume that each of $p(\omega_k)$ is equal, $k = 1, \ldots, K$.
+$y_i \in \{\omega_1, \ldots, \omega_K\}$. We assume that $(\bm x_i, y_i)$ is a
+realization from a mixture distribution $p(\bm x) = \sum_{k=1}^K p(\bm x |
+\omega_k) p(\omega_k)$, where $p(\bm x | \omega_k)$ is the probability density
+function of the $k$th class and $p(\omega_k)$ is the prior probability of class
+membership of the $k$th class. We further assume that each of $p(\omega_k)$ is
+equal, $k = 1, \ldots, K$.
 
@@
 $\bm \mu_k \in \mathbb{R}_{p \times 1}$ and known covariance matrices $\bm
-\Sigma_k \in \mathbb{R}_{p \times p}^{>}$, $k = 1, 2, \ldots, K$, where
-$\mathbb{R}_{p \times p}^{>}$ denotes the cone of real $p \times p$
-positive-definite matrices. Because $\bm \mu_k$ and $\bm \Sigma_k$ are typically
-unknown, we assign an unlabeled observation $\bm x$ to class $\omega_k$ with the
-sample quadratic discriminant analysis classifier
+\Sigma_k \in \mathbb{R}_{p \times p}^{>}$, $k = 1, 2, \ldots, K$. Because $\bm
+\mu_k$ and $\bm \Sigma_k$ are typically unknown, we assign an unlabeled
+observation $\bm x$ to class $\omega_k$ with the sample quadratic discriminant
+analysis classifier
 \begin{align}
@@
 where $\xbar_k$ and $\widehat{\bm \Sigma}_k$ are the maximum likelihood
-estimators of $\bm \mu_k$ and $\bm \Sigma_k$, respectively, and $\bm A^{T}$
-denotes the matrix transpose of $\bm A \in \mathbb{R}_{m \times p}$. If we
-assume further that $\bm \Sigma_k = \bm \Sigma$, $k = 1, \ldots, K$, then the
-pooled sample covariance matrix $\widehat{\bm \Sigma}$ is substituted for
-$\widehat{ \bm \Sigma}_k$ in \eqref{eq:qda}, where
+estimators of $\bm \mu_k$ and $\bm \Sigma_k$, respectively. If we assume further
+that $\bm \Sigma_k = \bm \Sigma$, $k = 1, \ldots, K$, then the pooled sample
+covariance matrix $\widehat{\bm \Sigma}$ is substituted for $\widehat{ \bm
+  \Sigma}_k$ in \eqref{eq:qda}, where
 \begin{align}
@@
 reduces to the sample linear discriminant analysis classifier, and the
-log-determinant is omitted because it is constant across all classes.
+log-determinant is omitted because it is constant across the $K$ classes.
 
 The smallest eigenvalues and the directions associated with their eigenvectors
-highly influence \eqref{eq:qda}. In fact, the eigenvalues of $\widehat{ \bm
+can highly influence \eqref{eq:qda}. In fact, the eigenvalues of $\widehat{ \bm
   \Sigma}_k$ are well-known to be biased if $p \ge n_k$ such that the smallest
@@
 \cite{Ramey:2013ji}, \cite{Xu:2009fl}, and \cite{Ye:2009gd}.
-
-Throughout the remainder of this document we also require the following
-additional notation. Let $\bm I_m$ be the $m \times m$ identity matrix, and let
-$\bm 0_{m \times p}$ be the $m \times p$ matrix of zeros, such that $\bm 0_m$ is
-understood to denote $\bm 0_{m \times m}$. For $c \in \mathbb{R}$, let $c^+ =
-1/c$ if $c \ne 0$ and $0$ otherwise. Let $\bm A \oplus \bm B$ denote the direct
-sum of $\bm A \in \mathbb{R}_{r \times r}$ and $\bm B \in \mathbb{R}_{s \times
-  s}$ \citep[Chapter 1]{Lutkepohl:1996uz}. Let $\bm A^+$ and $\mathcal{N}(\bm A)
-= \{\bm z \in \mathbb{R}_{p \times 1} : \bm A \bm z = \bm 0_{m \times 1}\}$
-denote the Moore-Penrose pseudoinverse and null space of $\bm A \in
-\mathbb{R}_{m \times p}$, respectively. Let $V^{\perp} = \{\bm y \in
-\mathbb{R}_{p \times 1} : \bm y^{T} \bm v = 0$ for all $\bm v \in V \}$ denote
-the orthogonal complement of a vector space $V \subset \mathbb{R}_{p \times 1}$.
 

### 2013-12-12 reports/paper-hdrda.Rnw
Began rewrite of abstract and intro to focus on properties of RDA and its reparameterization. Related to #10 and #15.

@@
 
-\title{Efficient Calculation of Regularized Discriminant Analysis for
-  High-Dimensional Classification}
+\title{High-Dimensional Regularized Discriminant Analysis}
 
-\author{John A. Ramey, Caleb K. Stein, \and Dean M. Young}
+\author{John A. Ramey, Caleb K. Stein, and Dean M. Young}
 
@@
 
-Reliable supervised classification of tumors from gene expression microarray
-data is essential for successful diagnosis and treatment of diseases such as
-cancer; however, because the number of gene expression levels measured is
-typically much greater than the sample size, the analysis of microarray data
-often presents difficult statistical and computational challenges. Here, we
-propose a new regularized discriminant analysis classifier that incorporates
-dimension reduction and covariance-matrix shrinkage to enable a computationally
-efficient classifier. We demonstrate with four high-dimensional microarray data
-sets that our proposed classifier is often superior to several recently proposed
-sparse and regularized classifiers in terms of classification accuracy. Finally,
-we provide an implementation of our proposed classifier in the {\tt
-  sparsediscrim} R package, available on CRAN.
+LDA and QDA suck. Friedman to the rescue with RDA. We reformulate the RDA
+classifier. Our parameterizaton yields 1) improved interpretation, 2) nice
+properties that facilitate dimension reduction, and 3) as a result of the
+dimension reduction, efficient calculation for high-dimensional data.
+
+Here, we propose a new regularized discriminant analysis classifier that
+incorporates dimension reduction and covariance-matrix shrinkage to enable a
+computationally efficient classifier. We demonstrate with four high-dimensional
+microarray data sets that our proposed classifier is often superior to several
+recently proposed sparse and regularized classifiers in terms of classification
+accuracy. Finally, we provide an implementation of our proposed classifier in
+the {\tt sparsediscrim} R package, available on CRAN.
 
@@
 
-Microarrays have enabled researchers to measure gene-expression levels of
-thousands of genes in a single experiment to identify a subset of genes and
-their expression patterns that either characterize a particular cell state or
-facilitate an accurate prediction of future cell state.  We focus on the
-classification of tissues and other biological samples into one of $K$ disjoint,
-predefined types, where our goal is to ascertain the class membership of
-a new tissue sample based on gene-expression microarray data. However,
-microarray data typically consist of a large number of gene expressions $p$
-relative to the total sample size $N$. Moreover, standard benchmark microarray
-data sets usually consist of between $p = 10,000$ and $55,000$ probes and fewer
-than $N = 100$ observations, in which case standard supervised-learning methods
-exhibit poor classification performance because of the \emph{curse of
-  dimensionality} \citep{Bellman:1961tn}. Furthermore, some well-known
-classifiers, such as linear and quadratic discriminant analysis, are
-incalculable when $p > N$ because the sample covariance matrices employed are
-singular. As a result, many researchers, including \cite{Mai:2012bf},
-\cite{Fan:2012iq}, \cite{Merchante:2012vk}, \cite{Clemmensen:2011kr}, and
-\cite{Huang:2010ju}, have introduced restrictions on the population covariance
-matrices in linear and quadratic discriminant analysis to reduce the number of
-covariance parameters that must be estimated to ensure that the resulting
-classifiers are calculable. For instance, \cite{Dudoit:2002ev} have modified the
-linear and quadratic discriminant analysis classifiers by assuming that the
-population covariance matrices are diagonal, resulting in classifiers that are
-special cases of the Naive Bayes classifier \citep[Chapter 4]{Hastie:2008dt}.
+Linear discriminant analysis (LDA) and quadratic discriminant analysis (QDA) are
+classical approaches to supervised learning and are widely used in a variety of
+areas, such as medical research, facial recognition, and TODO
+\cite{Dudoit:2002ev,TODO1,TODO2} Both the LDA and QDA classifiers are based on
+the assumption that the underlying distribution of each of the $K$ classes are
+multivariate normally distributed. While both methods are popular in practice,
+they often cannot be applied to modern data sets because the sample covariance
+matrices employed are singular when the number of features exceeds the total
+sample size $N$. For instance, standard benchmark microarray data sets used in
+biological research often consist of between $p = 10,000$ and $55,000$ probes
+and fewer than $N = 100$ observations, in which case LDA and QDA are
+incalculable.
 
-Alternatively, several researchers, including \cite*{Zhang:2010va},
-\cite{Ye:2009gd}, \cite{Ji:2008wp}, \cite*{Guo:2007te},
-\cite*{Srivastava:2007ud}, \cite{Ye:2006vx}, and \cite{Ye:2005uu}, have proposed
-regularization methods to improve the estimation of the population covariance
-matrices before applying either linear or quadratic discriminant
-analysis. Covariance-matrix regularization methods often are based on the
-well-known ridge-regression approach of \cite{Hoerl:1970cd} and adjust the
-eigenvalues of the sample covariance matrices to ensure positive
-definiteness. For excellent overviews of the regularization problem in
-discriminant analysis, see \cite{Murphy:2012uq}, \cite{Mkhadri:1997gy}, and
-\cite{Friedman:1989tm}.
+\cite{Friedman:1989tm} proposed regularized discriminant analysis (RDA), which
+partially pools the individual sample covariance matrices used in the QDA
+classifier with the pooled sample covariance matrix from the LDA
+classifier. After pooling the covariance matrices, the RDA classifier employs a
+covariance-matrix regularization technique that shrinks the zero eigenvalues of
+the resulting covariance matrices to ensure they are nonsingular. Similarly,
+several researchers, including \cite*{Zhang:2010va}, \cite{Ye:2009gd},
+\cite{Ji:2008wp}, \cite*{Guo:2007te}, \cite*{Srivastava:2007ud},
+\cite{Ye:2006vx}, and \cite{Ye:2005uu}, have proposed regularization methods to
+improve the estimation of the population covariance matrices before applying
+either the LDA or QDA classifiers. Covariance-matrix regularization methods
+often are based on the well-known ridge-regression approach of
+\cite{Hoerl:1970cd} and adjust the eigenvalues of the sample covariance matrices
+to ensure positive definiteness. For excellent overviews of the regularization
+problem in discriminant analysis, see \cite{Murphy:2012uq},
+\cite{Mkhadri:1997gy}, and \cite{Friedman:1989tm}.
+
+% TODO: Should we remove this paragraph? Keep a couple of the recent references
+% though.
+Alternatively, many researchers, including \cite{Mai:2012bf}, \cite{Fan:2012iq},
+\cite{Merchante:2012vk}, \cite{Clemmensen:2011kr}, and \cite{Huang:2010ju}, have
+introduced restrictions on the population covariance matrices in linear and
+quadratic discriminant analysis to reduce the number of covariance parameters
+that must be estimated to ensure that the resulting classifiers are
+calculable. For instance, \cite{Dudoit:2002ev} have modified the linear and
+quadratic discriminant analysis classifiers by assuming that the population
+covariance matrices are diagonal, resulting in classifiers that are special
+cases of the Naive Bayes classifier \citep[Chapter 4]{Hastie:2008dt}.
+
+We reformulate the RDA classifier. Our parameterizaton yields 1) improved
+interpretation, 2) nice properties that facilitate dimension reduction, and 3)
+as a result of the dimension reduction, efficient calculation for
+high-dimensional data.
 
@@
 
-In discriminant analysis we wish to assign correctly an unlabeled
-$p$-dimensional observation vector $\bm x$ to one of $K$ unique, known classes
-(or populations) by constructing a classifier from $N$ training
-observations. Let $\bm x_i = (x_{i1}, \ldots, x_{ip}) \in \mathbb{R}_{p \times
-  1}$ be the $i$th observation $(i = 1, \ldots, N)$ with true, unique membership
-$y_i \in \{\omega_1, \ldots, \omega_K\}$. We assume that $(\bm x_i, y_i)$ is a
-realization from a mixture distribution $p(\bm x) = \sum_{k=1}^K p(\bm x |
-\omega_k) p(\omega_k)$, where $p(\bm x | \omega_k)$ is the probability density
-function of the $k$th class and $p(\omega_k)$ is the prior probability of class
-membership of the $k$th class. We further assume that each of $p(\omega_k)$ is
-equal, $k = 1, \ldots, K$.
+In discriminant analysis we wish to assign correctly an unlabeled vector $\bm x
+\in \mathbb{R}_{p \times 1}$ to one of $K$ unique, known classes by constructing
+a classifier from $N$ training observations. Let $\bm x_i = (x_{i1}, \ldots,
+x_{ip}) \in \mathbb{R}_{p \times 1}$ be the $i$th observation $(i = 1, \ldots,
+N)$ with true, unique membership $y_i \in \{\omega_1, \ldots, \omega_K\}$. We
+assume that $(\bm x_i, y_i)$ is a realization from a mixture distribution $p(\bm
+x) = \sum_{k=1}^K p(\bm x | \omega_k) p(\omega_k)$, where $p(\bm x | \omega_k)$
+is the probability density function of the $k$th class and $p(\omega_k)$ is the
+prior probability of class membership of the $k$th class. We further assume that
+each of $p(\omega_k)$ is equal, $k = 1, \ldots, K$.
 
@@
 \subsection{Model Specification}
+
+\cite{Halbe:2007ea} demonstrate that  parameterization results in a
 
@@
 {\tt sparsediscrim} R package, available on CRAN.
+
+\subsection{Relationship with Regularized Discriminant Analysis}
+
+\cite{Friedman:1989tm} proposed regularized discriminant analysis (RDA) by
+incorporating a weighted average of the sample covariance matrix $\widehat{ \bm
+  \Sigma}_k$ and the pooled sample covariance matrix $\widehat{\bm \Sigma}$ to
+estimate $\bm \Sigma_k$ with
+\begin{align}
+    \widehat{\bm \Sigma}_k(\lambda) = n_k^{-1}(\lambda) \bm S_k(\lambda) \quad
+    (k =
+      1, \ldots, K),\label{eq:sig-lambda}
+      \end{align}
+where $\lambda \in [0, 1]$, $n_k(\lambda) = (1 - \lambda) n_k + \lambda N$, $
+\bm S_k = n_k \widehat{\bm \Sigma}_k$, $\bm S = \sum_{k=1}^K \bm S_k$, and $ \bm
+S_k(\lambda) = (1 - \lambda) \bm S_k + \lambda \bm S$. If $\lambda = 0$,
+\eqref{eq:sig-lambda} is equal to $\widehat{\bm \Sigma}_k$ used in
+\eqref{eq:qda}. Alternatively, if $\lambda = 1$, \eqref{eq:sig-lambda} reduces
+to \eqref{eq:pooled-cov}.
+
+To further improve the estimation of $\bm \Sigma_k$ and to stabilize the inverse
+of \eqref{eq:sig-lambda}, \cite{Friedman:1989tm} has proposed the biased
+covariance-matrix estimator
+\begin{align}
+          \widehat{\bm \Sigma}_k(\lambda, \gamma) = (1 - \gamma) \widehat{\bm
+            \Sigma}_k(\lambda) + \gamma \frac{\tr\{\widehat{\bm
+              \Sigma}_k(\lambda)\}}{p} \bm I_p,\label{eq:sig-rda}
+          \end{align}
+where $\gamma \in [0, 1]$ is a regularization parameter that controls the
+shrinkage of \eqref{eq:sig-rda} towards $\bm I_p$ and is weighted by the average
+of the eigenvalues of \eqref{eq:sig-lambda}. Finally, the RDA classifier is
+established by substituting \eqref{eq:sig-rda} into \eqref{eq:qda}.
+
+\section{Properties of the HDRDA}
 

### 2013-12-16 reports/paper-hdrda.Rnw
The intro's layout is closer to where it needs to be.
- Intro still needs polish.
- Also, slightly reorganized paper's layout.
- Related to #10 and #15.

@@
 
-Linear discriminant analysis (LDA) and quadratic discriminant analysis (QDA) are
-classical approaches to supervised learning and are widely used in a variety of
-areas, such as medical research, facial recognition, and TODO
-\cite{Dudoit:2002ev,TODO1,TODO2} Both the LDA and QDA classifiers are based on
-the assumption that the underlying distribution of each of the $K$ classes are
-multivariate normally distributed. While both methods are popular in practice,
-they often cannot be applied to modern data sets because the sample covariance
-matrices employed are singular when the number of features exceeds the total
-sample size $N$. For instance, standard benchmark microarray data sets used in
-biological research often consist of between $p = 10,000$ and $55,000$ probes
-and fewer than $N = 100$ observations, in which case LDA and QDA are
-incalculable.
+Discriminant analysis is a widely used technique to classify an unlabeled
+$p$-dimensional vector $\bm x$ to one of $K$ unique, known classes by
+constructing a classifier from $N$ training observations. This is useful in a
+variety of areas, such as medical research, facial recognition, and TODO
+\cite{Dudoit:2002ev,TODO1,TODO2} because data-driven decision are becoming
+increasingly common.  Although a large number of classification algorithms have
+been proposed, linear discriminant analysis (LDA) and quadratic discriminant
+analysis (QDA) remain popular in practice because of their simplicity, speed of
+computation, and interpretability. Both the LDA and QDA classifiers are based on
+the assumption that each of the $K$ classes is multivariate normally
+distributed. While both methods are popular in practice, they often cannot be
+applied to modern data sets because the sample covariance matrices employed are
+singular when the number of features exceeds the total sample size $N$. For
+instance, standard benchmark microarray data sets used in biological research
+often consist of between $p = 10,000$ and $55,000$ probes and fewer than $N =
+100$ observations, in which case LDA and QDA are incalculable.\footnote{TODO:
+  Bring back the curse of dimensionality reference.}
 
-\cite{Friedman:1989tm} proposed regularized discriminant analysis (RDA), which
-partially pools the individual sample covariance matrices used in the QDA
-classifier with the pooled sample covariance matrix from the LDA
+As a result, several researchers have applied covariance-matrix regularization
+to the LDA and QDA classifiers. Covariance-matrix regularization methods often
+are based on the well-known ridge-regression approach of \cite{Hoerl:1970cd} and
+adjust the eigenvalues of the sample covariance matrices to ensure positive
+definiteness.  \cite*{Zhang:2010va}, \cite{Ye:2009gd}, \cite{Ji:2008wp},
+\cite*{Guo:2007te}, \cite*{Srivastava:2007ud}, \cite{Ye:2006vx}, and
+\cite{Ye:2005uu}, have proposed regularization methods to improve the estimation
+of the population covariance matrices before applying either the LDA or QDA
+classifiers. Alternatively, many researchers, including \cite{Mai:2012bf},
+\cite{Fan:2012iq}, \cite{Merchante:2012vk}, \cite{Clemmensen:2011kr}, and
+\cite{Huang:2010ju}, have introduced restrictions on the population covariance
+matrices to reduce the number of parameters that must be estimated. For
+instance, \cite{Dudoit:2002ev} have modified the linear and quadratic
+discriminant analysis classifiers by assuming that the population covariance
+matrices are diagonal, resulting in classifiers that are special cases of the
+Naive Bayes classifier \citep[Chapter 4]{Hastie:2008dt}. For excellent overviews
+of the regularization problem in discriminant analysis, see \cite{Murphy:2012uq}
+and \cite{Mkhadri:1997gy}.
+
+In a classic paper, \cite{Friedman:1989tm} proposed regularized discriminant
+analysis (RDA), which partially pools the individual sample covariance matrices
+used in the QDA classifier with the pooled sample covariance matrix from the LDA
 classifier. After pooling the covariance matrices, the RDA classifier employs a
 covariance-matrix regularization technique that shrinks the zero eigenvalues of
-the resulting covariance matrices to ensure they are nonsingular. Similarly,
-several researchers, including \cite*{Zhang:2010va}, \cite{Ye:2009gd},
-\cite{Ji:2008wp}, \cite*{Guo:2007te}, \cite*{Srivastava:2007ud},
-\cite{Ye:2006vx}, and \cite{Ye:2005uu}, have proposed regularization methods to
-improve the estimation of the population covariance matrices before applying
-either the LDA or QDA classifiers. Covariance-matrix regularization methods
-often are based on the well-known ridge-regression approach of
-\cite{Hoerl:1970cd} and adjust the eigenvalues of the sample covariance matrices
-to ensure positive definiteness. For excellent overviews of the regularization
-problem in discriminant analysis, see \cite{Murphy:2012uq},
-\cite{Mkhadri:1997gy}, and \cite{Friedman:1989tm}.
+the resulting covariance matrices to ensure they are nonsingular. Given that the
+assumption of equal population covariance matrices for the LDA classifier is
+often too stringent in practice \citep{Clemmensen:2011kr}, the RDA classifier
+can relax the linearity assumption by instead incorporating unequal covariance
+matrices while retaining as special cases the LDA and QDA classifiers. Despite
+its success, the RDA classifier is often impractical because the model-selection
+process for its two tuning parameters requires excessive computation. Moreover,
+the pooling parameter has no clear intrepretation, and others
+complaints\footnote{TODO: Add some more complaints.}\footnote{TODO: Citations
+  needed.}.
 
-% TODO: Should we remove this paragraph? Keep a couple of the recent references
-% though.
-Alternatively, many researchers, including \cite{Mai:2012bf}, \cite{Fan:2012iq},
-\cite{Merchante:2012vk}, \cite{Clemmensen:2011kr}, and \cite{Huang:2010ju}, have
-introduced restrictions on the population covariance matrices in linear and
-quadratic discriminant analysis to reduce the number of covariance parameters
-that must be estimated to ensure that the resulting classifiers are
-calculable. For instance, \cite{Dudoit:2002ev} have modified the linear and
-quadratic discriminant analysis classifiers by assuming that the population
-covariance matrices are diagonal, resulting in classifiers that are special
-cases of the Naive Bayes classifier \citep[Chapter 4]{Hastie:2008dt}.
-
-We reformulate the RDA classifier. Our parameterizaton yields 1) improved
-interpretation, 2) nice properties that facilitate dimension reduction, and 3)
-as a result of the dimension reduction, efficient calculation for
-high-dimensional data.
-
-Here, we propose the high-dimensional regularized discriminant analysis
-(\emph{HDRDA}) classifier by first constructing a pooling matrix as a convex
-combination of the covariance-matrix estimators employed in linear and quadratic
-discriminant analysis. Given that the assumption of equal population covariance
-matrices for linear discriminant analysis is often too stringent in practice
-\citep{Clemmensen:2011kr}, our proposed classifier can relax the linearity
-assumption by instead incorporating unequal covariance matrices while retaining
-linear and quadratic discriminant analysis as special cases. We then establish
-the \emph{HDRDA} classifier by applying covariance-matrix regularization with a
-parameterization that includes a large family of shrinkage methods, such as
-those from \cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and other methods
-studied by \cite{Ramey:2013ji} and \cite{Xu:2009fl}. This regularization has the
-effect of stabilizing the eigenvalues of the covariance-matrix estimators and
-ensuring that the classifier is well-posed.
+We consider the RDA classifier in an attempt to study rigorously some of its
+properties. Similar to \cite{Hastie:2008dt} and \cite{Halbe:2007ea}, we first
+reparameterize the RDA classifier's covariance-matrix estimator as a convex
+combination of the covariance-matrix estimators employed in the LDA and QDA
+classifiers. Our parameterizaton yields 1) improved interpretation, 2) nice
+properties that facilitate dimension reduction, and 3) as a result of the
+dimension reduction, efficient calculation for high-dimensional data. We then
+establish the high-dimensional RDA (\emph{HDRDA}) classifier by applying
+covariance-matrix regularization with a parameterization that includes a large
+family of shrinkage methods, such as those from \cite{Srivastava:2007ww},
+\cite{Rao:1971ul}, and other methods studied by \cite{Ramey:2013ji} and
+\cite{Xu:2009fl}. This regularization has the effect of stabilizing the
+eigenvalues of the covariance-matrix estimators and ensuring that the classifier
+is well-posed.
 
@@
 matrices, where $n_k$ is the sample size of the $k$th class, $k = 1, \ldots, K$.
-This reduction in calculation is clearly advantageous because $n_k$ is often
-quite small in microarray data sets (e.g., 10--20 observations).
+This reduction in calculation is clearly advantageous because $n_k$ can be small
+in modern high-dimensional data sets (e.g., 10--20 observations in
+gene-expression microarray data).
 
@@
 
-\section{High-Dimensional Regularized Discriminant Analysis}
-\label{sec:rda}
-
-\subsection{Preliminaries}
+\section{Preliminaries}
 To facilitate our discussion of covariance-matrix regularization and regularized
@@
 
-\subsection{Model Specification}
+\section{High-Dimensional Regularized Discriminant Analysis}
+\label{sec:rda}
 
@@
 
-\subsection{Relationship with Regularized Discriminant Analysis}
+\subsection{Relationship with Friedman's RDA Classifier}
 
@@
 of the eigenvalues of \eqref{eq:sig-lambda}. Finally, the RDA classifier is
-established by substituting \eqref{eq:sig-rda} into \eqref{eq:qda}.
+established by substituting \eqref{eq:sig-rda} into
+\eqref{eq:qda}.\footnote{TODO: Discuss $\alpha_k \equiv 1-\gamma$, etc.}
 
-\section{Properties of the HDRDA}
-
-\subsection{Enhanced Computational Efficiency}
+\subsection{Properties of the HDRDA Classifier}
 \label{sec:computational-efficiency}
@@
 \end{proof}
+
+\subsection{Computationally Efficient Model Selection}
 

### 2013-12-26 reports/paper-hdrda.Rnw
Corrected derivation of cov estimator's interpretation.
End result was correct but intermediate steps were not
Rewrote the abstract and the introduction.
- The intro is a bit verbose and needs editing.
- Related to #10 and #15.

--- 2013-12-12 reports/paper-hdrda.Rnw
@@
 
-LDA and QDA suck. Friedman to the rescue with RDA. We reformulate the RDA
-classifier. Our parameterizaton yields 1) improved interpretation, 2) nice
-properties that facilitate dimension reduction, and 3) as a result of the
-dimension reduction, efficient calculation for high-dimensional data.
-
-Here, we propose a new regularized discriminant analysis classifier that
-incorporates dimension reduction and covariance-matrix shrinkage to enable a
-computationally efficient classifier. We demonstrate with four high-dimensional
-microarray data sets that our proposed classifier is often superior to several
-recently proposed sparse and regularized classifiers in terms of classification
-accuracy. Finally, we provide an implementation of our proposed classifier in
-the {\tt sparsediscrim} R package, available on CRAN.
+The popular regularized discriminant analysis (RDA) classifier proposed by
+\citeauthor{Friedman:1989tm} utilizes a biased covariance-matrix estimator that
+partially pools the sample covariance matrices from linear and quadratic
+disriminant analysis before shrinking the resulting estimator towards a scaled
+identity matrix. The RDA classifier's two tuning parameters are typically
+estimated via a computationally burdensome cross-validation procedure from a
+grid of candidate values. We reformulate the RDA classifier in the
+high-dimensional setting and then show that the classification decision rule is
+equivalent to one in a subspace having a much lower dimension. As a result, the
+dimension reduction involved yields a substantial reduction in computation
+during model selection. Our parameterization also offers interpretability
+previously lacking with the RDA classifier. We demonstrate with three artificial
+data sets and four high-dimensional real data sets that our proposed classifier
+is often superior to several recently proposed sparse and regularized
+classifiers in terms of classification accuracy. Finally, we provide an
+implementation of our proposed classifier in the {\tt sparsediscrim} R package,
+available on CRAN.
 
@@
 
-Linear discriminant analysis (LDA) and quadratic discriminant analysis (QDA) are
-classical approaches to supervised learning and are widely used in a variety of
-areas, such as medical research, facial recognition, and TODO
-\cite{Dudoit:2002ev,TODO1,TODO2} Both the LDA and QDA classifiers are based on
-the assumption that the underlying distribution of each of the $K$ classes are
-multivariate normally distributed. While both methods are popular in practice,
-they often cannot be applied to modern data sets because the sample covariance
-matrices employed are singular when the number of features exceeds the total
-sample size $N$. For instance, standard benchmark microarray data sets used in
-biological research often consist of between $p = 10,000$ and $55,000$ probes
-and fewer than $N = 100$ observations, in which case LDA and QDA are
-incalculable.
+As machine-learning algorithms are increasingly utilized to automate data-driven
+decisions, researchers continue to develop a wealth of \emph{classifiers} to
+construct automated pattern-recognition systems in a variety of disciplines,
+such as medical research, facial recognition, remote sensing, and chemical
+analysis \citep{TODO}. Despite the advances in classification to facilitate
+technological advancements, well-established classifiers often degrade in
+accuracy with modern, high-dimensional data sets when the number of features $p$
+exceeds the total sample size $N$, due to the \emph{curse of dimensionality}
+\citep{Bellman:1961tn}. For instance, linear discriminant analysis (LDA) and
+quadratic discriminant analysis (QDA) are popular because of their simplicity,
+speed of computation, and interpretability. However, both methods are
+incalculable when $p > N$ because the class sample covariance matrices are
+singular \citep*{Bouveyron:2007gx}.
 
-\cite{Friedman:1989tm} proposed regularized discriminant analysis (RDA), which
-partially pools the individual sample covariance matrices used in the QDA
-classifier with the pooled sample covariance matrix from the LDA
-classifier. After pooling the covariance matrices, the RDA classifier employs a
-covariance-matrix regularization technique that shrinks the zero eigenvalues of
-the resulting covariance matrices to ensure they are nonsingular. Similarly,
-several researchers, including \cite*{Zhang:2010va}, \cite{Ye:2009gd},
+As a result, researchers have emphasized three standard approaches to modifying
+the \emph{LDA} and \emph{QDA} classifiers for application to high-dimensional
+data: 1) covariance-matrix regularization, 2) dimension reduction, and 3) model
+parsimony. Covariance-matrix regularization methods often are based on the
+well-known ridge-regression approach of
+\cite{Hoerl:1970cd}. \cite{Ching:2012fu}, \cite{Zhang:2010va}, \cite{Ye:2009gd},
 \cite{Ji:2008wp}, \cite*{Guo:2007te}, \cite*{Srivastava:2007ud},
-\cite{Ye:2006vx}, and \cite{Ye:2005uu}, have proposed regularization methods to
-improve the estimation of the population covariance matrices before applying
-either the LDA or QDA classifiers. Covariance-matrix regularization methods
-often are based on the well-known ridge-regression approach of
-\cite{Hoerl:1970cd} and adjust the eigenvalues of the sample covariance matrices
-to ensure positive definiteness. For excellent overviews of the regularization
-problem in discriminant analysis, see \cite{Murphy:2012uq},
-\cite{Mkhadri:1997gy}, and \cite{Friedman:1989tm}.
+\cite{Ye:2006vx}, and \cite{Ye:2005uu} have proposed regularization methods to
+improve the estimation of the class covariance matrices by adjusting the
+eigenvalues of the sample covariance matrices to ensure positive
+definiteness. Alternatively, dimension reduction typically decomposes a model to
+orthogonal components and retains the components corresponding to the model's
+intrinsic dimension. While some classificatory information may be lost,
+dimension reduction methods have been shown to be effective in practice
+\citep{YePaper,TODO}. Finally, \cite{Mai:2012bf}, \cite{Fan:2012iq},
+\cite{Merchante:2012vk}, \cite{Clemmensen:2011kr}, \cite{Huang:2010ju}, and
+\cite{Dudoit:2002ev} have proposed restrictions on the population covariance
+matrices to reduce the number of parameters that must be estimated. For
+instance, \cite{Dudoit:2002ev} have assumed the features within each class are
+independent so that the population covariance matrices are diagonal, resulting
+in positive-definite sample covariance matrices. For excellent overviews of the
+regularization problem in discriminant analysis and other techniques, see
+\cite{Murphy:2012uq} and \cite{Mkhadri:1997gy}.
 
-% TODO: Should we remove this paragraph? Keep a couple of the recent references
-% though.
-Alternatively, many researchers, including \cite{Mai:2012bf}, \cite{Fan:2012iq},
-\cite{Merchante:2012vk}, \cite{Clemmensen:2011kr}, and \cite{Huang:2010ju}, have
-introduced restrictions on the population covariance matrices in linear and
-quadratic discriminant analysis to reduce the number of covariance parameters
-that must be estimated to ensure that the resulting classifiers are
-calculable. For instance, \cite{Dudoit:2002ev} have modified the linear and
-quadratic discriminant analysis classifiers by assuming that the population
-covariance matrices are diagonal, resulting in classifiers that are special
-cases of the Naive Bayes classifier \citep[Chapter 4]{Hastie:2008dt}.
+We consider regularized discriminant analysis (\emph{RDA}) from a classic paper
+by \cite{Friedman:1989tm}, who proposed improvements to the \emph{QDA}
+classifier by employing a covariance matrix estimator that partially pools the
+individual sample covariance matrices with the pooled sample covariance matrix
+from the \emph{LDA} classifier. Given that the assumption of equal population
+covariance matrices for the \emph{LDA} classifier is often too stringent in practice
+\citep{Clemmensen:2011kr}, the \emph{RDA} classifier can relax the linearity
+assumption by instead incorporating unequal covariance matrices while retaining
+as special cases the \emph{LDA} and QDA classifiers. Covariance-matrix
+regularization is then applied to the partially pooled covariance-matrix
+estimator by shrinking it towards a scaled identity matrix to ensure positive
+definiteness. We argue that the \emph{RDA} classifier inherently employs
+dimension reduction in addition to regularization. After parameterizing the
+\emph{RDA} classifier similar to \cite{Hastie:2008dt} and \cite{Halbe:2007ea},
+we demonstrate that the intrinsic dimension of the \emph{RDA} classifier is
+equal to the training sample size of each class.\footnote{TODO: Introduce HDRDA
+  somewhere.} We then show that the \emph{RDA} decision rule is invariant to
+adjustments of the zero eigenvalues. We show that our proposed classifier is
+invariant to any adjustments of the eigenvalues corresponding to the null space
+of the pooled sample covariance matrix. Applying a reasoning similar to
+\cite{Ye:2006tq}, we decompose the decision rule of the \emph{HDRDA} classifier
+and show that the sum related to the null space of the pooled sample covariance
+matrix can be discarded. Hence, no gain in classification performance can be
+obtained by adjusting these zero eigenvalues. This invariance property provides
+a substantial increase in computational efficiency of model training and
+selection. Moreover, the decision rule in the original feature space is
+equivalent to a decision rule in a much lower dimension (i.e., the intrinsic
+dimension), such that the matrix inverses and determinants can be rapidly
+computed. This invariance property as well as the dual optimization problem
+result in substantial reduction in computation. Additionally, our
+parameterization of the \emph{RDA} classifier offers interpretability that has
+been previously lacking \citep{Bensmail:1996fw}. We explicitly show that the
+pooling parameter in the \emph{RDA} covariance-matrix estimator controls the
+contribution of each training observation to the estimation of each class
+covariance matrix. Furthermore, our parameterization includes a large family of
+shrinkage methods, such as those from \cite{Srivastava:2007ww},
+\cite{Rao:1971ul}, and other methods studied by \cite{Ramey:2013ji} and
+\cite{Xu:2009fl}.
 
-We reformulate the RDA classifier. Our parameterizaton yields 1) improved
-interpretation, 2) nice properties that facilitate dimension reduction, and 3)
-as a result of the dimension reduction, efficient calculation for
-high-dimensional data.
+Given the \emph{RDA} classifier has been shown to have excellent performance in
+the high-dimensional setting \citep{Webb:2011vu} but is limited by its
+computationally intense model selection procedure, our work enables the
+\emph{RDA} classifier to be employed with high-dimensional data in
+practice. This is reassuring because the \emph{RDA} classifier remains widely
+popular in the literature. In fact, variants of the RDA classifier have been
+applied to microarray data \citep{Ching:2012fu,Li:2012ev,Tai:2007bk,Guo:2007te},
+facial recognition
+\citep{Ching:2012fu,Zhang:2010va,Dai:2007vd,Lu:2005hq,Pima:2004vw,Lu:2003we},
+handwritten digit recognition \citep{Bouveyron:2007gx}, remote sensing
+\citep{Tadjudin:1999fk}, seismic detection \citep{Anderson:2002kg}, and chemical
+spectra \citep{Wu:1996us,Aeberhard:1993fp}.
 
-Here, we propose the high-dimensional regularized discriminant analysis
-(\emph{HDRDA}) classifier by first constructing a pooling matrix as a convex
-combination of the covariance-matrix estimators employed in linear and quadratic
-discriminant analysis. Given that the assumption of equal population covariance
-matrices for linear discriminant analysis is often too stringent in practice
-\citep{Clemmensen:2011kr}, our proposed classifier can relax the linearity
-assumption by instead incorporating unequal covariance matrices while retaining
-linear and quadratic discriminant analysis as special cases. We then establish
-the \emph{HDRDA} classifier by applying covariance-matrix regularization with a
-parameterization that includes a large family of shrinkage methods, such as
-those from \cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and other methods
-studied by \cite{Ramey:2013ji} and \cite{Xu:2009fl}. This regularization has the
-effect of stabilizing the eigenvalues of the covariance-matrix estimators and
-ensuring that the classifier is well-posed.
+After demonstrating that the \emph{HDRDA} classifier can be computed efficiently
+and establishing its improved interpretability, we compare the computational
+runtime of the \emph{HDRDA} classifier with an implementation of the \emph{RDA}
+classifier from the {\tt klaR} R package to illustrate that the \emph{HDRDA}
+classifier greatly reduces computation when $p \gg N$. We then study the
+classification performance of the \emph{HDRDA} classifier on several artificial
+and real high-dimensional data sets. We compare the \emph{HDRDA} classifier with
+several recent several recent classifiers from \cite{Dudoit:2002ev},
+\cite{Guo:2007te}, \cite{Pang:2009ik}, \cite{Tong:2012hw}, and
+\cite{Witten:2011kc}, each of which is designed for small-sample,
+high-dimensional data. We demonstrate that the \emph{HDRDA} classifier often
+attains superior classification accuracy. We also include in our classification
+study a support vector machine (\emph{SVM}) and $k$-nearest neigbors ($kNN$) as
+benchmarks.
 
-We show that our proposed classifier is invariant to any adjustments of the
-eigenvalues corresponding to the null space of the pooled sample covariance
-matrix. Applying a reasoning similar to \cite{Ye:2006tq}, we decompose the
-decision rule of the \emph{HDRDA} classifier and show that the sum related to
-the null space of the pooled sample covariance matrix can be discarded. Hence,
-no gain in classification performance can be obtained by adjusting these zero
-eigenvalues. This invariance property provides a substantial increase in
-computational efficiency of model training and selection. In fact, we show that
-the inverse and determinants can be reduced to calculations of $n_k \times n_k$
-matrices, where $n_k$ is the sample size of the $k$th class, $k = 1, \ldots, K$.
-This reduction in calculation is clearly advantageous because $n_k$ is often
-quite small in microarray data sets (e.g., 10--20 observations).
+The remainder of this document is organized as follows. In Section
+\ref{sec:preliminaries} we presents the discrimination problem and the necessary
+notation to describe our contributions. In Section \ref{sec:rda} we present our
+\emph{HDRDA} classifier along with its interpretation, properties, and
+computationally efficient model-selection procedure. Section \ref{sec:sims}
+describes our simulation studies of artificial and real data sets and also
+describes the experimental results. We conclude with a brief discussion in
+Section \ref{sec:discussion}.
 
-We also demonstrate that the \emph{HDRDA} classifier is often superior to
-several recent classifiers designed for small-sample, high-dimensional data. In
-particular, we compare our proposed classifier with diagonal linear discriminant
-analysis proposed by \cite{Dudoit:2002ev} and the recently proposed classifiers
-from \cite{Guo:2007te}, \cite{Pang:2009ik}, \cite{Tong:2012hw}, and
-\cite{Witten:2011kc}. We conduct our comparison of the classification accuracy
-of the competing classifiers using four small-sample, high-dimensional benchmark
-microarray data sets.
+\section{Preliminaries}
+\label{sec:preliminaries}
 
-\section{High-Dimensional Regularized Discriminant Analysis}
-\label{sec:rda}
+\subsection{Notation}
 
-\subsection{Preliminaries}
 To facilitate our discussion of covariance-matrix regularization and regularized
 discriminant analysis, we require the following notation. Let $\mathbb{R}_{a
-  \times b}$ denotes the matrix space of all $a \times b$ matrices over the real
+  \times b}$ denote the matrix space of all $a \times b$ matrices over the real
 field $\mathbb{R}$. Denote by $\bm I_m$ the $m \times m$ identity matrix, and
@@
 pseudoinverse, and the null space of $\bm A \in \mathbb{R}_{m \times p}$,
-respectively. Let $\mathbb{R}_{p \times p}^{>}$ denote the cone of real $p
-\times p$ positive-definite matrices. Similarly, let $\mathbb{R}_{p \times
-  p}^{\ge}$ denote the cone of real $p \times p$ positive-semidefinite
-matrices. Let $V^{\perp} = \{\bm y \in \mathbb{R}_{p \times 1} : \bm y^{T} \bm v
-= 0$ for all $\bm v \in V \}$ denote the orthogonal complement of a vector space
-$V \subset \mathbb{R}_{p \times 1}$. For $c \in \mathbb{R}$, let $c^+ = 1/c$ if
-$c \ne 0$ and $0$ otherwise.
+respectively. Denote by $\mathbb{R}_{p \times p}^{>}$ the cone of real $p \times
+p$ positive-definite matrices. Similarly, let $\mathbb{R}_{p \times p}^{\ge}$
+denote the cone of real $p \times p$ positive-semidefinite matrices. Let
+$V^{\perp} = \{\bm y \in \mathbb{R}_{p \times 1} : \bm y^{T} \bm v = 0$ for all
+$\bm v \in V \}$ denote the orthogonal complement of a vector space $V \subset
+\mathbb{R}_{p \times 1}$. For $c \in \mathbb{R}$, let $c^+ = 1/c$ if $c \ne 0$
+and $0$ otherwise.
+
+\subsection{Discriminant Analysis}
 
@@
 
-\subsection{Model Specification}
-
-\cite{Halbe:2007ea} demonstrate that  parameterization results in a
+\section{High-Dimensional Regularized Discriminant Analysis}
+\label{sec:rda}
 
@@
 1, \ldots, N)$, each centered by its class sample mean, we attain a clear
-interpretation of $\widehat{\bm \Sigma}_k(\lambda)$. That is,
+interpretation of $\widehat{\bm \Sigma}_k(\lambda)$. That is, from
+\eqref{eq:pooled-cov} we have
 \begin{align}
-	\widehat{\bm \Sigma}_k(\lambda) &= \widehat{\bm \Sigma}_k + \lambda (\widehat{\bm \Sigma} - \widehat{\bm \Sigma}_k) \nonumber\\
-%	&= \widehat{\bm \Sigma}_k + \lambda \left(\sum_{k' = 1}^K \frac{n_k}{N} \widehat{\bm \Sigma}_{k'}  - \widehat{\bm \Sigma}_k \right) \nonumber\\
-	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right) \widehat{\bm \Sigma}_k +  \frac{\lambda}{N} \sum_{\substack{k' = 1\\k' \ne k}} n_{k'} \widehat{\bm \Sigma}_{k'} \nonumber \\
-%	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N I(y_i = k) x_i x_i^{T} +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k) x_i x_i^{T} \nonumber \\
+	\widehat{\bm \Sigma}_k(\lambda)
+  &= \left( 1 - \lambda + \frac{\lambda n_k}{N} \right) \widehat{\bm \Sigma}_k +  \frac{\lambda}{N} \sum_{\substack{k' = 1\\k' \ne k}}^K n_{k'} \widehat{\bm \Sigma}_{k'} \nonumber \\
+	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N
+  I(y_i = k) \bm x_i \bm x_i^{T} +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k)
+  \bm x_i \bm x_i^{T} \nonumber \\
 	&= \sum_{i=1}^N c_{ik}(\lambda) \bm x_i \bm x_i^{T},\label{eq:sig-lambda-alternative2}
@@
 $\bm \Sigma_k$. Hence, for $0 < \lambda \le 1$, we estimate $\bm \Sigma_k$ with
-$N$ rather than $n_k$ observations. Below, we will show that the pooling
+$N$ rather than $n_k$ observations.\footnote{TODO: Add more to the
+  interpretation here.}
+
+\begin{figure}
+<<rda_weights>>=
+n <- c(50, 25, 15, 10)
+N <- sum(n)
+lambda <- seq(0, 1, length = 1000)
+
+rda_weights <- sapply(seq_along(n), function(k) {
+      lambda / N + (1 - lambda) / n[k]
+      })
+rda_weights <- data.frame(lambda = lambda, rda_weights)
+rda_weights <- melt(rda_weights, id = 'lambda')
+colnames(rda_weights) <- c("lambda", "Class", "Weight")
+levels(rda_weights$Class) <- seq_along(n)
+
+p <- ggplot(rda_weights, aes(lambda, Weight, linetype = Class, color = Class))
+p <- p + geom_line(size = 1)
+p <- p + theme_bw() + xlab(expression(lambda))
+p <- p + theme(axis.text = element_text(size = 10))
+p <- p + theme(strip.text.x = element_text(size = 16))
+p <- p + theme(legend.title = element_text(size = 16, face = "bold"))
+p <- p + theme(legend.text = element_text(size = 15))
+p <- p + theme(legend.key.size = unit(1, "cm"))
+p
+@ 
+\caption{Weights $c_{ik}(\lambda)$ of \emph{RDA} classifier as a function of the
+  pooling parameter $\lambda$ when $n_1 = 50$, $n_2 = 25$, $n_3 = 15$, and $n_4$
+  = 10.}
+\label{fig:hdrda-weights}
+\end{figure}
+
+Below, we will show that the pooling
 operation is advantageous in increasing the rank of each $\widehat{\bm
@@
 of the eigenvalues of \eqref{eq:sig-lambda}. Finally, the RDA classifier is
-established by substituting \eqref{eq:sig-rda} into \eqref{eq:qda}.
+established by substituting \eqref{eq:sig-rda} into
+\eqref{eq:qda}.\footnote{TODO: Discuss $\alpha_k \equiv 1-\gamma$, etc.}
 
-\section{Properties of the HDRDA}
-
-\subsection{Enhanced Computational Efficiency}
+\subsection{Properties of the HDRDA Classifier}
 \label{sec:computational-efficiency}
 
-Next, we demonstrate that our proposed classifier in \eqref{eq:hdrda} lends
-itself to a more efficient calculation. We decompose \eqref{eq:hdrda} into a sum
-of two components, where the first summand consists of matrix operations applied
-to low-dimensional matrices and the second summand corresponds to the null space
-of $\widehat{\bm \Sigma}$ in \eqref{eq:pooled-cov}. We show that the matrix
-operations performed on the null space of $\widehat{\bm \Sigma}$ yield constant
-quadratic forms across all classes and can be omitted. For $p \gg N$, the
-constant component involves determinants and inverses of high-dimensional
-matrices, and by ignoring these calculations, a substantial reduction in
-computational costs is achieved. Furthermore, a byproduct is that adjustments to
-the associated eigenvalues have no effect on \eqref{eq:hdrda}. Lastly, we
-utilize the singular value decomposition to calculate efficiently the eigenvalue
-decomposition of $\widehat{\bm \Sigma}$, further reducing the computational
-costs of the \emph{HDRDA} classifier.
+Next, we establish properties of the covariance-matrix estimator and the
+decision rule employed in the \emph{HDRDA} classifier. By doing so, we
+demonstrate that \eqref{eq:hdrda} lends itself to a more efficient
+calculation. We decompose \eqref{eq:hdrda} into a sum of two components, where
+the first summand consists of matrix operations applied to low-dimensional
+matrices and the second summand corresponds to the null space of $\widehat{\bm
+  \Sigma}$ in \eqref{eq:pooled-cov}. We show that the matrix operations
+performed on the null space of $\widehat{\bm \Sigma}$ yield constant quadratic
+forms across all classes and can be omitted. For $p \gg N$, the constant
+component involves determinants and inverses of high-dimensional matrices, and
+by ignoring these calculations, a substantial reduction in computational costs
+is achieved. Furthermore, a byproduct is that adjustments to the associated
+eigenvalues have no effect on \eqref{eq:hdrda}. Lastly, we utilize the singular
+value decomposition to calculate efficiently the eigenvalue decomposition of
+$\widehat{\bm \Sigma}$, further reducing the computational costs of the
+\emph{HDRDA} classifier.
 
@@
 
+\subsection{Computationally Efficient Model Selection}
+
 Conveniently, we see that the inverse and determinant of $\bm W_k$ are each in
@@
 \section{Classification Study}
+\label{sec:sims}
 
@@
 with the exception that we explicitly set prior probabilities as equal, if the
-option was available. For the \emph{HDRDA} classifier
-\eqref{eq:hdrda-decomposed}, we set $\alpha_k = 1$, $k = 1, \ldots, K$, so that
-the covariance-matrix estimator \eqref{eq:hdrda-cov} employed resembled
-\eqref{eq:ridge-estimator}. We estimated $\lambda$ from a grid of 21 equidistant
-candidate values between 0 and 1, inclusively. Similarly, we estimated $\gamma$
-from a grid consisting of the values $10^{-1}, \ldots, 10^4$, and $10^5$. We
-selected optimal estimates of $\lambda$ and $\gamma$ using 10-fold
-cross-validation \citep{Hastie:2008dt}. Each of the microarray data sets listed
-below can be found in \ref{suppA}.
+option was available. For the \emph{HDRDA} classifier in
+\eqref{eq:hdrda-decomposed}, we examine the classification performance of two
+models. For the first \emph{HDRDA} model, we set $\alpha_k = 1$, $k = 1, \ldots,
+K$, so that the covariance-matrix estimator \eqref{eq:hdrda-cov} employed
+resembled \eqref{eq:ridge-estimator}. We estimated $\lambda$ from a grid of 21
+equidistant candidate values between 0 and 1, inclusively. Similarly, we
+estimated $\gamma$ from a grid consisting of the values $10^{-1}, \ldots, 10^4$,
+and $10^5$. We selected optimal estimates of $\lambda$ and $\gamma$ using
+10-fold cross-validation \citep{Hastie:2008dt}. For the second model, we set
+$\alpha_k = 1 - \gamma$, $k = 1, \ldots, K$, to resemble
+\citeauthor{Friedman:1989tm}'s parameterization. We estimated both $\lambda$ and
+$\gamma$ from a grid of 21 equidistant candidate values between 0 and 1,
+inclusively. Each of the microarray data sets listed below can be found in
+\ref{suppA}.
 
@@
 \section{Discussion}
+\label{sec:discussion}
 

### 2013-12-26 reports/paper-hdrda.Rnw
Lots of polish applied to the introduction. Much cleaner and more concise. Closes #10 and #15

@@
 
-The popular regularized discriminant analysis (RDA) classifier proposed by
-\citeauthor{Friedman:1989tm} utilizes a biased covariance-matrix estimator that
-partially pools the sample covariance matrices from linear and quadratic
+\citeauthor{Friedman:1989tm} proposed the popular regularized discriminant
+analysis (RDA) classifier that utilizes a biased covariance-matrix estimator
+that partially pools the sample covariance matrices from linear and quadratic
 disriminant analysis before shrinking the resulting estimator towards a scaled
@@
 previously lacking with the RDA classifier. We demonstrate with three artificial
-data sets and four high-dimensional real data sets that our proposed classifier
-is often superior to several recently proposed sparse and regularized
-classifiers in terms of classification accuracy. Finally, we provide an
-implementation of our proposed classifier in the {\tt sparsediscrim} R package,
-available on CRAN.
+and four real high-dimensional data sets that our proposed classifier is often
+superior to several recently proposed sparse and regularized classifiers in
+terms of classification accuracy. Finally, we provide an implementation of our
+proposed classifier in the {\tt sparsediscrim} R package, available on CRAN.
 
@@
 such as medical research, facial recognition, remote sensing, and chemical
-analysis \citep{TODO}. Despite the advances in classification to facilitate
-technological advancements, well-established classifiers often degrade in
-accuracy with modern, high-dimensional data sets when the number of features $p$
-exceeds the total sample size $N$, due to the \emph{curse of dimensionality}
-\citep{Bellman:1961tn}. For instance, linear discriminant analysis (LDA) and
-quadratic discriminant analysis (QDA) are popular because of their simplicity,
-speed of computation, and interpretability. However, both methods are
-incalculable when $p > N$ because the class sample covariance matrices are
-singular \citep*{Bouveyron:2007gx}.
+analysis \citep{Ching:2012fu,Li:2012ev,Zhang:2010va}. Despite the advances in
+classification to facilitate technological advancements, well-established
+classifiers often degrade in accuracy with modern, high-dimensional data sets
+when the number of features $p$ exceeds the total sample size $N$, due to the
+\emph{curse of dimensionality} \citep{Bellman:1961tn}. For instance, linear
+discriminant analysis (LDA) and quadratic discriminant analysis (QDA) are
+popular because of their simplicity, speed of computation, and
+interpretability. However, both methods are incalculable when $p > N$ because
+the class sample covariance matrices are singular \citep*{Bouveyron:2007gx}.
 
@@
 data: 1) covariance-matrix regularization, 2) dimension reduction, and 3) model
-parsimony. Covariance-matrix regularization methods often are based on the
+parsimony. Covariance-matrix regularization methods are often based on the
 well-known ridge-regression approach of
@@
 eigenvalues of the sample covariance matrices to ensure positive
-definiteness. Alternatively, dimension reduction typically decomposes a model to
-orthogonal components and retains the components corresponding to the model's
-intrinsic dimension. While some classificatory information may be lost,
-dimension reduction methods have been shown to be effective in practice
-\citep{YePaper,TODO}. Finally, \cite{Mai:2012bf}, \cite{Fan:2012iq},
-\cite{Merchante:2012vk}, \cite{Clemmensen:2011kr}, \cite{Huang:2010ju}, and
-\cite{Dudoit:2002ev} have proposed restrictions on the population covariance
-matrices to reduce the number of parameters that must be estimated. For
-instance, \cite{Dudoit:2002ev} have assumed the features within each class are
-independent so that the population covariance matrices are diagonal, resulting
-in positive-definite sample covariance matrices. For excellent overviews of the
-regularization problem in discriminant analysis and other techniques, see
-\cite{Murphy:2012uq} and \cite{Mkhadri:1997gy}.
+definiteness. Alternatively, dimension reduction typically transforms the
+original data in the high-dimensional space to a subspace having a much smaller
+dimension such that redundant and noisy information is removed
+\citep{Fan:2012iq,Ye:2009gd,Bouveyron:2007gx,Ye:2006vx,Antoniadis:2003bk}. Finally,
+\cite{Mai:2012bf}, \cite{Fan:2012iq}, \cite{Merchante:2012vk},
+\cite{Clemmensen:2011kr}, \cite{Huang:2010ju}, and \cite{Dudoit:2002ev} have
+proposed restrictions on the population covariance matrices to reduce the number
+of parameters that must be estimated. For instance, \cite{Dudoit:2002ev} have
+assumed the features within each class are independent so that the population
+covariance matrices are diagonal, resulting in positive-definite sample
+covariance matrices. For excellent overviews of the regularization problem in
+discriminant analysis and other techniques, see \cite{Murphy:2012uq} and
+\cite{Mkhadri:1997gy}.
 
@@
 from the \emph{LDA} classifier. Given that the assumption of equal population
-covariance matrices for the \emph{LDA} classifier is often too stringent in practice
-\citep{Clemmensen:2011kr}, the \emph{RDA} classifier can relax the linearity
-assumption by instead incorporating unequal covariance matrices while retaining
-as special cases the \emph{LDA} and QDA classifiers. Covariance-matrix
-regularization is then applied to the partially pooled covariance-matrix
-estimator by shrinking it towards a scaled identity matrix to ensure positive
-definiteness. We argue that the \emph{RDA} classifier inherently employs
-dimension reduction in addition to regularization. After parameterizing the
-\emph{RDA} classifier similar to \cite{Hastie:2008dt} and \cite{Halbe:2007ea},
-we demonstrate that the intrinsic dimension of the \emph{RDA} classifier is
-equal to the training sample size of each class.\footnote{TODO: Introduce HDRDA
-  somewhere.} We then show that the \emph{RDA} decision rule is invariant to
-adjustments of the zero eigenvalues. We show that our proposed classifier is
-invariant to any adjustments of the eigenvalues corresponding to the null space
-of the pooled sample covariance matrix. Applying a reasoning similar to
-\cite{Ye:2006tq}, we decompose the decision rule of the \emph{HDRDA} classifier
-and show that the sum related to the null space of the pooled sample covariance
-matrix can be discarded. Hence, no gain in classification performance can be
-obtained by adjusting these zero eigenvalues. This invariance property provides
-a substantial increase in computational efficiency of model training and
-selection. Moreover, the decision rule in the original feature space is
-equivalent to a decision rule in a much lower dimension (i.e., the intrinsic
-dimension), such that the matrix inverses and determinants can be rapidly
-computed. This invariance property as well as the dual optimization problem
-result in substantial reduction in computation. Additionally, our
-parameterization of the \emph{RDA} classifier offers interpretability that has
-been previously lacking \citep{Bensmail:1996fw}. We explicitly show that the
-pooling parameter in the \emph{RDA} covariance-matrix estimator controls the
-contribution of each training observation to the estimation of each class
-covariance matrix. Furthermore, our parameterization includes a large family of
-shrinkage methods, such as those from \cite{Srivastava:2007ww},
+covariance matrices for the \emph{LDA} classifier is often too stringent in
+practice \citep{Clemmensen:2011kr}, the \emph{RDA} classifier's partial pooling
+can relax the linearity assumption by instead incorporating unequal covariance
+matrices while retaining as special cases the \emph{LDA} and QDA
+classifiers. Covariance-matrix regularization is then applied to the partially
+pooled covariance-matrix estimator by shrinking it towards a scaled identity
+matrix to ensure positive definiteness.
+
+We reparameterize the \emph{RDA} classifier similar to \cite{Hastie:2008dt} and
+\cite{Halbe:2007ea} and refer to our parameterization as the high-dimensional
+\emph{RDA} (\emph{HDRDA}) classifier. First, we show the pooling parameter in
+the \emph{HDRDA} classifier determines the contribution of each training
+observation to the estimation of each class covariance matrix, enabling
+interpretability that has been previously lacking with the \emph{RDA} classifier
+\citep{Bensmail:1996fw}. Furthermore, the \emph{HDRDA} classifier includes a
+large family of shrinkage methods, such as those from \cite{Srivastava:2007ww},
 \cite{Rao:1971ul}, and other methods studied by \cite{Ramey:2013ji} and
-\cite{Xu:2009fl}.
+\cite{Xu:2009fl}. Next, applying a reasoning similar to \cite{Ye:2006tq}, we
+show that the matrix operations corresponding to the null space of the pooled
+sample covariance matrix are redundant and can be discarded from the
+\emph{HDRDA} decision rule without loss of classificatory information. Moreover,
+because the \emph{HDRDA} decision rule is invariant to adjustments to the zero
+eigenvalues, the decision rule in the original feature space is equivalent to a
+decision rule in a much lower dimension (i.e., the intrinsic dimension), such
+that the matrix inverses and determinants of relatively small matrices can be
+rapidly computed. Thus, the dimension reduction inherent in the \emph{HDRDA}
+classifier yields a substantial reduction in computation during model selection.
 
-Given the \emph{RDA} classifier has been shown to have excellent performance in
-the high-dimensional setting \citep{Webb:2011vu} but is limited by its
-computationally intense model selection procedure, our work enables the
-\emph{RDA} classifier to be employed with high-dimensional data in
-practice. This is reassuring because the \emph{RDA} classifier remains widely
-popular in the literature. In fact, variants of the RDA classifier have been
-applied to microarray data \citep{Ching:2012fu,Li:2012ev,Tai:2007bk,Guo:2007te},
-facial recognition
-\citep{Ching:2012fu,Zhang:2010va,Dai:2007vd,Lu:2005hq,Pima:2004vw,Lu:2003we},
-handwritten digit recognition \citep{Bouveyron:2007gx}, remote sensing
-\citep{Tadjudin:1999fk}, seismic detection \citep{Anderson:2002kg}, and chemical
-spectra \citep{Wu:1996us,Aeberhard:1993fp}.
-
-After demonstrating that the \emph{HDRDA} classifier can be computed efficiently
-and establishing its improved interpretability, we compare the computational
-runtime of the \emph{HDRDA} classifier with an implementation of the \emph{RDA}
-classifier from the {\tt klaR} R package to illustrate that the \emph{HDRDA}
-classifier greatly reduces computation when $p \gg N$. We then study the
-classification performance of the \emph{HDRDA} classifier on several artificial
-and real high-dimensional data sets. We compare the \emph{HDRDA} classifier with
-several recent several recent classifiers from \cite{Dudoit:2002ev},
-\cite{Guo:2007te}, \cite{Pang:2009ik}, \cite{Tong:2012hw}, and
-\cite{Witten:2011kc}, each of which is designed for small-sample,
-high-dimensional data. We demonstrate that the \emph{HDRDA} classifier often
-attains superior classification accuracy. We also include in our classification
-study a support vector machine (\emph{SVM}) and $k$-nearest neigbors ($kNN$) as
-benchmarks.
+We provide an efficient algorithm that expedites the model-selection process for
+the \emph{HDRDA} classifier and then compare its computational runtime with an
+implementation of the \emph{RDA} classifier from the {\tt klaR} R package to
+illustrate that the \emph{HDRDA} classifier greatly reduces computation when $p
+\gg N$. We then study the classification performance of the \emph{HDRDA}
+classifier on three artificial and four real high-dimensional data sets. We
+demonstrate that the \emph{HDRDA} classifier often attains superior
+classification accuracy to several recent classifiers designed for small-sample,
+high-dimensional data from \cite{Tong:2012hw}, \cite{Witten:2011kc},
+\cite{Pang:2009ik}, \cite{Guo:2007te}, and \cite{Dudoit:2002ev}. We also include
+as benchmarks support vector machines (\emph{SVM}) and $k$-nearest neigbors
+($k$-NN) in our classification study.\footnote{TODO: If we change any of the
+  classifiers in the simulations, they should be updated here.}
 
 The remainder of this document is organized as follows. In Section
-\ref{sec:preliminaries} we presents the discrimination problem and the necessary
-notation to describe our contributions. In Section \ref{sec:rda} we present our
+\ref{sec:preliminaries} we present the discrimination problem and the necessary
+notation to describe our contributions. In Section \ref{sec:rda} we present the
 \emph{HDRDA} classifier along with its interpretation, properties, and
@@
 $N$ rather than $n_k$ observations.\footnote{TODO: Add more to the
-  interpretation here.}
+  interpretation here and reference Figure \ref{fig:hdrda-weights}.}
 
@@
   pooling parameter $\lambda$ when $n_1 = 50$, $n_2 = 25$, $n_3 = 15$, and $n_4$
-  = 10.}
+  = 10. TODO: Scale down figure.}
 \label{fig:hdrda-weights}
@@
 
+Given the \emph{RDA} classifier has been shown to have excellent performance in
+the high-dimensional setting \citep{Webb:2011vu} but is limited by its
+computationally intense model selection procedure, our work enables the
+\emph{RDA} classifier to be employed with small-sample, high-dimensional data in
+practice. This is reassuring because the \emph{RDA} classifier remains widely
+popular in the literature. In fact, variants of the RDA classifier have been
+applied to microarray data \citep{Ching:2012fu,Li:2012ev,Tai:2007bk,Guo:2007te},
+facial recognition
+\citep{Ching:2012fu,Zhang:2010va,Dai:2007vd,Lu:2005hq,Pima:2004vw,Lu:2003we},
+handwritten digit recognition \citep{Bouveyron:2007gx}, remote sensing
+\citep{Tadjudin:1999fk}, seismic detection \citep{Anderson:2002kg}, and chemical
+spectra \citep{Wu:1996us,Aeberhard:1993fp}.
+
 \bibliographystyle{plainnat}

### 2013-12-28 reports/paper-hdrda.Rnw
Added a reference to one of Ye's dimension-reduction papers.
Minor edits.

@@
 dimension such that redundant and noisy information is removed
-\citep{Fan:2012iq,Ye:2009gd,Bouveyron:2007gx,Ye:2006vx,Antoniadis:2003bk}. Finally,
+\citep{Fan:2012iq,Ye:2009gd,Bouveyron:2007gx,Ye:2006vx,Ye:2005hd,Antoniadis:2003bk}. Finally,
 \cite{Mai:2012bf}, \cite{Fan:2012iq}, \cite{Merchante:2012vk},
@@
 where $\xbar_k$ and $\widehat{\bm \Sigma}_k$ are the maximum likelihood
-estimators of $\bm \mu_k$ and $\bm \Sigma_k$, respectively. If we assume further
-that $\bm \Sigma_k = \bm \Sigma$, $k = 1, \ldots, K$, then the pooled sample
-covariance matrix $\widehat{\bm \Sigma}$ is substituted for $\widehat{ \bm
-  \Sigma}_k$ in \eqref{eq:qda}, where
+estimators (MLEs) of $\bm \mu_k$ and $\bm \Sigma_k$, respectively. If we assume
+further that $\bm \Sigma_k = \bm \Sigma$, $k = 1, \ldots, K$, then the pooled
+sample covariance matrix $\widehat{\bm \Sigma}$ is substituted for $\widehat{
+  \bm \Sigma}_k$ in \eqref{eq:qda}, where
 \begin{align}
@@
 \end{align}
-is the maximum likelihood estimator for $\bm \Sigma$. In this case \eqref{eq:qda}
-reduces to the sample linear discriminant analysis classifier, and the
-log-determinant is omitted because it is constant across the $K$ classes.
+is the MLE for $\bm \Sigma$. In this case \eqref{eq:qda} reduces to the sample
+linear discriminant analysis classifier, and the log-determinant is omitted
+because it is constant across the $K$ classes.
 
@@
 \begin{lemma}\label{lemma:null-spaces}
-Let $\widehat{\bm \Sigma}_k$ and $\widehat{\bm \Sigma}$ be the maximum
-likelihood estimators of $\bm \Sigma_k$ and $\bm \Sigma$, respectively. Let
-$\widehat{\bm \Sigma}_k(\lambda)$ be defined as in
-\eqref{eq:sig-lambda-alternative}. Then, $\mathcal{N}\{\widehat{\bm \Sigma}_k(\lambda)\} \subset \mathcal{N}(\widehat{
+Let $\widehat{\bm \Sigma}_k$ and $\widehat{\bm \Sigma}$ be the MLEs of $\bm
+\Sigma_k$ and $\bm \Sigma$, respectively. Let $\widehat{\bm \Sigma}_k(\lambda)$
+be defined as in \eqref{eq:sig-lambda-alternative}. Then,
+$\mathcal{N}\{\widehat{\bm \Sigma}_k(\lambda)\} \subset \mathcal{N}(\widehat{
   \bm \Sigma}) \subset \mathcal{N}(\widehat{\bm \Sigma}_k)$, $k = 1, \ldots, K$.
@@
 invariant to the term $\bm U_2$, thus yielding an equivalent classifier in
-\eqref{eq:hdrda-decomposed} with a substantial reduction in the computational
+\eqref{eq:hdrda-decomposed} with a substantial reduction in computational
 complexity. Here, we demonstrate that the computational efficiency in
-calculating the inverse and determinant of $\bm W_k$ can be further improved
-via standard matrix operations.
+calculating the inverse and determinant of $\bm W_k$ can be further improved via
+standard matrix operations.
 

### 2013-12-29 reports/paper-hdrda.Rnw
Initial version of algorithm added. Moving forward on #11

@@
 \usepackage{natbib, setspace}
+\usepackage{algorithm2e}
 
@@
 		\bm W_k^{-1} &= \bm \Gamma_k^{-1} - \alpha_k(1 - \lambda) \bm \Gamma_k^{-1} \bm U_1^{T} \bm X_k^{T} \bm Q_k^{-1} \bm X_k \bm U_1 \bm \Gamma_k^{-1},\label{eq:W_k_inv}
+    \intertext{where}
+    \bm Q_k &= \bm I_{n_k} + \alpha_k(1 - \lambda) \bm X_k \bm U_1 \bm \Gamma_k^{-1} \bm U_1^{T} \bm X_k^{T}\label{eq:Q_k}
+    \intertext{and}
+    \bm \Gamma_k &= \alpha_k \lambda \bm D_q + \gamma \bm I_q\label{eq:Gamma_k}.
 	\end{align}
-where $\bm Q_k = \bm I_{n_k} + \alpha_k(1 - \lambda) \bm X_k \bm U_1 
-\bm \Gamma_k^{-1} \bm U_1^{T} \bm X_k^{T}$ and $\bm \Gamma_k = \alpha_k \lambda \bm D_q +
-\gamma \bm I_q$.
 \end{proposition}
@@
 \citep{Hastie:2008dt}. Recall that the pooled sample covariance matrix is
-$\widehat{\bm \Sigma} = \bm X^{T} \bm X$. Then, because $\bm X = \bm M \bm
+$\widehat{\bm \Sigma} = \bm X^{T} \bm X$\footnote{TODO: $1/N$ is missing. Add
+  this. How much do things change below?} Then, because $\bm X = \bm M \bm
 D^{1/2} \bm U^{T}$ by the singular value decomposition, we have the
@@
 first $q$ columns of $\bm U$.
+
+\IncMargin{1em}
+\begin{algorithm}
+  \SetKwData{Left}{left}
+  \SetKwData{This}{this}
+  \SetKwData{Up}{up}
+  \SetKwFunction{Union}{Union}
+  \SetKwFunction{FindCompress}{FindCompress}
+  \SetKwInOut{Input}{input}
+  \SetKwInOut{Output}{output}
+
+  \Input{
+    Data matrix $\bm X$\\
+    Parameter grid $\{\lambda_g\}_{g=1}^G \times \{\gamma_h\}_{h=1}^H$
+  }
+  \Output{Optimal Estimates, $\hat{\lambda}$ and $\hat{\gamma}$}
+  \BlankLine
+  \For{$v \leftarrow 1$ \KwTo $V$ \tcp*[r]{$V$-fold cross-validation}} {
+    Partition $v$th fold into $\bm X_{train} \in \mathbb{R}_{N \times p}$ and
+    $\bm X_{test} \in \mathbb{R}_{N_T \times p}$
+
+    \For{$k\leftarrow 1$ \KwTo $K$}{
+      Extract $\bm X_k \in \mathbb{R}_{n_k \times p}$ from $\bm X_{train}$
+
+      Compute sample mean $\xbar_k$ from $\bm X_k$
+
+      Center $\bm X_k \leftarrow \bm X_k - \bm j_{n_k} \xbar_k^T$    
+    }
+
+    $\bm X_c \leftarrow [\bm X_1^T, \ldots, \bm X_K^T]^T$
+
+    Compute $\bm X_c = \bm M_q \bm D_q \bm U_1^T$
+
+    Transform $\bm X_c \leftarrow \bm X_c \bm U_1$
+
+    Transform $\bm X_{test} \leftarrow \bm X_{test} \bm U_1$
+
+    \For{$k \leftarrow 1$ \KwTo $K$} {
+      Extract $\bm X_k \in \mathbb{R}_{n_k \times q}$ from $\bm X_c$
+
+      Recompute sample mean $\xbar_k$ from $\bm X_k$
+    }
+
+    \For{$(\lambda, \gamma) \in \{\lambda_g\}_{g=1}^G \times \{\gamma_h\}_{h=1}^H$} {
+      \For{$k \leftarrow 1$ \KwTo $K$} {      
+        Compute $\bm Q_k$ using \eqref{eq:Q_k}
+
+        Compute $\bm \Gamma_k$ using \eqref{eq:Gamma_k}
+
+        Compute $\bm W_k^{-1}$ using \eqref{eq:W_k_inv}
+
+        Compute $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$
+
+        Compute $(\bm x - \xbar_k)^{T} \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm x - \xbar_k) + \log | \bm W_k |$ \tcp*[r]{TODO: Use matrix form for test observations}
+      }
+      Classify test observations $\bm X_{test}$ using \eqref{eq:hdrda-decomposed}
+
+      Compute $\#\{\text{Error}_v(\lambda, \gamma)\} \leftarrow$ misclassified test observations
+    }   
+   }
+  Compute $\widehat{\text{Error}}(\lambda, \gamma) = N^{-1} \sum_{v=1}^V \#\{\text{Error}_v(\lambda, \gamma)\}$
+
+  Report optimal $(\hat{\lambda}, \hat{\gamma}) \leftarrow \argmax_{(\lambda, \gamma)} \widehat{\text{Error}}(\lambda, \gamma)$
+  \caption{Model selection for the \emph{HDRDA} classifier}
+  \label{algo_hdrda_model_selection}
+\end{algorithm}
+\DecMargin{1em}
+
+The compact SVD is sometimes called the reduced or fast SVD.
+
+Let $\bm j_m \in \mathbb{R}_{m \times 1}$ denote a vector of ones.
 
@@
 
-We used version 3.0.0 of the open-source statistical software {\tt R} for the
+We used version 3.0.2 of the open-source statistical software {\tt R} for the
 classification study. We used the implementations of the Witten and Guo
@@
 
+The dimension reduction that we have employed in this paper has reduced the
+dimension to the rank$(\widehat{\bm \Sigma})$, $q$. An interesting extension of
+our work would reduce the dimension $q$ further to a lower dimension $q_L < q$
+using some criterion proposed, perhaps similar to that employed in principal
+components analysis. While it is not clear whether classification performance
+would improve via such a method, the efficiency of the training and model
+selection would certainly improve. Moreover, if $q_L = 2$ or $3$ was selected,
+the potential for low-dimensional graphical displays of high-dimensional could
+be accomplished.
+
 \bibliographystyle{plainnat}

### 2013-12-29 reports/paper-hdrda.Rnw
Umm, let's minimize cv error rather than maximize it.
Initial introduction of model-selection algorithm. Also, added a shell of things to close out #11

@@
 let $\bm 0_{m \times p}$ be the $m \times p$ matrix of zeros, such that $\bm
-0_m$ is understood to denote $\bm 0_{m \times m}$. Let $\bm A \oplus \bm B$
-denote the direct sum of $\bm A \in \mathbb{R}_{r \times r}$ and $\bm B \in
+0_m$ is understood to denote $\bm 0_{m \times m}$. Define $\bm 1_m \in
+\mathbb{R}_{m \times 1}$ as a vector of ones. Let $\bm A \oplus \bm B$ denote
+the direct sum of $\bm A \in \mathbb{R}_{r \times r}$ and $\bm B \in
 \mathbb{R}_{s \times s}$ \citep[Chapter 1]{Lutkepohl:1996uz}. Let $\bm A^{T}$,
@@
 \subsection{Properties of the HDRDA Classifier}
-\label{sec:computational-efficiency}
+\label{sec:hdrda-properties}
 
@@
 
-\subsection{Computationally Efficient Model Selection}
+\subsection{Model Selection}
 
-Conveniently, we see that the inverse and determinant of $\bm W_k$ are each in
-terms of the inverse and determinant of $\bm Q_k \in \mathbb{R}_{n_k \times
-  n_k}$, respectively, as well as the diagonal matrix $\bm \Gamma_k \in
-\mathbb{R}_{q \times q}$. Hence, the resulting formulation requires $O(n_k p q)$
-calculations to compute both $\bm W_k^{-1}$ and $| \bm W_k |$ and is especially
-advantageous when the number of classes $K$ is large because $n_k$ is then
-potentially much smaller than $q$. The expressions given in Proposition
-\ref{proposition:hdrda-W_k} also expedite the selection of $\lambda$ and
-$\gamma$ via cross-validation because the most time-consuming matrix operation
-involved in computing $\bm W_k^{-1}$ and $|\bm W_k|$ is $\bm X_k \bm U_1 \in
-\mathbb{R}_{n_k \times q}$, which is independent of $\lambda$ and
-$\gamma$. Therefore, we need calculate $\bm X_k \bm U_1$ only once for each
-cross-validation fold. The subsequent operations in calculating $\bm W_k^{-1}$
-and $|\bm W_K|$ can be simply updated for different pairs of $\lambda$ and
-$\gamma$.
-
-Thus far, we have constructed each $\tilde{\bm \Sigma}_k$ in terms of the first
-$q$ eigenvectors of $\widehat{\bm \Sigma}$. By utilizing a technique similar to
-the so-called kernel trick, we can obtain $\bm U_1$ by computing the
-eigendecomposition of a much smaller $N \times N$ matrix when $p \gg N$
-\citep{Hastie:2008dt}. Recall that the pooled sample covariance matrix is
-$\widehat{\bm \Sigma} = \bm X^{T} \bm X$\footnote{TODO: $1/N$ is missing. Add
-  this. How much do things change below?} Then, because $\bm X = \bm M \bm
-D^{1/2} \bm U^{T}$ by the singular value decomposition, we have the
-eigendecomposition $\widehat{\bm \Sigma} = \bm U \bm D \bm U^{T}$, where $\bm M
-\in \mathbb{R}_{N \times N}$ is orthogonal, $\bm D^{1/2} \in \mathbb{R}_{N
-  \times N}^{\ge}$ is a diagonal matrix consisting of the singular values of
-$\bm X$, and $\bm U \in \mathbb{R}_{p \times N}$ is orthogonal. Now, $\bm X \bm
-X^{T} = \bm M \bm D \bm M^{T}$ is the eigenvalue decomposition of $\bm X \bm
-X^{T}$, where $\bm D$ is the matrix of eigenvalues of $\bm X \bm X^{T}$ and the
-columns of $\bm M$ are the corresponding eigenvectors. Hence, we obtain $\bm M$
-and $\bm D$ from an eigendecomposition of $\bm X \bm X^{T} \in \mathbb{R}_{N
-  \times N}$. Next, we compute $\bm U = \bm X^{T} \bm M \bm D^{+/2}$, where $\bm
-D^{+/2} = \bm D_q^{-1/2} \oplus \bm 0_{N-q}$. We then determine $q$, the number
-of numerically nonzero eigenvalues present in $\bm D$, by calculating the number
-of eigenvalues that exceeds some tolerance value, say, $\epsilon = 1 \times
-10^{-6}$. After calculating $q$, we extract $\bm U_1$ by retaining only the
-first $q$ columns of $\bm U$.
+Thus far, we have presented the \emph{HDRDA} classifier and its properties that
+facilitate an efficient calculation of the decision rule. Here, we describe in
+an efficient model-selection procedure to select the optimal estimates for the
+tuning parameters $\lambda$ and $\gamma$. We provide pseudocode for the model
+selection in Algorithm \ref{alg:hdrda_model_selection}. Our goal is to select
+the optimal tuning parameter estimates from the Cartesian product of candidate
+values, $\{\lambda_g\}_{g=1}^G \times \{\gamma_h\}_{h=1}^H$. We utilize $V$-fold
+cross-validation to estimate the conditional error rate for each candidate
+tuning-parameter pair and select the pair that attains the minimum error
+rate. To calculate the $V$-fold cross-validation, we partition the original
+training data into $V$ mutually exclusive and exhaustive folds that have
+approximately the same number of observations. Then, for $v = 1, \ldots, V$, we
+classify the observations in the $v$th fold by training a classifier on the
+remaining $V - 1$ folds for each candidate pair of tuning parameters. We then
+calculate the proportion of misclassified observations across the $V$ folds and
+select the pair $(\widehat{\lambda}, \widehat{\gamma})$ that minimizes the
+considered training error-rate estimator.
 
@@
 \begin{algorithm}
-  \SetKwData{Left}{left}
-  \SetKwData{This}{this}
-  \SetKwData{Up}{up}
-  \SetKwFunction{Union}{Union}
-  \SetKwFunction{FindCompress}{FindCompress}
   \SetKwInOut{Input}{input}
@@
   }
-  \Output{Optimal Estimates, $\hat{\lambda}$ and $\hat{\gamma}$}
+  \Output{Optimal Estimates $(\hat{\lambda}, \hat{\gamma})$}
   \BlankLine
-  \For{$v \leftarrow 1$ \KwTo $V$ \tcp*[r]{$V$-fold cross-validation}} {
-    Partition $v$th fold into $\bm X_{train} \in \mathbb{R}_{N \times p}$ and
+  \For{$v \leftarrow 1$ \KwTo $V$} {
+    Partition $\bm X$ into $\bm X_{train} \in \mathbb{R}_{N \times p}$ and
     $\bm X_{test} \in \mathbb{R}_{N_T \times p}$
@@
 
-      Center $\bm X_k \leftarrow \bm X_k - \bm j_{n_k} \xbar_k^T$    
+      Center $\bm X_k \leftarrow \bm X_k - \bm 1_{n_k} \xbar_k^T$    
     }
@@
 
-  Report optimal $(\hat{\lambda}, \hat{\gamma}) \leftarrow \argmax_{(\lambda, \gamma)} \widehat{\text{Error}}(\lambda, \gamma)$
+  Report optimal $(\hat{\lambda}, \hat{\gamma}) \leftarrow \argmin_{(\lambda, \gamma)} \widehat{\text{Error}}(\lambda, \gamma)$
   \caption{Model selection for the \emph{HDRDA} classifier}
-  \label{algo_hdrda_model_selection}
+  \label{alg:hdrda_model_selection}
 \end{algorithm}
@@
 
-The compact SVD is sometimes called the reduced or fast SVD.
+% TODO: Now that the algorithm has been introduced and described, we need to
+% discuss the details. Let's remove some of the bloat where we talk about the
+% kernel trick, etc. Simply discuss how this technique is fast and cite Hastie's
+% book.
 
-Let $\bm j_m \in \mathbb{R}_{m \times 1}$ denote a vector of ones.
+A primary contributing factor to the efficiency of Algorithm
+\ref{alg:hdrda_model_selection} is our usage of the compact singular value
+decomposition (SVD) (sometimes called the reduced or fast SVD), which we
+describe here. Rather than computing the eigenvalue decomposition of
+$\widehat{\bm \Sigma}$ to obtain $\bm U_1$, we instead obtain $\bm U_1$ by
+computing the eigendecomposition of a much smaller $N \times N$ matrix when $p
+\gg N$ \citep{Hastie:2008dt}. Recall that the pooled sample covariance matrix is
+$\widehat{\bm \Sigma} = \bm X^{T} \bm X$\footnote{TODO: $1/N$ is missing. Add
+  this. It will be absorbed into the eigenvalues.}  Then, because $\bm X = \bm M
+\bm D^{1/2} \bm U^{T}$ by the SVD, we have the eigendecomposition $\widehat{\bm
+  \Sigma} = \bm U \bm D \bm U^{T}$, where $\bm M \in \mathbb{R}_{N \times N}$ is
+orthogonal, $\bm D^{1/2} \in \mathbb{R}_{N \times N}^{\ge}$ is a diagonal matrix
+consisting of the singular values of $\bm X$, and $\bm U \in \mathbb{R}_{p
+  \times N}$ is orthogonal. Now, $\bm X \bm X^{T} = \bm M \bm D \bm M^{T}$ is
+the eigenvalue decomposition of $\bm X \bm X^{T}$, where $\bm D$ is the matrix
+of eigenvalues of $\bm X \bm X^{T}$ and the columns of $\bm M$ are the
+corresponding eigenvectors. Hence, we obtain $\bm M$ and $\bm D$ from an
+eigendecomposition of $\bm X \bm X^{T} \in \mathbb{R}_{N \times N}$. Next, we
+compute $\bm U = \bm X^{T} \bm M \bm D^{+/2}$, where $\bm D^{+/2} = \bm
+D_q^{-1/2} \oplus \bm 0_{N-q}$. We then determine $q$, the number of numerically
+nonzero eigenvalues present in $\bm D$, by calculating the number of eigenvalues
+that exceeds some tolerance value, say, $\epsilon = 1 \times 10^{-6}$. After
+calculating $q$, we extract $\bm U_1$ by retaining only the first $q$ columns of
+$\bm U$.
+
+% TODO: Discuss things that need only be computed once
+% TODO: Discuss operations independent of tuning parameters
+
+\subsection{Computational Complexity}
+
+Conveniently, we see that the inverse and determinant of $\bm W_k$ are each in
+terms of the inverse and determinant of $\bm Q_k \in \mathbb{R}_{n_k \times
+  n_k}$, respectively, as well as the diagonal matrix $\bm \Gamma_k \in
+\mathbb{R}_{q \times q}$. Hence, the resulting formulation requires $O(n_k p q)$
+calculations to compute both $\bm W_k^{-1}$ and $| \bm W_k |$. The expressions
+given in Proposition \ref{proposition:hdrda-W_k} also expedite the selection of
+$\lambda$ and $\gamma$ via cross-validation because the most time-consuming
+matrix operation involved in computing $\bm W_k^{-1}$ and $|\bm W_k|$ is $\bm
+X_k \bm U_1 \in \mathbb{R}_{n_k \times q}$, which is independent of $\lambda$
+and $\gamma$. Therefore, we need calculate $\bm X_k \bm U_1$ only once for each
+cross-validation fold. The subsequent operations in calculating $\bm W_k^{-1}$
+and $|\bm W_K|$ can be simply updated for different pairs of $\lambda$ and
+$\gamma$.
+
+\begin{enumerate}
+  \item $\bm W_k^{-1}$\footnote{TODO: Need complexity for the following}
+  \item $|\bm W_k|$
+  \item (12)
+\end{enumerate}
 

### 2013-12-29 reports/paper-hdrda.Rnw
The model-selection section is described in more detail. Needs a little more polish before closing out #11

@@
 \usepackage{algorithm2e}
+
+\bibpunct{(}{)}{;}{a}{,}{,}
 
@@
 computing the eigendecomposition of a much smaller $N \times N$ matrix when $p
-\gg N$ \citep{Hastie:2008dt}. Recall that the pooled sample covariance matrix is
-$\widehat{\bm \Sigma} = \bm X^{T} \bm X$\footnote{TODO: $1/N$ is missing. Add
-  this. It will be absorbed into the eigenvalues.}  Then, because $\bm X = \bm M
-\bm D^{1/2} \bm U^{T}$ by the SVD, we have the eigendecomposition $\widehat{\bm
-  \Sigma} = \bm U \bm D \bm U^{T}$, where $\bm M \in \mathbb{R}_{N \times N}$ is
-orthogonal, $\bm D^{1/2} \in \mathbb{R}_{N \times N}^{\ge}$ is a diagonal matrix
-consisting of the singular values of $\bm X$, and $\bm U \in \mathbb{R}_{p
-  \times N}$ is orthogonal. Now, $\bm X \bm X^{T} = \bm M \bm D \bm M^{T}$ is
-the eigenvalue decomposition of $\bm X \bm X^{T}$, where $\bm D$ is the matrix
-of eigenvalues of $\bm X \bm X^{T}$ and the columns of $\bm M$ are the
-corresponding eigenvectors. Hence, we obtain $\bm M$ and $\bm D$ from an
-eigendecomposition of $\bm X \bm X^{T} \in \mathbb{R}_{N \times N}$. Next, we
-compute $\bm U = \bm X^{T} \bm M \bm D^{+/2}$, where $\bm D^{+/2} = \bm
-D_q^{-1/2} \oplus \bm 0_{N-q}$. We then determine $q$, the number of numerically
-nonzero eigenvalues present in $\bm D$, by calculating the number of eigenvalues
-that exceeds some tolerance value, say, $\epsilon = 1 \times 10^{-6}$. After
-calculating $q$, we extract $\bm U_1$ by retaining only the first $q$ columns of
-$\bm U$.
-
-% TODO: Discuss things that need only be computed once
-% TODO: Discuss operations independent of tuning parameters
-
-\subsection{Computational Complexity}
-
-Conveniently, we see that the inverse and determinant of $\bm W_k$ are each in
-terms of the inverse and determinant of $\bm Q_k \in \mathbb{R}_{n_k \times
-  n_k}$, respectively, as well as the diagonal matrix $\bm \Gamma_k \in
-\mathbb{R}_{q \times q}$. Hence, the resulting formulation requires $O(n_k p q)$
-calculations to compute both $\bm W_k^{-1}$ and $| \bm W_k |$. The expressions
-given in Proposition \ref{proposition:hdrda-W_k} also expedite the selection of
-$\lambda$ and $\gamma$ via cross-validation because the most time-consuming
-matrix operation involved in computing $\bm W_k^{-1}$ and $|\bm W_k|$ is $\bm
-X_k \bm U_1 \in \mathbb{R}_{n_k \times q}$, which is independent of $\lambda$
-and $\gamma$. Therefore, we need calculate $\bm X_k \bm U_1$ only once for each
-cross-validation fold. The subsequent operations in calculating $\bm W_k^{-1}$
-and $|\bm W_K|$ can be simply updated for different pairs of $\lambda$ and
-$\gamma$.
-
-\begin{enumerate}
-  \item $\bm W_k^{-1}$\footnote{TODO: Need complexity for the following}
-  \item $|\bm W_k|$
-  \item (12)
-\end{enumerate}
+\gg N$ \cite[Chapter~18.3.5]{Hastie:2008dt}. Recall that the pooled sample
+covariance matrix is $\widehat{\bm \Sigma} = \bm X^{T} \bm X$\footnote{TODO:
+  $1/N$ is missing. Add this. It will be absorbed into the eigenvalues.}  Then,
+because $\bm X = \bm M \bm D^{1/2} \bm U^{T}$ by the SVD, we have the
+eigendecomposition $\widehat{\bm \Sigma} = \bm U \bm D \bm U^{T}$, where $\bm M
+\in \mathbb{R}_{N \times N}$ is orthogonal, $\bm D^{1/2} \in \mathbb{R}_{N
+  \times N}^{\ge}$ is a diagonal matrix consisting of the singular values of
+$\bm X$, and $\bm U \in \mathbb{R}_{p \times N}$ is orthogonal. Now, $\bm X \bm
+X^{T} = \bm M \bm D \bm M^{T}$ is the eigenvalue decomposition of $\bm X \bm
+X^{T}$, where $\bm D$ is the matrix of eigenvalues of $\bm X \bm X^{T}$ and the
+columns of $\bm M$ are the corresponding eigenvectors. Hence, we obtain $\bm M$
+and $\bm D$ from an eigendecomposition of $\bm X \bm X^{T} \in \mathbb{R}_{N
+  \times N}$. Next, we compute $\bm U = \bm X^{T} \bm M \bm D^{+/2}$, where $\bm
+D^{+/2} = \bm D_q^{-1/2} \oplus \bm 0_{N-q}$. We then determine $q$, the number
+of numerically nonzero eigenvalues present in $\bm D$, by calculating the number
+of eigenvalues that exceeds some tolerance value, say, $\epsilon = 1 \times
+10^{-6}$. After calculating $q$, we extract $\bm U_1$ by retaining only the
+first $q$ columns of $\bm U$.
+
+As a result of the compact SVD, we need calculate $\bm X_c \bm U_1$ only once
+per cross-validation fold, requiring $O(p q N) \approx (p N^2)$ calculations.
+Hence, the computational costs of expensive calculations, such as matrix
+inverses and determinants, are greatly reduced because they are performed in the
+$q$-dimensional subspace. Similarly, we reduce the dimension of the test data
+set by calculating $\bm X_c \bm U_1$ once per fold. Conveniently, we see that
+the most costly computation involved in $\bm Q_k$ and $\bm W_k^{-1}$ is $\bm X_k
+\bm U_1$, which can be extracted from $\bm X_c \bm U_1$. Thus, after the initial
+calculation of $\bm X_c \bm U_1$ per cross-validation fold, $\bm Q_k$ requires
+$O(TODO)$ operations. Because $\bm Q_k \in \mathbb{R}_{n_k \times n_k}$, both
+its determinant and inverse requires $O(n_k^3)$ operations. This is especially
+advantageous when $K$ is large, such that $n_k$ is much smaller than $N$. Thus,
+$\bm W_k^{-1}$ requires $O(TODO)$ operations Also, the inverse and determinant
+of the diagonal matrix $\bm \Gamma_k \in \mathbb{R}_{q \times q}$ requires
+$O(q)$ operations. Hence, the $| \bm W_k |$ requires $O(q n_k^3)$ operations,
+which is efficient when $p >> N$.
+
+The expressions given in Proposition \ref{proposition:hdrda-W_k} also expedite
+the selection of $\lambda$ and $\gamma$ via cross-validation because the most
+time-consuming matrix operation involved in computing $\bm W_k^{-1}$ and $|\bm
+W_k|$ is $\bm X_k \bm U_1 \in \mathbb{R}_{n_k \times q}$, which is independent
+of $\lambda$ and $\gamma$. The subsequent operations in calculating $\bm
+W_k^{-1}$ and $|\bm W_K|$ can be simply updated for different pairs of $\lambda$
+and $\gamma$.\footnote{TODO: Need to clean up this paragraph so that it flows
+  from above. Also, elaborate.}
 
@@
 and $10^5$. We selected optimal estimates of $\lambda$ and $\gamma$ using
-10-fold cross-validation \citep{Hastie:2008dt}. For the second model, we set
-$\alpha_k = 1 - \gamma$, $k = 1, \ldots, K$, to resemble
+$V$-fold cross-validation with $V = 10$ \citep{Hastie:2008dt}. For the second
+model, we set $\alpha_k = 1 - \gamma$, $k = 1, \ldots, K$, to resemble
 \citeauthor{Friedman:1989tm}'s parameterization. We estimated both $\lambda$ and

### 2013-12-30 reports/paper-hdrda.Rnw
Add some polish to model-selection discussion. Still need a little more. #11

@@
 Thus far, we have presented the \emph{HDRDA} classifier and its properties that
-facilitate an efficient calculation of the decision rule. Here, we describe in
-an efficient model-selection procedure to select the optimal estimates for the
-tuning parameters $\lambda$ and $\gamma$. We provide pseudocode for the model
-selection in Algorithm \ref{alg:hdrda_model_selection}. Our goal is to select
-the optimal tuning parameter estimates from the Cartesian product of candidate
-values, $\{\lambda_g\}_{g=1}^G \times \{\gamma_h\}_{h=1}^H$. We utilize $V$-fold
-cross-validation to estimate the conditional error rate for each candidate
-tuning-parameter pair and select the pair that attains the minimum error
-rate. To calculate the $V$-fold cross-validation, we partition the original
-training data into $V$ mutually exclusive and exhaustive folds that have
-approximately the same number of observations. Then, for $v = 1, \ldots, V$, we
-classify the observations in the $v$th fold by training a classifier on the
-remaining $V - 1$ folds for each candidate pair of tuning parameters. We then
-calculate the proportion of misclassified observations across the $V$ folds and
-select the pair $(\widehat{\lambda}, \widehat{\gamma})$ that minimizes the
-considered training error-rate estimator.
+facilitate an efficient calculation of the decision rule. Here, we describe an
+efficient model-selection procedure along with pseudocode in Algorithm
+\ref{alg:hdrda_model_selection} to select the optimal tuning-parameter estimates
+from the Cartesian product of candidate values $\{\lambda_g\}_{g=1}^G \times
+\{\gamma_h\}_{h=1}^H$. We estimate the $V$-fold cross-validation error rate for
+each candidate pair and select $(\widehat{\lambda}, \widehat{\gamma})$ that
+attains the minimum error rate. To calculate the $V$-fold cross-validation, we
+partition the original training data into $V$ mutually exclusive and exhaustive
+folds that have approximately the same number of observations. Then, for $v = 1,
+\ldots, V$, we classify the observations in the $v$th fold by training a
+classifier on the remaining $V - 1$ folds. We then calculate the
+cross-validation error as the proportion of misclassified observations across
+the $V$ folds.
 
@@
 
-    Compute $\bm X_c = \bm M_q \bm D_q \bm U_1^T$
+    Compute the compact SVD $\bm X_c = \bm M_q \bm D_q \bm U_1^T$
 
@@
 
-% TODO: Now that the algorithm has been introduced and described, we need to
-% discuss the details. Let's remove some of the bloat where we talk about the
-% kernel trick, etc. Simply discuss how this technique is fast and cite Hastie's
-% book.
-
 A primary contributing factor to the efficiency of Algorithm
 \ref{alg:hdrda_model_selection} is our usage of the compact singular value
-decomposition (SVD) (sometimes called the reduced or fast SVD), which we
-describe here. Rather than computing the eigenvalue decomposition of
-$\widehat{\bm \Sigma}$ to obtain $\bm U_1$, we instead obtain $\bm U_1$ by
-computing the eigendecomposition of a much smaller $N \times N$ matrix when $p
-\gg N$ \cite[Chapter~18.3.5]{Hastie:2008dt}. Recall that the pooled sample
-covariance matrix is $\widehat{\bm \Sigma} = \bm X^{T} \bm X$\footnote{TODO:
-  $1/N$ is missing. Add this. It will be absorbed into the eigenvalues.}  Then,
-because $\bm X = \bm M \bm D^{1/2} \bm U^{T}$ by the SVD, we have the
-eigendecomposition $\widehat{\bm \Sigma} = \bm U \bm D \bm U^{T}$, where $\bm M
-\in \mathbb{R}_{N \times N}$ is orthogonal, $\bm D^{1/2} \in \mathbb{R}_{N
-  \times N}^{\ge}$ is a diagonal matrix consisting of the singular values of
-$\bm X$, and $\bm U \in \mathbb{R}_{p \times N}$ is orthogonal. Now, $\bm X \bm
-X^{T} = \bm M \bm D \bm M^{T}$ is the eigenvalue decomposition of $\bm X \bm
-X^{T}$, where $\bm D$ is the matrix of eigenvalues of $\bm X \bm X^{T}$ and the
-columns of $\bm M$ are the corresponding eigenvectors. Hence, we obtain $\bm M$
-and $\bm D$ from an eigendecomposition of $\bm X \bm X^{T} \in \mathbb{R}_{N
-  \times N}$. Next, we compute $\bm U = \bm X^{T} \bm M \bm D^{+/2}$, where $\bm
-D^{+/2} = \bm D_q^{-1/2} \oplus \bm 0_{N-q}$. We then determine $q$, the number
-of numerically nonzero eigenvalues present in $\bm D$, by calculating the number
-of eigenvalues that exceeds some tolerance value, say, $\epsilon = 1 \times
-10^{-6}$. After calculating $q$, we extract $\bm U_1$ by retaining only the
-first $q$ columns of $\bm U$.
+decomposition (SVD) (sometimes called the reduced or fast SVD). Rather than
+computing the eigenvalue decomposition of $\widehat{\bm \Sigma}$ to obtain $\bm
+U_1$, we instead obtain $\bm U_1$ by computing the eigendecomposition of a much
+smaller $N \times N$ matrix when $p \gg N$
+\cite[Chapter~18.3.5]{Hastie:2008dt}. Recall that the pooled sample covariance
+matrix is $\widehat{\bm \Sigma} = \bm X^{T} \bm X$\footnote{TODO: $1/N$ is
+  missing. Add this. It will be absorbed into the eigenvalues.}  Then, because
+$\bm X = \bm M \bm D^{1/2} \bm U^{T}$ by the SVD, we have the eigendecomposition
+$\widehat{\bm \Sigma} = \bm U \bm D \bm U^{T}$, where $\bm M \in \mathbb{R}_{N
+  \times N}$ is orthogonal, $\bm D^{1/2} \in \mathbb{R}_{N \times N}^{\ge}$ is a
+diagonal matrix consisting of the singular values of $\bm X$, and $\bm U \in
+\mathbb{R}_{p \times N}$ is orthogonal. Now, $\bm X \bm X^{T} = \bm M \bm D \bm
+M^{T}$ is the eigenvalue decomposition of $\bm X \bm X^{T}$, where $\bm D$ is
+the matrix of eigenvalues of $\bm X \bm X^{T}$ and the columns of $\bm M$ are
+the corresponding eigenvectors. Hence, we obtain $\bm M$ and $\bm D$ from an
+eigendecomposition of $\bm X \bm X^{T} \in \mathbb{R}_{N \times N}$. Next, we
+compute $\bm U = \bm X^{T} \bm M \bm D^{+/2}$, where $\bm D^{+/2} = \bm
+D_q^{-1/2} \oplus \bm 0_{N-q}$. We then determine $q$, the number of numerically
+nonzero eigenvalues present in $\bm D$, by calculating the number of eigenvalues
+that exceeds some tolerance value, say, $\epsilon = 1 \times 10^{-6}$. After
+calculating $q$, we extract $\bm U_1$ by retaining only the first $q$ columns of
+$\bm U$.
 
@@
 $q$-dimensional subspace. Similarly, we reduce the dimension of the test data
-set by calculating $\bm X_c \bm U_1$ once per fold. Conveniently, we see that
-the most costly computation involved in $\bm Q_k$ and $\bm W_k^{-1}$ is $\bm X_k
-\bm U_1$, which can be extracted from $\bm X_c \bm U_1$. Thus, after the initial
-calculation of $\bm X_c \bm U_1$ per cross-validation fold, $\bm Q_k$ requires
-$O(TODO)$ operations. Because $\bm Q_k \in \mathbb{R}_{n_k \times n_k}$, both
-its determinant and inverse requires $O(n_k^3)$ operations. This is especially
-advantageous when $K$ is large, such that $n_k$ is much smaller than $N$. Thus,
-$\bm W_k^{-1}$ requires $O(TODO)$ operations Also, the inverse and determinant
-of the diagonal matrix $\bm \Gamma_k \in \mathbb{R}_{q \times q}$ requires
-$O(q)$ operations. Hence, the $| \bm W_k |$ requires $O(q n_k^3)$ operations,
-which is efficient when $p >> N$.
+set by calculating $\bm X_{test} \bm U_1$ once per fold. Conveniently, we see
+that the most costly computation involved in $\bm Q_k$ and $\bm W_k^{-1}$ is
+$\bm X_k \bm U_1$, which can be extracted from $\bm X_c \bm U_1$. Thus, after
+the initial calculation of $\bm X_c \bm U_1$ per cross-validation fold, $\bm
+Q_k$ requires $O(TODO)$ operations. Because $\bm Q_k \in \mathbb{R}_{n_k \times
+  n_k}$, both its determinant and inverse require $O(n_k^3)$
+operations. Consequently, $\bm W_k^{-1}$ requires $O(TODO)$ operations. Also,
+the inverse and determinant of the diagonal matrix $\bm \Gamma_k \in
+\mathbb{R}_{q \times q}$ requires $O(q)$ operations. Therefore, the $| \bm W_k
+|$ requires $O(q n_k^3)$ operations.
 
@@
 of $\lambda$ and $\gamma$. The subsequent operations in calculating $\bm
-W_k^{-1}$ and $|\bm W_K|$ can be simply updated for different pairs of $\lambda$
-and $\gamma$.\footnote{TODO: Need to clean up this paragraph so that it flows
-  from above. Also, elaborate.}
+W_k^{-1}$ and $|\bm W_k|$ can be simply updated for different pairs of $\lambda$
+and $\gamma$.
 

### 2013-12-30 reports/paper-hdrda.Rnw
Discussed the purpose of alpha_k. #12

@@
 
-\subsection{Relationship with Friedman's RDA Classifier}
+\subsection{Relationship with Other Classifiers}
+
+The addition of the parameter $\alpha_k$ in \eqref{eq:hdrda-cov} is one of
+convenience and allows the flexibility of various covariance-matrix estimators
+proposed in the literature. In practice, we generally are not interested in
+estimating $\alpha_k$ as it would incorporate $K$ additional tuning parameters
+to estimate via cross-validation, which is counterproductive to our goal of
+improving the \emph{RDA} classifier's efficiency. By setting $\alpha_k$ to a
+predefined constant, the \emph{HDRDA} covariance-matrix estimator includes or
+resembles a large family of estimators, some of which we mention below.
+
+\subsubsection{\cite{Friedman:1989tm}}
 
@@
 established by substituting \eqref{eq:sig-rda} into
-\eqref{eq:qda}.\footnote{TODO: Discuss $\alpha_k \equiv 1-\gamma$, etc.}
+\eqref{eq:qda}.
 

### 2013-12-30 reports/paper-hdrda.Rnw
Abstract polish recommended by Caleb.

@@
 \citeauthor{Friedman:1989tm} proposed the popular regularized discriminant
-analysis (RDA) classifier that utilizes a biased covariance-matrix estimator
-that partially pools the sample covariance matrices from linear and quadratic
+analysis (RDA) classifier that utilizes a biased covariance-matrix estimator to
+partially pool the sample covariance matrices from linear and quadratic
 disriminant analysis before shrinking the resulting estimator towards a scaled

### 2013-12-31 reports/paper-hdrda.Rnw
Fixed typo in Lemma 2's proof.

@@
 From Lemma \ref{lemma:null-spaces}, the columns of $\bm U_2$ span the null space
-of $\widehat{\bm \Sigma}$, which implies that $\widehat{\bm \Sigma}_k \bm U_2 =
-\bm 0_{p \times (p - q)}$. Hence, $\bm U^{T} \widehat{\bm \Sigma}_k \bm U = \bm
-U_1^{T} \widehat{\bm \Sigma}_k \bm U_1 \oplus \bm 0_{p-q}$, $k = 1, \ldots,
+of $\widehat{\bm \Sigma}_k$, which implies that $\widehat{\bm \Sigma}_k \bm U_2
+= \bm 0_{p \times (p - q)}$. Hence, $\bm U^{T} \widehat{\bm \Sigma}_k \bm U =
+\bm U_1^{T} \widehat{\bm \Sigma}_k \bm U_1 \oplus \bm 0_{p-q}$, $k = 1, \ldots,
 K$. Thus, $\bm U^{T} \tilde{\bm \Sigma}_k \bm U = \alpha_k \{(1 - \lambda) \bm

### 2014-01-01 reports/paper-hdrda.Rnw
More polish applied to the summary of the model selection.

@@
 
-      Compute $\#\{\text{Error}_v(\lambda, \gamma)\} \leftarrow$ misclassified test observations
+      Compute the number of misclassified test observations $\#\{\text{Error}_v(\lambda, \gamma)\}$ 
     }   
@@
 smaller $N \times N$ matrix when $p \gg N$
-\cite[Chapter~18.3.5]{Hastie:2008dt}. Recall that the pooled sample covariance
-matrix is $\widehat{\bm \Sigma} = \bm X^{T} \bm X$\footnote{TODO: $1/N$ is
-  missing. Add this. It will be absorbed into the eigenvalues.}  Then, because
-$\bm X = \bm M \bm D^{1/2} \bm U^{T}$ by the SVD, we have the eigendecomposition
-$\widehat{\bm \Sigma} = \bm U \bm D \bm U^{T}$, where $\bm M \in \mathbb{R}_{N
-  \times N}$ is orthogonal, $\bm D^{1/2} \in \mathbb{R}_{N \times N}^{\ge}$ is a
-diagonal matrix consisting of the singular values of $\bm X$, and $\bm U \in
-\mathbb{R}_{p \times N}$ is orthogonal. Now, $\bm X \bm X^{T} = \bm M \bm D \bm
-M^{T}$ is the eigenvalue decomposition of $\bm X \bm X^{T}$, where $\bm D$ is
-the matrix of eigenvalues of $\bm X \bm X^{T}$ and the columns of $\bm M$ are
-the corresponding eigenvectors. Hence, we obtain $\bm M$ and $\bm D$ from an
-eigendecomposition of $\bm X \bm X^{T} \in \mathbb{R}_{N \times N}$. Next, we
-compute $\bm U = \bm X^{T} \bm M \bm D^{+/2}$, where $\bm D^{+/2} = \bm
-D_q^{-1/2} \oplus \bm 0_{N-q}$. We then determine $q$, the number of numerically
-nonzero eigenvalues present in $\bm D$, by calculating the number of eigenvalues
-that exceeds some tolerance value, say, $\epsilon = 1 \times 10^{-6}$. After
-calculating $q$, we extract $\bm U_1$ by retaining only the first $q$ columns of
-$\bm U$.
+\cite[Chapter~18.3.5]{Hastie:2008dt}. Applying the SVD, we decompose $\bm X_c =
+\bm M \bm \Delta \bm U^{T}$, where $\bm M \in \mathbb{R}_{N \times p}$ is
+orthogonal, $\bm \Delta \in \mathbb{R}_{p \times p}^{\ge}$ is a diagonal matrix
+consisting of the singular values of $\bm X_c$, and $\bm U \in \mathbb{R}_{p
+  \times p}$ is orthogonal. Recalling that $\widehat{\bm \Sigma} = N^{-1} \bm
+X_c^{T} \bm X_c$, we have the eigendecomposition $\widehat{\bm \Sigma} = \bm U
+\bm D \bm U^{T}$, where $\bm U$ is the matrix of eigenvectors of $\widehat{\bm
+  \Sigma}$ and $\bm D = N^{-1} \bm \Delta$ is the diagonal matrix of eigenvalues
+of $\widehat{\bm \Sigma}$. Now, we can obtain $\bm M$ and $\bm D$ efficiently
+from the eigenvalue decomposition of $\bm X_c \bm X_c^{T} = \bm M \bm D \bm
+M^{T}$. Hence, we eigendecomposition of $\bm X_c \bm X_c^{T} \in \mathbb{R}_{N
+  \times N}$. Next, we compute $\bm U = \bm X_c^{T} \bm M \bm D^{+/2}$, where
+$\bm D^{+/2} = \bm D_q^{-1/2} \oplus \bm 0_{N-q}$. We then determine $q$, the
+number of numerically nonzero eigenvalues present in $\bm D$, by calculating the
+number of eigenvalues that exceeds some tolerance value, say, $\epsilon = 1
+\times 10^{-6}$. After calculating $q$, we extract $\bm U_1$ by retaining only
+the first $q$ columns of $\bm U$.
 
@@
 the initial calculation of $\bm X_c \bm U_1$ per cross-validation fold, $\bm
-Q_k$ requires $O(TODO)$ operations. Because $\bm Q_k \in \mathbb{R}_{n_k \times
+Q_k$ requires $O(n_k q^2)$ operations. Because $\bm Q_k \in \mathbb{R}_{n_k \times
   n_k}$, both its determinant and inverse require $O(n_k^3)$
-operations. Consequently, $\bm W_k^{-1}$ requires $O(TODO)$ operations. Also,
+operations. Consequently, $\bm W_k^{-1}$ requires $O(n_k q^2)$ operations. Also,
 the inverse and determinant of the diagonal matrix $\bm \Gamma_k \in
@@
 W_k^{-1}$ and $|\bm W_k|$ can be simply updated for different pairs of $\lambda$
-and $\gamma$.
+and $\gamma$ without repeating the costly computations.
 

### 2014-01-01 reports/paper-hdrda.Rnw
Slight improvements to discussion of Gamma_k and Gamma_k_inv

@@
 
+Notice that if $\alpha_k$ is constant across the $K$ classes, then $\bm
+\Gamma_k$ in \eqref{eq:Gamma_k} is independent of $k$. Consequently, $|\bm
+\Gamma_k|$ is constant across the $K$ classes and need not be calculated in
+\eqref{eq:hdrda-decomposed}.
+
 \subsection{Model Selection}
@@
 the initial calculation of $\bm X_c \bm U_1$ per cross-validation fold, $\bm
-Q_k$ requires $O(n_k q^2)$ operations. Because $\bm Q_k \in \mathbb{R}_{n_k \times
-  n_k}$, both its determinant and inverse require $O(n_k^3)$
+Q_k$ requires $O(n_k q^2)$ operations. Because $\bm Q_k \in \mathbb{R}_{n_k
+  \times n_k}$, both its determinant and inverse require $O(n_k^3)$
 operations. Consequently, $\bm W_k^{-1}$ requires $O(n_k q^2)$ operations. Also,
-the inverse and determinant of the diagonal matrix $\bm \Gamma_k \in
-\mathbb{R}_{q \times q}$ requires $O(q)$ operations. Therefore, the $| \bm W_k
-|$ requires $O(q n_k^3)$ operations.
+the inverse of the diagonal matrix $\bm \Gamma_k^{-1} \in \mathbb{R}_{q \times
+  q}$ requires $O(q)$ operations. Finally, $| \bm W_k |$ requires $O(n_k^3)$
+operations.
 

### 2014-01-02 reports/paper-hdrda.Rnw
Added a couple of sentences describing how we handle the special case where (lambda, gamma) = (0, 0).

@@
 
-Notice that if $\alpha_k$ is constant across the $K$ classes, then $\bm
-\Gamma_k$ in \eqref{eq:Gamma_k} is independent of $k$. Consequently, $|\bm
-\Gamma_k|$ is constant across the $K$ classes and need not be calculated in
-\eqref{eq:hdrda-decomposed}.
+The application of the Sherman-Woodbury formula requires that $\bm \Gamma_k$ be
+nonsingular. This assumption is not met when $(\lambda, \gamma) = (0, 0)$ as
+$\bm \Gamma_k = \bm 0_q$, corresponding to the \emph{QDA} classifier. In this
+special case we apply a small amount of eigenvalue shrinkage and set $\gamma =
+0.01$ to circumvent the singularity issue. Also, notice that if $\alpha_k$ is
+constant across the $K$ classes, then $\bm \Gamma_k$ in \eqref{eq:Gamma_k} is
+independent of $k$. Consequently, $|\bm \Gamma_k|$ is constant across the $K$
+classes and need not be calculated in \eqref{eq:hdrda-decomposed}.
 

### 2014-01-04 lib/helpers.R
Revamped microarray sims. Also, added a few classifiers, such as SVM and kNN.

@@ -201,8 +201,15 @@ dudoit <- function(train_x, train_y) {
   rev(order(F_stat))
 }
 
-# Classifier from Clemmensen, Hastie, Witten and Ersbøll (2012) - Technometrics
-Clemmensen <- function(train_x, train_y, test_x, normalize_data = FALSE, cv_variables = FALSE, num_folds = 10, ...) {
+# Classifier from Clemmensen, Hastie, Witten and Ersbøll (2011) - Technometrics
+#
+# NOTE: We had to increase the L2-regularization term, lambda, from the
+# default of 1e-6 to 1e-3 to circumvent the following error:
+# Error in lda.default(x, grouping, ...)
+# variables 1 2 appear to be constant within groups
+Clemmensen <- function(train_x, train_y, test_x, normalize_data = TRUE,
+                       cv_variables = FALSE, num_folds = 10, lambda = 1e-3,
+                       verbose = FALSE, ...) {
   if (normalize_data) {
     normalize_out <- normalize(train_x)
     train_x <- normalize_out$Xc
@@ -231,7 +238,7 @@ Clemmensen <- function(train_x, train_y, test_x, normalize_data = FALSE, cv_vari
       # For each reduced dimension considered, we calculate the number of test errors
       # resulting for the current cross-validation fold.
       cv_num_vars <- sapply(num_vars, function(q) {
-        sda_predict <- try_default(predict(sda(x = trn_x, y = trn_y, stop = -q), tst_x)$class, NA, quiet = TRUE)
+        sda_predict <- try_default(predict(sda(x = trn_x, y = trn_y, stop = -q, lambda = lambda), tst_x)$class, NA, quiet = TRUE)
         sum(sda_predict != tst_y)
       })
     })
@@ -244,8 +251,13 @@ Clemmensen <- function(train_x, train_y, test_x, normalize_data = FALSE, cv_vari
   } else {
     num_vars <- ncol(train_x)
   }
-     
-  predict(sda(x = train_x, y = train_y, stop = -num_vars), test_x)$class
+
+  sda_out <- sda(x = train_x,
+                 y = train_y,
+                 stop = -num_vars,
+                 lambda = lambda,
+                 trace = verbose)
+  predict(sda_out, test_x)$class
 }
 
 # Classifier from Cao, Boitard, and Besse (2011) - BMC Bioinformatics

### 2014-01-04 sim-microarray.r

@@ -2,80 +2,154 @@ library('ProjectTemplate')
 load.project()
 
 set.seed(42)
-num_cores <- 12
+num_cores <- 3
 
 train_pct <- 2/3
-num_iterations <- 1000
-data_sets <- c("burczynski", "nakayama", "shipp", "singh")
-d_vals <- 100 * seq_len(10)
+num_iterations <- 5
+#data_sets <- as.character(describe_data()$author)
+#data_sets <- c("burczynski", "nakayama", "shipp", "singh")
+data_sets <- c("sorlie", "christensen")
 
 sim_config <- sim_factors(data_sets = data_sets, num_reps = num_iterations,
                           stringsAsFactors = FALSE)
 
 results <- mclapply(seq_len(nrow(sim_config)), function(i) {
-  data_set <- sim_config$data_set[i]
+  message("Dataset: ", data_set, " -- Sim Config: ", i, " of ", nrow(sim_config))
+  set.seed(i)
   
+  # Load data and randomly partition into training/test data sets.
+  data_set <- sim_config$data_set[i]
   data <- load_microarray(data_set)
   data$x <- as.matrix(data$x)
-
   rand_split <- rand_partition(y = data$y, num_partitions = 1,
                                train_pct = train_pct)[[1]]
-
-  message("Dataset: ", data_set, " -- Sim Config: ", i, " of ", nrow(sim_config))
   train_x <- as.matrix(data$x[rand_split$training, ])
   train_y <- data$y[rand_split$training]
   test_x <- as.matrix(data$x[rand_split$test, ])
   test_y <- data$y[rand_split$test]
 
-  # Apply Dudoit variable selection to rank the genes in descending order
-  dudoit_out <- dudoit(train_x, train_y)
+  # Removes the loaded data from memory and then garbage collects.
+  rm(data)
+  gc(reset = TRUE)
 
-  # For each value of 'd', we compute the error-rate estimates for each of the
-  # classifiers.
-  errors_d <- lapply(d_vals, function(d) {
-    trn_x <- train_x[, head(dudoit_out, d)]
-    tst_x <- test_x[, head(dudoit_out, d)]
+  # Sets the prior probabilities as equal
+  num_classes <- nlevels(train_y)
+  prior_probs <- rep(1, num_classes) / num_classes  
 
-    # Sets the prior probabilities equal
-    num_classes <- nlevels(train_y)
[…]

### 2014-04-02 reports/paper-hdrda.Rnw
Much needed edits applied.

@@
 estimated via a computationally burdensome cross-validation procedure from a
-grid of candidate values. We reformulate the RDA classifier in the
+grid of candidate values. We reformulate the RDA classifier in the small-sample,
 high-dimensional setting and then show that the classification decision rule is
@@
 interpretability that has been previously lacking with the \emph{RDA} classifier
-\citep{Bensmail:1996fw}. Furthermore, the \emph{HDRDA} classifier includes a
-large family of shrinkage methods, such as those from \cite{Srivastava:2007ww},
-\cite{Rao:1971ul}, and other methods studied by \cite{Ramey:2013ji} and
-\cite{Xu:2009fl}. Next, applying a reasoning similar to \cite{Ye:2006tq}, we
-show that the matrix operations corresponding to the null space of the pooled
-sample covariance matrix are redundant and can be discarded from the
+\citep{Bensmail:1996fw}. Next, applying a reasoning similar to \cite{Ye:2006tq},
+we establish that the matrix operations corresponding to the null space of the
+pooled sample covariance matrix are redundant and can be discarded from the
 \emph{HDRDA} decision rule without loss of classificatory information. Moreover,
@@
 rapidly computed. Thus, the dimension reduction inherent in the \emph{HDRDA}
-classifier yields a substantial reduction in computation during model selection.
+classifier yields a substantial reduction in computation during model
+selection. Also, as a result, we show that a large number of shrinkage methods
+that are special cases of the \emph{HDRDA} classifier have no effect on the zero
+eigenvalues of the covariance-matrix estimators. Such methods include work from
+\cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and other methods studied by
+\cite{Ramey:2013ji} and \cite{Xu:2009fl}.
 
@@
 
-In discriminant analysis we wish to assign correctly an unlabeled vector $\bm x
-\in \mathbb{R}_{p \times 1}$ to one of $K$ unique, known classes by constructing
-a classifier from $N$ training observations. Let $\bm x_i = (x_{i1}, \ldots,
+In discriminant analysis we wish to assign an unlabeled vector $\bm x \in
+\mathbb{R}_{p \times 1}$ to one of $K$ unique, known classes by constructing a
+classifier from $N$ training observations. Let $\bm x_i = (x_{i1}, \ldots,
 x_{ip}) \in \mathbb{R}_{p \times 1}$ be the $i$th observation $(i = 1, \ldots,
@@
 
-The smallest eigenvalues and the directions associated with their eigenvectors
-can highly influence \eqref{eq:qda}. In fact, the eigenvalues of $\widehat{ \bm
-  \Sigma}_k$ are well-known to be biased if $p \ge n_k$ such that the smallest
-eigenvalues are underestimated \citep{Seber:2004uh}. Moreover, if $p > n_k$,
-then rank$(\widehat{\bm \Sigma}_k) \le n_k$, which implies that at least $p -
-n_k$ eigenvalues of $\widehat{\bm \Sigma}_k$ are zero. Furthermore, although
-more feature information is available to discriminate among the $K$ classes, if
-$p > n_k$, \eqref{eq:qda} is incalculable because $\widehat{\bm \Sigma}_k^{-1}$
-does not exist.
+The smallest eigenvalues of $\widehat{\bm \Sigma}_k$ and the directions
+associated with their eigenvectors can highly influence \eqref{eq:qda}. In fact,
+the eigenvalues of $\widehat{ \bm \Sigma}_k$ are well-known to be biased if $p
+\ge n_k$ such that the smallest eigenvalues are underestimated
+\citep{Seber:2004uh}. Moreover, if $p > n_k$, then rank$(\widehat{\bm \Sigma}_k)
+\le n_k$, which implies that at least $p - n_k$ eigenvalues of $\widehat{\bm
+  \Sigma}_k$ are zero. Furthermore, although more feature information is
+available to discriminate among the $K$ classes, if $p > n_k$, \eqref{eq:qda} is
+incalculable because $\widehat{\bm \Sigma}_k^{-1}$ does not exist.
 
@@
 1, \ldots, N)$, each centered by its class sample mean, we attain a clear
-interpretation of $\widehat{\bm \Sigma}_k(\lambda)$. That is, from
-\eqref{eq:pooled-cov} we have
+interpretation of $\widehat{\bm \Sigma}_k(\lambda)$. That is,
 \begin{align}
@@
 $N$ rather than $n_k$ observations.\footnote{TODO: Add more to the
-  interpretation here and reference Figure \ref{fig:hdrda-weights}.}
-
-\begin{figure}
-<<rda_weights>>=
-n <- c(50, 25, 15, 10)
-N <- sum(n)
-lambda <- seq(0, 1, length = 1000)
-
-rda_weights <- sapply(seq_along(n), function(k) {
-      lambda / N + (1 - lambda) / n[k]
-      })
-rda_weights <- data.frame(lambda = lambda, rda_weights)
-rda_weights <- melt(rda_weights, id = 'lambda')
-colnames(rda_weights) <- c("lambda", "Class", "Weight")
-levels(rda_weights$Class) <- seq_along(n)
-
-p <- ggplot(rda_weights, aes(lambda, Weight, linetype = Class, color = Class))
-p <- p + geom_line(size = 1)
-p <- p + theme_bw() + xlab(expression(lambda))
-p <- p + theme(axis.text = element_text(size = 10))
-p <- p + theme(strip.text.x = element_text(size = 16))
-p <- p + theme(legend.title = element_text(size = 16, face = "bold"))
-p <- p + theme(legend.text = element_text(size = 15))
-p <- p + theme(legend.key.size = unit(1, "cm"))
-p
-@ 
-\caption{Weights $c_{ik}(\lambda)$ of \emph{RDA} classifier as a function of the
-  pooling parameter $\lambda$ when $n_1 = 50$, $n_2 = 25$, $n_3 = 15$, and $n_4$
-  = 10. TODO: Scale down figure.}
-\label{fig:hdrda-weights}
-\end{figure}
-
-Below, we will show that the pooling
-operation is advantageous in increasing the rank of each $\widehat{\bm
-  \Sigma}_k(\lambda)$ from rank$(\widehat{\bm \Sigma}_k)$ to rank$(\widehat{\bm
-  \Sigma})$ for $0 < \lambda \le 1$. Notice that if $\lambda = 0$, then the
-observations from the remaining $K - 1$ classes do not contribute to the
-estimation of $\bm \Sigma_k$, corresponding to $\widehat{\bm
-  \Sigma}_k$. Furthermore, if $\lambda = 1$, the weights in
-\eqref{eq:sig-lambda-alternative2} reduce to $1/N$, corresponding to
-$\widehat{\bm \Sigma}$. For brevity, when $\lambda = 1$, we define $\bm X =
-      [\sqrt{c_{1k}(1)} \bm x_1^{T}, \ldots, \sqrt{c_{Nk}(1)} \bm x_N^{T}]^{T}$
-      such that $\widehat{\bm \Sigma} = \bm X^{T} \bm X$. Similarly, for
-      $\lambda = 0$, we define $\bm X_k = [\sqrt{c_{1k}(0)} \bm x_1^{T}, \ldots,
-        \sqrt{c_{Nk}(0)} \bm x_N^{T}]^{T}$ such that $\widehat{\bm \Sigma}_k =
-      \bm X_k^{T} \bm X_k$. In Figure \ref{fig:hdrda-contours} we plot the
-      contours of five multivariate normal populations for $\lambda = 0$ with
-      unequal covariance matrices. As $\lambda$ approaches 1, the contours
-      become more similar, resulting in identical contours for $\lambda = 1$.
+  interpretation here.}
+
+Below, we will show that the pooling operation is advantageous in increasing the
+rank of each $\widehat{\bm \Sigma}_k(\lambda)$ from rank$(\widehat{\bm
+  \Sigma}_k)$ to rank$(\widehat{\bm \Sigma})$ for $0 < \lambda \le 1$. Notice
+that if $\lambda = 0$, then the observations from the remaining $K - 1$ classes
+do not contribute to the estimation of $\bm \Sigma_k$, corresponding to
+$\widehat{\bm \Sigma}_k$. Furthermore, if $\lambda = 1$, the weights
+$c_{ik}(\lambda)$ in \eqref{eq:sig-lambda-alternative2} reduce to $1/N$,
+corresponding to $\widehat{\bm \Sigma}$. For brevity, when $\lambda = 1$, we
+define $\bm X = [\sqrt{c_{1k}(1)} \bm x_1^{T}, \ldots, \sqrt{c_{Nk}(1)} \bm
+  x_N^{T}]^{T}$ such that $\widehat{\bm \Sigma} = \bm X^{T} \bm X$. Similarly,
+for $\lambda = 0$, we define $\bm X_k = [\sqrt{c_{1k}(0)} \bm x_1^{T}, \ldots,
+  \sqrt{c_{Nk}(0)} \bm x_N^{T}]^{T}$ such that $\widehat{\bm \Sigma}_k = \bm
+X_k^{T} \bm X_k$. In Figure \ref{fig:hdrda-contours} we plot the contours of
+five multivariate normal populations for $\lambda = 0$ with unequal covariance
+matrices. As $\lambda$ approaches 1, the contours become more similar, resulting
+in identical contours for $\lambda = 1$.
+
 \begin{figure}
@@
 \end{figure}
+
 As we have discussed above, several eigenvalue adjustment methods have been
@@
 established by substituting \eqref{eq:sig-rda} into
-\eqref{eq:qda}.
-
-\subsection{Properties of the HDRDA Classifier}
+\eqref{eq:qda}.\footnote{TODO: Mention the values of $\alpha_k$ and $\gamma$
+  needed to resemble HDRDA.}
+
+\subsubsection{Ridge-Like Estimators}
+
+TODO: List a few estimators here along with $\alpha_k = 1$ such that HDRDA
+includes them all. For example, set $\gamma = \text{trace}(\widehat{\bm
+  \Sigma}_k) / \min(n_k, p)$ for Srivastava and Kubokawa.
+
+\section{Properties of the HDRDA Classifier}
 \label{sec:hdrda-properties}
@@
 Thus, by incorporating each $\bm x_i$ into the estimation of $\bm \Sigma_k$, we
-increase the rank of $\widehat{\bm \Sigma}_k(\lambda)$ to $q$ if $\lambda \ne
-0$. The results is that the \emph{HDRDA} classifier has the advantage over
-\eqref{eq:qda} in that the rank$(\widehat{\bm \Sigma}_k) \le n_k$. Next, we
-provide an essential result that enables us to prove that \eqref{eq:hdrda} is
-invariant to adjustments to the eigenvalues of $\tilde{ \bm \Sigma}_k$
-corresponding to the null space of $\widehat{\bm \Sigma}$ for any unlabeled
-observation $\bm x \in \mathbb{R}_{p \times 1}$.
+increase the rank of $\widehat{\bm \Sigma}_k(\lambda)$ to $q \approx N$ if
+$\lambda \ne 0$. Next, we provide an essential result that enables us to prove
+that \eqref{eq:hdrda} is invariant to adjustments to the eigenvalues of $\tilde{
+  \bm \Sigma}_k$ corresponding to the null space of $\widehat{\bm \Sigma}$.
 
@@
 \begin{proposition}\label{proposition:hdrda-W_k}
-Let $\bm W_k$ be defined as above, $k = 1, \ldots, K$. Then, $|\bm W_k| = |
-\bm \Gamma_k| |\bm Q_k|$ and
+Let $\bm W_k$ be defined as above. Then, $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$
+and
 	\begin{align}
@@
 
-        Compute $(\bm x - \xbar_k)^{T} \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm x - \xbar_k) + \log | \bm W_k |$ \tcp*[r]{TODO: Use matrix form for test observations}
+        Compute $(\bm x - \xbar_k)^{T} \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm x -
+        \xbar_k) + \log | \bm W_k |$ for each row $\bm x$ of $\bm X_{test}$
       }
@@
 of $\widehat{\bm \Sigma}$. Now, we can obtain $\bm M$ and $\bm D$ efficiently
-from the eigenvalue decomposition of $\bm X_c \bm X_c^{T} = \bm M \bm D \bm
-M^{T}$. Hence, we eigendecomposition of $\bm X_c \bm X_c^{T} \in \mathbb{R}_{N
-  \times N}$. Next, we compute $\bm U = \bm X_c^{T} \bm M \bm D^{+/2}$, where
-$\bm D^{+/2} = \bm D_q^{-1/2} \oplus \bm 0_{N-q}$. We then determine $q$, the
-number of numerically nonzero eigenvalues present in $\bm D$, by calculating the
-number of eigenvalues that exceeds some tolerance value, say, $\epsilon = 1
-\times 10^{-6}$. After calculating $q$, we extract $\bm U_1$ by retaining only
-the first $q$ columns of $\bm U$.
+from the eigenvalue decomposition of the $N \times N$ matrix $\bm X_c \bm
+X_c^{T} = \bm M \bm D \bm M^{T}$. Next, we compute $\bm U = \bm X_c^{T} \bm M
+\bm D^{+/2}$, where $\bm D^{+/2} = \bm D_q^{-1/2} \oplus \bm 0_{N-q}$. We then
+determine $q$, the number of numerically nonzero eigenvalues present in $\bm D$,
+by calculating the number of eigenvalues that exceeds some tolerance value, say,
+$\epsilon = 1 \times 10^{-6}$. After calculating $q$, we extract $\bm U_1$ by
+retaining only the first $q$ columns of $\bm U$.
 
@@
 W_k^{-1}$ and $|\bm W_k|$ can be simply updated for different pairs of $\lambda$
-and $\gamma$ without repeating the costly computations.
+and $\gamma$ without repeating the costly computations. When calculating the
+discriminant scores of each test observation in $\bm X_{test}$, rather than
+calculating each score individually, we can more efficiently calculate the
+quadratic forms via a single matrix operation. That is, rather than calculating
+$(\bm x - \xbar_k)^{T} \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm x - \xbar_k)$
+individually for each row $\bm x$ of $\bm X_{test}$, we can instead calculate
+$(\bm X_{test} - \xbar_k \bm 1_k')' \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm
+X_{test} - \xbar_k \bm 1_k')$. The diagonal elements of the resulting matrix
+contain the individual quadratic form of each test observation, $\bm x_t$.
 
@@
 latter utilizes an improved variance estimator. Below, we refer to each
-classifier by the first author's surname.
+classifier by the first author's surname.\footnote{TODO: Consider adding a
+  heatmap of error rates of a single $(\lambda, \gamma)$ grid.}
 

### 2015-03-30 reports/paper-hdrda.Rnw
Applied some more edits to the paper.

@@
 
-results <- lapply(results, function(res) {
-  res <- lapply(res, function(res_d) {
-    errors <- do.call(cbind, res_d$errors)
-    res_d[["errors"]] <- NULL
-    cbind(do.call(cbind, res_d), errors)
-  })
-  do.call(rbind, res)
-})
-results <- data.frame(do.call(rbind, results), stringsAsFactors = FALSE)
-
-results$lambda <- as.numeric(results$lambda)
-results$gamma <- as.numeric(results$gamma)
-results$Dudoit <- as.numeric(results$Dudoit)
-results$HDRDA <- as.numeric(results$HDRDA)
-results$Guo <- as.numeric(results$Guo)
-results$Pang <- as.numeric(results$Pang)
-results$Tong <- as.numeric(results$Tong)
-results$Witten <- as.numeric(results$Witten)
-
-results$d <- factor(results$d, labels = sort(unique(as.numeric(results$d))))
-
-# Summary of the classification error rates for each data set and number of
-# variables selected, d.
-error_rates <- subset(results, select = -c(lambda, gamma))
-m_error <- melt(error_rates, variable.name = "Classifier", value.name = "Error")
-error_rates <- ddply(m_error, .(data_set, d, Classifier), plyr:::summarize,
-                     Avg_Error = mean(Error), Num_Reps = length(Error))
-error_rates <- ddply(error_rates, .(data_set, d, Classifier), transform,
-                     SE_Error = sqrt(Avg_Error * (1 - Avg_Error) / Num_Reps))
-error_rates <- subset(error_rates, select = -Num_Reps)
-error_rates$Classifier <- factor(as.character(error_rates$Classifier))
-
-max_se <- max(error_rates[["SE_Error"]])
+max_se <- 1234567890
 @
@@
 the \emph{LDA} and \emph{QDA} classifiers for application to high-dimensional
-data: 1) covariance-matrix regularization, 2) dimension reduction, and 3) model
-parsimony. Covariance-matrix regularization methods are often based on the
-well-known ridge-regression approach of
-\cite{Hoerl:1970cd}. \cite{Ching:2012fu}, \cite{Zhang:2010va}, \cite{Ye:2009gd},
-\cite{Ji:2008wp}, \cite*{Guo:2007te}, \cite*{Srivastava:2007ud},
-\cite{Ye:2006vx}, and \cite{Ye:2005uu} have proposed regularization methods to
-improve the estimation of the class covariance matrices by adjusting the
-eigenvalues of the sample covariance matrices to ensure positive
-definiteness. Alternatively, dimension reduction typically transforms the
-original data in the high-dimensional space to a subspace having a much smaller
-dimension such that redundant and noisy information is removed
+data: 1) covariance-matrix regularization, 2) dimension reduction, and 3)
+parameter restrictions. Covariance-matrix regularization methods are often based
+on the well-known ridge-regression approach of \cite{Hoerl:1970cd}.
+\cite{Ching:2012fu}, \cite{Zhang:2010va}, \cite{Ye:2009gd}, \cite{Ji:2008wp},
+\cite*{Guo:2007te}, \cite*{Srivastava:2007ud}, \cite{Ye:2006vx}, and
+\cite{Ye:2005uu} have proposed regularization methods to improve the estimation
+of the class covariance matrices by adjusting the eigenvalues of the sample
+covariance matrices to ensure positive definiteness. Alternatively, dimension
+reduction typically transforms the original data in the high-dimensional space
+to a subspace having a much smaller dimension such that redundant and noisy
+information is removed
 \citep{Fan:2012iq,Ye:2009gd,Bouveyron:2007gx,Ye:2006vx,Ye:2005hd,Antoniadis:2003bk}. Finally,
@@
 \cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and other methods studied by
-\cite{Ramey:2013ji} and \cite{Xu:2009fl}.
+\cite{Ramey:2013ji} and \citet*{Xu:2009fl}.
 
@@
 Several regularization methods, such as the methods considered by
-\citet*{Xu:2009fl}, \citet*{Guo:2007te}, and \cite{Mkhadri:1995jp} have been
+\cite{Xu:2009fl}, \citet*{Guo:2007te}, and \cite{Mkhadri:1995jp} have been
 proposed in the literature to adjust the eigenvalues of $\widehat{\bm \Sigma}_k$
@@
 \end{align}
-where $c_{ik}(\lambda) = \lambda N^{-1} + (1 - \lambda)n_k^{-1}I(y_i = k)$. We
-can interpret \eqref{eq:sig-lambda-alternative2} as a covariance-matrix
-estimator that borrows from $\widehat{\bm \Sigma}$ in \eqref{eq:pooled-cov} to
-better estimate $ \bm \Sigma_k$. From \eqref{eq:sig-lambda-alternative2},
-$\lambda$ weights the contribution of each of the $N$ observations in estimating
-$\bm \Sigma_k$. Hence, for $0 < \lambda \le 1$, we estimate $\bm \Sigma_k$ with
-$N$ rather than $n_k$ observations.\footnote{TODO: Add more to the
-  interpretation here.}
+where $c_{ik}(\lambda) = \lambda N^{-1} + (1 - \lambda)n_k^{-1}I(y_i = k)$.
+From \eqref{eq:sig-lambda-alternative2}, $\lambda$ weights the contribution of
+each of the $N$ observations in estimating $\bm \Sigma_k$ from all $K$ classes
+rather than $n_k$ observations from a single class. As a result, we can
+interpret \eqref{eq:sig-lambda-alternative2} as a covariance-matrix estimator
+that borrows from $\widehat{\bm \Sigma}$ in \eqref{eq:pooled-cov} to estimate $
+\bm \Sigma_k$.
 
@@
 define $\bm X = [\sqrt{c_{1k}(1)} \bm x_1^{T}, \ldots, \sqrt{c_{Nk}(1)} \bm
-  x_N^{T}]^{T}$ such that $\widehat{\bm \Sigma} = \bm X^{T} \bm X$. Similarly,
+  x_N^{T}]^{T}$ such that $\widehat{\bm \Sigma} = N^{-1} \bm X^{T} \bm X$. Similarly,
 for $\lambda = 0$, we define $\bm X_k = [\sqrt{c_{1k}(0)} \bm x_1^{T}, \ldots,
-  \sqrt{c_{Nk}(0)} \bm x_N^{T}]^{T}$ such that $\widehat{\bm \Sigma}_k = \bm
+  \sqrt{c_{Nk}(0)} \bm x_N^{T}]^{T}$ such that $\widehat{\bm \Sigma}_k = n_k^{-1} \bm
 X_k^{T} \bm X_k$. In Figure \ref{fig:hdrda-contours} we plot the contours of
@@
 rda_contours(lambda = 0)
-rda_contours(lambda = 1/4)
-rda_contours(lambda = 3/4)
+rda_contours(lambda = 0.2)
+rda_contours(lambda = 0.4)
+rda_contours(lambda = 0.6)
+rda_contours(lambda = 0.8)
 rda_contours(lambda = 1)
@@
 if $\alpha_k = 1 - \gamma$, then \eqref{eq:hdrda-cov} has a form similar to the
-\emph{RDA} classifier from \cite{Friedman:1989tm}. Now, substituting
+\emph{RDA} classifier from \cite{Friedman:1989tm}. Substituting
 \eqref{eq:hdrda-cov} into \eqref{eq:qda}, we define the \emph{HDRDA} classifier
@@
   \Sigma}_k$. Following \cite{Friedman:1989tm}, we select $\lambda$ and $\gamma$
-from a grid of candidate models via cross-validation \citep{Hastie:2008dt}. We
-provide an implementation of \eqref{eq:hdrda} in the {\tt hdrda} function in the
-{\tt sparsediscrim} R package, available on CRAN.
+from a grid of candidate models via cross-validation. We provide an
+implementation of \eqref{eq:hdrda} in the {\tt hdrda} function in the {\tt
+  sparsediscrim} R package, available on CRAN.
 
-\subsection{Relationship with Other Classifiers}
+The choice of $\alpha_k$ in \eqref{eq:hdrda-cov} is one of convenience and
+allows the flexibility of various covariance-matrix estimators proposed in the
+literature. In practice, we generally are not interested in estimating
+$\alpha_k$ as it would incorporate $K$ additional tuning parameters to estimate
+via cross-validation, which is counterproductive to our goal of improving the
+\emph{RDA} classifier's efficiency. By setting $\alpha_k$ to an appropriate
+constant, the \emph{HDRDA} covariance-matrix estimator includes or resembles a
+large family of estimators. Notice that if $\alpha_k = 1$ and $\lambda = 1$,
+\eqref{eq:hdrda-cov} is equivalent to the standard ridge-like covariance-matrix
+estimator in \eqref{eq:ridge-estimator}, and various estimators proposed in the
+literature can be obtained by selecting $\gamma$ accordingly. For instance, with
+$\gamma = \tr\{ \widehat{\bm \Sigma} \} / \min(N, p)$, we obtain the estimator
+from \cite{Srivastava:2007ww}.
 
-The addition of the parameter $\alpha_k$ in \eqref{eq:hdrda-cov} is one of
-convenience and allows the flexibility of various covariance-matrix estimators
-proposed in the literature. In practice, we generally are not interested in
-estimating $\alpha_k$ as it would incorporate $K$ additional tuning parameters
-to estimate via cross-validation, which is counterproductive to our goal of
-improving the \emph{RDA} classifier's efficiency. By setting $\alpha_k$ to a
-predefined constant, the \emph{HDRDA} covariance-matrix estimator includes or
-resembles a large family of estimators, some of which we mention below.
-
-\subsubsection{\cite{Friedman:1989tm}}
-
-\cite{Friedman:1989tm} proposed regularized discriminant analysis (RDA) by
+Perhaps the most well-known of the ridge-like estimators is from
+\cite{Friedman:1989tm}, who proposed regularized discriminant analysis (RDA) by
 incorporating a weighted average of the sample covariance matrix $\widehat{ \bm
@@
 \eqref{eq:qda}. Alternatively, if $\lambda = 1$, \eqref{eq:sig-lambda} reduces
-to \eqref{eq:pooled-cov}.
-
-To further improve the estimation of $\bm \Sigma_k$ and to stabilize the inverse
-of \eqref{eq:sig-lambda}, \cite{Friedman:1989tm} has proposed the biased
-covariance-matrix estimator
+to \eqref{eq:pooled-cov}. To further improve the estimation of $\bm \Sigma_k$
+and to stabilize the inverse of \eqref{eq:sig-lambda}, \cite{Friedman:1989tm}
+has proposed the biased covariance-matrix estimator
 \begin{align}
@@
 of the eigenvalues of \eqref{eq:sig-lambda}. Finally, the RDA classifier is
-established by substituting \eqref{eq:sig-rda} into
-\eqref{eq:qda}.\footnote{TODO: Mention the values of $\alpha_k$ and $\gamma$
-  needed to resemble HDRDA.}
+established by substituting \eqref{eq:sig-rda} into \eqref{eq:qda}.
 
-\subsubsection{Ridge-Like Estimators}
-
-TODO: List a few estimators here along with $\alpha_k = 1$ such that HDRDA
-includes them all. For example, set $\gamma = \text{trace}(\widehat{\bm
-  \Sigma}_k) / \min(n_k, p)$ for Srivastava and Kubokawa.
+Our proposed \emph{HDRDA} classifier resembles \eqref{eq:sig-rda} but improves
+the interpretation of the pooling parameter $\lambda$ as discussed
+above. Moreover, as we discuss below, our formulation of \emph{HDRDA} classifier
+lends itself to rapid estimation in high-dimensional classification.
 
@@
   1.2.5]{Kollo:2005vp}. Now, because $\bm x_i \in \mathcal{C}(\widehat{ \bm
-  \Sigma})$ $(i = 1, \ldots, N)$, $\bm U_2^{T} \bm x_i = \bm 0_{(p-q) \times
-  1}$.  Hence, $\bm 0_{(p-q) \times 1} = \sum_{i=1}^N \beta_i \bm U_2^{T}\bm x_i
-= \bm U_2^{T}(\xbar_k - \xbar_{k'})$, where $\beta_i = (n_k n_{k'})^{-1} \{
-I(y_i = k) n_{k'} - I(y_i = k') n_k \}$. Therefore, $\bm U_2^{T} (\bm x -
-\xbar_k) = \bm U_2^{T} (\bm x - \xbar_{k'})$.
+  \Sigma})$ $(i = 1, \ldots, N)$, $\bm U_2^{T} \bm x_i = \bm 0_{p-q}$.  Hence,
+$\bm 0_{p-q} = \sum_{i=1}^N \beta_i \bm U_2^{T}\bm x_i = \bm U_2^{T}(\xbar_k -
+\xbar_{k'})$, where $\beta_i = (n_k n_{k'})^{-1} \{ I(y_i = k) n_{k'} - I(y_i =
+k') n_k \}$. Therefore, $\bm U_2^{T} (\bm x - \xbar_k) = \bm U_2^{T} (\bm x -
+\xbar_{k'})$.
 \end{proof}
@@
 
-Thus far, we have shown that our proposed classifier in \eqref{eq:hdrda} is
-invariant to the term $\bm U_2$, thus yielding an equivalent classifier in
-\eqref{eq:hdrda-decomposed} with a substantial reduction in computational
-complexity. Here, we demonstrate that the computational efficiency in
-calculating the inverse and determinant of $\bm W_k$ can be further improved via
-standard matrix operations.
+As we can see, we are able to avoid the time-consuming inverses and determinants
+of $p \times p$ covariance matrices in \eqref{eq:hdrda} and instead calculating
+these same operations on $\bm W_k \in \mathbb{R}_{q \times q}$ in
+\eqref{eq:hdrda-decomposed}. The substantial computational improvements arise
+because our proposed classifier in \eqref{eq:hdrda} is invariant to the term
+$\bm U_2$, thus yielding an equivalent classifier in \eqref{eq:hdrda-decomposed}
+with a substantial reduction in computational complexity. Here, we demonstrate
+that the computational efficiency in calculating the inverse and determinant of
+$\bm W_k$ can be further improved via standard matrix operations by showing that
+the inverses and determinants of $\bm W_k$ can be performed on matrices of size
+$n_k \times n_k$.
 
@@
 	\begin{align}
-		\bm W_k^{-1} &= \bm \Gamma_k^{-1} - \alpha_k(1 - \lambda) \bm \Gamma_k^{-1} \bm U_1^{T} \bm X_k^{T} \bm Q_k^{-1} \bm X_k \bm U_1 \bm \Gamma_k^{-1},\label{eq:W_k_inv}
+		\bm W_k^{-1} &= \bm \Gamma_k^{-1} - n_k^{-1} \alpha_k(1 - \lambda) \bm \Gamma_k^{-1} \bm U_1^{T} \bm X_k^{T} \bm Q_k^{-1} \bm X_k \bm U_1 \bm \Gamma_k^{-1},\label{eq:W_k_inv}
     \intertext{where}
-    \bm Q_k &= \bm I_{n_k} + \alpha_k(1 - \lambda) \bm X_k \bm U_1 \bm \Gamma_k^{-1} \bm U_1^{T} \bm X_k^{T}\label{eq:Q_k}
+    \bm Q_k &= \bm I_{n_k} + n_k^{-1} \alpha_k(1 - \lambda) \bm X_k \bm U_1 \bm \Gamma_k^{-1} \bm U_1^{T} \bm X_k^{T}\label{eq:Q_k}
     \intertext{and}
@@
 \begin{proof}
-To calculate $|\bm W_k|$, we apply Theorem 18.1.1 from \cite{Harville:2008wja},
-which states that $|\bm A + \bm B \bm T \bm C| = |\bm A| |\bm T| |\bm T^{-1} +
-\bm C \bm A^{-1} \bm B|$. Thus, setting $\bm A = \bm \Gamma_k$, $\bm B =
-\alpha_k (1 - \lambda) \bm U_1^{T} \bm X_k^{T}$, $\bm T = \bm I_{n_k}$, and $\bm
-C = \bm X_k \bm U_1$, we have $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$. Also,
-\eqref{eq:W_k_inv} follows from the well-known Sherman-Woodbury formula
-\citep[Theorem 18.2.8]{Harville:2008wja}.
+First, we write $\bm W_k = n_k^{-1} \alpha_k (1 - \lambda) \bm U_1^{T} \bm
+X_k^{T} \bm X_k \bm U_1 + \bm \Gamma_k$. To calculate $|\bm W_k|$, we apply
+Theorem 18.1.1 from \cite{Harville:2008wja}, which states that $|\bm A + \bm B
+\bm T \bm C| = |\bm A| |\bm T| |\bm T^{-1} + \bm C \bm A^{-1} \bm B|$. Thus,
+setting $\bm A = \bm \Gamma_k$, $\bm B = \alpha_k (1 - \lambda) \bm U_1^{T} \bm
+X_k^{T}$, $\bm T = \bm I_{n_k}$, and $\bm C = \bm X_k \bm U_1$, we have $|\bm
+W_k| = |\bm \Gamma_k| |\bm Q_k|$. Similarly, \eqref{eq:W_k_inv} follows from the
+well-known Sherman-Woodbury formula \citep[Theorem 18.2.8]{Harville:2008wja} by
+noting that $(\bm A + \bm B \bm T \bm C)^{-1} = \bm A^{-1} - \bm A^{-1} \bm B
+(\bm T^{-1} + \bm C \bm A^{-1} \bm B)^{-1} \bm C \bm A^{-1}$.
 \end{proof}
@@
 The application of the Sherman-Woodbury formula requires that $\bm \Gamma_k$ be
-nonsingular. This assumption is not met when $(\lambda, \gamma) = (0, 0)$ as
-$\bm \Gamma_k = \bm 0_q$, corresponding to the \emph{QDA} classifier. In this
-special case we apply a small amount of eigenvalue shrinkage and set $\gamma =
-0.01$ to circumvent the singularity issue. Also, notice that if $\alpha_k$ is
-constant across the $K$ classes, then $\bm \Gamma_k$ in \eqref{eq:Gamma_k} is
-independent of $k$. Consequently, $|\bm \Gamma_k|$ is constant across the $K$
-classes and need not be calculated in \eqref{eq:hdrda-decomposed}.
+nonsingular. This assumption is not met when $(\lambda, \gamma) = (0, 0)$
+because $\bm \Gamma_k = \bm 0_q$. In this case we instead use the formulation in
+\eqref{eq:hdrda-decomposed}. Also, notice that if $\alpha_k$ is constant across
+the $K$ classes, then $\bm \Gamma_k$ in \eqref{eq:Gamma_k} is independent of
+$k$. Consequently, $|\bm \Gamma_k|$ is constant across the $K$ classes and need
+not be calculated in \eqref{eq:hdrda-decomposed}.
 
@@
 rate for each classifier. Repeating this process 1000 times, we computed the
-average of the error-rate estimates for each classifier. We remark that our
-simulation design avoids the selection bias discussed in \cite{Ambroise:2002fa}.
+average of the error-rate estimates for each classifier.
 
-We used version 3.0.2 of the open-source statistical software {\tt R} for the
+We used version 3.1.0 of the open-source statistical software {\tt R} for the
 classification study. We used the implementations of the Witten and Guo
@@
 {\tt sparsediscrim} package. We used the default settings for each classifier
-with the exception that we explicitly set prior probabilities as equal, if the
-option was available. For the \emph{HDRDA} classifier in
+with the exception that we explicitly set prior probabilities as equal if the
+classifier allowed the option. For the \emph{HDRDA} classifier in
 \eqref{eq:hdrda-decomposed}, we examine the classification performance of two
@@
 $\gamma$ from a grid of 21 equidistant candidate values between 0 and 1,
-inclusively. Each of the microarray data sets listed below can be found in
-\ref{suppA}.
+inclusively. Here, we provide a concise description of each high-dimensional
+data set examined in our classification study.
 
@@
 
-\cite{Burczynski:2006ik} acquired the peripheral blood mononuclear cells (PBMC)
+\cite{Burczynski:2006ik} acquired peripheral blood mononuclear cells (PBMC)
 through hybridization to microarrays from 127 individuals resulting in 22,283
@@
 
-\cite{Nakayama:2007fl} acquired the gene expression through an oligonucleotide
-microarray from 105 samples of 10 types of soft tissue tumors. This included 16
-samples of synovial sarcoma (SS), 19 samples of myxoid/round cell liposarcoma
-(MLS), 3 samples of lipoma, 3 samples of well-differentiated liposarcoma (WDLS),
-15 samples of dedifferentiated liposarcoma (DDLS), 15 samples of
-myxofibrosarcoma (MFS), 6 samples of leiomyosarcoma (LMS), 3 samples of
-malignant nerve sheathe tumor (MPNST), 4 samples of fibrosarcoma (FS), and 21
-samples of malignant fibrous histiocytoma (MFH). \cite{Nakayama:2007fl}
-determined from their data that these 10 types fell into 4 broader groups: (1)
-SS; (2) MLS; (3) Lipoma, WDLS, and part of DDLS; (4) Spindle cell and
-pleomorophic sarcomas including DDLS, MFS, LMS, MPNST, FS, and MFH. Following
-\cite{Witten:2011kc}, we restrict our analysis to the five tumor types having at
-least 15 observations.
+\cite{Nakayama:2007fl} acquired 105 gene-expression samples of 10 types of soft
+tissue tumors through an oligonucleotide microarray, including 16 samples of
+synovial sarcoma (SS), 19 samples of myxoid/round cell liposarcoma (MLS), 3
+samples of lipoma, 3 samples of well-differentiated liposarcoma (WDLS), 15
+samples of dedifferentiated liposarcoma (DDLS), 15 samples of myxofibrosarcoma
+(MFS), 6 samples of leiomyosarcoma (LMS), 3 samples of malignant nerve sheathe
+tumor (MPNST), 4 samples of fibrosarcoma (FS), and 21 samples of malignant
+fibrous histiocytoma (MFH). \cite{Nakayama:2007fl} determined from their data
+that these 10 types fell into 4 broader groups: (1) SS; (2) MLS; (3) Lipoma,
+WDLS, and part of DDLS; (4) Spindle cell and pleomorophic sarcomas including
+DDLS, MFS, LMS, MPNST, FS, and MFH. Following \cite{Witten:2011kc}, we restrict
+our analysis to the five tumor types having at least 15 observations.
 
@@
 <<error_rates, fig.width=8, fig.height=8, out.width='.65\\linewidth'>>=
-error_rates$data_set <- Hmisc:::capitalize(error_rates$data_set)
-p <- ggplot(error_rates, aes(x = d, y = Avg_Error, group = Classifier))
-p <- p + geom_line(aes(color = Classifier), size = 1, linetype = 2)
-p <- p + geom_point(aes(color = Classifier, shape = Classifier), size = 3)
-p <- p + facet_wrap(~ data_set) + ylim(c(0, max(error_rates$Avg_Error))) + theme_bw()
-p <- p + xlab("Number of Variables Selected, d") + ylab("Average Error Rate")
-p <- p + theme(axis.text = element_text(size = 10))
-p <- p + theme(axis.title.x = element_text(size = 17))
-p <- p + theme(axis.title.y = element_text(size = 17))
-p <- p + theme(strip.text.x = element_text(size = 16))
-p <- p + theme(legend.title = element_text(size = 16, face = "bold"))
-p <- p + theme(legend.text = element_text(size = 15))
-p <- p + theme(legend.key.size = unit(1, "cm"))
-p
+print(NULL)
 @ 
@@
 
-In our simulation study, we have compared our proposed \emph{HDRDA} classifier
-to several recently proposed classifiers, including the Dudoit, Pang, Tong, and
-Guo classifiers, each of which assumes independence of the features conditional
-on class membership. Despite the rapid computational performance of diagonal
-classifiers and the reduction in the number of parameters to estimate,
-\cite{Mai:2012bf} and \cite{Fan:2012iq} have noted that diagonal classifiers can
-often yield inferior classification performance to other classification
-methods. We have shown that the \emph{HDRDA} classifier often yields superior
-classification accuracy to the diagonal classifiers considered, confirming the
-assertions of \cite{Mai:2012bf} and \cite{Fan:2012iq}. Furthermore, the
-implementation of \eqref{eq:hdrda-decomposed} in the {\tt sparsediscrim} R
-package was comparable to that of the diagonal classifiers in terms of
-computational runtime. Moreover, when preliminary variable selection was applied
-to the data in our classification study, the time to perform model selection
-with our proposed classifier was nearly instantaneous. Hence, our proposed
-\emph{HDRDA} classifier is competitive in terms of classification accuracy for
-modern small-sample, high-dimensional data sets and requires substantially less
-computing time than standard regularized classifiers, such as regularized
-discriminant analysis proposed by \cite{Friedman:1989tm}.
+We have compared our proposed \emph{HDRDA} classifier to several recently
+proposed classifiers, including the Dudoit, Pang, Tong, and Guo classifiers,
+each of which assumes independence of the features conditional on class
+membership. Despite the rapid computational performance of diagonal classifiers
+and the reduction in the number of parameters to estimate, \cite{Mai:2012bf} and
+\cite{Fan:2012iq} have noted that diagonal classifiers can often yield inferior
+classification performance to other classification methods. We have shown that
+the \emph{HDRDA} classifier often yields superior classification accuracy to the
+diagonal classifiers considered, confirming the assertions of \cite{Mai:2012bf}
+and \cite{Fan:2012iq}. Furthermore, the implementation of
+\eqref{eq:hdrda-decomposed} in the {\tt sparsediscrim} R package was comparable
+to that of the diagonal classifiers in terms of computational runtime. Moreover,
+when preliminary variable selection was applied to the data in our
+classification study, the time to perform model selection with our proposed
+classifier was nearly instantaneous. Hence, our proposed \emph{HDRDA} classifier
+is competitive in terms of classification accuracy for modern small-sample,
+high-dimensional data sets and requires substantially less computing time than
+standard regularized classifiers, such as regularized discriminant analysis
+proposed by \cite{Friedman:1989tm}.
 

### 2015-03-30 sim-microarray.r

@@ -4,8 +4,8 @@ load.project()
 set.seed(42)
 num_cores <- 4
 
-train_pct <- 0.6
-num_iterations <- 100
+train_pct <- 2/3
+num_iterations <- 1000
 
 data_sets <- c("burczynski", "nakayama", "shipp", "singh")
 sim_config <- rep(data_sets, each=num_iterations)

### 2015-03-30 requirements.r
Updated README. Added requirements script

@@ -1,0 +1,12 @@
+source("http://bioconductor.org/biocLite.R")
+biocLite("multtest", ask=FALSE)
+
+cran_packages <- c('futile.logger', 'reshape2', 'plyr', 'ggplot2', 'stringr',
+                   'rda', 'penalizedLDA', 'xtable', 'Hmisc', 'mixtools',
+                   'gridExtra', 'sparseLDA', 'e1071', 'klaR', 'kernlab',
+                   'devtools')
+install.packages(cran_packages, dep=TRUE)
+
+github_packages <- c('ramhiser/datamicroarray', 'ramhiser/sparsediscrim')
+library(devtools)
+devtools::install_github(github_packages)

### 2015-03-30 sim-block-autocorrelation.r
Added sim code for block autocorrelated MVN populations

@@ -1,0 +1,163 @@
+library('ProjectTemplate')
+load.project()
+
+set.seed(42)
+num_cores <- 2
+
+num_iterations <- 1000
+
+K <- 3
+sample_sizes <- rep(25, K)
+autocorrelations <- c(0.1, 0.5, 0.9)
+feature_dim <- seq(50, 500, by=50)
+block_size <- 25
+
+# The number of observations to generate from each population as a test data set.
+test_sizes <- rep(1000, K)
+
+sim_config <- rep(feature_dim, each=num_iterations)
+
+flog.logger("sim", INFO, appender=appender.file('sim-block-autocorrelation.log'))
+
+results <- mclapply(seq_along(sim_config), function(i) {
+  p <- sim_config[i]
+  flog.info("Dimension: %s -- Sim:  %s of %s", p, i, length(sim_config), name="sim")
+  set.seed(i)
+
+  mu1 <- rep(0, p)
+  mu2 <- c(rep(1/2, p), rep(0, p - 50))
+  mu3 <- c(rep(-1/2, p), rep(0, p - 50))
+  num_blocks <- p / block_size
+
+  # Generates training data
+  train_data <- generate_blockdiag(n=sample_sizes,
+                                   mu=cbind(mu1, mu2, mu3),
+                                   block_size=block_size,
+                                   num_blocks=num_blocks,
+                                   rho=autocorrelations)
+
+  train_x <- train_data$x
+  train_y <- train_data$y
+
+  # Generates test data
+  test_data <- generate_blockdiag(n=test_sizes,
+                                  mu=cbind(mu1, mu2, mu3),
+                                  block_size=block_size,
+                                  num_blocks=num_blocks,
+                                  rho=autocorrelations)
+
+  test_x <- test_data$x
+  test_y <- test_data$y
+
+  # Removes the loaded data from memory and then garbage collects.
+  rm(train_data)
+  rm(test_data)
+  gc(reset = TRUE)
+
+  # Sets the prior probabilities as equal
+  num_classes <- nlevels(train_y)
+  prior_probs <- rep(1, num_classes) / num_classes
+
+  # kNN
+  knn_errors <- try({
+      tune.knn_out <- tune.knn(x = train_x,
+                               y = train_y,
+                               k = 1:5,
+                               tunecontrol = tune.control(sampling = "cross", cross = 10))
+      knn_out <- knn(train = train_x,
[…]

### 2015-03-30 requirements.r
Added random forest to the sims.

@@ -4,7 +4,7 @@ biocLite("multtest", ask=FALSE)
 cran_packages <- c('futile.logger', 'reshape2', 'plyr', 'ggplot2', 'stringr',
                    'rda', 'penalizedLDA', 'xtable', 'Hmisc', 'mixtools',
                    'gridExtra', 'sparseLDA', 'e1071', 'klaR', 'kernlab',
-                   'devtools')
+                   'devtools', 'randomForest')
 install.packages(cran_packages, dep=TRUE)
 
 github_packages <- c('ramhiser/datamicroarray', 'ramhiser/sparsediscrim')

### 2015-03-30 sim-block-autocorrelation.r

@@ -71,6 +71,16 @@ results <- mclapply(seq_along(sim_config), function(i) {
       mean(knn_out != test_y)
   })
 
+  # Random Forest
+  rf_errors <- try({
+    rf_out <- randomForest(x=train_x,
+                           y=train_y,
+                           ntree=500,
+                           maxnodes=100)
+    rf_predict <- predict(rf_out, test_x)
+    mean(rf_predict != test_y)
+  })
+
   # SVM with Radial Basis Functions
   ksvm_radial_errors <- try({
       ksvm_out <- ksvm(x = train_x,
@@ -144,6 +154,7 @@ results <- mclapply(seq_along(sim_config), function(i) {
     HDRDA_Convex = hdrda_convex_errors,
     kNN = knn_errors,
     Pang = Pang_errors,
+    Random_Forest = rf_errors,
     SVM_Radial = ksvm_radial_errors,
     Tong = Tong_errors,
     Witten = Witten_errors

### 2015-03-30 sim-microarray.r

@@ -32,7 +32,7 @@ results <- mclapply(seq_along(sim_config), function(i) {
 
   # Sets the prior probabilities as equal
   num_classes <- nlevels(train_y)
-  prior_probs <- rep(1, num_classes) / num_classes  
+  prior_probs <- rep(1, num_classes) / num_classes
 
   # kNN
   knn_errors <- try({
@@ -45,6 +45,16 @@ results <- mclapply(seq_along(sim_config), function(i) {
                      test = test_x,
                      k = tune.knn_out$best.model$k)
       mean(knn_out != test_y)
+  })
+
+  # Random Forest
+  rf_errors <- try({
+    rf_out <- randomForest(x=train_x,
+                           y=train_y,
+                           ntree=500,
+                           maxnodes=100)
+    rf_predict <- predict(rf_out, test_x)
+    mean(rf_predict != test_y)
   })
 
   # SVM with Radial Basis Functions
@@ -120,6 +130,7 @@ results <- mclapply(seq_along(sim_config), function(i) {
     HDRDA_Convex = hdrda_convex_errors,
     kNN = knn_errors,
     Pang = Pang_errors,
+    Random_Forest = rf_errors,
     SVM_Radial = ksvm_radial_errors,
     Tong = Tong_errors,
     Witten = Witten_errors

### 2015-03-30 lib/helpers.R
Initial contaminated MVN sim. Needs some TLC

@@ -272,7 +272,7 @@ Clemmensen <- function(train_x, train_y, test_x, normalize_data = TRUE,
 }
 
 # Classifier from Cao, Boitard, and Besse (2011) - BMC Bioinformatics
-# We use the 'max.dist' 
+# We use the 'max.dist'
 Cao <- function(train_x, train_y, test_x) {
   plsda_out <- plsda(X = train_x, Y = train_y)
   plsda_predict <- predict(plsda_out, test_x, method = "max.dist")
@@ -292,7 +292,7 @@ Guo <- function(train_x, train_y, test_x, prior = NULL) {
 
   # Annoying: Class labels must be specified as integers
   train_y_int <- as.integer(train_y)
-  
+
   rda_out <- rda:::rda(x = train_x, y = train_y_int, prior = prior)
   rda_cv_out <- rda:::rda.cv(rda_out, x = train_x, y = train_y_int, prior = prior)
 
@@ -320,3 +320,61 @@ Guo <- function(train_x, train_y, test_x, prior = NULL) {
   factor(rda_class, levels = seq_along(group_names), labels = group_names)
 }
 
+
+#' Generates contaminated MVN populations with block-autocorrelated cov matrices
+#'
+#' @param n vector of the sample sizes of each class. The length of \code{n}
+#' determines the number of classes \code{K}.
+#' @param mu matrix containing the mean vectors for each class. Expected to have
+#' \code{p} rows and \code{K} columns.
+#' @param num_blocks the number of block matrices. See details.
+#' @param block_size the dimensions of the square block matrix. See details.
+#' @param rho vector of the values of the autocorrelation parameter for each
+#' class covariance matrix. Must equal the length of \code{n} (i.e., equal to
+#' \code{K}).
+#' @param uncontaminated_var vector of the variance coefficients for each class
+#' covariance matrix. Must equal the length of \code{n} (i.e., equal to
+#' \code{K}).
+#' @param contaminated_var vector of the variance coefficients for each class
+#' covariance matrix when an observation is contaminated. Must equal the length
+#' of \code{n} (i.e., equal to \code{K}).
+#' @param contaminated_prob probability that an observation is drawn from its
+#' class' contaminated distribution
+#' @return named list with elements:
+#' \itemize{
+#'   \item \code{x}: matrix of observations with \code{n} rows and \code{p}
+#' columns
+#'   \item \code{y}: vector of class labels that indicates class membership for
[…]

### 2015-03-30 sim-block-contaminated.r

@@ -1,0 +1,225 @@
+library('ProjectTemplate')
+load.project()
+
+set.seed(42)
+num_cores <- 2
+
+#num_iterations <- 1000
+num_iterations <- 10
+
+K <- 3
+sample_sizes <- rep(25, K)
+autocorrelations <- c(0.1, 0.5, 0.9)
+#feature_dim <- seq(50, 500, by=50)
+feature_dim <- 50
+block_size <- 25
+
+# The number of observations to generate from each population as a test data set.
+test_sizes <- rep(1000, K)
+
+sim_config <- rep(feature_dim, each=num_iterations)
+
+flog.logger("sim", INFO, appender=appender.file('sim-block-autocorrelation-contaminated.log'))
+
+n <- sample_sizes
+K <- length(n)
+uncontaminated_var <- rep(1, K)
+contaminated_var <- rep(100, K)
+
+contamination_prob <- 0.1
+
+num_uncontaminated <- rbinom(rep(1, K), n, prob=1 - contamination_prob)
+num_contaminated <- n - num_uncontaminated
+
+
+# TODO:
+# 1. Fix error when 0 contaminated samples for a class
+# 2. Update simulation to iterate through several values of p as well as
+#    contamination probs 0.05, 0.1, 0.15. NOTE: Original block sims are
+#    contamination prob = 0. Might be useful to draw this connection in paper.
+# 3. Finalize sim code
+# 4. Try code locally with small number of reps
+# 5. Run all the sims on EC2 instance
+
+uncontaminated_data <- generate_blockdiag(n=num_uncontaminated,
+                                          mu=cbind(mu1, mu2, mu3),
+                                          block_size=block_size,
+                                          num_blocks=num_blocks,
+                                          rho=autocorrelations,
+                                          sigma2=uncontaminated_var)
+
+contaminated_data <- generate_blockdiag(n=num_contaminated,
+                                        mu=cbind(mu1, mu2, mu3),
+                                        block_size=block_size,
+                                        num_blocks=num_blocks,
+                                        rho=autocorrelations,
+                                        sigma2=contaminated_var)
+
+
+contaminated_data <- generate_contaminated(n=sample_sizes,
+                                           mu=cbind(mu1, mu2, mu3),
+                                           block_size=block_size,
+                                           num_blocks=num_blocks,
[…]

### 2015-04-02 lib/helpers.R
Finished up contaminated MVN sims.

@@ -358,23 +358,25 @@ generate_contaminated <- function(n, mu, num_blocks, block_size, rho,
   mu <- as.matrix(mu)
 
   K <- length(n)
-  num_uncontaminated <- rbinom(rep(1, K), n, prob=1 - contamination_prob)
-  num_contaminated <- n - num_uncontaminated
+  num_contaminated <- rbinom(rep(1, K), n, prob=contamination_prob)
+  num_uncontaminated <- n - num_contaminated
 
   uncontaminated_data <- generate_blockdiag(n=num_uncontaminated,
-                                            mu=cbind(mu1, mu2, mu3),
+                                            mu=mu,
                                             block_size=block_size,
                                             num_blocks=num_blocks,
                                             rho=autocorrelations,
                                             sigma2=uncontaminated_var)
 
   contaminated_data <- generate_blockdiag(n=num_contaminated,
-                                          mu=cbind(mu1, mu2, mu3),
+                                          mu=mu,
                                           block_size=block_size,
                                           num_blocks=num_blocks,
                                           rho=autocorrelations,
                                           sigma2=contaminated_var)
 
-  list(x=rbind(uncontaminated_data$x, contaminated_data$x),
-       y=c(uncontaminated_data$y, contaminated_data$y))
+  x <- rbind(uncontaminated_data$x, contaminated_data$x)
+  y <- c(as.character(uncontaminated_data$y),
+         as.character(contaminated_data$y))
+  list(x=x, y=factor(y))
 }

### 2015-04-02 requirements.r

@@ -7,6 +7,7 @@ cran_packages <- c('futile.logger', 'reshape2', 'plyr', 'ggplot2', 'stringr',
                    'devtools', 'randomForest')
 install.packages(cran_packages, dep=TRUE)
 
-github_packages <- c('ramhiser/datamicroarray', 'ramhiser/sparsediscrim')
+github_packages <- c('ramhiser/datamicroarray', 'ramhiser/sparsediscrim',
+                     'ramhiser/itertools2')
 library(devtools)
 devtools::install_github(github_packages)

### 2015-04-02 sim-block-contaminated.r

@@ -4,75 +4,34 @@ load.project()
 set.seed(42)
 num_cores <- 2
 
-#num_iterations <- 1000
-num_iterations <- 10
+num_iterations <- 1000
 
 K <- 3
 sample_sizes <- rep(25, K)
 autocorrelations <- c(0.1, 0.5, 0.9)
-#feature_dim <- seq(50, 500, by=50)
-feature_dim <- 50
+feature_dim <- seq(50, 500, by=50)
 block_size <- 25
+contamination_probs <- c(0.05, 0.1, 0.15)
+uncontaminated_var <- rep(1, K)
+contaminated_var <- rep(100, K)
 
 # The number of observations to generate from each population as a test data set.
 test_sizes <- rep(1000, K)
 
-sim_config <- rep(feature_dim, each=num_iterations)
+iter_product <- itertools2::iproduct(p=feature_dim,
+                                     contamination_prob=contamination_probs,
+                                     times=num_iterations)
+sim_config <- itertools2::ienum(iter_product)
+num_iterations <- length(feature_dim) * length(contamination_probs) * num_iterations
 
-flog.logger("sim", INFO, appender=appender.file('sim-block-autocorrelation-contaminated.log'))
+flog.logger("sim", INFO, appender=appender.file('sim-block-contaminated.log'))
 
-n <- sample_sizes
-K <- length(n)
-uncontaminated_var <- rep(1, K)
-contaminated_var <- rep(100, K)
-
-contamination_prob <- 0.1
-
-num_uncontaminated <- rbinom(rep(1, K), n, prob=1 - contamination_prob)
-num_contaminated <- n - num_uncontaminated
-
-
-# TODO:
-# 1. Fix error when 0 contaminated samples for a class
-# 2. Update simulation to iterate through several values of p as well as
-#    contamination probs 0.05, 0.1, 0.15. NOTE: Original block sims are
-#    contamination prob = 0. Might be useful to draw this connection in paper.
-# 3. Finalize sim code
-# 4. Try code locally with small number of reps
-# 5. Run all the sims on EC2 instance
-
-uncontaminated_data <- generate_blockdiag(n=num_uncontaminated,
-                                          mu=cbind(mu1, mu2, mu3),
-                                          block_size=block_size,
-                                          num_blocks=num_blocks,
-                                          rho=autocorrelations,
-                                          sigma2=uncontaminated_var)
-
-contaminated_data <- generate_blockdiag(n=num_contaminated,
-                                        mu=cbind(mu1, mu2, mu3),
[…]

### 2015-10-18 sim-microarray.r
Rm'd unnecessary classifiers/bloat from data sims

@@ -40,22 +40,6 @@ results <- mclapply(seq_along(sim_config), function(i) {
   num_classes <- nlevels(train_y)
   prior_probs <- rep(1, num_classes) / num_classes
 
-  # kNN
-  knn_errors <- try({
-      tune.knn_out <- tune.knn(x=train_x,
-                               y=train_y,
-                               k=1:5,
-                               tunecontrol=tune.control(sampling="cross", cross=10))
-      knn_out <- knn(train=train_x,
-                     cl=train_y,
-                     test=test_x,
-                     k=tune.knn_out$best.model$k)
-      mean(knn_out != test_y)
-  })
-
-  flog.info("kNN Error Rate: %s -- Data Set: %s -- Sim Config: %s of %s",
-            knn_errors, data_set, i, length(sim_config), name="sim")
-
   # Random Forest
   rf_errors <- try({
     rf_out <- randomForest(x=train_x,
@@ -65,23 +49,10 @@ results <- mclapply(seq_along(sim_config), function(i) {
     rf_predict <- predict(rf_out, test_x)
     mean(rf_predict != test_y)
   })
+  rm("rf_out")
 
   flog.info("Random Forest Error Rate: %s -- Data Set: %s -- Sim Config: %s of %s",
             rf_errors, data_set, i, length(sim_config), name="sim")
-
-  # SVM with Radial Basis Functions
-  ksvm_radial_errors <- try({
-      ksvm_out <- ksvm(x=train_x,
-                       y=train_y,
-                       kernel="rbfdot",
-                       kpar="automatic",
-                       cross=10)
-      ksvm_radial <- predict(ksvm_out, test_x)
-      mean(ksvm_radial != test_y)
-  })
-
-  flog.info("SVM Error Rate: %s -- Data Set: %s -- Sim Config: %s of %s",
-            ksvm_radial_errors, data_set, i, length(sim_config), name="sim")
 
   # Witten and Tibshirani (2011) - JRSS B
   Witten_errors <- try({
@@ -90,6 +61,7 @@ results <- mclapply(seq_along(sim_config), function(i) {
                                       test_x=test_x)
       mean(Witten_out$predictions != test_y)
   })
+  rm("Witten_out")
 
   flog.info("Witten Error Rate: %s -- Data Set: %s -- Sim Config: %s of %s",
[…]

### 2015-10-19 sim-microarray.r
Increased num_iterations to 250 in data sims

@@ -5,7 +5,7 @@ set.seed(42)
 num_cores <- 8
 
 train_pct <- 2/3
-num_iterations <- 50
+num_iterations <- 250
 
 data_sets <- c("burczynski", "nakayama", "shipp", "singh")
 sim_config <- rep(data_sets, each=num_iterations)

### 2015-10-21 reports/paper-hdrda.Rnw
New description of sim design: rough for now.

@@
 
-\author{John A. Ramey, Caleb K. Stein, and Dean M. Young}
+\author{John A. Ramey, Caleb K. Stein, Phil D. Young, and Dean M. Young}
 
@@
 high-dimensional data from \cite{Tong:2012hw}, \cite{Witten:2011kc},
-\cite{Pang:2009ik}, \cite{Guo:2007te}, and \cite{Dudoit:2002ev}. We also include
-as benchmarks support vector machines (\emph{SVM}) and $k$-nearest neigbors
-($k$-NN) in our classification study.\footnote{TODO: If we change any of the
-  classifiers in the simulations, they should be updated here.}
+\cite{Pang:2009ik}, and \cite{Guo:2007te}. We also include as a benchmark the
+random forest from \cite{Breiman:2001fb} because \cite{FernandezDelgado:2014ul}
+has concluded that the random forest is often superior to other classifiers in
+benchmark studies.
 
@@
 
-In this section, we describe four high-dimensional microarray data sets and
-compare our proposed classifier with five classifiers recently proposed for
-small-sample, high-dimensional data. Our competitors include diagonal linear
-discriminant analysis from \cite{Dudoit:2002ev}, shrunken centroids regularized
-discriminant analysis from \cite{Guo:2007te}, and penalized linear discriminant
-analysis from \cite{Witten:2011kc}. We also consider two modifications of
-diagonal linear discriminant analysis from \cite{Tong:2012hw} and
-\cite{Pang:2009ik}, where the former employs an improved mean estimator and the
-latter utilizes an improved variance estimator. Below, we refer to each
-classifier by the first author's surname.\footnote{TODO: Consider adding a
-  heatmap of error rates of a single $(\lambda, \gamma)$ grid.}
-
-We evaluated the classification performance of each classifier by randomly
-partitioning the data set under consideration such that $2/3$ of the
-observations were allocated as training data and the remaining $1/3$ of the
-observations were allocated as a test data set. For each random partition, we
-applied the feature-selection method from \cite{Dudoit:2002ev} to the training
-data to choose the $d$ features having the largest ratio of their between-group
-to within-group sums of squares.  We then filtered the test data set with the
-corresponding $d$ features. After training each classifier from the filtered
-training data, we classified the filtered test data sets and computed the
-proportion of mislabeled test observations to estimate the classification error
-rate for each classifier. Repeating this process 1000 times, we computed the
-average of the error-rate estimates for each classifier.
-
-We used version 3.1.0 of the open-source statistical software {\tt R} for the
+In this section, we compare our proposed classifier with four classifiers
+recently proposed for small-sample, high-dimensional data along with the
+random-forest classifier from \cite{Breiman:2001fb} as a benchmark. Our
+competitors include shrunken centroids regularized discriminant analysis from
+\cite{Guo:2007te} and penalized linear discriminant analysis from
+\cite{Witten:2011kc}. We also consider two modifications of diagonal linear
+discriminant analysis from \cite{Tong:2012hw} and \cite{Pang:2009ik}, where the
+former employs an improved mean estimator and the latter utilizes an improved
+variance estimator. Below, we refer to each classifier by the first author's
+surname.\footnote{TODO: Consider adding a heatmap of error rates of a single
+  $(\lambda, \gamma)$ grid.}
+
+We used version 3.2.2 of the open-source statistical software {\tt R} for the
 classification study. We used the implementations of the Witten and Guo
-classifiers from the {\tt penalizedLDA} and {\tt rda} packages, respectively, on
-CRAN. Because the {\tt rda} package does not perform the authors' ``Min-Min''
-rule automatically, we applied this rule in our {\tt R} code. We also utilized
-our implementations of the Dudoit, Pang, and Tong classifiers available in the
-{\tt sparsediscrim} package. We used the default settings for each classifier
-with the exception that we explicitly set prior probabilities as equal if the
-classifier allowed the option. For the \emph{HDRDA} classifier in
-\eqref{eq:hdrda-decomposed}, we examine the classification performance of two
-models. For the first \emph{HDRDA} model, we set $\alpha_k = 1$, $k = 1, \ldots,
-K$, so that the covariance-matrix estimator \eqref{eq:hdrda-cov} employed
-resembled \eqref{eq:ridge-estimator}. We estimated $\lambda$ from a grid of 21
-equidistant candidate values between 0 and 1, inclusively. Similarly, we
-estimated $\gamma$ from a grid consisting of the values $10^{-1}, \ldots, 10^4$,
-and $10^5$. We selected optimal estimates of $\lambda$ and $\gamma$ using
-$V$-fold cross-validation with $V = 10$ \citep{Hastie:2008dt}. For the second
-model, we set $\alpha_k = 1 - \gamma$, $k = 1, \ldots, K$, to resemble
-\citeauthor{Friedman:1989tm}'s parameterization. We estimated both $\lambda$ and
-$\gamma$ from a grid of 21 equidistant candidate values between 0 and 1,
-inclusively. Here, we provide a concise description of each high-dimensional
+classifiers from the {\tt penalizedLDA} and {\tt rda} packages, respectively.
+Because the {\tt rda} package does not perform the authors' ``Min-Min'' rule
+automatically, we applied this rule within our {\tt R} code. We also utilized
+the implementations of the Pang and Tong classifiers available in the {\tt
+  sparsediscrim} package. We used the implementation of the random-forest
+classifier from the {\tt randomForest} package. For each classifer we explicitly
+set prior probabilities as equal if applicable. All other classifier options
+were set to their default settings.
+
+For the \emph{HDRDA} classifier in \eqref{eq:hdrda-decomposed}, we examine the
+classification performance of two models. For the first \emph{HDRDA} model, we
+set $\alpha_k = 1$, $k = 1, \ldots, K$, so that the covariance-matrix estimator
+\eqref{eq:hdrda-cov} employed resembled \eqref{eq:ridge-estimator}. We estimated
+$\lambda$ from a grid of 21 equidistant candidate values between 0 and 1,
+inclusively. Similarly, we estimated $\gamma$ from a grid consisting of the
+values $10^{-1}, \ldots, 10^4$, and $10^5$. We selected optimal estimates of
+$\lambda$ and $\gamma$ using $V$-fold cross-validation with $V = 10$
+\citep{Hastie:2008dt}. For the second model, we set $\alpha_k = 1 - \gamma$, $k
+= 1, \ldots, K$, to resemble \citeauthor{Friedman:1989tm}'s parameterization. We
+estimated both $\lambda$ and $\gamma$ from a grid of 21 equidistant candidate
+values between 0 and 1, inclusively.
+
+\subsection{Simulation Study}
+
+We compared the \emph{HDRDA} classifer to the five competing classifiers in a
+simulaton study. We simulated $K = 3$ $p$-dimensional multivariate normal
+populations with a common positive-definite covariance matrix having a
+block-diagonal covariance structure, which is intended to resemble
+gene-expression data. That is, within each block of pathways, genes are
+positively correlated and that this correlation decays as a function of the
+distance between any two genes. Thus, we simulated $K = 3$ multivariate normal
+populations with the common positive-definite covariance matrix
+
+\begin{align*}
+  \bm\Sigma = \bm\Sigma^{(\rho)} \oplus \bm\Sigma^{(-\rho)} \oplus \ldots \oplus \bm\Sigma^{(\rho)},
+\end{align*}
+
+where the $(i,j)$th entry of $\bm\Sigma^{(\rho)} \in \mathbb{R}_{25 \times 25}$ is
+
+\begin{align*}
+\bm\Sigma_{ij}^{(\rho)} = \{ \rho^{|i - j|} \}_{1 \le i,j \le 25},
+\end{align*}
+
+altering positive and negative values of $\rho$ are used from block to block. In
+our simulation study, we considered three different values of $\rho$, 0.1, 0.5,
+and 0.9, to explore the effect of autocorrelation on classification accuracy,
+especially given that the classifiers from \cite{Tong:2012hw} and
+\cite{Pang:2009ik} assume independence among features.
+
+Furthermore, we set the population of mean of class 1 to be a $p$-dimensional
+vector of zeros. Next, we set class 2's population mean to a vector equaling 1/2
+for the first 50 features and 0 after ($\mu_j = 1/2$ if $j \le 50$ and $\mu_j =
+0$ if $j > 50$). Similarly, we set class 3's population mean to a vector equaling -1/2
+for the first 50 features and 0 after ($\mu_j = -1/2$ if $j \le 50$ and $\mu_j =
+0$ if $j > 50$). The three populations differ in their mean vectors in the first
+50 features corresponding to the first two blocks. No difference in the means
+occurs in the remaining blocks.
+
+The block-diagonal simulation design considered is widely used within the
+high-dimensional classification literature, including the studies by
+\cite{Ramey:2013ji}, \cite{Witten:2011kc}, and \cite{Guo:2007te}. While the
+design is indeed standard, the simulation configuration is based on an ideal
+scenario where no artifacts, often observed in real data, are present. As a
+result, we wished to investigate the effect of outliers on the high-dimensional
+classifiers. To do this, we generalized the block-diagonal simulation
+configuration by sampling from a $p$-dimensional multivariate
+$\epsilon$-contaminated normal distribution, where $\epsilon in [0, 1]$. For
+$\epsilon = 0$, we have the benchmark block-diagonal simulations considered in
+the literature. As $\epsilon$ is increased, the number of outliers is increased
+(on average). We consider the values of $\epsilon = 0, 0.05, 0.1, 0.15,
+0.2$. For the contamination simulation, we define the probability density
+function of the $k$th class to be
+
+\begin{align*}
+  p(\bm x | \omega_k) = (1 - \epsilon) N_p(\bm x | \bm \mu_k, \bm \Sigma) + \epsilon N_p(\bm x | \bm \mu_k, 100 * \bm \Sigma)
+\end{align*}
+
+From each of the $K=3$ populations, we sampled 25 training observations ($n_k =
+25$ for all $k$) and 1000 test observations. After training each classifier on
+the training data, we classified the test data sets and computed the proportion
+of mislabeled test observations to estimate the classification error rate for
+each classifier. Repeating this process 250 times, we computed the average of
+the error-rate estimates for each classifier. We allowed the number of features
+to vary from $p = 100$ to $p = 500$ in increments of 100 to examine the increase
+in dimension while maintaining a small sample size.
+
+Despite the claims of \cite{FernandezDelgado:2014ul} that the random forest
+classifier is typically superior, we observed that the classifiers designed for
+high-dimensional data were indeed superior.
+
+\subsection{Application to Gene Expression Data}
+
+We compared the \emph{HDRDA} classifer to the five competing classifiers on four
+benchmark gene-expression microarray data sets. We evaluated the classification
+performance of each classifier by randomly partitioning the data set under
+consideration such that $2/3$ of the observations were allocated as training
+data and the remaining $1/3$ of the observations were allocated as a test data
+set. After training each classifier on the training data, we classified the test
+data sets and computed the proportion of mislabeled test observations to
+estimate the classification error rate for each classifier. Repeating this
+process 250 times, we computed the average of the error-rate estimates for each
+classifier. Here, we provide a concise description of each high-dimensional
 data set examined in our classification study.
 
-\subsection{\cite{Burczynski:2006ik} Data Set}
+\subsubsection{\cite{Burczynski:2006ik} Data Set}
 
@@
 
-\subsection{\cite{Nakayama:2007fl} Data Set}
+\subsubsection{\cite{Nakayama:2007fl} Data Set}
 
@@
 
-\subsection{\cite{Shipp:2002ka} Data Set}
+\subsubsection{\cite{Shipp:2002ka} Data Set}
 
@@
 
-\subsection{\cite{Singh:2002fh} Data Set}
+\subsubsection{\cite{Singh:2002fh} Data Set}
 

### 2015-10-24 reports/paper-hdrda.Rnw
Improvements to the intro.

--- 2015-03-30 reports/paper-hdrda.Rnw
@@
 As machine-learning algorithms are increasingly utilized to automate data-driven
-decisions, researchers continue to develop a wealth of \emph{classifiers} to
-construct automated pattern-recognition systems in a variety of disciplines,
-such as medical research, facial recognition, remote sensing, and chemical
-analysis \citep{Ching:2012fu,Li:2012ev,Zhang:2010va}. Despite the advances in
+decisions, researchers continue to develop a wealth of \emph{classifiers}
+employed in a variety of disciplines, such as medical research, facial
+recognition, remote sensing, and chemical analysis
+\citep{Ching:2012fu,Li:2012ev,Zhang:2010va}. Despite the advances in
 classification to facilitate technological advancements, well-established
@@
 
-We consider regularized discriminant analysis (\emph{RDA}) from a classic paper
-by \cite{Friedman:1989tm}, who proposed improvements to the \emph{QDA}
-classifier by employing a covariance matrix estimator that partially pools the
-individual sample covariance matrices with the pooled sample covariance matrix
-from the \emph{LDA} classifier. Given that the assumption of equal population
-covariance matrices for the \emph{LDA} classifier is often too stringent in
-practice \citep{Clemmensen:2011kr}, the \emph{RDA} classifier's partial pooling
-can relax the linearity assumption by instead incorporating unequal covariance
-matrices while retaining as special cases the \emph{LDA} and QDA
-classifiers. Covariance-matrix regularization is then applied to the partially
-pooled covariance-matrix estimator by shrinking it towards a scaled identity
-matrix to ensure positive definiteness.
+Here, we present a new classification method intended for the $p > N$ case that
+builds on the classic regularized discriminant analysis (\emph{RDA}) proposed by
+\cite{Friedman:1989tm}. Our proposed high-dimensional \emph{RDA} (\emph{HDRDA})
+classifier employs a biased covariance matrix estimator that partially pools the
+individual sample covariance matrices from the \emph{QDA} classifier with the
+pooled sample covariance matrix from the \emph{LDA} classifier. We then shrink
+the resulting covariance-matrix estimator towards a scaled identity matrix to
+ensure positive definiteness. By reparameterizing the \emph{RDA} classifier
+similar to \cite{Hastie:2008dt} and \cite{Halbe:2007ea}, we first show the
+pooling parameter in the \emph{HDRDA} classifier determines the contribution of
+each training observation to the estimation of each class covariance matrix,
+enabling interpretability that has been previously lacking with the \emph{RDA}
+classifier \citep{Bensmail:1996fw}.
 
-We reparameterize the \emph{RDA} classifier similar to \cite{Hastie:2008dt} and
-\cite{Halbe:2007ea} and refer to our parameterization as the high-dimensional
-\emph{RDA} (\emph{HDRDA}) classifier. First, we show the pooling parameter in
-the \emph{HDRDA} classifier determines the contribution of each training
-observation to the estimation of each class covariance matrix, enabling
-interpretability that has been previously lacking with the \emph{RDA} classifier
-\citep{Bensmail:1996fw}. Next, applying a reasoning similar to \cite{Ye:2006tq},
-we establish that the matrix operations corresponding to the null space of the
-pooled sample covariance matrix are redundant and can be discarded from the
-\emph{HDRDA} decision rule without loss of classificatory information. Moreover,
-because the \emph{HDRDA} decision rule is invariant to adjustments to the zero
-eigenvalues, the decision rule in the original feature space is equivalent to a
-decision rule in a much lower dimension (i.e., the intrinsic dimension), such
-that the matrix inverses and determinants of relatively small matrices can be
-rapidly computed. Thus, the dimension reduction inherent in the \emph{HDRDA}
-classifier yields a substantial reduction in computation during model
-selection. Also, as a result, we show that a large number of shrinkage methods
-that are special cases of the \emph{HDRDA} classifier have no effect on the zero
+Next, we achieve a substantial reduction in dimension to reduce greatly the
+computation during model selection by applying a reasoning similar to
+\cite{Ye:2006tq}. We establish that the matrix operations corresponding to the
+null space of the pooled sample covariance matrix are redundant and can be
+discarded from the \emph{HDRDA} decision rule without loss of classificatory
+information. Moreover, because the \emph{HDRDA} decision rule is invariant to
+adjustments to the zero eigenvalues, the decision rule in the original feature
+space is equivalent to a decision rule in a much lower dimension, such that the
+matrix inverses and determinants of relatively small matrices can be rapidly
+computed. Furthermore, we show that a large number of shrinkage methods that are
+special cases of the \emph{HDRDA} classifier have no effect on the zero
 eigenvalues of the covariance-matrix estimators. Such methods include work from
@@
 We provide an efficient algorithm that expedites the model-selection process for
-the \emph{HDRDA} classifier and then compare its computational runtime with an
-implementation of the \emph{RDA} classifier from the {\tt klaR} R package to
-illustrate that the \emph{HDRDA} classifier greatly reduces computation when $p
-\gg N$. We then study the classification performance of the \emph{HDRDA}
-classifier on three artificial and four real high-dimensional data sets. We
-demonstrate that the \emph{HDRDA} classifier often attains superior
-classification accuracy to several recent classifiers designed for small-sample,
-high-dimensional data from \cite{Tong:2012hw}, \cite{Witten:2011kc},
-\cite{Pang:2009ik}, \cite{Guo:2007te}, and \cite{Dudoit:2002ev}. We also include
-as benchmarks support vector machines (\emph{SVM}) and $k$-nearest neigbors
-($k$-NN) in our classification study.\footnote{TODO: If we change any of the
-  classifiers in the simulations, they should be updated here.}
+the \emph{HDRDA} classifier when $p \gg N$. We then study the classification
+performance of the \emph{HDRDA} classifier on three artificial and four real
+high-dimensional data sets. We demonstrate that the \emph{HDRDA} classifier
+often attains superior classification accuracy to several recent classifiers
+designed for small-sample, high-dimensional data from \cite{Tong:2012hw},
+\cite{Witten:2011kc}, \cite{Pang:2009ik}, \cite{Guo:2007te}, and
+\cite{Dudoit:2002ev}. We also include as benchmarks support vector machines
+(\emph{SVM}) and $k$-nearest neigbors ($k$-NN) in our classification
+study.\footnote{TODO: If we change any of the classifiers in the simulations,
+  they should be updated here.}
 

### 2015-10-24 reports/paper-hdrda.Rnw

--- 2015-10-21 reports/paper-hdrda.Rnw
@@
 
-We consider regularized discriminant analysis (\emph{RDA}) from a classic paper
-by \cite{Friedman:1989tm}, who proposed improvements to the \emph{QDA}
-classifier by employing a covariance matrix estimator that partially pools the
-individual sample covariance matrices with the pooled sample covariance matrix
-from the \emph{LDA} classifier. Given that the assumption of equal population
-covariance matrices for the \emph{LDA} classifier is often too stringent in
-practice \citep{Clemmensen:2011kr}, the \emph{RDA} classifier's partial pooling
-can relax the linearity assumption by instead incorporating unequal covariance
-matrices while retaining as special cases the \emph{LDA} and QDA
-classifiers. Covariance-matrix regularization is then applied to the partially
-pooled covariance-matrix estimator by shrinking it towards a scaled identity
-matrix to ensure positive definiteness.
-
-We reparameterize the \emph{RDA} classifier similar to \cite{Hastie:2008dt} and
-\cite{Halbe:2007ea} and refer to our parameterization as the high-dimensional
-\emph{RDA} (\emph{HDRDA}) classifier. First, we show the pooling parameter in
-the \emph{HDRDA} classifier determines the contribution of each training
-observation to the estimation of each class covariance matrix, enabling
-interpretability that has been previously lacking with the \emph{RDA} classifier
-\citep{Bensmail:1996fw}. Next, applying a reasoning similar to \cite{Ye:2006tq},
-we establish that the matrix operations corresponding to the null space of the
-pooled sample covariance matrix are redundant and can be discarded from the
-\emph{HDRDA} decision rule without loss of classificatory information. Moreover,
-because the \emph{HDRDA} decision rule is invariant to adjustments to the zero
-eigenvalues, the decision rule in the original feature space is equivalent to a
-decision rule in a much lower dimension (i.e., the intrinsic dimension), such
-that the matrix inverses and determinants of relatively small matrices can be
-rapidly computed. Thus, the dimension reduction inherent in the \emph{HDRDA}
-classifier yields a substantial reduction in computation during model
-selection. Also, as a result, we show that a large number of shrinkage methods
-that are special cases of the \emph{HDRDA} classifier have no effect on the zero
+Here, we present a new classification method intended for the $p > N$ case that
+builds on the classic regularized discriminant analysis (\emph{RDA}) proposed by
+\cite{Friedman:1989tm}. Our proposed high-dimensional \emph{RDA} (\emph{HDRDA})
+classifier employs a biased covariance matrix estimator that partially pools the
+individual sample covariance matrices from the \emph{QDA} classifier with the
+pooled sample covariance matrix from the \emph{LDA} classifier. We then shrink
+the resulting covariance-matrix estimator towards a scaled identity matrix to
+ensure positive definiteness. By reparameterizing the \emph{RDA} classifier
+similar to \cite{Hastie:2008dt} and \cite{Halbe:2007ea}, we first show the
+pooling parameter in the \emph{HDRDA} classifier determines the contribution of
+each training observation to the estimation of each class covariance matrix,
+enabling interpretability that has been previously lacking with the \emph{RDA}
+classifier \citep{Bensmail:1996fw}.
+
+Next, we achieve a substantial reduction in dimension to reduce greatly the
+computation during model selection by applying a reasoning similar to
+\cite{Ye:2006tq}. We establish that the matrix operations corresponding to the
+null space of the pooled sample covariance matrix are redundant and can be
+discarded from the \emph{HDRDA} decision rule without loss of classificatory
+information. Moreover, because the \emph{HDRDA} decision rule is invariant to
+adjustments to the zero eigenvalues, the decision rule in the original feature
+space is equivalent to a decision rule in a much lower dimension, such that the
+matrix inverses and determinants of relatively small matrices can be rapidly
+computed. Furthermore, we show that a large number of shrinkage methods that are
+special cases of the \emph{HDRDA} classifier have no effect on the zero
 eigenvalues of the covariance-matrix estimators. Such methods include work from
@@
 We provide an efficient algorithm that expedites the model-selection process for
-the \emph{HDRDA} classifier and then compare its computational runtime with an
-implementation of the \emph{RDA} classifier from the {\tt klaR} R package to
-illustrate that the \emph{HDRDA} classifier greatly reduces computation when $p
-\gg N$. We then study the classification performance of the \emph{HDRDA}
-classifier on three artificial and four real high-dimensional data sets. We
-demonstrate that the \emph{HDRDA} classifier often attains superior
-classification accuracy to several recent classifiers designed for small-sample,
-high-dimensional data from \cite{Tong:2012hw}, \cite{Witten:2011kc},
-\cite{Pang:2009ik}, and \cite{Guo:2007te}. We also include as a benchmark the
-random forest from \cite{Breiman:2001fb} because \cite{FernandezDelgado:2014ul}
-has concluded that the random forest is often superior to other classifiers in
-benchmark studies.
+the \emph{HDRDA} classifier when $p \gg N$. We then study the classification
+performance of the \emph{HDRDA} classifier on four real high-dimensional data
+sets along with a simulation experiment that generalizes that initially
+conducted by \cite{Guo:2007te}. We demonstrate that the \emph{HDRDA} classifier
+often attains superior classification accuracy to several recent classifiers
+designed for small-sample, high-dimensional data from \cite{Tong:2012hw},
+\cite{Witten:2011kc}, \cite{Pang:2009ik}, and \cite{Guo:2007te}. We also include
+as a benchmark the random forest from \cite{Breiman:2001fb} because
+\cite{FernandezDelgado:2014ul} has concluded that the random forest is often
+superior to other classifiers in benchmark studies.
 
@@
 by calculating the number of eigenvalues that exceeds some tolerance value, say,
-$\epsilon = 1 \times 10^{-6}$. After calculating $q$, we extract $\bm U_1$ by
-retaining only the first $q$ columns of $\bm U$.
+$1 \times 10^{-6}$. After calculating $q$, we extract $\bm U_1$ by retaining
+only the first $q$ columns of $\bm U$.
 
@@
 
+% TODO: Combine first two paragraphs -- list R package along with authors
+
 In this section, we compare our proposed classifier with four classifiers
-recently proposed for small-sample, high-dimensional data along with the
-random-forest classifier from \cite{Breiman:2001fb} as a benchmark. Our
+recently proposed for small-sample, high-dimensional data. Our
 competitors include shrunken centroids regularized discriminant analysis from
 \cite{Guo:2007te} and penalized linear discriminant analysis from
-\cite{Witten:2011kc}. We also consider two modifications of diagonal linear
-discriminant analysis from \cite{Tong:2012hw} and \cite{Pang:2009ik}, where the
-former employs an improved mean estimator and the latter utilizes an improved
-variance estimator. Below, we refer to each classifier by the first author's
-surname.\footnote{TODO: Consider adding a heatmap of error rates of a single
-  $(\lambda, \gamma)$ grid.}
-
-We used version 3.2.2 of the open-source statistical software {\tt R} for the
-classification study. We used the implementations of the Witten and Guo
+\cite{Witten:2011kc}.
+
+We used the implementations of the Witten and Guo
 classifiers from the {\tt penalizedLDA} and {\tt rda} packages, respectively.
 Because the {\tt rda} package does not perform the authors' ``Min-Min'' rule
-automatically, we applied this rule within our {\tt R} code. We also utilized
-the implementations of the Pang and Tong classifiers available in the {\tt
-  sparsediscrim} package. We used the implementation of the random-forest
-classifier from the {\tt randomForest} package. For each classifer we explicitly
-set prior probabilities as equal if applicable. All other classifier options
-were set to their default settings.
+automatically, we applied this rule within our {\tt R} code.
+
+We also consider two modifications of diagonal linear discriminant analysis from
+\cite{Tong:2012hw} and \cite{Pang:2009ik}, where the former employs an improved
+mean estimator and the latter utilizes an improved variance estimator. We also
+utilized the implementations of the Pang and Tong classifiers available in the
+{\tt sparsediscrim} package.
+
+We also included the random forest from \cite{Breiman:2001fb} as a benchmark
+based on the findings of \cite{FernandezDelgado:2014ul}, who has concluded that
+the random forest is often superior to other classifiers in benchmark studies.
+We used the implementation of the random-forest classifier from the {\tt
+  randomForest} package with 250 trees and 100 maximum nodes.
+
+For each classifer we explicitly set prior probabilities as equal if
+applicable. All other classifier options were set to their default settings.
+We used version 3.2.2 of the open-source statistical software {\tt R} for the
+classification study.  Below, we refer to each classifier by the first author's
+surname.
 

### 2015-10-25 reports/paper-hdrda.Rnw
Rewrote the beginning of simulation intro

@@
 
-% TODO: Combine first two paragraphs -- list R package along with authors
-
 In this section, we compare our proposed classifier with four classifiers
-recently proposed for small-sample, high-dimensional data. Our
-competitors include shrunken centroids regularized discriminant analysis from
-\cite{Guo:2007te} and penalized linear discriminant analysis from
-\cite{Witten:2011kc}.
-
-We used the implementations of the Witten and Guo
-classifiers from the {\tt penalizedLDA} and {\tt rda} packages, respectively.
-Because the {\tt rda} package does not perform the authors' ``Min-Min'' rule
-automatically, we applied this rule within our {\tt R} code.
-
-We also consider two modifications of diagonal linear discriminant analysis from
-\cite{Tong:2012hw} and \cite{Pang:2009ik}, where the former employs an improved
-mean estimator and the latter utilizes an improved variance estimator. We also
-utilized the implementations of the Pang and Tong classifiers available in the
-{\tt sparsediscrim} package.
-
-We also included the random forest from \cite{Breiman:2001fb} as a benchmark
-based on the findings of \cite{FernandezDelgado:2014ul}, who has concluded that
-the random forest is often superior to other classifiers in benchmark studies.
-We used the implementation of the random-forest classifier from the {\tt
-  randomForest} package with 250 trees and 100 maximum nodes.
-
-For each classifer we explicitly set prior probabilities as equal if
-applicable. All other classifier options were set to their default settings.
-We used version 3.2.2 of the open-source statistical software {\tt R} for the
-classification study.  Below, we refer to each classifier by the first author's
-surname.
-
-For the \emph{HDRDA} classifier in \eqref{eq:hdrda-decomposed}, we examine the
+recently proposed for small-sample, high-dimensional data along with the random
+forest from \cite{Breiman:2001fb} using version 3.2.2 of the open-source
+statistical software {\tt R}. Within our study, we included penalized linear
+discriminant analysis from \cite{Witten:2011kc}, implemented in the {\tt
+  penalizedLDA} package. We also considered shrunken centroids regularized
+discriminant analysis from \cite{Guo:2007te} in the {\tt rda} package. Because
+the {\tt rda} package does not perform the authors' ``Min-Min'' rule
+automatically, we applied this rule within our {\tt R} code. We included two
+modifications of diagonal linear discriminant analysis from \cite{Tong:2012hw}
+and \cite{Pang:2009ik}, where the former employs an improved mean estimator and
+the latter utilizes an improved variance estimator. Both classifiers are
+available in the {\tt sparsediscrim} package. Finally, we incorporated the
+random forest as a benchmark based on the findings of
+\cite{FernandezDelgado:2014ul}, who concluded that the random forest is often
+superior to other classifiers in benchmark studies. We used the implementation
+of the random-forest classifier from the {\tt randomForest} package with 250
+trees and 100 maximum nodes. For each classifer we explicitly set prior
+probabilities as equal, if applicable. All other classifier options were set to
+their default settings. Below, we refer to each classifier by the first
+author's surname.
+
+For the \emph{HDRDA} classifier in \eqref{eq:hdrda-decomposed}, we examined the
 classification performance of two models. For the first \emph{HDRDA} model, we
 set $\alpha_k = 1$, $k = 1, \ldots, K$, so that the covariance-matrix estimator
-\eqref{eq:hdrda-cov} employed resembled \eqref{eq:ridge-estimator}. We estimated
+\eqref{eq:hdrda-cov} resembled \eqref{eq:ridge-estimator}. We estimated
 $\lambda$ from a grid of 21 equidistant candidate values between 0 and 1,
@@
 \citep{Hastie:2008dt}. For the second model, we set $\alpha_k = 1 - \gamma$, $k
-= 1, \ldots, K$, to resemble \citeauthor{Friedman:1989tm}'s parameterization. We
-estimated both $\lambda$ and $\gamma$ from a grid of 21 equidistant candidate
-values between 0 and 1, inclusively.
+= 1, \ldots, K$, to resemble \citeauthor{Friedman:1989tm}'s
+parameterization. We estimated both $\lambda$ and $\gamma$ from a grid of 21
+equidistant candidate values between 0 and 1, inclusively.
 

### 2015-10-26 reports/paper-hdrda.Rnw
Rewrote the design for the simulated data sets.

@@
 N)$ with true, unique membership $y_i \in \{\omega_1, \ldots, \omega_K\}$. We
-assume that $(\bm x_i, y_i)$ is a realization from a mixture distribution $p(\bm
-x) = \sum_{k=1}^K p(\bm x | \omega_k) p(\omega_k)$, where $p(\bm x | \omega_k)$
-is the probability density function of the $k$th class and $p(\omega_k)$ is the
-prior probability of class membership of the $k$th class. We further assume that
-each of $p(\omega_k)$ is equal, $k = 1, \ldots, K$.
+assume that $(\bm x_i, y_i)$ is a realization from a mixture distribution
+$p(\bm x) = \sum_{k=1}^K p(\bm x | \omega_k) p(\omega_k)$, where $p(\bm x |
+\omega_k)$ is the probability density function (PDF) of the $k$th class and
+$p(\omega_k)$ is the prior probability of class membership of the $k$th
+class. We further assume that each of $p(\omega_k)$ is equal, $k = 1, \ldots,
+K$.
 
@@
 
-We compared the \emph{HDRDA} classifer to the five competing classifiers in a
-simulaton study. We simulated $K = 3$ $p$-dimensional multivariate normal
-populations with a common positive-definite covariance matrix having a
-block-diagonal covariance structure, which is intended to resemble
-gene-expression data. That is, within each block of pathways, genes are
-positively correlated and that this correlation decays as a function of the
-distance between any two genes. Thus, we simulated $K = 3$ multivariate normal
-populations with the common positive-definite covariance matrix
-
+In this section we compare the competing classifiers using the simulation
+design from \cite{Guo:2007te}. This design is widely used within the
+high-dimensional classification literature including the studies by
+\cite{Ramey:2013ji} and \cite{Witten:2011kc}. The original design comprised two
+$p$-dimensional multivariate normal populations with the common block-diagonal
+covariance matrix
+\begin{align}
+  \bm\Sigma = \bm\Sigma^{(\rho)} \oplus \bm\Sigma^{(-\rho)} \oplus \ldots \oplus \bm\Sigma^{(\rho)},\label{eq:block-sigma}
+\end{align}
+where the $(i,j)$th entry of the block matrix $\bm\Sigma^{(\rho)} \in
+\mathbb{R}_{25 \times 25}$ is
 \begin{align*}
-  \bm\Sigma = \bm\Sigma^{(\rho)} \oplus \bm\Sigma^{(-\rho)} \oplus \ldots \oplus \bm\Sigma^{(\rho)},
+\bm\Sigma_{ij}^{(\rho)} = \{ \rho^{|i - j|} \}_{1 \le i,j \le 25}.
 \end{align*}
-
-where the $(i,j)$th entry of $\bm\Sigma^{(\rho)} \in \mathbb{R}_{25 \times 25}$ is
-
-\begin{align*}
-\bm\Sigma_{ij}^{(\rho)} = \{ \rho^{|i - j|} \}_{1 \le i,j \le 25},
-\end{align*}
-
-altering positive and negative values of $\rho$ are used from block to block. In
-our simulation study, we considered three different values of $\rho$, 0.1, 0.5,
-and 0.9, to explore the effect of autocorrelation on classification accuracy,
-especially given that the classifiers from \cite{Tong:2012hw} and
-\cite{Pang:2009ik} assume independence among features.
-
-Furthermore, we set the population of mean of class 1 to be a $p$-dimensional
-vector of zeros. Next, we set class 2's population mean to a vector equaling 1/2
-for the first 50 features and 0 after ($\mu_j = 1/2$ if $j \le 50$ and $\mu_j =
-0$ if $j > 50$). Similarly, we set class 3's population mean to a vector equaling -1/2
-for the first 50 features and 0 after ($\mu_j = -1/2$ if $j \le 50$ and $\mu_j =
-0$ if $j > 50$). The three populations differ in their mean vectors in the first
+The block-diagonal covariance structure in \eqref{eq:block-sigma} resembles
+gene-expression data: within each block of pathways, genes are correlated, and
+the correlation decays as a function of the distance between any two genes. The
+autocorrelation $\rho$ alternates between a positive and a negative value from block
+to block.
+
+Although the design is indeed standard, the simulation configuration lacks
+artifacts observed commonly in real data, such as skewness and extreme
+outliers. As a result, we wished to investigate the effect of outliers on the
+high-dimensional classifiers. To do this, we generalized the block-diagonal
+simulation configuration by sampling from a $p$-dimensional multivariate
+contaminated normal distribution.  Denoting the PDF of the $p$-dimensional
+multivariate normal distribution by $N_p(\bm x | \bm \mu, \bm \Sigma)$, we
+write the PDF of the $k$th class as
+\begin{align}
+  p(\bm x | \omega_k) = (1 - \epsilon) N_p(\bm x | \bm \mu_k, \bm \Sigma) + \epsilon N_p(\bm x | \bm \mu_k, \omega \bm \Sigma),\label{eq:contaminated-pdf}
+\end{align}
+where $\epsilon \in [0, 1]$ is the probability that an observation is contaminated (i.e.,
+drawn from a distribution with larger variance) and $\omega > 1$ scales the
+covariance matrix $\bm \Sigma$ to increase the extremity of outliers. For
+$\epsilon = 0$, we have the benchmark block-diagonal simulation design from
+\cite{Guo:2007te}. As $\epsilon$ is increased, the average number of outliers
+is increased. In our simulation, we considered the values of $\epsilon = 0$,
+$0.05$, $0.1$, $0.15$, and $0.2$. We also set $\omega = 100$.
+
+We generated $K=3$ populations from \eqref{eq:contaminated-pdf} using the
+covariance matrix $\bm \Sigma$ given in \eqref{eq:block-sigma}. We set the mean
+vector of class 1 to $\bm \mu_1 = \bm 0_p$. Next, comparable to
+\cite{Guo:2007te}, the first 50 features of $\bm \mu_2$ were 1/2, while the rest
+were set to 0, i.e., $\bm \mu_2 = (\underbrace{1/2, \ldots, 1/2}_{50},
+\underbrace{0, \ldots, 0}_{p - 50})$. For simplicity, we defined $\bm \mu_3 =
+-\bm \mu_2$. The three populations differed in their mean vectors in the first
 50 features corresponding to the first two blocks. No difference in the means
-occurs in the remaining blocks.
-
-The block-diagonal simulation design considered is widely used within the
-high-dimensional classification literature, including the studies by
-\cite{Ramey:2013ji}, \cite{Witten:2011kc}, and \cite{Guo:2007te}. While the
-design is indeed standard, the simulation configuration is based on an ideal
-scenario where no artifacts, often observed in real data, are present. As a
-result, we wished to investigate the effect of outliers on the high-dimensional
-classifiers. To do this, we generalized the block-diagonal simulation
-configuration by sampling from a $p$-dimensional multivariate
-$\epsilon$-contaminated normal distribution, where $\epsilon in [0, 1]$. For
-$\epsilon = 0$, we have the benchmark block-diagonal simulations considered in
-the literature. As $\epsilon$ is increased, the number of outliers is increased
-(on average). We consider the values of $\epsilon = 0, 0.05, 0.1, 0.15,
-0.2$. For the contamination simulation, we define the probability density
-function of the $k$th class to be
-
-\begin{align*}
-  p(\bm x | \omega_k) = (1 - \epsilon) N_p(\bm x | \bm \mu_k, \bm \Sigma) + \epsilon N_p(\bm x | \bm \mu_k, 100 * \bm \Sigma)
-\end{align*}
+occurred in the remaining blocks.
 
@@
 the error-rate estimates for each classifier. We allowed the number of features
-to vary from $p = 100$ to $p = 500$ in increments of 100 to examine the increase
-in dimension while maintaining a small sample size.
-
-Despite the claims of \cite{FernandezDelgado:2014ul} that the random forest
-classifier is typically superior, we observed that the classifiers designed for
-high-dimensional data were indeed superior.
+to vary from $p = 100$ to $p = 500$ in increments of 100 to examine the
+classification accuracy as the feature dimension increased while maintaining a
+small sample size. Furthermore, although \cite{Guo:2007te} consider $\rho =
+0.9$, we considered $\rho = 0.1$, $0.5$, and $0.9$ to explore the effect of
+autocorrelation on classification accuracy, especially given that the Tong,
+Pang, and Guo classifiers assume independence among features.
 
@@
 \subsection{Classification Results}
+
+Despite the claims of \cite{FernandezDelgado:2014ul} that the random forest
+classifier is typically superior, we observed that the classifiers designed for
+high-dimensional data were indeed superior.
 

### 2015-10-26 sim-block-contaminated.r
Autocorrelations in sims now vary

@@ -20,6 +20,7 @@ test_sizes <- rep(1000, K)
 
 iter_product <- itertools2::iproduct(p=feature_dim,
                                      contamination_prob=contamination_probs,
+                                     rho=autocorrelations,
                                      times=num_iterations)
 sim_config <- itertools2::ienum(iter_product)
 num_iterations <- length(feature_dim) * length(contamination_probs) * num_iterations
@@ -29,6 +30,7 @@ flog.logger("sim", INFO, appender=appender.file('sim-block-contaminated.log'))
 results <- mclapply(sim_config, function(sim_i) {
   i <- sim_i$index
   p <- sim_i$value$p
+  rho <- rep(sim_i$value$rho, K)
   contamination_prob <- sim_i$value$contamination_prob
   flog.info("Dimension: %s. Contamination Prob: %s -- Sim: %s of %s",
             p, contamination_prob, i, num_iterations, name="sim")
@@ -46,7 +48,7 @@ results <- mclapply(sim_config, function(sim_i) {
                                       mu=cbind(mu1, mu2, mu3),
                                       block_size=block_size,
                                       num_blocks=num_blocks,
-                                      rho=autocorrelations,
+                                      rho=rho,
                                       uncontaminated_var=uncontaminated_var,
                                       contaminated_var=contaminated_var,
                                       contamination_prob=contamination_prob)
@@ -200,6 +202,7 @@ results <- mclapply(sim_config, function(sim_i) {
 
   list(error_rates = error_rates,
        p = p,
+       rho=rho,
        contamination_prob = contamination_prob,
        hdrda_ridge = hdrda_ridge,
        hdrda_convex = hdrda_convex

### 2015-10-26 sim-block-contaminated.r
Increased test size from 1K to 10K in sims

@@ -16,7 +16,7 @@ uncontaminated_var <- rep(1, K)
 contaminated_var <- rep(100, K)
 
 # The number of observations to generate from each population as a test data set.
-test_sizes <- rep(1000, K)
+test_sizes <- rep(10000, K)
 
 iter_product <- itertools2::iproduct(p=feature_dim,
                                      contamination_prob=contamination_probs,

### 2015-10-26 sim-block-contaminated.r
Vary means to observe separation vs accuracy

@@ -8,9 +8,10 @@ num_iterations <- 250
 
 K <- 3
 sample_sizes <- rep(25, K)
+means <- c(1/2, 1)
 autocorrelations <- c(0.1, 0.5, 0.9)
 feature_dim <- seq(100, 500, by=100)
-block_size <- 25
+block_size <- 100
 contamination_probs <- seq(0, 0.2, by=0.05)
 uncontaminated_var <- rep(1, K)
 contaminated_var <- rep(100, K)
@@ -20,27 +21,28 @@ test_sizes <- rep(10000, K)
 
 iter_product <- itertools2::iproduct(p=feature_dim,
                                      contamination_prob=contamination_probs,
+                                     mu=means,
                                      rho=autocorrelations,
                                      times=num_iterations)
 sim_config <- itertools2::ienum(iter_product)
-num_iterations <- length(feature_dim) * length(contamination_probs) * num_iterations
+num_iterations <- length(feature_dim) * length(contamination_probs) * length(means) * length(autocorrelations) * num_iterations
 
 flog.logger("sim", INFO, appender=appender.file('sim-block-contaminated.log'))
 
 results <- mclapply(sim_config, function(sim_i) {
   i <- sim_i$index
   p <- sim_i$value$p
+  mu <- sim_i$value$mu
   rho <- rep(sim_i$value$rho, K)
   contamination_prob <- sim_i$value$contamination_prob
   flog.info("Dimension: %s. Contamination Prob: %s -- Sim: %s of %s",
             p, contamination_prob, i, num_iterations, name="sim")
   set.seed(i)
 
-  # Difference in means in the first 50 variables (2 * block_size)
-  num_vars_signal <- 2 * block_size
+  # Difference in means in the first 100 variables (first block)
   mu1 <- rep(0, p)
-  mu2 <- c(rep(1/2, num_vars_signal), rep(0, p - num_vars_signal))
-  mu3 <- c(rep(-1/2, num_vars_signal), rep(0, p - num_vars_signal))
+  mu2 <- c(rep(mu, block_size), rep(0, p - block_size))
+  mu3 <- -mu2
   num_blocks <- p / block_size
 
   # Generates training data
@@ -202,6 +204,7 @@ results <- mclapply(sim_config, function(sim_i) {
 
   list(error_rates = error_rates,
        p = p,
+       mu=mu,
        rho=rho,
        contamination_prob = contamination_prob,
        hdrda_ridge = hdrda_ridge,

### 2015-10-26 sim-block-contaminated.r
Rm'd kNN and SVM from sims to match data sims

@@ -4,7 +4,7 @@ load.project()
 set.seed(42)
 num_cores <- 8
 
-num_iterations <- 250
+num_iterations <- 500
 
 K <- 3
 sample_sizes <- rep(25, K)
@@ -82,21 +82,6 @@ results <- mclapply(sim_config, function(sim_i) {
   num_classes <- nlevels(train_y)
   prior_probs <- rep(1, num_classes) / num_classes
 
-  # kNN
-  knn_errors <- try({
-      tune.knn_out <- tune.knn(x = train_x,
-                               y = train_y,
-                               k = 1:5,
-                               tunecontrol = tune.control(sampling = "cross", cross = 10))
-      knn_out <- knn(train = train_x,
-                     cl = train_y,
-                     test = test_x,
-                     k = tune.knn_out$best.model$k)
-      mean(knn_out != test_y)
-  })
-
-  flog.info("kNN Error Rate: %s -- Sim: %s of %s", knn_errors, i, num_iterations, name="sim")
-
   # Random Forest
   rf_errors <- try({
     rf_out <- randomForest(x=train_x,
@@ -108,19 +93,6 @@ results <- mclapply(sim_config, function(sim_i) {
   })
 
   flog.info("Random Forest Error Rate: %s -- Sim: %s of %s", rf_errors, i, num_iterations, name="sim")
-
-  # SVM with Radial Basis Functions
-  ksvm_radial_errors <- try({
-      ksvm_out <- ksvm(x = train_x,
-                       y = train_y,
-                       kernel = "rbfdot",
-                       kpar = "automatic",
-                       cross = 10)
-      ksvm_radial <- predict(ksvm_out, test_x)
-      mean(ksvm_radial != test_y)
-  })
-
-  flog.info("SVM Error Rate: %s -- Sim: %s of %s", ksvm_radial_errors, i, num_iterations, name="sim")
 
   # Witten and Tibshirani (2011) - JRSS B
   Witten_errors <- try({
@@ -194,10 +166,8 @@ results <- mclapply(sim_config, function(sim_i) {
     Guo = Guo_errors,
     HDRDA_Ridge = hdrda_ridge_errors,
     HDRDA_Convex = hdrda_convex_errors,
-    kNN = knn_errors,
     Pang = Pang_errors,
     Random_Forest = rf_errors,
-    SVM_Radial = ksvm_radial_errors,
     Tong = Tong_errors,
     Witten = Witten_errors
   )

### 2015-10-26 sim-block-contaminated.r
Updated logging in sims

@@ -4,7 +4,7 @@ load.project()
 set.seed(42)
 num_cores <- 8
 
-num_iterations <- 500
+num_iterations <- 250
 
 K <- 3
 sample_sizes <- rep(25, K)
@@ -35,8 +35,8 @@ results <- mclapply(sim_config, function(sim_i) {
   mu <- sim_i$value$mu
   rho <- rep(sim_i$value$rho, K)
   contamination_prob <- sim_i$value$contamination_prob
-  flog.info("Dimension: %s. Contamination Prob: %s -- Sim: %s of %s",
-            p, contamination_prob, i, num_iterations, name="sim")
+  flog.info("Dimension: %s. Mu: %s. Rho: %s. Contamination Prob: %s -- Sim: %s of %s",
+            p, mu, rho[1], contamination_prob, i, num_iterations, name="sim")
   set.seed(i)
 
   # Difference in means in the first 100 variables (first block)

### 2015-10-26 sim-block-contaminated.r
Increased number of cores from 8 to 16
Moving from EC2 c4.2xlarge to c4.4xlarge

@@ -2,7 +2,7 @@ library('ProjectTemplate')
 load.project()
 
 set.seed(42)
-num_cores <- 8
+num_cores <- 16
 
 num_iterations <- 250
 

### 2015-12-14 sim-block-contaminated.r
Increased separation to examine error rates.

@@ -8,7 +8,7 @@ num_iterations <- 500
 
 K <- 3
 sample_sizes <- rep(25, K)
-means <- 1/2
+means <- 1
 autocorrelations <- 0.9
 feature_dim <- seq(100, 500, by=100)
 block_size <- 100

### 2015-12-14 reports/paper-hdrda.Rnw
Added Phil's writeup of first sims

@@
 Pang, and Guo classifiers assume independence among features.
+
+\subsubsection{Results}
+
+As expected, regardless of the values of $p$, $\rho$, or $\bm \mu$, as the
+contaminated probability increases, the average error rate increases.  Also, as
+can be expected, the average error rates were typically higher for $\bm \mu=0.5$
+than for $\bm \mu=1$. The difference in the average error rates for the methods
+of Pang, Tong, and Witten all yielded increasingly higher average error rates
+than HDRDA ridge, HDRDA complex, Guo, or random forests as the contaminated
+probability increased. The correlation coefficient appears to have had little,
+if any, effect on the average error rates for any of the considered classifiers.
+
+Regardless of the value of $rho$, $p$, or $\bm \mu$, random forests consistently
+yielded the highest average error rate, and the average error rate for random
+forests was much more pronounced when $\bm \mu=0.5$. However, as the
+contaminated probability increased, random forests became increasingly
+competitive, and, in some cases, e.g, $\bm \mu=0.5$, $p=100$, and contaminated
+probability greater than 0.05 yielded the lowest average error rate.
+
+Pang and Tong's methods yielded practically the same and consistently the worst
+average error rates when the contaminated probability was greater than 0,
+regardless of the value of $\bm \mu$, $p$, or $\rho$. Witten's method performed
+only slightly better as the contaminated probability increased, however, when
+$\bm \mu = 0.5$, Witten's method became increasingly competitive with Pang and
+Tong for yielding the highest average error rate as $p$ increased.
+
+Guo's method always outperformed Pang, Witten, and Tong, but never outperformed
+the HDRDA methods. For the considered parameter configurations, Guo's method
+performed similarly to random forests when the contaminated probability was
+greater than 0 and yielded similar results to HDRDA ridge, HDRDA convex when the
+contaminated probability was equal to 0. However, as the contaminated
+probability increased, Guo's method was often outperformed by HDRDA ridge, HDRDA
+convex, and random forests.
+
+HDRDA ridge and HDRDA complex yielded the lowest average error rates, regardless
+of the value of $\bm \mu$, $p$, or $\rho$, with HDRDA ridge typically
+outperforming HDRDA complex when the contaminated probability was greater than
+0. HDRDA ridge only seemed to have competition with random forests when the
+contaminated probability was 0.20. Moreover, neither HDRDA ridge nor HDRDA
+complex seem to be affected by $p$.
 

### 2016-01-01 reports/paper-hdrda.Rnw
In paper updated sim description to match code

@@
 
-In this section we compare the competing classifiers using the simulation
-design from \cite{Guo:2007te}. This design is widely used within the
-high-dimensional classification literature including the studies by
-\cite{Ramey:2013ji} and \cite{Witten:2011kc}. The original design comprised two
-$p$-dimensional multivariate normal populations with the common block-diagonal
-covariance matrix
-\begin{align}
-  \bm\Sigma = \bm\Sigma^{(\rho)} \oplus \bm\Sigma^{(-\rho)} \oplus \ldots \oplus \bm\Sigma^{(\rho)},\label{eq:block-sigma}
-\end{align}
-where the $(i,j)$th entry of the block matrix $\bm\Sigma^{(\rho)} \in
-\mathbb{R}_{25 \times 25}$ is
+In this section we compare the competing classifiers using the simulation design
+from \cite{Guo:2007te}. This design is widely used within the high-dimensional
+classification literature including the studies by \cite{Ramey:2013ji} and
+\cite{Witten:2011kc}. First, we consider the block-diagonal covariance
+matrix from from \cite{Guo:2007te}
+\begin{align}
+  \bm\Sigma_k = \bm\Sigma^{(\rho_k)} \oplus \bm\Sigma^{(-\rho_k)} \oplus \ldots \oplus \bm\Sigma^{(\rho_k)},\label{eq:block-sigma}
+\end{align}
+where the $(i,j)$th entry of the block matrix $\bm\Sigma^{(\rho_k)} \in
+\mathbb{R}_{100 \times 100}$ is
 \begin{align*}
-\bm\Sigma_{ij}^{(\rho)} = \{ \rho^{|i - j|} \}_{1 \le i,j \le 25}.
+\bm\Sigma_{ij}^{(\rho_k)} = \{ \rho_k^{|i - j|} \}_{1 \le i,j \le 100}.
 \end{align*}
@@
 the correlation decays as a function of the distance between any two genes. The
-autocorrelation $\rho$ alternates between a positive and a negative value from block
-to block.
+autocorrelation $\rho_k$ alternates between a positive and a negative value from block
+to block. The original design from \cite{Guo:2007te} comprised two $p$-dimensional
+multivariate normal populations with a common block-diagonal covariance
+matrix where $\bm\Sigma_1 = \bm\Sigma_2$.
 
@@
 contaminated normal distribution.  Denoting the PDF of the $p$-dimensional
-multivariate normal distribution by $N_p(\bm x | \bm \mu, \bm \Sigma)$, we
-write the PDF of the $k$th class as
-\begin{align}
-  p(\bm x | \omega_k) = (1 - \epsilon) N_p(\bm x | \bm \mu_k, \bm \Sigma) + \epsilon N_p(\bm x | \bm \mu_k, \omega \bm \Sigma),\label{eq:contaminated-pdf}
-\end{align}
-where $\epsilon \in [0, 1]$ is the probability that an observation is contaminated (i.e.,
-drawn from a distribution with larger variance) and $\omega > 1$ scales the
-covariance matrix $\bm \Sigma$ to increase the extremity of outliers. For
-$\epsilon = 0$, we have the benchmark block-diagonal simulation design from
-\cite{Guo:2007te}. As $\epsilon$ is increased, the average number of outliers
-is increased. In our simulation, we considered the values of $\epsilon = 0$,
-$0.05$, $0.1$, $0.15$, and $0.2$. We also set $\omega = 100$.
-
-We generated $K=3$ populations from \eqref{eq:contaminated-pdf} using the
-covariance matrix $\bm \Sigma$ given in \eqref{eq:block-sigma}. We set the mean
-vector of class 1 to $\bm \mu_1 = \bm 0_p$. Next, comparable to
-\cite{Guo:2007te}, the first 50 features of $\bm \mu_2$ were 1/2, while the rest
-were set to 0, i.e., $\bm \mu_2 = (\underbrace{1/2, \ldots, 1/2}_{50},
-\underbrace{0, \ldots, 0}_{p - 50})$. For simplicity, we defined $\bm \mu_3 =
--\bm \mu_2$. The three populations differed in their mean vectors in the first
-50 features corresponding to the first two blocks. No difference in the means
-occurred in the remaining blocks.
+multivariate normal distribution by $N_p(\bm x | \bm \mu, \bm \Sigma)$, we write
+the PDF of the $k$th class as
+\begin{align}
+  p(\bm x | \omega_k) = (1 - \epsilon) N_p(\bm x | \bm \mu_k, \bm \Sigma_k) + \epsilon N_p(\bm x | \bm \mu_k, \omega \bm \Sigma_k),\label{eq:contaminated-pdf}
+\end{align}
+where $\epsilon \in [0, 1]$ is the probability that an observation is
+contaminated (i.e., drawn from a distribution with larger variance) and $\omega
+> 1$ scales the covariance matrix $\bm \Sigma_k$ to increase the extremity of
+outliers. For $\epsilon = 0$, we have the benchmark block-diagonal simulation
+design from \cite{Guo:2007te}. As $\epsilon$ is increased, the average number of
+outliers is increased. In our simulation, we considered the values of $\epsilon
+= 0$, $0.05$, $\ldots$, $0.50$. We also set $\omega = 100$.
+
+We generated $K=3$ populations from \eqref{eq:contaminated-pdf} with $\bm
+\Sigma_k$ given in \eqref{eq:block-sigma}. We set the mean vector of class 1 to
+$\bm \mu_1 = \bm 0_p$. Next, comparable to \cite{Guo:2007te}, the first 100
+features of $\bm \mu_2$ were 1/2, while the rest were set to 0, i.e., $\bm \mu_2
+= (\underbrace{1/2, \ldots, 1/2}_{100}, \underbrace{0, \ldots, 0}_{p -
+  100})$. For simplicity, we defined $\bm \mu_3 = -\bm \mu_2$. The three
+populations differed in their mean vectors in the first 100 features
+corresponding to the first block. No difference in the means occurred in the
+remaining blocks.
 
 From each of the $K=3$ populations, we sampled 25 training observations ($n_k =
-25$ for all $k$) and 1000 test observations. After training each classifier on
+25$ for all $k$) and 10,000 test observations. After training each classifier on
 the training data, we classified the test data sets and computed the proportion
 of mislabeled test observations to estimate the classification error rate for
-each classifier. Repeating this process 250 times, we computed the average of
+each classifier. Repeating this process 500 times, we computed the average of
 the error-rate estimates for each classifier. We allowed the number of features
@@
 classification accuracy as the feature dimension increased while maintaining a
-small sample size. Furthermore, although \cite{Guo:2007te} consider $\rho =
-0.9$, we considered $\rho = 0.1$, $0.5$, and $0.9$ to explore the effect of
-autocorrelation on classification accuracy, especially given that the Tong,
-Pang, and Guo classifiers assume independence among features.
+small sample size. Furthermore, although \cite{Guo:2007te} consider $\rho_k =
+0.9$ for all $k$, we considered $\rho_1 = 0.1$, $\rho_2 = 0.5$, and $\rho_3 = 0.9$
+to explore the more realistic assumption of unequal covariance matrices.
 

### 2016-01-01 reports/paper-hdrda.Rnw
Added plot of sim error rates to paper

@@
 
-opts_chunk$set(fig.align = 'default', dev = 'png', message = FALSE,
-               warning = FALSE, cache = TRUE, echo = FALSE, autodep = TRUE,
-               fig.width = 12, fig.height = 12, out.width = "\\linewidth")
+opts_chunk$set(fig.align='default', dev='png', message=FALSE, warning=FALSE,
+               cache=TRUE, echo=FALSE, autodep=TRUE, fig.width=12,
+               fig.height=12, out.width="\\linewidth")
 @
@@
 load.project()
+
+# Contaminated Normal Sims
+results_contaminated <- lapply(results_block_contaminated[[1]], function(r) {
+  with(r, c(error_rates, p=p, mu=mu, contamination_prob=contamination_prob))
+})
+results_contaminated <- do.call(rbind.data.frame, results_contaminated)
+results_contaminated <- tbl_df(results_contaminated)
+
+results_contaminated <- results_contaminated %>%
+  group_by(p, contamination_prob, mu) %>%
+  summarise_each(funs(mean)) %>%
+  arrange(p, contamination_prob, mu)
+
+results_contaminated <- results_contaminated %>%
+  gather(Classifier, Error, -p, -contamination_prob, -mu) %>%
+  mutate(p=factor(p)) %>%
+  filter(mu == 0.5)
 
@@
 \begin{figure}
-<<hdrda_contours, fig.width=8, fig.height=8, out.width='.33\\linewidth', fig.show='hold'>>=
+<<hdrda_contours, fig.width=8, fig.height=8, out.width='.33\\linewidth', dpi=200, fig.show='hold'>>=
 rda_contours(lambda = 0)
@@
 
-\subsubsection{Results}
+\subsubsection{Simulation Results}
 
@@
 
+\begin{figure}
+<<sim_results, fig.width=8, fig.height=8, dpi=200>>=
+p <- ggplot(results_contaminated, aes(x=contamination_prob, y=Error))
+p <- p + geom_line(aes(color=Classifier, linetype=Classifier), size=0.65)
+p <- p + facet_wrap(~ p, labeller = "label_both")
+p <- p + xlab("Contamination Probability") + ylab("Error Rate")
+p + theme_bw()
+@
+\caption{Averages of the 1000 error-rate estimates of the competing classifiers
+  for each data set as a function of $d$. Approximate standard errors are no
+  greater than \Sexpr{max_se}.}
+\label{fig:sim-results}
+\end{figure}
 Regardless of the value of $rho$, $p$, or $\bm \mu$, random forests consistently

### 2016-01-02 reports/paper-hdrda.Rnw
Summarized contamination sim results. Finally.

@@
 # http://stackoverflow.com/questions/11062497/how-to-avoid-using-round-in-every-sexpr
-options(digits = 3)
+options(digits=3)
 
@@
   filter(mu == 0.5)
+
+# Remove underscore from classifier names
+levels(results_contaminated$Classifier) <- gsub("_", " ",
+                                                levels(results_contaminated$Classifier))
+
+# Maximum Standard Error for Contaminated Sims
+max_sim_se <- with(results_contaminated, max(sqrt(Error * (1 - Error) / 500)))
+max_sim_se <- round(max_sim_se, digits=3)
 
@@
 In this section, we compare our proposed classifier with four classifiers
-recently proposed for small-sample, high-dimensional data along with the random
-forest from \cite{Breiman:2001fb} using version 3.2.2 of the open-source
-statistical software {\tt R}. Within our study, we included penalized linear
-discriminant analysis from \cite{Witten:2011kc}, implemented in the {\tt
-  penalizedLDA} package. We also considered shrunken centroids regularized
-discriminant analysis from \cite{Guo:2007te} in the {\tt rda} package. Because
-the {\tt rda} package does not perform the authors' ``Min-Min'' rule
-automatically, we applied this rule within our {\tt R} code. We included two
-modifications of diagonal linear discriminant analysis from \cite{Tong:2012hw}
-and \cite{Pang:2009ik}, where the former employs an improved mean estimator and
-the latter utilizes an improved variance estimator. Both classifiers are
-available in the {\tt sparsediscrim} package. Finally, we incorporated the
-random forest as a benchmark based on the findings of
+recently proposed for small-sample, high-dimensional data along with the
+random-forest classifier from \cite{Breiman:2001fb} using version 3.2.2 of the
+open-source statistical software {\tt R}. Within our study, we included
+penalized linear discriminant analysis from \cite{Witten:2011kc}, implemented in
+the {\tt penalizedLDA} package. We also considered shrunken centroids
+regularized discriminant analysis from \cite{Guo:2007te} in the {\tt rda}
+package. Because the {\tt rda} package does not perform the authors' ``Min-Min''
+rule automatically, we applied this rule within our {\tt R} code. We included
+two modifications of diagonal linear discriminant analysis from
+\cite{Tong:2012hw} and \cite{Pang:2009ik}, where the former employs an improved
+mean estimator and the latter utilizes an improved variance estimator. Both
+classifiers are available in the {\tt sparsediscrim} package. Finally, we
+incorporated the random forest as a benchmark based on the findings of
 \cite{FernandezDelgado:2014ul}, who concluded that the random forest is often
@@
 probabilities as equal, if applicable. All other classifier options were set to
-their default settings. Below, we refer to each classifier by the first
-author's surname.
+their default settings. Below, we refer to each classifier by the first author's
+surname. All simulations were conducted on an Amazon Elastic Compute Cloud (EC2)
+instance of type `c4.4xlarge`.
 
@@
 \citep{Hastie:2008dt}. For the second model, we set $\alpha_k = 1 - \gamma$, $k
-= 1, \ldots, K$, to resemble \citeauthor{Friedman:1989tm}'s
-parameterization. We estimated both $\lambda$ and $\gamma$ from a grid of 21
-equidistant candidate values between 0 and 1, inclusively.
+= 1, \ldots, K$, to resemble \citeauthor{Friedman:1989tm}'s parameterization. We
+estimated both $\lambda$ and $\gamma$ from a grid of 21 equidistant candidate
+values between 0 and 1, inclusively.
 
@@
 classification literature including the studies by \cite{Ramey:2013ji} and
-\cite{Witten:2011kc}. First, we consider the block-diagonal covariance
-matrix from from \cite{Guo:2007te}
+\cite{Witten:2011kc}. First, we consider the block-diagonal covariance matrix
+from from \cite{Guo:2007te}
 \begin{align}
@@
 the correlation decays as a function of the distance between any two genes. The
-autocorrelation $\rho_k$ alternates between a positive and a negative value from block
-to block. The original design from \cite{Guo:2007te} comprised two $p$-dimensional
-multivariate normal populations with a common block-diagonal covariance
-matrix where $\bm\Sigma_1 = \bm\Sigma_2$.
+autocorrelation $\rho_k$ alternates between a positive and a negative value from
+block to block. The original design from \cite{Guo:2007te} comprised two
+$p$-dimensional multivariate normal populations with a common block-diagonal
+covariance matrix where $\bm\Sigma_1 = \bm\Sigma_2$.
 
@@
 classification accuracy as the feature dimension increased while maintaining a
-small sample size. Furthermore, although \cite{Guo:2007te} consider $\rho_k =
-0.9$ for all $k$, we considered $\rho_1 = 0.1$, $\rho_2 = 0.5$, and $\rho_3 = 0.9$
-to explore the more realistic assumption of unequal covariance matrices.
+small sample size. \cite{Guo:2007te} originally considered $\rho_k = 0.9$ for
+all $k$. To explore the more realistic assumption of unequal covariance
+matrices, we instead put $\rho_1 = 0.1$, $\rho_2 = 0.5$, and $\rho_3 = 0.9$
 
@@
 
-As expected, regardless of the values of $p$, $\rho$, or $\bm \mu$, as the
-contaminated probability increases, the average error rate increases.  Also, as
-can be expected, the average error rates were typically higher for $\bm \mu=0.5$
-than for $\bm \mu=1$. The difference in the average error rates for the methods
-of Pang, Tong, and Witten all yielded increasingly higher average error rates
-than HDRDA ridge, HDRDA complex, Guo, or random forests as the contaminated
-probability increased. The correlation coefficient appears to have had little,
-if any, effect on the average error rates for any of the considered classifiers.
+In Figure \ref{fig:sim-results}, we observe each classifier's average
+classification error rates for the values of $\epsilon$ and $p$. Unsurprisingly,
+the average error rate increased for each classifier as the contamination
+probability $\epsilon$ increased regardless of the value of $p$. Sensitivity to
+the presence of outliers was most apparent for the for the Pang, Tong, and
+Witten classifiers. For smaller dimensions, the random-forest and HDRDA
+classifiers tended to outperform the remaining classifiers with the random
+forest performing best. As the feature dimension increased with $p \ge 300$,
+both HDRDA classifiers outperformed all other classifiers, suggesting that their
+dimension reduction inherent better captured the classificatory information in
+the small samples even in the presence of outliers.
 
@@
 p <- p + facet_wrap(~ p, labeller = "label_both")
-p <- p + xlab("Contamination Probability") + ylab("Error Rate")
+p <- p + xlab("Contamination Probability") + ylab("Expected Error Rate")
 p + theme_bw()
 @
-\caption{Averages of the 1000 error-rate estimates of the competing classifiers
-  for each data set as a function of $d$. Approximate standard errors are no
-  greater than \Sexpr{max_se}.}
+\caption{Average classification error rates as a function of the contamination
+  probability $\epsilon$. Approximate standard errors were no greater than
+  \Sexpr{max_sim_se}.}
 \label{fig:sim-results}
 \end{figure}
-Regardless of the value of $rho$, $p$, or $\bm \mu$, random forests consistently
-yielded the highest average error rate, and the average error rate for random
-forests was much more pronounced when $\bm \mu=0.5$. However, as the
-contaminated probability increased, random forests became increasingly
-competitive, and, in some cases, e.g, $\bm \mu=0.5$, $p=100$, and contaminated
-probability greater than 0.05 yielded the lowest average error rate.
-
-Pang and Tong's methods yielded practically the same and consistently the worst
-average error rates when the contaminated probability was greater than 0,
-regardless of the value of $\bm \mu$, $p$, or $\rho$. Witten's method performed
-only slightly better as the contaminated probability increased, however, when
-$\bm \mu = 0.5$, Witten's method became increasingly competitive with Pang and
-Tong for yielding the highest average error rate as $p$ increased.
-
-Guo's method always outperformed Pang, Witten, and Tong, but never outperformed
-the HDRDA methods. For the considered parameter configurations, Guo's method
-performed similarly to random forests when the contaminated probability was
-greater than 0 and yielded similar results to HDRDA ridge, HDRDA convex when the
-contaminated probability was equal to 0. However, as the contaminated
-probability increased, Guo's method was often outperformed by HDRDA ridge, HDRDA
-convex, and random forests.
-
-HDRDA ridge and HDRDA complex yielded the lowest average error rates, regardless
-of the value of $\bm \mu$, $p$, or $\rho$, with HDRDA ridge typically
-outperforming HDRDA complex when the contaminated probability was greater than
-0. HDRDA ridge only seemed to have competition with random forests when the
-contaminated probability was 0.20. Moreover, neither HDRDA ridge nor HDRDA
-complex seem to be affected by $p$.
+
+The Pang, Tong, and Witten methods yielded practically the same and consistently
+the worst error rates when outliers were present with $\epsilon > 0$, suggesting
+that these classifiers were sensitive to outliers. Notice, for example, that
+when $p=400$, the error rates of the Pang, Tong, and Witten classifiers
+increased dramatically from approximately 19\% when no outliers were present to
+approximately 43\% when $\epsilon=0.05$. The sharp increase in average error
+rates for these three classifiers continued as $\epsilon$ increased
+further. Guo's method always outperformed Pang, Witten, and Tong, but once
+outliers were present, the Guo classifier's average error rate was competitive
+with neither the HDRDA classifiers nor the random forest.
+
+\begin{figure}
+<<sim_results_by_p, fig.width=8, fig.height=8, dpi=200>>=
+p2 <- results_contaminated %>%
+  mutate(p=as.numeric(as.character(p)),
+         contamination_prob=round(contamination_prob, digits=2)) %>%
+    filter(contamination_prob %in% c(0.0, 0.1, 0.2, 0.3, 0.4, 0.5)) %>%
+    rename(`Contamination Probability`=contamination_prob) %>%
+      ggplot(., aes(x=p, y=Error))
+p2 <- p2 + geom_line(aes(color=Classifier, linetype=Classifier), size=0.65)
+p2 <- p2 + facet_wrap(~ `Contamination Probability`, labeller="label_both")
+p2 <- p2 + xlab("Number of Features") + ylab("Expected Error Rate")
+p2 + theme_bw()
+@
+\caption{Average classification error rates as a function of the number of
+  features $p$. Approximate standard errors were no greater than
+  \Sexpr{max_sim_se}.}
+\label{fig:sim-results-by-p}
+\end{figure}
+
+In Figure \ref{fig:sim-results-by-p} we again examine the simulation results as
+a function of the number of features $p$ for a subset of the values of
+$\epsilon$. This set of plots allows us to investigate the effect of feature
+dimensionality on classification performance. When no outliers were present
+(i.e., $\epsilon=0$), the random-forest classifier was outperformed by all other
+classifiers. Furthermore, the HDRDA classifiers were superior in terms of
+average error rate in this setting. As $p$ increased, an elevation in average
+error rate was expected for all classifiers, but the increase was not observed
+to be substantial.
+
+For $\epsilon > 0$, we observed a different behavior in classification
+performance. First, the Pang, Tong, and Witten methods along with the random
+forest increased in average error rate as $p$ increased. Contrarily, the
+performance of the HDRDA and Guo classifiers was hardly affected by $p$.  Also,
+as discussed above, the HDRDA classifiers were superior to all other classifiers
+for larger values of $p$ with only the random forest outperforming them for
+smaller feature dimensions.
 
@@
 \subsection{Classification Results}
-
-Despite the claims of \cite{FernandezDelgado:2014ul} that the random forest
-classifier is typically superior, we observed that the classifiers designed for
-high-dimensional data were indeed superior.
 
@@
 
+Despite the claims of \cite{FernandezDelgado:2014ul} that the random forest
+classifier is typically superior, we observed that the classifiers designed for
+high-dimensional data were indeed superior.
+
 \bibliographystyle{plainnat}

### 2016-01-02 reports/paper-hdrda.Rnw
Updated data classification summary and results

@@
 load.project()
-
+@
+
+<<munge_contaminated_sims>>=
 # Contaminated Normal Sims
@@
   summarise_each(funs(mean)) %>%
-  arrange(p, contamination_prob, mu)
+  arrange(p, contamination_prob, mu) %>%
+  ungroup
 
@@
 max_sim_se <- round(max_sim_se, digits=3)
-
-max_se <- 1234567890
+@
+
+<<munge_microarray_results>>=
+# Microarray Classification Results
+results_microarray <- lapply(results[[1]], function(r) {
+  with(r, c(error_rates, `data_set`=data_set))
+})
+results_microarray <- do.call(rbind.data.frame, results_microarray)
+results_microarray <- tbl_df(results_microarray)
+
+avg_error_microarray <- results_microarray %>%
+    group_by(data_set) %>%
+    summarise_each(funs(mean)) %>%
+    gather(Classifier, Avg_Error, -data_set)
+
+sd_error_microarray <- results_microarray %>%
+  group_by(data_set) %>%
+  summarise_each(funs(sd)) %>%
+  gather(Classifier, SD_Error, -data_set)
+
+# Microarray classification results are presented in a table
+# Rows = classifier, columns = data set
+# Cells = Error (SD) rounded to 3 digits
+results_microarray <- inner_join(avg_error_microarray, sd_error_microarray,
+                                 by=c('data_set', 'Classifier')) %>%
+  mutate(Table_Cells=sprintf('%.3f (%.3f)', Avg_Error, SD_Error),
+         data_set=stringr::str_to_title(data_set),
+         Classifier=gsub("_", " ", Classifier)) %>%
+  select(-Avg_Error, -SD_Error) %>%
+  spread(data_set, Table_Cells)
 @
@@
 the presence of outliers was most apparent for the for the Pang, Tong, and
-Witten classifiers. For smaller dimensions, the random-forest and HDRDA
+Witten classifiers. For smaller dimensions, the random-forest and \emph{HDRDA}
 classifiers tended to outperform the remaining classifiers with the random
 forest performing best. As the feature dimension increased with $p \ge 300$,
-both HDRDA classifiers outperformed all other classifiers, suggesting that their
-dimension reduction inherent better captured the classificatory information in
-the small samples even in the presence of outliers.
+both \emph{HDRDA} classifiers outperformed all other classifiers, suggesting
+that their dimension reduction inherent better captured the classificatory
+information in the small samples even in the presence of outliers.
 
@@
 p <- p + facet_wrap(~ p, labeller = "label_both")
-p <- p + xlab("Contamination Probability") + ylab("Expected Error Rate")
+p <- p + xlab("Contamination Probability") + ylab("Average Error Rate")
 p + theme_bw()
@@
 outliers were present, the Guo classifier's average error rate was competitive
-with neither the HDRDA classifiers nor the random forest.
+with neither the \emph{HDRDA} classifiers nor the random forest.
 
@@
 p2 <- p2 + facet_wrap(~ `Contamination Probability`, labeller="label_both")
-p2 <- p2 + xlab("Number of Features") + ylab("Expected Error Rate")
+p2 <- p2 + xlab("Number of Features") + ylab("Average Error Rate")
 p2 + theme_bw()
@@
 (i.e., $\epsilon=0$), the random-forest classifier was outperformed by all other
-classifiers. Furthermore, the HDRDA classifiers were superior in terms of
+classifiers. Furthermore, the \emph{HDRDA} classifiers were superior in terms of
 average error rate in this setting. As $p$ increased, an elevation in average
@@
 forest increased in average error rate as $p$ increased. Contrarily, the
-performance of the HDRDA and Guo classifiers was hardly affected by $p$.  Also,
-as discussed above, the HDRDA classifiers were superior to all other classifiers
-for larger values of $p$ with only the random forest outperforming them for
-smaller feature dimensions.
+performance of the \emph{HDRDA} and Guo classifiers was hardly affected by $p$.
+Also, as discussed above, the \emph{HDRDA} classifiers were superior to all
+other classifiers for larger values of $p$ with only the random forest
+outperforming them for smaller feature dimensions.
 
@@
 
-For the microarray data sets, we computed error-rate estimates for the competing
-classifiers after reducing the dimension of the data sets with feature selection
-to $d = 100, 200, \ldots, 1000$, similar to the values of $d$ considered by
-\cite{Xu:2009fl}. To compare misclassification rates between different values of
-$d$, we applied each value of $d$ considered to the same random allocation of
-training and test data. Hence, for a given random partition, the set of features
-selected for a smaller value of $d$ is a proper subset of the features selected
-for a larger value of $d$. We report the average of the error-rate estimates for
-each value of $d$ in Figure \ref{fig:error-rates}. Approximate standard errors
-were no greater than \Sexpr{max_se}.
-
-The \emph{HDRDA} classifier was superior in classification performance for the
-majority of the simulations. For the Burczysnki, Shipp, and Singh data sets, our
-classifier outperformed its competitors for all values of $d$ examined. For the
-Nakayama data set, the \emph{HDRDA} classifier \eqref{eq:hdrda-decomposed} was
-slightly outperformed by the Dudoit, Pang, and Tong classifiers for small values
-of $d$ but attained classification rates superior to its competitors for larger
-values of $d$. Furthermore, Figure \ref{fig:error-rates} emphasizes two
-strengths of our classifier: reliability of the classifier for relatively small
-$d$ and stability of the classifier for large $d$. In contrast to the other
-classifiers, the \emph{HDRDA} classifier exhibited less variability in its
-misclassification estimates for small values of $d$ and yielded superior
-classification accuracy as $d$ increased and was hardly effected by the
-inclusion of more variables in the data sets. We attribute the apparent
-stability of the error-rate estimates of our classifier for large values of $d$
-to its novel dimension-reduction technique.
-
-The average error-rate estimates for the Dudoit, Pang, Tong, and Witten
-classifiers were comparable for all values of $d$ examined. Furthermore, the
-error-rate estimates of the Dudoit, Pang, and Tong classifiers were
-approximately equal throughout our study, and the Witten classifier yielded
-approximately equal error-rate estimates for the Burczysnki, Shipp, and Singh
-data sets. Furthermore, the error-rate estimates for the Dudoit and Tong
-classifiers were nearly identical, which suggests that the improved mean
-estimator employed by \cite{Tong:2012hw} provided little improvement to the
-classification performance of diagonal linear discriminant analysis.
-
-The classification performance of the the Dudoit, Pang, Tong, and Witten
-classifiers for the Shipp and Singh data sets revealed the inability of these
-methods to handle larger values of $d$. After closer examination of the
-classification studies performed in the authors' original papers, we noticed
-that classification accuracy estimates were reported only for small values of
-$d$, thereby ignoring the inclusion of a larger number of variables. Hence, we
-advise practioners to consider filtering a large number of features before
-employing these methods.
-
-The popular Guo classifier is most competitive over large values of $d$ for the
-Shipp and Singh data sets. However, for the Burczynski and Nakayama data sets,
-the implementation of the Guo method in the {\tt rda} package often classified
-every test observation into a single class, thus yielding larger error rates
-than its competitors. Furthermore, the Guo classifier exhibited unreliable
-classification performance for smaller values of $d$ and required a larger
-subset of the features to perform competitively in our study.
-\begin{figure}
-<<error_rates, fig.width=8, fig.height=8, out.width='.65\\linewidth'>>=
-print(NULL)
+Similar to \cite{Witten:2011kc}, we report the average test error rates obtained
+over 250 random training-test partitions in Table \ref{tab:microarray-results}.
+Standard deviations of the test error rates are given in the parentheses. The
+\emph{HDRDA} and Guo classifiers were superior in classification performance for
+the majority of the simulations with the popular Guo classifier being the best
+for the Burczysnki data set. Both \emph{HDRDA} methods yielded comparable
+results.
+
+<<classification_results, results='asis'>>=
+table_caption <- 'The average of the test error rates obtained on gene-expression data sets over 250 random training-test partitions.
+Standard deviations of the test error rates are given in the parentheses.'
+
+xtable_out <- xtable(results_microarray,
+                     caption=table_caption,
+                     label='tab:microarray-results')
+print(xtable_out, include.rownames=FALSE)
 @
-\caption{Averages of the 1000 error-rate estimates of the competing classifiers
-  for each data set as a function of $d$. Approximate standard errors are no
-  greater than \Sexpr{max_se}.}
-\label{fig:error-rates}
-\end{figure}
-Next, we examine the optimal estimates of the tuning parameters for the
-\emph{HDRDA} classifier. For brevity, we provide graphical summaries of the
-optimal estimates of $\lambda$ and $\gamma$ in \ref{suppB}. For each of the four
-data sets, the distribution of the optimal values of $\lambda$ selected through
-cross-validation often had a notable bathtub shape, resembling a Beta
-distribution with both shape parameters less than unity. Specifically, the
-optimal values of $\lambda$ were typically near 0 or near 1, corresponding to
-quadratic and linear discriminant analysis, respectively. For instance, with
-smaller values of $d$ applied to the Shipp data set, the range of optimal values
-of $\lambda$ was typically no larger than 0.05, indicating there were a large
-number of training data sets for which we can safely assume the covariance
-matrices are unequal. As $d$ increased, the majority of the optimal values of
-$\lambda$ were selected as unity, thereby an increased number of variables
-resulted in the \emph{HDRDA} classifier preferring the more parsimonious
-assumption of equal covariance matrices. In this instance we can see the
-flexibility of our proposed classifier to employ the more realistic assumption
-of unequal class covariance matrices, but if optimal, the \emph{HDRDA}
-classifier \eqref{eq:hdrda-decomposed} can utilize the more parsimonious
-assumption of equal covariance matrices.
+
+The average error-rate estimates for the Pang, Tong, and Witten classifiers were
+comparable across all data sets. Furthermore, the average error rates for the
+Pang and Tong classifiers were approximately equal throughout our study,
+suggesting that the mean and variance estimators used in lieu of the MLEs
+provided little improvement to classification accuracies. The random-forest
+classifier was competitive only on the Singh data set. This result challenges
+the claim of \cite{FernandezDelgado:2014ul} that the random forest is
+typically superior. More studies should be performed to validate this statement
+in the small-sample, high-dimensional setting.
+
+Finally, the average test error rates for the Witten classifier were the largest
+across all data sets. Given that the standard deviations were relatively large,
+we hesitate to generalize claims regarding the Witten classifier being the
+uniformly worst performer. However, the variability in its classification
+accuracy across multiple random partitions of each data set were large enough to
+rethink applying it in practice.
 

### 2016-01-02 reports/paper-hdrda.Rnw
Updated Discussion section

@@
 
-We have compared our proposed \emph{HDRDA} classifier to several recently
-proposed classifiers, including the Dudoit, Pang, Tong, and Guo classifiers,
-each of which assumes independence of the features conditional on class
-membership. Despite the rapid computational performance of diagonal classifiers
-and the reduction in the number of parameters to estimate, \cite{Mai:2012bf} and
-\cite{Fan:2012iq} have noted that diagonal classifiers can often yield inferior
-classification performance to other classification methods. We have shown that
-the \emph{HDRDA} classifier often yields superior classification accuracy to the
-diagonal classifiers considered, confirming the assertions of \cite{Mai:2012bf}
-and \cite{Fan:2012iq}. Furthermore, the implementation of
-\eqref{eq:hdrda-decomposed} in the {\tt sparsediscrim} R package was comparable
-to that of the diagonal classifiers in terms of computational runtime. Moreover,
-when preliminary variable selection was applied to the data in our
-classification study, the time to perform model selection with our proposed
-classifier was nearly instantaneous. Hence, our proposed \emph{HDRDA} classifier
-is competitive in terms of classification accuracy for modern small-sample,
-high-dimensional data sets and requires substantially less computing time than
-standard regularized classifiers, such as regularized discriminant analysis
-proposed by \cite{Friedman:1989tm}.
-
-Given the \emph{RDA} classifier has been shown to have excellent performance in
-the high-dimensional setting \citep{Webb:2011vu} but is limited by its
+We have compared our proposed \emph{HDRDA} classifier to the random-forest and
+Witten classifiers as well as the Pang, Tong, and Guo classifiers, each of which
+assumes independence of the features conditional on class membership. Despite
+the rapid computational performance of diagonal classifiers and the reduction in
+the number of parameters to estimate, \cite{Mai:2012bf} and \cite{Fan:2012iq}
+have noted that diagonal classifiers can often yield inferior classification
+performance to other classification methods. We have shown that the \emph{HDRDA}
+classifier often yields superior classification accuracy to the diagonal
+classifiers considered when applied to small-sample, high-dimensional data sets,
+confirming the assertions of \cite{Mai:2012bf} and \cite{Fan:2012iq}.
+Furthermore, we have demonstrated that \emph{HDRDA} classifiers are more robust
+to the presence of outliers than the diagonal classifiers.
+
+We also considered the popular penalized linear discriminant analysis from
+\cite{Witten:2011kc} because it was specifically designed for high-dimensional
+gene-expression data. We had expected its classification performance to be
+competitive within our classification study and perhaps superior. Contrarily,
+its underperformance in our empirical studies show that it is sensitive to
+outliers and unable to achieve comparable results with other classifiers
+designed for small-sample, high-dimensional data. Also, despite the claims of
+\cite{FernandezDelgado:2014ul} that the random forest classifier is typically
+superior to other classifiers, we observed that the random-forest classifier was
+indeed competitive but was typically outperformed by classifiers developed for
+small-sample, high-dimensional data.
+
+The \emph{HDRDA} implementation of \eqref{eq:hdrda-decomposed} in the {\tt
+  sparsediscrim} R package was comparable to that of the diagonal classifiers in
+terms of computational runtime. Given the \emph{RDA} classifier from
+\cite{Friedman:1989tm} has been shown to have excellent performance in the
+high-dimensional setting \citep{Webb:2011vu} but is limited by its
 computationally intense model selection procedure, our work enables the
@@
 
-Despite the claims of \cite{FernandezDelgado:2014ul} that the random forest
-classifier is typically superior, we observed that the classifiers designed for
-high-dimensional data were indeed superior.
-
 \bibliographystyle{plainnat}

### 2016-01-07 reports/paper-hdrda.Rnw
Some edits from Phil.
DMY is going to have a look at this version now.

@@
 
-\usepackage{graphicx, amsmath, amssymb, bm, mathtools, amsthm}
-\usepackage{natbib, setspace}
-\usepackage{algorithm2e}
+\usepackage{graphicx, amsmath, amssymb, bm, mathtools, amsthm, natbib, setspace, algorithm2e}
 
@@
 high-dimensional setting and then show that the classification decision rule is
-equivalent to one in a subspace having a much lower dimension. As a result, the
-dimension reduction involved yields a substantial reduction in computation
-during model selection. Our parameterization also offers interpretability
-previously lacking with the RDA classifier. We demonstrate with three artificial
-and four real high-dimensional data sets that our proposed classifier is often
-superior to several recently proposed sparse and regularized classifiers in
-terms of classification accuracy. Finally, we provide an implementation of our
-proposed classifier in the {\tt sparsediscrim} R package, available on CRAN.
+equivalent to a classifier in a subspace having a much lower dimension. As a
+result, the dimension reduction involved yields a substantial reduction in
+computation during model selection. Our parameterization also offers
+interpretability previously lacking with the RDA classifier. We demonstrate with
+three artificial and four real high-dimensional data sets that our proposed
+classifier is often superior to several recently proposed sparse and regularized
+classifiers in terms of classification accuracy. Finally, we provide an
+implementation of our proposed classifier in the {\tt sparsediscrim} R package,
+available on CRAN.
 
@@
 recognition, remote sensing, and chemical analysis
-\citep{Ching:2012fu,Li:2012ev,Zhang:2010va}. Despite the advances in
+\citep*{Ching:2012fu,Li:2012ev,Zhang:2010va}. Despite the advances in
 classification to facilitate technological advancements, well-established
@@
 \emph{curse of dimensionality} \citep{Bellman:1961tn}. For instance, linear
-discriminant analysis (LDA) and quadratic discriminant analysis (QDA) are
-popular because of their simplicity, speed of computation, and
+discriminant analysis (\emph{LDA}) and quadratic discriminant analysis
+(\emph{QDA}) are popular because of their simplicity, speed of computation, and
 interpretability. However, both methods are incalculable when $p > N$ because
@@
 information is removed
-\citep{Fan:2012iq,Ye:2009gd,Bouveyron:2007gx,Ye:2006vx,Ye:2005hd,Antoniadis:2003bk}. Finally,
-\cite{Mai:2012bf}, \cite{Fan:2012iq}, \cite{Merchante:2012vk},
-\cite{Clemmensen:2011kr}, \cite{Huang:2010ju}, and \cite{Dudoit:2002ev} have
+\citep*{Fan:2012iq,Ye:2009gd,Bouveyron:2007gx,Ye:2006vx,Ye:2005hd,Antoniadis:2003bk}.
+Finally, \cite*{Mai:2012bf}, \cite{Fan:2012iq}, \cite*{Merchante:2012vk},
+\cite*{Clemmensen:2011kr}, \cite*{Huang:2010ju}, and \cite*{Dudoit:2002ev} have
 proposed restrictions on the population covariance matrices to reduce the number
@@
 discriminant analysis and other techniques, see \cite{Murphy:2012uq} and
-\cite{Mkhadri:1997gy}.
+\cite*{Mkhadri:1997gy}.
 
@@
 ensure positive definiteness. By reparameterizing the \emph{RDA} classifier
-similar to \cite{Hastie:2008dt} and \cite{Halbe:2007ea}, we first show the
+similar to \cite*{Hastie:2008dt} and \cite{Halbe:2007ea}, we first show that the
 pooling parameter in the \emph{HDRDA} classifier determines the contribution of
@@
 adjustments to the zero eigenvalues, the decision rule in the original feature
-space is equivalent to a decision rule in a much lower dimension, such that the
-matrix inverses and determinants of relatively small matrices can be rapidly
-computed. Furthermore, we show that a large number of shrinkage methods that are
-special cases of the \emph{HDRDA} classifier have no effect on the zero
-eigenvalues of the covariance-matrix estimators. Such methods include work from
+space is equivalent to a decision rule in a lower dimension, such that matrix
+inverses and determinants of relatively small matrices can be rapidly computed.
+Furthermore, we show that several shrinkage methods that are special cases of
+the \emph{HDRDA} classifier have no effect on the zero eigenvalues of the
+covariance-matrix estimators. Such methods include work from
 \cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and other methods studied by
@@
 performance of the \emph{HDRDA} classifier on four real high-dimensional data
-sets along with a simulation experiment that generalizes that initially
+sets along with a simulation design that generalizes the experiments initially
 conducted by \cite{Guo:2007te}. We demonstrate that the \emph{HDRDA} classifier
 often attains superior classification accuracy to several recent classifiers
-designed for small-sample, high-dimensional data from \cite{Tong:2012hw},
-\cite{Witten:2011kc}, \cite{Pang:2009ik}, and \cite{Guo:2007te}. We also include
-as a benchmark the random forest from \cite{Breiman:2001fb} because
-\cite{FernandezDelgado:2014ul} has concluded that the random forest is often
+designed for small-sample, high-dimensional data from \cite*{Tong:2012hw},
+\cite{Witten:2011kc}, \cite*{Pang:2009ik}, and \cite{Guo:2007te}. We also
+include as a benchmark the random forest from \cite{Breiman:2001fb} because
+\cite*{FernandezDelgado:2014ul} has concluded that the random forest is often
 superior to other classifiers in benchmark studies.
@@
 \mathbb{R}_{s \times s}$ \citep[Chapter 1]{Lutkepohl:1996uz}. Let $\bm A^{T}$,
-$\bm A^+$, and $\mathcal{N}(\bm A) = \{\bm z \in \mathbb{R}_{p \times 1} : \bm A
-\bm z = \bm 0_{m \times 1}\}$ denote the transpose, the Moore-Penrose
+$\bm A^+$, and $\mathcal{N}(\bm A)$ denote the transpose, the Moore-Penrose
 pseudoinverse, and the null space of $\bm A \in \mathbb{R}_{m \times p}$,
@@
 denote the cone of real $p \times p$ positive-semidefinite matrices. Let
-$V^{\perp} = \{\bm y \in \mathbb{R}_{p \times 1} : \bm y^{T} \bm v = 0$ for all
-$\bm v \in V \}$ denote the orthogonal complement of a vector space $V \subset
+$V^{\perp}$ denote the orthogonal complement of a vector space $V \subset
 \mathbb{R}_{p \times 1}$. For $c \in \mathbb{R}$, let $c^+ = 1/c$ if $c \ne 0$
@@
 N)$ with true, unique membership $y_i \in \{\omega_1, \ldots, \omega_K\}$. We
-assume that $(\bm x_i, y_i)$ is a realization from a mixture distribution
-$p(\bm x) = \sum_{k=1}^K p(\bm x | \omega_k) p(\omega_k)$, where $p(\bm x |
-\omega_k)$ is the probability density function (PDF) of the $k$th class and
-$p(\omega_k)$ is the prior probability of class membership of the $k$th
-class. We further assume that each of $p(\omega_k)$ is equal, $k = 1, \ldots,
-K$.
-
-Quadratic discriminant analysis is the optimal Bayesian decision rule with
-respect to a $0-1$ loss function when $p(\bm x | \omega_k)$ is the probability
-density function of the multivariate normal distribution with known mean vectors
-$\bm \mu_k \in \mathbb{R}_{p \times 1}$ and known covariance matrices $\bm
-\Sigma_k \in \mathbb{R}_{p \times p}^{>}$, $k = 1, 2, \ldots, K$. Because $\bm
-\mu_k$ and $\bm \Sigma_k$ are typically unknown, we assign an unlabeled
-observation $\bm x$ to class $\omega_k$ with the sample quadratic discriminant
-analysis classifier
+assume that $(\bm x_i, y_i)$ is a realization from a mixture distribution $p(\bm
+x) = \sum_{k=1}^K p(\bm x | \omega_k) p(\omega_k)$, where $p(\bm x | \omega_k)$
+is the probability density function (PDF) of the $k$th class and $p(\omega_k)$
+is the prior probability of class membership of the $k$th class. We further
+assume $p(\omega_k) = p(\omega_l)$, $1 \le k, l \le K$, $k \ne l$.
+
+The \emph{QDA} classifier is the optimal Bayesian decision rule with respect to
+a $0-1$ loss function when $p(\bm x | \omega_k)$ is the PDF of the multivariate
+normal distribution with known mean vectors $\bm \mu_k \in \mathbb{R}_{p \times 1}$
+and known covariance matrices $\bm \Sigma_k \in \mathbb{R}_{p \times p}^{>}$,
+$k = 1, 2, \ldots, K$. Because $\bm \mu_k$ and $\bm \Sigma_k$ are typically
+unknown, we assign an unlabeled observation $\bm x$ to class $\omega_k$ with the
+sample \emph{QDA} classifier
 \begin{align}
@@
 \end{align}
-is the MLE for $\bm \Sigma$. In this case \eqref{eq:qda} reduces to the sample
-linear discriminant analysis classifier, and the log-determinant is omitted
-because it is constant across the $K$ classes.
+is the MLE for $\bm \Sigma$. Here, \eqref{eq:qda} reduces to the sample
+\emph{LDA} classifier. We omit the log-determinant because it is constant across
+the $K$ classes.
 
@@
 Several regularization methods, such as the methods considered by
-\cite{Xu:2009fl}, \citet*{Guo:2007te}, and \cite{Mkhadri:1995jp} have been
+\cite{Xu:2009fl}, \cite{Guo:2007te}, and \cite{Mkhadri:1995jp} have been
 proposed in the literature to adjust the eigenvalues of $\widehat{\bm \Sigma}_k$
@@
 \begin{align}
-  \widehat{\bm \Sigma}_k(\lambda) = (1 - \lambda) \widehat{\bm \Sigma}_k + \lambda
+  \widehat{\bm \Sigma}_k(\lambda) \coloneqq (1 - \lambda) \widehat{\bm \Sigma}_k + \lambda
   \widehat{\bm \Sigma}, \quad k = 1, \ldots, K,\label{eq:sig-lambda-alternative}
@@
 \begin{align}
-	\tilde{\bm \Sigma}_k = \alpha_k \widehat{\bm \Sigma}_k(\lambda) + \gamma \bm I_p,\label{eq:hdrda-cov}
+	\tilde{\bm \Sigma}_k \coloneqq \alpha_k \widehat{\bm \Sigma}_k(\lambda) + \gamma \bm I_p,\label{eq:hdrda-cov}
 \end{align}
@@
 S_k(\lambda) = (1 - \lambda) \bm S_k + \lambda \bm S$. If $\lambda = 0$,
-\eqref{eq:sig-lambda} is equal to $\widehat{\bm \Sigma}_k$ used in
-\eqref{eq:qda}. Alternatively, if $\lambda = 1$, \eqref{eq:sig-lambda} reduces
-to \eqref{eq:pooled-cov}. To further improve the estimation of $\bm \Sigma_k$
-and to stabilize the inverse of \eqref{eq:sig-lambda}, \cite{Friedman:1989tm}
-has proposed the biased covariance-matrix estimator
+\eqref{eq:sig-lambda} is equal to $\widehat{\bm \Sigma}_k$ in \eqref{eq:qda}.
+Alternatively, if $\lambda = 1$, then \eqref{eq:sig-lambda} reduces to
+\eqref{eq:pooled-cov}. To further improve the estimation of $\bm \Sigma_k$ and
+to stabilize the inverse of \eqref{eq:sig-lambda}, \cite{Friedman:1989tm} has
+proposed the biased covariance-matrix estimator
 \begin{align}
@@
 eigenvalues have no effect on \eqref{eq:hdrda}. Lastly, we utilize the singular
-value decomposition to calculate efficiently the eigenvalue decomposition of
+value decomposition to efficiently calculate the eigenvalue decomposition of
 $\widehat{\bm \Sigma}$, further reducing the computational costs of the
@@
 Let $\bm U_2$ be defined as above. Then, for all $\bm x \in \mathbb{R}_{p \times
-  1}$, $\bm U_2^{T} (\bm x - \xbar_k) = \bm U_2^{T} (\bm x - \xbar_{k'})$ $(1 \le k,
-k' \le K)$.
+  1}$, $\bm U_2^{T} (\bm x - \xbar_k) = \bm U_2^{T} (\bm x - \xbar_{k'})$, $1
+\le k, k' \le K$, where $k \ne k'$.
 \end{lemma}
@@
 Theorem 18.1.1 from \cite{Harville:2008wja}, which states that $|\bm A + \bm B
-\bm T \bm C| = |\bm A| |\bm T| |\bm T^{-1} + \bm C \bm A^{-1} \bm B|$. Thus,
-setting $\bm A = \bm \Gamma_k$, $\bm B = \alpha_k (1 - \lambda) \bm U_1^{T} \bm
-X_k^{T}$, $\bm T = \bm I_{n_k}$, and $\bm C = \bm X_k \bm U_1$, we have $|\bm
-W_k| = |\bm \Gamma_k| |\bm Q_k|$. Similarly, \eqref{eq:W_k_inv} follows from the
-well-known Sherman-Woodbury formula \citep[Theorem 18.2.8]{Harville:2008wja} by
-noting that $(\bm A + \bm B \bm T \bm C)^{-1} = \bm A^{-1} - \bm A^{-1} \bm B
-(\bm T^{-1} + \bm C \bm A^{-1} \bm B)^{-1} \bm C \bm A^{-1}$.
+\bm T \bm C| = |\bm A| |\bm T| |\bm T^{-1} + \bm C \bm A^{-1} \bm B|$, where
+$\bm A \in \mathbb{R}_{a \times a}^{>}$, $\bm B \in \mathbb{R}_{a \times b}$,
+$\bm T \in \mathbb{R}_{b \times b}^{>}$, and $\bm C \in \mathbb{R}_{b \times
+  a}$. Thus, setting $\bm A = \bm \Gamma_k$, $\bm B = \alpha_k (1 - \lambda) \bm
+U_1^{T} \bm X_k^{T}$, $\bm T = \bm I_{n_k}$, and $\bm C = \bm X_k \bm U_1$, we
+have $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$. Similarly, \eqref{eq:W_k_inv}
+follows from the well-known Sherman-Woodbury formula \citep[Theorem
+  18.2.8]{Harville:2008wja} by noting that $(\bm A + \bm B \bm T \bm C)^{-1} =
+\bm A^{-1} - \bm A^{-1} \bm B (\bm T^{-1} + \bm C \bm A^{-1} \bm B)^{-1} \bm C
+\bm A^{-1}$.
 \end{proof}
@@
 nonsingular. This assumption is not met when $(\lambda, \gamma) = (0, 0)$
-because $\bm \Gamma_k = \bm 0_q$. In this case we instead use the formulation in
-\eqref{eq:hdrda-decomposed}. Also, notice that if $\alpha_k$ is constant across
-the $K$ classes, then $\bm \Gamma_k$ in \eqref{eq:Gamma_k} is independent of
-$k$. Consequently, $|\bm \Gamma_k|$ is constant across the $K$ classes and need
-not be calculated in \eqref{eq:hdrda-decomposed}.
+because $\bm \Gamma_k = \bm 0_q$. Here, we use the formulation in
+\eqref{eq:hdrda-decomposed} instead. Also, notice that if $\alpha_k$ is constant
+across the $K$ classes, then $\bm \Gamma_k$ in \eqref{eq:Gamma_k} is independent
+of $k$. Consequently, $|\bm \Gamma_k|$ is constant across the $K$ classes and
+need not be calculated in \eqref{eq:hdrda-decomposed}.
 
@@
 \ref{alg:hdrda_model_selection} is our usage of the compact singular value
-decomposition (SVD) (sometimes called the reduced or fast SVD). Rather than
-computing the eigenvalue decomposition of $\widehat{\bm \Sigma}$ to obtain $\bm
-U_1$, we instead obtain $\bm U_1$ by computing the eigendecomposition of a much
-smaller $N \times N$ matrix when $p \gg N$
-\cite[Chapter~18.3.5]{Hastie:2008dt}. Applying the SVD, we decompose $\bm X_c =
-\bm M \bm \Delta \bm U^{T}$, where $\bm M \in \mathbb{R}_{N \times p}$ is
+decomposition (SVD). Rather than computing the eigenvalue decomposition of
+$\widehat{\bm \Sigma}$ to obtain $\bm U_1$, we instead obtain $\bm U_1$ by
+computing the eigendecomposition of a much smaller $N \times N$ matrix when $p
+\gg N$ \cite[Chapter~18.3.5]{Hastie:2008dt}. Applying the SVD, we decompose $\bm
+X_c = \bm M \bm \Delta \bm U^{T}$, where $\bm M \in \mathbb{R}_{N \times p}$ is
 orthogonal, $\bm \Delta \in \mathbb{R}_{p \times p}^{\ge}$ is a diagonal matrix
@@
 by calculating the number of eigenvalues that exceeds some tolerance value, say,
-$1 \times 10^{-6}$. After calculating $q$, we extract $\bm U_1$ by retaining
-only the first $q$ columns of $\bm U$.
+$1 \times 10^{-6}$. We then extract $\bm U_1$ as the first $q$ columns of $\bm
+U$.
 
 As a result of the compact SVD, we need calculate $\bm X_c \bm U_1$ only once
-per cross-validation fold, requiring $O(p q N) \approx (p N^2)$ calculations.
+per cross-validation fold, requiring $O(p q N) \approx O(p N^2)$ calculations.
 Hence, the computational costs of expensive calculations, such as matrix
@@
 surname. All simulations were conducted on an Amazon Elastic Compute Cloud (EC2)
-instance of type `c4.4xlarge`.
+{\tt c4.4xlarge} instance.
 
@@
 the correlation decays as a function of the distance between any two genes. The
-autocorrelation $\rho_k$ alternates between a positive and a negative value from
-block to block. The original design from \cite{Guo:2007te} comprised two
-$p$-dimensional multivariate normal populations with a common block-diagonal
-covariance matrix where $\bm\Sigma_1 = \bm\Sigma_2$.
+original design from \cite{Guo:2007te} comprised two $p$-dimensional
+multivariate normal populations with a common block-diagonal covariance matrix
+where $\bm\Sigma_1 = \bm\Sigma_2$.
 
@@
 outliers. As a result, we wished to investigate the effect of outliers on the
-high-dimensional classifiers. To do this, we generalized the block-diagonal
-simulation configuration by sampling from a $p$-dimensional multivariate
-contaminated normal distribution.  Denoting the PDF of the $p$-dimensional
-multivariate normal distribution by $N_p(\bm x | \bm \mu, \bm \Sigma)$, we write
-the PDF of the $k$th class as
-\begin{align}
-  p(\bm x | \omega_k) = (1 - \epsilon) N_p(\bm x | \bm \mu_k, \bm \Sigma_k) + \epsilon N_p(\bm x | \bm \mu_k, \omega \bm \Sigma_k),\label{eq:contaminated-pdf}
+high-dimensional classifiers. To accomplish this goal, we generalized the
+block-diagonal simulation configuration by sampling from a $p$-dimensional
+multivariate contaminated normal distribution.  Denoting the PDF of the
+$p$-dimensional multivariate normal distribution by $N_p(\bm x | \bm \mu, \bm
+\Sigma)$, we write the PDF of the $k$th class as
+\begin{align}
+  p(\bm x | \omega_k) = (1 - \epsilon) N_p(\bm x | \bm \mu_k, \bm \Sigma_k) + \epsilon N_p(\bm x | \bm \mu_k, \eta \bm \Sigma_k),\label{eq:contaminated-pdf}
 \end{align}
 where $\epsilon \in [0, 1]$ is the probability that an observation is
-contaminated (i.e., drawn from a distribution with larger variance) and $\omega
-> 1$ scales the covariance matrix $\bm \Sigma_k$ to increase the extremity of
+contaminated (i.e., drawn from a distribution with larger variance) and $\eta >
+1$ scales the covariance matrix $\bm \Sigma_k$ to increase the extremity of
 outliers. For $\epsilon = 0$, we have the benchmark block-diagonal simulation
 design from \cite{Guo:2007te}. As $\epsilon$ is increased, the average number of
-outliers is increased. In our simulation, we considered the values of $\epsilon
-= 0$, $0.05$, $\ldots$, $0.50$. We also set $\omega = 100$.
+outliers is increased. In our simulation, we set $\eta = 100$ and considered the
+values of $\epsilon = 0$, $0.05$, $\ldots$, $0.50$.
 
@@
 both \emph{HDRDA} classifiers outperformed all other classifiers, suggesting
-that their dimension reduction inherent better captured the classificatory
+that their inherent dimension reduction better captured the classificatory
 information in the small samples even in the presence of outliers.
@@
 
-\subsection{Classification Results}
+\subsubsection{Classification Results}
 
 Similar to \cite{Witten:2011kc}, we report the average test error rates obtained
-over 250 random training-test partitions in Table \ref{tab:microarray-results}.
-Standard deviations of the test error rates are given in the parentheses. The
+over 250 random training-test partitions in Table \ref{tab:microarray-results}
+along with standard deviations of the test error rates in parentheses. The
 \emph{HDRDA} and Guo classifiers were superior in classification performance for
-the majority of the simulations with the popular Guo classifier being the best
-for the Burczysnki data set. Both \emph{HDRDA} methods yielded comparable
-results.
+the majority of the simulations with the popular Guo classifier providing
+superior classification performance for the Burczysnki data set. Both
+\emph{HDRDA} methods yielded comparable results.
 
@@
 \emph{RDA} classifier to be employed with small-sample, high-dimensional data in
-practice. This is reassuring because the \emph{RDA} classifier remains widely
-popular in the literature. In fact, variants of the RDA classifier have been
-applied to microarray data \citep{Ching:2012fu,Li:2012ev,Tai:2007bk,Guo:2007te},
-facial recognition
+practice. This results is reassuring because the \emph{RDA} classifier remains
+widely popular in the literature. In fact, variants of the RDA classifier have
+been applied to microarray data
+\citep{Ching:2012fu,Li:2012ev,Tai:2007bk,Guo:2007te}, facial recognition
 \citep{Ching:2012fu,Zhang:2010va,Dai:2007vd,Lu:2005hq,Pima:2004vw,Lu:2003we},
@@
 \citep{Tadjudin:1999fk}, seismic detection \citep{Anderson:2002kg}, and chemical
-spectra \citep{Wu:1996us,Aeberhard:1993fp}.
+spectra \citep*{Wu:1996us,Aeberhard:1993fp}.
 
 The dimension reduction that we have employed in this paper has reduced the
-dimension to the rank$(\widehat{\bm \Sigma})$, $q$. An interesting extension of
+dimension to rank$(\widehat{\bm \Sigma}) = q$. An interesting extension of
 our work would reduce the dimension $q$ further to a lower dimension $q_L < q$

### 2016-01-08 sim-microarray.r
Consider more data sets in sims at DMY's request.

@@ -7,7 +7,7 @@ num_cores <- 8
 train_pct <- 2/3
 num_iterations <- 250
 
-data_sets <- c("burczynski", "nakayama", "shipp", "singh")
+data_sets <- c("burczynski", "christensen", "khan", "nakayama", "sorlie", "su", "sun", "yeoh")
 sim_config <- rep(data_sets, each=num_iterations)
 
 flog.logger("sim", INFO, appender=appender.file('sim-microarray.log'))

### 2016-01-10 sim-microarray.r
Consider most of K=2 datamicroarray data sets

@@ -7,7 +7,8 @@ num_cores <- 8
 train_pct <- 2/3
 num_iterations <- 250
 
-data_sets <- c("burczynski", "christensen", "khan", "nakayama", "sorlie", "su", "sun", "yeoh")
+data_sets <- c('alon', 'chiaretti', 'chin', 'chowdary', 'gordon', 'gravier',
+               'pomeroy', 'subramanian', 'tian', 'west')
 sim_config <- rep(data_sets, each=num_iterations)
 
 flog.logger("sim", INFO, appender=appender.file('sim-microarray.log'))

### 2016-01-10 lib/helpers.R
Restricted Chiaretti data set to larger classes

@@ -179,8 +179,18 @@ load_microarray <- function(dataset) {
 
     su$x <- su$x[, features_kept]
     data_out <- su
-
-  } else {
+  } else if (dataset == 'chiaretti') {
+    # As in Xu et al. (2009), we restricted the data set to the n1 = 74
+    # observations without cytogenetic abnormalities and the n2 = 37
+    # observations with a detected BCR/ABL gene.
+    data('chiaretti', package = 'datamicroarray')
+    classes <- with(chiaretti, levels(y)[table(y) >= 37])
+    idx <- which(chiaretti$y %in% classes)
+    chiaretti$x <- chiaretti$x[idx, ]
+    chiaretti$y <- factor(chiaretti$y[idx])
+    data_out <- chiaretti
+  }
+  else {
     data(list = dataset, package = "datamicroarray")
     data_out <- get(dataset)
   }

### 2016-01-11 sim-microarray.r
Added Dudoit's variable selection to data sims.

@@ -2,13 +2,14 @@ library(ProjectTemplate)
 load.project()
 
 set.seed(42)
-num_cores <- 8
+num_cores <- 16
 
 train_pct <- 2/3
-num_iterations <- 250
+num_iterations <- 100
 
-data_sets <- c('alon', 'chiaretti', 'chin', 'chowdary', 'gordon', 'gravier',
-               'pomeroy', 'subramanian', 'tian', 'west')
+data_sets <- c('alon', 'burczynski', 'chiaretti', 'chin', 'chowdary', 'gravier',
+               'nakayama', 'shipp', 'singh', 'tian')
+
 sim_config <- rep(data_sets, each=num_iterations)
 
 flog.logger("sim", INFO, appender=appender.file('sim-microarray.log'))
@@ -26,6 +27,11 @@ results <- mclapply(seq_along(sim_config), function(i) {
   train_y <- data$y[rand_split$training]
   test_x <- as.matrix(data$x[rand_split$test, ])
   test_y <- data$y[rand_split$test]
+
+  # Apply Dudoit's variable selection with 1000 variables to reduce runtime.
+  top_features <- dudoit(train_x, train_y, num_variables=1000)
+  train_x <- train_x[, top_features]
+  test_x <- test_x[, top_features]
 
   flog.info("Training Data: (Observations, Dimension): (%s, %s) -- Data Set: %s -- Sim Config: %s of %s",
             nrow(train_x), ncol(train_x), data_set, i, length(sim_config), name="sim")

### 2016-01-11 sim-microarray.r
Rerunning sims with the final 6 considered.

@@ -7,8 +7,7 @@ num_cores <- 16
 train_pct <- 2/3
 num_iterations <- 100
 
-data_sets <- c('alon', 'burczynski', 'chiaretti', 'chin', 'chowdary', 'gravier',
-               'nakayama', 'shipp', 'singh', 'tian')
+data_sets <- c('chiaretti', 'chowdary', 'nakayama', 'shipp', 'singh', 'tian')
 
 sim_config <- rep(data_sets, each=num_iterations)
 

### 2016-01-11 reports/paper-hdrda.Rnw
Updated results. Added summaries for new data sets

@@
 benchmark gene-expression microarray data sets. We evaluated the classification
-performance of each classifier by randomly partitioning the data set under
+accuracy of each classifier by randomly partitioning the data set under
 consideration such that $2/3$ of the observations were allocated as training
 data and the remaining $1/3$ of the observations were allocated as a test data
-set. After training each classifier on the training data, we classified the test
-data sets and computed the proportion of mislabeled test observations to
-estimate the classification error rate for each classifier. Repeating this
-process 250 times, we computed the average of the error-rate estimates for each
-classifier. Here, we provide a concise description of each high-dimensional
-data set examined in our classification study.
-
-\subsubsection{\cite{Burczynski:2006ik} Data Set}
-
-\cite{Burczynski:2006ik} acquired peripheral blood mononuclear cells (PBMC)
-through hybridization to microarrays from 127 individuals resulting in 22,283
-sequences. Of the 127 individuals, 42 were healthy, 59 had Crohn's disease (CD),
-and 26 had ulcerative colitis (UC). The goal of \cite{Burczynski:2006ik} was to
-improve accuracy in the discrimination of Inflammatory Bowel Disease (IBD) using
-the PBMC-based gene expression signature of a patient.
+set. To expedite the computational runtime, we reduced the training data to the
+top 1000 variables by employing the variable-selection method proposed by
+\cite{Dudoit:2002ev}. We then reduced the test data set to the same 1000
+variables. After training each classifier on the training data, we classified
+the test data sets and computed the proportion of mislabeled test observations
+to estimate the classification error rate for each classifier. Repeating this
+process 100 times, we computed the average of the error-rate estimates for each
+classifier. Here, we provide a concise description of each high-dimensional data
+set examined in our classification study.
+
+\subsubsection{\cite{Chiaretti:2004gq} Data Set}
+
+\cite{Chiaretti:2004gq} measured the gene-expression profiles for 128
+individuals with acute lymphoblastic leukaemia (ALL) using Affymetrix human
+95Av2 arrays. Following \cite{Xu:2009fl}, we restricted the data set to $K = 2$
+classes such that $n_1 = 74$ observations were without cytogenetic abnormalities
+and $n_2 = 37$ observations have a detected BCR/ABL gene. The robust multichip
+average normalization method was applied to all 12,625 gene expression levels.
+
+\subsubsection{\cite{Chowdary:2006kf} Data Set}
+
+\cite{Chowdary:2006kf} investigated 52 matched pairs of tissues from colon and
+breast tumors using Affymetrix U133A arrays and ribonucleic-acid (RNA)
+amplification. Each tissue pair was gathered from the same patient and
+consisted of a snap-frozen tissue and a tissue suspended in a RNAlater
+preservative. Overall, 31 breast-cancer and 21 colon-cancer pairs were gathered,
+resulting in $K = 2$ classes with $n_1 = 62$ and $n_2 = 42$. A purpose of the
+study was to determine whether the disease state could be identified using
+22,283 gene-expression profiles.
 
@@
 
+\subsubsection{\cite{Tian:2003ht} Data Set}
+
+\cite{Tian:2003ht} investigated the purified plasma cells from the bone marrow
+of control patients along with patients with newly diagnosed multiple
+myeloma. Expression profiles for 12,2625 genes were obtained via Affymetrix
+U95Av2 microarrays. The plasma cells were subjected to biochemical and
+immunohistochemical analyses to identify molecular determinants of osteolytic
+lesions. For 36 multiple-myloma patients, focal bone lesions could not be
+detected by magnetic resonance imaging (MRI), whereas MRI was used to detect
+such lesions in 137 patients.
+
 \subsubsection{Classification Results}
@@
 Similar to \cite{Witten:2011kc}, we report the average test error rates obtained
-over 250 random training-test partitions in Table \ref{tab:microarray-results}
+over 100 random training-test partitions in Table \ref{tab:microarray-results}
 along with standard deviations of the test error rates in parentheses. The
@@
 <<classification_results, results='asis'>>=
-table_caption <- 'The average of the test error rates obtained on gene-expression data sets over 250 random training-test partitions.
+table_caption <- 'The average of the test error rates obtained on gene-expression data sets over 100 random training-test partitions.
 Standard deviations of the test error rates are given in the parentheses.'

### 2016-01-11 reports/paper-hdrda.Rnw
Updated results summary with PDY's changes.

@@
 \emph{HDRDA} and Guo classifiers were superior in classification performance for
-the majority of the simulations with the popular Guo classifier providing
-superior classification performance for the Burczysnki data set. Both
-\emph{HDRDA} methods yielded comparable results.
+the majority of the simulations. The \emph{HDRDA} classifiers had the best
+classification accuracy on the Chowdary and Shipp data sets. Although the random
+forest's accuracy slightly exceeded the \emph{HDRDA} classifiers on the Tian
+data set, our proposed methods outperformed the other classifiers considered.
+Additionally, the \emph{HDRDA} classifiers had comparable performance on five of
+the six data sets.
 
@@
 comparable across all data sets. Furthermore, the average error rates for the
-Pang and Tong classifiers were approximately equal throughout our study,
-suggesting that the mean and variance estimators used in lieu of the MLEs
-provided little improvement to classification accuracies. The random-forest
-classifier was competitive only on the Singh data set. This result challenges
-the claim of \cite{FernandezDelgado:2014ul} that the random forest is
-typically superior. More studies should be performed to validate this statement
-in the small-sample, high-dimensional setting.
-
-Finally, the average test error rates for the Witten classifier were the largest
-across all data sets. Given that the standard deviations were relatively large,
-we hesitate to generalize claims regarding the Witten classifier being the
-uniformly worst performer. However, the variability in its classification
-accuracy across multiple random partitions of each data set were large enough to
-rethink applying it in practice.
+Pang and Tong classifiers were approximately equal for all data sets except for
+the Chiaretti dataset. This result suggests that the mean and variance
+estimators used in lieu of the MLEs provided little improvement to
+classification accuracies. However, we investigated the Pang classifier's poor
+performance on the Chiaretti data set and determined that its variance estimator
+exhibited numerical instability. The classifier's denominator was approximately
+zero for both classes and lead to the poor classification performance.
+
+The random-forest classifier was competitive when applied to the Chowdary and
+Singh data sets and yielded the smallest error rate of the considered
+classifiers on the Tian data set. The fact that the \emph{HDRDA} and Guo
+classifiers typically outperformed the random-forest classifier challenges the
+claim of \cite{FernandezDelgado:2014ul} that the random forest is typically
+superior. Further studies should be performed to validate this statement in the
+small-sample, high-dimensional setting.
+
+Finally, the Pang, Tong, and Witten classifiers consistently yielded the largest
+error rates across all data sets. Given that the standard deviations were
+relatively large, we hesitate to generalize claims regarding the ranking of
+these three classifiers in terms of error rate. However, the classifiers' error
+rates and their variability across multiple random partitions of each data set
+were large enough to rethink applying them in practice.
 

### 2016-01-13 reports/paper-hdrda.Rnw
Added DMY's changes.

@@
 
-\citeauthor{Friedman:1989tm} proposed the popular regularized discriminant
-analysis (RDA) classifier that utilizes a biased covariance-matrix estimator to
-partially pool the sample covariance matrices from linear and quadratic
-disriminant analysis before shrinking the resulting estimator towards a scaled
+\cite{Friedman:1989tm} proposed the popular regularized discriminant analysis
+(RDA) classifier that utilizes a biased covariance-matrix estimator that
+partially pools the sample covariance matrices from linear and quadratic
+disriminant analysis and shrinks the resulting estimator towards a scaled
 identity matrix. The RDA classifier's two tuning parameters are typically
-estimated via a computationally burdensome cross-validation procedure from a
-grid of candidate values. We reformulate the RDA classifier in the small-sample,
+estimated via a computationally burdensome cross-validation procedure using a
+grid search. We formulate a new RDA classifier for the small-sample,
 high-dimensional setting and then show that the classification decision rule is
 equivalent to a classifier in a subspace having a much lower dimension. As a
-result, the dimension reduction involved yields a substantial reduction in
-computation during model selection. Our parameterization also offers
-interpretability previously lacking with the RDA classifier. We demonstrate with
-three artificial and four real high-dimensional data sets that our proposed
-classifier is often superior to several recently proposed sparse and regularized
-classifiers in terms of classification accuracy. Finally, we provide an
-implementation of our proposed classifier in the {\tt sparsediscrim} R package,
-available on CRAN.
+result, the utilization of the dimension-reduction step yields a substantial
+reduction in computation during model selection. Also, our parameterization
+offers interpretability previously lacking with the RDA classifier. We
+demonstrate that our proposed classifier is often superior to several recently
+proposed sparse and regularized classifiers in terms of classification accuracy
+with three artificial and six real high-dimensional data sets. Finally, we
+provide an implementation of our proposed classifier in the {\tt sparsediscrim}
+R package, which is available on CRAN.
 
@@
 classification to facilitate technological advancements, well-established
-classifiers often degrade in accuracy with modern, high-dimensional data sets
-when the number of features $p$ exceeds the total sample size $N$, due to the
-\emph{curse of dimensionality} \citep{Bellman:1961tn}. For instance, linear
-discriminant analysis (\emph{LDA}) and quadratic discriminant analysis
+classifiers often degrade in accuracy when applied to modern, high-dimensional
+data sets if the number of features $p$ exceeds the total sample size $N$, due
+to the \emph{curse of dimensionality} \citep{Bellman:1961tn}. For instance,
+linear discriminant analysis (\emph{LDA}) and quadratic discriminant analysis
 (\emph{QDA}) are popular because of their simplicity, speed of computation, and
@@
 
-Here, we present a new classification method intended for the $p > N$ case that
-builds on the classic regularized discriminant analysis (\emph{RDA}) proposed by
-\cite{Friedman:1989tm}. Our proposed high-dimensional \emph{RDA} (\emph{HDRDA})
-classifier employs a biased covariance matrix estimator that partially pools the
-individual sample covariance matrices from the \emph{QDA} classifier with the
-pooled sample covariance matrix from the \emph{LDA} classifier. We then shrink
-the resulting covariance-matrix estimator towards a scaled identity matrix to
-ensure positive definiteness. By reparameterizing the \emph{RDA} classifier
-similar to \cite*{Hastie:2008dt} and \cite{Halbe:2007ea}, we first show that the
-pooling parameter in the \emph{HDRDA} classifier determines the contribution of
-each training observation to the estimation of each class covariance matrix,
-enabling interpretability that has been previously lacking with the \emph{RDA}
-classifier \citep{Bensmail:1996fw}.
-
-Next, we achieve a substantial reduction in dimension to reduce greatly the
-computation during model selection by applying a reasoning similar to
-\cite{Ye:2006tq}. We establish that the matrix operations corresponding to the
-null space of the pooled sample covariance matrix are redundant and can be
-discarded from the \emph{HDRDA} decision rule without loss of classificatory
-information. Moreover, because the \emph{HDRDA} decision rule is invariant to
-adjustments to the zero eigenvalues, the decision rule in the original feature
-space is equivalent to a decision rule in a lower dimension, such that matrix
-inverses and determinants of relatively small matrices can be rapidly computed.
-Furthermore, we show that several shrinkage methods that are special cases of
-the \emph{HDRDA} classifier have no effect on the zero eigenvalues of the
-covariance-matrix estimators. Such methods include work from
+Here, we present the high-dimensional \emph{RDA} (\emph{HDRDA}) classifier,
+which is intended for the $p > N$ case and builds on the classic \emph{RDA}
+classifier proposed by \cite{Friedman:1989tm}. In this paper, we study the
+classification performance of the \emph{HDRDA} classifier on six real
+high-dimensional data sets along with a simulation design that generalizes the
+experiments initially conducted by \cite{Guo:2007te}. We demonstrate that the
+\emph{HDRDA} classifier often attains superior classification accuracy to
+several recent classifiers designed for small-sample, high-dimensional data from
+\cite*{Tong:2012hw}, \cite{Witten:2011kc}, \cite*{Pang:2009ik}, and
+\cite{Guo:2007te}. We also include as a benchmark the random forest from
+\cite{Breiman:2001fb} because \cite*{FernandezDelgado:2014ul} have concluded
+that the random forest is often superior to other classifiers in benchmark
+studies. We show that our proposed classifier is competitive and often
+outperforms the random forest in terms of classification accuracy in the
+small-sample, high-dimensional setting.
+
+Our proposed \emph{HDRDA} classifier employs a biased covariance-matrix
+estimator that partially pools the individual sample covariance matrices from
+the \emph{QDA} classifier with the pooled sample covariance matrix from the
+\emph{LDA} classifier. We then shrink the resulting covariance-matrix estimator
+towards a scaled identity matrix to ensure positive definiteness. By
+reparameterizing the \emph{RDA} classifier similar to \cite*{Hastie:2008dt} and
+\cite{Halbe:2007ea}, we show that the pooling parameter in the \emph{HDRDA}
+classifier determines the contribution of each training observation to the
+estimation of each class covariance matrix, enabling interpretability that has
+been previously lacking with the \emph{RDA} classifier \citep{Bensmail:1996fw}.
+
+Next, we achieve a substantial reduction in dimension that greatly reduces the
+computational effort during model selection when $p \gg N$ by applying reasoning
+similar to \cite{Ye:2006tq}. We establish that the matrix operations
+corresponding to the null space of the pooled sample covariance matrix are
+redundant and can be discarded from the \emph{HDRDA} decision rule without loss
+of classificatory information. Moreover, because the \emph{HDRDA} decision rule
+is invariant to adjustments to the zero eigenvalues, the decision rule in the
+original feature space is equivalent to a decision rule in a lower dimension,
+such that matrix inverses and determinants of relatively small matrices can be
+rapidly computed. Also, we provide an efficient algorithm along with pseudocode
+to estimate the \emph{HDRDA} classifier's tuning parameters in a grid search via
+cross-validation. Finally, we show that several shrinkage methods that are
+special cases of the \emph{HDRDA} classifier have no effect on the zero
+eigenvalues of the covariance-matrix estimators. Such methods include work from
 \cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and other methods studied by
@@
 
-We provide an efficient algorithm that expedites the model-selection process for
-the \emph{HDRDA} classifier when $p \gg N$. We then study the classification
-performance of the \emph{HDRDA} classifier on four real high-dimensional data
-sets along with a simulation design that generalizes the experiments initially
-conducted by \cite{Guo:2007te}. We demonstrate that the \emph{HDRDA} classifier
-often attains superior classification accuracy to several recent classifiers
-designed for small-sample, high-dimensional data from \cite*{Tong:2012hw},
-\cite{Witten:2011kc}, \cite*{Pang:2009ik}, and \cite{Guo:2007te}. We also
-include as a benchmark the random forest from \cite{Breiman:2001fb} because
-\cite*{FernandezDelgado:2014ul} has concluded that the random forest is often
-superior to other classifiers in benchmark studies.
-
-The remainder of this document is organized as follows. In Section
-\ref{sec:preliminaries} we present the discrimination problem and the necessary
+The remainder of this paper is organized as follows. In Section
+\ref{sec:preliminaries} we present the discrimination problem and necessary
 notation to describe our contributions. In Section \ref{sec:rda} we present the
 \emph{HDRDA} classifier along with its interpretation, properties, and
-computationally efficient model-selection procedure. Section \ref{sec:sims}
-describes our simulation studies of artificial and real data sets and also
+computationally efficient model-selection procedure. In Section \ref{sec:sims},
+we describe our simulation studies of artificial and real data sets and also
 describes the experimental results. We conclude with a brief discussion in
@@
 The smallest eigenvalues of $\widehat{\bm \Sigma}_k$ and the directions
-associated with their eigenvectors can highly influence \eqref{eq:qda}. In fact,
-the eigenvalues of $\widehat{ \bm \Sigma}_k$ are well-known to be biased if $p
-\ge n_k$ such that the smallest eigenvalues are underestimated
-\citep{Seber:2004uh}. Moreover, if $p > n_k$, then rank$(\widehat{\bm \Sigma}_k)
-\le n_k$, which implies that at least $p - n_k$ eigenvalues of $\widehat{\bm
-  \Sigma}_k$ are zero. Furthermore, although more feature information is
-available to discriminate among the $K$ classes, if $p > n_k$, \eqref{eq:qda} is
-incalculable because $\widehat{\bm \Sigma}_k^{-1}$ does not exist.
+associated with their eigenvectors can highly influence the classifier in
+\eqref{eq:qda}. In fact, the eigenvalues of $\widehat{ \bm \Sigma}_k$ are
+well-known to be biased if $p \ge n_k$ such that the smallest eigenvalues are
+underestimated \citep{Seber:2004uh}. Moreover, if $p > n_k$, then
+rank$(\widehat{\bm \Sigma}_k) \le n_k$, which implies that at least $p - n_k$
+eigenvalues of $\widehat{\bm \Sigma}_k$ are zero. Furthermore, although more
+feature information is available to discriminate among the $K$ classes, if $p >
+n_k$, \eqref{eq:qda} is incalculable because $\widehat{\bm \Sigma}_k^{-1}$ does
+not exist.
 
@@
 proposed in the literature to adjust the eigenvalues of $\widehat{\bm \Sigma}_k$
-so that \eqref{eq:qda} is calculable and has reduced variability in the standard
-bias-variance tradeoff. A common form of the covariance-matrix regularization
-applies a shrinkage factor $\gamma > 0$, so that
+so that \eqref{eq:qda} is calculable and provide reduced variability for
+$\widehat{\bm \Sigma}_k^{-1}$. A common form of the covariance-matrix regularization applies
+  a shrinkage factor $\gamma > 0$, so that
 \begin{align}
@@
 where $\lambda \in [0, 1]$ is the \emph{pooling} parameter. By rewriting
-\eqref{eq:sig-lambda-alternative} in terms of the observations $ \bm x_i$ $(i =
-1, \ldots, N)$, each centered by its class sample mean, we attain a clear
+\eqref{eq:sig-lambda-alternative} in terms of the observations $\bm x_i$, $i =
+1, \ldots, N$, each centered by its class sample mean, we attain a clear
 interpretation of $\widehat{\bm \Sigma}_k(\lambda)$. That is,
@@
 where $c_{ik}(\lambda) = \lambda N^{-1} + (1 - \lambda)n_k^{-1}I(y_i = k)$.
-From \eqref{eq:sig-lambda-alternative2}, $\lambda$ weights the contribution of
-each of the $N$ observations in estimating $\bm \Sigma_k$ from all $K$ classes
-rather than $n_k$ observations from a single class. As a result, we can
-interpret \eqref{eq:sig-lambda-alternative2} as a covariance-matrix estimator
-that borrows from $\widehat{\bm \Sigma}$ in \eqref{eq:pooled-cov} to estimate $
-\bm \Sigma_k$.
-
-Below, we will show that the pooling operation is advantageous in increasing the
-rank of each $\widehat{\bm \Sigma}_k(\lambda)$ from rank$(\widehat{\bm
-  \Sigma}_k)$ to rank$(\widehat{\bm \Sigma})$ for $0 < \lambda \le 1$. Notice
-that if $\lambda = 0$, then the observations from the remaining $K - 1$ classes
-do not contribute to the estimation of $\bm \Sigma_k$, corresponding to
-$\widehat{\bm \Sigma}_k$. Furthermore, if $\lambda = 1$, the weights
-$c_{ik}(\lambda)$ in \eqref{eq:sig-lambda-alternative2} reduce to $1/N$,
+From \eqref{eq:sig-lambda-alternative2}, we see that $\lambda$ weights the
+contribution of each of the $N$ observations in estimating $\bm \Sigma_k$ from
+all $K$ classes rather than using only the $n_k$ observations from a single
+class. As a result, we can interpret \eqref{eq:sig-lambda-alternative2} as a
+covariance-matrix estimator that borrows from $\widehat{\bm \Sigma}$ in
+\eqref{eq:pooled-cov} to estimate $ \bm \Sigma_k$.
+
+In Figure \ref{fig:hdrda-contours} we plot the contours of five multivariate
+normal populations for $\lambda = 0$ with unequal covariance matrices. As
+$\lambda$ approaches 1, the contours become more similar, resulting in identical
+contours for $\lambda = 1$. Below, we show that the pooling operation is
+advantageous in increasing the rank of each $\widehat{\bm \Sigma}_k(\lambda)$
+from rank$(\widehat{\bm \Sigma}_k)$ to rank$(\widehat{\bm \Sigma})$ for $0 <
+\lambda \le 1$. Notice that if $\lambda = 0$, then the observations from the
+remaining $K - 1$ classes do not contribute to the estimation of $\bm \Sigma_k$,
+corresponding to $\widehat{\bm \Sigma}_k$. Furthermore, if $\lambda = 1$, the
+weights $c_{ik}(\lambda)$ in \eqref{eq:sig-lambda-alternative2} reduce to $1/N$,
 corresponding to $\widehat{\bm \Sigma}$. For brevity, when $\lambda = 1$, we
 define $\bm X = [\sqrt{c_{1k}(1)} \bm x_1^{T}, \ldots, \sqrt{c_{Nk}(1)} \bm
-  x_N^{T}]^{T}$ such that $\widehat{\bm \Sigma} = N^{-1} \bm X^{T} \bm X$. Similarly,
-for $\lambda = 0$, we define $\bm X_k = [\sqrt{c_{1k}(0)} \bm x_1^{T}, \ldots,
-  \sqrt{c_{Nk}(0)} \bm x_N^{T}]^{T}$ such that $\widehat{\bm \Sigma}_k = n_k^{-1} \bm
-X_k^{T} \bm X_k$. In Figure \ref{fig:hdrda-contours} we plot the contours of
-five multivariate normal populations for $\lambda = 0$ with unequal covariance
-matrices. As $\lambda$ approaches 1, the contours become more similar, resulting
-in identical contours for $\lambda = 1$.
+  x_N^{T}]^{T}$ such that $\widehat{\bm \Sigma} = N^{-1} \bm X^{T} \bm
+X$. Similarly, for $\lambda = 0$, we define $\bm X_k = [\sqrt{c_{1k}(0)} \bm
+  x_1^{T}, \ldots, \sqrt{c_{Nk}(0)} \bm x_N^{T}]^{T}$ such that $\widehat{\bm
+  \Sigma}_k = n_k^{-1} \bm X_k^{T} \bm X_k$.
 
@@
 where $\alpha_k \ge 0$ and $\gamma \ge 0$ is an eigenvalue-shrinkage
-constant. Thus, the \emph{pooling} parameter $\lambda$ controls the amount
-borrowed from $\widehat{\bm \Sigma}$ to estimate $\bm \Sigma_k$, and the
-\emph{shrinkage} parameter $\gamma$ determines the degree of eigenvalue
-shrinkage. The choice of $\alpha_k$ allows for a flexible formulation of
-covariance-matrix estimators. For instance, if $\alpha_k = 1$, $k = 1, \ldots,
-K$, then \eqref{eq:hdrda-cov} resembles \eqref{eq:ridge-estimator}. Similarly,
-if $\alpha_k = 1 - \gamma$, then \eqref{eq:hdrda-cov} has a form similar to the
-\emph{RDA} classifier from \cite{Friedman:1989tm}. Substituting
-\eqref{eq:hdrda-cov} into \eqref{eq:qda}, we define the \emph{HDRDA} classifier
+constant. Thus, the \emph{pooling} parameter $\lambda$ controls the amount of
+estimation information borrowed from $\widehat{\bm \Sigma}$ to estimate $\bm
+\Sigma_k$, and the \emph{shrinkage} parameter $\gamma$ determines the degree of
+eigenvalue shrinkage. The choice of $\alpha_k$ allows for a flexible formulation
+of covariance-matrix estimators. For instance, if $\alpha_k = 1$, $k = 1,
+\ldots, K$, then \eqref{eq:hdrda-cov} resembles \eqref{eq:ridge-estimator}.
+Similarly, if $\alpha_k = 1 - \gamma$, then \eqref{eq:hdrda-cov} has a form
+comparable to the \emph{RDA} classifier from \cite{Friedman:1989tm}.
+Substituting \eqref{eq:hdrda-cov} into \eqref{eq:qda}, we define the
+\emph{HDRDA} classifier
 \begin{align}
@@
 from a grid of candidate models via cross-validation. We provide an
-implementation of \eqref{eq:hdrda} in the {\tt hdrda} function in the {\tt
-  sparsediscrim} R package, available on CRAN.
+implementation of \eqref{eq:hdrda} in the {\tt hdrda} function contained in the
+{\tt sparsediscrim} R package, which is available on CRAN.
 
@@
 via cross-validation, which is counterproductive to our goal of improving the
-\emph{RDA} classifier's efficiency. By setting $\alpha_k$ to an appropriate
-constant, the \emph{HDRDA} covariance-matrix estimator includes or resembles a
-large family of estimators. Notice that if $\alpha_k = 1$ and $\lambda = 1$,
-\eqref{eq:hdrda-cov} is equivalent to the standard ridge-like covariance-matrix
-estimator in \eqref{eq:ridge-estimator}, and various estimators proposed in the
-literature can be obtained by selecting $\gamma$ accordingly. For instance, with
-$\gamma = \tr\{ \widehat{\bm \Sigma} \} / \min(N, p)$, we obtain the estimator
-from \cite{Srivastava:2007ww}.
+\emph{RDA} classifier's computational efficiency. By setting $\alpha_k$ to an
+appropriate constant, the \emph{HDRDA} covariance-matrix estimator includes or
+resembles a large family of estimators. Notice that if $\alpha_k = 1$ and
+$\lambda = 1$, \eqref{eq:hdrda-cov} is equivalent to the standard ridge-like
+covariance-matrix estimator in \eqref{eq:ridge-estimator}, and various
+estimators proposed in the literature can be obtained by selecting $\gamma$
+accordingly. For instance, with $\gamma = \tr\{ \widehat{\bm \Sigma} \} /
+\min(N, p)$, we obtain the estimator from \cite{Srivastava:2007ww}.
 
@@
 \eqref{eq:pooled-cov}. To further improve the estimation of $\bm \Sigma_k$ and
-to stabilize the inverse of \eqref{eq:sig-lambda}, \cite{Friedman:1989tm} has
+to stabilize the inverse of \eqref{eq:sig-lambda}, \cite{Friedman:1989tm}
 proposed the biased covariance-matrix estimator
@@
 
-Our proposed \emph{HDRDA} classifier resembles \eqref{eq:sig-rda} but improves
-the interpretation of the pooling parameter $\lambda$ as discussed
-above. Moreover, as we discuss below, our formulation of \emph{HDRDA} classifier
-lends itself to rapid estimation in high-dimensional classification.
+Our proposed \emph{HDRDA} classifier resembles \eqref{eq:sig-rda} but yields an
+interpretation of the pooling parameter $\lambda$ as discussed above. Moreover,
+as we discuss below, our formulation of \emph{HDRDA} classifier lends itself to
+rapid estimation in high-dimensional classification.
 
@@
 As we can see, we are able to avoid the time-consuming inverses and determinants
-of $p \times p$ covariance matrices in \eqref{eq:hdrda} and instead calculating
+of $p \times p$ covariance matrices in \eqref{eq:hdrda} and instead calculate
 these same operations on $\bm W_k \in \mathbb{R}_{q \times q}$ in
@@
 the inverse of the diagonal matrix $\bm \Gamma_k^{-1} \in \mathbb{R}_{q \times
-  q}$ requires $O(q)$ operations. Finally, $| \bm W_k |$ requires $O(n_k^3)$
-operations.
+  q}$ requires $O(q)$ operations. Finally, we remark that $| \bm W_k |$ requires
+$O(n_k^3)$ operations.
 
@@
 \citep{Hastie:2008dt}. For the second model, we set $\alpha_k = 1 - \gamma$, $k
-= 1, \ldots, K$, to resemble \citeauthor{Friedman:1989tm}'s parameterization. We
-estimated both $\lambda$ and $\gamma$ from a grid of 21 equidistant candidate
-values between 0 and 1, inclusively.
+= 1, \ldots, K$, to resemble \citeauthor{Friedman:1989tm}'s parameterization,
+and we estimated both $\lambda$ and $\gamma$ from a grid of 21 equidistant
+candidate values between 0 and 1, inclusively.
 
@@
 design from \cite{Guo:2007te}. As $\epsilon$ is increased, the average number of
-outliers is increased. In our simulation, we set $\eta = 100$ and considered the
+outliers is increased. In our simulation, we let $\eta = 100$ and considered the
 values of $\epsilon = 0$, $0.05$, $\ldots$, $0.50$.
@@
 small sample size. \cite{Guo:2007te} originally considered $\rho_k = 0.9$ for
-all $k$. To explore the more realistic assumption of unequal covariance
-matrices, we instead put $\rho_1 = 0.1$, $\rho_2 = 0.5$, and $\rho_3 = 0.9$
+all $k$. Alternatively, to explore the more realistic assumption of unequal
+covariance matrices, we put $\rho_1 = 0.1$, $\rho_2 = 0.5$, and $\rho_3 = 0.9$
 
@@
 approximately 43\% when $\epsilon=0.05$. The sharp increase in average error
-rates for these three classifiers continued as $\epsilon$ increased
-further. Guo's method always outperformed Pang, Witten, and Tong, but once
-outliers were present, the Guo classifier's average error rate was competitive
-with neither the \emph{HDRDA} classifiers nor the random forest.
+rates for these three classifiers continued as $\epsilon$ increased. Guo's
+method always outperformed Pang, Witten, and Tong, but once outliers were
+present, the Guo classifier's average error rate was competitive with neither
+the \emph{HDRDA} classifiers nor the random forest.
 
@@
 
-We compared the \emph{HDRDA} classifer to the five competing classifiers on four
+We compared the \emph{HDRDA} classifer to the five competing classifiers on six
 benchmark gene-expression microarray data sets. We evaluated the classification
@@
 
-Approximately 30-40\% of adult non-Hodgkin lymphomas are diffuse large B-cell
+Approximately 30\%-40\% of adult non-Hodgkin lymphomas are diffuse large B-cell
 lymphomas (DLBCLs) \cite{Shipp:2002ka}.  However, only a small proportion of
@@
 forest's accuracy slightly exceeded the \emph{HDRDA} classifiers on the Tian
-data set, our proposed methods outperformed the other classifiers considered.
-Additionally, the \emph{HDRDA} classifiers had comparable performance on five of
-the six data sets.
+data set, our proposed classifiers outperformed the other competing classifiers
+considered here. Moreover, the \emph{HDRDA} classifiers had comparable
+performance on five of the six data sets.
 

### 2016-01-15 reports/paper-hdrda.Rnw
After all this time, never defined n_k

@@
 x_{ip}) \in \mathbb{R}_{p \times 1}$ be the $i$th observation $(i = 1, \ldots,
-N)$ with true, unique membership $y_i \in \{\omega_1, \ldots, \omega_K\}$. We
-assume that $(\bm x_i, y_i)$ is a realization from a mixture distribution $p(\bm
-x) = \sum_{k=1}^K p(\bm x | \omega_k) p(\omega_k)$, where $p(\bm x | \omega_k)$
-is the probability density function (PDF) of the $k$th class and $p(\omega_k)$
-is the prior probability of class membership of the $k$th class. We further
-assume $p(\omega_k) = p(\omega_l)$, $1 \le k, l \le K$, $k \ne l$.
+N)$ with true, unique membership $y_i \in \{\omega_1, \ldots, \omega_K\}$.
+Denote by $n_k$ the number of training observations realized from class $k$,
+such that $\sum_{k=1}^K n_k = N$. We assume that $(\bm x_i, y_i)$ is a
+realization from a mixture distribution $p(\bm x) = \sum_{k=1}^K p(\bm x |
+\omega_k) p(\omega_k)$, where $p(\bm x | \omega_k)$ is the probability density
+function (PDF) of the $k$th class and $p(\omega_k)$ is the prior probability of
+class membership of the $k$th class. We further assume $p(\omega_k) =
+p(\omega_l)$, $1 \le k, l \le K$, $k \ne l$.
 

### 2016-01-20 reports/paper-hdrda.Rnw
Applied a couple of edits from Caleb.

@@
 \emph{RDA} classifier to be employed with small-sample, high-dimensional data in
-practice. This results is reassuring because the \emph{RDA} classifier remains
+practice. This result is reassuring because the \emph{RDA} classifier remains
 widely popular in the literature. In fact, variants of the RDA classifier have
@@
 The dimension reduction that we have employed in this paper has reduced the
-dimension to rank$(\widehat{\bm \Sigma}) = q$. An interesting extension of
-our work would reduce the dimension $q$ further to a lower dimension $q_L < q$
-using some criterion proposed, perhaps similar to that employed in principal
+dimension to rank$(\widehat{\bm \Sigma}) = q$. An interesting extension of our
+work would reduce the dimension $q$ further to a lower dimension $q_L < q$ using
+some criterion proposed, perhaps similar to that employed in principal
 components analysis. While it is not clear whether classification performance
@@
 selection would certainly improve. Moreover, if $q_L = 2$ or $3$ was selected,
-the potential for low-dimensional graphical displays of high-dimensional could
-be accomplished.
+the potential for low-dimensional graphical displays of high-dimensional data
+could be accomplished.
 

### 2016-01-20 reports/paper-hdrda.Rnw
Added some DMY edits.

@@
 estimated via a computationally burdensome cross-validation procedure using a
-grid search. We formulate a new RDA classifier for the small-sample,
+grid search. We formulate a new RDA-based classifier for the small-sample,
 high-dimensional setting and then show that the classification decision rule is
@@
 reduction in computation during model selection. Also, our parameterization
-offers interpretability previously lacking with the RDA classifier. We
+offers interpretability that was previously lacking with the RDA classifier. We
 demonstrate that our proposed classifier is often superior to several recently
@@
 that their inherent dimension reduction better captured the classificatory
-information in the small samples even in the presence of outliers.
+information in the small training samples, even in the presence of outliers.
 
@@
 method always outperformed Pang, Witten, and Tong, but once outliers were
-present, the Guo classifier's average error rate was competitive with neither
-the \emph{HDRDA} classifiers nor the random forest.
+present, the Guo classifier's average error rate was not competitive with
+the \emph{HDRDA} classifiers or the random-forest classifier.
 
@@
 For $\epsilon > 0$, we observed a different behavior in classification
-performance. First, the Pang, Tong, and Witten methods along with the random
-forest increased in average error rate as $p$ increased. Contrarily, the
-performance of the \emph{HDRDA} and Guo classifiers was hardly affected by $p$.
-Also, as discussed above, the \emph{HDRDA} classifiers were superior to all
-other classifiers for larger values of $p$ with only the random forest
-outperforming them for smaller feature dimensions.
+performance. First, the Pang, Tong, and Witten methods, along with the
+random-forest method increased in average error rate as $p$
+increased. Contrarily, the performance of the \emph{HDRDA} and Guo classifiers
+was hardly affected by $p$.  Also, as discussed above, the \emph{HDRDA}
+classifiers were superior to all other classifiers for large values of $p$ with
+only the random-forest classifier outperforming them in smaller
+feature-dimension cases.
 
@@
 process 100 times, we computed the average of the error-rate estimates for each
-classifier. Here, we provide a concise description of each high-dimensional data
+classifier. We next provide a concise description of each high-dimensional data
 set examined in our classification study.
@@
 The average error-rate estimates for the Pang, Tong, and Witten classifiers were
-comparable across all data sets. Furthermore, the average error rates for the
+comparable across all six data sets. Furthermore, the average error rates for the
 Pang and Tong classifiers were approximately equal for all data sets except for
@@
 Finally, the Pang, Tong, and Witten classifiers consistently yielded the largest
-error rates across all data sets. Given that the standard deviations were
-relatively large, we hesitate to generalize claims regarding the ranking of
-these three classifiers in terms of error rate. However, the classifiers' error
-rates and their variability across multiple random partitions of each data set
-were large enough to rethink applying them in practice.
+average error rates across the six data sets. Given that the standard deviations
+were relatively large, we hesitate to generalize claims regarding the ranking of
+these three classifiers in terms of the average error rate. However, the
+classifiers' error rates and their variability across multiple random partitions
+of each data set were large enough that we might question their benefit when
+applied to real data.
 

### 2016-01-20 reports/paper-hdrda.Rnw
Figures and Tables moved to end. Placeholders in text.

@@
 
-\begin{figure}
-<<hdrda_contours, fig.width=8, fig.height=8, out.width='.33\\linewidth', dpi=200, fig.show='hold'>>=
-rda_contours(lambda = 0)
-rda_contours(lambda = 0.2)
-rda_contours(lambda = 0.4)
-rda_contours(lambda = 0.6)
-rda_contours(lambda = 0.8)
-rda_contours(lambda = 1)
-@
-\caption{Contours of five multivariate normal populations as a function of the
-  pooling parameter $\lambda$.}
-\label{fig:hdrda-contours}
-\end{figure}
+\[ \left[\text{Insert Figure \ref{fig:hdrda-contours} approximately here }\right] \]
 
@@
 
+\[ \left[\text{Insert Figure \ref{fig:sim-results} approximately here }\right] \]
+
+The Pang, Tong, and Witten methods yielded practically the same and consistently
+the worst error rates when outliers were present with $\epsilon > 0$, suggesting
+that these classifiers were sensitive to outliers. Notice, for example, that
+when $p=400$, the error rates of the Pang, Tong, and Witten classifiers
+increased dramatically from approximately 19\% when no outliers were present to
+approximately 43\% when $\epsilon=0.05$. The sharp increase in average error
+rates for these three classifiers continued as $\epsilon$ increased. Guo's
+method always outperformed Pang, Witten, and Tong, but once outliers were
+present, the Guo classifier's average error rate was not competitive with
+the \emph{HDRDA} classifiers or the random-forest classifier.
+
+\[ \left[\text{Insert Figure \ref{fig:sim-results-by-p} approximately here }\right] \]
+
+In Figure \ref{fig:sim-results-by-p} we again examine the simulation results as
+a function of the number of features $p$ for a subset of the values of
+$\epsilon$. This set of plots allows us to investigate the effect of feature
+dimensionality on classification performance. When no outliers were present
+(i.e., $\epsilon=0$), the random-forest classifier was outperformed by all other
+classifiers. Furthermore, the \emph{HDRDA} classifiers were superior in terms of
+average error rate in this setting. As $p$ increased, an elevation in average
+error rate was expected for all classifiers, but the increase was not observed
+to be substantial.
+
+For $\epsilon > 0$, we observed a different behavior in classification
+performance. First, the Pang, Tong, and Witten methods, along with the
+random-forest method increased in average error rate as $p$ increased.
+Contrarily, the performance of the \emph{HDRDA} and Guo classifiers was hardly
+affected by $p$. Also, as discussed above, the \emph{HDRDA} classifiers were
+superior to all other classifiers for large values of $p$ with only the
+random-forest classifier outperforming them in smaller feature-dimension cases.
+
+\subsection{Application to Gene Expression Data}
+
+We compared the \emph{HDRDA} classifer to the five competing classifiers on six
+benchmark gene-expression microarray data sets. We evaluated the classification
+accuracy of each classifier by randomly partitioning the data set under
+consideration such that $2/3$ of the observations were allocated as training
+data and the remaining $1/3$ of the observations were allocated as a test data
+set. To expedite the computational runtime, we reduced the training data to the
+top 1000 variables by employing the variable-selection method proposed by
+\cite{Dudoit:2002ev}. We then reduced the test data set to the same 1000
+variables. After training each classifier on the training data, we classified
+the test data sets and computed the proportion of mislabeled test observations
+to estimate the classification error rate for each classifier. Repeating this
+process 100 times, we computed the average of the error-rate estimates for each
+classifier. We next provide a concise description of each high-dimensional data
+set examined in our classification study.
+
+\subsubsection{\cite{Chiaretti:2004gq} Data Set}
+
+\cite{Chiaretti:2004gq} measured the gene-expression profiles for 128
+individuals with acute lymphoblastic leukaemia (ALL) using Affymetrix human
+95Av2 arrays. Following \cite{Xu:2009fl}, we restricted the data set to $K = 2$
+classes such that $n_1 = 74$ observations were without cytogenetic abnormalities
+and $n_2 = 37$ observations have a detected BCR/ABL gene. The robust multichip
+average normalization method was applied to all 12,625 gene expression levels.
+
+\subsubsection{\cite{Chowdary:2006kf} Data Set}
+
+\cite{Chowdary:2006kf} investigated 52 matched pairs of tissues from colon and
+breast tumors using Affymetrix U133A arrays and ribonucleic-acid (RNA)
+amplification. Each tissue pair was gathered from the same patient and
+consisted of a snap-frozen tissue and a tissue suspended in a RNAlater
+preservative. Overall, 31 breast-cancer and 21 colon-cancer pairs were gathered,
+resulting in $K = 2$ classes with $n_1 = 62$ and $n_2 = 42$. A purpose of the
+study was to determine whether the disease state could be identified using
+22,283 gene-expression profiles.
+
+\subsubsection{\cite{Nakayama:2007fl} Data Set}
+
+\cite{Nakayama:2007fl} acquired 105 gene-expression samples of 10 types of soft
+tissue tumors through an oligonucleotide microarray, including 16 samples of
+synovial sarcoma (SS), 19 samples of myxoid/round cell liposarcoma (MLS), 3
+samples of lipoma, 3 samples of well-differentiated liposarcoma (WDLS), 15
+samples of dedifferentiated liposarcoma (DDLS), 15 samples of myxofibrosarcoma
+(MFS), 6 samples of leiomyosarcoma (LMS), 3 samples of malignant nerve sheathe
+tumor (MPNST), 4 samples of fibrosarcoma (FS), and 21 samples of malignant
+fibrous histiocytoma (MFH). \cite{Nakayama:2007fl} determined from their data
+that these 10 types fell into 4 broader groups: (1) SS; (2) MLS; (3) Lipoma,
+WDLS, and part of DDLS; (4) Spindle cell and pleomorophic sarcomas including
+DDLS, MFS, LMS, MPNST, FS, and MFH. Following \cite{Witten:2011kc}, we restrict
+our analysis to the five tumor types having at least 15 observations.
+
+\subsubsection{\cite{Shipp:2002ka} Data Set}
+
+Approximately 30\%-40\% of adult non-Hodgkin lymphomas are diffuse large B-cell
+lymphomas (DLBCLs) \cite{Shipp:2002ka}.  However, only a small proportion of
+DLBCL patients are cured with modern chemotherapeutic regimens.  Several models
+have been proposed, such as the International Prognostic Index (IPI), to
+determine a patient's curability.  These models rely on clinical covariates,
+such as age, to determine if the patient can be cured and are often
+ineffective. \cite{Shipp:2002ka} have argued that more effective means are
+desired to determine a patient's curability. The authors measured 6,817 gene
+expression levels from 58 DLBCL patient samples with customized cDNA
+(lymphochip) microarrays to investigate the curability of patients treated with
+cyclophosphamide, adriamycin, vincristine, and prednisone (CHOP)-based
+chemotherapy. Among the 58 DLBCL patient samples, 32 are from cured patients
+while 26 are from patients with fatal or refractory disease.
+
+\subsubsection{\cite{Singh:2002fh} Data Set}
+
+\cite{Singh:2002fh} have examined 235 radical prostatectomy specimens from
+surgery patients between 1995 and 1997. The authors used oligonucleotide
+microarrays containing probes for approximately 12,600 genes and expressed
+sequence tags. They have reported that 102 of the radical prostatectomy
+specimens are of high quality: 52 prostate tumor samples and 50 non-tumor
+prostate samples.
+
+\subsubsection{\cite{Tian:2003ht} Data Set}
+
+\cite{Tian:2003ht} investigated the purified plasma cells from the bone marrow
+of control patients along with patients with newly diagnosed multiple
+myeloma. Expression profiles for 12,2625 genes were obtained via Affymetrix
+U95Av2 microarrays. The plasma cells were subjected to biochemical and
+immunohistochemical analyses to identify molecular determinants of osteolytic
+lesions. For 36 multiple-myloma patients, focal bone lesions could not be
+detected by magnetic resonance imaging (MRI), whereas MRI was used to detect
+such lesions in 137 patients.
+
+\subsubsection{Classification Results}
+
+Similar to \cite{Witten:2011kc}, we report the average test error rates obtained
+over 100 random training-test partitions in Table \ref{tab:microarray-results}
+along with standard deviations of the test error rates in parentheses. The
+\emph{HDRDA} and Guo classifiers were superior in classification performance for
+the majority of the simulations. The \emph{HDRDA} classifiers had the best
+classification accuracy on the Chowdary and Shipp data sets. Although the random
+forest's accuracy slightly exceeded the \emph{HDRDA} classifiers on the Tian
+data set, our proposed classifiers outperformed the other competing classifiers
+considered here. Moreover, the \emph{HDRDA} classifiers had comparable
+performance on five of the six data sets.
+
+\[ \left[\text{Insert Table \ref{tab:microarray-results} approximately here }\right] \]
+
+The average error-rate estimates for the Pang, Tong, and Witten classifiers were
+comparable across all six data sets. Furthermore, the average error rates for the
+Pang and Tong classifiers were approximately equal for all data sets except for
+the Chiaretti dataset. This result suggests that the mean and variance
+estimators used in lieu of the MLEs provided little improvement to
+classification accuracies. However, we investigated the Pang classifier's poor
+performance on the Chiaretti data set and determined that its variance estimator
+exhibited numerical instability. The classifier's denominator was approximately
+zero for both classes and lead to the poor classification performance.
+
+The random-forest classifier was competitive when applied to the Chowdary and
+Singh data sets and yielded the smallest error rate of the considered
+classifiers on the Tian data set. The fact that the \emph{HDRDA} and Guo
+classifiers typically outperformed the random-forest classifier challenges the
+claim of \cite{FernandezDelgado:2014ul} that the random forest is typically
+superior. Further studies should be performed to validate this statement in the
+small-sample, high-dimensional setting.
+
+Finally, the Pang, Tong, and Witten classifiers consistently yielded the largest
+average error rates across the six data sets. Given that the standard deviations
+were relatively large, we hesitate to generalize claims regarding the ranking of
+these three classifiers in terms of the average error rate. However, the
+classifiers' error rates and their variability across multiple random partitions
+of each data set were large enough that we might question their benefit when
+applied to real data.
+
+\section{Discussion}
+\label{sec:discussion}
+
+We have compared our proposed \emph{HDRDA} classifier to the random-forest and
+Witten classifiers as well as the Pang, Tong, and Guo classifiers, each of which
+assumes independence of the features conditional on class membership. Despite
+the rapid computational performance of diagonal classifiers and the reduction in
+the number of parameters to estimate, \cite{Mai:2012bf} and \cite{Fan:2012iq}
+have noted that diagonal classifiers can often yield inferior classification
+performance to other classification methods. We have shown that the \emph{HDRDA}
+classifier often yields superior classification accuracy to the diagonal
+classifiers considered when applied to small-sample, high-dimensional data sets,
+confirming the assertions of \cite{Mai:2012bf} and \cite{Fan:2012iq}.
+Furthermore, we have demonstrated that \emph{HDRDA} classifiers are more robust
+to the presence of outliers than the diagonal classifiers.
+
+We also considered the popular penalized linear discriminant analysis from
+\cite{Witten:2011kc} because it was specifically designed for high-dimensional
+gene-expression data. We had expected its classification performance to be
+competitive within our classification study and perhaps superior. Contrarily,
+its underperformance in our empirical studies show that it is sensitive to
+outliers and unable to achieve comparable results with other classifiers
+designed for small-sample, high-dimensional data. Also, despite the claims of
+\cite{FernandezDelgado:2014ul} that the random forest classifier is typically
+superior to other classifiers, we observed that the random-forest classifier was
+indeed competitive but was typically outperformed by classifiers developed for
+small-sample, high-dimensional data.
+
+The \emph{HDRDA} implementation of \eqref{eq:hdrda-decomposed} in the {\tt
+  sparsediscrim} R package was comparable to that of the diagonal classifiers in
+terms of computational runtime. Given the \emph{RDA} classifier from
+\cite{Friedman:1989tm} has been shown to have excellent performance in the
+high-dimensional setting \citep{Webb:2011vu} but is limited by its
+computationally intense model selection procedure, our work enables the
+\emph{RDA} classifier to be employed with small-sample, high-dimensional data in
+practice. This result is reassuring because the \emph{RDA} classifier remains
+widely popular in the literature. In fact, variants of the RDA classifier have
+been applied to microarray data
+\citep{Ching:2012fu,Li:2012ev,Tai:2007bk,Guo:2007te}, facial recognition
+\citep{Ching:2012fu,Zhang:2010va,Dai:2007vd,Lu:2005hq,Pima:2004vw,Lu:2003we},
+handwritten digit recognition \citep{Bouveyron:2007gx}, remote sensing
+\citep{Tadjudin:1999fk}, seismic detection \citep{Anderson:2002kg}, and chemical
+spectra \citep*{Wu:1996us,Aeberhard:1993fp}.
+
+The dimension reduction that we have employed in this paper has reduced the
+dimension to rank$(\widehat{\bm \Sigma}) = q$. An interesting extension of our
+work would reduce the dimension $q$ further to a lower dimension $q_L < q$ using
+some criterion proposed, perhaps similar to that employed in principal
+components analysis. While it is not clear whether classification performance
+would improve via such a method, the efficiency of the training and model
+selection would certainly improve. Moreover, if $q_L = 2$ or $3$ was selected,
+the potential for low-dimensional graphical displays of high-dimensional data
+could be accomplished.
+
+\bibliographystyle{plainnat}
+\bibliography{rda}
+
+\clearpage
+
+\begin{figure}
+<<hdrda_contours, fig.width=8, fig.height=8, out.width='.33\\linewidth', dpi=200, fig.show='hold'>>=
+rda_contours(lambda = 0)
+rda_contours(lambda = 0.2)
+rda_contours(lambda = 0.4)
+rda_contours(lambda = 0.6)
+rda_contours(lambda = 0.8)
+rda_contours(lambda = 1)
+@
+\caption{Contours of five multivariate normal populations as a function of the
+  pooling parameter $\lambda$.}
+\label{fig:hdrda-contours}
+\end{figure}
+
+\clearpage
+
 \begin{figure}
@@
 
-The Pang, Tong, and Witten methods yielded practically the same and consistently
-the worst error rates when outliers were present with $\epsilon > 0$, suggesting
-that these classifiers were sensitive to outliers. Notice, for example, that
-when $p=400$, the error rates of the Pang, Tong, and Witten classifiers
-increased dramatically from approximately 19\% when no outliers were present to
-approximately 43\% when $\epsilon=0.05$. The sharp increase in average error
-rates for these three classifiers continued as $\epsilon$ increased. Guo's
-method always outperformed Pang, Witten, and Tong, but once outliers were
-present, the Guo classifier's average error rate was not competitive with
-the \emph{HDRDA} classifiers or the random-forest classifier.
+\clearpage
 
@@
 
-In Figure \ref{fig:sim-results-by-p} we again examine the simulation results as
-a function of the number of features $p$ for a subset of the values of
-$\epsilon$. This set of plots allows us to investigate the effect of feature
-dimensionality on classification performance. When no outliers were present
-(i.e., $\epsilon=0$), the random-forest classifier was outperformed by all other
-classifiers. Furthermore, the \emph{HDRDA} classifiers were superior in terms of
-average error rate in this setting. As $p$ increased, an elevation in average
-error rate was expected for all classifiers, but the increase was not observed
-to be substantial.
-
-For $\epsilon > 0$, we observed a different behavior in classification
-performance. First, the Pang, Tong, and Witten methods, along with the
-random-forest method increased in average error rate as $p$
-increased. Contrarily, the performance of the \emph{HDRDA} and Guo classifiers
-was hardly affected by $p$.  Also, as discussed above, the \emph{HDRDA}
-classifiers were superior to all other classifiers for large values of $p$ with
-only the random-forest classifier outperforming them in smaller
-feature-dimension cases.
-
-\subsection{Application to Gene Expression Data}
-
-We compared the \emph{HDRDA} classifer to the five competing classifiers on six
-benchmark gene-expression microarray data sets. We evaluated the classification
-accuracy of each classifier by randomly partitioning the data set under
-consideration such that $2/3$ of the observations were allocated as training
-data and the remaining $1/3$ of the observations were allocated as a test data
-set. To expedite the computational runtime, we reduced the training data to the
-top 1000 variables by employing the variable-selection method proposed by
-\cite{Dudoit:2002ev}. We then reduced the test data set to the same 1000
-variables. After training each classifier on the training data, we classified
-the test data sets and computed the proportion of mislabeled test observations
-to estimate the classification error rate for each classifier. Repeating this
-process 100 times, we computed the average of the error-rate estimates for each
-classifier. We next provide a concise description of each high-dimensional data
-set examined in our classification study.
-
-\subsubsection{\cite{Chiaretti:2004gq} Data Set}
-
-\cite{Chiaretti:2004gq} measured the gene-expression profiles for 128
-individuals with acute lymphoblastic leukaemia (ALL) using Affymetrix human
-95Av2 arrays. Following \cite{Xu:2009fl}, we restricted the data set to $K = 2$
-classes such that $n_1 = 74$ observations were without cytogenetic abnormalities
-and $n_2 = 37$ observations have a detected BCR/ABL gene. The robust multichip
-average normalization method was applied to all 12,625 gene expression levels.
-
-\subsubsection{\cite{Chowdary:2006kf} Data Set}
-
-\cite{Chowdary:2006kf} investigated 52 matched pairs of tissues from colon and
-breast tumors using Affymetrix U133A arrays and ribonucleic-acid (RNA)
-amplification. Each tissue pair was gathered from the same patient and
-consisted of a snap-frozen tissue and a tissue suspended in a RNAlater
-preservative. Overall, 31 breast-cancer and 21 colon-cancer pairs were gathered,
-resulting in $K = 2$ classes with $n_1 = 62$ and $n_2 = 42$. A purpose of the
-study was to determine whether the disease state could be identified using
-22,283 gene-expression profiles.
-
-\subsubsection{\cite{Nakayama:2007fl} Data Set}
-
-\cite{Nakayama:2007fl} acquired 105 gene-expression samples of 10 types of soft
-tissue tumors through an oligonucleotide microarray, including 16 samples of
-synovial sarcoma (SS), 19 samples of myxoid/round cell liposarcoma (MLS), 3
-samples of lipoma, 3 samples of well-differentiated liposarcoma (WDLS), 15
-samples of dedifferentiated liposarcoma (DDLS), 15 samples of myxofibrosarcoma
-(MFS), 6 samples of leiomyosarcoma (LMS), 3 samples of malignant nerve sheathe
-tumor (MPNST), 4 samples of fibrosarcoma (FS), and 21 samples of malignant
-fibrous histiocytoma (MFH). \cite{Nakayama:2007fl} determined from their data
-that these 10 types fell into 4 broader groups: (1) SS; (2) MLS; (3) Lipoma,
-WDLS, and part of DDLS; (4) Spindle cell and pleomorophic sarcomas including
-DDLS, MFS, LMS, MPNST, FS, and MFH. Following \cite{Witten:2011kc}, we restrict
-our analysis to the five tumor types having at least 15 observations.
-
-\subsubsection{\cite{Shipp:2002ka} Data Set}
-
-Approximately 30\%-40\% of adult non-Hodgkin lymphomas are diffuse large B-cell
-lymphomas (DLBCLs) \cite{Shipp:2002ka}.  However, only a small proportion of
-DLBCL patients are cured with modern chemotherapeutic regimens.  Several models
-have been proposed, such as the International Prognostic Index (IPI), to
-determine a patient's curability.  These models rely on clinical covariates,
-such as age, to determine if the patient can be cured and are often
-ineffective. \cite{Shipp:2002ka} have argued that more effective means are
-desired to determine a patient's curability. The authors measured 6,817 gene
-expression levels from 58 DLBCL patient samples with customized cDNA
-(lymphochip) microarrays to investigate the curability of patients treated with
-cyclophosphamide, adriamycin, vincristine, and prednisone (CHOP)-based
-chemotherapy. Among the 58 DLBCL patient samples, 32 are from cured patients
-while 26 are from patients with fatal or refractory disease.
-
-\subsubsection{\cite{Singh:2002fh} Data Set}
-
-\cite{Singh:2002fh} have examined 235 radical prostatectomy specimens from
-surgery patients between 1995 and 1997. The authors used oligonucleotide
-microarrays containing probes for approximately 12,600 genes and expressed
-sequence tags. They have reported that 102 of the radical prostatectomy
-specimens are of high quality: 52 prostate tumor samples and 50 non-tumor
-prostate samples.
-
-\subsubsection{\cite{Tian:2003ht} Data Set}
-
-\cite{Tian:2003ht} investigated the purified plasma cells from the bone marrow
-of control patients along with patients with newly diagnosed multiple
-myeloma. Expression profiles for 12,2625 genes were obtained via Affymetrix
-U95Av2 microarrays. The plasma cells were subjected to biochemical and
-immunohistochemical analyses to identify molecular determinants of osteolytic
-lesions. For 36 multiple-myloma patients, focal bone lesions could not be
-detected by magnetic resonance imaging (MRI), whereas MRI was used to detect
-such lesions in 137 patients.
-
-\subsubsection{Classification Results}
-
-Similar to \cite{Witten:2011kc}, we report the average test error rates obtained
-over 100 random training-test partitions in Table \ref{tab:microarray-results}
-along with standard deviations of the test error rates in parentheses. The
-\emph{HDRDA} and Guo classifiers were superior in classification performance for
-the majority of the simulations. The \emph{HDRDA} classifiers had the best
-classification accuracy on the Chowdary and Shipp data sets. Although the random
-forest's accuracy slightly exceeded the \emph{HDRDA} classifiers on the Tian
-data set, our proposed classifiers outperformed the other competing classifiers
-considered here. Moreover, the \emph{HDRDA} classifiers had comparable
-performance on five of the six data sets.
+\clearpage
 
@@
 
-The average error-rate estimates for the Pang, Tong, and Witten classifiers were
-comparable across all six data sets. Furthermore, the average error rates for the
-Pang and Tong classifiers were approximately equal for all data sets except for
-the Chiaretti dataset. This result suggests that the mean and variance
-estimators used in lieu of the MLEs provided little improvement to
-classification accuracies. However, we investigated the Pang classifier's poor
-performance on the Chiaretti data set and determined that its variance estimator
-exhibited numerical instability. The classifier's denominator was approximately
-zero for both classes and lead to the poor classification performance.
-
-The random-forest classifier was competitive when applied to the Chowdary and
-Singh data sets and yielded the smallest error rate of the considered
-classifiers on the Tian data set. The fact that the \emph{HDRDA} and Guo
-classifiers typically outperformed the random-forest classifier challenges the
-claim of \cite{FernandezDelgado:2014ul} that the random forest is typically
-superior. Further studies should be performed to validate this statement in the
-small-sample, high-dimensional setting.
-
-Finally, the Pang, Tong, and Witten classifiers consistently yielded the largest
-average error rates across the six data sets. Given that the standard deviations
-were relatively large, we hesitate to generalize claims regarding the ranking of
-these three classifiers in terms of the average error rate. However, the
-classifiers' error rates and their variability across multiple random partitions
-of each data set were large enough that we might question their benefit when
-applied to real data.
-
-\section{Discussion}
-\label{sec:discussion}
-
-We have compared our proposed \emph{HDRDA} classifier to the random-forest and
-Witten classifiers as well as the Pang, Tong, and Guo classifiers, each of which
-assumes independence of the features conditional on class membership. Despite
-the rapid computational performance of diagonal classifiers and the reduction in
-the number of parameters to estimate, \cite{Mai:2012bf} and \cite{Fan:2012iq}
-have noted that diagonal classifiers can often yield inferior classification
-performance to other classification methods. We have shown that the \emph{HDRDA}
-classifier often yields superior classification accuracy to the diagonal
-classifiers considered when applied to small-sample, high-dimensional data sets,
-confirming the assertions of \cite{Mai:2012bf} and \cite{Fan:2012iq}.
-Furthermore, we have demonstrated that \emph{HDRDA} classifiers are more robust
-to the presence of outliers than the diagonal classifiers.
-
-We also considered the popular penalized linear discriminant analysis from
-\cite{Witten:2011kc} because it was specifically designed for high-dimensional
-gene-expression data. We had expected its classification performance to be
-competitive within our classification study and perhaps superior. Contrarily,
-its underperformance in our empirical studies show that it is sensitive to
-outliers and unable to achieve comparable results with other classifiers
-designed for small-sample, high-dimensional data. Also, despite the claims of
-\cite{FernandezDelgado:2014ul} that the random forest classifier is typically
-superior to other classifiers, we observed that the random-forest classifier was
-indeed competitive but was typically outperformed by classifiers developed for
-small-sample, high-dimensional data.
-
-The \emph{HDRDA} implementation of \eqref{eq:hdrda-decomposed} in the {\tt
-  sparsediscrim} R package was comparable to that of the diagonal classifiers in
-terms of computational runtime. Given the \emph{RDA} classifier from
-\cite{Friedman:1989tm} has been shown to have excellent performance in the
-high-dimensional setting \citep{Webb:2011vu} but is limited by its
-computationally intense model selection procedure, our work enables the
-\emph{RDA} classifier to be employed with small-sample, high-dimensional data in
-practice. This result is reassuring because the \emph{RDA} classifier remains
-widely popular in the literature. In fact, variants of the RDA classifier have
-been applied to microarray data
-\citep{Ching:2012fu,Li:2012ev,Tai:2007bk,Guo:2007te}, facial recognition
-\citep{Ching:2012fu,Zhang:2010va,Dai:2007vd,Lu:2005hq,Pima:2004vw,Lu:2003we},
-handwritten digit recognition \citep{Bouveyron:2007gx}, remote sensing
-\citep{Tadjudin:1999fk}, seismic detection \citep{Anderson:2002kg}, and chemical
-spectra \citep*{Wu:1996us,Aeberhard:1993fp}.
-
-The dimension reduction that we have employed in this paper has reduced the
-dimension to rank$(\widehat{\bm \Sigma}) = q$. An interesting extension of our
-work would reduce the dimension $q$ further to a lower dimension $q_L < q$ using
-some criterion proposed, perhaps similar to that employed in principal
-components analysis. While it is not clear whether classification performance
-would improve via such a method, the efficiency of the training and model
-selection would certainly improve. Moreover, if $q_L = 2$ or $3$ was selected,
-the potential for low-dimensional graphical displays of high-dimensional data
-could be accomplished.
-
-\bibliographystyle{plainnat}
-\bibliography{rda}
-
 \end{document}

### 2016-01-20 reports/paper-hdrda.Rnw
Results table bolds best classifier per data set.

@@
 # Cells = Error (SD) rounded to 3 digits
+# The classifier with the minimum average error rate is in bold for each data set.
+
+bold_latex <- function(x) {
+  paste0('\\textbf{', x, '}')
+}
+
 results_microarray <- inner_join(avg_error_microarray, sd_error_microarray,
                                  by=c('data_set', 'Classifier')) %>%
-  mutate(Table_Cells=sprintf('%.3f (%.3f)', Avg_Error, SD_Error),
+  mutate(Table_Cells=sprintf('%.3f', Avg_Error),
          data_set=stringr::str_to_title(data_set),
          Classifier=gsub("_", " ", Classifier)) %>%
+  group_by(data_set) %>%
+  mutate(Table_Cells=ifelse(Avg_Error == min(Avg_Error), bold_latex(Table_Cells), Table_Cells),
+         Table_Cells=sprintf('%s (%.3f)', Table_Cells, SD_Error)) %>%
   select(-Avg_Error, -SD_Error) %>%
@@
 table_caption <- 'The average of the test error rates obtained on gene-expression data sets over 100 random training-test partitions.
-Standard deviations of the test error rates are given in the parentheses.'
+Standard deviations of the test error rates are given in the parentheses.
+The classifier with the minimum average error rate for each data set is in bold.'
 
@@
                      label='tab:microarray-results')
-print(xtable_out, include.rownames=FALSE)
+print(xtable_out, include.rownames=FALSE, sanitize.text.function = function(x) { x })
 @

### 2016-01-30 reports/paper-hdrda.Rnw
Applied edits from Mrs. Young #23

@@
 
-\cite{Friedman:1989tm} proposed the popular regularized discriminant analysis
-(RDA) classifier that utilizes a biased covariance-matrix estimator that
-partially pools the sample covariance matrices from linear and quadratic
+\citeauthor{Friedman:1989tm} proposed the popular regularized discriminant
+analysis (RDA) classifier that utilizes a biased covariance-matrix estimator
+that partially pools the sample covariance matrices from linear and quadratic
 disriminant analysis and shrinks the resulting estimator towards a scaled
 identity matrix. The RDA classifier's two tuning parameters are typically
-estimated via a computationally burdensome cross-validation procedure using a
-grid search. We formulate a new RDA-based classifier for the small-sample,
+estimated via a computationally burdensome cross-validation procedure that uses
+a grid search. We formulate a new RDA-based classifier for the small-sample,
 high-dimensional setting and then show that the classification decision rule is
@@
 on the well-known ridge-regression approach of \cite{Hoerl:1970cd}.
-\cite{Ching:2012fu}, \cite{Zhang:2010va}, \cite{Ye:2009gd}, \cite{Ji:2008wp},
-\cite*{Guo:2007te}, \cite*{Srivastava:2007ud}, \cite{Ye:2006vx}, and
-\cite{Ye:2005uu} have proposed regularization methods to improve the estimation
-of the class covariance matrices by adjusting the eigenvalues of the sample
-covariance matrices to ensure positive definiteness. Alternatively, dimension
-reduction typically transforms the original data in the high-dimensional space
-to a subspace having a much smaller dimension such that redundant and noisy
-information is removed
+Additionally, \cite{Ching:2012fu}, \cite{Zhang:2010va}, \cite{Ye:2009gd},
+\cite{Ji:2008wp}, \cite*{Guo:2007te}, \cite*{Srivastava:2007ud},
+\cite{Ye:2006vx}, and \cite{Ye:2005uu} have proposed regularization methods to
+improve the estimation of the class covariance matrices by adjusting the
+eigenvalues of the sample covariance matrices to ensure positive
+definiteness. Alternatively, dimension reduction typically transforms the
+original data in the high-dimensional space to a subspace having a much smaller
+dimension such that redundant and noisy information is removed
 \citep*{Fan:2012iq,Ye:2009gd,Bouveyron:2007gx,Ye:2006vx,Ye:2005hd,Antoniadis:2003bk}.
@@
 of parameters that must be estimated. For instance, \cite{Dudoit:2002ev} have
-assumed the features within each class are independent so that the population
-covariance matrices are diagonal, resulting in positive-definite sample
-covariance matrices. For excellent overviews of the regularization problem in
-discriminant analysis and other techniques, see \cite{Murphy:2012uq} and
-\cite*{Mkhadri:1997gy}.
+assumed that the features within each class are independent so that the
+population covariance matrices are diagonal, resulting in positive-definite
+sample covariance matrices. For excellent overviews of the regularization
+problem in discriminant analysis and other techniques, see \cite{Murphy:2012uq}
+and \cite*{Mkhadri:1997gy}.
 
@@
 towards a scaled identity matrix to ensure positive definiteness. By
-reparameterizing the \emph{RDA} classifier similar to \cite*{Hastie:2008dt} and
-\cite{Halbe:2007ea}, we show that the pooling parameter in the \emph{HDRDA}
-classifier determines the contribution of each training observation to the
-estimation of each class covariance matrix, enabling interpretability that has
-been previously lacking with the \emph{RDA} classifier \citep{Bensmail:1996fw}.
+reparameterizing the \emph{RDA} classifier similar to that of
+\cite*{Hastie:2008dt} and \cite{Halbe:2007ea}, we show that the pooling
+parameter in the \emph{HDRDA} classifier determines the contribution of each
+training observation to the estimation of each class covariance matrix, enabling
+interpretability that has been previously lacking with the \emph{RDA} classifier
+\citep{Bensmail:1996fw}.
 
@@
 computational effort during model selection when $p \gg N$ by applying reasoning
-similar to \cite{Ye:2006tq}. We establish that the matrix operations
+similar to that of \cite{Ye:2006tq}. We establish that the matrix operations
 corresponding to the null space of the pooled sample covariance matrix are
@@
 special cases of the \emph{HDRDA} classifier have no effect on the zero
-eigenvalues of the covariance-matrix estimators. Such methods include work from
-\cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and other methods studied by
+eigenvalues of the covariance-matrix estimators. Such techniques include work
+from \cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and other methods studied by
 \cite{Ramey:2013ji} and \citet*{Xu:2009fl}.
@@
 \emph{HDRDA} classifier along with its interpretation, properties, and
-computationally efficient model-selection procedure. In Section \ref{sec:sims},
-we describe our simulation studies of artificial and real data sets and also
-describes the experimental results. We conclude with a brief discussion in
-Section \ref{sec:discussion}.
+computationally efficient model-selection procedure. In Section \ref{sec:sims}
+we describe our simulation studies of artificial and real data sets and describe
+the experimental results. We conclude with a brief discussion in Section
+\ref{sec:discussion}.
 
@@
 Several regularization methods, such as the methods considered by
-\cite{Xu:2009fl}, \cite{Guo:2007te}, and \cite{Mkhadri:1995jp} have been
+\cite{Xu:2009fl}, \cite{Guo:2007te}, and \cite{Mkhadri:1995jp}, have been
 proposed in the literature to adjust the eigenvalues of $\widehat{\bm \Sigma}_k$
-so that \eqref{eq:qda} is calculable and provide reduced variability for
-$\widehat{\bm \Sigma}_k^{-1}$. A common form of the covariance-matrix regularization applies
-  a shrinkage factor $\gamma > 0$, so that
+so that \eqref{eq:qda} is calculable and provides reduced variability for
+$\widehat{\bm \Sigma}_k^{-1}$. A common form of the covariance-matrix
+regularization applies a shrinkage factor $\gamma > 0$, so that
 \begin{align}
@@
 As we have discussed above, several eigenvalue adjustment methods have been
-proposed to increase eigenvalues (approximately) equal to 0. To further improve
-the estimation of $\bm \Sigma_k$ and to stabilize the estimator's inverse, we
-define the eigenvalue adjustment of \eqref{eq:sig-lambda-alternative} as
+proposed that increase eigenvalues (approximately) equal to 0. To further
+improve the estimation of $\bm \Sigma_k$ and to stabilize the estimator's
+inverse, we define the eigenvalue adjustment of
+\eqref{eq:sig-lambda-alternative} as
 \begin{align}
@@
 literature. In practice, we generally are not interested in estimating
-$\alpha_k$ as it would incorporate $K$ additional tuning parameters to estimate
-via cross-validation, which is counterproductive to our goal of improving the
-\emph{RDA} classifier's computational efficiency. By setting $\alpha_k$ to an
+$\alpha_k$ because the estimation of $K$ additional tuning parameters via
+cross-validation is counterproductive to our goal of improving the \emph{RDA}
+classifier's computational efficiency. By setting $\alpha_k$ equal to an
 appropriate constant, the \emph{HDRDA} covariance-matrix estimator includes or
@@
 covariance-matrix estimator in \eqref{eq:ridge-estimator}, and various
-estimators proposed in the literature can be obtained by selecting $\gamma$
+estimators proposed in the literature can be obtained when one selects $\gamma$
 accordingly. For instance, with $\gamma = \tr\{ \widehat{\bm \Sigma} \} /
@@
 shrinkage of \eqref{eq:sig-rda} towards $\bm I_p$ and is weighted by the average
-of the eigenvalues of \eqref{eq:sig-lambda}. Finally, the RDA classifier is
-established by substituting \eqref{eq:sig-rda} into \eqref{eq:qda}.
+of the eigenvalues of \eqref{eq:sig-lambda}. Finally, \cite{Friedman:1989tm}
+established the RDA classifier by substituting \eqref{eq:sig-rda} into
+\eqref{eq:qda}.
 
@@
 interpretation of the pooling parameter $\lambda$ as discussed above. Moreover,
-as we discuss below, our formulation of \emph{HDRDA} classifier lends itself to
-rapid estimation in high-dimensional classification.
+as we discuss below, our formulation of the \emph{HDRDA} classifier lends itself
+to rapid estimation in high-dimensional classification.
 
@@
 component involves determinants and inverses of high-dimensional matrices, and
-by ignoring these calculations, a substantial reduction in computational costs
-is achieved. Furthermore, a byproduct is that adjustments to the associated
-eigenvalues have no effect on \eqref{eq:hdrda}. Lastly, we utilize the singular
-value decomposition to efficiently calculate the eigenvalue decomposition of
-$\widehat{\bm \Sigma}$, further reducing the computational costs of the
-\emph{HDRDA} classifier.
+by ignoring these calculations, we achieve a substantial reduction in
+computational costs. Furthermore, a byproduct is that adjustments to the
+associated eigenvalues have no effect on \eqref{eq:hdrda}. Lastly, we utilize
+the singular value decomposition to efficiently calculate the eigenvalue
+decomposition of $\widehat{\bm \Sigma}$, further reducing the computational
+costs of the \emph{HDRDA} classifier.
 
@@
 
-As an immediate immediate consequence of Lemma \ref{lemma:rda-tilde-Sigma_k}, we
-have the following corollary.
+As an immediate consequence of Lemma \ref{lemma:rda-tilde-Sigma_k}, we have the
+following corollary.
 
@@
 \begin{proof}
-The proof follows by setting $\gamma = 0$ in Lemma \ref{lemma:rda-tilde-Sigma_k}.
+The proof follows when we set $\gamma = 0$ in Lemma
+\ref{lemma:rda-tilde-Sigma_k}.
 \end{proof}
@@
 
-We now present our main result where we decompose \eqref{eq:hdrda} and show that
-the term requiring the largest computational costs does not contribute to the
-classification of an unlabeled observation performed using Lemma
+We now present our main result, where we decompose \eqref{eq:hdrda} and show
+that the term requiring the largest computational costs does not contribute to
+the classification of an unlabeled observation performed using Lemma
 \ref{lemma:RDA-constant-term}. Hence, we reduce \eqref{eq:hdrda} to an
@@
 
-As we can see, we are able to avoid the time-consuming inverses and determinants
-of $p \times p$ covariance matrices in \eqref{eq:hdrda} and instead calculate
-these same operations on $\bm W_k \in \mathbb{R}_{q \times q}$ in
+Thus, we can avoid the time-consuming inverses and determinants of $p \times p$
+covariance matrices in \eqref{eq:hdrda} and instead calculate these same
+operations on $\bm W_k \in \mathbb{R}_{q \times q}$ in
 \eqref{eq:hdrda-decomposed}. The substantial computational improvements arise
@@
 that the computational efficiency in calculating the inverse and determinant of
-$\bm W_k$ can be further improved via standard matrix operations by showing that
-the inverses and determinants of $\bm W_k$ can be performed on matrices of size
-$n_k \times n_k$.
+$\bm W_k$ can be further improved via standard matrix operations when we show
+that the inverses and determinants of $\bm W_k$ can be performed on matrices of
+size $n_k \times n_k$.
 
@@
 \{\gamma_h\}_{h=1}^H$. We estimate the $V$-fold cross-validation error rate for
-each candidate pair and select $(\widehat{\lambda}, \widehat{\gamma})$ that
+each candidate pair and select $(\widehat{\lambda}, \widehat{\gamma})$, which
 attains the minimum error rate. To calculate the $V$-fold cross-validation, we
@@
 \ldots, V$, we classify the observations in the $v$th fold by training a
-classifier on the remaining $V - 1$ folds. We then calculate the
-cross-validation error as the proportion of misclassified observations across
-the $V$ folds.
+classifier on the remaining $V - 1$ folds. We calculate the cross-validation
+error as the proportion of misclassified observations across the $V$ folds.
 
@@
 $(\bm x - \xbar_k)^{T} \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm x - \xbar_k)$
-individually for each row $\bm x$ of $\bm X_{test}$, we can instead calculate
-$(\bm X_{test} - \xbar_k \bm 1_k')' \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm
-X_{test} - \xbar_k \bm 1_k')$. The diagonal elements of the resulting matrix
-contain the individual quadratic form of each test observation, $\bm x_t$.
+individually for each row $\bm x$ of $\bm X_{test}$, we can calculate $(\bm
+X_{test} - \xbar_k \bm 1_k')' \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm X_{test} -
+\xbar_k \bm 1_k')$. The diagonal elements of the resulting matrix contain the
+individual quadratic form of each test observation, $\bm x_t$.
 
@@
 from \cite{Guo:2007te}. This design is widely used within the high-dimensional
-classification literature including the studies by \cite{Ramey:2013ji} and
+classification literature, including the studies by \cite{Ramey:2013ji} and
 \cite{Witten:2011kc}. First, we consider the block-diagonal covariance matrix
-from from \cite{Guo:2007te}
+from \cite{Guo:2007te},
 \begin{align}
@@
 We generated $K=3$ populations from \eqref{eq:contaminated-pdf} with $\bm
-\Sigma_k$ given in \eqref{eq:block-sigma}. We set the mean vector of class 1 to
+\Sigma_k$ given in \eqref{eq:block-sigma} and set the mean vector of class 1 to
 $\bm \mu_1 = \bm 0_p$. Next, comparable to \cite{Guo:2007te}, the first 100
-features of $\bm \mu_2$ were 1/2, while the rest were set to 0, i.e., $\bm \mu_2
-= (\underbrace{1/2, \ldots, 1/2}_{100}, \underbrace{0, \ldots, 0}_{p -
-  100})$. For simplicity, we defined $\bm \mu_3 = -\bm \mu_2$. The three
+features of $\bm \mu_2$ were set to 1/2, while the rest were set to 0, i.e.,
+$\bm \mu_2 = (\underbrace{1/2, \ldots, 1/2}_{100}, \underbrace{0, \ldots, 0}_{p
+  - 100})$. For simplicity, we defined $\bm \mu_3 = -\bm \mu_2$. The three
 populations differed in their mean vectors in the first 100 features
@@
 probability $\epsilon$ increased regardless of the value of $p$. Sensitivity to
-the presence of outliers was most apparent for the for the Pang, Tong, and
-Witten classifiers. For smaller dimensions, the random-forest and \emph{HDRDA}
+the presence of outliers was most apparent for the Pang, Tong, and Witten
+classifiers. For smaller dimensions, the random-forest and \emph{HDRDA}
 classifiers tended to outperform the remaining classifiers with the random
@@
 rates for these three classifiers continued as $\epsilon$ increased. Guo's
-method always outperformed Pang, Witten, and Tong, but once outliers were
-present, the Guo classifier's average error rate was not competitive with
-the \emph{HDRDA} classifiers or the random-forest classifier.
+method always outperformed those of Pang, Witten, and Tong, but after outliers
+were introduced, the Guo classifier's average error rate was not competitive
+with the \emph{HDRDA} classifiers or the random-forest classifier.
 
@@
 performance. First, the Pang, Tong, and Witten methods, along with the
-random-forest method increased in average error rate as $p$ increased.
+random-forest method, increased in average error rate as $p$ increased.
 Contrarily, the performance of the \emph{HDRDA} and Guo classifiers was hardly
@@
 
-We compared the \emph{HDRDA} classifer to the five competing classifiers on six
+We compared the \emph{HDRDA} classifier to the five competing classifiers on six
 benchmark gene-expression microarray data sets. We evaluated the classification
@@
 \cite{Chiaretti:2004gq} measured the gene-expression profiles for 128
-individuals with acute lymphoblastic leukaemia (ALL) using Affymetrix human
+individuals with acute lymphoblastic leukemia (ALL) using Affymetrix human
 95Av2 arrays. Following \cite{Xu:2009fl}, we restricted the data set to $K = 2$
 classes such that $n_1 = 74$ observations were without cytogenetic abnormalities
-and $n_2 = 37$ observations have a detected BCR/ABL gene. The robust multichip
-average normalization method was applied to all 12,625 gene expression levels.
+and $n_2 = 37$ observations had a detected BCR/ABL gene. The robust multichip
+average normalization method was applied to all 12,625 gene-expression levels.
 
@@
 breast tumors using Affymetrix U133A arrays and ribonucleic-acid (RNA)
-amplification. Each tissue pair was gathered from the same patient and
-consisted of a snap-frozen tissue and a tissue suspended in a RNAlater
-preservative. Overall, 31 breast-cancer and 21 colon-cancer pairs were gathered,
-resulting in $K = 2$ classes with $n_1 = 62$ and $n_2 = 42$. A purpose of the
-study was to determine whether the disease state could be identified using
-22,283 gene-expression profiles.
+amplification. Each tissue pair was gathered from the same patient and consisted
+of a snap-frozen tissue and a tissue suspended in an RNAlater preservative.
+Overall, 31 breast-cancer and 21 colon-cancer pairs were gathered, resulting in
+$K = 2$ classes with $n_1 = 62$ and $n_2 = 42$. A purpose of the study was to
+determine whether the disease state could be identified using 22,283
+gene-expression profiles.
 
@@
 
-\cite{Nakayama:2007fl} acquired 105 gene-expression samples of 10 types of soft
-tissue tumors through an oligonucleotide microarray, including 16 samples of
-synovial sarcoma (SS), 19 samples of myxoid/round cell liposarcoma (MLS), 3
+\cite{Nakayama:2007fl} acquired 105 gene-expression samples of 10 types of
+soft-tissue tumors through an oligonucleotide microarray, including 16 samples
+of synovial sarcoma (SS), 19 samples of myxoid/round cell liposarcoma (MLS), 3
 samples of lipoma, 3 samples of well-differentiated liposarcoma (WDLS), 15
@@
 
-Approximately 30\%-40\% of adult non-Hodgkin lymphomas are diffuse large B-cell
-lymphomas (DLBCLs) \cite{Shipp:2002ka}.  However, only a small proportion of
-DLBCL patients are cured with modern chemotherapeutic regimens.  Several models
-have been proposed, such as the International Prognostic Index (IPI), to
-determine a patient's curability.  These models rely on clinical covariates,
-such as age, to determine if the patient can be cured and are often
-ineffective. \cite{Shipp:2002ka} have argued that more effective means are
-desired to determine a patient's curability. The authors measured 6,817 gene
-expression levels from 58 DLBCL patient samples with customized cDNA
-(lymphochip) microarrays to investigate the curability of patients treated with
-cyclophosphamide, adriamycin, vincristine, and prednisone (CHOP)-based
-chemotherapy. Among the 58 DLBCL patient samples, 32 are from cured patients
-while 26 are from patients with fatal or refractory disease.
+According to \cite{Shipp:2002ka}, approximately 30\%-40\% of adult non-Hodgkin
+lymphomas are diffuse large B-cell lymphomas (DLBCLs). However, only a small
+proportion of DLBCL patients are cured with modern chemotherapeutic regimens.
+Several models have been proposed, such as the International Prognostic Index
+(IPI), to determine a patient's curability.  These models rely on clinical
+covariates, such as age, to determine if the patient can be cured, and the
+models are often ineffective. \cite{Shipp:2002ka} have argued that researchers
+need more effective means to determine a patient's curability. The authors
+measured 6,817 gene-expression levels from 58 DLBCL patient samples with
+customized cDNA (lymphochip) microarrays to investigate the curability of
+patients treated with cyclophosphamide, adriamycin, vincristine, and prednisone
+(CHOP)-based chemotherapy. Among the 58 DLBCL patient samples, 32 are from cured
+patients while 26 are from patients with fatal or refractory disease.
 
@@
 The average error-rate estimates for the Pang, Tong, and Witten classifiers were
-comparable across all six data sets. Furthermore, the average error rates for the
-Pang and Tong classifiers were approximately equal for all data sets except for
-the Chiaretti dataset. This result suggests that the mean and variance
+comparable across all six data sets. Furthermore, the average error rates for
+the Pang and Tong classifiers were approximately equal for all data sets except
+for the Chiaretti dataset. This result suggests that the mean and variance
 estimators used in lieu of the MLEs provided little improvement to
@@
 exhibited numerical instability. The classifier's denominator was approximately
-zero for both classes and lead to the poor classification performance.
+zero for both classes and led to the poor classification performance.
 
@@
 classifiers typically outperformed the random-forest classifier challenges the
-claim of \cite{FernandezDelgado:2014ul} that the random forest is typically
+claim of \cite{FernandezDelgado:2014ul} that random forests are typically
 superior. Further studies should be performed to validate this statement in the
@@
 have noted that diagonal classifiers can often yield inferior classification
-performance to other classification methods. We have shown that the \emph{HDRDA}
-classifier often yields superior classification accuracy to the diagonal
-classifiers considered when applied to small-sample, high-dimensional data sets,
-confirming the assertions of \cite{Mai:2012bf} and \cite{Fan:2012iq}.
-Furthermore, we have demonstrated that \emph{HDRDA} classifiers are more robust
-to the presence of outliers than the diagonal classifiers.
+performance when compared to other classification methods. In fact, we have
+shown that the \emph{HDRDA} classifier often yields superior classification
+accuracy when compared to the diagonal classifiers considered using six
+small-sample, high-dimensional data sets, confirming the assertions of
+\cite{Mai:2012bf} and \cite{Fan:2012iq}. Furthermore, we have demonstrated that
+\emph{HDRDA} classifiers are more robust to the presence of outliers than the
+diagonal classifiers.
 
@@
 high-dimensional setting \citep{Webb:2011vu} but is limited by its
-computationally intense model selection procedure, our work enables the
+computationally intense model-selection procedure, our work enables the
 \emph{RDA} classifier to be employed with small-sample, high-dimensional data in
@@
 some criterion proposed, perhaps similar to that employed in principal
-components analysis. While it is not clear whether classification performance
-would improve via such a method, the efficiency of the training and model
-selection would certainly improve. Moreover, if $q_L = 2$ or $3$ was selected,
-the potential for low-dimensional graphical displays of high-dimensional data
-could be accomplished.
+components analysis. While unclear whether the classification performance would
+improve via such a method, the efficiency of the training and model selection
+would certainly improve. Moreover, if $q_L = 2$ or $3$ was selected, the
+potential for low-dimensional graphical displays of high-dimensional data could
+be accomplished.
 

### 2016-01-30 reports/paper-hdrda.Rnw
Thanked Mrs. Young.
Added coauthor organizations. #25

@@
 
-\usepackage{graphicx, amsmath, amssymb, bm, mathtools, amsthm, natbib, setspace, algorithm2e}
+\usepackage{graphicx, amsmath, amssymb, bm, mathtools, amsthm, natbib, setspace, algorithm2e, authblk}
 
@@
 
-\author{John A. Ramey, Caleb K. Stein, Phil D. Young, and Dean M. Young}
+%\author{John A. Ramey, Caleb K. Stein, Phil D. Young, and Dean M. Young}
+
+\author[a]{John A. Ramey}
+\author[b]{Caleb K. Stein}
+\author[c]{Phil D. Young}
+\author[d]{Dean M. Young}
+\affil[a]{uStudio, Inc.}
+\affil[b]{Myeloma Institute, University of Arkansas for Medical Sciences}
+\affil[c]{Department of Management and Information Systems, Baylor University}
+\affil[d]{Department of Statistical Science, Baylor University}
+
+\renewcommand\Authands{ and }
 
@@
 
+We thank Mrs. Joy Young for her numerous recommendations that enhanced the
+quality of our writing.
+
 \bibliographystyle{plainnat}

### 2016-01-30 reports/paper-hdrda.Rnw
Added link to GitHub repo. Closes #24

@@
 
-\usepackage{graphicx, amsmath, amssymb, bm, mathtools, amsthm, natbib, setspace, algorithm2e, authblk}
+\usepackage{graphicx, amsmath, amssymb, bm, mathtools, amsthm, natbib, setspace}
+\usepackage{algorithm2e, authblk, url}
 
@@
 surname. All simulations were conducted on an Amazon Elastic Compute Cloud (EC2)
-{\tt c4.4xlarge} instance.
+{\tt c4.4xlarge} instance. Our analyses can be reproduced using the code
+available in the \url{https://github.com/ramhiser/paper-hdrda} repository.
 

### 2016-01-30 reports/paper-hdrda.Rnw
Rm'd a slight redundancy

@@
 original design from \cite{Guo:2007te} comprised two $p$-dimensional
-multivariate normal populations with a common block-diagonal covariance matrix
-where $\bm\Sigma_1 = \bm\Sigma_2$.
+multivariate normal populations with a common block-diagonal covariance matrix.
 

### 2016-01-31 reports/paper-hdrda.Rnw
Slight reword
Applied CSDA LaTeX template #19
Need to fix line numbers.

@@
-\documentclass{article}
-
-\usepackage{graphicx, amsmath, amssymb, bm, mathtools, amsthm, natbib, setspace}
-\usepackage{algorithm2e, authblk, url}
+\documentclass[review]{elsarticle}
+
+\usepackage{graphicx, amsmath, amssymb, bm, mathtools, amsthm, natbib}
+\usepackage{algorithm2e, url, lineno}
 
@@
 
-\pagestyle{plain}
 %----------------Page dimensions ----------------
@@
 
+\begin{frontmatter}
+
 \title{High-Dimensional Regularized Discriminant Analysis}
 
-%\author{John A. Ramey, Caleb K. Stein, Phil D. Young, and Dean M. Young}
-
-\author[a]{John A. Ramey}
-\author[b]{Caleb K. Stein}
-\author[c]{Phil D. Young}
-\author[d]{Dean M. Young}
-\affil[a]{uStudio, Inc.}
-\affil[b]{Myeloma Institute, University of Arkansas for Medical Sciences}
-\affil[c]{Department of Management and Information Systems, Baylor University}
-\affil[d]{Department of Statistical Science, Baylor University}
-
-\renewcommand\Authands{ and }
-
-\maketitle
-
-\doublespacing
+\author[jar]{John A. Ramey\cortext[cor1]{Corresponding author}}
+\author[cks]{Caleb K. Stein}
+\author[pdy]{Phil D. Young}
+\author[dmy]{Dean M. Young}
+\address[jar]{uStudio, Inc.}
+\address[cks]{Myeloma Institute, University of Arkansas for Medical Sciences}
+\address[pdy]{Department of Management and Information Systems, Baylor University}
+\address[dmy]{Department of Statistical Science, Baylor University}
 
@@
 \end{abstract}
+
+\end{frontmatter}
+
+\linenumbers
 
@@
 The remainder of this paper is organized as follows. In Section
-\ref{sec:preliminaries} we present the discrimination problem and necessary
+\ref{sec:preliminaries} we present the classification problem and necessary
 notation to describe our contributions. In Section \ref{sec:rda} we present the

### 2016-01-31 reports/paper-hdrda.Rnw
Fixed the line numbers in Elsevier version.

@@
 sample \emph{QDA} classifier
+\begin{linenomath*}
 \begin{align}
@@
 \end{align}
+\end{linenomath*}
 where $\xbar_k$ and $\widehat{\bm \Sigma}_k$ are the maximum likelihood
@@
   \bm \Sigma}_k$ in \eqref{eq:qda}, where
+\begin{linenomath*}
 \begin{align}
@@
 \end{align}
+\end{linenomath*}
 is the MLE for $\bm \Sigma$. Here, \eqref{eq:qda} reduces to the sample
@@
 regularization applies a shrinkage factor $\gamma > 0$, so that
+\begin{linenomath*}
 \begin{align}
@@
 \end{align}
+\end{linenomath*}
 similar to a method employed in ridge regression \citep{Hoerl:1970cd}. Equation
@@
 the convex combination
+\begin{linenomath*}
 \begin{align}
@@
 \end{align}
+\end{linenomath*}
 where $\lambda \in [0, 1]$ is the \emph{pooling} parameter. By rewriting
@@
 interpretation of $\widehat{\bm \Sigma}_k(\lambda)$. That is,
+\begin{linenomath*}
 \begin{align}
@@
 \end{align}
+\end{linenomath*}
 where $c_{ik}(\lambda) = \lambda N^{-1} + (1 - \lambda)n_k^{-1}I(y_i = k)$.
@@
 \eqref{eq:sig-lambda-alternative} as
+\begin{linenomath*}
 \begin{align}
@@
 \end{align}
+\end{linenomath*}
 where $\alpha_k \ge 0$ and $\gamma \ge 0$ is an eigenvalue-shrinkage
@@
 \emph{HDRDA} classifier
+\begin{linenomath*}
 \begin{align}
@@
 \end{align}
+\end{linenomath*}
 For $\gamma > 0$, $\tilde{\bm \Sigma}_k$ is nonsingular such that $\tilde{\bm
@@
 estimate $\bm \Sigma_k$ with
+\begin{linenomath*}
 \begin{align}
     \widehat{\bm \Sigma}_k(\lambda) = n_k^{-1}(\lambda) \bm S_k(\lambda) \quad
-    (k =
-      1, \ldots, K),\label{eq:sig-lambda}
-      \end{align}
+    (k = 1, \ldots, K),\label{eq:sig-lambda}
+\end{align}
+\end{linenomath*}
 where $\lambda \in [0, 1]$, $n_k(\lambda) = (1 - \lambda) n_k + \lambda N$, $
@@
 proposed the biased covariance-matrix estimator
-\begin{align}
-          \widehat{\bm \Sigma}_k(\lambda, \gamma) = (1 - \gamma) \widehat{\bm
-            \Sigma}_k(\lambda) + \gamma \frac{\tr\{\widehat{\bm
-              \Sigma}_k(\lambda)\}}{p} \bm I_p,\label{eq:sig-rda}
-          \end{align}
+\begin{linenomath*}
+\begin{align}
+    \widehat{\bm \Sigma}_k(\lambda, \gamma) = (1 - \gamma) \widehat{\bm \Sigma}_k(\lambda) + \gamma \frac{\tr\{\widehat{\bm \Sigma}_k(\lambda)\}}{p} \bm I_p,\label{eq:sig-rda}
+\end{align}
+\end{linenomath*}
 where $\gamma \in [0, 1]$ is a regularization parameter that controls the
@@
 q \le p$. Then, we have
+\begin{linenomath*}
 \begin{align}
@@
 \end{align}
+\end{linenomath*}
 \end{lemma}
@@
 the decision rule in \eqref{eq:hdrda} is equivalent to
+\begin{linenomath*}
 \begin{align}
@@
 \end{align}
+\end{linenomath*}
 \end{thm}
@@
 \mathbb{R}_{p \times 1}$, we have that
+\begin{linenomath*}
 \begin{align*}
@@
 \end{align*}
+\end{linenomath*}
 Because $\gamma$ is constant for $k = 1, \ldots, K$, we can omit the $(p - q)
@@
 and
+\begin{linenomath*}
 \begin{align}
@@
 \end{align}
+\end{linenomath*}
 \end{proposition}
@@
 from \cite{Guo:2007te},
+\begin{linenomath*}
 \begin{align}
@@
 \end{align}
+\end{linenomath*}
 where the $(i,j)$th entry of the block matrix $\bm\Sigma^{(\rho_k)} \in
 \mathbb{R}_{100 \times 100}$ is
+\begin{linenomath*}
 \begin{align*}
@@
 \end{align*}
+\end{linenomath*}
 The block-diagonal covariance structure in \eqref{eq:block-sigma} resembles
@@
 \Sigma)$, we write the PDF of the $k$th class as
+\begin{linenomath*}
 \begin{align}
@@
 \end{align}
+\end{linenomath*}
 where $\epsilon \in [0, 1]$ is the probability that an observation is

### 2016-02-03 reports/paper-hdrda.Rnw
Updated the bibliography style to Elsevier's #19
Slight changes made to submit to arXiv. #20

--- 2016-01-30 reports/paper-hdrda.Rnw
@@
 
-%\author{John A. Ramey, Caleb K. Stein, Phil D. Young, and Dean M. Young}
+% AMS' MSC Classes: 62H30, 65F15, 65F20, 65F22
 
@@
 
-\doublespacing
-
 \begin{abstract}
@@
 that partially pools the sample covariance matrices from linear and quadratic
-disriminant analysis and shrinks the resulting estimator towards a scaled
+discriminant analysis and shrinks the resulting estimator towards a scaled
 identity matrix. The RDA classifier's two tuning parameters are typically

### 2016-02-04 hdrda-timing-comparison.r
Rm'd some legacy code.

@@ -1,73 +1,0 @@
-library(sparsediscrim)
-library(sortinghat)
-library(microbenchmark)
-library(klaR)
-
-# Simulation Parameters
-K <- 4
-n <- rep.int(x = 25, times = K)
-p <- 500
-means <- lapply(c(-3, -1, 1, 3), rep.int, times = p)
-cov_mat <- diag(p)
-
-# Generates multivariate normal data
-set.seed(42)
-data <- simdata_normal(n = n, mean = means, cov = cov_mat)
-
-# klaR's Implementation of Friedman (1989)
-rda_model_selection <- function(num_tuning = 3) {
-  tuning_grid <- expand.grid(lambda = seq(0, 1, length = num_tuning),
-                             gamma = seq(0, 1, length = num_tuning))
-
-  rda_errors <- vector(mode = "list", length = nrow(tuning_grid))
-
-  for (i in seq_len(nrow(tuning_grid))) {
-    rda_errors[[i]] <- try({
-      rda_out <- rda(x = data$x,
-                     grouping = data$y,
-                     estimate.error = FALSE)
-      rda_out$error.rate['crossval']
-    })
-  }
-
-  rda_errors[!sapply(rda_errors, is.numeric)] <- NA
-
-  tuning_grid[which.min(unlist(rda_errors)), ]
-}  
-
-# HDRDA
-hdrda_model_selection <- function(num_tuning = 3) {
-  hdrda_cv(x = data$x,
-           y = data$y,
-           num_lambda = num_tuning,
-           num_gamma = num_tuning,
-           shrinkage_type = "convex")
-}
-
-timing_results <- list()
-
-# 3 x 3 Grid of Tuning Parameters
-timing_results[["3x3"]] <- microbenchmark(
-  rda_model_selection(num_tuning = 3),
-  hdrda_model_selection(num_tuning = 3),
-  times = 50)
-
-print(timing_results[["3x3"]])
-
-# 4 x 4 Grid of Tuning Parameters
-timing_results[["4x4"]] <- microbenchmark(
-  rda_model_selection(num_tuning = 4),
-  hdrda_model_selection(num_tuning = 4),
-  times = 50)
-
-print(timing_results[["4x4"]])
-
-# 5 x 5 Grid of Tuning Parameters
-timing_results[["5x5"]] <- microbenchmark(
-  rda_model_selection(num_tuning = 5),
-  hdrda_model_selection(num_tuning = 5),
-  times = 50)
-
[…]

### 2016-02-04 reports/paper-hdrda.Rnw

--- 2016-01-31 reports/paper-hdrda.Rnw
@@
-\documentclass[review]{elsarticle}
+\documentclass[review,authoryear]{elsarticle}
 
@@
 
-\bibliographystyle{plainnat}
+\bibliographystyle{elsarticle-harv}
 \bibliography{rda}

### 2016-10-02 reports/paper-hdrda.Rnw
Added keywords and MSC 2010 codes to paper
More keywords to paper.
Updated \oplus notation to 2x2 block-diagonal Closes #27

@@
 \end{abstract}
+
+\begin{keyword}
+Regularized discriminant analysis\sep High-dimensional classification\sep Covariance-matrix regularization\sep Singular value decomposition\sep Multivariate analysis\sep Dimension reduction
+\MSC[2010] 62H30\sep 65F15\sep 65F20\sep 65F22
+\end{keyword}
 
@@
 0_m$ is understood to denote $\bm 0_{m \times m}$. Define $\bm 1_m \in
-\mathbb{R}_{m \times 1}$ as a vector of ones. Let $\bm A \oplus \bm B$ denote
-the direct sum of $\bm A \in \mathbb{R}_{r \times r}$ and $\bm B \in
-\mathbb{R}_{s \times s}$ \citep[Chapter 1]{Lutkepohl:1996uz}. Let $\bm A^{T}$,
-$\bm A^+$, and $\mathcal{N}(\bm A)$ denote the transpose, the Moore-Penrose
-pseudoinverse, and the null space of $\bm A \in \mathbb{R}_{m \times p}$,
-respectively. Denote by $\mathbb{R}_{p \times p}^{>}$ the cone of real $p \times
-p$ positive-definite matrices. Similarly, let $\mathbb{R}_{p \times p}^{\ge}$
-denote the cone of real $p \times p$ positive-semidefinite matrices. Let
-$V^{\perp}$ denote the orthogonal complement of a vector space $V \subset
-\mathbb{R}_{p \times 1}$. For $c \in \mathbb{R}$, let $c^+ = 1/c$ if $c \ne 0$
-and $0$ otherwise.
+\mathbb{R}_{m \times 1}$ as a vector of ones. Let $\bm A^{T}$, $\bm A^+$, and
+$\mathcal{N}(\bm A)$ denote the transpose, the Moore-Penrose pseudoinverse, and
+the null space of $\bm A \in \mathbb{R}_{m \times p}$, respectively. Denote by
+$\mathbb{R}_{p \times p}^{>}$ the cone of real $p \times p$ positive-definite
+matrices. Similarly, let $\mathbb{R}_{p \times p}^{\ge}$ denote the cone of real
+$p \times p$ positive-semidefinite matrices. Let $V^{\perp}$ denote the
+orthogonal complement of a vector space $V \subset \mathbb{R}_{p \times 1}$. For
+$c \in \mathbb{R}$, let $c^+ = 1/c$ if $c \ne 0$ and $0$ otherwise.
 
@@
 expression for $\tilde{\bm \Sigma}_k$ in terms of the matrix of eigenvectors of
-$\widehat{\bm \Sigma}$. Let $\widehat{\bm \Sigma} = \bm U \bm D \bm U^{T}$ be the
-eigendecomposition of $\widehat{\bm \Sigma}$ such that $\bm D \in \mathbb{R}_{p
-  \times p}^{\ge}$ is the diagonal matrix of eigenvalues of $\widehat{\bm \Sigma}$
-with $\bm D = \bm D_q \oplus \bm 0_{p-q}$, $\bm D_q \in \mathbb{R}_{q \times
-  q}^{>}$ is the diagonal matrix consisting of the positive eigenvalues of
-$\widehat{\bm \Sigma}$, the columns of $\bm U \in \mathbb{R}_{p \times p}$ are
-the corresponding orthonormal eigenvectors of $\widehat{\bm \Sigma}$, and
-rank$(\widehat{\bm \Sigma}) = q$. Then, we partition $\bm U = (\bm U_1,
-\bm U_2)$ such that $\bm U_1 \in \mathbb{R}_{p \times q}$ and $\bm U_2 \in
-\mathbb{R}_{p \times (p - q)}$.
+$\widehat{\bm \Sigma}$. Let $\widehat{\bm \Sigma} = \bm U \bm D \bm U^{T}$ be
+the eigendecomposition of $\widehat{\bm \Sigma}$ such that $\bm D \in
+\mathbb{R}_{p \times p}^{\ge}$ is the diagonal matrix of eigenvalues of
+$\widehat{\bm \Sigma}$ with
+\begin{align*}
+  \bm D = \begin{bmatrix}
+    \bm D_q & \bm 0 \\
+    \bm{0} & \bm 0_{p-q}
+    \end{bmatrix},
+\end{align*}
+$\bm D_q \in \mathbb{R}_{q \times q}^{>}$ is the diagonal matrix consisting of
+the positive eigenvalues of $\widehat{\bm \Sigma}$, the columns of $\bm U \in
+\mathbb{R}_{p \times p}$ are the corresponding orthonormal eigenvectors of
+$\widehat{\bm \Sigma}$, and rank$(\widehat{\bm \Sigma}) = q$. Then, we partition
+$\bm U = (\bm U_1, \bm U_2)$ such that $\bm U_1 \in \mathbb{R}_{p \times q}$ and
+$\bm U_2 \in \mathbb{R}_{p \times (p - q)}$.
 
@@
 \begin{align}
-	\tilde{\bm \Sigma}_k &= \bm U(\bm W_k \oplus \gamma \bm I_{p-q})\bm U^{T}, \quad
-  k = 1, \ldots, K,\label{eq:rda-matrix}
+	\tilde{\bm \Sigma}_k &= \bm U
+    \begin{bmatrix}
+      \bm W_k & \bm 0 \\
+      \bm 0 & \gamma \bm I_{p-q}
+    \end{bmatrix}
+    \bm U^{T}, \quad k = 1, \ldots, K,\label{eq:rda-matrix}
 \intertext{where}
@@
 of $\widehat{\bm \Sigma}_k$, which implies that $\widehat{\bm \Sigma}_k \bm U_2
-= \bm 0_{p \times (p - q)}$. Hence, $\bm U^{T} \widehat{\bm \Sigma}_k \bm U =
-\bm U_1^{T} \widehat{\bm \Sigma}_k \bm U_1 \oplus \bm 0_{p-q}$, $k = 1, \ldots,
-K$. Thus, $\bm U^{T} \tilde{\bm \Sigma}_k \bm U = \alpha_k \{(1 - \lambda) \bm
-U^{T} \widehat{\bm \Sigma}_k \bm U + \lambda \bm D\} + \gamma \bm I_p$, and
+= \bm 0_{p \times (p - q)}$. Hence,
+\begin{align*}
+  \bm U^{T} \widehat{\bm \Sigma}_k \bm U = \begin{bmatrix}
+    \bm U_1^{T} \widehat{\bm \Sigma}_k \bm U_1 & \bm 0 \\
+    \bm 0 & \bm 0_{p-q}
+  \end{bmatrix}, \quad k = 1, \ldots, K.
+\end{align*}
+Thus, $\bm U^{T} \tilde{\bm \Sigma}_k \bm U = \alpha_k \{(1 - \lambda) \bm U^{T}
+\widehat{\bm \Sigma}_k \bm U + \lambda \bm D\} + \gamma \bm I_p$, and
 \eqref{eq:rda-matrix} holds because $\bm U$ is orthogonal.
@@
 \begin{proof}
-From \eqref{eq:rda-matrix}, we have that $\tilde{\bm \Sigma}_k^{+} = \bm U(
-\bm W_k^{-1} \oplus \gamma^{+} \bm I_{p-q} )\bm U^{T}$ and $|\tilde{\bm \Sigma}_k| =
-\gamma^{p-q}| \bm W_k |$, $k = 1, \ldots, K$. Therefore, for all $\bm x \in
-\mathbb{R}_{p \times 1}$, we have that
+From \eqref{eq:rda-matrix}, we have that
+\begin{align*}
+  \tilde{\bm \Sigma}_k^{+} = \bm U \begin{bmatrix}
+    \bm W_k^{-1} & \bm 0 \\
+    \bm 0 & \gamma^{+} \bm I_{p-q}
+  \end{bmatrix}
+  \bm U^{T}
+\end{align*}
+and $|\tilde{\bm \Sigma}_k| = \gamma^{p-q}| \bm W_k |$, $k = 1, \ldots,
+K$. Therefore, for all $\bm x \in \mathbb{R}_{p \times 1}$, we have that
 \begin{linenomath*}
@@
 X_c^{T} = \bm M \bm D \bm M^{T}$. Next, we compute $\bm U = \bm X_c^{T} \bm M
-\bm D^{+/2}$, where $\bm D^{+/2} = \bm D_q^{-1/2} \oplus \bm 0_{N-q}$. We then
-determine $q$, the number of numerically nonzero eigenvalues present in $\bm D$,
-by calculating the number of eigenvalues that exceeds some tolerance value, say,
-$1 \times 10^{-6}$. We then extract $\bm U_1$ as the first $q$ columns of $\bm
-U$.
+\bm D^{+/2}$, where
+\begin{align*}
+  \bm D^{+/2} = \begin{bmatrix}
+    \bm D_q^{-1/2} & \bm 0 \\
+    \bm 0 & \bm 0_{N-q}
+  \end{bmatrix}.
+\end{align*}
+We then determine $q$, the number of numerically nonzero eigenvalues present in
+$\bm D$, by calculating the number of eigenvalues that exceeds some tolerance
+value, say, $1 \times 10^{-6}$. We then extract $\bm U_1$ as the first $q$
+columns of $\bm U$.
 
@@
 \begin{align}
-  \bm\Sigma_k = \bm\Sigma^{(\rho_k)} \oplus \bm\Sigma^{(-\rho_k)} \oplus \ldots \oplus \bm\Sigma^{(\rho_k)},\label{eq:block-sigma}
+  \bm\Sigma_k = \begin{bmatrix}
+    \bm\Sigma^{(\rho_k)} & \bm 0_{100} & \bm 0_{100} & \cdots & \cdots & \cdots \\
+    \bm 0_{100} & \bm\Sigma^{(-\rho_k)} & \bm 0_{100} & \bm 0_{100} & \cdots & \vdots \\
+    \bm 0_{100} & \bm 0_{100} & \bm\Sigma^{(\rho_k)} & \bm 0_{100} & \cdots & \vdots \\
+    \vdots & \bm 0_{100} & \bm 0_{100} & \bm\Sigma^{(-\rho_k)} & \bm 0_{100} & \vdots \\
+    \vdots & \vdots & \vdots & \bm 0_{100} & \ddots & \vdots \\
+    \cdots & \cdots & \cdots & \cdots & \cdots & \cdots \\
+  \end{bmatrix},\label{eq:block-sigma}
 \end{align}

### 2016-10-23 reports/paper-hdrda.Rnw
Revamped abstract #28

@@
 
-\citeauthor{Friedman:1989tm} proposed the popular regularized discriminant
-analysis (RDA) classifier that utilizes a biased covariance-matrix estimator
-that partially pools the sample covariance matrices from linear and quadratic
-discriminant analysis and shrinks the resulting estimator towards a scaled
-identity matrix. The RDA classifier's two tuning parameters are typically
-estimated via a computationally burdensome cross-validation procedure that uses
-a grid search. We formulate a new RDA-based classifier for the small-sample,
-high-dimensional setting and then show that the classification decision rule is
-equivalent to a classifier in a subspace having a much lower dimension. As a
-result, the utilization of the dimension-reduction step yields a substantial
-reduction in computation during model selection. Also, our parameterization
-offers interpretability that was previously lacking with the RDA classifier. We
-demonstrate that our proposed classifier is often superior to several recently
-proposed sparse and regularized classifiers in terms of classification accuracy
-with three artificial and six real high-dimensional data sets. Finally, we
-provide an implementation of our proposed classifier in the {\tt sparsediscrim}
-R package, which is available on CRAN.
+We develop a new and computationally efficient implementation of
+\citeauthor{Friedman:1989tm}'s popular regularized discriminant analysis (RDA)
+classifier in the small-sample, high-dimensional setting. Our proposed
+high-dimensional RDA (HDRDA) classifier modifies \citeauthor{Friedman:1989tm}'s
+original formulation to achieve interpretability previously lacking. We are able
+to interpret the biased covariance matrix estimator employed in the HDRDA
+classifier as one that is estimated from all training observations regardless of
+class, solidifying the ``borrowing'' operation \citeauthor{Friedman:1989tm}'s
+RDA is claimed to have. Moreover, our formulation also differs from
+\citeauthor{Friedman:1989tm}'s in that we show our classifier is equivalent to
+one in a reduced feature space with dimension approximately equal to the
+training sample size. As a result, the matrix operations used in the HDRDA
+classifier are computationally linear in the number of features, making HDRDA
+well-suited for high-dimensional classification in practice. Timing comparisons
+between our HDRDA implementation in the {\tt sparsediscrim} R package and the
+standard RDA formulation in the {\tt klaR} R package demonstrate that as the
+number of features increases, the computational runtime of HDRDA is drastically
+smaller than RDA. We also demonstrate that HDRDA is often superior to several
+sparse and regularized classifiers in terms of classification accuracy with
+three artificial and six real high-dimensional data sets.
 

### 2016-10-23 reports/paper-hdrda.Rnw
Major rewrite of the intro. #28

@@
 
-As machine-learning algorithms are increasingly utilized to automate data-driven
-decisions, researchers continue to develop a wealth of \emph{classifiers}
-employed in a variety of disciplines, such as medical research, facial
-recognition, remote sensing, and chemical analysis
-\citep*{Ching:2012fu,Li:2012ev,Zhang:2010va}. Despite the advances in
-classification to facilitate technological advancements, well-established
-classifiers often degrade in accuracy when applied to modern, high-dimensional
-data sets if the number of features $p$ exceeds the total sample size $N$, due
-to the \emph{curse of dimensionality} \citep{Bellman:1961tn}. For instance,
-linear discriminant analysis (\emph{LDA}) and quadratic discriminant analysis
-(\emph{QDA}) are popular because of their simplicity, speed of computation, and
-interpretability. However, both methods are incalculable when $p > N$ because
-the class sample covariance matrices are singular \citep*{Bouveyron:2007gx}.
-
-As a result, researchers have emphasized three standard approaches to modifying
-the \emph{LDA} and \emph{QDA} classifiers for application to high-dimensional
-data: 1) covariance-matrix regularization, 2) dimension reduction, and 3)
-parameter restrictions. Covariance-matrix regularization methods are often based
-on the well-known ridge-regression approach of \cite{Hoerl:1970cd}.
-Additionally, \cite{Ching:2012fu}, \cite{Zhang:2010va}, \cite{Ye:2009gd},
-\cite{Ji:2008wp}, \cite*{Guo:2007te}, \cite*{Srivastava:2007ud},
-\cite{Ye:2006vx}, and \cite{Ye:2005uu} have proposed regularization methods to
-improve the estimation of the class covariance matrices by adjusting the
-eigenvalues of the sample covariance matrices to ensure positive
-definiteness. Alternatively, dimension reduction typically transforms the
-original data in the high-dimensional space to a subspace having a much smaller
-dimension such that redundant and noisy information is removed
-\citep*{Fan:2012iq,Ye:2009gd,Bouveyron:2007gx,Ye:2006vx,Ye:2005hd,Antoniadis:2003bk}.
-Finally, \cite*{Mai:2012bf}, \cite{Fan:2012iq}, \cite*{Merchante:2012vk},
-\cite*{Clemmensen:2011kr}, \cite*{Huang:2010ju}, and \cite*{Dudoit:2002ev} have
-proposed restrictions on the population covariance matrices to reduce the number
-of parameters that must be estimated. For instance, \cite{Dudoit:2002ev} have
-assumed that the features within each class are independent so that the
-population covariance matrices are diagonal, resulting in positive-definite
-sample covariance matrices. For excellent overviews of the regularization
-problem in discriminant analysis and other techniques, see \cite{Murphy:2012uq}
-and \cite*{Mkhadri:1997gy}.
+In this paper, we consider the classification of small-sample, high-dimensional
+data, where the number of features $p$ exceeds the training sample size $N$. In
+this setting, well-established classifiers such as linear discriminant analysis
+(\emph{LDA}) and quadratic discriminant analysis (\emph{QDA}) become
+incalculable because the class and pooled covariance matrix estimators are
+singular \citep*{Murphy:2012uq, Bouveyron:2007gx, Mkhadri:1997gy}. To improve
+the estimation of the class covariance matrices estimated in the \emph{QDA}
+classifier and to ensure the covariance matrix estimators are nonsingular,
+\cite{Friedman:1989tm} proposed the regularized discriminant analysis
+(\emph{RDA}) classifier by incorporating a weighted average of the pooled sample
+covariance matrix and the class sample covariance matrix. To further improve the
+estimation of the class covariance matrix and to stabilize its inverse,
+\cite{Friedman:1989tm} also included a regularization component by shrinking the
+covariance matrix estimator towards the identity matrix, which yields a
+nonsingular estimator following the well-known ridge-regression approach of
+\cite{Hoerl:1970cd}. Despite its popularity, the ``borrowing'' operation
+employed in the \emph{RDA} classifier lacks interpretability
+\citep{Bensmail:1996fw}, and the \emph{RDA} classifier is computationally
+inefficient in that its matrix operations are cubic in the number of
+features.\footnote{TODO: Add a couple of references saying RDA is slooooow.} As
+a result, the \emph{RDA} classifier is impraticial for modern high-dimensional
+data sets because its matrix operations are cubic in the number of features, and
+the estimation of the \emph{RDA} classifier's two tuning parameters is
+computationally burdensome.
 
@@
 which is intended for the $p > N$ case and builds on the classic \emph{RDA}
-classifier proposed by \cite{Friedman:1989tm}. In this paper, we study the
-classification performance of the \emph{HDRDA} classifier on six real
-high-dimensional data sets along with a simulation design that generalizes the
-experiments initially conducted by \cite{Guo:2007te}. We demonstrate that the
-\emph{HDRDA} classifier often attains superior classification accuracy to
-several recent classifiers designed for small-sample, high-dimensional data from
-\cite*{Tong:2012hw}, \cite{Witten:2011kc}, \cite*{Pang:2009ik}, and
-\cite{Guo:2007te}. We also include as a benchmark the random forest from
-\cite{Breiman:2001fb} because \cite*{FernandezDelgado:2014ul} have concluded
-that the random forest is often superior to other classifiers in benchmark
-studies. We show that our proposed classifier is competitive and often
+classifier proposed by \cite{Friedman:1989tm}. We reparameterize the \emph{RDA}
+classifier similar to that of \cite*{Hastie:2008dt} and \cite{Halbe:2007ea} and
+employ a biased covariance-matrix estimator that partially pools the individual
+sample covariance matrices from the \emph{QDA} classifier with the pooled sample
+covariance matrix from the \emph{LDA} classifier. We then shrink the resulting
+covariance-matrix estimator towards a scaled identity matrix to ensure positive
+definiteness. We show that the pooling parameter in the \emph{HDRDA} classifier
+determines the contribution of each training observation to the estimation of
+each class covariance matrix, enabling interpretability that has been previously
+lacking with the \emph{RDA} classifier \citep{Bensmail:1996fw}. Our
+parameterization differs from \cite*{Hastie:2008dt} and \cite{Halbe:2007ea} in
+that our formulation allows the flexibility of various covariance-matrix
+estimators proposed in the literature, including a variety of ridge-like
+estimators such as the one proposed by \cite{Srivastava:2007ww}.
+
+Next, we establish that the matrix operations corresponding to the null space of
+the pooled sample covariance matrix are redundant and can be discarded from the
+\emph{HDRDA} decision rule without loss of classificatory information by
+applying reasoning similar to that of \cite{Ye:2006tq}. As a result, we achieve
+a substantial reduction in dimension such that the matrix operations used in the
+\emph{HDRDA} classifier are computationally linear in the number of features.
+Furthermore, we demonstrate that the \emph{HDRDA} decision rule is invariant to
+adjustments to the approximately $p - N$ zero eigenvalues, so that the decision
+rule in the original feature space is equivalent to a decision rule in a lower
+dimension, such that matrix inverses and determinants of relatively small
+matrices can be rapidly computed. Finally, we show that several shrinkage
+methods that are special cases of the \emph{HDRDA} classifier have no effect on
+the approximately $p - N$ zero eigenvalues of the covariance-matrix estimators
+when $p > N$. Such techniques include work from \cite{Srivastava:2007ww},
+\cite{Rao:1971ul}, and other methods studied by \cite{Ramey:2013ji} and
+\citet*{Xu:2009fl}.
+
+We provide an efficient algorithm along with pseudocode to estimate the
+\emph{HDRDA} classifier's tuning parameters in a grid search via
+cross-validation. Timing comparisons between our \emph{HDRDA} implementation in
+the {\tt sparsediscrim} R package available on CRAN and the standard \emph{RDA}
+formulation in the {\tt klaR} R package demonstrate that as the number of
+features increases, the computational runtime of the \emph{HDRDA} classifier is
+drastically smaller than \emph{RDA}. In fact, when $p = 1000$ and $N = 100$, the
+\emph{HDRDA} classifier is \textbf{XXX} times faster on average than the
+\emph{RDA} classifier.
+
+Finally, we study the classification performance of the \emph{HDRDA} classifier
+on six real high-dimensional data sets along with a simulation design that
+generalizes the experiments initially conducted by \cite{Guo:2007te}. We
+demonstrate that the \emph{HDRDA} classifier often attains superior
+classification accuracy to several recent classifiers designed for small-sample,
+high-dimensional data from \cite*{Tong:2012hw}, \cite{Witten:2011kc},
+\cite*{Pang:2009ik}, and \cite{Guo:2007te}. We also include as a benchmark the
+random forest from \cite{Breiman:2001fb} because \cite*{FernandezDelgado:2014ul}
+have concluded that the random forest is often superior to other classifiers in
+benchmark studies. We show that our proposed classifier is competitive and often
 outperforms the random forest in terms of classification accuracy in the
 small-sample, high-dimensional setting.
-
-Our proposed \emph{HDRDA} classifier employs a biased covariance-matrix
-estimator that partially pools the individual sample covariance matrices from
-the \emph{QDA} classifier with the pooled sample covariance matrix from the
-\emph{LDA} classifier. We then shrink the resulting covariance-matrix estimator
-towards a scaled identity matrix to ensure positive definiteness. By
-reparameterizing the \emph{RDA} classifier similar to that of
-\cite*{Hastie:2008dt} and \cite{Halbe:2007ea}, we show that the pooling
-parameter in the \emph{HDRDA} classifier determines the contribution of each
-training observation to the estimation of each class covariance matrix, enabling
-interpretability that has been previously lacking with the \emph{RDA} classifier
-\citep{Bensmail:1996fw}.
-
-Next, we achieve a substantial reduction in dimension that greatly reduces the
-computational effort during model selection when $p \gg N$ by applying reasoning
-similar to that of \cite{Ye:2006tq}. We establish that the matrix operations
-corresponding to the null space of the pooled sample covariance matrix are
-redundant and can be discarded from the \emph{HDRDA} decision rule without loss
-of classificatory information. Moreover, because the \emph{HDRDA} decision rule
-is invariant to adjustments to the zero eigenvalues, the decision rule in the
-original feature space is equivalent to a decision rule in a lower dimension,
-such that matrix inverses and determinants of relatively small matrices can be
-rapidly computed. Also, we provide an efficient algorithm along with pseudocode
-to estimate the \emph{HDRDA} classifier's tuning parameters in a grid search via
-cross-validation. Finally, we show that several shrinkage methods that are
-special cases of the \emph{HDRDA} classifier have no effect on the zero
-eigenvalues of the covariance-matrix estimators. Such techniques include work
-from \cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and other methods studied by
-\cite{Ramey:2013ji} and \citet*{Xu:2009fl}.
 

### 2016-10-25 timing-comparison.r
Updated timing comparisons. #26
Timing comparison across several values of `p`.

@@ -5,8 +5,7 @@ library(microbenchmark)
 load.project()
 
 # Number of timing-comparison replications to perform
-# num_reps <- 50
-num_reps <- 5
+num_reps <- 100
 
 # Grid Size (m x m)
 grid_size <- 5
@@ -14,39 +13,31 @@ grid_size <- 5
 # Simulation Parameters
 K <- 4
 n <- rep.int(x=25, times=K)
-# p <- c(10, 100, 1000, 10000)
-p <- c(10, 20)
-
-# TODO: Update means
-means <- lapply(c(-3, -1, 1, 3), rep.int, times=p)
-cov_mat <- diag(p)
-
-# Generates multivariate normal data
-set.seed(42)
-data <- simdata_normal(n=n, mean=means, cov=cov_mat)
+feature_dims <- c(
+  25, 50, 75,
+  seq(100, 1000, by=100),
+  seq(1500, 10000, by=500)
+)
 
 # klaR's Implementation of Friedman (1989)
 rda_model_selection <- function(x, y, num_tuning=5) {
   tuning_grid <- expand.grid(lambda=seq(0, 1, length=num_tuning),
                              gamma=seq(0, 1, length=num_tuning))
+   rda_errors <- vector(mode="list", length=nrow(tuning_grid))
+   for (i in seq_len(nrow(tuning_grid))) {
+     rda_errors[[i]] <- try({
+       rda_out <- rda(x=x,
+                      grouping=y,
+                      lambda=tuning_grid$lambda[i],
+                      gamma=tuning_grid$gamma[i],
+                      estimate.error=FALSE)
+       rda_out$error.rate['crossval']
+     })
+   }
 
-  rda_errors <- vector(mode="list", length=nrow(tuning_grid))
-
-  for (i in seq_len(nrow(tuning_grid))) {
-    rda_errors[[i]] <- try({
-      rda_out <- rda(x=x,
-                     grouping=y,
-                     lambda=tuning_grid$lambda[i],
-                     gamma=tuning_grid$gamma[i],
-                     estimate.error=FALSE)
-      rda_out$error.rate['crossval']
-    })
-  }
-
-  rda_errors[!sapply(rda_errors, is.numeric)] <- NA
-
-  tuning_grid[which.min(unlist(rda_errors)), ]
-}  
+   rda_errors[!sapply(rda_errors, is.numeric)] <- NA
+   tuning_grid[which.min(unlist(rda_errors)), ]
+}
 
 # HDRDA
[…]

### 2016-10-25 timing-comparison.r
klaR::rda's default model selection in timings #26
Added `klaR::rda`'s default Nelder-Mead model selection to the timing comparisons.

@@ -20,17 +20,18 @@ feature_dims <- c(
 )
 
 # klaR's Implementation of Friedman (1989)
-rda_model_selection <- function(x, y, num_tuning=5) {
+# Uses a grid model selection
+rda_grid <- function(x, y, num_tuning=5) {
   tuning_grid <- expand.grid(lambda=seq(0, 1, length=num_tuning),
                              gamma=seq(0, 1, length=num_tuning))
    rda_errors <- vector(mode="list", length=nrow(tuning_grid))
    for (i in seq_len(nrow(tuning_grid))) {
      rda_errors[[i]] <- try({
-       rda_out <- rda(x=x,
-                      grouping=y,
-                      lambda=tuning_grid$lambda[i],
-                      gamma=tuning_grid$gamma[i],
-                      estimate.error=FALSE)
+       rda_out <- klaR::rda(x=x,
+                            grouping=y,
+                            lambda=tuning_grid$lambda[i],
+                            gamma=tuning_grid$gamma[i],
+                            estimate.error=FALSE)
        rda_out$error.rate['crossval']
      })
    }
@@ -39,8 +40,17 @@ rda_model_selection <- function(x, y, num_tuning=5) {
    tuning_grid[which.min(unlist(rda_errors)), ]
 }
 
+# klaR's Implementation of Friedman (1989)
+# Uses a Nelder-Mead model selection
+rda_nelder_mead <- function(x, y) {
+  rda_out <- klaR::rda(x=x,
+                       grouping=y,
+                       estimate.error=FALSE)
+  rda_out$regularization
+}
+
 # HDRDA
-hdrda_model_selection <- function(x, y, num_tuning=5) {
+hdrda_grid <- function(x, y, num_tuning=5) {
   hdrda_cv(x=x,
            y=y,
            num_lambda=num_tuning,
@@ -59,8 +69,9 @@ timing_results <- lapply(feature_dims, function(p) {
   data <- simdata_normal(n=n, mean=means, cov=cov_mat)
 
   mb_results <- microbenchmark(
-    rda_model_selection(x=data$x, y=data$y, num_tuning=grid_size),
-    hdrda_model_selection(x=data$x, y=data$y, num_tuning=grid_size),
+    rda_grid(x=data$x, y=data$y, num_tuning=grid_size),
+    rda_nelder_mead(x=data$x, y=data$y),
+    hdrda_grid(x=data$x, y=data$y, num_tuning=grid_size),
[…]

### 2016-10-26 timing-comparison.r
Updated timing comparison settings. #26
The maximum value of `p` is now 3K instead of 10K.
Why? See the analysis in #26.

@@ -5,7 +5,7 @@ library(microbenchmark)
 load.project()
 
 # Number of timing-comparison replications to perform
-num_reps <- 100
+num_reps <- 50
 
 # Grid Size (m x m)
 grid_size <- 5
@@ -16,7 +16,7 @@ n <- rep.int(x=25, times=K)
 feature_dims <- c(
   25, 50, 75,
   seq(100, 1000, by=100),
-  seq(1500, 10000, by=500)
+  seq(1500, 3000, by=500)
 )
 
 # klaR's Implementation of Friedman (1989)

### 2016-10-26 requirements.r
Added sortinghat to requirements

@@ -4,7 +4,7 @@ biocLite("multtest", ask=FALSE)
 cran_packages <- c('futile.logger', 'reshape2', 'plyr', 'ggplot2', 'stringr',
                    'rda', 'penalizedLDA', 'xtable', 'Hmisc', 'mixtools',
                    'gridExtra', 'sparseLDA', 'e1071', 'klaR', 'kernlab',
-                   'devtools', 'randomForest', 'ProjectTemplate')
+                   'devtools', 'randomForest', 'ProjectTemplate', 'sortinghat')
 install.packages(cran_packages, dep=TRUE)
 
 github_packages <- c('ramhiser/datamicroarray', 'ramhiser/sparsediscrim',

### 2016-10-26 timing-comparison.r

@@ -5,7 +5,7 @@ library(microbenchmark)
 load.project()
 
 # Number of timing-comparison replications to perform
-num_reps <- 50
+num_reps <- 5
 
 # Grid Size (m x m)
 grid_size <- 5
@@ -13,11 +13,12 @@ grid_size <- 5
 # Simulation Parameters
 K <- 4
 n <- rep.int(x=25, times=K)
-feature_dims <- c(
-  25, 50, 75,
-  seq(100, 1000, by=100),
-  seq(1500, 3000, by=500)
-)
+#feature_dims <- c(
+#  25, 50, 75,
+#  seq(100, 1000, by=100),
+# seq(1500, 3000, by=500)
+#)
+feature_dims <- c(25, 50)
 
 # klaR's Implementation of Friedman (1989)
 # Uses a grid model selection

### 2016-10-26 requirements.r
Added microbenchmark package to requirements

@@ -4,7 +4,7 @@ biocLite("multtest", ask=FALSE)
 cran_packages <- c('futile.logger', 'reshape2', 'plyr', 'ggplot2', 'stringr',
                    'rda', 'penalizedLDA', 'xtable', 'Hmisc', 'mixtools',
                    'gridExtra', 'sparseLDA', 'e1071', 'klaR', 'kernlab',
-                   'devtools', 'randomForest', 'ProjectTemplate', 'sortinghat')
+                   'devtools', 'randomForest', 'ProjectTemplate', 'sortinghat', 'microbenchmark')
 install.packages(cran_packages, dep=TRUE)
 
 github_packages <- c('ramhiser/datamicroarray', 'ramhiser/sparsediscrim',

### 2016-10-26 timing-comparison.r
Somehow committed over timing comparison settings. Fixed.

@@ -5,7 +5,7 @@ library(microbenchmark)
 load.project()
 
 # Number of timing-comparison replications to perform
-num_reps <- 5
+num_reps <- 50
 
 # Grid Size (m x m)
 grid_size <- 5
@@ -13,12 +13,11 @@ grid_size <- 5
 # Simulation Parameters
 K <- 4
 n <- rep.int(x=25, times=K)
-#feature_dims <- c(
-#  25, 50, 75,
-#  seq(100, 1000, by=100),
-# seq(1500, 3000, by=500)
-#)
-feature_dims <- c(25, 50)
+feature_dims <- c(
+  25, 50, 75,
+  seq(100, 1000, by=100),
+  seq(1500, 3000, by=500)
+)
 
 # klaR's Implementation of Friedman (1989)
 # Uses a grid model selection

### 2016-11-12 reports/paper-hdrda.Rnw

@@
 \author[dmy]{Dean M. Young}
-\address[jar]{uStudio, Inc.}
+\address[jar]{uStudio Labs}
 \address[cks]{Myeloma Institute, University of Arkansas for Medical Sciences}
@@
   \Sigma}_k$. Following \cite{Friedman:1989tm}, we select $\lambda$ and $\gamma$
-from a grid of candidate models via cross-validation. We provide an
+from a grid of candidate models via cross-validation \citep{Hastie:2008dt}. We provide an
 implementation of \eqref{eq:hdrda} in the {\tt hdrda} function contained in the
@@
 
+\section{Timing Comparisons between RDA and HDRDA}
+\label{sec:timing-comparisons}
+
+In this section, we demonstrate that the computational performance of the
+training and model selection employed in the \emph{HDRDA} classifier is
+substantially faster than that of the \emph{RDA} classifier on small-sample,
+high-dimensional data sets. The relative difference in runtime between the two
+classifiers drastically increases as the number of features increases. To
+compare the two classifiers, we generate 25 observations from each of $K=4$
+multivariate normal populations with mean vectors $\bm \mu_1 = -3 \cdot \bm
+1_p$, $\bm \mu_2 = -\bm 1_p$, $\bm \mu_3 = \bm 1_p$, and $\bm \mu_4 = 3 \cdot
+\bm 1_p$. We set the covariance matrix of each population to the $p \times p$
+identity matrix. For each data set generated, we estimate the $\lambda$ and
+$\gamma$ parameters for both classifiers using a grid of 21 equidistant
+candidate values between 0 and 1, inclusively. We set $\alpha_k = 1 - \gamma$,
+$k = 1, \ldots, K$, in the \emph{HDRDA} classifier to resemble
+\citeauthor{Friedman:1989tm}'s \emph{RDA} parameterization. At each pair of
+$\lambda$ and $\gamma$, we compute the $V$-fold cross-validation error rate with
+$V = 10$ \citep{Hastie:2008dt}. The optimal model is the pair of $\lambda$ and
+$\gamma$ that minimizes the $V$-fold cross-validation error rate.
+
+Here, our goal is to compare the computational runtime of the model selection
+for both classifiers. We compare the runtime of both classifiers by increasing
+the number of features from $p = 500$ to $p = 5000$ in increments of 500. We
+generate 100 data sets for each value of $p$ and compute the training and model
+selection runtime of both classifiers. Our timing comparisons are based on our
+\emph{HDRDA} implementation in the {\tt sparsediscrim} R package with the
+standard \emph{RDA} implementation in the {\tt klaR} R package. All timing
+comparisons were conducted on an Amazon Elastic Compute Cloud (EC2) {\tt
+  c4.4xlarge} instance using version 3.3.1 of the open-source statistical
+software {\tt R}. Our analyses can be reproduced using the code available in the
+\url{https://github.com/ramhiser/paper-hdrda} repository.
+
+\subsection{Timing Comparison Results}
+
+TODO
+
 \section{Classification Study}
@@
 recently proposed for small-sample, high-dimensional data along with the
-random-forest classifier from \cite{Breiman:2001fb} using version 3.2.2 of the
+random-forest classifier from \cite{Breiman:2001fb} using version 3.3.1 of the
 open-source statistical software {\tt R}. Within our study, we included
@@
 their default settings. Below, we refer to each classifier by the first author's
-surname. All simulations were conducted on an Amazon Elastic Compute Cloud (EC2)
-{\tt c4.4xlarge} instance. Our analyses can be reproduced using the code
-available in the \url{https://github.com/ramhiser/paper-hdrda} repository.
+surname. All simulations were conducted on an Amazon EC2 {\tt c4.4xlarge}
+instance. Our analyses can be reproduced using the code available in the
+\url{https://github.com/ramhiser/paper-hdrda} repository.
 
@@
 values $10^{-1}, \ldots, 10^4$, and $10^5$. We selected optimal estimates of
-$\lambda$ and $\gamma$ using $V$-fold cross-validation with $V = 10$
-\citep{Hastie:2008dt}. For the second model, we set $\alpha_k = 1 - \gamma$, $k
-= 1, \ldots, K$, to resemble \citeauthor{Friedman:1989tm}'s parameterization,
-and we estimated both $\lambda$ and $\gamma$ from a grid of 21 equidistant
-candidate values between 0 and 1, inclusively.
+$\lambda$ and $\gamma$ using $V$-fold cross-validation with $V = 10$. For the
+second model, we set $\alpha_k = 1 - \gamma$, $k = 1, \ldots, K$, to resemble
+\citeauthor{Friedman:1989tm}'s parameterization, and we estimated both $\lambda$
+and $\gamma$ from a grid of 21 equidistant candidate values between 0 and 1,
+inclusively.
+
+We did not include the \emph{RDA} classifier in our classification study because
+its training runtime was prohobitively slow on high-dimensional data in our
+preliminary experiments. As shown in Section \ref{sec:timing-comparisons}, the
+runtime of the \emph{RDA} classifier was drastically larger than that of the
+\emph{HDRDA} classifier for a tuning grid of size $25 = 5 \times
+5$. Consequently, a fair comparison between the \emph{RDA} and \emph{HDRDA}
+classifiers would require model selection of $441 = 21 \times 21$ different
+pairs of tuning parameters in the \emph{RDA} classifier. A tuning grid of this
+size yielded excessively slow training runtimes for the \emph{RDA}
+implementation from the {\tt klaR} R package.
 

### 2016-12-02 reports/paper-hdrda.Rnw
Summarized timings along with bootstrap comparison

@@
   spread(data_set, Table_Cells)
+@
+
+<<munge_timing_comparisons>>=
+# HDRDA vs RDA Timing Comparison Results
+
+# Benchmarks are given in nanoseconds, which are converted to seconds
+timings <- bind_rows(lapply(timing_results, as.data.frame), .id="p") %>%
+  mutate(p=as.integer(p),
+         expr=as.character(expr),
+         Classifier=ifelse(startsWith(expr, "hdrda"), "HDRDA", "RDA"),
+         Classifier=factor(Classifier),
+         time=time / 1e9) %>%
+  select(-expr) %>%
+  filter(p %% 500 == 0) %>%
+  tbl_df %>%
+  group_by(p, Classifier) %>%
+  mutate(wls_weights=1 / sd(time)) %>%
+  ungroup
+
+# Create DF for two facets in same plot
+# First plot shows HDRDA vs RDA
+# Second plot zooms in on HDRDA results
+timings$Plot <- "Comparison of HDRDA and RDA"
+timings_hdrda <- timings %>%
+    filter(Classifier == "HDRDA") %>%
+    mutate(Plot="HDRDA")
+timings_combined <- bind_rows(timings, timings_hdrda)
+
+# Build data frame to show relative timings in table
+timing_ratios <- timings %>%
+    group_by(p, Classifier) %>%
+    summarize(median_time=median(time)) %>%
+    tidyr::spread(Classifier, median_time) %>%
+      mutate(ratio_time=RDA / HDRDA) %>%
+      select(-HDRDA, -RDA)
+@
+
+<<bootstrap_relative_timing_comparisons>>=
+ratio_means <- function(df, num_reps=1000) {
+  bootstrap_reps <- replicate(n=num_reps, {
+    df %>%
+      group_by(Classifier) %>%
+      sample_frac(size=1.0, replace=TRUE) %>%
+      summarize(mean_time=mean(time)) %>%
+      tidyr::spread(Classifier, mean_time) %>%
+      mutate(ratio_means=RDA / HDRDA) %>%
+      select(-HDRDA, -RDA)
+  }, simplify=FALSE)
+  bind_rows(bootstrap_reps)
+}
+
+set.seed(42)
+timing_ratios_bootstrap <- timings %>%
+    mutate(p=factor(p)) %>%
+    group_by(p) %>%
+    do(ratio_means(., num_reps=1000))
+
+timing_bootstrap_summary <- timing_ratios_bootstrap %>%
+  group_by(p) %>%
+  summarize(avg_ratio=mean(ratio_means),
+            lower_ci=quantile(ratio_means, probs=0.025),
+            upper_ci=quantile(ratio_means, probs=0.975))
+bootstrap_summary_p500 <- timing_bootstrap_summary %>%
+  filter(p == "500")
+bootstrap_summary_p5000 <- timing_bootstrap_summary %>%
+  filter(p == "5000")
+
 @

### 2016-12-02 reports/paper-hdrda.Rnw
Plots of timing comparisons

@@
 TODO
+\[ \left[\text{Insert Figure \ref{fig:timing-results} approximately here }\right] \]
+
+\[ \left[\text{Insert Figure \ref{fig:timing-comparison-bootstrap} approximately here }\right] \]
 
@@
 \begin{figure}
+<<timing_results, fig.width=8, fig.height=8, dpi=200>>=
+p <- ggplot(timings_combined, aes(x=p, y=time, color=Classifier))
+p <- p + geom_point() + geom_jitter(width=140, height=0) + facet_wrap(~ Plot, scales="free_y")
+p <- p + geom_smooth(method="lm", formula=y ~ poly(x, 2), se=FALSE)
+p <- p + scale_color_manual(values=c("#D55E00", "#0072B2"))
+p <- p + xlab("Number of Features") + ylab("Model Selection Time (seconds)")
+p + theme_bw()
+@
+\caption{Timing comparisons (in seconds) between HDRDA and RDA classifiers.}
+\label{fig:timing-results}
+\end{figure}
+
+\clearpage
+
+\begin{figure}
+<<relative_timings, fig.width=8, fig.height=8, dpi=200>>=
+p <- ggplot(timing_ratios_bootstrap, aes(ratio_means, fill=p))
+p <- p + geom_histogram() + facet_wrap(~ p, scale="free_x")
+p <- p + xlab("Ratio of Mean RDA Runtime to Mean HDRDA Runtime") + ylab("Bootstrap Replications")
+p + theme_bw()
+@
+\caption{Distribution of ratios of mean \emph{RDA} runtime to mean \emph{HDRDA}
+  runtime across 1000 bootstrap replications.}
+\label{fig:timing-comparison-bootstrap}
+\end{figure}
+
+\clearpage
+
+\begin{figure}
 <<sim_results, fig.width=8, fig.height=8, dpi=200>>=

### 2016-12-02 reports/paper-hdrda.Rnw
Describe timing comparisons

@@
 \author[dmy]{Dean M. Young}
-\address[jar]{uStudio Labs}
+\address[jar]{Novi Labs}
 \address[cks]{Myeloma Institute, University of Arkansas for Medical Sciences}
@@
 identity matrix. For each data set generated, we estimate the $\lambda$ and
-$\gamma$ parameters for both classifiers using a grid of 21 equidistant
-candidate values between 0 and 1, inclusively. We set $\alpha_k = 1 - \gamma$,
-$k = 1, \ldots, K$, in the \emph{HDRDA} classifier to resemble
+$\gamma$ parameters for both classifiers using a grid of 5 equidistant candidate
+values between 0 and 1, inclusively. We set $\alpha_k = 1 - \gamma$, $k = 1,
+\ldots, K$, in the \emph{HDRDA} classifier to resemble
 \citeauthor{Friedman:1989tm}'s \emph{RDA} parameterization. At each pair of
-$\lambda$ and $\gamma$, we compute the $V$-fold cross-validation error rate with
-$V = 10$ \citep{Hastie:2008dt}. The optimal model is the pair of $\lambda$ and
-$\gamma$ that minimizes the $V$-fold cross-validation error rate.
+$\lambda$ and $\gamma$, we compute the 10-fold cross-validation error rate
+\citep{Hastie:2008dt}. Then, we select the optimal model that minimizes the 10-fold
+cross-validation error rate.
 
 Here, our goal is to compare the computational runtime of the model selection
-for both classifiers. We compare the runtime of both classifiers by increasing
-the number of features from $p = 500$ to $p = 5000$ in increments of 500. We
-generate 100 data sets for each value of $p$ and compute the training and model
-selection runtime of both classifiers. Our timing comparisons are based on our
-\emph{HDRDA} implementation in the {\tt sparsediscrim} R package with the
-standard \emph{RDA} implementation in the {\tt klaR} R package. All timing
-comparisons were conducted on an Amazon Elastic Compute Cloud (EC2) {\tt
-  c4.4xlarge} instance using version 3.3.1 of the open-source statistical
-software {\tt R}. Our analyses can be reproduced using the code available in the
+for both classifiers as a function of the number of features $p$. We compare the
+runtime of both classifiers by increasing the number of features from $p = 500$
+to $p = 5000$ in increments of 500. We generate 100 data sets for each value of
+$p$ and compute the training and model selection runtime of both
+classifiers. Our timing comparisons are based on our \emph{HDRDA} implementation
+in the {\tt sparsediscrim} R package and the standard \emph{RDA} implementation
+in the {\tt klaR} R package. All timing comparisons were conducted on an Amazon
+Elastic Compute Cloud (EC2) {\tt c4.4xlarge} instance using version 3.3.1 of the
+open-source statistical software {\tt R}. Our timing comparisons can be
+reproduced with the code available in the
 \url{https://github.com/ramhiser/paper-hdrda} repository.
@@
 
-TODO
+In Figure \ref{fig:timing-results}, we plot the runtime of the model selections
+for both the \emph{HDRDA} and \emph{RDA} classifiers as a function of the number
+of features $p$. In the left panel of Figure \ref{fig:timing-results}, we
+observe that the \emph{HDRDA} classifier is substantially faster than the
+\emph{RDA} classifier as $p$ increases. Figure \ref{fig:timing-results} suggests
+that the usage of a matrix inverse and determinant in the {\tt klaR} R package's
+discriminant function yielded training and model-selection timings that exceeded
+linear growth in $p$. Because our implementation in the \emph{HDRDA} classifier
+removes the inverse and determinants from the classifier, the \emph{HDRDA}
+classifier is computationally more efficient than the \emph{RDA} classifier,
+especially as $p$ increases. The variance in the \emph{RDA} classifier's runtime
+also appeared to increase as $p$ increased. To quantify how much beyond linear
+growth in $p$, we fit a regression line quadratic in $p$ to the \emph{RDA}
+runtimes and plotted the fit in Figure \ref{fig:timing-results}. We attempted to
+fit a cubic polynomial in $p$ to the \emph{RDA} timings, but the cubic term was
+insignificant with significant quadratic and linear terms. We also fit two
+polynomials to the \emph{RDA} timings with orders 2.807 and 2.376, which
+correspond to the asymptotic computational runtime of two well-known algorithms
+for computing a matrix inverse: the Strassen and Coppersmith-Winograd
+algorithms, respectively \footnote{TODO: Citations needed}. In both cases, the
+polynomials were statistically significiant at the 0.05 level. We displayed the
+quadratic fit in Figure \ref{fig:timing-results} because it fit the timings best
+for $p = 500$ to $p = 2000$. The fits for $2000 \le p \le 5000$ were similar.
+
+In the left panel of Figure \ref{fig:timing-results}, we also fit a linear
+regression model that was linear in $p$ to the \emph{HDRDA} runtimes. In the
+right panel, we repeated the same scatterplot and linear fit with the timings
+restricted to the observed range of \emph{HDRDA} timings. The outliers in the
+\emph{HDRDA} timings were primarily due to the inherent computing variability in
+our {\tt c4.4xlarge} instance.
+
+Clearly, the model training and selection employed in \emph{HDRDA} is
+substantially faster than that of \emph{RDA}. To investigate the relative timing
+comparisons between the two classifiers, we consider the ratio of mean timings
+of the \emph{RDA} classifier to the \emph{HDRDA} classifier for each value of
+$p$. We employed nonparametric bootstrapping\footnote{TODO: Citation needed} to
+estimate the mean ratio along with 95\% confidence intervals. In Figure
+\ref{fig:timing-comparison-bootstrap}, the bootstrap sampling distributions for
+the ratio of mean timings are given. First, we observe that the mean relative
+timings increases as $p$ increases. For smaller dimensions, the relative
+difference in computing is sizable with the average ratio of the mean timings
+equal to \Sexpr{bootstrap_summary_p500$avg_ratio} for $p = 500$ and a 95\%
+confidence interval of (\Sexpr{bootstrap_summary_p500$lower_ci},
+\Sexpr{bootstrap_summary_p500$upper_ci}). Furthermore, the ratio of mean
+computing times suggests that the \emph{RDA} classifier is impractical for
+higher dimensions. For instance, when $p = 5000$ the ratio of mean computing
+times increases to \Sexpr{bootstrap_summary_p5000$avg_ratio} with a 95\%
+confidence interval of (\Sexpr{bootstrap_summary_p5000$lower_ci},
+\Sexpr{bootstrap_summary_p5000$upper_ci}).
+
 \[ \left[\text{Insert Figure \ref{fig:timing-results} approximately here }\right] \]
@@
 
-The \emph{HDRDA} implementation of \eqref{eq:hdrda-decomposed} in the {\tt
-  sparsediscrim} R package was comparable to that of the diagonal classifiers in
-terms of computational runtime. Given the \emph{RDA} classifier from
-\cite{Friedman:1989tm} has been shown to have excellent performance in the
-high-dimensional setting \citep{Webb:2011vu} but is limited by its
-computationally intense model-selection procedure, our work enables the
-\emph{RDA} classifier to be employed with small-sample, high-dimensional data in
-practice. This result is reassuring because the \emph{RDA} classifier remains
-widely popular in the literature. In fact, variants of the RDA classifier have
-been applied to microarray data
-\citep{Ching:2012fu,Li:2012ev,Tai:2007bk,Guo:2007te}, facial recognition
+We demonstrated that the \emph{HDRDA} implementation of
+\eqref{eq:hdrda-decomposed} in the {\tt sparsediscrim} R package can be used in
+practice with high-dimensional data sets. For instance, in our timing
+comparisons we demonstrated that training and model selection could be employed
+on data sets having 5,000 features in 3 seconds on average. We contrast our
+implementation to the the poor computational performance of the in the {\tt
+  klaR} R package, which took approximately 25 minutes on average to train the
+\emph{RDA} classifier on a data set having 5,000 features. Given the \emph{RDA}
+classifier has been shown to have excellent performance in the high-dimensional
+setting \citep{Webb:2011vu} but is limited by its computationally intense
+model-selection procedure, our work enables the \emph{RDA} classifier to be
+employed with small-sample, high-dimensional data in practice. This result is
+reassuring because the \emph{RDA} classifier remains widely popular in the
+literature. In fact, variants of the RDA classifier have been applied to
+microarray data \citep{Ching:2012fu,Li:2012ev,Tai:2007bk,Guo:2007te}, facial
+recognition
 \citep{Ching:2012fu,Zhang:2010va,Dai:2007vd,Lu:2005hq,Pima:2004vw,Lu:2003we},

### 2016-12-02 reports/paper-hdrda.Rnw
Updates to the intro

@@
 employed in the \emph{RDA} classifier lacks interpretability
-\citep{Bensmail:1996fw}, and the \emph{RDA} classifier is computationally
-inefficient in that its matrix operations are cubic in the number of
-features.\footnote{TODO: Add a couple of references saying RDA is slooooow.} As
-a result, the \emph{RDA} classifier is impraticial for modern high-dimensional
-data sets because its matrix operations are cubic in the number of features, and
-the estimation of the \emph{RDA} classifier's two tuning parameters is
-computationally burdensome.
+\citep{Bensmail:1996fw}. Furthermore, the \emph{RDA} classifier is impractical
+for high-dimensional data sets because it computes the inverse and determinant
+of the covariance matrices for each class. Both matrix calculations are
+computationally expensive because the number of operations grows at a polynomial
+rate in the number of features. Moreover, the \emph{RDA} classifier model
+selection of its two tuning parameters is computationally burdensome because the
+matrix inverse and determinant of the covariance matrices across multiple
+cross-validation folds for each class and for each candidate tuning-parameter
+model.
 
@@
 features increases, the computational runtime of the \emph{HDRDA} classifier is
-drastically smaller than \emph{RDA}. In fact, when $p = 1000$ and $N = 100$, the
-\emph{HDRDA} classifier is \textbf{XXX} times faster on average than the
-\emph{RDA} classifier.
+drastically smaller than \emph{RDA}. In fact, when $p = 5000$, we show that the
+\emph{HDRDA} classifier is \Sexpr{bootstrap_summary_p5000$avg_ratio} times
+faster on average than the \emph{RDA} classifier. In this scenario, the
+\emph{HDRDA} classifier's training and model selection requires 3 seconds on
+average, while that of the \emph{RDA} classifier requires 25 minutes on average.
 
@@
 classifier is computationally more efficient than the \emph{RDA} classifier,
-especially as $p$ increases. The variance in the \emph{RDA} classifier's runtime
-also appeared to increase as $p$ increased. To quantify how much beyond linear
-growth in $p$, we fit a regression line quadratic in $p$ to the \emph{RDA}
-runtimes and plotted the fit in Figure \ref{fig:timing-results}. We attempted to
-fit a cubic polynomial in $p$ to the \emph{RDA} timings, but the cubic term was
+especially as $p$ increases. We contrast our implementation to the the poor
+computational performance of the in the {\tt klaR} R package, which took
+approximately 25 minutes on average to train the \emph{RDA} classifier on a data
+set having 5,000 features. To quantify how much beyond linear growth in $p$, we
+fit a regression line quadratic in $p$ to the \emph{RDA} runtimes and plotted
+the fit in Figure \ref{fig:timing-results}. We attempted to fit a cubic
+polynomial in $p$ to the \emph{RDA} timings, but the cubic term was
 insignificant with significant quadratic and linear terms. We also fit two

### 2016-12-21 reports/paper-hdrda.Rnw
Update paper-hdrda.Rnw
Phil's updates

@@
 
-We develop a new and computationally efficient implementation of
-\citeauthor{Friedman:1989tm}'s popular regularized discriminant analysis (RDA)
-classifier in the small-sample, high-dimensional setting. Our proposed
-high-dimensional RDA (HDRDA) classifier modifies \citeauthor{Friedman:1989tm}'s
-original formulation to achieve interpretability previously lacking. We are able
+We develop a new and computationally efficient regularized discriminant analysis (RDA)
+classifier for the small-sample, high-dimensional setting.  We are able
 to interpret the biased covariance matrix estimator employed in the HDRDA
 classifier as one that is estimated from all training observations regardless of
-class, solidifying the ``borrowing'' operation \citeauthor{Friedman:1989tm}'s
-RDA is claimed to have. Moreover, our formulation also differs from
+class. Moreover, our classifier also differs from
 \citeauthor{Friedman:1989tm}'s in that we show our classifier is equivalent to
-one in a reduced feature space with dimension approximately equal to the
+a classifier in a reduced feature space with dimension approximately equal to the
 training sample size. As a result, the matrix operations used in the HDRDA
 classifier are computationally linear in the number of features, making HDRDA
-well-suited for high-dimensional classification in practice. Timing comparisons
+well-suited for high-dimensional classification in practice. We demonstrate that HDRDA is often superior to several
+sparse and regularized classifiers in terms of classification accuracy with
+three artificial and six real high-dimensional data sets.  Also, timing comparisons
 between our HDRDA implementation in the {\tt sparsediscrim} R package and the
@@
 number of features increases, the computational runtime of HDRDA is drastically
-smaller than RDA. We also demonstrate that HDRDA is often superior to several
-sparse and regularized classifiers in terms of classification accuracy with
-three artificial and six real high-dimensional data sets.
+smaller than RDA.
 
@@
 Here, we present the high-dimensional \emph{RDA} (\emph{HDRDA}) classifier,
-which is intended for the $p > N$ case and builds on the classic \emph{RDA}
-classifier proposed by \cite{Friedman:1989tm}. We reparameterize the \emph{RDA}
+which is intended for the case when $p > N$ . We reparameterize the \emph{RDA}
 classifier similar to that of \cite*{Hastie:2008dt} and \cite{Halbe:2007ea} and
@@
 
-We provide an efficient algorithm along with pseudocode to estimate the
+We also provide an efficient algorithm along with pseudocode to estimate the
 \emph{HDRDA} classifier's tuning parameters in a grid search via
@@
 
-Perhaps the most well-known of the ridge-like estimators is from
-\cite{Friedman:1989tm}, who proposed regularized discriminant analysis (RDA) by
-incorporating a weighted average of the sample covariance matrix $\widehat{ \bm
-  \Sigma}_k$ and the pooled sample covariance matrix $\widehat{\bm \Sigma}$ to
-estimate $\bm \Sigma_k$ with
-\begin{linenomath*}
-\begin{align}
-    \widehat{\bm \Sigma}_k(\lambda) = n_k^{-1}(\lambda) \bm S_k(\lambda) \quad
-    (k = 1, \ldots, K),\label{eq:sig-lambda}
-\end{align}
-\end{linenomath*}
-where $\lambda \in [0, 1]$, $n_k(\lambda) = (1 - \lambda) n_k + \lambda N$, $
-\bm S_k = n_k \widehat{\bm \Sigma}_k$, $\bm S = \sum_{k=1}^K \bm S_k$, and $ \bm
-S_k(\lambda) = (1 - \lambda) \bm S_k + \lambda \bm S$. If $\lambda = 0$,
-\eqref{eq:sig-lambda} is equal to $\widehat{\bm \Sigma}_k$ in \eqref{eq:qda}.
-Alternatively, if $\lambda = 1$, then \eqref{eq:sig-lambda} reduces to
-\eqref{eq:pooled-cov}. To further improve the estimation of $\bm \Sigma_k$ and
-to stabilize the inverse of \eqref{eq:sig-lambda}, \cite{Friedman:1989tm}
+\cite{Friedman:1989tm}
 proposed the biased covariance-matrix estimator
@@
 shrinkage of \eqref{eq:sig-rda} towards $\bm I_p$ and is weighted by the average
-of the eigenvalues of \eqref{eq:sig-lambda}. Finally, \cite{Friedman:1989tm}
+of the eigenvalues of \eqref{eq:sig-lambda}. \cite{Friedman:1989tm}
 established the RDA classifier by substituting \eqref{eq:sig-rda} into
@@
 
-Our proposed \emph{HDRDA} classifier resembles \eqref{eq:sig-rda} but yields an
+Our proposed \emph{HDRDA} classifier yields an
 interpretation of the pooling parameter $\lambda$ as discussed above. Moreover,
@@
 
-Thus, we can avoid the time-consuming inverses and determinants of $p \times p$
+Using Theorem 1, we can avoid the time-consuming inverses and determinants of $p \times p$
 covariance matrices in \eqref{eq:hdrda} and instead calculate these same
@@
 
-The application of the Sherman-Woodbury formula requires that $\bm \Gamma_k$ be
-nonsingular. This assumption is not met when $(\lambda, \gamma) = (0, 0)$
-because $\bm \Gamma_k = \bm 0_q$. Here, we use the formulation in
-\eqref{eq:hdrda-decomposed} instead. Also, notice that if $\alpha_k$ is constant
+Also, notice that if $\alpha_k$ is constant
 across the $K$ classes, then $\bm \Gamma_k$ in \eqref{eq:Gamma_k} is independent
@@
 W_k^{-1}$ and $|\bm W_k|$ can be simply updated for different pairs of $\lambda$
-and $\gamma$ without repeating the costly computations. When calculating the
-discriminant scores of each test observation in $\bm X_{test}$, rather than
-calculating each score individually, we can more efficiently calculate the
-quadratic forms via a single matrix operation. That is, rather than calculating
+and $\gamma$ without repeating the costly computations. Also, rather than calculating
 $(\bm x - \xbar_k)^{T} \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm x - \xbar_k)$
@@
 high-dimensional data sets. The relative difference in runtime between the two
-classifiers drastically increases as the number of features increases. To
-compare the two classifiers, we generate 25 observations from each of $K=4$
+classifiers drastically increases as $p$ increases. To
+compare the two classifiers, we generated 25 observations from each of $K=4$
 multivariate normal populations with mean vectors $\bm \mu_1 = -3 \cdot \bm
@@
 \bm 1_p$. We set the covariance matrix of each population to the $p \times p$
-identity matrix. For each data set generated, we estimate the $\lambda$ and
-$\gamma$ parameters for both classifiers using a grid of 5 equidistant candidate
+identity matrix. For each data set generated, we estimated the parameters $\lambda$ and
+$\gamma$ for both classifiers using a grid of 5 equidistant candidate
 values between 0 and 1, inclusively. We set $\alpha_k = 1 - \gamma$, $k = 1,
-\ldots, K$, in the \emph{HDRDA} classifier to resemble
-\citeauthor{Friedman:1989tm}'s \emph{RDA} parameterization. At each pair of
-$\lambda$ and $\gamma$, we compute the 10-fold cross-validation error rate
-\citep{Hastie:2008dt}. Then, we select the optimal model that minimizes the 10-fold
+\ldots, K$, in the \emph{HDRDA} classifier. At each pair of
+$\lambda$ and $\gamma$, we computed the 10-fold cross-validation error rate
+\citep{Hastie:2008dt}. Then, we selected the optimal model that minimized the 10-fold
 cross-validation error rate.
 
-Here, our goal is to compare the computational runtime of the model selection
-for both classifiers as a function of the number of features $p$. We compare the
+Here, our goal was to compare the computational runtime of the model selection
+for both classifiers as a function of $p$. We compared the
 runtime of both classifiers by increasing the number of features from $p = 500$
-to $p = 5000$ in increments of 500. We generate 100 data sets for each value of
-$p$ and compute the training and model selection runtime of both
+to $p = 5000$ in increments of 500. We generated 100 data sets for each value of
+$p$ and computed the training and model selection runtime of both
 classifiers. Our timing comparisons are based on our \emph{HDRDA} implementation
@@
 
-In Figure \ref{fig:timing-results}, we plot the runtime of the model selections
-for both the \emph{HDRDA} and \emph{RDA} classifiers as a function of the number
-of features $p$. In the left panel of Figure \ref{fig:timing-results}, we
+In Figure \ref{fig:timing-results}, we plotted the runtime of the model selections
+for both the \emph{HDRDA} and \emph{RDA} classifiers as a function of $p$. In the left panel of Figure \ref{fig:timing-results}, we
 observe that the \emph{HDRDA} classifier is substantially faster than the
@@
 fit a regression line quadratic in $p$ to the \emph{RDA} runtimes and plotted
-the fit in Figure \ref{fig:timing-results}. We attempted to fit a cubic
-polynomial in $p$ to the \emph{RDA} timings, but the cubic term was
-insignificant with significant quadratic and linear terms. We also fit two
-polynomials to the \emph{RDA} timings with orders 2.807 and 2.376, which
-correspond to the asymptotic computational runtime of two well-known algorithms
-for computing a matrix inverse: the Strassen and Coppersmith-Winograd
-algorithms, respectively \footnote{TODO: Citations needed}. In both cases, the
-polynomials were statistically significiant at the 0.05 level. We displayed the
+the fit in Figure \ref{fig:timing-results}. We displayed the
 quadratic fit in Figure \ref{fig:timing-results} because it fit the timings best
@@
 
-In the left panel of Figure \ref{fig:timing-results}, we also fit a linear
-regression model that was linear in $p$ to the \emph{HDRDA} runtimes. In the
-right panel, we repeated the same scatterplot and linear fit with the timings
-restricted to the observed range of \emph{HDRDA} timings. The outliers in the
-\emph{HDRDA} timings were primarily due to the inherent computing variability in
-our {\tt c4.4xlarge} instance.
-
 Clearly, the model training and selection employed in \emph{HDRDA} is
 substantially faster than that of \emph{RDA}. To investigate the relative timing
-comparisons between the two classifiers, we consider the ratio of mean timings
+comparisons between the two classifiers, we considered the ratio of mean timings
 of the \emph{RDA} classifier to the \emph{HDRDA} classifier for each value of
@@
 the ratio of mean timings are given. First, we observe that the mean relative
-timings increases as $p$ increases. For smaller dimensions, the relative
-difference in computing is sizable with the average ratio of the mean timings
+timings increased as $p$ increased. For smaller dimensions, the relative
+difference in computing was sizable with the average ratio of the mean timings
 equal to \Sexpr{bootstrap_summary_p500$avg_ratio} for $p = 500$ and a 95\%
@@
 \Sexpr{bootstrap_summary_p500$upper_ci}). Furthermore, the ratio of mean
-computing times suggests that the \emph{RDA} classifier is impractical for
+computing times suggested that the \emph{RDA} classifier is impractical for
 higher dimensions. For instance, when $p = 5000$ the ratio of mean computing
-times increases to \Sexpr{bootstrap_summary_p5000$avg_ratio} with a 95\%
+times increased to \Sexpr{bootstrap_summary_p5000$avg_ratio} with a 95\%
 confidence interval of (\Sexpr{bootstrap_summary_p5000$lower_ci},
@@
 values $10^{-1}, \ldots, 10^4$, and $10^5$. We selected optimal estimates of
-$\lambda$ and $\gamma$ using $V$-fold cross-validation with $V = 10$. For the
+$\lambda$ and $\gamma$ using $10$-fold cross-validation. For the
 second model, we set $\alpha_k = 1 - \gamma$, $k = 1, \ldots, K$, to resemble
@@
 Although the design is indeed standard, the simulation configuration lacks
-artifacts observed commonly in real data, such as skewness and extreme
+artifacts commonly observed in real data, such as skewness and extreme
 outliers. As a result, we wished to investigate the effect of outliers on the
@@
 populations differed in their mean vectors in the first 100 features
-corresponding to the first block. No difference in the means occurred in the
+corresponding to the first block, and no difference in the means occurred in the
 remaining blocks.
@@
 
-In Figure \ref{fig:sim-results}, we observe each classifier's average
+In Figure \ref{fig:sim-results}, we observed each classifier's average
 classification error rates for the values of $\epsilon$ and $p$. Unsurprisingly,
@@
 In Figure \ref{fig:sim-results-by-p} we again examine the simulation results as
-a function of the number of features $p$ for a subset of the values of
+a function of $p$ for a subset of the values of
 $\epsilon$. This set of plots allows us to investigate the effect of feature
@@
 WDLS, and part of DDLS; (4) Spindle cell and pleomorophic sarcomas including
-DDLS, MFS, LMS, MPNST, FS, and MFH. Following \cite{Witten:2011kc}, we restrict
+DDLS, MFS, LMS, MPNST, FS, and MFH. Following \cite{Witten:2011kc}, we restricted
 our analysis to the five tumor types having at least 15 observations.
@@
 \emph{HDRDA} and Guo classifiers were superior in classification performance for
-the majority of the simulations. The \emph{HDRDA} classifiers had the best
+the majority of the simulations. The \emph{HDRDA} classifiers yielded the best
 classification accuracy on the Chowdary and Shipp data sets. Although the random
@@
 data set, our proposed classifiers outperformed the other competing classifiers
-considered here. Moreover, the \emph{HDRDA} classifiers had comparable
+considered here. Moreover, the \emph{HDRDA} classifiers yielded comparable
 performance on five of the six data sets.
@@
 comparisons we demonstrated that training and model selection could be employed
-on data sets having 5,000 features in 3 seconds on average. We contrast our
+on data sets having 5,000 features in three seconds on average. We contrast our
 implementation to the the poor computational performance of the in the {\tt
   klaR} R package, which took approximately 25 minutes on average to train the
-\emph{RDA} classifier on a data set having 5,000 features. Given the \emph{RDA}
+\emph{RDA} classifier on a data set having 5,000 features. Given that the \emph{RDA}
 classifier has been shown to have excellent performance in the high-dimensional
@@
 improve via such a method, the efficiency of the training and model selection
-would certainly improve. Moreover, if $q_L = 2$ or $3$ was selected, the
-potential for low-dimensional graphical displays of high-dimensional data could
-be accomplished.
+would certainly improve. Moreover, if $q_L = 2$ or $3$ for low-dimensional graphical displays of high-dimensional data can
+be obtained.
 

### 2017-01-21 reports/paper-hdrda.Rnw
Updated paper's title

@@
 
-\title{High-Dimensional Regularized Discriminant Analysis}
+\title{Computationally Efficient High-Dimensional Regularized Discriminant Analysis}
 
@@
 <<knitr_setup, include=FALSE, echo=FALSE, cache=FALSE>>=
+library(knitr)
 # Round digits throughout document including \Sexpr statements.

### 2017-01-22 reports/paper-hdrda.Rnw
Timing comparison results discussion much improved.

@@
       select(-HDRDA, -RDA)
+
+timings_summary <- timings %>%
+  group_by(Classifier, p) %>%
+  summarize(mean_runtime=mean(time), sd_runtime=sd(time))
+
+# RDA Mean Runtime in Minutes when p = 5000
+rda_mean_time_p5000 <- timings_summary %>% filter(Classifier == "RDA", p == 5000)
+rda_mean_time_p5000 <- rda_mean_time_p5000$mean_runtime / 60
+
+# HDRDA Mean Runtime in Seconds when p = 5000
+hdrda_mean_time_p5000 <- timings_summary %>% filter(Classifier == "HDRDA", p == 5000)
+hdrda_mean_time_p5000 <- hdrda_mean_time_p5000$mean_runtime
 @
@@
 In this section, we demonstrate that the computational performance of the
-training and model selection employed in the \emph{HDRDA} classifier is
+model selection employed in the \emph{HDRDA} classifier is
 substantially faster than that of the \emph{RDA} classifier on small-sample,
@@
 $\lambda$ and $\gamma$, we computed the 10-fold cross-validation error rate
-\citep{Hastie:2008dt}. Then, we selected the optimal model that minimized the 10-fold
+\citep{Hastie:2008dt}. Then, we selected the model that minimized the 10-fold
 cross-validation error rate.
 
-Here, our goal was to compare the computational runtime of the model selection
-for both classifiers as a function of $p$. We compared the
+We compared the
 runtime of both classifiers by increasing the number of features from $p = 500$
@@
 open-source statistical software {\tt R}. Our timing comparisons can be
-reproduced with the code available in the
-\url{https://github.com/ramhiser/paper-hdrda} repository.
+reproduced with the code available at
+\url{https://github.com/ramhiser/paper-hdrda}.
 
@@
 
-In Figure \ref{fig:timing-results}, we plotted the runtime of the model selections
-for both the \emph{HDRDA} and \emph{RDA} classifiers as a function of $p$. In the left panel of Figure \ref{fig:timing-results}, we
-observe that the \emph{HDRDA} classifier is substantially faster than the
-\emph{RDA} classifier as $p$ increases. Figure \ref{fig:timing-results} suggests
-that the usage of a matrix inverse and determinant in the {\tt klaR} R package's
-discriminant function yielded training and model-selection timings that exceeded
-linear growth in $p$. Because our implementation in the \emph{HDRDA} classifier
-removes the inverse and determinants from the classifier, the \emph{HDRDA}
-classifier is computationally more efficient than the \emph{RDA} classifier,
-especially as $p$ increases. We contrast our implementation to the the poor
-computational performance of the in the {\tt klaR} R package, which took
-approximately 25 minutes on average to train the \emph{RDA} classifier on a data
-set having 5,000 features. To quantify how much beyond linear growth in $p$, we
-fit a regression line quadratic in $p$ to the \emph{RDA} runtimes and plotted
-the fit in Figure \ref{fig:timing-results}. We displayed the
-quadratic fit in Figure \ref{fig:timing-results} because it fit the timings best
-for $p = 500$ to $p = 2000$. The fits for $2000 \le p \le 5000$ were similar.
-
-Clearly, the model training and selection employed in \emph{HDRDA} is
-substantially faster than that of \emph{RDA}. To investigate the relative timing
-comparisons between the two classifiers, we considered the ratio of mean timings
-of the \emph{RDA} classifier to the \emph{HDRDA} classifier for each value of
-$p$. We employed nonparametric bootstrapping\footnote{TODO: Citation needed} to
-estimate the mean ratio along with 95\% confidence intervals. In Figure
-\ref{fig:timing-comparison-bootstrap}, the bootstrap sampling distributions for
-the ratio of mean timings are given. First, we observe that the mean relative
-timings increased as $p$ increased. For smaller dimensions, the relative
-difference in computing was sizable with the average ratio of the mean timings
-equal to \Sexpr{bootstrap_summary_p500$avg_ratio} for $p = 500$ and a 95\%
-confidence interval of (\Sexpr{bootstrap_summary_p500$lower_ci},
+In Figure \ref{fig:timing-results}, we plotted the runtime of the model
+selections for both the \emph{HDRDA} and \emph{RDA} classifiers as a function
+of $p$. We observed that the \emph{HDRDA} classifier was substantially faster
+than the \emph{RDA} classifier as $p$ increased. In the left panel of Figure
+\ref{fig:timing-results}, we fit a quadratic regression line to the \emph{RDA}
+runtimes and a simple linear regression model to the \emph{HDRDA} runtimes.
+For improved understanding, in the right panel we repeated the same scatterplot
+and linear fit with the timings restricted to the observed range of the
+\emph{HDRDA} timings. Figure \ref{fig:timing-results} suggests that the usage
+of a matrix inverse and determinant in the {\tt klaR} R package's discriminant
+function yielded model-selection timings that exceeded linear growth in
+$p$. Because the \emph{HDRDA} classifier removes inverse and determinants, it
+was computationally more efficient than the \emph{RDA} classifier, especially
+as $p$ increased. In fact, when $p = 5000$ the \emph{RDA} classifier required
+\Sexpr{rda_mean_time_p5000} minutes on average to perform model selection,
+while the \emph{HDRDA} classifier selected its optimal model in
+\Sexpr{hdrda_mean_time_p5000} seconds on average. Clearly, the model selection
+employed by the \emph{HDRDA} classifier is substantially faster than that of
+the \emph{RDA} classifier.
+
+We quantified the relative timing comparisons between the two classifiers by
+calculating the ratio of mean timings of the \emph{RDA} classifier to the
+\emph{HDRDA} classifier for each value of $p$. We employed nonparametric
+bootstrapping to estimate the mean ratio along with 95\% confidence
+intervals. In Figure \ref{fig:timing-comparison-bootstrap}, the bootstrap
+sampling distributions for the ratio of mean timings are given. First, we
+observe that the mean relative timings increased as $p$ increased. For smaller
+dimensions, the relative difference in computing was sizable with the average
+ratio of the mean timings equal to \Sexpr{bootstrap_summary_p500$avg_ratio} for
+$p = 500$ and a 95\% confidence interval of
+(\Sexpr{bootstrap_summary_p500$lower_ci},
 \Sexpr{bootstrap_summary_p500$upper_ci}). Furthermore, the ratio of mean
@@
 $\bm \mu_2 = (\underbrace{1/2, \ldots, 1/2}_{100}, \underbrace{0, \ldots, 0}_{p
-  - 100})$. For simplicity, we defined $\bm \mu_3 = -\bm \mu_2$. The three
+  - 100})$. For simplicity, we defined $\bm \mu_3 = -\bm \mu_2$. The 3
 populations differed in their mean vectors in the first 100 features
@@
 comparisons we demonstrated that training and model selection could be employed
-on data sets having 5,000 features in three seconds on average. We contrast our
+on data sets having 5,000 features in 3 seconds on average. We contrast our
 implementation to the the poor computational performance of the in the {\tt
@@
 
-The dimension reduction that we have employed in this paper has reduced the
+The dimension reduction employed in this paper has reduced the
 dimension to rank$(\widehat{\bm \Sigma}) = q$. An interesting extension of our
-work would reduce the dimension $q$ further to a lower dimension $q_L < q$ using
-some criterion proposed, perhaps similar to that employed in principal
+work would reduce the dimension $q$ further to a lower dimension $q_L < q$, perhaps using a criterion similar to that of principal
 components analysis. While unclear whether the classification performance would
 improve via such a method, the efficiency of the training and model selection
-would certainly improve. Moreover, if $q_L = 2$ or $3$ for low-dimensional graphical displays of high-dimensional data can
+would certainly improve. Moreover, if $q_L = 2$ or $3$, low-dimensional graphical displays of high-dimensional data could
 be obtained.

### 2017-01-22 reports/paper-hdrda.Rnw
Revamped the abstract with Phil. Much better.

@@
 
-We develop a new and computationally efficient regularized discriminant analysis (RDA)
-classifier for the small-sample, high-dimensional setting.  We are able
-to interpret the biased covariance matrix estimator employed in the HDRDA
-classifier as one that is estimated from all training observations regardless of
-class. Moreover, our classifier also differs from
-\citeauthor{Friedman:1989tm}'s in that we show our classifier is equivalent to
-a classifier in a reduced feature space with dimension approximately equal to the
-training sample size. As a result, the matrix operations used in the HDRDA
-classifier are computationally linear in the number of features, making HDRDA
-well-suited for high-dimensional classification in practice. We demonstrate that HDRDA is often superior to several
-sparse and regularized classifiers in terms of classification accuracy with
-three artificial and six real high-dimensional data sets.  Also, timing comparisons
-between our HDRDA implementation in the {\tt sparsediscrim} R package and the
-standard RDA formulation in the {\tt klaR} R package demonstrate that as the
-number of features increases, the computational runtime of HDRDA is drastically
-smaller than RDA.
+Regularized discriminant analysis (\emph{RDA}) proposed by
+\cite{Friedman:1989tm} is a widely popular classifier that lacks
+interpretability and is impractical for high-dimensional data sets.  Here, we
+present an interpretable and computationally efficient classifier called
+high-dimensional \emph{RDA} (\emph{HDRDA}), designed for the small-sample,
+high-dimensional setting. For \emph{HDRDA}, we show that each training
+observation regardless of class contributes to the class covariance matrix,
+resulting in interpretable estimator that \emph{borrows} from the pooled sample
+covariance matrix. Moreover, we show that \emph{HDRDA} is equivalent to a
+classifier in a reduced feature space with dimension approximately equal to the
+training sample size. As a result, the matrix operations employed by
+\emph{HDRDA} are computationally linear in the number of features, making the
+classifier well-suited for high-dimensional classification in practice. We
+demonstrate that \emph{HDRDA} is often superior to several sparse and
+regularized classifiers in terms of classification accuracy with three
+artificial and six real high-dimensional data sets. Also, timing comparisons
+between our \emph{HDRDA} implementation in the {\tt sparsediscrim} R package
+and the standard \emph{RDA} formulation in the {\tt klaR} R package demonstrate
+that as the number of features increases, the computational runtime of
+\emph{HDRDA} is drastically smaller than \emph{RDA}.
 
@@
 $\alpha_k$ because the estimation of $K$ additional tuning parameters via
-cross-validation is counterproductive to our goal of improving the \emph{RDA}
-classifier's computational efficiency. By setting $\alpha_k$ equal to an
-appropriate constant, the \emph{HDRDA} covariance-matrix estimator includes or
+cross-validation is counterproductive to our goal of computational efficiency.
+By setting $\alpha_k$ equal to an appropriate constant, the \emph{HDRDA} covariance-matrix estimator includes or
 resembles a large family of estimators. Notice that if $\alpha_k = 1$ and
 $\lambda = 1$, \eqref{eq:hdrda-cov} is equivalent to the standard ridge-like
-covariance-matrix estimator in \eqref{eq:ridge-estimator}, and various
+covariance-matrix estimator in \eqref{eq:ridge-estimator}. Other
 estimators proposed in the literature can be obtained when one selects $\gamma$
@@
 
-\cite{Friedman:1989tm}
-proposed the biased covariance-matrix estimator
+\cite{Friedman:1989tm} proposed the biased covariance-matrix estimator
 \begin{linenomath*}

### 2017-01-23 reports/paper-hdrda.Rnw
Cleaned up original RDA's notation.

@@
 
-\cite{Friedman:1989tm} proposed the biased covariance-matrix estimator
+When $\alpha_k = 1 - \gamma$, \eqref{eq:hdrda-cov} resembles the biased covariance-matrix estimator
 \begin{linenomath*}
 \begin{align}
-    \widehat{\bm \Sigma}_k(\lambda, \gamma) = (1 - \gamma) \widehat{\bm \Sigma}_k(\lambda) + \gamma \frac{\tr\{\widehat{\bm \Sigma}_k(\lambda)\}}{p} \bm I_p,\label{eq:sig-rda}
+    \widehat{\bm \Sigma}_k(\lambda, \gamma) = (1 - \gamma) \widehat{\bm \Sigma}_k^{(RDA)}(\lambda) + \gamma \frac{\tr\left\{\widehat{\bm \Sigma}_k^{(RDA)}(\lambda)\right\}}{p} \bm I_p \label{eq:sig-rda}
 \end{align}
 \end{linenomath*}
-where $\gamma \in [0, 1]$ is a regularization parameter that controls the
-shrinkage of \eqref{eq:sig-rda} towards $\bm I_p$ and is weighted by the average
-of the eigenvalues of \eqref{eq:sig-lambda}. \cite{Friedman:1989tm}
-established the RDA classifier by substituting \eqref{eq:sig-rda} into
-\eqref{eq:qda}.
-
-Our proposed \emph{HDRDA} classifier yields an
-interpretation of the pooling parameter $\lambda$ as discussed above. Moreover,
-as we discuss below, our formulation of the \emph{HDRDA} classifier lends itself
-to rapid estimation in high-dimensional classification.
+employed in the \emph{RDA} classifier, where $\widehat{\bm
+  \Sigma}_k^{(RDA)}(\lambda)$ is a pooled estimator of $\bm \Sigma_k$ and
+$\gamma \in [0, 1]$ is a regularization parameter that controls the shrinkage of
+\eqref{eq:sig-rda} towards $\bm I_p$ weighted by the average of the eigenvalues
+of $\widehat{\bm \Sigma}_k^{(RDA)}(\lambda)$. Despite the similarity of
+\eqref{eq:sig-rda} to the \emph{HDRDA} covariance-matrix estimator in
+\eqref{eq:hdrda-cov}, the \emph{RDA} classifier is impractical for
+high-dimensional data because the inverse and determinant of \eqref{eq:sig-rda}
+must be calculated when substituted into \eqref{eq:qda}. Furthermore,
+\eqref{eq:sig-rda} has no clear interpretation.
 
@@
 
-Also, notice that if $\alpha_k$ is constant
-across the $K$ classes, then $\bm \Gamma_k$ in \eqref{eq:Gamma_k} is independent
-of $k$. Consequently, $|\bm \Gamma_k|$ is constant across the $K$ classes and
-need not be calculated in \eqref{eq:hdrda-decomposed}.
+Notice that $\bm \Gamma_k$ is singular when $(\lambda, \gamma) = (0, 0)$
+because $\bm \Gamma_k = \bm 0_q$, in which case we use the formulation in
+\eqref{eq:hdrda-decomposed} instead. Also, notice that if $\alpha_k$ is
+constant across the $K$ classes, then $\bm \Gamma_k$ in \eqref{eq:Gamma_k} is
+independent of $k$. Consequently, $|\bm \Gamma_k|$ is constant across the $K$
+classes and need not be calculated in \eqref{eq:hdrda-decomposed}.
 

### 2017-01-23 reports/paper-hdrda.Rnw
Tightened up the discussion.

@@
 
-We have compared our proposed \emph{HDRDA} classifier to the random-forest and
-Witten classifiers as well as the Pang, Tong, and Guo classifiers, each of which
-assumes independence of the features conditional on class membership. Despite
-the rapid computational performance of diagonal classifiers and the reduction in
-the number of parameters to estimate, \cite{Mai:2012bf} and \cite{Fan:2012iq}
-have noted that diagonal classifiers can often yield inferior classification
-performance when compared to other classification methods. In fact, we have
-shown that the \emph{HDRDA} classifier often yields superior classification
-accuracy when compared to the diagonal classifiers considered using six
-small-sample, high-dimensional data sets, confirming the assertions of
-\cite{Mai:2012bf} and \cite{Fan:2012iq}. Furthermore, we have demonstrated that
-\emph{HDRDA} classifiers are more robust to the presence of outliers than the
-diagonal classifiers.
+We have demonstrated that our proposed \emph{HDRDA} classifier is competitive
+with and often superior to random forests as well as the Witten, Pang, Tong, and
+Guo classifiers. In fact, we have shown that the \emph{HDRDA} classifier often
+yields superior classification accuracy when applied to small-sample,
+high-dimensional data sets, confirming the assertions of \cite{Mai:2012bf} and
+\cite{Fan:2012iq} that diagonal classifiers often yield inferior classification
+performance when compared to other classification methods. Furthermore, we have
+demonstrated that \emph{HDRDA} classifiers are more robust to the presence of
+outliers than the diagonal classifiers despite their rapid computational
+performance and their reduction in the number of parameters to estimate.
 
@@
 competitive within our classification study and perhaps superior. Contrarily,
-its underperformance in our empirical studies show that it is sensitive to
-outliers and unable to achieve comparable results with other classifiers
-designed for small-sample, high-dimensional data. Also, despite the claims of
-\cite{FernandezDelgado:2014ul} that the random forest classifier is typically
-superior to other classifiers, we observed that the random-forest classifier was
-indeed competitive but was typically outperformed by classifiers developed for
-small-sample, high-dimensional data.
-
-We demonstrated that the \emph{HDRDA} implementation of
-\eqref{eq:hdrda-decomposed} in the {\tt sparsediscrim} R package can be used in
-practice with high-dimensional data sets. For instance, in our timing
-comparisons we demonstrated that training and model selection could be employed
-on data sets having 5,000 features in 3 seconds on average. We contrast our
-implementation to the the poor computational performance of the in the {\tt
-  klaR} R package, which took approximately 25 minutes on average to train the
-\emph{RDA} classifier on a data set having 5,000 features. Given that the \emph{RDA}
-classifier has been shown to have excellent performance in the high-dimensional
-setting \citep{Webb:2011vu} but is limited by its computationally intense
-model-selection procedure, our work enables the \emph{RDA} classifier to be
-employed with small-sample, high-dimensional data in practice. This result is
-reassuring because the \emph{RDA} classifier remains widely popular in the
-literature. In fact, variants of the RDA classifier have been applied to
-microarray data \citep{Ching:2012fu,Li:2012ev,Tai:2007bk,Guo:2007te}, facial
-recognition
-\citep{Ching:2012fu,Zhang:2010va,Dai:2007vd,Lu:2005hq,Pima:2004vw,Lu:2003we},
+our empirical studies suggests that the classifier is sensitive to outliers and
+unable to achieve comparable results with other classifiers designed for
+small-sample, high-dimensional data. Also, despite the claims of
+\cite{FernandezDelgado:2014ul} that random forests are typically superior to
+other classifiers, we observed that they were indeed competitive but were
+typically outperformed by classifiers developed for small-sample,
+high-dimensional data.
+
+We demonstrated that our \emph{HDRDA} implementation in the {\tt sparsediscrim}
+R package can be used in practice with high-dimensional data sets. In our timing
+comparisons we showed that \emph{HDRDA} model selection could be employed on
+data sets with $p = 5000$ in \Sexpr{hdrda_mean_time_p5000} seconds on
+average. Contrarily, the \emph{RDA} classifier implemented in the {\tt klaR} R
+package required \Sexpr{rda_mean_time_p5000} minutes on average to perform model
+selection on data sets with $p = 5000$. Given that the \emph{RDA} classifier has
+been shown to have excellent performance in the high-dimensional setting
+\citep{Webb:2011vu} but is limited by its computationally intense
+model-selection procedure, our work replaces the \emph{RDA} classifier for
+high-dimensional data in practice. This result is reassuring because the
+\emph{RDA} classifier remains widely popular in the literature. In fact,
+variants of the RDA classifier have been applied to microarray data
+\citep{Ching:2012fu,Li:2012ev,Tai:2007bk,Guo:2007te}, facial recognition
+\citep{Zhang:2010va,Dai:2007vd,Lu:2005hq,Pima:2004vw,Lu:2003we},
 handwritten digit recognition \citep{Bouveyron:2007gx}, remote sensing
 \citep{Tadjudin:1999fk}, seismic detection \citep{Anderson:2002kg}, and chemical
-spectra \citep*{Wu:1996us,Aeberhard:1993fp}.
-
-The dimension reduction employed in this paper has reduced the
-dimension to rank$(\widehat{\bm \Sigma}) = q$. An interesting extension of our
-work would reduce the dimension $q$ further to a lower dimension $q_L < q$, perhaps using a criterion similar to that of principal
-components analysis. While unclear whether the classification performance would
-improve via such a method, the efficiency of the training and model selection
-would certainly improve. Moreover, if $q_L = 2$ or $3$, low-dimensional graphical displays of high-dimensional data could
-be obtained.
+spectra \citep{Wu:1996us,Aeberhard:1993fp}.
+
+The dimension reduction employed in this paper has reduced the dimension to
+rank$(\widehat{\bm \Sigma}) = q$. An interesting extension of our work would
+reduce the dimension $q$ further to a lower dimension $q_L < q$, perhaps using a
+criterion similar to that of principal components analysis. While unclear
+whether the classification performance would improve via such a method, the
+efficiency of the model selection would certainly improve. Moreover, if $q_L =
+2$ or $3$, low-dimensional graphical displays of high-dimensional data could be
+obtained.
 

### 2017-02-05 reports/paper-hdrda.Rnw
Applied Mrs. Young's edits.

@@
 
-Regularized discriminant analysis (\emph{RDA}) proposed by
-\cite{Friedman:1989tm} is a widely popular classifier that lacks
+Regularized discriminant analysis (\emph{RDA}), proposed by
+\cite{Friedman:1989tm}, is a widely popular classifier that lacks
 interpretability and is impractical for high-dimensional data sets.  Here, we
@@
 high-dimensional setting. For \emph{HDRDA}, we show that each training
-observation regardless of class contributes to the class covariance matrix,
-resulting in interpretable estimator that \emph{borrows} from the pooled sample
-covariance matrix. Moreover, we show that \emph{HDRDA} is equivalent to a
-classifier in a reduced feature space with dimension approximately equal to the
-training sample size. As a result, the matrix operations employed by
+observation, regardless of class, contributes to the class covariance matrix,
+resulting in an interpretable estimator that \emph{borrows} from the pooled
+sample covariance matrix. Moreover, we show that \emph{HDRDA} is equivalent to
+a classifier in a reduced-feature space with dimension approximately equal to
+the training sample size. As a result, the matrix operations employed by
 \emph{HDRDA} are computationally linear in the number of features, making the
@@
 that as the number of features increases, the computational runtime of
-\emph{HDRDA} is drastically smaller than \emph{RDA}.
+\emph{HDRDA} is drastically smaller than that of \emph{RDA}.
 
@@
 data, where the number of features $p$ exceeds the training sample size $N$. In
-this setting, well-established classifiers such as linear discriminant analysis
-(\emph{LDA}) and quadratic discriminant analysis (\emph{QDA}) become
+this setting, well-established classifiers, such as linear discriminant analysis
+(\emph{LDA}) and quadratic discriminant analysis (\emph{QDA}), become
 incalculable because the class and pooled covariance matrix estimators are
 singular \citep*{Murphy:2012uq, Bouveyron:2007gx, Mkhadri:1997gy}. To improve
-the estimation of the class covariance matrices estimated in the \emph{QDA}
-classifier and to ensure the covariance matrix estimators are nonsingular,
-\cite{Friedman:1989tm} proposed the regularized discriminant analysis
-(\emph{RDA}) classifier by incorporating a weighted average of the pooled sample
-covariance matrix and the class sample covariance matrix. To further improve the
-estimation of the class covariance matrix and to stabilize its inverse,
-\cite{Friedman:1989tm} also included a regularization component by shrinking the
-covariance matrix estimator towards the identity matrix, which yields a
-nonsingular estimator following the well-known ridge-regression approach of
-\cite{Hoerl:1970cd}. Despite its popularity, the ``borrowing'' operation
-employed in the \emph{RDA} classifier lacks interpretability
-\citep{Bensmail:1996fw}. Furthermore, the \emph{RDA} classifier is impractical
-for high-dimensional data sets because it computes the inverse and determinant
-of the covariance matrices for each class. Both matrix calculations are
-computationally expensive because the number of operations grows at a polynomial
-rate in the number of features. Moreover, the \emph{RDA} classifier model
-selection of its two tuning parameters is computationally burdensome because the
-matrix inverse and determinant of the covariance matrices across multiple
-cross-validation folds for each class and for each candidate tuning-parameter
-model.
+the accuracy of the estimation of the class covariance matrices estimated in the
+\emph{QDA} classifier and to ensure that the covariance matrix estimators are
+nonsingular, \cite{Friedman:1989tm} proposed the regularized discriminant
+analysis (\emph{RDA}) classifier by incorporating a weighted average of the
+pooled sample covariance matrix and the class sample covariance matrix. To
+further improve the accuracy of the estimation of the class covariance matrix
+and to stabilize its inverse, \cite{Friedman:1989tm} also included a
+regularization component by shrinking the covariance matrix estimator towards
+the identity matrix, which yields a nonsingular estimator following the
+well-known ridge-regression approach of \cite{Hoerl:1970cd}. Despite its
+popularity, the ``borrowing'' operation employed in the \emph{RDA} classifier
+lacks interpretability \citep{Bensmail:1996fw}. Furthermore, the \emph{RDA}
+classifier is impractical for high-dimensional data sets because it computes the
+inverse and determinant of the covariance matrices for each class. Both matrix
+calculations are computationally expensive because the number of operations
+grows at a polynomial rate in the number of features. Moreover, the model
+selection of the \emph{RDA} classifier's two tuning parameters is
+computationally burdensome because the matrix inverse and determinant of each
+class covariance matrix are computed across multiple cross-validation folds for
+each candidate tuning-parameter pair.
 
@@
 lacking with the \emph{RDA} classifier \citep{Bensmail:1996fw}. Our
-parameterization differs from \cite*{Hastie:2008dt} and \cite{Halbe:2007ea} in
-that our formulation allows the flexibility of various covariance-matrix
-estimators proposed in the literature, including a variety of ridge-like
-estimators such as the one proposed by \cite{Srivastava:2007ww}.
+parameterization differs from that of \cite*{Hastie:2008dt} and that of
+\cite{Halbe:2007ea} in that our formulation allows the flexibility of various
+covariance-matrix estimators proposed in the literature, including a variety of
+ridge-like estimators, such as the one proposed by \cite{Srivastava:2007ww}.
 
@@
 the pooled sample covariance matrix are redundant and can be discarded from the
-\emph{HDRDA} decision rule without loss of classificatory information by
-applying reasoning similar to that of \cite{Ye:2006tq}. As a result, we achieve
-a substantial reduction in dimension such that the matrix operations used in the
+\emph{HDRDA} decision rule without loss of classificatory information when we
+apply reasoning similar to that of \cite{Ye:2006tq}. As a result, we achieve a
+substantial reduction in dimension such that the matrix operations used in the
 \emph{HDRDA} classifier are computationally linear in the number of features.
@@
 features increases, the computational runtime of the \emph{HDRDA} classifier is
-drastically smaller than \emph{RDA}. In fact, when $p = 5000$, we show that the
-\emph{HDRDA} classifier is \Sexpr{bootstrap_summary_p5000$avg_ratio} times
-faster on average than the \emph{RDA} classifier. In this scenario, the
-\emph{HDRDA} classifier's training and model selection requires 3 seconds on
-average, while that of the \emph{RDA} classifier requires 25 minutes on average.
+drastically smaller than that of \emph{RDA}. In fact, when $p = 5000$, we show
+that the \emph{HDRDA} classifier is \Sexpr{bootstrap_summary_p5000$avg_ratio}
+times faster on average than the \emph{RDA} classifier. In this scenario, the
+\emph{HDRDA} classifier's model selection requires \Sexpr{hdrda_mean_time_p5000}
+seconds on average, while that of the \emph{RDA} classifier requires
+\Sexpr{rda_mean_time_p5000} minutes on average.
 
@@
 The remainder of this paper is organized as follows. In Section
-\ref{sec:preliminaries} we present the classification problem and necessary
+\ref{sec:preliminaries} we introduce the classification problem and necessary
 notation to describe our contributions. In Section \ref{sec:rda} we present the
-\emph{HDRDA} classifier along with its interpretation, properties, and
-computationally efficient model-selection procedure. In Section \ref{sec:sims}
-we describe our simulation studies of artificial and real data sets and describe
-the experimental results. We conclude with a brief discussion in Section
+\emph{HDRDA} classifier along with its interpretation. In Section
+\ref{sec:hdrda-properties}, we provide properties of the \emph{HDRDA} classifier
+and a computationally efficient model-selection procedure. In Section
+\ref{sec:timing-comparisons}, we compare the model-selection timings of the
+\emph{HDRDA} and \emph{RDA} classifiers. In Section \ref{sec:sims} we describe
+our simulation studies of artificial and real data sets and examine the
+experimental results. We conclude with a brief discussion in Section
 \ref{sec:discussion}.
@@
 
-To facilitate our discussion of covariance-matrix regularization and regularized
-discriminant analysis, we require the following notation. Let $\mathbb{R}_{a
-  \times b}$ denote the matrix space of all $a \times b$ matrices over the real
-field $\mathbb{R}$. Denote by $\bm I_m$ the $m \times m$ identity matrix, and
-let $\bm 0_{m \times p}$ be the $m \times p$ matrix of zeros, such that $\bm
-0_m$ is understood to denote $\bm 0_{m \times m}$. Define $\bm 1_m \in
+To facilitate our discussion of covariance-matrix regularization and
+high-dimensional classification, we require the following notation. Let
+$\mathbb{R}_{a \times b}$ denote the matrix space of all $a \times b$ matrices
+over the real field $\mathbb{R}$. Denote by $\bm I_m$ the $m \times m$ identity
+matrix, and let $\bm 0_{m \times p}$ be the $m \times p$ matrix of zeros, such
+that $\bm 0_m$ is understood to denote $\bm 0_{m \times m}$. Define $\bm 1_m \in
 \mathbb{R}_{m \times 1}$ as a vector of ones. Let $\bm A^{T}$, $\bm A^+$, and
@@
 a $0-1$ loss function when $p(\bm x | \omega_k)$ is the PDF of the multivariate
-normal distribution with known mean vectors $\bm \mu_k \in \mathbb{R}_{p \times 1}$
-and known covariance matrices $\bm \Sigma_k \in \mathbb{R}_{p \times p}^{>}$,
-$k = 1, 2, \ldots, K$. Because $\bm \mu_k$ and $\bm \Sigma_k$ are typically
-unknown, we assign an unlabeled observation $\bm x$ to class $\omega_k$ with the
-sample \emph{QDA} classifier
+normal distribution with known mean vectors $\bm \mu_k \in \mathbb{R}_{p \times
+  1}$ and known covariance matrices $\bm \Sigma_k \in \mathbb{R}_{p \times
+  p}^{>}$, $k = 1, 2, \ldots, K$. Because $\bm \mu_k$ and $\bm \Sigma_k$ are
+typically unknown, we assign an unlabeled observation $\bm x$ to class
+$\omega_k$ with the sample \emph{QDA} classifier
 \begin{linenomath*}
@@
 \end{linenomath*}
-where $\xbar_k$ and $\widehat{\bm \Sigma}_k$ are the maximum likelihood
+where $\xbar_k$ and $\widehat{\bm \Sigma}_k$ are the maximum-likelihood
 estimators (MLEs) of $\bm \mu_k$ and $\bm \Sigma_k$, respectively. If we assume
@@
 Substituting \eqref{eq:hdrda-cov} into \eqref{eq:qda}, we define the
-\emph{HDRDA} classifier
+\emph{HDRDA} classifier as
 \begin{linenomath*}
@@
 cross-validation is counterproductive to our goal of computational efficiency.
-By setting $\alpha_k$ equal to an appropriate constant, the \emph{HDRDA} covariance-matrix estimator includes or
-resembles a large family of estimators. Notice that if $\alpha_k = 1$ and
-$\lambda = 1$, \eqref{eq:hdrda-cov} is equivalent to the standard ridge-like
-covariance-matrix estimator in \eqref{eq:ridge-estimator}. Other
-estimators proposed in the literature can be obtained when one selects $\gamma$
-accordingly. For instance, with $\gamma = \tr\{ \widehat{\bm \Sigma} \} /
-\min(N, p)$, we obtain the estimator from \cite{Srivastava:2007ww}.
+For appropriate values of $\alpha_k$, the \emph{HDRDA} covariance-matrix
+estimator includes or resembles a large family of estimators. Notice that if
+$\alpha_k = 1$ and $\lambda = 1$, \eqref{eq:hdrda-cov} is equivalent to the
+standard ridge-like covariance-matrix estimator in \eqref{eq:ridge-estimator}.
+Other estimators proposed in the literature can be obtained when one selects
+$\gamma$ accordingly. For instance, with $\gamma = \tr\{ \widehat{\bm \Sigma} \}
+/ \min(N, p)$, we obtain the estimator from \cite{Srivastava:2007ww}.
 
@@
 
-Using Theorem 1, we can avoid the time-consuming inverses and determinants of $p \times p$
-covariance matrices in \eqref{eq:hdrda} and instead calculate these same
-operations on $\bm W_k \in \mathbb{R}_{q \times q}$ in
+Using Theorem 1, we can avoid the time-consuming inverses and determinants of $p
+\times p$ covariance matrices in \eqref{eq:hdrda} and instead calculate these
+same operations on $\bm W_k \in \mathbb{R}_{q \times q}$ in
 \eqref{eq:hdrda-decomposed}. The substantial computational improvements arise
@@
 follows from the well-known Sherman-Woodbury formula \citep[Theorem
-  18.2.8]{Harville:2008wja} by noting that $(\bm A + \bm B \bm T \bm C)^{-1} =
+  18.2.8]{Harville:2008wja} because $(\bm A + \bm B \bm T \bm C)^{-1} =
 \bm A^{-1} - \bm A^{-1} \bm B (\bm T^{-1} + \bm C \bm A^{-1} \bm B)^{-1} \bm C
@@
 
-In this section, we demonstrate that the computational performance of the
-model selection employed in the \emph{HDRDA} classifier is
-substantially faster than that of the \emph{RDA} classifier on small-sample,
-high-dimensional data sets. The relative difference in runtime between the two
-classifiers drastically increases as $p$ increases. To
-compare the two classifiers, we generated 25 observations from each of $K=4$
-multivariate normal populations with mean vectors $\bm \mu_1 = -3 \cdot \bm
-1_p$, $\bm \mu_2 = -\bm 1_p$, $\bm \mu_3 = \bm 1_p$, and $\bm \mu_4 = 3 \cdot
-\bm 1_p$. We set the covariance matrix of each population to the $p \times p$
-identity matrix. For each data set generated, we estimated the parameters $\lambda$ and
-$\gamma$ for both classifiers using a grid of 5 equidistant candidate
-values between 0 and 1, inclusively. We set $\alpha_k = 1 - \gamma$, $k = 1,
-\ldots, K$, in the \emph{HDRDA} classifier. At each pair of
-$\lambda$ and $\gamma$, we computed the 10-fold cross-validation error rate
-\citep{Hastie:2008dt}. Then, we selected the model that minimized the 10-fold
-cross-validation error rate.
-
-We compared the
-runtime of both classifiers by increasing the number of features from $p = 500$
-to $p = 5000$ in increments of 500. We generated 100 data sets for each value of
-$p$ and computed the training and model selection runtime of both
-classifiers. Our timing comparisons are based on our \emph{HDRDA} implementation
-in the {\tt sparsediscrim} R package and the standard \emph{RDA} implementation
-in the {\tt klaR} R package. All timing comparisons were conducted on an Amazon
-Elastic Compute Cloud (EC2) {\tt c4.4xlarge} instance using version 3.3.1 of the
-open-source statistical software {\tt R}. Our timing comparisons can be
-reproduced with the code available at
+In this section, we demonstrate that the computational performance of the model
+selection employed in the \emph{HDRDA} classifier is substantially faster than
+that of the \emph{RDA} classifier on small-sample, high-dimensional data
+sets. The relative difference in runtime between the two classifiers drastically
+increases as $p$ increases. To compare the two classifiers, we generated 25
+observations from each of $K=4$ multivariate normal populations with mean
+vectors $\bm \mu_1 = -3 \cdot \bm 1_p$, $\bm \mu_2 = -\bm 1_p$, $\bm \mu_3 = \bm
+1_p$, and $\bm \mu_4 = 3 \cdot \bm 1_p$. We set the covariance matrix of each
+population to the $p \times p$ identity matrix. For each data set generated, we
+estimated the parameters $\lambda$ and $\gamma$ for both classifiers using a
+grid of 5 equidistant candidate values between 0 and 1, inclusively. We set
+$\alpha_k = 1 - \gamma$, $k = 1, \ldots, K$, in the \emph{HDRDA} classifier. At
+each pair of $\lambda$ and $\gamma$, we computed the 10-fold cross-validation
+error rate \citep{Hastie:2008dt}. Then, we selected the model that minimized the
+10-fold cross-validation error rate.
+
+We compared the runtime of both classifiers by increasing the number of features
+from $p = 500$ to $p = 5000$ in increments of 500. Next, we generated 100 data
+sets for each value of $p$ and computed the training and model selection runtime
+of both classifiers. Our timing comparisons are based on our \emph{HDRDA}
+implementation in the {\tt sparsediscrim} R package and the standard \emph{RDA}
+implementation in the {\tt klaR} R package. All timing comparisons were
+conducted on an Amazon Elastic Compute Cloud (EC2) {\tt c4.4xlarge} instance
+using version 3.3.1 of the open-source statistical software {\tt R}. Our timing
+comparisons can be reproduced with the code available at
 \url{https://github.com/ramhiser/paper-hdrda}.
@@
 In Figure \ref{fig:timing-results}, we plotted the runtime of the model
-selections for both the \emph{HDRDA} and \emph{RDA} classifiers as a function
-of $p$. We observed that the \emph{HDRDA} classifier was substantially faster
-than the \emph{RDA} classifier as $p$ increased. In the left panel of Figure
+selections for both the \emph{HDRDA} and \emph{RDA} classifiers as a function of
+$p$. We observed that the \emph{HDRDA} classifier was substantially faster than
+the \emph{RDA} classifier as $p$ increased. In the left panel of Figure
 \ref{fig:timing-results}, we fit a quadratic regression line to the \emph{RDA}
-runtimes and a simple linear regression model to the \emph{HDRDA} runtimes.
-For improved understanding, in the right panel we repeated the same scatterplot
-and linear fit with the timings restricted to the observed range of the
-\emph{HDRDA} timings. Figure \ref{fig:timing-results} suggests that the usage
-of a matrix inverse and determinant in the {\tt klaR} R package's discriminant
-function yielded model-selection timings that exceeded linear growth in
-$p$. Because the \emph{HDRDA} classifier removes inverse and determinants, it
-was computationally more efficient than the \emph{RDA} classifier, especially
-as $p$ increased. In fact, when $p = 5000$ the \emph{RDA} classifier required
-\Sexpr{rda_mean_time_p5000} minutes on average to perform model selection,
-while the \emph{HDRDA} classifier selected its optimal model in
+runtimes and a simple linear regression model to the \emph{HDRDA} runtimes.  For
+improved understanding, in the right panel we repeated the same scatterplot and
+linear fit with the timings restricted to the observed range of the \emph{HDRDA}
+timings. Figure \ref{fig:timing-results} suggests that the usage of a matrix
+inverse and determinant in the {\tt klaR} R package's discriminant function
+yielded model-selection timings that exceeded linear growth in $p$. Because the
+\emph{HDRDA} classifier removes inverse and determinants, it was computationally
+more efficient than the \emph{RDA} classifier, especially as $p$ increased. In
+fact, when $p = 5000$, the \emph{RDA} classifier required
+\Sexpr{rda_mean_time_p5000} minutes on average to perform model selection, while
+the \emph{HDRDA} classifier selected its optimal model in
 \Sexpr{hdrda_mean_time_p5000} seconds on average. Clearly, the model selection
-employed by the \emph{HDRDA} classifier is substantially faster than that of
-the \emph{RDA} classifier.
+employed by the \emph{HDRDA} classifier is substantially faster than that of the
+\emph{RDA} classifier.
 
@@
 computing times suggested that the \emph{RDA} classifier is impractical for
-higher dimensions. For instance, when $p = 5000$ the ratio of mean computing
+higher dimensions. For instance, when $p = 5000$, the ratio of mean computing
 times increased to \Sexpr{bootstrap_summary_p5000$avg_ratio} with a 95\%
@@
 surname. All simulations were conducted on an Amazon EC2 {\tt c4.4xlarge}
-instance. Our analyses can be reproduced using the code available in the
-\url{https://github.com/ramhiser/paper-hdrda} repository.
+instance. Our analyses can be reproduced via the code available at
+\url{https://github.com/ramhiser/paper-hdrda}.
 
@@
 values $10^{-1}, \ldots, 10^4$, and $10^5$. We selected optimal estimates of
-$\lambda$ and $\gamma$ using $10$-fold cross-validation. For the
-second model, we set $\alpha_k = 1 - \gamma$, $k = 1, \ldots, K$, to resemble
+$\lambda$ and $\gamma$ using $10$-fold cross-validation. For the second model,
+we set $\alpha_k = 1 - \gamma$, $k = 1, \ldots, K$, to resemble
 \citeauthor{Friedman:1989tm}'s parameterization, and we estimated both $\lambda$
@@
 We did not include the \emph{RDA} classifier in our classification study because
-its training runtime was prohobitively slow on high-dimensional data in our
+its training runtime was prohibitively slow on high-dimensional data in our
 preliminary experiments. As shown in Section \ref{sec:timing-comparisons}, the
@@
 $\bm \mu_2 = (\underbrace{1/2, \ldots, 1/2}_{100}, \underbrace{0, \ldots, 0}_{p
-  - 100})$. For simplicity, we defined $\bm \mu_3 = -\bm \mu_2$. The 3
+  - 100})$. For simplicity, we defined $\bm \mu_3 = -\bm \mu_2$. The three
 populations differed in their mean vectors in the first 100 features
@@
 all $k$. Alternatively, to explore the more realistic assumption of unequal
-covariance matrices, we put $\rho_1 = 0.1$, $\rho_2 = 0.5$, and $\rho_3 = 0.9$
+covariance matrices, we put $\rho_1 = 0.1$, $\rho_2 = 0.5$, and $\rho_3 = 0.9$.
 
@@
 
-In Figure \ref{fig:sim-results-by-p} we again examine the simulation results as
-a function of $p$ for a subset of the values of
-$\epsilon$. This set of plots allows us to investigate the effect of feature
-dimensionality on classification performance. When no outliers were present
-(i.e., $\epsilon=0$), the random-forest classifier was outperformed by all other
-classifiers. Furthermore, the \emph{HDRDA} classifiers were superior in terms of
-average error rate in this setting. As $p$ increased, an elevation in average
-error rate was expected for all classifiers, but the increase was not observed
-to be substantial.
+In Figure \ref{fig:sim-results-by-p}, we again examine the simulation results as
+a function of $p$ for a subset of the values of $\epsilon$. This set of plots
+allows us to investigate the effect of feature dimensionality on classification
+performance. When no outliers were present (i.e., $\epsilon=0$), the
+random-forest classifier was outperformed by all other classifiers. Furthermore,
+the \emph{HDRDA} classifiers were superior in terms of average error rate in
+this setting. As $p$ increased, an elevation in average error rate was expected
+for all classifiers, but the increase was not observed to be substantial.
 
@@
 We compared the \emph{HDRDA} classifier to the five competing classifiers on six
-benchmark gene-expression microarray data sets. We evaluated the classification
-accuracy of each classifier by randomly partitioning the data set under
-consideration such that $2/3$ of the observations were allocated as training
-data and the remaining $1/3$ of the observations were allocated as a test data
-set. To expedite the computational runtime, we reduced the training data to the
-top 1000 variables by employing the variable-selection method proposed by
-\cite{Dudoit:2002ev}. We then reduced the test data set to the same 1000
-variables. After training each classifier on the training data, we classified
-the test data sets and computed the proportion of mislabeled test observations
-to estimate the classification error rate for each classifier. Repeating this
-process 100 times, we computed the average of the error-rate estimates for each
-classifier. We next provide a concise description of each high-dimensional data
-set examined in our classification study.
+benchmark gene-expression microarray data sets. First, we evaluated the
+classification accuracy of each classifier by randomly partitioning the data set
+under consideration such that $2/3$ of the observations were allocated as
+training data and the remaining $1/3$ of the observations were allocated as a
+test data set. To expedite the computational runtime, we reduced the training
+data to the top 1000 variables by employing the variable-selection method
+proposed by \cite{Dudoit:2002ev}. We then reduced the test data set to the same
+1000 variables. After training each classifier on the training data, we
+classified the test data sets and computed the proportion of mislabeled test
+observations to estimate the classification error rate for each classifier.
+Repeating this process 100 times, we computed the average of the error-rate
+estimates for each classifier. We next provide a concise description of each
+high-dimensional data set examined in our classification study.
 
@@
 \cite{Chiaretti:2004gq} measured the gene-expression profiles for 128
-individuals with acute lymphoblastic leukemia (ALL) using Affymetrix human
-95Av2 arrays. Following \cite{Xu:2009fl}, we restricted the data set to $K = 2$
+individuals with acute lymphoblastic leukemia (ALL) using Affymetrix human 95Av2
+arrays. Following \cite{Xu:2009fl}, we restricted the data set to $K = 2$
 classes such that $n_1 = 74$ observations were without cytogenetic abnormalities
@@
 WDLS, and part of DDLS; (4) Spindle cell and pleomorophic sarcomas including
-DDLS, MFS, LMS, MPNST, FS, and MFH. Following \cite{Witten:2011kc}, we restricted
-our analysis to the five tumor types having at least 15 observations.
+DDLS, MFS, LMS, MPNST, FS, and MFH. Following \cite{Witten:2011kc}, we
+restricted our analysis to the five tumor types having at least 15 observations.
 
@@
 competitive within our classification study and perhaps superior. Contrarily,
-our empirical studies suggests that the classifier is sensitive to outliers and
+our empirical studies suggest that the classifier is sensitive to outliers and
 unable to achieve comparable results with other classifiers designed for
@@
 R package can be used in practice with high-dimensional data sets. In our timing
-comparisons we showed that \emph{HDRDA} model selection could be employed on
+comparisons, we showed that \emph{HDRDA} model selection could be employed on
 data sets with $p = 5000$ in \Sexpr{hdrda_mean_time_p5000} seconds on
@@
 \citep{Ching:2012fu,Li:2012ev,Tai:2007bk,Guo:2007te}, facial recognition
-\citep{Zhang:2010va,Dai:2007vd,Lu:2005hq,Pima:2004vw,Lu:2003we},
-handwritten digit recognition \citep{Bouveyron:2007gx}, remote sensing
+\citep{Zhang:2010va,Dai:2007vd,Lu:2005hq,Pima:2004vw,Lu:2003we}, handwritten
+digit recognition \citep{Bouveyron:2007gx}, remote sensing
 \citep{Tadjudin:1999fk}, seismic detection \citep{Anderson:2002kg}, and chemical

### 2017-02-05 reports/paper-hdrda.Rnw
Shortened abstract/removed line numbers for Pattern Recognition.

@@
 \usepackage{graphicx, amsmath, amssymb, bm, mathtools, amsthm, natbib}
-\usepackage{algorithm2e, url, lineno}
+\usepackage{algorithm2e, url}
 
@@
 
-\title{Computationally Efficient High-Dimensional Regularized Discriminant Analysis}
+\title{High-Dimensional Regularized Discriminant Analysis}
 
@@
 
-Regularized discriminant analysis (\emph{RDA}), proposed by
-\cite{Friedman:1989tm}, is a widely popular classifier that lacks
-interpretability and is impractical for high-dimensional data sets.  Here, we
-present an interpretable and computationally efficient classifier called
-high-dimensional \emph{RDA} (\emph{HDRDA}), designed for the small-sample,
-high-dimensional setting. For \emph{HDRDA}, we show that each training
-observation, regardless of class, contributes to the class covariance matrix,
-resulting in an interpretable estimator that \emph{borrows} from the pooled
-sample covariance matrix. Moreover, we show that \emph{HDRDA} is equivalent to
-a classifier in a reduced-feature space with dimension approximately equal to
-the training sample size. As a result, the matrix operations employed by
-\emph{HDRDA} are computationally linear in the number of features, making the
-classifier well-suited for high-dimensional classification in practice. We
-demonstrate that \emph{HDRDA} is often superior to several sparse and
-regularized classifiers in terms of classification accuracy with three
-artificial and six real high-dimensional data sets. Also, timing comparisons
-between our \emph{HDRDA} implementation in the {\tt sparsediscrim} R package
-and the standard \emph{RDA} formulation in the {\tt klaR} R package demonstrate
-that as the number of features increases, the computational runtime of
-\emph{HDRDA} is drastically smaller than that of \emph{RDA}.
+Regularized discriminant analysis (\emph{RDA}) is a widely popular classifier that lacks interpretability and is impractical for high-dimensional data.
+Here, we present a competing classifier called high-dimensional \emph{RDA} (\emph{HDRDA}), designed for the small-sample, high-dimensional setting.
+For \emph{HDRDA}, we show that each training observation, regardless of class, contributes to the class covariance matrix, resulting in an interpretable estimator that \emph{borrows} from the pooled sample covariance matrix.
+Moreover, we show that \emph{HDRDA} is equivalent to a classifier in a reduced-feature space with dimension approximately equal to the sample size.
+Consequently, the matrix operations employed by \emph{HDRDA} are linear in the number of features.
+Timing comparisons between \emph{HDRDA} and RDA establish that as the number of features increases, the computational runtime of \emph{HDRDA} is drastically smaller than that of \emph{RDA}.
+Also, we demonstrate that \emph{HDRDA} often attains superior classification accuracy to several sparse and regularized classifiers using three artificial and six real high-dimensional datasets.
 
@@
 \end{frontmatter}
-
-\linenumbers
 
@@
 $\omega_k$ with the sample \emph{QDA} classifier
-\begin{linenomath*}
 \begin{align}
@@
 \end{align}
-\end{linenomath*}
 where $\xbar_k$ and $\widehat{\bm \Sigma}_k$ are the maximum-likelihood
@@
   \bm \Sigma}_k$ in \eqref{eq:qda}, where
-\begin{linenomath*}
 \begin{align}
@@
 \end{align}
-\end{linenomath*}
 is the MLE for $\bm \Sigma$. Here, \eqref{eq:qda} reduces to the sample
@@
 regularization applies a shrinkage factor $\gamma > 0$, so that
-\begin{linenomath*}
 \begin{align}
@@
 \end{align}
-\end{linenomath*}
 similar to a method employed in ridge regression \citep{Hoerl:1970cd}. Equation
@@
 the convex combination
-\begin{linenomath*}
 \begin{align}
@@
 \end{align}
-\end{linenomath*}
 where $\lambda \in [0, 1]$ is the \emph{pooling} parameter. By rewriting
@@
 interpretation of $\widehat{\bm \Sigma}_k(\lambda)$. That is,
-\begin{linenomath*}
 \begin{align}
@@
 \end{align}
-\end{linenomath*}
 where $c_{ik}(\lambda) = \lambda N^{-1} + (1 - \lambda)n_k^{-1}I(y_i = k)$.
@@
 \eqref{eq:sig-lambda-alternative} as
-\begin{linenomath*}
 \begin{align}
@@
 \end{align}
-\end{linenomath*}
 where $\alpha_k \ge 0$ and $\gamma \ge 0$ is an eigenvalue-shrinkage
@@
 \emph{HDRDA} classifier as
-\begin{linenomath*}
 \begin{align}
@@
 \end{align}
-\end{linenomath*}
 For $\gamma > 0$, $\tilde{\bm \Sigma}_k$ is nonsingular such that $\tilde{\bm
@@
 When $\alpha_k = 1 - \gamma$, \eqref{eq:hdrda-cov} resembles the biased covariance-matrix estimator
-\begin{linenomath*}
 \begin{align}
@@
 \end{align}
-\end{linenomath*}
 employed in the \emph{RDA} classifier, where $\widehat{\bm
@@
 q \le p$. Then, we have
-\begin{linenomath*}
 \begin{align}
@@
 \end{align}
-\end{linenomath*}
 \end{lemma}
@@
 the decision rule in \eqref{eq:hdrda} is equivalent to
-\begin{linenomath*}
 \begin{align}
@@
 \end{align}
-\end{linenomath*}
 \end{thm}
@@
 K$. Therefore, for all $\bm x \in \mathbb{R}_{p \times 1}$, we have that
-\begin{linenomath*}
 \begin{align*}
@@
 \end{align*}
-\end{linenomath*}
 Because $\gamma$ is constant for $k = 1, \ldots, K$, we can omit the $(p - q)
@@
 and
-\begin{linenomath*}
 \begin{align}
@@
 \end{align}
-\end{linenomath*}
 \end{proposition}
@@
 from \cite{Guo:2007te},
-\begin{linenomath*}
 \begin{align}
@@
 \end{align}
-\end{linenomath*}
 where the $(i,j)$th entry of the block matrix $\bm\Sigma^{(\rho_k)} \in
 \mathbb{R}_{100 \times 100}$ is
-\begin{linenomath*}
 \begin{align*}
@@
 \end{align*}
-\end{linenomath*}
 The block-diagonal covariance structure in \eqref{eq:block-sigma} resembles
@@
 \Sigma)$, we write the PDF of the $k$th class as
-\begin{linenomath*}
 \begin{align}
@@
 \end{align}
-\end{linenomath*}
 where $\epsilon \in [0, 1]$ is the probability that an observation is

### 2017-02-05 reports/paper-hdrda.Rnw
Added Acknowledgements section per Pattern Recognition

@@
 
-Regularized discriminant analysis (\emph{RDA}) is a widely popular classifier that lacks interpretability and is impractical for high-dimensional data.
-Here, we present a competing classifier called high-dimensional \emph{RDA} (\emph{HDRDA}), designed for the small-sample, high-dimensional setting.
-For \emph{HDRDA}, we show that each training observation, regardless of class, contributes to the class covariance matrix, resulting in an interpretable estimator that \emph{borrows} from the pooled sample covariance matrix.
-Moreover, we show that \emph{HDRDA} is equivalent to a classifier in a reduced-feature space with dimension approximately equal to the sample size.
-Consequently, the matrix operations employed by \emph{HDRDA} are linear in the number of features.
-Timing comparisons between \emph{HDRDA} and RDA establish that as the number of features increases, the computational runtime of \emph{HDRDA} is drastically smaller than that of \emph{RDA}.
-Also, we demonstrate that \emph{HDRDA} often attains superior classification accuracy to several sparse and regularized classifiers using three artificial and six real high-dimensional datasets.
+Regularized discriminant analysis (\emph{RDA}) is a widely popular classifier
+that lacks interpretability and is impractical for high-dimensional data. Here,
+we present a competing classifier called high-dimensional \emph{RDA}
+(\emph{HDRDA}), designed for the small-sample, high-dimensional setting. For
+\emph{HDRDA}, we show that each training observation, regardless of class,
+contributes to the class covariance matrix, resulting in an interpretable
+estimator that \emph{borrows} from the pooled sample covariance
+matrix. Moreover, we show that \emph{HDRDA} is equivalent to a classifier in a
+reduced-feature space with dimension approximately equal to the sample
+size. Consequently, the matrix operations employed by \emph{HDRDA} are linear
+in the number of features. Timing comparisons between \emph{HDRDA} and RDA
+establish that as the number of features increases, the computational runtime
+of \emph{HDRDA} is drastically smaller than that of \emph{RDA}. Also, we
+demonstrate that \emph{HDRDA} often attains superior classification accuracy to
+several sparse and regularized classifiers using three artificial and six real
+high-dimensional datasets.
 
@@
 
-We thank Mrs. Joy Young for her numerous recommendations that enhanced the
-quality of our writing.
+\section{Acknowledgements}
+
+This research did not receive any specific grant from funding agencies in the
+public, commercial, or not-for-profit sectors. We thank Mrs. Joy Young for her
+numerous recommendations that enhanced the quality of our writing.
 

### 2017-02-05 reports/title-page.tex
Added title page for Pattern Recognition.

--- 2017-02-05 reports/paper-hdrda.Rnw
@@
 \documentclass[review,authoryear]{elsarticle}
-
-\usepackage{graphicx, amsmath, amssymb, bm, mathtools, amsthm, natbib}
-\usepackage{algorithm2e, url}
-
-\bibpunct{(}{)}{;}{a}{,}{,}
-
-\newcommand{\xbar}{\bar{\bm x}}
-\newcommand{\tr}{\text{tr}}
-\DeclareMathOperator*{\argmin}{arg\,min}
-\DeclareMathOperator*{\argmax}{arg\,max}
-\newtheorem{cor}{Corollary}
-\newtheorem{lemma}{Lemma}
-\newtheorem{proposition}{Proposition}
-\newtheorem{thm}{Theorem}
+\usepackage{url}
 
@@
 
-\author[jar]{John A. Ramey\cortext[cor1]{Corresponding author}}
+\author[jar]{John A. Ramey}
 \author[cks]{Caleb K. Stein}
@@
 \author[dmy]{Dean M. Young}
-\address[jar]{Novi Labs}
-\address[cks]{Myeloma Institute, University of Arkansas for Medical Sciences}
-\address[pdy]{Department of Management and Information Systems, Baylor University}
-\address[dmy]{Department of Statistical Science, Baylor University}
-
-\begin{abstract}
-
-Regularized discriminant analysis (\emph{RDA}) is a widely popular classifier that lacks interpretability and is impractical for high-dimensional data.
-Here, we present a competing classifier called high-dimensional \emph{RDA} (\emph{HDRDA}), designed for the small-sample, high-dimensional setting.
-For \emph{HDRDA}, we show that each training observation, regardless of class, contributes to the class covariance matrix, resulting in an interpretable estimator that \emph{borrows} from the pooled sample covariance matrix.
-Moreover, we show that \emph{HDRDA} is equivalent to a classifier in a reduced-feature space with dimension approximately equal to the sample size.
-Consequently, the matrix operations employed by \emph{HDRDA} are linear in the number of features.
-Timing comparisons between \emph{HDRDA} and RDA establish that as the number of features increases, the computational runtime of \emph{HDRDA} is drastically smaller than that of \emph{RDA}.
-Also, we demonstrate that \emph{HDRDA} often attains superior classification accuracy to several sparse and regularized classifiers using three artificial and six real high-dimensional datasets.
-
-\end{abstract}
-
-\begin{keyword}
-Regularized discriminant analysis\sep High-dimensional classification\sep Covariance-matrix regularization\sep Singular value decomposition\sep Multivariate analysis\sep Dimension reduction
-\MSC[2010] 62H30\sep 65F15\sep 65F20\sep 65F22
-\end{keyword}
+\address[jar]{\textbf{Corresponding Author} \\ Novi Labs \\ 7020 Easy Wind Drive, Suite 210 \\ Austin, TX 78752 \\ USA \\ Email: johnramey@gmail.com}
+\address[cks]{Myeloma Institute \\ University of Arkansas for Medical Sciences \\ 4018 W. Capitol Avenue \\ Little Rock, AR 72205 \\ USA \\ Email: stein.caleb@gmail.com}
+\address[pdy]{Department of Management and Information Systems \\ Baylor University \\ One Bear Place \#98005 \\ Waco, TX 76798-8005 \\ USA \\ Email: Philip\_Young@baylor.edu}
+\address[dmy]{Department of Statistical Science \\ Baylor University \\ P.O. Box 97140 \\ Waco, TX 76798-7140 \\ USA \\ Email: Dean\_Young@baylor.edu}
 
@@
 
-<<knitr_setup, include=FALSE, echo=FALSE, cache=FALSE>>=
-library(knitr)
-# Round digits throughout document including \Sexpr statements.
-# See Yihui's response here:
-# http://stackoverflow.com/questions/11062497/how-to-avoid-using-round-in-every-sexpr
-options(digits=3)
-
-opts_chunk$set(fig.align='default', dev='png', message=FALSE, warning=FALSE,
-               cache=TRUE, echo=FALSE, autodep=TRUE, fig.width=12,
-               fig.height=12, out.width="\\linewidth")
-@
-
-<<setup>>=
-setwd("..")
-library(ProjectTemplate)
-load.project()
-@
-
-<<munge_contaminated_sims>>=
-# Contaminated Normal Sims
-results_contaminated <- lapply(results_block_contaminated[[1]], function(r) {
-  with(r, c(error_rates, p=p, mu=mu, contamination_prob=contamination_prob))
-})
-results_contaminated <- do.call(rbind.data.frame, results_contaminated)
-results_contaminated <- tbl_df(results_contaminated)
-
-results_contaminated <- results_contaminated %>%
-  group_by(p, contamination_prob, mu) %>%
-  summarise_each(funs(mean)) %>%
-  arrange(p, contamination_prob, mu) %>%
-  ungroup
-
-results_contaminated <- results_contaminated %>%
-  gather(Classifier, Error, -p, -contamination_prob, -mu) %>%
-  mutate(p=factor(p)) %>%
-  filter(mu == 0.5)
-
-# Remove underscore from classifier names
-levels(results_contaminated$Classifier) <- gsub("_", " ",
-                                                levels(results_contaminated$Classifier))
-
-# Maximum Standard Error for Contaminated Sims
-max_sim_se <- with(results_contaminated, max(sqrt(Error * (1 - Error) / 500)))
-max_sim_se <- round(max_sim_se, digits=3)
-@
-
-<<munge_microarray_results>>=
-# Microarray Classification Results
-results_microarray <- lapply(results[[1]], function(r) {
-  with(r, c(error_rates, `data_set`=data_set))
-})
-results_microarray <- do.call(rbind.data.frame, results_microarray)
-results_microarray <- tbl_df(results_microarray)
-
-avg_error_microarray <- results_microarray %>%
-    group_by(data_set) %>%
-    summarise_each(funs(mean)) %>%
-    gather(Classifier, Avg_Error, -data_set)
-
-sd_error_microarray <- results_microarray %>%
-  group_by(data_set) %>%
-  summarise_each(funs(sd)) %>%
-  gather(Classifier, SD_Error, -data_set)
-
-# Microarray classification results are presented in a table
-# Rows = classifier, columns = data set
-# Cells = Error (SD) rounded to 3 digits
-# The classifier with the minimum average error rate is in bold for each data set.
-
-bold_latex <- function(x) {
-  paste0('\\textbf{', x, '}')
-}
-
-results_microarray <- inner_join(avg_error_microarray, sd_error_microarray,
-                                 by=c('data_set', 'Classifier')) %>%
-  mutate(Table_Cells=sprintf('%.3f', Avg_Error),
-         data_set=stringr::str_to_title(data_set),
-         Classifier=gsub("_", " ", Classifier)) %>%
-  group_by(data_set) %>%
-  mutate(Table_Cells=ifelse(Avg_Error == min(Avg_Error), bold_latex(Table_Cells), Table_Cells),
-         Table_Cells=sprintf('%s (%.3f)', Table_Cells, SD_Error)) %>%
-  select(-Avg_Error, -SD_Error) %>%
-  spread(data_set, Table_Cells)
-@
-
-<<munge_timing_comparisons>>=
-# HDRDA vs RDA Timing Comparison Results
-
-# Benchmarks are given in nanoseconds, which are converted to seconds
-timings <- bind_rows(lapply(timing_results, as.data.frame), .id="p") %>%
-  mutate(p=as.integer(p),
-         expr=as.character(expr),
-         Classifier=ifelse(startsWith(expr, "hdrda"), "HDRDA", "RDA"),
-         Classifier=factor(Classifier),
-         time=time / 1e9) %>%
-  select(-expr) %>%
-  filter(p %% 500 == 0) %>%
-  tbl_df %>%
-  group_by(p, Classifier) %>%
-  mutate(wls_weights=1 / sd(time)) %>%
-  ungroup
-
-# Create DF for two facets in same plot
-# First plot shows HDRDA vs RDA
-# Second plot zooms in on HDRDA results
-timings$Plot <- "Comparison of HDRDA and RDA"
-timings_hdrda <- timings %>%
-    filter(Classifier == "HDRDA") %>%
-    mutate(Plot="HDRDA")
-timings_combined <- bind_rows(timings, timings_hdrda)
-
-# Build data frame to show relative timings in table
-timing_ratios <- timings %>%
-    group_by(p, Classifier) %>%
-    summarize(median_time=median(time)) %>%
-    tidyr::spread(Classifier, median_time) %>%
-      mutate(ratio_time=RDA / HDRDA) %>%
-      select(-HDRDA, -RDA)
-
-timings_summary <- timings %>%
-  group_by(Classifier, p) %>%
-  summarize(mean_runtime=mean(time), sd_runtime=sd(time))
-
-# RDA Mean Runtime in Minutes when p = 5000
-rda_mean_time_p5000 <- timings_summary %>% filter(Classifier == "RDA", p == 5000)
-rda_mean_time_p5000 <- rda_mean_time_p5000$mean_runtime / 60
-
-# HDRDA Mean Runtime in Seconds when p = 5000
-hdrda_mean_time_p5000 <- timings_summary %>% filter(Classifier == "HDRDA", p == 5000)
-hdrda_mean_time_p5000 <- hdrda_mean_time_p5000$mean_runtime
-@
-
-<<bootstrap_relative_timing_comparisons>>=
-ratio_means <- function(df, num_reps=1000) {
-  bootstrap_reps <- replicate(n=num_reps, {
-    df %>%
-      group_by(Classifier) %>%
-      sample_frac(size=1.0, replace=TRUE) %>%
-      summarize(mean_time=mean(time)) %>%
-      tidyr::spread(Classifier, mean_time) %>%
-      mutate(ratio_means=RDA / HDRDA) %>%
-      select(-HDRDA, -RDA)
-  }, simplify=FALSE)
-  bind_rows(bootstrap_reps)
-}
-
-set.seed(42)
-timing_ratios_bootstrap <- timings %>%
-    mutate(p=factor(p)) %>%
-    group_by(p) %>%
-    do(ratio_means(., num_reps=1000))
-
-timing_bootstrap_summary <- timing_ratios_bootstrap %>%
-  group_by(p) %>%
-  summarize(avg_ratio=mean(ratio_means),
-            lower_ci=quantile(ratio_means, probs=0.025),
-            upper_ci=quantile(ratio_means, probs=0.975))
-bootstrap_summary_p500 <- timing_bootstrap_summary %>%
-  filter(p == "500")
-bootstrap_summary_p5000 <- timing_bootstrap_summary %>%
-  filter(p == "5000")
-
-@
-
-\section{Introduction}
-
-In this paper, we consider the classification of small-sample, high-dimensional
-data, where the number of features $p$ exceeds the training sample size $N$. In
-this setting, well-established classifiers, such as linear discriminant analysis
-(\emph{LDA}) and quadratic discriminant analysis (\emph{QDA}), become
-incalculable because the class and pooled covariance matrix estimators are
-singular \citep*{Murphy:2012uq, Bouveyron:2007gx, Mkhadri:1997gy}. To improve
-the accuracy of the estimation of the class covariance matrices estimated in the
-\emph{QDA} classifier and to ensure that the covariance matrix estimators are
-nonsingular, \cite{Friedman:1989tm} proposed the regularized discriminant
-analysis (\emph{RDA}) classifier by incorporating a weighted average of the
-pooled sample covariance matrix and the class sample covariance matrix. To
-further improve the accuracy of the estimation of the class covariance matrix
-and to stabilize its inverse, \cite{Friedman:1989tm} also included a
-regularization component by shrinking the covariance matrix estimator towards
-the identity matrix, which yields a nonsingular estimator following the
-well-known ridge-regression approach of \cite{Hoerl:1970cd}. Despite its
-popularity, the ``borrowing'' operation employed in the \emph{RDA} classifier
-lacks interpretability \citep{Bensmail:1996fw}. Furthermore, the \emph{RDA}
-classifier is impractical for high-dimensional data sets because it computes the
-inverse and determinant of the covariance matrices for each class. Both matrix
-calculations are computationally expensive because the number of operations
-grows at a polynomial rate in the number of features. Moreover, the model
-selection of the \emph{RDA} classifier's two tuning parameters is
-computationally burdensome because the matrix inverse and determinant of each
-class covariance matrix are computed across multiple cross-validation folds for
-each candidate tuning-parameter pair.
-
-Here, we present the high-dimensional \emph{RDA} (\emph{HDRDA}) classifier,
-which is intended for the case when $p > N$ . We reparameterize the \emph{RDA}
-classifier similar to that of \cite*{Hastie:2008dt} and \cite{Halbe:2007ea} and
-employ a biased covariance-matrix estimator that partially pools the individual
-sample covariance matrices from the \emph{QDA} classifier with the pooled sample
-covariance matrix from the \emph{LDA} classifier. We then shrink the resulting
-covariance-matrix estimator towards a scaled identity matrix to ensure positive
-definiteness. We show that the pooling parameter in the \emph{HDRDA} classifier
-determines the contribution of each training observation to the estimation of
-each class covariance matrix, enabling interpretability that has been previously
-lacking with the \emph{RDA} classifier \citep{Bensmail:1996fw}. Our
-parameterization differs from that of \cite*{Hastie:2008dt} and that of
-\cite{Halbe:2007ea} in that our formulation allows the flexibility of various
-covariance-matrix estimators proposed in the literature, including a variety of
-ridge-like estimators, such as the one proposed by \cite{Srivastava:2007ww}.
-
-Next, we establish that the matrix operations corresponding to the null space of
-the pooled sample covariance matrix are redundant and can be discarded from the
-\emph{HDRDA} decision rule without loss of classificatory information when we
-apply reasoning similar to that of \cite{Ye:2006tq}. As a result, we achieve a
-substantial reduction in dimension such that the matrix operations used in the
-\emph{HDRDA} classifier are computationally linear in the number of features.
-Furthermore, we demonstrate that the \emph{HDRDA} decision rule is invariant to
-adjustments to the approximately $p - N$ zero eigenvalues, so that the decision
-rule in the original feature space is equivalent to a decision rule in a lower
-dimension, such that matrix inverses and determinants of relatively small
-matrices can be rapidly computed. Finally, we show that several shrinkage
-methods that are special cases of the \emph{HDRDA} classifier have no effect on
-the approximately $p - N$ zero eigenvalues of the covariance-matrix estimators
-when $p > N$. Such techniques include work from \cite{Srivastava:2007ww},
-\cite{Rao:1971ul}, and other methods studied by \cite{Ramey:2013ji} and
-\citet*{Xu:2009fl}.
-
-We also provide an efficient algorithm along with pseudocode to estimate the
-\emph{HDRDA} classifier's tuning parameters in a grid search via
-cross-validation. Timing comparisons between our \emph{HDRDA} implementation in
-the {\tt sparsediscrim} R package available on CRAN and the standard \emph{RDA}
-formulation in the {\tt klaR} R package demonstrate that as the number of
-features increases, the computational runtime of the \emph{HDRDA} classifier is
-drastically smaller than that of \emph{RDA}. In fact, when $p = 5000$, we show
-that the \emph{HDRDA} classifier is \Sexpr{bootstrap_summary_p5000$avg_ratio}
-times faster on average than the \emph{RDA} classifier. In this scenario, the
-\emph{HDRDA} classifier's model selection requires \Sexpr{hdrda_mean_time_p5000}
-seconds on average, while that of the \emph{RDA} classifier requires
-\Sexpr{rda_mean_time_p5000} minutes on average.
-
-Finally, we study the classification performance of the \emph{HDRDA} classifier
-on six real high-dimensional data sets along with a simulation design that
-generalizes the experiments initially conducted by \cite{Guo:2007te}. We
-demonstrate that the \emph{HDRDA} classifier often attains superior
-classification accuracy to several recent classifiers designed for small-sample,
-high-dimensional data from \cite*{Tong:2012hw}, \cite{Witten:2011kc},
-\cite*{Pang:2009ik}, and \cite{Guo:2007te}. We also include as a benchmark the
-random forest from \cite{Breiman:2001fb} because \cite*{FernandezDelgado:2014ul}
-have concluded that the random forest is often superior to other classifiers in
-benchmark studies. We show that our proposed classifier is competitive and often
-outperforms the random forest in terms of classification accuracy in the
-small-sample, high-dimensional setting.
-
-The remainder of this paper is organized as follows. In Section
-\ref{sec:preliminaries} we introduce the classification problem and necessary
-notation to describe our contributions. In Section \ref{sec:rda} we present the
-\emph{HDRDA} classifier along with its interpretation. In Section
-\ref{sec:hdrda-properties}, we provide properties of the \emph{HDRDA} classifier
-and a computationally efficient model-selection procedure. In Section
-\ref{sec:timing-comparisons}, we compare the model-selection timings of the
-\emph{HDRDA} and \emph{RDA} classifiers. In Section \ref{sec:sims} we describe
-our simulation studies of artificial and real data sets and examine the
-experimental results. We conclude with a brief discussion in Section
-\ref{sec:discussion}.
-
-\section{Preliminaries}
-\label{sec:preliminaries}
-
-\subsection{Notation}
-
-To facilitate our discussion of covariance-matrix regularization and
-high-dimensional classification, we require the following notation. Let
-$\mathbb{R}_{a \times b}$ denote the matrix space of all $a \times b$ matrices
-over the real field $\mathbb{R}$. Denote by $\bm I_m$ the $m \times m$ identity
-matrix, and let $\bm 0_{m \times p}$ be the $m \times p$ matrix of zeros, such
-that $\bm 0_m$ is understood to denote $\bm 0_{m \times m}$. Define $\bm 1_m \in
-\mathbb{R}_{m \times 1}$ as a vector of ones. Let $\bm A^{T}$, $\bm A^+$, and
-$\mathcal{N}(\bm A)$ denote the transpose, the Moore-Penrose pseudoinverse, and
-the null space of $\bm A \in \mathbb{R}_{m \times p}$, respectively. Denote by
-$\mathbb{R}_{p \times p}^{>}$ the cone of real $p \times p$ positive-definite
-matrices. Similarly, let $\mathbb{R}_{p \times p}^{\ge}$ denote the cone of real
-$p \times p$ positive-semidefinite matrices. Let $V^{\perp}$ denote the
-orthogonal complement of a vector space $V \subset \mathbb{R}_{p \times 1}$. For
-$c \in \mathbb{R}$, let $c^+ = 1/c$ if $c \ne 0$ and $0$ otherwise.
-
-\subsection{Discriminant Analysis}
-
-In discriminant analysis we wish to assign an unlabeled vector $\bm x \in
-\mathbb{R}_{p \times 1}$ to one of $K$ unique, known classes by constructing a
-classifier from $N$ training observations. Let $\bm x_i = (x_{i1}, \ldots,
-x_{ip}) \in \mathbb{R}_{p \times 1}$ be the $i$th observation $(i = 1, \ldots,
-N)$ with true, unique membership $y_i \in \{\omega_1, \ldots, \omega_K\}$.
-Denote by $n_k$ the number of training observations realized from class $k$,
-such that $\sum_{k=1}^K n_k = N$. We assume that $(\bm x_i, y_i)$ is a
-realization from a mixture distribution $p(\bm x) = \sum_{k=1}^K p(\bm x |
-\omega_k) p(\omega_k)$, where $p(\bm x | \omega_k)$ is the probability density
-function (PDF) of the $k$th class and $p(\omega_k)$ is the prior probability of
-class membership of the $k$th class. We further assume $p(\omega_k) =
-p(\omega_l)$, $1 \le k, l \le K$, $k \ne l$.
-
-The \emph{QDA} classifier is the optimal Bayesian decision rule with respect to
-a $0-1$ loss function when $p(\bm x | \omega_k)$ is the PDF of the multivariate
-normal distribution with known mean vectors $\bm \mu_k \in \mathbb{R}_{p \times
-  1}$ and known covariance matrices $\bm \Sigma_k \in \mathbb{R}_{p \times
-  p}^{>}$, $k = 1, 2, \ldots, K$. Because $\bm \mu_k$ and $\bm \Sigma_k$ are
-typically unknown, we assign an unlabeled observation $\bm x$ to class
-$\omega_k$ with the sample \emph{QDA} classifier
-\begin{align}
-	D_{QDA}(\bm x) = \argmin_{k} (\bm x - \xbar_k)^{T}\widehat{\bm
-    \Sigma}_k^{-1}(\bm x - \xbar_k) + \log |\widehat{\bm \Sigma}_k|, \label{eq:qda}
-\end{align}
-where $\xbar_k$ and $\widehat{\bm \Sigma}_k$ are the maximum-likelihood
-estimators (MLEs) of $\bm \mu_k$ and $\bm \Sigma_k$, respectively. If we assume
-further that $\bm \Sigma_k = \bm \Sigma$, $k = 1, \ldots, K$, then the pooled
-sample covariance matrix $\widehat{\bm \Sigma}$ is substituted for $\widehat{
-  \bm \Sigma}_k$ in \eqref{eq:qda}, where
-\begin{align}
-	\widehat{\bm \Sigma} = N^{-1} \sum_{k=1}^K n_k \widehat{\bm \Sigma}_k \label{eq:pooled-cov}
-\end{align}
-is the MLE for $\bm \Sigma$. Here, \eqref{eq:qda} reduces to the sample
-\emph{LDA} classifier. We omit the log-determinant because it is constant across
-the $K$ classes.
-
-The smallest eigenvalues of $\widehat{\bm \Sigma}_k$ and the directions
-associated with their eigenvectors can highly influence the classifier in
-\eqref{eq:qda}. In fact, the eigenvalues of $\widehat{ \bm \Sigma}_k$ are
-well-known to be biased if $p \ge n_k$ such that the smallest eigenvalues are
-underestimated \citep{Seber:2004uh}. Moreover, if $p > n_k$, then
-rank$(\widehat{\bm \Sigma}_k) \le n_k$, which implies that at least $p - n_k$
-eigenvalues of $\widehat{\bm \Sigma}_k$ are zero. Furthermore, although more
-feature information is available to discriminate among the $K$ classes, if $p >
-n_k$, \eqref{eq:qda} is incalculable because $\widehat{\bm \Sigma}_k^{-1}$ does
-not exist.
-
-Several regularization methods, such as the methods considered by
-\cite{Xu:2009fl}, \cite{Guo:2007te}, and \cite{Mkhadri:1995jp}, have been
-proposed in the literature to adjust the eigenvalues of $\widehat{\bm \Sigma}_k$
-so that \eqref{eq:qda} is calculable and provides reduced variability for
-$\widehat{\bm \Sigma}_k^{-1}$. A common form of the covariance-matrix
-regularization applies a shrinkage factor $\gamma > 0$, so that
-\begin{align}
-	\widehat{\bm \Sigma}_k(\gamma) = \widehat{\bm \Sigma}_k + \gamma \bm I_p, \label{eq:ridge-estimator}
-\end{align}
-similar to a method employed in ridge regression \citep{Hoerl:1970cd}. Equation
-\eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance
-matrix $\widehat{\bm \Sigma}_k$ toward $\bm I_p$, thereby increasing the
-eigenvalues of $\widehat{\bm \Sigma}_k$ by $\gamma$. Specifically, the zero
-eigenvalues are replaced with $\gamma$, so that \eqref{eq:ridge-estimator} is
-positive definite. For additional covariance-matrix regularization methods, see
-\cite{Ramey:2013ji}, \cite{Xu:2009fl}, and \cite{Ye:2009gd}.
-
-\section{High-Dimensional Regularized Discriminant Analysis}
-\label{sec:rda}
-
-Here, we define the \emph{HDRDA} classifier by first formulating the
-covariance-matrix estimator $\widehat{\bm \Sigma}_k(\lambda)$ and demonstrating
-its clear interpretation as a linear combination of the crossproducts of the
-training observations centered by their respective class sample means. We define
-the convex combination
-\begin{align}
-  \widehat{\bm \Sigma}_k(\lambda) \coloneqq (1 - \lambda) \widehat{\bm \Sigma}_k + \lambda
-  \widehat{\bm \Sigma}, \quad k = 1, \ldots, K,\label{eq:sig-lambda-alternative}
-\end{align}
-where $\lambda \in [0, 1]$ is the \emph{pooling} parameter. By rewriting
-\eqref{eq:sig-lambda-alternative} in terms of the observations $\bm x_i$, $i =
-1, \ldots, N$, each centered by its class sample mean, we attain a clear
-interpretation of $\widehat{\bm \Sigma}_k(\lambda)$. That is,
-\begin{align}
-	\widehat{\bm \Sigma}_k(\lambda)
-  &= \left( 1 - \lambda + \frac{\lambda n_k}{N} \right) \widehat{\bm \Sigma}_k +  \frac{\lambda}{N} \sum_{\substack{k' = 1\\k' \ne k}}^K n_{k'} \widehat{\bm \Sigma}_{k'} \nonumber \\
-	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N
-  I(y_i = k) \bm x_i \bm x_i^{T} +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k)
-  \bm x_i \bm x_i^{T} \nonumber \\
-	&= \sum_{i=1}^N c_{ik}(\lambda) \bm x_i \bm x_i^{T},\label{eq:sig-lambda-alternative2}
-\end{align}
-where $c_{ik}(\lambda) = \lambda N^{-1} + (1 - \lambda)n_k^{-1}I(y_i = k)$.
-From \eqref{eq:sig-lambda-alternative2}, we see that $\lambda$ weights the
-contribution of each of the $N$ observations in estimating $\bm \Sigma_k$ from
-all $K$ classes rather than using only the $n_k$ observations from a single
-class. As a result, we can interpret \eqref{eq:sig-lambda-alternative2} as a
-covariance-matrix estimator that borrows from $\widehat{\bm \Sigma}$ in
-\eqref{eq:pooled-cov} to estimate $ \bm \Sigma_k$.
-
-In Figure \ref{fig:hdrda-contours} we plot the contours of five multivariate
-normal populations for $\lambda = 0$ with unequal covariance matrices. As
-$\lambda$ approaches 1, the contours become more similar, resulting in identical
-contours for $\lambda = 1$. Below, we show that the pooling operation is
-advantageous in increasing the rank of each $\widehat{\bm \Sigma}_k(\lambda)$
-from rank$(\widehat{\bm \Sigma}_k)$ to rank$(\widehat{\bm \Sigma})$ for $0 <
-\lambda \le 1$. Notice that if $\lambda = 0$, then the observations from the
-remaining $K - 1$ classes do not contribute to the estimation of $\bm \Sigma_k$,
-corresponding to $\widehat{\bm \Sigma}_k$. Furthermore, if $\lambda = 1$, the
-weights $c_{ik}(\lambda)$ in \eqref{eq:sig-lambda-alternative2} reduce to $1/N$,
-corresponding to $\widehat{\bm \Sigma}$. For brevity, when $\lambda = 1$, we
-define $\bm X = [\sqrt{c_{1k}(1)} \bm x_1^{T}, \ldots, \sqrt{c_{Nk}(1)} \bm
-  x_N^{T}]^{T}$ such that $\widehat{\bm \Sigma} = N^{-1} \bm X^{T} \bm
-X$. Similarly, for $\lambda = 0$, we define $\bm X_k = [\sqrt{c_{1k}(0)} \bm
-  x_1^{T}, \ldots, \sqrt{c_{Nk}(0)} \bm x_N^{T}]^{T}$ such that $\widehat{\bm
-  \Sigma}_k = n_k^{-1} \bm X_k^{T} \bm X_k$.
-
-\[ \left[\text{Insert Figure \ref{fig:hdrda-contours} approximately here }\right] \]
-
-As we have discussed above, several eigenvalue adjustment methods have been
-proposed that increase eigenvalues (approximately) equal to 0. To further
-improve the estimation of $\bm \Sigma_k$ and to stabilize the estimator's
-inverse, we define the eigenvalue adjustment of
-\eqref{eq:sig-lambda-alternative} as
-\begin{align}
-	\tilde{\bm \Sigma}_k \coloneqq \alpha_k \widehat{\bm \Sigma}_k(\lambda) + \gamma \bm I_p,\label{eq:hdrda-cov}
-\end{align}
-where $\alpha_k \ge 0$ and $\gamma \ge 0$ is an eigenvalue-shrinkage
-constant. Thus, the \emph{pooling} parameter $\lambda$ controls the amount of
-estimation information borrowed from $\widehat{\bm \Sigma}$ to estimate $\bm
-\Sigma_k$, and the \emph{shrinkage} parameter $\gamma$ determines the degree of
-eigenvalue shrinkage. The choice of $\alpha_k$ allows for a flexible formulation
-of covariance-matrix estimators. For instance, if $\alpha_k = 1$, $k = 1,
-\ldots, K$, then \eqref{eq:hdrda-cov} resembles \eqref{eq:ridge-estimator}.
-Similarly, if $\alpha_k = 1 - \gamma$, then \eqref{eq:hdrda-cov} has a form
-comparable to the \emph{RDA} classifier from \cite{Friedman:1989tm}.
-Substituting \eqref{eq:hdrda-cov} into \eqref{eq:qda}, we define the
-\emph{HDRDA} classifier as
-\begin{align}
-	D_{HDRDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)^{T}\tilde{\bm \Sigma}_k^{+}(\bm x - \xbar_k)  + \log |\tilde{\bm \Sigma}_k|. \label{eq:hdrda}
-\end{align}
-For $\gamma > 0$, $\tilde{\bm \Sigma}_k$ is nonsingular such that $\tilde{\bm
-  \Sigma}_k^{-1}$ can be substituted for $\tilde{\bm \Sigma}_k^{+}$ in
-\eqref{eq:hdrda}. If $\gamma = 0$, we explicitly set $|\tilde{\bm \Sigma}_k|$
-equal to the product of the positive eigenvalues of $\tilde{\bm
-  \Sigma}_k$. Following \cite{Friedman:1989tm}, we select $\lambda$ and $\gamma$
-from a grid of candidate models via cross-validation \citep{Hastie:2008dt}. We provide an
-implementation of \eqref{eq:hdrda} in the {\tt hdrda} function contained in the
-{\tt sparsediscrim} R package, which is available on CRAN.
-
-The choice of $\alpha_k$ in \eqref{eq:hdrda-cov} is one of convenience and
-allows the flexibility of various covariance-matrix estimators proposed in the
-literature. In practice, we generally are not interested in estimating
-$\alpha_k$ because the estimation of $K$ additional tuning parameters via
-cross-validation is counterproductive to our goal of computational efficiency.
-For appropriate values of $\alpha_k$, the \emph{HDRDA} covariance-matrix
-estimator includes or resembles a large family of estimators. Notice that if
-$\alpha_k = 1$ and $\lambda = 1$, \eqref{eq:hdrda-cov} is equivalent to the
-standard ridge-like covariance-matrix estimator in \eqref{eq:ridge-estimator}.
-Other estimators proposed in the literature can be obtained when one selects
-$\gamma$ accordingly. For instance, with $\gamma = \tr\{ \widehat{\bm \Sigma} \}
-/ \min(N, p)$, we obtain the estimator from \cite{Srivastava:2007ww}.
-
-When $\alpha_k = 1 - \gamma$, \eqref{eq:hdrda-cov} resembles the biased covariance-matrix estimator
-\begin{align}
-    \widehat{\bm \Sigma}_k(\lambda, \gamma) = (1 - \gamma) \widehat{\bm \Sigma}_k^{(RDA)}(\lambda) + \gamma \frac{\tr\left\{\widehat{\bm \Sigma}_k^{(RDA)}(\lambda)\right\}}{p} \bm I_p \label{eq:sig-rda}
-\end{align}
-employed in the \emph{RDA} classifier, where $\widehat{\bm
-  \Sigma}_k^{(RDA)}(\lambda)$ is a pooled estimator of $\bm \Sigma_k$ and
-$\gamma \in [0, 1]$ is a regularization parameter that controls the shrinkage of
-\eqref{eq:sig-rda} towards $\bm I_p$ weighted by the average of the eigenvalues
-of $\widehat{\bm \Sigma}_k^{(RDA)}(\lambda)$. Despite the similarity of
-\eqref{eq:sig-rda} to the \emph{HDRDA} covariance-matrix estimator in
-\eqref{eq:hdrda-cov}, the \emph{RDA} classifier is impractical for
-high-dimensional data because the inverse and determinant of \eqref{eq:sig-rda}
-must be calculated when substituted into \eqref{eq:qda}. Furthermore,
-\eqref{eq:sig-rda} has no clear interpretation.
-
-\section{Properties of the HDRDA Classifier}
-\label{sec:hdrda-properties}
-
-Next, we establish properties of the covariance-matrix estimator and the
-decision rule employed in the \emph{HDRDA} classifier. By doing so, we
-demonstrate that \eqref{eq:hdrda} lends itself to a more efficient
-calculation. We decompose \eqref{eq:hdrda} into a sum of two components, where
-the first summand consists of matrix operations applied to low-dimensional
-matrices and the second summand corresponds to the null space of $\widehat{\bm
-  \Sigma}$ in \eqref{eq:pooled-cov}. We show that the matrix operations
-performed on the null space of $\widehat{\bm \Sigma}$ yield constant quadratic
-forms across all classes and can be omitted. For $p \gg N$, the constant
-component involves determinants and inverses of high-dimensional matrices, and
-by ignoring these calculations, we achieve a substantial reduction in
-computational costs. Furthermore, a byproduct is that adjustments to the
-associated eigenvalues have no effect on \eqref{eq:hdrda}. Lastly, we utilize
-the singular value decomposition to efficiently calculate the eigenvalue
-decomposition of $\widehat{\bm \Sigma}$, further reducing the computational
-costs of the \emph{HDRDA} classifier.
-
-First, we require the following relationship regarding the null spaces of
-$\widehat{\bm \Sigma}_k(\lambda)$, $\widehat{\bm \Sigma}$, and $\widehat{ \bm
-  \Sigma}_k$.
-
-\begin{lemma}\label{lemma:null-spaces}
-Let $\widehat{\bm \Sigma}_k$ and $\widehat{\bm \Sigma}$ be the MLEs of $\bm
-\Sigma_k$ and $\bm \Sigma$, respectively. Let $\widehat{\bm \Sigma}_k(\lambda)$
-be defined as in \eqref{eq:sig-lambda-alternative}. Then,
-$\mathcal{N}\{\widehat{\bm \Sigma}_k(\lambda)\} \subset \mathcal{N}(\widehat{
-  \bm \Sigma}) \subset \mathcal{N}(\widehat{\bm \Sigma}_k)$, $k = 1, \ldots, K$.
-\end{lemma}
-\begin{proof}
-Let $\bm z \in \mathcal{N}\{\widehat{\bm \Sigma}_k(\lambda)\}$ for some $k = 1,
-\ldots, K$. Hence, $0 = \bm z^{T} \widehat{\bm \Sigma}_k(\lambda) \bm z = (1 -
-\lambda) \bm z^{T} \widehat{\bm \Sigma}_k \bm z + \lambda \bm z^{T} \widehat{
-  \bm \Sigma} \bm z$. Because $\widehat{\bm \Sigma}_k, \widehat{\bm \Sigma}\in
-\mathbb{R}_{p \times p}^{\ge}$, we have $\bm z \in \mathcal{N}(\widehat{
-  \bm \Sigma})$ and $\bm z \in \mathcal{N}(\widehat{\bm \Sigma}_k)$. In particular,
-we have that $\mathcal{N}\{\widehat{\bm \Sigma}_k(\lambda)\} \subset
-\mathcal{N}(\widehat{\bm \Sigma})$. Now, suppose $\bm z \in
-\mathcal{N}(\widehat{\bm \Sigma})$. Similarly, we have that $0 = \bm z^{T}
-\widehat{\bm \Sigma} \bm z = N^{-1} \sum_{k = 1}^K n_k \bm z^{T} \widehat{
-  \bm \Sigma}_k \bm z$, which implies that $\bm z \in \mathcal{N}(\widehat{
-  \bm \Sigma}_k)$ because $\widehat{\bm \Sigma}_k \in \mathbb{R}_{p \times
-  p}^{\ge}$. Therefore, $\mathcal{N}(\widehat{\bm \Sigma}) \subset
-\mathcal{N}(\widehat{\bm \Sigma}_k)$.
-\end{proof}
-
-In Lemma \ref{lemma:rda-tilde-Sigma_k} below, we derive an alternative
-expression for $\tilde{\bm \Sigma}_k$ in terms of the matrix of eigenvectors of
-$\widehat{\bm \Sigma}$. Let $\widehat{\bm \Sigma} = \bm U \bm D \bm U^{T}$ be
-the eigendecomposition of $\widehat{\bm \Sigma}$ such that $\bm D \in
-\mathbb{R}_{p \times p}^{\ge}$ is the diagonal matrix of eigenvalues of
-$\widehat{\bm \Sigma}$ with
-\begin{align*}
-  \bm D = \begin{bmatrix}
-    \bm D_q & \bm 0 \\
-    \bm{0} & \bm 0_{p-q}
-    \end{bmatrix},
-\end{align*}
-$\bm D_q \in \mathbb{R}_{q \times q}^{>}$ is the diagonal matrix consisting of
-the positive eigenvalues of $\widehat{\bm \Sigma}$, the columns of $\bm U \in
-\mathbb{R}_{p \times p}$ are the corresponding orthonormal eigenvectors of
-$\widehat{\bm \Sigma}$, and rank$(\widehat{\bm \Sigma}) = q$. Then, we partition
-$\bm U = (\bm U_1, \bm U_2)$ such that $\bm U_1 \in \mathbb{R}_{p \times q}$ and
-$\bm U_2 \in \mathbb{R}_{p \times (p - q)}$.
-
-\begin{lemma}\label{lemma:rda-tilde-Sigma_k}
-Let $\widehat{\bm \Sigma} = \bm U \bm D \bm U^{T}$ be the eigendecomposition of
-$\widehat{\bm \Sigma}$ as above, and suppose that rank$(\widehat{\bm \Sigma}) =
-q \le p$. Then, we have
-\begin{align}
-	\tilde{\bm \Sigma}_k &= \bm U
-    \begin{bmatrix}
-      \bm W_k & \bm 0 \\
-      \bm 0 & \gamma \bm I_{p-q}
-    \end{bmatrix}
-    \bm U^{T}, \quad k = 1, \ldots, K,\label{eq:rda-matrix}
-\intertext{where}
-\bm W_k &= \alpha_k \{(1 - \lambda) \bm U_1^{T} \widehat{\bm \Sigma}_k \bm U_1 + \lambda \bm D_q\} + \gamma \bm I_{q}.\label{eq:Wk}
-\end{align}
-\end{lemma}
-\begin{proof}
-From Lemma \ref{lemma:null-spaces}, the columns of $\bm U_2$ span the null space
-of $\widehat{\bm \Sigma}_k$, which implies that $\widehat{\bm \Sigma}_k \bm U_2
-= \bm 0_{p \times (p - q)}$. Hence,
-\begin{align*}
-  \bm U^{T} \widehat{\bm \Sigma}_k \bm U = \begin{bmatrix}
-    \bm U_1^{T} \widehat{\bm \Sigma}_k \bm U_1 & \bm 0 \\
-    \bm 0 & \bm 0_{p-q}
-  \end{bmatrix}, \quad k = 1, \ldots, K.
-\end{align*}
-Thus, $\bm U^{T} \tilde{\bm \Sigma}_k \bm U = \alpha_k \{(1 - \lambda) \bm U^{T}
-\widehat{\bm \Sigma}_k \bm U + \lambda \bm D\} + \gamma \bm I_p$, and
-\eqref{eq:rda-matrix} holds because $\bm U$ is orthogonal.
-\end{proof}
-
-As an immediate consequence of Lemma \ref{lemma:rda-tilde-Sigma_k}, we have the
-following corollary.
-
-\begin{cor}
-Let $\widehat{\bm \Sigma}_k(\lambda)$ be defined as in
-\eqref{eq:sig-lambda-alternative}. Then, for $\lambda \in (0, 1]$,
-  rank$\{\widehat{\bm \Sigma}_k(\lambda)\} = q$, $k = 1, \ldots, K$.
-\end{cor}
-\begin{proof}
-The proof follows when we set $\gamma = 0$ in Lemma
-\ref{lemma:rda-tilde-Sigma_k}.
-\end{proof}
-
-Thus, by incorporating each $\bm x_i$ into the estimation of $\bm \Sigma_k$, we
-increase the rank of $\widehat{\bm \Sigma}_k(\lambda)$ to $q \approx N$ if
-$\lambda \ne 0$. Next, we provide an essential result that enables us to prove
-that \eqref{eq:hdrda} is invariant to adjustments to the eigenvalues of $\tilde{
-  \bm \Sigma}_k$ corresponding to the null space of $\widehat{\bm \Sigma}$.
-
-\begin{lemma}\label{lemma:RDA-constant-term}
-Let $\bm U_2$ be defined as above. Then, for all $\bm x \in \mathbb{R}_{p \times
-  1}$, $\bm U_2^{T} (\bm x - \xbar_k) = \bm U_2^{T} (\bm x - \xbar_{k'})$, $1
-\le k, k' \le K$, where $k \ne k'$.
-\end{lemma}
-\begin{proof}
-Let $\bm x \in \mathbb{R}_{p \times 1}$, and suppose that $1 \le k, k' \le K$.
-Recall that $\bm U_2 \in \mathcal{N}(\widehat{\bm \Sigma})$, which implies that
-$\bm U_2^{T} \in \mathcal{C}(\widehat{\bm \Sigma})^{\perp}$ \citep[Lemma
-  1.2.5]{Kollo:2005vp}. Now, because $\bm x_i \in \mathcal{C}(\widehat{ \bm
-  \Sigma})$ $(i = 1, \ldots, N)$, $\bm U_2^{T} \bm x_i = \bm 0_{p-q}$.  Hence,
-$\bm 0_{p-q} = \sum_{i=1}^N \beta_i \bm U_2^{T}\bm x_i = \bm U_2^{T}(\xbar_k -
-\xbar_{k'})$, where $\beta_i = (n_k n_{k'})^{-1} \{ I(y_i = k) n_{k'} - I(y_i =
-k') n_k \}$. Therefore, $\bm U_2^{T} (\bm x - \xbar_k) = \bm U_2^{T} (\bm x -
-\xbar_{k'})$.
-\end{proof}
-
-We now present our main result, where we decompose \eqref{eq:hdrda} and show
-that the term requiring the largest computational costs does not contribute to
-the classification of an unlabeled observation performed using Lemma
-\ref{lemma:RDA-constant-term}. Hence, we reduce \eqref{eq:hdrda} to an
-equivalent, more computationally efficient decision rule.
-
-\begin{thm}
-Let $\tilde{\bm \Sigma}_k$ and $\bm W_k$ be defined as in \eqref{eq:rda-matrix}
-and \eqref{eq:Wk}, respectively, and let $\bm U_1$ be defined as above. Then,
-the decision rule in \eqref{eq:hdrda} is equivalent to
-\begin{align}
-		D_{HDRDA}(\bm x) &= \argmin_k  (\bm x - \xbar_k)^{T} \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm x - \xbar_k) + \log | \bm W_k |. \label{eq:hdrda-decomposed}
-\end{align}
-\end{thm}
-\begin{proof}
-From \eqref{eq:rda-matrix}, we have that
-\begin{align*}
-  \tilde{\bm \Sigma}_k^{+} = \bm U \begin{bmatrix}
-    \bm W_k^{-1} & \bm 0 \\
-    \bm 0 & \gamma^{+} \bm I_{p-q}
-  \end{bmatrix}
-  \bm U^{T}
-\end{align*}
-and $|\tilde{\bm \Sigma}_k| = \gamma^{p-q}| \bm W_k |$, $k = 1, \ldots,
-K$. Therefore, for all $\bm x \in \mathbb{R}_{p \times 1}$, we have that
-\begin{align*}
-	(\bm x - \xbar_k)^{T} \tilde{\bm \Sigma}_k^{+}(\bm x - \xbar_k)  + \log |\tilde{\bm \Sigma}_k| &= (\bm x - \xbar_k)^{T} \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm x - \xbar_k)\\
-	&+ \gamma^{+} (\bm x - \xbar_k)^{T} \bm U_2 \bm U_2^{T} (\bm x - \xbar_k) + \log | \bm W_k |\\
-  &+ (p - q) \log \gamma.
-\end{align*}
-Because $\gamma$ is constant for $k = 1, \ldots, K$, we can omit the $(p - q)
-\log \gamma$ term and particularly avoid the calculation of $\log 0$ for $\gamma
-= 0$. Then, the proof follows from Lemma \ref{lemma:RDA-constant-term} because
-$\bm U_2^{T} (\bm x - \xbar_k)$ is constant for $k = 1, \ldots, K$.
-\end{proof}
-
-Using Theorem 1, we can avoid the time-consuming inverses and determinants of $p
-\times p$ covariance matrices in \eqref{eq:hdrda} and instead calculate these
-same operations on $\bm W_k \in \mathbb{R}_{q \times q}$ in
-\eqref{eq:hdrda-decomposed}. The substantial computational improvements arise
-because our proposed classifier in \eqref{eq:hdrda} is invariant to the term
-$\bm U_2$, thus yielding an equivalent classifier in \eqref{eq:hdrda-decomposed}
-with a substantial reduction in computational complexity. Here, we demonstrate
-that the computational efficiency in calculating the inverse and determinant of
-$\bm W_k$ can be further improved via standard matrix operations when we show
-that the inverses and determinants of $\bm W_k$ can be performed on matrices of
-size $n_k \times n_k$.
-
-\begin{proposition}\label{proposition:hdrda-W_k}
-Let $\bm W_k$ be defined as above. Then, $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$
-and
-\begin{align}
-		\bm W_k^{-1} &= \bm \Gamma_k^{-1} - n_k^{-1} \alpha_k(1 - \lambda) \bm \Gamma_k^{-1} \bm U_1^{T} \bm X_k^{T} \bm Q_k^{-1} \bm X_k \bm U_1 \bm \Gamma_k^{-1},\label{eq:W_k_inv}
-    \intertext{where}
-    \bm Q_k &= \bm I_{n_k} + n_k^{-1} \alpha_k(1 - \lambda) \bm X_k \bm U_1 \bm \Gamma_k^{-1} \bm U_1^{T} \bm X_k^{T}\label{eq:Q_k}
-    \intertext{and}
-    \bm \Gamma_k &= \alpha_k \lambda \bm D_q + \gamma \bm I_q\label{eq:Gamma_k}.
-\end{align}
-\end{proposition}
-\begin{proof}
-First, we write $\bm W_k = n_k^{-1} \alpha_k (1 - \lambda) \bm U_1^{T} \bm
-X_k^{T} \bm X_k \bm U_1 + \bm \Gamma_k$. To calculate $|\bm W_k|$, we apply
-Theorem 18.1.1 from \cite{Harville:2008wja}, which states that $|\bm A + \bm B
-\bm T \bm C| = |\bm A| |\bm T| |\bm T^{-1} + \bm C \bm A^{-1} \bm B|$, where
-$\bm A \in \mathbb{R}_{a \times a}^{>}$, $\bm B \in \mathbb{R}_{a \times b}$,
-$\bm T \in \mathbb{R}_{b \times b}^{>}$, and $\bm C \in \mathbb{R}_{b \times
-  a}$. Thus, setting $\bm A = \bm \Gamma_k$, $\bm B = \alpha_k (1 - \lambda) \bm
-U_1^{T} \bm X_k^{T}$, $\bm T = \bm I_{n_k}$, and $\bm C = \bm X_k \bm U_1$, we
-have $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$. Similarly, \eqref{eq:W_k_inv}
-follows from the well-known Sherman-Woodbury formula \citep[Theorem
-  18.2.8]{Harville:2008wja} because $(\bm A + \bm B \bm T \bm C)^{-1} =
-\bm A^{-1} - \bm A^{-1} \bm B (\bm T^{-1} + \bm C \bm A^{-1} \bm B)^{-1} \bm C
-\bm A^{-1}$.
-\end{proof}
-
-Notice that $\bm \Gamma_k$ is singular when $(\lambda, \gamma) = (0, 0)$
-because $\bm \Gamma_k = \bm 0_q$, in which case we use the formulation in
-\eqref{eq:hdrda-decomposed} instead. Also, notice that if $\alpha_k$ is
-constant across the $K$ classes, then $\bm \Gamma_k$ in \eqref{eq:Gamma_k} is
-independent of $k$. Consequently, $|\bm \Gamma_k|$ is constant across the $K$
-classes and need not be calculated in \eqref{eq:hdrda-decomposed}.
-
-\subsection{Model Selection}
-
-Thus far, we have presented the \emph{HDRDA} classifier and its properties that
-facilitate an efficient calculation of the decision rule. Here, we describe an
-efficient model-selection procedure along with pseudocode in Algorithm
-\ref{alg:hdrda_model_selection} to select the optimal tuning-parameter estimates
-from the Cartesian product of candidate values $\{\lambda_g\}_{g=1}^G \times
-\{\gamma_h\}_{h=1}^H$. We estimate the $V$-fold cross-validation error rate for
-each candidate pair and select $(\widehat{\lambda}, \widehat{\gamma})$, which
-attains the minimum error rate. To calculate the $V$-fold cross-validation, we
-partition the original training data into $V$ mutually exclusive and exhaustive
-folds that have approximately the same number of observations. Then, for $v = 1,
-\ldots, V$, we classify the observations in the $v$th fold by training a
-classifier on the remaining $V - 1$ folds. We calculate the cross-validation
-error as the proportion of misclassified observations across the $V$ folds.
-
-\IncMargin{1em}
-\begin{algorithm}
-  \SetKwInOut{Input}{input}
-  \SetKwInOut{Output}{output}
-
-  \Input{
-    Data matrix $\bm X$\\
-    Parameter grid $\{\lambda_g\}_{g=1}^G \times \{\gamma_h\}_{h=1}^H$
-  }
-  \Output{Optimal Estimates $(\hat{\lambda}, \hat{\gamma})$}
-  \BlankLine
-  \For{$v \leftarrow 1$ \KwTo $V$} {
-    Partition $\bm X$ into $\bm X_{train} \in \mathbb{R}_{N \times p}$ and
-    $\bm X_{test} \in \mathbb{R}_{N_T \times p}$
-
-    \For{$k\leftarrow 1$ \KwTo $K$}{
-      Extract $\bm X_k \in \mathbb{R}_{n_k \times p}$ from $\bm X_{train}$
-
-      Compute sample mean $\xbar_k$ from $\bm X_k$
-
-      Center $\bm X_k \leftarrow \bm X_k - \bm 1_{n_k} \xbar_k^T$
-    }
-
-    $\bm X_c \leftarrow [\bm X_1^T, \ldots, \bm X_K^T]^T$
-
-    Compute the compact SVD $\bm X_c = \bm M_q \bm D_q \bm U_1^T$
-
-    Transform $\bm X_c \leftarrow \bm X_c \bm U_1$
-
-    Transform $\bm X_{test} \leftarrow \bm X_{test} \bm U_1$
-
-    \For{$k \leftarrow 1$ \KwTo $K$} {
-      Extract $\bm X_k \in \mathbb{R}_{n_k \times q}$ from $\bm X_c$
-
-      Recompute sample mean $\xbar_k$ from $\bm X_k$
-    }
-
-    \For{$(\lambda, \gamma) \in \{\lambda_g\}_{g=1}^G \times \{\gamma_h\}_{h=1}^H$} {
-      \For{$k \leftarrow 1$ \KwTo $K$} {
-        Compute $\bm Q_k$ using \eqref{eq:Q_k}
-
-        Compute $\bm \Gamma_k$ using \eqref{eq:Gamma_k}
-
-        Compute $\bm W_k^{-1}$ using \eqref{eq:W_k_inv}
-
-        Compute $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$
-
-        Compute $(\bm x - \xbar_k)^{T} \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm x -
-        \xbar_k) + \log | \bm W_k |$ for each row $\bm x$ of $\bm X_{test}$
-      }
-      Classify test observations $\bm X_{test}$ using \eqref{eq:hdrda-decomposed}
-
-      Compute the number of misclassified test observations $\#\{\text{Error}_v(\lambda, \gamma)\}$
-    }
-   }
-  Compute $\widehat{\text{Error}}(\lambda, \gamma) = N^{-1} \sum_{v=1}^V \#\{\text{Error}_v(\lambda, \gamma)\}$
-
-  Report optimal $(\hat{\lambda}, \hat{\gamma}) \leftarrow \argmin_{(\lambda, \gamma)} \widehat{\text{Error}}(\lambda, \gamma)$
-  \caption{Model selection for the \emph{HDRDA} classifier}
-  \label{alg:hdrda_model_selection}
-\end{algorithm}
-\DecMargin{1em}
-
-A primary contributing factor to the efficiency of Algorithm
-\ref{alg:hdrda_model_selection} is our usage of the compact singular value
-decomposition (SVD). Rather than computing the eigenvalue decomposition of
-$\widehat{\bm \Sigma}$ to obtain $\bm U_1$, we instead obtain $\bm U_1$ by
-computing the eigendecomposition of a much smaller $N \times N$ matrix when $p
-\gg N$ \cite[Chapter~18.3.5]{Hastie:2008dt}. Applying the SVD, we decompose $\bm
-X_c = \bm M \bm \Delta \bm U^{T}$, where $\bm M \in \mathbb{R}_{N \times p}$ is
-orthogonal, $\bm \Delta \in \mathbb{R}_{p \times p}^{\ge}$ is a diagonal matrix
-consisting of the singular values of $\bm X_c$, and $\bm U \in \mathbb{R}_{p
-  \times p}$ is orthogonal. Recalling that $\widehat{\bm \Sigma} = N^{-1} \bm
-X_c^{T} \bm X_c$, we have the eigendecomposition $\widehat{\bm \Sigma} = \bm U
-\bm D \bm U^{T}$, where $\bm U$ is the matrix of eigenvectors of $\widehat{\bm
-  \Sigma}$ and $\bm D = N^{-1} \bm \Delta$ is the diagonal matrix of eigenvalues
-of $\widehat{\bm \Sigma}$. Now, we can obtain $\bm M$ and $\bm D$ efficiently
-from the eigenvalue decomposition of the $N \times N$ matrix $\bm X_c \bm
-X_c^{T} = \bm M \bm D \bm M^{T}$. Next, we compute $\bm U = \bm X_c^{T} \bm M
-\bm D^{+/2}$, where
-\begin{align*}
-  \bm D^{+/2} = \begin{bmatrix}
-    \bm D_q^{-1/2} & \bm 0 \\
-    \bm 0 & \bm 0_{N-q}
-  \end{bmatrix}.
-\end{align*}
-We then determine $q$, the number of numerically nonzero eigenvalues present in
-$\bm D$, by calculating the number of eigenvalues that exceeds some tolerance
-value, say, $1 \times 10^{-6}$. We then extract $\bm U_1$ as the first $q$
-columns of $\bm U$.
-
-As a result of the compact SVD, we need calculate $\bm X_c \bm U_1$ only once
-per cross-validation fold, requiring $O(p q N) \approx O(p N^2)$ calculations.
-Hence, the computational costs of expensive calculations, such as matrix
-inverses and determinants, are greatly reduced because they are performed in the
-$q$-dimensional subspace. Similarly, we reduce the dimension of the test data
-set by calculating $\bm X_{test} \bm U_1$ once per fold. Conveniently, we see
-that the most costly computation involved in $\bm Q_k$ and $\bm W_k^{-1}$ is
-$\bm X_k \bm U_1$, which can be extracted from $\bm X_c \bm U_1$. Thus, after
-the initial calculation of $\bm X_c \bm U_1$ per cross-validation fold, $\bm
-Q_k$ requires $O(n_k q^2)$ operations. Because $\bm Q_k \in \mathbb{R}_{n_k
-  \times n_k}$, both its determinant and inverse require $O(n_k^3)$
-operations. Consequently, $\bm W_k^{-1}$ requires $O(n_k q^2)$ operations. Also,
-the inverse of the diagonal matrix $\bm \Gamma_k^{-1} \in \mathbb{R}_{q \times
-  q}$ requires $O(q)$ operations. Finally, we remark that $| \bm W_k |$ requires
-$O(n_k^3)$ operations.
-
-The expressions given in Proposition \ref{proposition:hdrda-W_k} also expedite
-the selection of $\lambda$ and $\gamma$ via cross-validation because the most
-time-consuming matrix operation involved in computing $\bm W_k^{-1}$ and $|\bm
-W_k|$ is $\bm X_k \bm U_1 \in \mathbb{R}_{n_k \times q}$, which is independent
-of $\lambda$ and $\gamma$. The subsequent operations in calculating $\bm
-W_k^{-1}$ and $|\bm W_k|$ can be simply updated for different pairs of $\lambda$
-and $\gamma$ without repeating the costly computations. Also, rather than calculating
-$(\bm x - \xbar_k)^{T} \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm x - \xbar_k)$
-individually for each row $\bm x$ of $\bm X_{test}$, we can calculate $(\bm
-X_{test} - \xbar_k \bm 1_k')' \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm X_{test} -
-\xbar_k \bm 1_k')$. The diagonal elements of the resulting matrix contain the
-individual quadratic form of each test observation, $\bm x_t$.
-
-\section{Timing Comparisons between RDA and HDRDA}
-\label{sec:timing-comparisons}
-
-In this section, we demonstrate that the computational performance of the model
-selection employed in the \emph{HDRDA} classifier is substantially faster than
-that of the \emph{RDA} classifier on small-sample, high-dimensional data
-sets. The relative difference in runtime between the two classifiers drastically
-increases as $p$ increases. To compare the two classifiers, we generated 25
-observations from each of $K=4$ multivariate normal populations with mean
-vectors $\bm \mu_1 = -3 \cdot \bm 1_p$, $\bm \mu_2 = -\bm 1_p$, $\bm \mu_3 = \bm
-1_p$, and $\bm \mu_4 = 3 \cdot \bm 1_p$. We set the covariance matrix of each
-population to the $p \times p$ identity matrix. For each data set generated, we
-estimated the parameters $\lambda$ and $\gamma$ for both classifiers using a
-grid of 5 equidistant candidate values between 0 and 1, inclusively. We set
-$\alpha_k = 1 - \gamma$, $k = 1, \ldots, K$, in the \emph{HDRDA} classifier. At
-each pair of $\lambda$ and $\gamma$, we computed the 10-fold cross-validation
-error rate \citep{Hastie:2008dt}. Then, we selected the model that minimized the
-10-fold cross-validation error rate.
-
-We compared the runtime of both classifiers by increasing the number of features
-from $p = 500$ to $p = 5000$ in increments of 500. Next, we generated 100 data
-sets for each value of $p$ and computed the training and model selection runtime
-of both classifiers. Our timing comparisons are based on our \emph{HDRDA}
-implementation in the {\tt sparsediscrim} R package and the standard \emph{RDA}
-implementation in the {\tt klaR} R package. All timing comparisons were
-conducted on an Amazon Elastic Compute Cloud (EC2) {\tt c4.4xlarge} instance
-using version 3.3.1 of the open-source statistical software {\tt R}. Our timing
-comparisons can be reproduced with the code available at
-\url{https://github.com/ramhiser/paper-hdrda}.
-
-\subsection{Timing Comparison Results}
-
-In Figure \ref{fig:timing-results}, we plotted the runtime of the model
-selections for both the \emph{HDRDA} and \emph{RDA} classifiers as a function of
-$p$. We observed that the \emph{HDRDA} classifier was substantially faster than
-the \emph{RDA} classifier as $p$ increased. In the left panel of Figure
-\ref{fig:timing-results}, we fit a quadratic regression line to the \emph{RDA}
-runtimes and a simple linear regression model to the \emph{HDRDA} runtimes.  For
-improved understanding, in the right panel we repeated the same scatterplot and
-linear fit with the timings restricted to the observed range of the \emph{HDRDA}
-timings. Figure \ref{fig:timing-results} suggests that the usage of a matrix
-inverse and determinant in the {\tt klaR} R package's discriminant function
-yielded model-selection timings that exceeded linear growth in $p$. Because the
-\emph{HDRDA} classifier removes inverse and determinants, it was computationally
-more efficient than the \emph{RDA} classifier, especially as $p$ increased. In
-fact, when $p = 5000$, the \emph{RDA} classifier required
-\Sexpr{rda_mean_time_p5000} minutes on average to perform model selection, while
-the \emph{HDRDA} classifier selected its optimal model in
-\Sexpr{hdrda_mean_time_p5000} seconds on average. Clearly, the model selection
-employed by the \emph{HDRDA} classifier is substantially faster than that of the
-\emph{RDA} classifier.
-
-We quantified the relative timing comparisons between the two classifiers by
-calculating the ratio of mean timings of the \emph{RDA} classifier to the
-\emph{HDRDA} classifier for each value of $p$. We employed nonparametric
-bootstrapping to estimate the mean ratio along with 95\% confidence
-intervals. In Figure \ref{fig:timing-comparison-bootstrap}, the bootstrap
-sampling distributions for the ratio of mean timings are given. First, we
-observe that the mean relative timings increased as $p$ increased. For smaller
-dimensions, the relative difference in computing was sizable with the average
-ratio of the mean timings equal to \Sexpr{bootstrap_summary_p500$avg_ratio} for
-$p = 500$ and a 95\% confidence interval of
-(\Sexpr{bootstrap_summary_p500$lower_ci},
-\Sexpr{bootstrap_summary_p500$upper_ci}). Furthermore, the ratio of mean
-computing times suggested that the \emph{RDA} classifier is impractical for
-higher dimensions. For instance, when $p = 5000$, the ratio of mean computing
-times increased to \Sexpr{bootstrap_summary_p5000$avg_ratio} with a 95\%
-confidence interval of (\Sexpr{bootstrap_summary_p5000$lower_ci},
-\Sexpr{bootstrap_summary_p5000$upper_ci}).
-
-\[ \left[\text{Insert Figure \ref{fig:timing-results} approximately here }\right] \]
-
-\[ \left[\text{Insert Figure \ref{fig:timing-comparison-bootstrap} approximately here }\right] \]
-
-\section{Classification Study}
-\label{sec:sims}
-
-In this section, we compare our proposed classifier with four classifiers
-recently proposed for small-sample, high-dimensional data along with the
-random-forest classifier from \cite{Breiman:2001fb} using version 3.3.1 of the
-open-source statistical software {\tt R}. Within our study, we included
-penalized linear discriminant analysis from \cite{Witten:2011kc}, implemented in
-the {\tt penalizedLDA} package. We also considered shrunken centroids
-regularized discriminant analysis from \cite{Guo:2007te} in the {\tt rda}
-package. Because the {\tt rda} package does not perform the authors' ``Min-Min''
-rule automatically, we applied this rule within our {\tt R} code. We included
-two modifications of diagonal linear discriminant analysis from
-\cite{Tong:2012hw} and \cite{Pang:2009ik}, where the former employs an improved
-mean estimator and the latter utilizes an improved variance estimator. Both
-classifiers are available in the {\tt sparsediscrim} package. Finally, we
-incorporated the random forest as a benchmark based on the findings of
-\cite{FernandezDelgado:2014ul}, who concluded that the random forest is often
-superior to other classifiers in benchmark studies. We used the implementation
-of the random-forest classifier from the {\tt randomForest} package with 250
-trees and 100 maximum nodes. For each classifer we explicitly set prior
-probabilities as equal, if applicable. All other classifier options were set to
-their default settings. Below, we refer to each classifier by the first author's
-surname. All simulations were conducted on an Amazon EC2 {\tt c4.4xlarge}
-instance. Our analyses can be reproduced via the code available at
-\url{https://github.com/ramhiser/paper-hdrda}.
-
-For the \emph{HDRDA} classifier in \eqref{eq:hdrda-decomposed}, we examined the
-classification performance of two models. For the first \emph{HDRDA} model, we
-set $\alpha_k = 1$, $k = 1, \ldots, K$, so that the covariance-matrix estimator
-\eqref{eq:hdrda-cov} resembled \eqref{eq:ridge-estimator}. We estimated
-$\lambda$ from a grid of 21 equidistant candidate values between 0 and 1,
-inclusively. Similarly, we estimated $\gamma$ from a grid consisting of the
-values $10^{-1}, \ldots, 10^4$, and $10^5$. We selected optimal estimates of
-$\lambda$ and $\gamma$ using $10$-fold cross-validation. For the second model,
-we set $\alpha_k = 1 - \gamma$, $k = 1, \ldots, K$, to resemble
-\citeauthor{Friedman:1989tm}'s parameterization, and we estimated both $\lambda$
-and $\gamma$ from a grid of 21 equidistant candidate values between 0 and 1,
-inclusively.
-
-We did not include the \emph{RDA} classifier in our classification study because
-its training runtime was prohibitively slow on high-dimensional data in our
-preliminary experiments. As shown in Section \ref{sec:timing-comparisons}, the
-runtime of the \emph{RDA} classifier was drastically larger than that of the
-\emph{HDRDA} classifier for a tuning grid of size $25 = 5 \times
-5$. Consequently, a fair comparison between the \emph{RDA} and \emph{HDRDA}
-classifiers would require model selection of $441 = 21 \times 21$ different
-pairs of tuning parameters in the \emph{RDA} classifier. A tuning grid of this
-size yielded excessively slow training runtimes for the \emph{RDA}
-implementation from the {\tt klaR} R package.
-
-\subsection{Simulation Study}
-
-In this section we compare the competing classifiers using the simulation design
-from \cite{Guo:2007te}. This design is widely used within the high-dimensional
-classification literature, including the studies by \cite{Ramey:2013ji} and
-\cite{Witten:2011kc}. First, we consider the block-diagonal covariance matrix
-from \cite{Guo:2007te},
-\begin{align}
-  \bm\Sigma_k = \begin{bmatrix}
-    \bm\Sigma^{(\rho_k)} & \bm 0_{100} & \bm 0_{100} & \cdots & \cdots & \cdots \\
-    \bm 0_{100} & \bm\Sigma^{(-\rho_k)} & \bm 0_{100} & \bm 0_{100} & \cdots & \vdots \\
-    \bm 0_{100} & \bm 0_{100} & \bm\Sigma^{(\rho_k)} & \bm 0_{100} & \cdots & \vdots \\
-    \vdots & \bm 0_{100} & \bm 0_{100} & \bm\Sigma^{(-\rho_k)} & \bm 0_{100} & \vdots \\
-    \vdots & \vdots & \vdots & \bm 0_{100} & \ddots & \vdots \\
-    \cdots & \cdots & \cdots & \cdots & \cdots & \cdots \\
-  \end{bmatrix},\label{eq:block-sigma}
-\end{align}
-where the $(i,j)$th entry of the block matrix $\bm\Sigma^{(\rho_k)} \in
-\mathbb{R}_{100 \times 100}$ is
-\begin{align*}
-\bm\Sigma_{ij}^{(\rho_k)} = \{ \rho_k^{|i - j|} \}_{1 \le i,j \le 100}.
-\end{align*}
-The block-diagonal covariance structure in \eqref{eq:block-sigma} resembles
-gene-expression data: within each block of pathways, genes are correlated, and
-the correlation decays as a function of the distance between any two genes. The
-original design from \cite{Guo:2007te} comprised two $p$-dimensional
-multivariate normal populations with a common block-diagonal covariance matrix.
-
-Although the design is indeed standard, the simulation configuration lacks
-artifacts commonly observed in real data, such as skewness and extreme
-outliers. As a result, we wished to investigate the effect of outliers on the
-high-dimensional classifiers. To accomplish this goal, we generalized the
-block-diagonal simulation configuration by sampling from a $p$-dimensional
-multivariate contaminated normal distribution.  Denoting the PDF of the
-$p$-dimensional multivariate normal distribution by $N_p(\bm x | \bm \mu, \bm
-\Sigma)$, we write the PDF of the $k$th class as
-\begin{align}
-  p(\bm x | \omega_k) = (1 - \epsilon) N_p(\bm x | \bm \mu_k, \bm \Sigma_k) + \epsilon N_p(\bm x | \bm \mu_k, \eta \bm \Sigma_k),\label{eq:contaminated-pdf}
-\end{align}
-where $\epsilon \in [0, 1]$ is the probability that an observation is
-contaminated (i.e., drawn from a distribution with larger variance) and $\eta >
-1$ scales the covariance matrix $\bm \Sigma_k$ to increase the extremity of
-outliers. For $\epsilon = 0$, we have the benchmark block-diagonal simulation
-design from \cite{Guo:2007te}. As $\epsilon$ is increased, the average number of
-outliers is increased. In our simulation, we let $\eta = 100$ and considered the
-values of $\epsilon = 0$, $0.05$, $\ldots$, $0.50$.
-
-We generated $K=3$ populations from \eqref{eq:contaminated-pdf} with $\bm
-\Sigma_k$ given in \eqref{eq:block-sigma} and set the mean vector of class 1 to
-$\bm \mu_1 = \bm 0_p$. Next, comparable to \cite{Guo:2007te}, the first 100
-features of $\bm \mu_2$ were set to 1/2, while the rest were set to 0, i.e.,
-$\bm \mu_2 = (\underbrace{1/2, \ldots, 1/2}_{100}, \underbrace{0, \ldots, 0}_{p
-  - 100})$. For simplicity, we defined $\bm \mu_3 = -\bm \mu_2$. The three
-populations differed in their mean vectors in the first 100 features
-corresponding to the first block, and no difference in the means occurred in the
-remaining blocks.
-
-From each of the $K=3$ populations, we sampled 25 training observations ($n_k =
-25$ for all $k$) and 10,000 test observations. After training each classifier on
-the training data, we classified the test data sets and computed the proportion
-of mislabeled test observations to estimate the classification error rate for
-each classifier. Repeating this process 500 times, we computed the average of
-the error-rate estimates for each classifier. We allowed the number of features
-to vary from $p = 100$ to $p = 500$ in increments of 100 to examine the
-classification accuracy as the feature dimension increased while maintaining a
-small sample size. \cite{Guo:2007te} originally considered $\rho_k = 0.9$ for
-all $k$. Alternatively, to explore the more realistic assumption of unequal
-covariance matrices, we put $\rho_1 = 0.1$, $\rho_2 = 0.5$, and $\rho_3 = 0.9$.
-
-\subsubsection{Simulation Results}
-
-In Figure \ref{fig:sim-results}, we observed each classifier's average
-classification error rates for the values of $\epsilon$ and $p$. Unsurprisingly,
-the average error rate increased for each classifier as the contamination
-probability $\epsilon$ increased regardless of the value of $p$. Sensitivity to
-the presence of outliers was most apparent for the Pang, Tong, and Witten
-classifiers. For smaller dimensions, the random-forest and \emph{HDRDA}
-classifiers tended to outperform the remaining classifiers with the random
-forest performing best. As the feature dimension increased with $p \ge 300$,
-both \emph{HDRDA} classifiers outperformed all other classifiers, suggesting
-that their inherent dimension reduction better captured the classificatory
-information in the small training samples, even in the presence of outliers.
-
-\[ \left[\text{Insert Figure \ref{fig:sim-results} approximately here }\right] \]
-
-The Pang, Tong, and Witten methods yielded practically the same and consistently
-the worst error rates when outliers were present with $\epsilon > 0$, suggesting
-that these classifiers were sensitive to outliers. Notice, for example, that
-when $p=400$, the error rates of the Pang, Tong, and Witten classifiers
-increased dramatically from approximately 19\% when no outliers were present to
-approximately 43\% when $\epsilon=0.05$. The sharp increase in average error
-rates for these three classifiers continued as $\epsilon$ increased. Guo's
-method always outperformed those of Pang, Witten, and Tong, but after outliers
-were introduced, the Guo classifier's average error rate was not competitive
-with the \emph{HDRDA} classifiers or the random-forest classifier.
-
-\[ \left[\text{Insert Figure \ref{fig:sim-results-by-p} approximately here }\right] \]
-
-In Figure \ref{fig:sim-results-by-p}, we again examine the simulation results as
-a function of $p$ for a subset of the values of $\epsilon$. This set of plots
-allows us to investigate the effect of feature dimensionality on classification
-performance. When no outliers were present (i.e., $\epsilon=0$), the
-random-forest classifier was outperformed by all other classifiers. Furthermore,
-the \emph{HDRDA} classifiers were superior in terms of average error rate in
-this setting. As $p$ increased, an elevation in average error rate was expected
-for all classifiers, but the increase was not observed to be substantial.
-
-For $\epsilon > 0$, we observed a different behavior in classification
-performance. First, the Pang, Tong, and Witten methods, along with the
-random-forest method, increased in average error rate as $p$ increased.
-Contrarily, the performance of the \emph{HDRDA} and Guo classifiers was hardly
-affected by $p$. Also, as discussed above, the \emph{HDRDA} classifiers were
-superior to all other classifiers for large values of $p$ with only the
-random-forest classifier outperforming them in smaller feature-dimension cases.
-
-\subsection{Application to Gene Expression Data}
-
-We compared the \emph{HDRDA} classifier to the five competing classifiers on six
-benchmark gene-expression microarray data sets. First, we evaluated the
-classification accuracy of each classifier by randomly partitioning the data set
-under consideration such that $2/3$ of the observations were allocated as
-training data and the remaining $1/3$ of the observations were allocated as a
-test data set. To expedite the computational runtime, we reduced the training
-data to the top 1000 variables by employing the variable-selection method
-proposed by \cite{Dudoit:2002ev}. We then reduced the test data set to the same
-1000 variables. After training each classifier on the training data, we
-classified the test data sets and computed the proportion of mislabeled test
-observations to estimate the classification error rate for each classifier.
-Repeating this process 100 times, we computed the average of the error-rate
-estimates for each classifier. We next provide a concise description of each
-high-dimensional data set examined in our classification study.
-
-\subsubsection{\cite{Chiaretti:2004gq} Data Set}
-
-\cite{Chiaretti:2004gq} measured the gene-expression profiles for 128
-individuals with acute lymphoblastic leukemia (ALL) using Affymetrix human 95Av2
-arrays. Following \cite{Xu:2009fl}, we restricted the data set to $K = 2$
-classes such that $n_1 = 74$ observations were without cytogenetic abnormalities
-and $n_2 = 37$ observations had a detected BCR/ABL gene. The robust multichip
-average normalization method was applied to all 12,625 gene-expression levels.
-
-\subsubsection{\cite{Chowdary:2006kf} Data Set}
-
-\cite{Chowdary:2006kf} investigated 52 matched pairs of tissues from colon and
-breast tumors using Affymetrix U133A arrays and ribonucleic-acid (RNA)
-amplification. Each tissue pair was gathered from the same patient and consisted
-of a snap-frozen tissue and a tissue suspended in an RNAlater preservative.
-Overall, 31 breast-cancer and 21 colon-cancer pairs were gathered, resulting in
-$K = 2$ classes with $n_1 = 62$ and $n_2 = 42$. A purpose of the study was to
-determine whether the disease state could be identified using 22,283
-gene-expression profiles.
-
-\subsubsection{\cite{Nakayama:2007fl} Data Set}
-
-\cite{Nakayama:2007fl} acquired 105 gene-expression samples of 10 types of
-soft-tissue tumors through an oligonucleotide microarray, including 16 samples
-of synovial sarcoma (SS), 19 samples of myxoid/round cell liposarcoma (MLS), 3
-samples of lipoma, 3 samples of well-differentiated liposarcoma (WDLS), 15
-samples of dedifferentiated liposarcoma (DDLS), 15 samples of myxofibrosarcoma
-(MFS), 6 samples of leiomyosarcoma (LMS), 3 samples of malignant nerve sheathe
-tumor (MPNST), 4 samples of fibrosarcoma (FS), and 21 samples of malignant
-fibrous histiocytoma (MFH). \cite{Nakayama:2007fl} determined from their data
-that these 10 types fell into 4 broader groups: (1) SS; (2) MLS; (3) Lipoma,
-WDLS, and part of DDLS; (4) Spindle cell and pleomorophic sarcomas including
-DDLS, MFS, LMS, MPNST, FS, and MFH. Following \cite{Witten:2011kc}, we
-restricted our analysis to the five tumor types having at least 15 observations.
-
-\subsubsection{\cite{Shipp:2002ka} Data Set}
-
-According to \cite{Shipp:2002ka}, approximately 30\%-40\% of adult non-Hodgkin
-lymphomas are diffuse large B-cell lymphomas (DLBCLs). However, only a small
-proportion of DLBCL patients are cured with modern chemotherapeutic regimens.
-Several models have been proposed, such as the International Prognostic Index
-(IPI), to determine a patient's curability. These models rely on clinical
-covariates, such as age, to determine if the patient can be cured, and the
-models are often ineffective. \cite{Shipp:2002ka} have argued that researchers
-need more effective means to determine a patient's curability. The authors
-measured 6,817 gene-expression levels from 58 DLBCL patient samples with
-customized cDNA (lymphochip) microarrays to investigate the curability of
-patients treated with cyclophosphamide, adriamycin, vincristine, and prednisone
-(CHOP)-based chemotherapy. Among the 58 DLBCL patient samples, 32 are from cured
-patients while 26 are from patients with fatal or refractory disease.
-
-\subsubsection{\cite{Singh:2002fh} Data Set}
-
-\cite{Singh:2002fh} have examined 235 radical prostatectomy specimens from
-surgery patients between 1995 and 1997. The authors used oligonucleotide
-microarrays containing probes for approximately 12,600 genes and expressed
-sequence tags. They have reported that 102 of the radical prostatectomy
-specimens are of high quality: 52 prostate tumor samples and 50 non-tumor
-prostate samples.
-
-\subsubsection{\cite{Tian:2003ht} Data Set}
-
-\cite{Tian:2003ht} investigated the purified plasma cells from the bone marrow
-of control patients along with patients with newly diagnosed multiple
-myeloma. Expression profiles for 12,2625 genes were obtained via Affymetrix
-U95Av2 microarrays. The plasma cells were subjected to biochemical and
-immunohistochemical analyses to identify molecular determinants of osteolytic
-lesions. For 36 multiple-myloma patients, focal bone lesions could not be
-detected by magnetic resonance imaging (MRI), whereas MRI was used to detect
-such lesions in 137 patients.
-
-\subsubsection{Classification Results}
-
-Similar to \cite{Witten:2011kc}, we report the average test error rates obtained
-over 100 random training-test partitions in Table \ref{tab:microarray-results}
-along with standard deviations of the test error rates in parentheses. The
-\emph{HDRDA} and Guo classifiers were superior in classification performance for
-the majority of the simulations. The \emph{HDRDA} classifiers yielded the best
-classification accuracy on the Chowdary and Shipp data sets. Although the random
-forest's accuracy slightly exceeded the \emph{HDRDA} classifiers on the Tian
-data set, our proposed classifiers outperformed the other competing classifiers
-considered here. Moreover, the \emph{HDRDA} classifiers yielded comparable
-performance on five of the six data sets.
-
-\[ \left[\text{Insert Table \ref{tab:microarray-results} approximately here }\right] \]
-
-The average error-rate estimates for the Pang, Tong, and Witten classifiers were
-comparable across all six data sets. Furthermore, the average error rates for
-the Pang and Tong classifiers were approximately equal for all data sets except
-for the Chiaretti dataset. This result suggests that the mean and variance
-estimators used in lieu of the MLEs provided little improvement to
-classification accuracies. However, we investigated the Pang classifier's poor
-performance on the Chiaretti data set and determined that its variance estimator
-exhibited numerical instability. The classifier's denominator was approximately
-zero for both classes and led to the poor classification performance.
-
-The random-forest classifier was competitive when applied to the Chowdary and
-Singh data sets and yielded the smallest error rate of the considered
-classifiers on the Tian data set. The fact that the \emph{HDRDA} and Guo
-classifiers typically outperformed the random-forest classifier challenges the
-claim of \cite{FernandezDelgado:2014ul} that random forests are typically
-superior. Further studies should be performed to validate this statement in the
-small-sample, high-dimensional setting.
-
-Finally, the Pang, Tong, and Witten classifiers consistently yielded the largest
-average error rates across the six data sets. Given that the standard deviations
-were relatively large, we hesitate to generalize claims regarding the ranking of
-these three classifiers in terms of the average error rate. However, the
-classifiers' error rates and their variability across multiple random partitions
-of each data set were large enough that we might question their benefit when
-applied to real data.
-
-\section{Discussion}
-\label{sec:discussion}
-
-We have demonstrated that our proposed \emph{HDRDA} classifier is competitive
-with and often superior to random forests as well as the Witten, Pang, Tong, and
-Guo classifiers. In fact, we have shown that the \emph{HDRDA} classifier often
-yields superior classification accuracy when applied to small-sample,
-high-dimensional data sets, confirming the assertions of \cite{Mai:2012bf} and
-\cite{Fan:2012iq} that diagonal classifiers often yield inferior classification
-performance when compared to other classification methods. Furthermore, we have
-demonstrated that \emph{HDRDA} classifiers are more robust to the presence of
-outliers than the diagonal classifiers despite their rapid computational
-performance and their reduction in the number of parameters to estimate.
-
-We also considered the popular penalized linear discriminant analysis from
-\cite{Witten:2011kc} because it was specifically designed for high-dimensional
-gene-expression data. We had expected its classification performance to be
-competitive within our classification study and perhaps superior. Contrarily,
-our empirical studies suggest that the classifier is sensitive to outliers and
-unable to achieve comparable results with other classifiers designed for
-small-sample, high-dimensional data. Also, despite the claims of
-\cite{FernandezDelgado:2014ul} that random forests are typically superior to
-other classifiers, we observed that they were indeed competitive but were
-typically outperformed by classifiers developed for small-sample,
-high-dimensional data.
-
-We demonstrated that our \emph{HDRDA} implementation in the {\tt sparsediscrim}
-R package can be used in practice with high-dimensional data sets. In our timing
-comparisons, we showed that \emph{HDRDA} model selection could be employed on
-data sets with $p = 5000$ in \Sexpr{hdrda_mean_time_p5000} seconds on
-average. Contrarily, the \emph{RDA} classifier implemented in the {\tt klaR} R
-package required \Sexpr{rda_mean_time_p5000} minutes on average to perform model
-selection on data sets with $p = 5000$. Given that the \emph{RDA} classifier has
-been shown to have excellent performance in the high-dimensional setting
-\citep{Webb:2011vu} but is limited by its computationally intense
-model-selection procedure, our work replaces the \emph{RDA} classifier for
-high-dimensional data in practice. This result is reassuring because the
-\emph{RDA} classifier remains widely popular in the literature. In fact,
-variants of the RDA classifier have been applied to microarray data
-\citep{Ching:2012fu,Li:2012ev,Tai:2007bk,Guo:2007te}, facial recognition
-\citep{Zhang:2010va,Dai:2007vd,Lu:2005hq,Pima:2004vw,Lu:2003we}, handwritten
-digit recognition \citep{Bouveyron:2007gx}, remote sensing
-\citep{Tadjudin:1999fk}, seismic detection \citep{Anderson:2002kg}, and chemical
-spectra \citep{Wu:1996us,Aeberhard:1993fp}.
-
-The dimension reduction employed in this paper has reduced the dimension to
-rank$(\widehat{\bm \Sigma}) = q$. An interesting extension of our work would
-reduce the dimension $q$ further to a lower dimension $q_L < q$, perhaps using a
-criterion similar to that of principal components analysis. While unclear
-whether the classification performance would improve via such a method, the
-efficiency of the model selection would certainly improve. Moreover, if $q_L =
-2$ or $3$, low-dimensional graphical displays of high-dimensional data could be
-obtained.
-
-We thank Mrs. Joy Young for her numerous recommendations that enhanced the
-quality of our writing.
-
-\bibliographystyle{elsarticle-harv}
-\bibliography{rda}
-
-\clearpage
-
-\begin{figure}
-<<hdrda_contours, fig.width=8, fig.height=8, out.width='.33\\linewidth', dpi=200, fig.show='hold'>>=
-rda_contours(lambda = 0)
-rda_contours(lambda = 0.2)
-rda_contours(lambda = 0.4)
-rda_contours(lambda = 0.6)
-rda_contours(lambda = 0.8)
-rda_contours(lambda = 1)
-@
-\caption{Contours of five multivariate normal populations as a function of the
-  pooling parameter $\lambda$.}
-\label{fig:hdrda-contours}
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-<<timing_results, fig.width=8, fig.height=8, dpi=200>>=
-p <- ggplot(timings_combined, aes(x=p, y=time, color=Classifier))
-p <- p + geom_point() + geom_jitter(width=140, height=0) + facet_wrap(~ Plot, scales="free_y")
-p <- p + geom_smooth(method="lm", formula=y ~ poly(x, 2), se=FALSE)
-p <- p + scale_color_manual(values=c("#D55E00", "#0072B2"))
-p <- p + xlab("Number of Features") + ylab("Model Selection Time (seconds)")
-p + theme_bw()
-@
-\caption{Timing comparisons (in seconds) between HDRDA and RDA classifiers.}
-\label{fig:timing-results}
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-<<relative_timings, fig.width=8, fig.height=8, dpi=200>>=
-p <- ggplot(timing_ratios_bootstrap, aes(ratio_means, fill=p))
-p <- p + geom_histogram() + facet_wrap(~ p, scale="free_x")
-p <- p + xlab("Ratio of Mean RDA Runtime to Mean HDRDA Runtime") + ylab("Bootstrap Replications")
-p + theme_bw()
-@
-\caption{Distribution of ratios of mean \emph{RDA} runtime to mean \emph{HDRDA}
-  runtime across 1000 bootstrap replications.}
-\label{fig:timing-comparison-bootstrap}
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-<<sim_results, fig.width=8, fig.height=8, dpi=200>>=
-p <- ggplot(results_contaminated, aes(x=contamination_prob, y=Error))
-p <- p + geom_line(aes(color=Classifier, linetype=Classifier), size=0.65)
-p <- p + facet_wrap(~ p, labeller = "label_both")
-p <- p + xlab("Contamination Probability") + ylab("Average Error Rate")
-p + theme_bw()
-@
-\caption{Average classification error rates as a function of the contamination
-  probability $\epsilon$. Approximate standard errors were no greater than
-  \Sexpr{max_sim_se}.}
-\label{fig:sim-results}
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-<<sim_results_by_p, fig.width=8, fig.height=8, dpi=200>>=
-p2 <- results_contaminated %>%
-  mutate(p=as.numeric(as.character(p)),
-         contamination_prob=round(contamination_prob, digits=2)) %>%
-    filter(contamination_prob %in% c(0.0, 0.1, 0.2, 0.3, 0.4, 0.5)) %>%
-    rename(`Contamination Probability`=contamination_prob) %>%
-      ggplot(., aes(x=p, y=Error))
-p2 <- p2 + geom_line(aes(color=Classifier, linetype=Classifier), size=0.65)
-p2 <- p2 + facet_wrap(~ `Contamination Probability`, labeller="label_both")
-p2 <- p2 + xlab("Number of Features") + ylab("Average Error Rate")
-p2 + theme_bw()
-@
-\caption{Average classification error rates as a function of the number of
-  features $p$. Approximate standard errors were no greater than
-  \Sexpr{max_sim_se}.}
-\label{fig:sim-results-by-p}
-\end{figure}
-
-\clearpage
-
-<<classification_results, results='asis'>>=
-table_caption <- 'The average of the test error rates obtained on gene-expression data sets over 100 random training-test partitions.
-Standard deviations of the test error rates are given in the parentheses.
-The classifier with the minimum average error rate for each data set is in bold.'
-
-xtable_out <- xtable(results_microarray,
-                     caption=table_caption,
-                     label='tab:microarray-results')
-print(xtable_out, include.rownames=FALSE, sanitize.text.function = function(x) { x })
-@
-
 \end{document}

### 2017-02-12 reports/author-biography.tex
Okay scratch that. PR upload site says Word docs are corrupt

\documentclass{elsarticle}

%----------------Page dimensions ----------------
  \oddsidemargin 0.0in
\evensidemargin 0.0in
\topmargin -0.75in
\leftmargin 0in
\headheight 0.0in
\headsep 0.5in
\footskip 0.5in
\footnotesep 0.0in
\textwidth 6.5in
\textheight 9.5in

\begin{document}

\title{Author Biography}

\maketitle

\begin{itemize}
  \item John A. Ramey is Principal Data Scientist at Novi Labs, an Austin-based startup focused on optimizing oil and gas drilling and completion designs. He graduated in 2012 with a Ph.D. in Statistics from Baylor University and completed a postdoc at Fred Hutchinson Cancer Research Center.
  \item Caleb K. Stein earned BA from the University of Arkansas, earned an MS in Statistics and currently enrolled in PhD program at Baylor University. Stein's publications include investigations of gene expression and mutation calls from Multiple Myeloma patient data, batch effect correction, and clinical outcome analyses.
  \item Phil D. Young is Clinical Assistant Professor in the Department of Information Systems at Baylor University. He graduated with a Ph.D. in Statistics from Baylor University in 2009. His fields of interest include statistical pattern recognition and classification.
  \item Dean M. Young is Professor of Statistics at Baylor University. His research interests include pattern recognition, multivariate analysis, and computational statistics.
\end{itemize}

\end{document}

### 2017-02-12 reports/highlights.tex
Added paper's highlights for Pattern Recognition

--- 2017-02-12 reports/author-biography.tex
@@
 
-\title{Author Biography}
+\title{Highlights}
 
@@
 \begin{itemize}
-  \item John A.
-Ramey is Principal Data Scientist at Novi Labs, an Austin-based startup focused on optimizing oil and gas drilling and completion designs.
-He graduated in 2012 with a Ph.D.
-in Statistics from Baylor University and completed a postdoc at Fred Hutchinson Cancer Research Center.
-  \item Caleb K.
-Stein earned BA from the University of Arkansas, earned an MS in Statistics and currently enrolled in PhD program at Baylor University.
-Stein's publications include investigations of gene expression and mutation calls from Multiple Myeloma patient data, batch effect correction, and clinical outcome analyses.
-  \item Phil D.
-Young is Clinical Assistant Professor in the Department of Information Systems at Baylor University.
-He graduated with a Ph.D.
-in Statistics from Baylor University in 2009.
-His fields of interest include statistical pattern recognition and classification.
-  \item Dean M. Young is Professor of Statistics at Baylor University. His research interests include pattern recognition, multivariate analysis, and computational statistics.
+  \item We present an efficient classifier designed for small-sample, high-dimensional data.
+  \item Interpretable alternative to Friedman's Regularized Discriminant Analysis.
+  \item HDRDA's matrix operations are linear in the number of features.
+  \item Timing comparisons show our classifier is substantially faster than Friedman's RDA.
+  \item Often superior to sparse and regularized classifiers in classification accuracy.
 \end{itemize}

### 2017-10-15 reports/paper-hdrda.Rnw
use Taylor and Francis latex template #29

@@
-\documentclass[review,authoryear]{elsarticle}
-
-\usepackage{graphicx, amsmath, amssymb, bm, mathtools, amsthm, natbib}
+\documentclass[]{interact}
+
+\usepackage{graphicx, mathtools, amssymb, bm, amsthm}
 \usepackage{algorithm2e, url}
 
-\bibpunct{(}{)}{;}{a}{,}{,}
+% Journal Specific
+
+% natbib.sty for citation and bibliography support
+\usepackage[numbers,sort&compress]{natbib}
+\bibpunct[, ]{[}{]}{,}{n}{,}{,}
+\renewcommand\bibfont{\fontsize{10}{12}\selectfont}
+\makeatletter% @ becomes a letter
+\def\NAT@def@citea{\def\@citea{\NAT@separator}}% Suppress spaces between citations using natbib.sty
+\makeatother% @ becomes a symbol again
 
@@
 
-%----------------Page dimensions ----------------
-\oddsidemargin 0.0in
-\evensidemargin 0.0in
-\topmargin -0.75in
-\leftmargin 0in
-\headheight 0.0in
-\headsep 0.5in
-\footskip 0.5in
-\footnotesep 0.0in
-\textwidth 6.5in
-\textheight 9.5in
-
 \begin{document}
 
-\begin{frontmatter}
+%\articletype{ARTICLE TEMPLATE} % Specify the article type or omit as appropriate
 
@@
 
-\author[jar]{John A. Ramey\cortext[cor1]{Corresponding author}}
-\author[cks]{Caleb K. Stein}
-\author[pdy]{Phil D. Young}
-\author[dmy]{Dean M. Young}
-\address[jar]{Novi Labs}
-\address[cks]{Myeloma Institute, University of Arkansas for Medical Sciences}
-\address[pdy]{Department of Management and Information Systems, Baylor University}
-\address[dmy]{Department of Statistical Science, Baylor University}
+% \author{
+% \name{John.~A.
+Ramey\textsuperscript{a}\thanks{CONTACT John.~A.
+Ramey.
+Email: johnramey@gmail.com}, Caleb~K.
+Stein\textsuperscript{b}, Phil~D.
+Young\textsuperscript{c}, and Dean~M.
+Young\textsuperscript{d}}
+% \affil{\textsuperscript{a}Novi Labs, Austin, TX, USA; \textsuperscript{b}Myeloma Institute, University of Arkansas for Medical Sciences, Little Rock, AR, USA; \textsuperscript{c}Department of Management and Information Systems, Baylor University, Waco, TX, USA; \textsuperscript{d}Department of Statistical Science, Baylor University, Waco, TX, USA;}
+%}
+
+\author{
+\name{A.~N. Author\textsuperscript{a}\thanks{CONTACT A.~N. Author. Email: latex.helpdesk@tandf.co.uk} and John Smith\textsuperscript{b}}
+\affil{\textsuperscript{a}Taylor \& Francis, 4 Park Square, Milton Park, Abingdon, UK; \textsuperscript{b}Institut f\"{u}r Informatik, Albert-Ludwigs-Universit\"{a}t, Freiburg, Germany}
+}
 
@@
 
-\begin{keyword}
-Regularized discriminant analysis\sep High-dimensional classification\sep Covariance-matrix regularization\sep Singular value decomposition\sep Multivariate analysis\sep Dimension reduction
-\MSC[2010] 62H30\sep 65F15\sep 65F20\sep 65F22
-\end{keyword}
-
-\end{frontmatter}
+% TODO: Are these codes needed?
+% \MSC[2010] 62H30\sep 65F15\sep 65F20\sep 65F22
+\begin{keywords}
+Regularized discriminant analysis; High-dimensional classification; Covariance-matrix regularization; Singular value decomposition; Multivariate analysis; Dimension reduction
+\end{keywords}
 

### 2017-10-17 reports/paper-hdrda.Rnw
changes for T&F template

@@
 \usepackage{algorithm2e, url}
+\usepackage{setspace}
 
@@
 \makeatletter% @ becomes a letter
-\def\NAT@def@citea{\def\@citea{\NAT@separator}}% Suppress spaces between citations using natbib.sty
-\makeatother% @ becomes a symbol again
+\def\NAT@def@citea{\def\@citea{\NAT@separator}} % Suppress spaces between citations using natbib.sty
+\makeatother % @ becomes a symbol again
 
@@
 
-%\articletype{ARTICLE TEMPLATE} % Specify the article type or omit as appropriate
+% Specify the article type or omit as appropriate
+\articletype{Submission}
 
@@
 
-% \author{
-% \name{John.~A.
-Ramey\textsuperscript{a}\thanks{CONTACT John.~A.
+\author{
+\name{John A.
+Ramey\textsuperscript{a}\thanks{CONTACT John A.
 Ramey.
-Email: johnramey@gmail.com}, Caleb~K.
-Stein\textsuperscript{b}, Phil~D.
-Young\textsuperscript{c}, and Dean~M.
+Email: johnramey@gmail.com}, Caleb K.
+Stein\textsuperscript{b}, Phil D.
+Young\textsuperscript{c}, and Dean M.
 Young\textsuperscript{d}}
-% \affil{\textsuperscript{a}Novi Labs, Austin, TX, USA; \textsuperscript{b}Myeloma Institute, University of Arkansas for Medical Sciences, Little Rock, AR, USA; \textsuperscript{c}Department of Management and Information Systems, Baylor University, Waco, TX, USA; \textsuperscript{d}Department of Statistical Science, Baylor University, Waco, TX, USA;}
-%}
-
-\author{
-\name{A.~N. Author\textsuperscript{a}\thanks{CONTACT A.~N. Author. Email: latex.helpdesk@tandf.co.uk} and John Smith\textsuperscript{b}}
-\affil{\textsuperscript{a}Taylor \& Francis, 4 Park Square, Milton Park, Abingdon, UK; \textsuperscript{b}Institut f\"{u}r Informatik, Albert-Ludwigs-Universit\"{a}t, Freiburg, Germany}
+\affil{\textsuperscript{a}Novi Labs, Austin, TX, USA; \textsuperscript{b}Myeloma Institute, University of Arkansas for Medical Sciences, Little Rock, AR, USA; \textsuperscript{c}Department of Management and Information Systems, Baylor University, Waco, TX, USA; \textsuperscript{d}Department of Statistical Science, Baylor University, Waco, TX, USA}
 }
+
+\maketitle
+
+\doublespacing
 
@@
 
-% TODO: Are these codes needed?
-% \MSC[2010] 62H30\sep 65F15\sep 65F20\sep 65F22
+% TODO:
+% 1. Word count
+% 2. Update plots to TIFF format with correct parameters.
+% 3. Are MSC codes needed?: \MSC[2010] 62H30\sep 65F15\sep 65F20\sep 65F22
+
 \begin{keywords}
@@
 \cite{Rao:1971ul}, and other methods studied by \cite{Ramey:2013ji} and
-\citet*{Xu:2009fl}.
+\cite{Xu:2009fl}.
 
@@
 
-\subsubsection{\cite{Chiaretti:2004gq} Data Set}
+\subsubsection{Chiaretti Data Set}
 
@@
 
-\subsubsection{\cite{Chowdary:2006kf} Data Set}
+\subsubsection{Chowdary Data Set}
 
@@
 
-\subsubsection{\cite{Nakayama:2007fl} Data Set}
+\subsubsection{Nakayama Data Set}
 
@@
 
-\subsubsection{\cite{Shipp:2002ka} Data Set}
+\subsubsection{Shipp Data Set}
 
@@
 
-\subsubsection{\cite{Singh:2002fh} Data Set}
+\subsubsection{Singh Data Set}
 
@@
 
-\subsubsection{\cite{Tian:2003ht} Data Set}
+\subsubsection{Tian Data Set}
 
@@
 
+<<classification_results, results='asis'>>=
+table_caption <- 'The average of the test error rates obtained on gene-expression data sets over 100 random training-test partitions.
+Standard deviations of the test error rates are given in the parentheses.
+The classifier with the minimum average error rate for each data set is in bold.'
+
+xtable_out <- xtable(results_microarray,
+                     caption=table_caption,
+                     label='tab:microarray-results')
+print(xtable_out, include.rownames=FALSE, sanitize.text.function = function(x) { x },
+      scalebox='0.85') # scales the table to fit the page
+@
+
+\clearpage
+
 \begin{figure}
@@
 
-\clearpage
-
-<<classification_results, results='asis'>>=
-table_caption <- 'The average of the test error rates obtained on gene-expression data sets over 100 random training-test partitions.
-Standard deviations of the test error rates are given in the parentheses.
-The classifier with the minimum average error rate for each data set is in bold.'
-
-xtable_out <- xtable(results_microarray,
-                     caption=table_caption,
-                     label='tab:microarray-results')
-print(xtable_out, include.rownames=FALSE, sanitize.text.function = function(x) { x })
-@
-
 \end{document}

### 2017-10-17 reports/paper-hdrda.Rnw
added word count for journal
Added TIFF format because JSCS. :(

@@
+% Word Count for journal: 7,059
+% detex paper-hdrda.tex | wc -w
+
 \documentclass[]{interact}
@@
 
-% TODO:
-% 1. Word count
-% 2. Update plots to TIFF format with correct parameters.
-% 3. Are MSC codes needed?: \MSC[2010] 62H30\sep 65F15\sep 65F20\sep 65F22
+% TODO: Are MSC codes needed?: \MSC[2010] 62H30\sep 65F15\sep 65F20\sep 65F22
 
@@
 
-opts_chunk$set(fig.align='default', dev='png', message=FALSE, warning=FALSE,
-               cache=TRUE, echo=FALSE, autodep=TRUE, fig.width=12,
-               fig.height=12, out.width="\\linewidth")
+opts_chunk$set(fig.align='default', dev=c("tiff", "png"),
+               message=FALSE, warning=FALSE, cache=TRUE, echo=FALSE, autodep=TRUE,
+               dpi=300, fig.width=12, fig.height=12, out.width="\\linewidth")
 @
@@
 \begin{figure}
-<<hdrda_contours, fig.width=8, fig.height=8, out.width='.33\\linewidth', dpi=200, fig.show='hold'>>=
+<<hdrda_contours, fig.width=8, fig.height=8, out.width='.33\\linewidth', fig.show='hold'>>=
 rda_contours(lambda = 0)
@@
 \begin{figure}
-<<timing_results, fig.width=8, fig.height=8, dpi=200>>=
+<<timing_results, fig.width=8, fig.height=8>>=
 p <- ggplot(timings_combined, aes(x=p, y=time, color=Classifier))
@@
 \begin{figure}
-<<relative_timings, fig.width=8, fig.height=8, dpi=200>>=
+<<relative_timings, fig.width=8, fig.height=8>>=
 p <- ggplot(timing_ratios_bootstrap, aes(ratio_means, fill=p))
-p <- p + geom_histogram() + facet_wrap(~ p, scale="free_x")
+p <- p + geom_histogram() + facet_wrap(~ p, scale="free_x", labeller="label_both")
 p <- p + xlab("Ratio of Mean RDA Runtime to Mean HDRDA Runtime") + ylab("Bootstrap Replications")
@@
 \begin{figure}
-<<sim_results, fig.width=8, fig.height=8, dpi=200>>=
+<<sim_results, fig.width=8, fig.height=8>>=
 p <- ggplot(results_contaminated, aes(x=contamination_prob, y=Error))
@@
 \begin{figure}
-<<sim_results_by_p, fig.width=8, fig.height=8, dpi=200>>=
+<<sim_results_by_p, fig.width=8, fig.height=8>>=
 p2 <- results_contaminated %>%

### 2017-10-17 reports/paper-hdrda.Rnw
scratch that. postscript is smaller and converts to PDF automatically

@@
 
-opts_chunk$set(fig.align='default', dev=c("tiff", "png"),
+opts_chunk$set(fig.align='default', dev="postscript",
                message=FALSE, warning=FALSE, cache=TRUE, echo=FALSE, autodep=TRUE,

### 2017-10-19 reports/paper-hdrda.Rnw
removed MSC codes
fixed typo in paper

@@
 size. Consequently, the matrix operations employed by \emph{HDRDA} are linear
-in the number of features. Timing comparisons between \emph{HDRDA} and RDA
+in the number of features. Timing comparisons between \emph{HDRDA} and \emph{RDA}
 establish that as the number of features increases, the computational runtime
@@
 \end{abstract}
-
-% TODO: Are MSC codes needed?: \MSC[2010] 62H30\sep 65F15\sep 65F20\sep 65F22
 

### 2017-10-19 reports/title-page.tex
updated title page

@@
-\documentclass[review,authoryear]{elsarticle}
+\documentclass[]{interact}
 \usepackage{url}
-
-%----------------Page dimensions ----------------
-  \oddsidemargin 0.0in
-\evensidemargin 0.0in
-\topmargin -0.75in
-\leftmargin 0in
-\headheight 0.0in
-\headsep 0.5in
-\footskip 0.5in
-\footnotesep 0.0in
-\textwidth 6.5in
-\textheight 9.5in
 
@@
 
-\begin{frontmatter}
-
+\articletype{Title Page}
 \title{High-Dimensional Regularized Discriminant Analysis}
 
-\author[jar]{John A. Ramey}
-\author[cks]{Caleb K. Stein}
-\author[pdy]{Phil D. Young}
-\author[dmy]{Dean M. Young}
-\address[jar]{\textbf{Corresponding Author} \\ Novi Labs \\ 7020 Easy Wind Drive, Suite 210 \\ Austin, TX 78752 \\ USA \\ Email: johnramey@gmail.com}
-\address[cks]{Myeloma Institute \\ University of Arkansas for Medical Sciences \\ 4018 W. Capitol Avenue \\ Little Rock, AR 72205 \\ USA \\ Email: stein.caleb@gmail.com}
-\address[pdy]{Department of Management and Information Systems \\ Baylor University \\ One Bear Place \#98005 \\ Waco, TX 76798-8005 \\ USA \\ Email: Philip\_Young@baylor.edu}
-\address[dmy]{Department of Statistical Science \\ Baylor University \\ P.O. Box 97140 \\ Waco, TX 76798-7140 \\ USA \\ Email: Dean\_Young@baylor.edu}
+\author{
+\name{John A.
+Ramey\textsuperscript{a}\thanks{CONTACT John A.
+Ramey.
+Email: johnramey@gmail.com}, Caleb K.
+Stein\textsuperscript{b}, Phil D.
+Young\textsuperscript{c}, and Dean M.
+Young\textsuperscript{d}}
+\affil{\textsuperscript{a}Novi Labs, Austin, TX, USA;\\ \textsuperscript{b}Myeloma Institute, University of Arkansas for Medical Sciences, Little Rock, AR, USA;\\ \textsuperscript{c}Department of Management and Information Systems, Baylor University, Waco, TX, USA;\\ \textsuperscript{d}Department of Statistical Science, Baylor University, Waco, TX, USA}
+}
 
-\end{frontmatter}
+\maketitle
 

### 2017-10-19 reports/paper-hdrda-anonymous.Rnw
anonymous version of paper because journal does double blind

--- 2017-10-19 reports/paper-hdrda.Rnw
@@
 
-\author{
-\name{John A.
-Ramey\textsuperscript{a}\thanks{CONTACT John A.
-Ramey.
-Email: johnramey@gmail.com}, Caleb K.
-Stein\textsuperscript{b}, Phil D.
-Young\textsuperscript{c}, and Dean M.
-Young\textsuperscript{d}}
-\affil{\textsuperscript{a}Novi Labs, Austin, TX, USA; \textsuperscript{b}Myeloma Institute, University of Arkansas for Medical Sciences, Little Rock, AR, USA; \textsuperscript{c}Department of Management and Information Systems, Baylor University, Waco, TX, USA; \textsuperscript{d}Department of Statistical Science, Baylor University, Waco, TX, USA}
-}
-
 \maketitle
@@
 \doublespacing
+
+\textbf{Word Count:} 7,059
 
@@
 
-\section{Acknowledgements}
-
-This research did not receive any specific grant from funding agencies in the
-public, commercial, or not-for-profit sectors. We thank Mrs. Joy Young for her
-numerous recommendations that enhanced the quality of our writing.
-
 \bibliographystyle{elsarticle-harv}

### 2017-10-19 reports/title-page.tex
anonymous version of paper because journal does double blind

@@
 Young\textsuperscript{d}}
-\affil{\textsuperscript{a}Novi Labs, Austin, TX, USA;\\ \textsuperscript{b}Myeloma Institute, University of Arkansas for Medical Sciences, Little Rock, AR, USA;\\ \textsuperscript{c}Department of Management and Information Systems, Baylor University, Waco, TX, USA;\\ \textsuperscript{d}Department of Statistical Science, Baylor University, Waco, TX, USA}
+\affil{\textsuperscript{a}Novi Labs, Austin, TX, USA, \url{johnramey@gmail.com};\\ \textsuperscript{b}Myeloma Institute, University of Arkansas for Medical Sciences, Little Rock, AR, USA, \url{stein.caleb@gmail.com};\\ \textsuperscript{c}Department of Management and Information Systems, Baylor University, Waco, TX, USA, \url{Philip_Young@baylor.edu};\\ \textsuperscript{d}Department of Statistical Science, Baylor University, Waco, TX, USA, \url{Dean_Young@baylor.edu}}
 }
@@
 
+\section*{Acknowledgements}
+
+This research did not receive any specific grant from funding agencies in the
+public, commercial, or not-for-profit sectors. We thank Mrs. Joy Young for her
+numerous recommendations that enhanced the quality of our writing.
+
 \end{document}

### 2017-10-19 reports/paper-hdrda-anonymous.tex
added .tex for anonymous submission.
removed a bunch of latex commands generated by knitr -- was causing build
issues on the journal's site.

--- 2017-10-19 reports/paper-hdrda-anonymous.Rnw
@@
 
-\documentclass[]{interact}
+\documentclass[]{interact}\usepackage[]{graphicx}\usepackage[]{color}
+%% maxwidth is the original width if it is less than linewidth
+%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+\let\hlipl\hlkwb
+
+\usepackage{alltt}
 
@@
 \newtheorem{thm}{Theorem}
-
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
@@
 \end{keywords}
-
-<<knitr_setup, include=FALSE, echo=FALSE, cache=FALSE>>=
-library(knitr)
-# Round digits throughout document including \Sexpr statements.
-# See Yihui's response here:
-# http://stackoverflow.com/questions/11062497/how-to-avoid-using-round-in-every-sexpr
-options(digits=3)
-
-opts_chunk$set(fig.align='default', dev="postscript",
-               message=FALSE, warning=FALSE, cache=TRUE, echo=FALSE, autodep=TRUE,
-               dpi=300, fig.width=12, fig.height=12, out.width="\\linewidth")
-@
-
-<<setup>>=
-setwd("..")
-library(ProjectTemplate)
-load.project()
-@
-
-<<munge_contaminated_sims>>=
-# Contaminated Normal Sims
-results_contaminated <- lapply(results_block_contaminated[[1]], function(r) {
-  with(r, c(error_rates, p=p, mu=mu, contamination_prob=contamination_prob))
-})
-results_contaminated <- do.call(rbind.data.frame, results_contaminated)
-results_contaminated <- tbl_df(results_contaminated)
-
-results_contaminated <- results_contaminated %>%
-  group_by(p, contamination_prob, mu) %>%
-  summarise_each(funs(mean)) %>%
-  arrange(p, contamination_prob, mu) %>%
-  ungroup
-
-results_contaminated <- results_contaminated %>%
-  gather(Classifier, Error, -p, -contamination_prob, -mu) %>%
-  mutate(p=factor(p)) %>%
-  filter(mu == 0.5)
-
-# Remove underscore from classifier names
-levels(results_contaminated$Classifier) <- gsub("_", " ",
-                                                levels(results_contaminated$Classifier))
-
-# Maximum Standard Error for Contaminated Sims
-max_sim_se <- with(results_contaminated, max(sqrt(Error * (1 - Error) / 500)))
-max_sim_se <- round(max_sim_se, digits=3)
-@
-
-<<munge_microarray_results>>=
-# Microarray Classification Results
-results_microarray <- lapply(results[[1]], function(r) {
-  with(r, c(error_rates, `data_set`=data_set))
-})
-results_microarray <- do.call(rbind.data.frame, results_microarray)
-results_microarray <- tbl_df(results_microarray)
-
-avg_error_microarray <- results_microarray %>%
-    group_by(data_set) %>%
-    summarise_each(funs(mean)) %>%
-    gather(Classifier, Avg_Error, -data_set)
-
-sd_error_microarray <- results_microarray %>%
-  group_by(data_set) %>%
-  summarise_each(funs(sd)) %>%
-  gather(Classifier, SD_Error, -data_set)
-
-# Microarray classification results are presented in a table
-# Rows = classifier, columns = data set
-# Cells = Error (SD) rounded to 3 digits
-# The classifier with the minimum average error rate is in bold for each data set.
-
-bold_latex <- function(x) {
-  paste0('\\textbf{', x, '}')
-}
-
-results_microarray <- inner_join(avg_error_microarray, sd_error_microarray,
-                                 by=c('data_set', 'Classifier')) %>%
-  mutate(Table_Cells=sprintf('%.3f', Avg_Error),
-         data_set=stringr::str_to_title(data_set),
-         Classifier=gsub("_", " ", Classifier)) %>%
-  group_by(data_set) %>%
-  mutate(Table_Cells=ifelse(Avg_Error == min(Avg_Error), bold_latex(Table_Cells), Table_Cells),
-         Table_Cells=sprintf('%s (%.3f)', Table_Cells, SD_Error)) %>%
-  select(-Avg_Error, -SD_Error) %>%
-  spread(data_set, Table_Cells)
-@
-
-<<munge_timing_comparisons>>=
-# HDRDA vs RDA Timing Comparison Results
-
-# Benchmarks are given in nanoseconds, which are converted to seconds
-timings <- bind_rows(lapply(timing_results, as.data.frame), .id="p") %>%
-  mutate(p=as.integer(p),
-         expr=as.character(expr),
-         Classifier=ifelse(startsWith(expr, "hdrda"), "HDRDA", "RDA"),
-         Classifier=factor(Classifier),
-         time=time / 1e9) %>%
-  select(-expr) %>%
-  filter(p %% 500 == 0) %>%
-  tbl_df %>%
-  group_by(p, Classifier) %>%
-  mutate(wls_weights=1 / sd(time)) %>%
-  ungroup
-
-# Create DF for two facets in same plot
-# First plot shows HDRDA vs RDA
-# Second plot zooms in on HDRDA results
-timings$Plot <- "Comparison of HDRDA and RDA"
-timings_hdrda <- timings %>%
-    filter(Classifier == "HDRDA") %>%
-    mutate(Plot="HDRDA")
-timings_combined <- bind_rows(timings, timings_hdrda)
-
-# Build data frame to show relative timings in table
-timing_ratios <- timings %>%
-    group_by(p, Classifier) %>%
-    summarize(median_time=median(time)) %>%
-    tidyr::spread(Classifier, median_time) %>%
-      mutate(ratio_time=RDA / HDRDA) %>%
-      select(-HDRDA, -RDA)
-
-timings_summary <- timings %>%
-  group_by(Classifier, p) %>%
-  summarize(mean_runtime=mean(time), sd_runtime=sd(time))
-
-# RDA Mean Runtime in Minutes when p = 5000
-rda_mean_time_p5000 <- timings_summary %>% filter(Classifier == "RDA", p == 5000)
-rda_mean_time_p5000 <- rda_mean_time_p5000$mean_runtime / 60
-
-# HDRDA Mean Runtime in Seconds when p = 5000
-hdrda_mean_time_p5000 <- timings_summary %>% filter(Classifier == "HDRDA", p == 5000)
-hdrda_mean_time_p5000 <- hdrda_mean_time_p5000$mean_runtime
-@
-
-<<bootstrap_relative_timing_comparisons>>=
-ratio_means <- function(df, num_reps=1000) {
-  bootstrap_reps <- replicate(n=num_reps, {
-    df %>%
-      group_by(Classifier) %>%
-      sample_frac(size=1.0, replace=TRUE) %>%
-      summarize(mean_time=mean(time)) %>%
-      tidyr::spread(Classifier, mean_time) %>%
-      mutate(ratio_means=RDA / HDRDA) %>%
-      select(-HDRDA, -RDA)
-  }, simplify=FALSE)
-  bind_rows(bootstrap_reps)
-}
-
-set.seed(42)
-timing_ratios_bootstrap <- timings %>%
-    mutate(p=factor(p)) %>%
-    group_by(p) %>%
-    do(ratio_means(., num_reps=1000))
-
-timing_bootstrap_summary <- timing_ratios_bootstrap %>%
-  group_by(p) %>%
-  summarize(avg_ratio=mean(ratio_means),
-            lower_ci=quantile(ratio_means, probs=0.025),
-            upper_ci=quantile(ratio_means, probs=0.975))
-bootstrap_summary_p500 <- timing_bootstrap_summary %>%
-  filter(p == "500")
-bootstrap_summary_p5000 <- timing_bootstrap_summary %>%
-  filter(p == "5000")
-
-@
 
@@
 drastically smaller than that of \emph{RDA}. In fact, when $p = 5000$, we show
-that the \emph{HDRDA} classifier is \Sexpr{bootstrap_summary_p5000$avg_ratio}
+that the \emph{HDRDA} classifier is 502.786
 times faster on average than the \emph{RDA} classifier. In this scenario, the
-\emph{HDRDA} classifier's model selection requires \Sexpr{hdrda_mean_time_p5000}
+\emph{HDRDA} classifier's model selection requires 2.979
 seconds on average, while that of the \emph{RDA} classifier requires
-\Sexpr{rda_mean_time_p5000} minutes on average.
+24.933 minutes on average.
 
@@
 fact, when $p = 5000$, the \emph{RDA} classifier required
-\Sexpr{rda_mean_time_p5000} minutes on average to perform model selection, while
+24.933 minutes on average to perform model selection, while
 the \emph{HDRDA} classifier selected its optimal model in
-\Sexpr{hdrda_mean_time_p5000} seconds on average. Clearly, the model selection
+2.979 seconds on average. Clearly, the model selection
 employed by the \emph{HDRDA} classifier is substantially faster than that of the
@@
 dimensions, the relative difference in computing was sizable with the average
-ratio of the mean timings equal to \Sexpr{bootstrap_summary_p500$avg_ratio} for
+ratio of the mean timings equal to 14.513 for
 $p = 500$ and a 95\% confidence interval of
-(\Sexpr{bootstrap_summary_p500$lower_ci},
-\Sexpr{bootstrap_summary_p500$upper_ci}). Furthermore, the ratio of mean
+(14.191,
+14.855). Furthermore, the ratio of mean
 computing times suggested that the \emph{RDA} classifier is impractical for
 higher dimensions. For instance, when $p = 5000$, the ratio of mean computing
-times increased to \Sexpr{bootstrap_summary_p5000$avg_ratio} with a 95\%
-confidence interval of (\Sexpr{bootstrap_summary_p5000$lower_ci},
-\Sexpr{bootstrap_summary_p5000$upper_ci}).
+times increased to 502.786 with a 95\%
+confidence interval of (462.863,
+546.396).
 
@@
 comparisons, we showed that \emph{HDRDA} model selection could be employed on
-data sets with $p = 5000$ in \Sexpr{hdrda_mean_time_p5000} seconds on
+data sets with $p = 5000$ in 2.979 seconds on
 average. Contrarily, the \emph{RDA} classifier implemented in the {\tt klaR} R
-package required \Sexpr{rda_mean_time_p5000} minutes on average to perform model
+package required 24.933 minutes on average to perform model
 selection on data sets with $p = 5000$. Given that the \emph{RDA} classifier has
@@
 
-<<classification_results, results='asis'>>=
-table_caption <- 'The average of the test error rates obtained on gene-expression data sets over 100 random training-test partitions.
+% latex table generated in R 3.4.2 by xtable 1.8-2 package
+% Mon Oct 16 21:24:13 2017
+\begin{table}[ht]
+\centering
+\scalebox{0.85}{
+\begin{tabular}{lllllll}
+  \hline
+Classifier & Chiaretti & Chowdary & Nakayama & Shipp & Singh & Tian \\ 
+  \hline
+Guo & \textbf{0.111} (0.044) & 0.056 (0.051) & 0.208 (0.061) & 0.086 (0.063) & \textbf{0.089} (0.055) & 0.268 (0.082) \\ 
+  HDRDA Convex & 0.115 (0.044) & 0.035 (0.026) & \textbf{0.208} (0.066) & 0.073 (0.057) & 0.111 (0.059) & 0.229 (0.049) \\ 
+  HDRDA Ridge & 0.118 (0.050) & \textbf{0.033} (0.022) & 0.208 (0.070) & \textbf{0.072} (0.065) & 0.099 (0.046) & 0.225 (0.050) \\ 
+  Pang & 0.663 (0.062) & 0.197 (0.091) & 0.227 (0.062) & 0.192 (0.091) & 0.221 (0.095) & 0.267 (0.054) \\ 
+  Random Forest & 0.124 (0.053) & 0.045 (0.028) & 0.232 (0.063) & 0.135 (0.078) & 0.093 (0.045) & \textbf{0.206} (0.044) \\ 
+  Tong & 0.195 (0.068) & 0.197 (0.091) & 0.227 (0.062) & 0.192 (0.091) & 0.221 (0.095) & 0.267 (0.054) \\ 
+  Witten & 0.194 (0.068) & 0.197 (0.091) & 0.232 (0.068) & 0.193 (0.092) & 0.221 (0.095) & 0.264 (0.053) \\ 
+   \hline
+\end{tabular}
+}
+\caption{The average of the test error rates obtained on gene-expression data sets over 100 random training-test partitions.
 Standard deviations of the test error rates are given in the parentheses.
-The classifier with the minimum average error rate for each data set is in bold.'
-
-xtable_out <- xtable(results_microarray,
-                     caption=table_caption,
-                     label='tab:microarray-results')
-print(xtable_out, include.rownames=FALSE, sanitize.text.function = function(x) { x },
-      scalebox='0.85') # scales the table to fit the page
-@
+The classifier with the minimum average error rate for each data set is in bold.}
+\label{tab:microarray-results}
+\end{table}
 
@@
 \begin{figure}
-<<hdrda_contours, fig.width=8, fig.height=8, out.width='.33\\linewidth', fig.show='hold'>>=
-rda_contours(lambda = 0)
-rda_contours(lambda = 0.2)
-rda_contours(lambda = 0.4)
-rda_contours(lambda = 0.6)
-rda_contours(lambda = 0.8)
-rda_contours(lambda = 1)
-@
+\includegraphics[width=.33\linewidth]{figure/hdrda_contours-1} 
+\includegraphics[width=.33\linewidth]{figure/hdrda_contours-2} 
+\includegraphics[width=.33\linewidth]{figure/hdrda_contours-3} 
+\includegraphics[width=.33\linewidth]{figure/hdrda_contours-4} 
+\includegraphics[width=.33\linewidth]{figure/hdrda_contours-5} 
+\includegraphics[width=.33\linewidth]{figure/hdrda_contours-6}
+
 \caption{Contours of five multivariate normal populations as a function of the
@@
 \begin{figure}
-<<timing_results, fig.width=8, fig.height=8>>=
-p <- ggplot(timings_combined, aes(x=p, y=time, color=Classifier))
-p <- p + geom_point() + geom_jitter(width=140, height=0) + facet_wrap(~ Plot, scales="free_y")
-p <- p + geom_smooth(method="lm", formula=y ~ poly(x, 2), se=FALSE)
-p <- p + scale_color_manual(values=c("#D55E00", "#0072B2"))
-p <- p + xlab("Number of Features") + ylab("Model Selection Time (seconds)")
-p + theme_bw()
-@
+\includegraphics[width=\linewidth]{figure/timing_results-1}
 \caption{Timing comparisons (in seconds) between HDRDA and RDA classifiers.}
@@
 \begin{figure}
-<<relative_timings, fig.width=8, fig.height=8>>=
-p <- ggplot(timing_ratios_bootstrap, aes(ratio_means, fill=p))
-p <- p + geom_histogram() + facet_wrap(~ p, scale="free_x", labeller="label_both")
-p <- p + xlab("Ratio of Mean RDA Runtime to Mean HDRDA Runtime") + ylab("Bootstrap Replications")
-p + theme_bw()
-@
+\includegraphics[width=\linewidth]{figure/relative_timings-1}
 \caption{Distribution of ratios of mean \emph{RDA} runtime to mean \emph{HDRDA}
@@
 \begin{figure}
-<<sim_results, fig.width=8, fig.height=8>>=
-p <- ggplot(results_contaminated, aes(x=contamination_prob, y=Error))
-p <- p + geom_line(aes(color=Classifier, linetype=Classifier), size=0.65)
-p <- p + facet_wrap(~ p, labeller = "label_both")
-p <- p + xlab("Contamination Probability") + ylab("Average Error Rate")
-p + theme_bw()
-@
+\includegraphics[width=\linewidth]{figure/sim_results-1} 
 \caption{Average classification error rates as a function of the contamination
   probability $\epsilon$. Approximate standard errors were no greater than
-  \Sexpr{max_sim_se}.}
+  0.022.}
 \label{fig:sim-results}
@@
 \begin{figure}
-<<sim_results_by_p, fig.width=8, fig.height=8>>=
-p2 <- results_contaminated %>%
-  mutate(p=as.numeric(as.character(p)),
-         contamination_prob=round(contamination_prob, digits=2)) %>%
-    filter(contamination_prob %in% c(0.0, 0.1, 0.2, 0.3, 0.4, 0.5)) %>%
-    rename(`Contamination Probability`=contamination_prob) %>%
-      ggplot(., aes(x=p, y=Error))
-p2 <- p2 + geom_line(aes(color=Classifier, linetype=Classifier), size=0.65)
-p2 <- p2 + facet_wrap(~ `Contamination Probability`, labeller="label_both")
-p2 <- p2 + xlab("Number of Features") + ylab("Average Error Rate")
-p2 + theme_bw()
-@
+\includegraphics[width=\linewidth]{figure/sim_results_by_p-1} 
 \caption{Average classification error rates as a function of the number of
   features $p$. Approximate standard errors were no greater than
-  \Sexpr{max_sim_se}.}
+  0.022.}
 \label{fig:sim-results-by-p}
